### 2014-05-02 Knitr/Noise.Rnw
Prepared the document for noise trade article

\documentclass[12pt, a4paper, oneside]{article} % Paper size, default font size and one-sided paper
%\graphicspath{{./Figures/}} % Specifies the directory where pictures are stored
%\usepackage[dcucite]{harvard}
\usepackage{rotating}
\usepackage{amsmath}
\usepackage{setspace}
\usepackage{pdflscape}
\usepackage[flushleft]{threeparttable}
\usepackage{multirow}
\usepackage{listings}
\usepackage[comma, sort&compress]{natbib}% Use the natbib reference package - read up on this to edit the reference style; if you want text (e.g. Smith et al., 2012) for the in-text references (instead of numbers), remove 'numbers' 
\usepackage{graphicx}
%\bibliographystyle{plainnat}
\bibliographystyle{agsm}
\usepackage[colorlinks = true, citecolor = blue, linkcolor = blue]{hyperref}
%\hypersetup{urlcolor=blue, colorlinks=true} % Colors hyperlinks in blue - change to black if annoying
%\renewcommand[\harvardurl]{URL: \url}
\begin{document}
\title{Noise trader: random walks or forced march?}
\author{Rob Hayward\footnote{University of Brighton Business School, Lewes Road, Brighton, BN2 4AT; Telephone 01273 642586.  rh49@brighton.ac.uk}}
\date{\today}
\maketitle
\begin{abstract}
Noise traders have been characterised as random providers of liquidity, as neutral equivalents of computor-generated random numbers.  hough noise traders may create risk that will deter rational traders from taking advantage of their knowledge of the fundamental value of securities (De-long) and thereby preventing them being driven into extinction as Friedman had asserted (Fridman), their presence  is usually seen as beneficial as they provide liquidity in a neutral fashion.  However, an analysis of noise trading that seeks to measure the weight of noise traders in the market and the strength of their opinin shows that herding is a common feature, highlighting the non-neutral influence that this speculative activity can impose on securities prices. 
\end{abstract}

\section{Introduction}
\section{Literature}
\section{Methods}
\section{Analysis}
\section{Conclusion}
\end{document}

### 2014-05-02 Rcode/Delong.R

@@ -1,0 +1,121 @@
+# Build the DeLong function made of 'rho' as the actual misperception, rhomu as
+# the average misperception, r as the risk free rate, mu is the weight of 
+# speculators in the total market participants, ra as the parameter for 
+# risk aversion, rhosd is the standard deviation of the misperception. 
+require(zoo)
+require(eventstudies)
+# Function to calculate price under DeLong model-----------------
+DeLong <- function(rho, rhomu, r, mu, ra, rhosd){
+  p = 100 + (mu * (rho - rhomu))/(1 + r) + (mu * rhomu)/r - 
+    ((2 * ra * mu^2 * rhosd^2)/(r * (1 + r)^2))
+  return(p)
+}
+# Set misperception as a random vaiable with a mean of 1 and a 
+# standard deviation of 1
+set.seed(1)
+rho <- rnorm(1000, 4, 2)
+p <- DeLong(rho = rho, rhomu = 4, r = 0.1, mu = 0.5,
+            ra = 0.3, rhosd = 2)
+length(p)
+plot(p, type = 'l', main = "Delong model with misperception as random variable")
+da <- cbind(p, rho)
+head(da)
+pz = as.zoo(da)
+# Create returns
+pz$r <- 100*diff(log(pz$p))
+str(pz)
+head(pz)
+# Create extreme as above 95th percentile
+pz$e = as.numeric(pz$rho >= quantile(pz$rho, 0.95, na.rm = TRUE))
+head(pz)
+eventslist.ex <- data.frame(index(pz$e[pz$e == 1]))
+eventslist.ex$unit <- as.character("r")
+colnames(eventslist.ex) <- c("when", "unit")
+dim(eventslist.ex)
+tail(eventslist.ex)
+# Align events
+a <- phys2eventtime(pz, eventslist.ex, width = 2)
+P.ex <- window(a$z.e, start = 0, end = 1)
+P.ex
+# Calculate descriptive statistics of rho extreme-------------------
+P.ext <- t(P.ex)
+P.ext
+mystats <- function(x, na.omit=FALSE){
+  if (na.omit)
+    x <- x[!is.na(x)]
+  m <- mean(x)
+  s <- sd(x) 
+  t <- m/s
+  n <- length(x)
+  skew <- sum((x-m)^3/s^3)/n
+  ps <- ((6*n*(n-1))/((n-1)*(n+1)*(n+3)))^0.5
[…]

### 2014-07-03 Knitr/Noise.Rnw
Updated the files commited earlier

@@
 Noise traders have been characterised as random providers of liquidity, as neutral equivalents of computor-generated random numbers.
-hough noise traders may create risk that will deter rational traders from taking advantage of their knowledge of the fundamental value of securities (De-long) and thereby preventing them being driven into extinction as Friedman had asserted (Fridman), their presence  is usually seen as beneficial as they provide liquidity in a neutral fashion.
+Though noise traders may create risk that will deter rational traders from taking advantage of their knowledge of the fundamental value of securities (De-long) and thereby preventing them being driven into extinction as Friedman had asserted (Fridman), their presence  is usually seen as beneficial as they provide liquidity in a neutral fashion.
 However, an analysis of noise trading that seeks to measure the weight of noise traders in the market and the strength of their opinin shows that herding is a common feature, highlighting the non-neutral influence that this speculative activity can impose on securities prices.
@@
 \section{Literature}
+\subsection{Notes}
+From "How noise trading..." in the other folder. 
+O'Hara and Saar suggest that there is a difference between noise traders and liquidity traders.
+Noise traders exert positive effects on liquidity but weaken price reversals.
+A tax on trading will reduce noise traders but will also reduce the influence of fundamental traders.
+
+Literature.
+Black (1986) highlights the positive effect of noise traders.
+Shliefer, Summers, DeLong etc identify the limits to arbitrage.
+The other writings are Battalio, Hatch and Jennings (1997), Harris and Schultz (1998) and Scheinkman and Xiong (2003).
+Lowenstein and Willard (2006).
+There is considerable debate about the precise role of noise trading in financial markets.
+There are two strands to the literature on noise trading that emerged in the 1980s:  microstricture literature will use the term "noise traders" and "liquidity traders" interchangeably to identify traders without information.
+The motivation of these traders is not identified.
+Examples of this are Glosten and Milgrom (1985) and Kyle (1985).
+It is usually assumed to be hedging or liquidity.
+There is also a \emph{limits-to-arbitrage} literature that uses the term to identify traders that are not movivated by standard explanations of fundamental or liquidity provision.
+This paper seeks to distinguish between those that trade for liquidty or heging and those that trade because of psychological biases.
+Fisher Black ``People who trade on noise
+are willing to trade even though from an objective point of view they would be better off 
+not trading. Perhaps they think the noise they are trading on is information. Or perhaps 
+they just like to trade``.
+
+In theory, traders may trade on mistaken fundamental information.
+Shliefer and Summmer (1990) use momentum.
+However, Kaniel, Saar and Titman (2006) find evidence of contrarian strategies.
+This is associated with short-term speculation.
+Does the activity drive the price away from fundamental value?
+Stiglitz (1990) is one of the arguments in favour of a Tobin tax.
+
+There is a quesiton about whether experimental markets (like this paper) are superior to actual markets where the identification of trader types is difficult.
+Theory - Delong, summers, shleifer etc,  Scheinkman and Xiong (2003).
+Lowenstein and Willard (2006).
+Empirical work:  (see, for example, Garvey and Murphy [2001], Linnainmaa [2003], Barber et.
+al.
+[2004], and Kaniel, Saar, and Titman [2006].
+
+For related papers on securities transaction taxes see Stiglitz [1989], Summers and Summers [1989], 
+Amihud and Mendelson [1993], Schwert and Seguin [1993], Umlauf [1993], Subrahmanyam [1998], Dow 
+and Rahi [2000], Habermeir and Kirilenko [2003], and Hinman [2003].
+
+Interesting difference between supplying and demanding liqudiity.
+Limit orders supply liquidity, market orders demand liqudity.
+The activity of liquidity traders will tend to change through the course of the session, moving from supplying to demanding liquidity.
+Noise traders tend to act in the opposite manner.
+They tend to supply more liquidity as the session progresses.
+Black (1986) noise traders put noise into prices.
+The effect of noise traders on liquidity: price impact of trades is lower when noise traders are present.
+However, this is mostly in the temporary impact on prices rather than the permanent impact.
+Therefore, noise traders tend to reduce the amount of volatility in prices because market orders do not temporarily have an effect and then reverse.
+The presence of noise traders will increase pricing errors.
+However, this depends on fundametal traders have valuable information.
+
+There are a number of different ways that noise traders could behave:  activing as SOES bandits (using the NASDAQ small order execution system to make profits) or by using information about the order book; by providing liqudity like a market-maker as a service to liqudiity-traders as in Grossman Miller (1988);  as irrational noise traders, using trading rules or popular models (Shiller 1984, 1990) positive feedback or contrarian.
+
+Footnote -  There is an empirical evidence that individual investors in various countries trade in a contrarian fashion (e.g., Choe, Kho, and Stulz [1999], Grinblatt and Keloharju [2000, 2001], Jackson [2003], Richards [2005], Kaniel, Saar, Titman [2006]).
+For experimental evidence consistent with contrarian behavior see
+Bloomfield, Tayler, and Zhou (2997)
+
+Can noise traders be viewed as just random noise.
+Dow and Gorton (2006) suggest that this is the case and that this is an outcrop of the rational expectations revolution.
+Informed traders profit by exploiting the diviations from equilibrium.
+This is the beneifts of Grossman-Stiglitz.
+What is the source of this noise?
+What are the insitiutions?
+The importance of noise traders here is that they represent a zero-mean, normally distributed random vaiable.
+This is also the description that is provided in Kyle (1985).
+Liquidity traders trade due to imbalance in the timing of consumption or from portfolio considerations.
+Shleifer, Summers and DeLong have trading activity that is affected by sentiment, models or "fads".
+This is important.
+How the average may not be zero.
+There is a range of behavioural biases that may account for this (see Barberis and Thaler [2003]).The literaure on hetrogenous traders also focuses on noise traders as trend chasers.
+These models (see
+for example Brock and Hommes [1998]; Lux [1995]; Lux and Marchesi [2000]) investigate bubbles and 
+crashes, and feature fundamental traders versus noise traders. In these models, it is the noise traders who 
+induce the aberrant market behavior.
+
 \section{Methods}

### 2014-09-25 Knitr/Paper.tex
Updated the Paper.tex file.  Rationalised the folder.  There are some additional notes elsewhere.

\documentclass[12pt, a4paper, oneside]{article} % Paper size, default font size and one-sided paper
%\usepackage[dcucite]{harvard}
\usepackage{tikz}
\usetikzlibrary{shapes, shadows, arrows}
\usepackage{rotating}
\usepackage{amsmath}
%\usepackage{setspace}
\usepackage{pdflscape}
\usepackage[flushleft]{threeparttable}
\usepackage{multirow}
\usepackage[comma, sort&compress]{natbib}% Use the natbib reference package - read up on this to edit the reference style; if you want text (e.g. Smith et al., 2012) for the in-text references (instead of numbers), remove 'numbers' 
\usepackage{graphicx}
\graphicspath{{./Figures/}} % Specifies the directory where pictures are stored
%\bibliographystyle{plainnat}
\usepackage{listings}
\bibliographystyle{agsm}
\usepackage[colorlinks = true, citecolor = blue, linkcolor = blue]{hyperref}
%\hypersetup{urlcolor=blue, colorlinks=true} % Colors hyperlinks in blue - change to black if annoying
\begin{document}
%\renewcommand[\harvardurl]{URL: \url}
\title{Speculation in foreign exchange: noise or information}
\author{Rob Hayward\footnote{University of Brighton Business School, Lewes Road, Brighton, BN2 4AT; Telephone 01273 642586.  rh49@brighton.ac.uk.  We are very grateful for the comments and suggestions made.  In particular, the EACES conference in Budapest between September 5th and 6th 2014, Andros Steinhoffer and.....}} 
\date{\today}
\maketitle
\begin{abstract}
An event study of extreme speculation is used to test the informational content of speculation.  Speculative extremes are ranked by weight of activity and intensity of sentiment.  In either case, the extreme says little about the future, suggesting that speculation is more than just noise and that there is some fundamental informational content. A unique dataset of risk-reversal skew on option prices is used to measure the weight of speculative sentiment.   
\end{abstract}

\section{Introduction}

There is a fundamental schism in the literature between those that view speculation as uninformed noise activity that can run in waves when affected by common decision-making bias or social influences, and those that view speculation as activity informed by fundamental knowledge that will stabilise markets.  In the former case, there is a risk that speculative activity will drive prices away from fundamental value as price changes and sentiment feed off each other.  At its most extreme, speculation in this interpretation is likely to turn into a bubble that may burst in a cascade where the previous run-up in price and sentiment is swiftly reversed.  However, the alternative view of speculation as an activity that is driven by fundamental knowledge and information would suggest that speculation acts to limit overshooting and divergence from fundamental value, thereby reducing volatility.  This second view says that it is the absence of speculators and the liquidity that they provide that encourages excess price movement and volatility.  The aim of this chapter is to use the foreign exchange market to assess the relationship between intensity or weight of speculation and the movement of prices.  If speculators are uninformed, their most extreme sentiment and their greatest weight in market activity should tend to cause overshooting and presage bubble-bursting, reversals and a crash; if speculators are informed, activity will drive prices towards fundamental value and equilibrium where volatility and crash risk should be reduced.

\subsection{Fundamental value}
Fundamental or intrinsic value is a difficult concept to apply to foreign exchange.     There is a weight of empirical evidence that standard economic models are not very effective in explaining the behaviour of exchange rates.  Meese and Rogoff found that in out-of-sample forecasts, a random walk performed as well as the four major economic models of foreign exchange behaviour that were employed \citep{Meese1983Empirical}.  These findings have been repeated in a number of similar, subsequent studies  \citep{Frankel1984tests},  \citep{Frankel1990Exchange}, \citep{Lyons1995Microstructure} and \citep{Sarno2006}.  In the previous chapter, the fundamental value of the exchange rate was assessed to be PPP.  However, even there the real exchange rate was not stable (see Figure \ref{fig:timeseries} and view the variable RTWI to see the evolution of the US real trade weighted index over the period 1985 to 2010).  Therefore, while the law of one price and purchasing power parity generally determines exchange rates in the long-run for currencies of countries at a similar level of economic development, in the short to medium-term there is still no robust economic model that can consistently explain exchange rate movements, let alone provide a forecast.

Noise-trading is the label attached to that trading or investment where decisions seem to have been made with no reference to fundamental information; activity is like a random noise.  This chapter provides an investigation into the effect of speculation in the foreign exchange market.  It seeks to assess the relationship between sentiment, speculation, information and exchange rate prices.  The DeLong-Summers model of noise trading is calibrated against daily and weekly returns for major foreign exchange rates \citep{Delong1990noise}.  Little support for the hypothesis that extreme levels of sentiment or extreme levels of speculative activity convey information about the future direction of prices is found.  This is something of a contrast to the evidence that has been found in some other markets.  See  \citep{DeBondtOver}, \citep{Chopra1992} and \citep{SteinOptions} amongst others (discussed more fully in Section \ref{secref:empiric} below).   The results of the experiment carried out here suggest that it is not sufficient just to measure the intensity of speculative sentiment or the weight of speculators to determine the divergence from intrinsic value and that the relationship between speculation and price is rather complex.  There is evidence here against the idea of speculation as purely noise.  Rather, it appears that speculation is part of the price discovery process.

\section{Literature review}  
In many ways this is a study about how efficient the market  is in processing information, specifically whether the presence of speculators improves or reduces the efficiency with which information is processed.  As such, there must first be some discussion of information and ideas about the efficiency with which it is incorporated into asset prices.

Bachelier is responsible for the first structured  study of informational efficiency in his 1900 doctoral analysis of the behaviour of French government bond, future and option prices \citep{Bachelier1900}.  His is apparently the first analysis that suggested that the evolution of securities' prices was a \emph{Martingale}, meaning that past events do not help to predict the future value, that the expected future value is equal to the current value and that the process is a \emph{fair game} for which the probability of winning or losing is equal.  A \emph{random walk} is a slightly stronger construct as each step is an independent and identically distributed random variable.  However, a martingale may contain past information about the higher moments.   So there may be, for example, volatility clustering. Brownian motion is named after the Scottish botanist Robert Brown and refers to the random movement of particles suspended in a liquid or gas.  See \citep{LeRoyMartingales} for an overview of market efficiency, random walks and Martingale processes.  Bachelier studied a set of bond prices and found that the potential returns of these $19^{th}$ century French government bonds could be represented by a probability distribution where the mean was zero once the gravitation towards parity at maturity and seasonal cycle of `dirty prices' with accrued interest had been accounted for.

\begin{quotation}
`` The determination of these fluctuations depends on an infinite number of factors; it is, therefore impossible to aspire to mathematical prediction of it.  Contradictory opinions concerning these changes diverge so much that at the same instant buyers believe in a price increase and sellers in a price decrease."
\end{quotation}

\citep[p. 17]{Bachelier1900}

Bachelier's ideas received little immediate attention and remained dormant until renewed interest in the process describing securities prices developed in the 1950s.  However, others were tentatively pursuing similar investigations.  In the 1930s Working assessed the empirical evidence for random behaviour of stock and commodity prices \citep{Working1934} and U.K. statistician Kendall studied the performance of weekly price data from the Dow Jones \citep{Kendall1953}.  Among those taking an interest in the subject were Roberts, who scorned the \emph{patterns} that were identified by professional investors while making the case for a focus on the change in price of securities rather than the price itself.  What we now call \emph{returns} were, argued Roberts, much more random than price levels.  He calibrated a model of random shocks to the price change, using the RAND Corporation's random number text \citep{RAND} and created graphs that were remarkably close to those produced by real stocks.  Nonetheless, Roberts warned his research would not prevent people from using past data to try to make investment decisions as ``chance behaviour itself provides 'patterns' that invite spurious interpretations'' \citep[p. 2]{Roberts1959}.

Towards the end of the decade, Osborne made the direct comparison between the price movement of stocks and the activity of small particles through the analysis of prices of a selection of stocks from the American and New York stock exchanges.  The data were collected from \emph{The Wall Street Journal} by a research company called F.W. Stephens for the period from 1924 to 1956.  The log of the price ratio for a range of different time periods and their distribution was compared to what would have been expected from something evolving with Brownian motion.  By plotting the ratio over time, Osborne showed that changes in the log of the price ratio was close to being normally distributed.  Osborne argued that the securities prices jumped around to find the point where buyers and sellers were balanced and overall opinion about the direction of prices was close to zero. Therefore, the evolution of prices was a like a sequence of independent random variables that were the consequence of transactions taking place.  The distribution function for this sequence of random variables was, according to Osborne's study, normal with a zero mean and a variance that increased steadily with time.  Osborne concluded that the random nature of price changes would ensure that buyers and sellers could only make a standard return on average  \citep{Osborne1959}.

Samuelson provided a theoretical foundation for the idea that securities prices would follow a random walk.  Starting with a series of simple axioms, he deduced that competitive markets would drive securities prices towards the expected value, where the expected value is defined as the projected mean of the process underlying the evolution of the price.  This introduced the idea of a \emph{Martingales} or the notion that the expected value of a security or financial asset would be equal to its current value.  If that is not the case, rational investors in a competitive market will take advantage of the deviation by buying when it is below the expected value and selling when it is above and, as a result of this activity, would drive the price towards the expected level and, in the process, remove the opportunity to make profits.   Using the backward iteration of expectations, the theory states that a stock price should equal the expected present value of future dividends. This work provided the underpinning and stimulus for another wave of empirical examinations but rests, of course, on the assumptions made about the competitive nature of markets and the ability of investors to form expectations about the future \citep{SamuelsonStockPrices}.

Amongst these empirical investigations, Cootner tested Bachelier's ideas on a range of financial markets, most notably the forward market for grain \citep{CootnerSpeculators} and  Fama found that the returns of stocks in the Dow Jones Industrial Average were very close to the benchmark of a random walk \citep{FamaStock}.  At this point the term \emph{Efficient Market Hypothesis} (EMH) became more widely used.  Jensen described market efficiency as

\begin{quotation}
``A market is efficient with respect to information set $\Omega_t$ if it is impossible to make economic profits by trading on the basis of information set $\Omega_t$".  
\end{quotation}
\citep[p. 95]{JensenEfficiency}.

More formally,

\begin{equation}
E[R_{t+1}, Q_{t+1} | \Omega_t] = 0
\end{equation}

where $E$ is the expectations operator, $Q_{t+1}$ is the risk premium at time $t+1$, $R_{t+1}$ is the return at time $t +1$ and $\Omega_t$ is the information set at time $t$ \citep[p. 95]{JensenEfficiency}.  Jensen also said "I believe that there is no other proposition in economics which has more solid empirical evidence supporting it than the Efficient Market Hypothesis" \citep[p. 93]{JensenEfficiency}.   Jensen was, however, identifying a number of inconsistencies and anomalies that had been uncovered and was anticipating a revolution that would overthrow the theory.  He was anticipating the paradigm shift that was to follow.

%this paragraph goes somewhere else. 
%Among these were positive returns following an earnings release \citep{Ball1978} and \citep{Watts1978};  a successful trading rule for buying closed-end funds trading at a discount to asset value \citep{Thompson1978}; discrepancies between stock and option prices on the NYSE \citep{Galai1978}; and, information in option prices \citep{ChirasManaster}.

\subsection{Failures of market efficiency}
Questions about the EMH had emerged gradually with doubts about some of the assumptions that Samuelson had used in the construction of his theory and with a rising number of indications that financial asset prices did not evolve in a totally random fashion.  There were three main strands to the question that arose:   empirical evidence that the price of securities did not follow a random walk made up of a series of independent and identically distributed random variables;  questions about the effects of institutional structures of financial markets on price;  the criticism of VNM expected utility theory as a valid model of investment decision-making.

\subsubsection{Empirical evidence}
The empirical evidence that has been gathered about the nature of the returns to securities reveals two primary issues.  The first is that serial correlation or trends are evident and the second is the leptokurtic and skewed distribution of these returns.  The former is the most damaging to the idea that financial markets can be informationally efficient as it suggests, contrary to the implications of the EMH, that future price changes could be partially foreseen by the movement of past price changes.\footnote{This of course depends on the assumption of \emph{risk neutrality} and the notion that investor decisions will be determined by the first two moments of the distribution of expected returns.  If the return distribution is skewed or fat tailed or there is serial correlation in the conditional variance of the returns, there may be unexplored opportunities to reduce risk or to increase risk-adjusted returns. The third and fourth moments of the distribution of returns can be extremely important in assessing the level of risk.}  The evidence seems to support the notion that prices can evolve in trends, and that there are times when positive price developments are more likely to be followed by further positive developments and that there are other times when positive price developments are more likely to be followed by a negative outcome.

There are a number of studies that find evidence that the reaction of stocks to information is less than instantaneous.  Ball and Brown seems to be the first study that identifies the upward drift in cumulative abnormal returns after some positive earnings news and downward drift in cumulative abnormal earnings after negative earnings news \citep{BallBrown}.  Bernard provides a summary of the evidence that stocks react slowly to earnings announcements.  He cites twenty examples of ``post earnings announcement drift'' \citep[p. 303]{BernardDrift}.  Part of this under-reaction comes from simply failing to identify what current earnings mean for subsequent quarters.    Bernard argues that the evidence is firmly on the side of earning reactions being too small and taking on average six months to be fully digested \citep[p. 305]{BernardDrift}.

Jagadesh and Titman tested the theory that there is momentum or serial correlation in the stock prices by creating portfolios that bought stocks that had performed well and sold those that had performed poorly over a period of between three months and one year before the selection date.  The portfolios record positive abnormal returns initially, but these returns start to disappear beyond the 12 month horizon.  The most successful strategy buys stock based on their performance over the previous 12 months and holds for a further 3 months.  The return here is 1.3\% per month \citep{Jagadeesh}.

Jeremy Stein also finds evidence of overshooting in the options market, what he calls ``the systematic tendency to overemphasize recent data at the expense of other information when making projections'' \citep[p 1011]{SteinOptions}. Here the term structure of implied volatility is inconsistent with rational behaviour or the pricing model.  Implied volatilities on the longer term options move almost step-in-step with the shorter ones despite the evidence of mean-reversion and swift decay of volatility shocks, suggesting an overreaction to news that spreads from the short end, where it is appropriate, to the long end, where it is not.   Short-term spikes in volatility are likely to be reversed, meaning that the long-dated options should be less affected than the short-dated options.

For a longer investment period, there is some evidence of negative serial correlation.  Eugene Fama and Kenneth French studied stock returns over the period of 1926 to 1985 and found slow mean-reversion or negative autocorrelation for return periods beyond one year.  This is similar to the cut-off period for positive momentum found in \citep{Jagadeesh}, suggesting some sort of threshold around 12 months.  Fama and French assert that up to forty percent of the 3-to-5 year return variation for small firms is explained by this mean reversion and up to twenty-five percent of the return variation of larger firms over the same time horizon \citep{FrenchFamaPermanent}.

De Bondt and Thaler investigate the issue of overshooting by building \emph{winning} and \emph{losing} portfolios  based on the best and worst performing stocks each year.  They use data from the Center for Research in Securities Prices (CRSP) on securities from the New York Stock Exchange for the period between January 1930 to December 1982.  They then compare the performance of the two portfolios to assess whether there is evidence that investors have contributed to an overshooting of the \emph{winning} stocks and an undershooting of the \emph{losing} stocks.  They reject the hypothesis that the returns from the two portfolios are the same and find that on average, over the three years following the portfolio contraction, the \emph{losing} portfolio returns 25 percentage points more than the \emph{winning} portfolio \citep{DeBondtOver}.

The idea that over-and-under shooting can be used to identify securities that have caught up in a wave of positive or negative sentiment has been repeated a number of times.  De Bondt and Thaler used the performance over the previous year to identify \emph{winners} and \emph{losers}, other have broken the portfolios into size-adjusted and beta-adjusted components to test the suggestion that the returns recorded are just a compensation for increased risk  \citep{Chopra1992}.  The main findings are that smaller firms account for the bulk of the gains but that overshooting is still evident.  There are other examples where a variety of measures, including the price-earnings ratio \citep{BasuPE} and cash flow to price \citep{Chan1991} have been used to assess the level of over-shooting.  It is shown that book-to-market and cash-flow ratios have the most positive impact on positive returns.  Lakonishok, Shleifer and Vishny provide an overview of \emph{contrarian} investment strategies \citep{LakonishokContrarian}.  This whole area is closely linked to the debate about the relative merits of \emph{value} and \emph{growth} stocks, with ample evidence from academic and practitioners that while growth stocks outperform value in the short-run, in line with the evidence of momentum, there are frequent reversals as over-reactions are reversed.

\subsubsection{Normal distribution}
The indications that the distribution of financial security returns are likely to be leptokurtic and skewed would not necessarily undermine the EMH as long as the distributions were independent.  However, this evidence that the distribution of returns is not normal, reinforces the view that standard analysis of investment based on the first two moments of the return distribution may not be adequate to understand risk.   This is something that will be more fully investigated in Chapter \ref{Chapter4} with a study of the carry trade.  In addition, if there is a pattern to the evolution of the higher moments over time, which appears to be the case, this can also suggest that there may be informational inefficiencies that can be exploited to make abnormal returns.

For example, volatility clustering means that a period when the returns to financial assets exhibit a high variance is much more likely to be followed by another period with high volatility than one where the volatility returns to some average level.  The creation of autoregressive conditional hetroscedasiticity (ARCH) and general autoregressive  conditional hetroscedasticity (GARCH) models have been designed to understand this process and provide greater precision about the effect that a shock to volatility will have on the system.  See \citep{Engle1982} for the original model, \citep{Bollerslev1986} for the generalised version, \citep{Engle1987} for an application to time-varying risk premia and \citep[pp. 657 - 665]{Hamilton} for a fuller overview of the subject.

Lo and MacKinlay directly tested the random walk hypothesis for weekly stock returns for the period between 1962 and 1985. The main aim of the study was to compare the ratio of different variances under the random walk assumption.  They found significant positive serial correlation for the market return index at weekly and monthly holding periods, indicating that returns were not independent.  The variance of a monthly return should be four times that of a weekly return if the evolution of price is a random walk.  However, there were clear deviations from this benchmark.    Asset price returns do not follow a random walk. \citep{LoMacKinlay1988}.

In 1959 Osborne found that returns were nearly normally distributed.  However, the quantile plots that he produced showed that there were slightly more extreme readings than would be expected \citep{Osborne1959}.  This was a finding that was repeated by a number of other studies.  With computer power lacking and measurement imprecise, these finding of a small deviation from normality were initially ignored.  Paul Cootner had already found that extreme returns were more likely to be recorded at the daily frequency than at the monthly level.  This, he suggested was related to the increased ability of speculators to take action at the lower frequency to limit extreme price movement \citep{Cootner1962}. However, the mathematician Benoit Mandelbrot rejected this idea, suggesting that the discrepancy between the evidence from the different frequencies was due to sample size (there being much more high-frequency data).  It was not what he had found with cotton prices.  Mandelbrot had made his own investigation of cotton prices and suggested that serial correlation in the returns was a significant problem and that an alternative distribution would be more appropriate for the returns as there were a number of extreme readings \citep{Mandelbrot1963} and \citep{Mandelbrot1971}.

\subsubsection{Institutional features}
One explanation for this serial correlation of returns is that information is absorbed into price through the process of trading and that this may cause a delay and therefore a gradual adjustment of price.  This idea is supported by research that was conducted by John Campbell, Sanford Grossman and Jiang Wang, which compared serial correlation with the volume of trading \citep{CampbellGrossmanWang}.  If a lack of trading causes information to be more gradually absorbed into price, there should be a negative relationship between trading volume and the degree of serial correlation.  This study found that for individual and stock indices, serial correlation declined with aggregate trading volume.  If this is correct, an increase in trading will increase market efficiency and points to a positive role for speculators.

\citep{LoMacKinlay1990} argue that one reason for the serial correlation in prices is the asynchronous trading system and the time lag between some stocks reacting to market news.  In other words, it could be that the underperforming stocks have not yet reacted to the news that the outperforming stocks have already incorporated.  However, this idea is rejected by Jagadeesh and Titman as a result of their method of decomposing returns into market and specific factors \citep[p. 73]{Jagadeesh}.  Indeed, Jagdeesh and Titman argue that the momentum that they find in stock price returns is consistent with delayed price reaction to firm-specific news \citep{Jagadeesh}

% Mandelbrot uses what he calls a \emph{stable Paretian} distribution to model returns.  This is a version of Levy's 1925 model and is the same as that used in chapter 4 by ref.  GED or something?  It is a general, stable distribution function with the parameters alpha, beta, x ? representing the tails, skewness, location and wideth.  A beta below 2 suggests flat tails in the distrubution. This is the distrubiton that Mandelbrot and Fama proposed. 
\subsubsection{Decision-making} 
As discussed in Section \ref{sec:prob} there are alternative views of the decision-making process.  Knight, Shackle, Keynes and others had suggested that that not all decisions could be subjected to a quantifiable, probabilisitic approach.  In addition, Simon updated the idea of \emph{economic man} with \emph{global rationality} to something that was consistent with the actual quality of information and computational capacities of economic agents.  Simon questioned the model of the \emph{economic man} with stable preferences, clear and voluminous information and computational skill.  Simon introduced the term \emph{bounded rationality} to express the limits to VNM rational-decision-making: there is only a certain amount of information that can be processed and probably only a certain amount of information that is available  \citep[p. 99]{Simon1955Behavioural}.    Simon wanted to use psychological evidence to support his ideas, but found it lacking at that time.  However, as more experimental work on the psychology of decision-making was carried out, empirical support was provided for many of the themes that Knight, Shackle, Keynes and Simon had already proposed.

%Of course, even those involved in the original construction of the EMH were aware that this was a model rather than a description of the decision-making process.  For example, back in 1959 Osborne stated that investment decisions would be made "as if" calculating the expected change in the log ratio of prices and

%\begin{quotation}
%"From what has been said about the anatomy of logical decisions (they need not be consciously logical, but this is the supposition as to how they are reached), we must suppose that for the buyer, his estimate of the expectation value of the change in value $(\Delta log_e P)$ is positive while the sellers estimate of the same quantity is negative"
%\end{quotation}
%\citep[p. 153]{Osborne1959}

%What is important is whether this model provides outcomes that are consistent with what is evident.  Again, it may be the case that in the long-term the standard model of rational decision-making provides a reasonable description of what happens in aggregate but there is a risk that at the individual level and on the path to the equilibrium, the process of groping around with limited information and limited ability to assess possible outcomes can create outcomes that diverge from that of the standard model.

%This needs to find a place in the introduction.   
%Herbert Simon introduced the term 'bounded rationality' to express the limits to rational-decision-making: there is only a certain amount of information that can be processed and probably only a certain amount of information that is available.  Simon was focused more on the decision-making process itself than the aggregate outcome of decisions that had been made.  He questioned the model of the "economic man" with stable preferences, clear and voluminous information and computational skill.   One of the key ideas that he developed was that of 'satisficing' rather than maximising utility.  It means that decisions are taken to achieve at least a particular outcome. In the field of investment, it may mean that decisions are taken to ensure that a minimum level of return is achieved rather than the maximum. This can reduce the amount of options that have to be assessed \citep{Simon1955Behavioural}.

%As we have seen, Knight, Shackle and Keynes %ref for introduction
%were particularly sceptical about rational decision-making.  Keynes argued that given the uncertainty that is attached to long-term investment projects, it is likely that investors will tend to use a short-cut to decision making that would be based on the assumption that the valuation or price will remain unchanged unless there is some strong reason to believe that it should be elsewhere.  Keynes' own statement of the EMH is 
%\begin{quotation}
%We are assuming, in effect, that the existing market valuation, however, arrived at, is uniquely correct in relation to our existing knowledge of the facts which will influence the yield of the investment, and that it will only change in proportion to changes in this knowledge. 
%\end{quotation}
%\citep[p. 152]{Keynes1936}.

%However, Keynes questioned this notion, arguing that the existing market valuation was more likely to be influenced by a variety of ephemeral, short-term fluctuations in sentiment and popular opinion with market valuations being based more on ignorance and subject to sudden fluctuations in the mass interpretation of the available information.  As a result, he felt that "the market would be subject to waves of optimistic and pessimistic sentiment" \citep[p. 33]{Keynes1936}.  One important aspect of this is the influence of public opinion and the importance of social pressures.  Keynes argues that the investment process is like a competition where the winner is the one that can guess what everyone else will choose.\footnote{The third degree "where we devote our intelligence to anticipating what average opinion expects average opinion to be" \citep[p. 156]{Keynes1936} has been achieved with 'sniffer algorithms' which try to identify the patterns used by other machines so that they can get a head start on opponents strategies.  See, for example, More or Less, 11 August 2011, BBC 4.}

\subsubsection{Behavioural finance}
In 2002 Daniel Kahneman shared the Nobel prize in economics for the work that he did with Amos Tversky. The Nobel committee cited the application of experiments and surveys to question the assumptions of economic rationality and the model of VNM expected utility and to build a more empirically-based theory of how decisions are made as a justification for the award \citep{Kahneman1979Prospect} and \citep{KahnemanThinking}.  Kahneman and Tversky present \emph{Prospect Theory} which is an alternative to the VNM expected utility theory.  The theory was developed from the results of a large number of experiments that Kahneman and Tverskey carried out to ask people what would be their decision given a choice of alternatives.  Kahneman and Tverskey then compared their results to those that would be expected from a VNM expected utility model.

Kahneman and Tversky developed a \emph{decision-weighting function} $\pi = \pi(p)$ which had a similar effect as the adjustments that Keynes makes to standard probability estimates using $q$ to cover risk aversion and $w$ to cover uncertainty. The essential difference between Prospect Theory and the VNM expected utility benchmark are that people underweigh outcomes that are probable compared to certainties, which contributes to relatively high risk-aversion in choices involving sure gains and risk-taking in choices involving sure losses \citep[p. 263]{Kahneman1979Prospect}.  In practical terms, for investors, this encourages the phenomenon of selling winners early to lock in gains and holding on to losers for too long, gambling in the face of adverse consequences.  In addition, the people making decisions in the experiments conducted by Kahneman and Tversky are concerned about changes in wealth rather than the level of wealth itself.  According to Kahneman and Tverskey, this explains why people are prepared to purchase insurance that will cover losses that are a relatively small proportion of their wealth, losses that they could presumably self-insure without the administration costs of a formal insurance company \citep[p. 286]{Kahneman1979Prospect}.

%can we cover what Shackle, Knight, Keynes and Simon said here....
More relevant for this study are the \emph{heuristics} or short-cuts to decision-making that have been identified by Kahneman, Tversky and other psychologists.   A range of investigations reveal that people deal with the problem of limited information and limited cognitive processing power by using these heuristics  that will economise on time and effort.  The heuristics allow decision to be made quickly with a limited amount of information.  However, they may also create some systematic biases in decision-making that can ensure that, in some situations, investors will make systematic mistakes in their evaluations.  If the mistakes are systematic, they should be identifiable and susceptible to exploitation by informed investors.  There are a wide range of heuristics that have been suggested, but this study will pay particular attention to \emph{conservatism} and \emph{representativeness}.

Conservatism is the tendency that people have to be cautious about changing their mind.  It is very similar to the \emph{convention} noted by Keynes in the \emph{General Theory} where he asserts that 
\begin{quotation}
``In practice we have tacitly agreed, as a rule, to fall back on what is, in truth, a \emph{convention}.  The essence of this convention - though it does not, of course, work quite so simply - lies in assuming that the existing state of affairs will continue indefinitely, except in so far as we have specific reason to expect a change."
\end{quotation}
\citep[p. 152]{Keynes1936}

There is a certain scepticism about new information and a caution about changing an established opinion. The Psychologist Edwards describes his findings relative to the benchmark of Bayesian decision-making.  
\begin{quotation}
``An abundance of research has shown that human beings are conservative processors of fallible information.  Such experiments compare human behavior to the outputs of Bayes's theorem, the formally optimal rule about how opinions (that is, probabilities) should be revised on the basis of new information.  It turns out that opinion change is very \emph{orderly}, and usually proportional to number calculated from Bayes's theorem  - but it is insufficient in amount.  A convenient first approximation to the data would say that it takes anywhere from two to five observations to do one observation's worth of work in inducing a subject to change his opinions."  
\end{quotation}
\citep[p. 359]{KahnemanUncertainty}

Different investors will interpret information in different ways and will require varying levels of evidence to convince them that they should change their opinion about the value of an investment.  This is likely to lead to a more gradual adjustment of securities prices to new information than would be expected if there were an immediate absorption of information into price as suggested by the EMH.  If investors are conservative in their decision-making, there will tend to be a gradual adjustment to information that would be consistent with trends and serial correlation in returns to securities that could be used to enhance returns.  If there are trends, there are inefficiencies in prices that can be exploited.

Another decision-making short-cut that has been identified by a number of psychological experiments is \emph{representativeness}.  This is the tendency to estimate the probability of an event by reference to its similarity to the parent population or similarity to the process that generated the parent population.  Base effects tend to be downgraded or ignored.  Kahneman presents the example of an accident that takes place in a town that has two taxi companies: 85\% of the cabs are green and 15\% are blue.  A witness identifies the taxi involved in the accident as being blue and tests of the witness reveal that she is correct in her assertion 80\% of the time.  What is the probability that the accident was caused by a blue cab?  Most people put the probability in the range of 80\% to 50\% but it is actually 41\%.

By Bayes' Rule (see Section \ref{secref:ex} for derivation). 
\begin{equation}
P(B|I) = \frac{P(I|B)P(B)}{P(I)}
\end{equation}

where $P(B|I)$ is the probability that the car was blue conditional on it having been identified as blue.  
\begin{align*}
P(I|B) &= 0.80 \text{   as there is 80\% accuracy}\\
P(B) & = 0.15 \text{   15\% blue cabs}\\
P(I|B)P(B) & = 0.80 \times 0.15\\
 &= 0.12\\
P(I|G)P(G) &= 0.2 \times 0.85 \\
&=0.17\\
P(I)&= P(I|B)P(B) + P(I|G)P(G) \text{   as B and G are the only options}\\
P(I) &= 0.12 + 0.17\\
P(B|I) &=0.41\\
\end{align*}
\citep{KahnemanCabs}

%Need more here.  Look for some psychology paper.  Is there one in the book, use taxi example?
There is too much influence from the immediate information about the witness and too little about the bigger picture.  The information about the proportion of blue and green cars is ignored.  However, Kahneman and Tverskey reveal that if the thought experiment is changed so that the participants are told that green taxis are involved in 85\% of the accidents, the information is regarded as relevant and it is incorporated into the decision-making process and probability estimates are very close to those of the Bayesian calculation.

Therefore, the representativeness heuristic also means that there is a departure from VNM expected utility outcomes and therefore some systematic biases that could be exploited by informed investors. This is consistent with the way that stock analysts over-react to new information \citep{DeBondtAnalysts} and more generally to the sort of myopic focus on the short-term that Keynes spoke about in Chapter Twelve of \emph{The General Theory} (see Section \ref{secref:noise} for fuller details). It tends to lead to over-shooting, over-reaction and the creation of bubbles.

\section{Methodology}
The aim of this study is to test whether the noise-trader model provides an appropriate explanation for speculators' activity in the  foreign exchange market and whether it can be used more generally to gain an understanding of speculative behaviour.   This will involve three steps:  the first is to assess the price implication of the noise-trader model, in particular the price performance when speculative activity becomes extreme; the second will involve measuring speculative activity in the foreign exchange market; finally, it will be assessed whether extreme speculative activity in the foreign exchange market is consistent with the noise-trader model.  If the speculative activity that is seen in the foreign exchange market produces results that are like the noise-trader model, it can be said that speculation is noise trading as Keynes asserted, if not it may be more informed as suggested by Friedman.  
%this can be referenced to the introduction.  
%The first clear outline of the term comes with Fisher Black who makes a contrast between "noise" and "information" \citep[p. 529]{BlackNoise}. Noise prevents expectations being perfect, its is the uncertainty about the consequences of monetary policy it is what is required for the smooth functioning of financial markets.  Without noise, the return on assets is known and there is no need to trade outside of liquidity requirements.   Noise traders trade on noise as if it is information.  Black offers two explanations for this phenomenon:  the first is that noise traders believe that the noise is information, and the second is that noise traders they gain utility from trading.  They may be risk-loving gamblers.  The studies by the likes of Barber and Odean, modelling and measuring overconfidence and its development, particularly amongst male day traders, support provide some empirical support for these type of explanation \citep{BarberOdean} and \citep{GervaisOdean}.    
% more may be required from Black and noise.

%This goes in the other section.  
%As argued before,  %ref?
%the idea of noise trading may be particularly important for the foreign exchange market where it is widely perceived that there are a relatively large number of traders engaged in technical analysis.  See \citep{FrootChartists} for a survey.  %this may go in the explanation for why fx is used.  Use this elsewhere. Add more?

%Noise is the vital ingredient that allows financial markets to work.  Noise is 'uncertainty' that Knight argued is essential to profits as opposed to risks that could be calculated and insured; noise is the 'uncertainty' that draws Keynes's speculators to focus on outguessing each other over the short term; noise is a consequence of Simon's \emph{Bounded Rationality}.  However, noise is also the encouragement to trade and therefore is a key contributor to liquidity and the existence of effective financial markets.

%The Grossman-Stiglitz model itself is very useful to understanding how and why information is acquired by investors and traders and what determines the relationship between informed and uninformed or noise traders.  Information is not freely available.  The price of a financial asset is a \emph{partially revealing equilibrium}.  This means that relationship between price and value contains a random element of noise.

%\begin{equation}
%V = p + \varepsilon
%\end{equation}

%where $V$ is the fundamental value, $p$ is the price and $\varepsilon$ is a normal random variable with a zero mean and a standard deviation of $\sigma_\varepsilon$.  As $\sigma_\varepsilon$ tends towards zero, price and fundamental value converge, the uncertainty and noise is removed and there is a \emph{fully revealing equilibrium}.  Therefore, the costly acquisition of information leads to market activity that pushes $p$ towards $V$ and there will be equilibrium not at the fully efficient point but where the cost of acquiring information is just equal to the benefit of its acquisition. This equilibrium will encompass informed traders who have acquired information and noise traders who rely on market price.  Noise traders will lose something to the informed but this will be, on average, equal to the cost of acquiring information; informed traders will make a return equal to the cost of information that is used.

The noise-trader model used in this study is from De Long, Shleifer, Summers and Waldeman \citep{Delong1990noise}. This is a derivative of the Samuelson overlapping generations model  with two-period agents making a portfolio choice \citep{Samuelson1958}.  There are two assets:  a safe asset $(s)$ which pays a set dividend $(d)$ and is in perfectly elastic supply; the risky asset $(u)$ also pays the same dividend but is not in elastic supply.  There are two types of investor or trader: sophisticated traders $(i)$ who know the fundamental value of the risky assets and noise-traders $(n)$ who misperceive the fundamental value.  In the absence of noise-traders, the assets are substitutes.  Each trader chooses their portfolio in the first period $(t)$ to maximise their utility on the basis of their belief about the price of the unsafe asset in the second period $(t + 1)$.  The representative noise-trader mistakes the price of the risky asset by a random variable

\begin{equation}
\rho_t \sim N(\rho^*, \sigma_{\rho^2})
\end{equation}

where $\rho^*$ is the average mistake of the noise-traders and $\sigma_{\rho}^2$ is the variance of the misperception.      Assuming an exponential utility function or constant absolute risk aversion\footnote{Though \citep{TobinLiquidity} used a quadratic utility function to show that investor preference could be defined by just the mean and the variance of the returns, the increasing absolute risk aversion this implies does not seem to be a suitable property as it implies people become more unwilling to take risk as they become more wealthy (\citep{Hicks1962} and \citep{ArrowRisk}.   Constant absolute risk aversion is the \citep{PrattCARA}, \citep{ArrowRisk}  measure based on $u(c) = 1 - e^{-\gamma c}$ where the \emph{coefficient of absolute risk aversion} $A(c) = -\frac{u''(c)}{u'(c)} = \gamma$.  This is a standard device.}, each type of investor chooses the proportion $(\lambda)$ of the risky asset to hold to maximise utility.

\begin{equation}
E(U) = \bar{w} - \gamma \sigma_{w}^2
\end{equation}

where $\bar{w}$ is end of period wealth, $\gamma$ is a measure of absolute risk aversion and $\sigma_w^2$ is the variance of wealth. Utility is a function of risk and return.  This depends on holdings of the risky asset.   For the informed trader this is,

\begin{equation}
E(U) = c_0 + \lambda_t^i[r + _tp_{t+1} - p_t(1 + r)] - \gamma (\lambda_t^i)^2 \left (_t\sigma^2_{p, t+1} \right ) 
\end{equation}

where $c_0$ is the endowment, following the notation in the original paper, $_tp_{t+1}$ is the expected price of the risky asset at $t + 1$, conditional on information at time $t$ and $_t\sigma^2_{\rho, t+1}$ is the expected variance of the misperception conditional on information at time $t$.

For the noise-trader utility is

\begin{equation}
E(U) = c_0 + \lambda_t^i[r + _tp_{t+1} - p_t(1 + r)] - \gamma (\lambda_t^i)^2 \left (_t\sigma^2_{p, t+1} \right ) + \lambda_t^i (\rho_i) 
\end{equation}

where the only difference from that of the sophisticated trader is the addition of the final term, which adds utility through the noise-traders' misperception of the expected return from holding $\lambda_t^i$ of the risky asset.

Maximising utility with respect to the share of the risky asset in the portfolio of the informed $(\lambda^i_t)$ and noise-traders $(\lambda^n_t)$ respectively,

\begin{subequations}
\begin{align}
\frac{\delta (U)}{\delta \lambda} &= 0 \\
\label{eq:NTM1}
\lambda_t^i &= \frac{r + _tp_{t+1} - (1 + r)p_t}{2\gamma(_t\sigma^2_{p,t +1})}\\
\lambda_t^n &= \frac{r + _tp_{t+1} - (1 + r)p_t}{2\gamma(_t\sigma^2_{t +1})} + \frac{\rho_t}{2\gamma(_t\sigma^2_{p,t+1})}
\label{eq:NTM2}
\end{align}
\end{subequations}

with the only difference between the two being the final term of equation \ref{eq:NTM2} which represents the misperception.

The old sell their holdings to the young so that demand of the young must equal unity. Equations \ref{eq:NTM1} and \ref{eq:NTM2} above imply that

\begin{subequations}
\begin{equation}
(1 - \mu)\lambda_t^i + \mu \lambda_t^n = 1
\end{equation}

where $\mu$ is the proportion of noise-traders to total in the market.

by substitution and re-arranging

\begin{equation}
(1 - \mu) \left [\frac{r + _tp_{t+1} - (1 + r)p_t}{2\gamma(_t\sigma^2_{p,t +1})} \right ] + \mu \left [ \frac{r + _tp_{t+1} - (1 + r)p_t}{2\gamma(_t\sigma^2_{p,t +1})} + \frac{\rho_t}{2\gamma(_t\sigma^2_{p,t+1})} \right ] = 1 
\end{equation}
\begin{equation}
1 = \frac{r + _tp_{t+1} - (1 + r)p_t}{2\gamma(_t\sigma^2_{p,t +1})} + \frac{\mu \rho_t}{2\gamma(_t\sigma^2_{p,t +1})}
\end{equation}
\begin{equation}
p_t = \frac{1}{1+r} \left [ r + _tp_{t+1} - 2\gamma(_t\sigma^2_{t+1}) + \mu\rho_t \right ]
\end{equation}
\end{subequations}

In words, the price is a function of the discounted value of the future dividend, the expected capital gain and the misperception of speculative noise-traders, less the perceived risk.  To obtain a more general solution, it is possible to substitute for future expected price, so

\begin{equation}
p_t = \frac{1}{1+r} \left [ r + \frac{1}{1 + r} \big \{  r + _tp_{t+2} - 2\gamma(_t\sigma^2_{p,t+2}) + \mu\rho_t \ \big \}  - 2\gamma(_t\sigma^2_{p,t+1}) + \mu\rho_t \right ]
\end{equation}

If this is repeated,  the price becomes the discounted value of the future dividends, the expected capital gain and the general misperception of noise-traders less risk.

\begin{equation}
p_t = 1 + \frac{\mu(\rho_t - \rho^*)}{1 + r} + \frac{ \mu \rho^*}{r} - \frac{2 \gamma}{r}\ \big (_t\sigma^2_{p,t+1} \big )
\label{eq:NP1}
\end{equation}

In words, the deviation from the true fundamental value of unity is a result of the discounted value of the noise-traders' misperception of the price $(p_t - \rho^*)$, multiplied by the weight of noise-traders in the total $(\mu)$ plus a perpetuity of the average noise-trader misperception $(\frac{ \mu \rho^*}{r} )$, adjusted for the risk.   This risk is a function of the volatility of the price which,  from equation \ref{eq:NP1} is a result of the variability of the misperception of speculative noise-traders.   All but the second term in equation \ref{eq:NP1} are constants, and if the variance of the price one period ahead  is equal to the variance of the misperception,

\begin{equation} 
 _t\sigma^2_{p,t+1} = \frac{\mu^2 \sigma_{\rho}^2}{1+r^2}  
\end{equation}

the final price equation becomes.

\begin{equation}
p_t = 1 + \frac{\mu(\rho_t - \rho^*)}{1 + r} + \frac{ \mu \rho^*}{r} - \frac{2 \gamma \mu^2 \sigma_{\rho}^2}{r(1 + r)^2}
\label{eq:NP2}
\end{equation}

De Long and the others use the utility function to compare the outcomes for the two types of investor for different realisations of the parameters.  They note four effects:  the \emph{Friedman Effect}, which is the tendency of informed investors to push prices back towards fundamental value; \emph{Price Pressure} as the way that speculative noise-traders buy more and drive the price higher as they become more bullish; \emph{Hold More},whereby the noise-traders hold more of the risky asset; \emph{Create Space} is the added risk that is caused by an increase in the variability of noise-trader belief.  The Friedman Effect and Price Pressure tend to reduce noise-trader utility as they work to create losses or reduced gains for noise-trader positions; Hold More and Create Space can increase noise-trader utility and profits \citep[pp. 14-15]{Delong1990noise}. 
%Put something on these four in the conclusion. 
A number of important ideas flow from the model.  The main findings are that noise-traders will not necessarily disappear as Friedman had suggested \citep{FriedmanPositive}.  While noise-traders will tend to get less utility than sophisticated traders, this does not necessarily drive them from the market.  The noise-traders expected to get higher utility due to their misperception.  The reduction in utility comes mainly from the fact that they are not compensated for taking risk that they are not aware of.  This is consistent with the type of over-confident, entrepreneurial misperception discussed in Section \ref{sec:prob} in relation to the returns to speculation.  If the noise-traders become aware of the risk that they are taking they are less likely to undertake the activity.

De Long and the others test a variety of specifications that include modest switching from one style of investment to another and changes in the weight of new traders after previous performance.  Some result in the disappearance of noise-traders while others suggest that they can become more dominant.  The model also explains the importance of investment with a long-term horizon as this is the only way that \emph{noise- trader risk} can be overcome.  If agents live for more than two periods, fundamental investors can make an arbitrage between safe and risky assets, driving the price back to fundamental value.  Therefore, investors or traders with a reputation that will allow resources and time to consider the long-term would be able to take full advantage of the noise-traders' misperception.  The model relies crucially on the institutional or psychological myopia. 
% read again and add more.

Noise-traders drive the price in the direction of their mis-perception.    The second and third terms in Equation \ref{eq:NP2} show the deviation from fundamental value due to the fact that the noise-trader mis-perception is not equal to zero but equal to $p^*$.  This means that the mis-perception is not random but biased in a positive fashion.  If the mis-perception is random, we are back to the random walk with noise-traders adding liquidity on each side of the market and the informed traders able to take advantage of any deviation of the price from its fundamental value.  There is still some risk that the outcome will be different from the fundamental value, because the noise-traders may still become more extreme in their misperception, but on average informed traders will make profit if they act.  This deviation in the price is what De Long and co-authors call price pressure.  While there is no fundamental risk in this model, price pressure causes risk for which there is no compensation and which is assumed by the noise-traders.  This is rather similar to the risk created by speculators in the Keynes models where additional uncertainty, in the Knightian sense, encourages more investors towards self-perpetuating myopic behaviour.

%One of the key assumptions of the model is the two period construction that means that final value is not determined.  This means that informed investors cannot just wait for the fundamental value to assert itself.  Therefore, the model is consistent with the sort of myopia that may be caused by impatience \citep{Keynes1936}, funding costs, social pressure and institutional features. % references needed.  
%\footnote{An good practical example is that of Tony Dye the Philips and Drew Fund Manager who refused to participate in the technology boom of the late 1990s.  The relative under-performance of the fund cause a loss of investors and his eventual resignation in March 2000.  This is the same date as the peak of the US technology index NASDAQ.  It is also consistent with the words of Citicorp Chief Executive Charles Prince who is reported to have said, "As long as the music's playing, you've got to get up and dance" when asked by the \emph{Financial Times} in July 2007 about the bank's lending activity.}

%is this paragraph necessary?
%The De Long and others do not go into the details of the misperception that is suffered by the noise traders.  There are a number of explanation for this deviation of the misperception from zero.  The first are the behavioural biases that were addressed in section \ref{sec:behavioural}.   If investors generally are suffering from the \emph{representativeness} heuristic an over-focus on the immediate information that the expense of the bigger picture or the base risk to securities value would cause a systematic deviation in favour or optimism or pessimism.  This effect can be exacerbated by the overconfidence that has been identified generally in the behavioural economic literature and by Odean specifically for investment markets.  Systematic bias creates price pressure which feeds back into a successful outcome that is regarded as being the consequence of skill, ability and good judgment and investment talent rather than luck and temporary circumstance.  %more may be required here Hirshliefer and Luo (2001) on the survival of noise traders.
%There is also the social pressure and a feedback mechanism that comes from taking an action that others are taking.  Therefore, if others have the same investment strategy, this can make investors more confident and, in some cases, will cause investors to infer private information from the actions of other investors who are basing investment decisions on the same style of imitation.  All the investors think that they are in good company and that the knowledge of the others fills the gaps where their own uncertainty begins. % need references for this.

\subsection{The implications of the noise-trader model}
The aim of this study is to test the implications of the noise-trader model.  There are two parameters from the noise-trader price Equation \ref{eq:NP2} that can be used to measure the amount of speculation:  the price misperception $(\rho)$ and the weight of noise-traders in the market $(\mu)$.  The aim of this section is to  identify the returns that are implied by the noise-trader model when either the intensity of the misperception or the weight of noise-traders is increased to extreme levels so that this can be used as a benchmark against which foreign exchange market activity can be judged.

To understand what the noise-trader model of speculative behaviour means for foreign exchange rates, a \emph{Monte Carlo} simulation is carried out with a variety of mis-perceptions and noise-trader weights.  Two time series are created.  The first series is based on a random noise-trade misperception, $(\rho^*)$.  This is called the Misperception Model (MPM). The second is based on a random weight of noise-traders $(\mu)$ and is called the Random Weight Model (RWM).   The simulation creates a 1000 period price series from Equation \ref{eq:NP2}. The dividend paid on the risky asset $(r)$ is posted at 10\%; the risk-aversion parameter $(\gamma)$ is assumed to be 0.3 (in line with that suggested in \citep[p. 127]{FrobesBehaviour}); starting from a price of 100, the average noise-trader mis-perception $(p^*)$ is a random variable in the first simulation with a mean of 4 and a variance $(\sigma_{\rho^*}^2)$ of 2, and has a fixed value of 4 in the second simulation; the weight of noise-traders $(\mu)$ is 50\% in the first case but is modelled as a two-period moving average of a normal random variable centred at 50\% with a standard deviation of 20\% and equally weighted average of its previous value (starting at 50\%) to allow smooth adjustment, in the second simulation.

A 1000 period prices simulation is created for each model and the returns for these series are calculated as the log change in price.  Those periods when mis-perception or the weight of speculators is \emph{extreme} are identified as those that are at or above the ninety-fifth percentile.  These extremes are the \emph{event} for an event study (see Section \ref{secref:event} for full details of the event study method) and the returns following the extreme are recorded.    The package \emph{eventstudies} \citep{eventstudies} aligns the defined events (in this case the extreme noise-trader activity) so that descriptive statistics and other analysis can be carried out on the period after the event has occurred.  
Descriptive statistics for the whole series of returns and for those when misperception or weight of noise-traders are presented in Table \ref{tabref:MC1}.  The process is then repeated for one thousand simulations of each model so that confidence intervals can be established for the estimates of the descriptive statistics.

\subsection{The results of simulation}
The results of the simulation are recorded in Table \ref{tabref:MC1}.  It shows that the mean return to an investment strategy in the whole of the sample is very close to zero.  However, for each of the extreme readings, there are large negative returns to follow. This is consistent with the structure of the model: noise-traders drive price above intrinsic value and subsequent returns will be negative; the more extreme the misperception or the weight of uninformed speculative noise-traders in the market, the greater the negative return that is likely to be recorded.

\begin{table}[h!]
\begin{threeparttable}
\caption{Results of the Monte Carlo Simulation of the Noise-Trader Model}
\begin{tabular}{p{6cm} | r r r r}
 \label{tabref:MC1}
                            & Base Case  & Extreme (MPM) & Extreme (RWM) \\
  \hline
\hline
Mean                     & -0.00 & -0.90*   & -0.49 \\
Standard Deviation  & 0.98 &  0.49*     &  0.42*\\
Skew                      &  -0.07 &  -0.16   &  -0.78*\\
Kurtosis                  &  -0.14 &  -0.38    &  -0.35\\
Number                 &  1000        &   50        & 50\\

\end{tabular}
\begin{tablenotes}
\small
\item Descriptive statistics for the De Long Noise-Trader model \citep{Delong1990noise}.  The Base Case measures the benchmark as returns from prices generated by Equation \ref{eq:NP2} for the whole sample.   The MPM measures returns after a period of extreme sentiment when the misperception of noise-traders ($\rho$) is equal to or above the 95th percentile of all the misperceptions; the RMW measures the returns following a recording of extreme weight of noise-traders ($\mu$) is equal to or above the 95th percentile of all noise-trader realisations.   The asterisk denotes significantly different from zero at the 95th confidence interval.   R Code used to create table is in Appendix \ref{AppendixB} Section \ref{secref:sim}. 
\end{tablenotes}
\end{threeparttable}
\end{table}

% this needs to be completed. 
\subsection{Measuring the intensity of speculative sentiment}
The measurement of speculative intensity must identify how strongly speculators believe that the market will move in a particular direction. In the absence of a daily survey of speculators, this study uses \emph{risk-reversal skew} as an objective measure of speculative sentiment rather than surveys of opinion.\footnote{There are a number of surveys of investor opinion that are regularly conducted by financial service information companies.  However, these are disparate and, in many cases, unsystematic, and it is not clear whether they measure speculative sentiment or investor opinion more generally.   The Wall Street Journal runs an annual survey of investor opinion, but this is focused mostly on US equities.  The \emph{Economist} magazine also solicits opinion, but in a relatively unstructured and \emph{ad hoc} fashion.  In the foreign exchange market, there is a broad set of market forecasts that was collected by the financial services company \emph{MMS International} (formerly Money Market Services) which polled a broad range of financial market participants on key exchange rates for one and four week forecast horizons. Frankel and Froot found that the median forecast from the MMS survey exhibited trend-following behaviour at the one to three week horizon while there is mean-reversion beyond that point.  Frankel and Froot also assess the 'Financial Report' that has been produced by \emph{The Economist} magazine and \emph{Currency Forecasters Digest} and find a wide dispersion of opinion and express some concern about the aggregation of such divergent forecasters \citep{FrootFrankelFDB} and \citep{FrankelFroot1987} \label{fnsurvey}.}    This is the relative price of currency put and call options for the same strike price.  The \emph{Put-Call-Parity} in the Black-Scholes option pricing model assumes that these will be the same \citep{Black1973Options}.  While the Black-Scholes model is the basis for valuing options, where the price of a non-dividend paying stock trades at a discount to the forward rate equal to the risk free rate, the forward rate in a currency option is determined by the relative interest rate differential between the two currencies.   The Garman-Kohlhagan adjustment that accounts for the difference in interest payments on the two currencies is used.   The spot discount is determined therefore by the relative interest rate differential for the appropriate maturity \citep{GarmanKohlhagan}.   This model allows the market practice of pricing options from the volatility implied by the model.  In other words, the Garman-Kohlhagan is used to extract the volatility that is implied by a specific option price given the strike and other parameters.

Determining the fundamental value of exchange rates has proved to be extremely difficult.  As PPP is, at best, a medium-term phenomenon and given the finding by Meese and Rogoff about the superiority of the random walk relative to other economic models and accepting for the moment the EMH, it will be assumed that the forward rate is the best estimate of the future value of the exchange rate.  Any forecast which deviates from the forward rate will in this case be counted as a misperception and the further away from the random walk the greater the misperception. \footnote{The expected  exchange rate should be equal on average to the forward rate once any risk premia has been take into account. Though it will be seen that UIP and the assumption that the future exchange rate is on average equal to the interest rate differential does not hold, the fact that daily data is being used here and the attention is on the most extreme divergences from UIP suggests that this will not greatly affect the results.}

This study uses a unique dataset of \emph{risk-reversals} for eleven major currency pairs.  This is the only period for which the data currently exist.  The risk-reversal data is very difficult to obtain in an over-the-counter market like foreign exchange where there is no centralised exchange to record and collect information, generally being the property of quoting institutions.  The risk-reversals are the average prices quoted at 10:00 GMT by major banks on the Thomson-Reuters composite option pages (OPT1)  for the period between January 2 1996 and September 6 2002.  The risk-reversals are the spread of implied volatility for a 1 month 25 delta at-the-money call option relative to the equivalent put.  A call option is the right to buy and a put option is the right to sell.  The delta of an option measures the effect of a change in the spot price on the value of the option.  This can range from zero for deeply out-of-the-money options to unity for deeply in the money strikes.  The at-the-money delta is 50 as there is a 50:50 chance of the option being exercised.  The benchmark for option risk-reversals is 25 delta.  This is a measure of how far the option is out of the money or how far the strike is from the forward rate.

If the exchange rate trajectory is expected to follow a random walk with drift towards the 1 month forward, implied volatility on the put and the call for the same delta should be equal.  Any deviation from this equivalence is an indication that a deviation from the random walk with drift is implied in the market price of options.\footnote{If the risk-reversal for a 1-month USD/CAD is quoted at 0.15-0.28\% and the implied volatility is 8.50\%, the market-maker would be willing to buy the 1-month 25 delta USD-put-CAD-call at 8.65\% and sell the USD-call-CAD-put at 8.50\%.  The dealer pays 0.15\%.  Alternatively, the dealer would sell a 25 delta call at 8.78\% and buy the USD-call-CAD-put at 8.50\%, earning 0.28\% \citep{Global}. }  Therefore, the relationship between implied volatility for calls and implied volatilities for puts for the same delta and maturity can be taken as an indication of how much more expensive calls are relative to the equivalent put or as the market bias towards puts over calls or, in this case, as an indication of market sentiment.  The larger the absolute level of the risk-reversal skew, the greater the relative price and, in equilibrium, the greater the belief, expressed through the action of market participants or the inventory of market-makers, that there will be a deviation from the forward rate.   If extreme deviations or speculations are noise they should presage a sharp reversal.  If speculation is informed, this is part of the price discovery process and extremes may be followed by calm conditions around a new equilibrium.

% The table may be pushed onwards.  Use t (i think). 
\begin{landscape}
\begin{table}[t]
\begin{threeparttable}
\caption{Descriptive Statistics for Daily Exchange Rate Returns: 2 January 1996 to 6 September 2002}
\begin{tabular}{l | r r  r r r r r r r r}
  \hline
 Statistic &  AUD & EURGBP & EURSEK & GBP & CHF & EURCHF & EURJPY & EUR & CAD & JPY \\
  \hline
Number & 1685 & 1675 & 1716 & 1667 & 1674 & 1674 & 1675 & 1668 & 1683 & 1675  \\
Mean & -0.0174 & -0.0187 & 0.0011 & 0.0025 & 0.0120 & -0.0039 & -0.0072 & -0.0172 & 0.0084 & 0.0076\\
Median &  0.0000 & -0.0298 & 0.0000& 0.0000 & 0.0114 & 0.0000 & 0.0000 & -0.0253 & 0.0000 & 0.0188\\
Standard deviation  & 0.7241 & 0.5542 & 0.0438 & 0.0489 & 0.6819 & 0.0277 & 0.8416 & 0.6338 & 0.0363 & 0.8170\\
Max & 5.8301 & 3.0543 & 2.1633 & 2.2573 & 3.5086 & 1.6477 & 4.0879 & 2.4136 & 2.2170 & 5.1788 \\
Min & -3.4666 & -2.2858 & -2.2900 & -2.2701 & -3.6525 & -1.8466 & -8.2848 & -2.2901 & -2.9707 & -10.6573 \\
Skewness  & 0.1206* & 0.2923* & 0.0432 & -0.0452 & -0.1423* & -0.4046* & -0.7388* & 0.1738* & -0.1095 & -1.7237*\\
Kurtosis & 4.17* & 1.88 & 2.43* & 1.48 & 1.96 & 5.84* & 7.08* & 1.12 & 5.11* & 21.52*\\
\hline
 \label{tabref:DS1}
\end{tabular}
\begin{tablenotes}
\small
\item Continuously compounded returns (multiplied by 100) for period 2 January 1996 to  September 2002 for US dollar per unit of Australian dollar (AUD), UK sterling per unit of Euro (EURGBP), Swedish Crown per unit of EUR (EURSEK); US dollar per unit of UK sterling (GBP);  Swiss franc per unit of US dollar (CHF); Swiss franc per unit of Euro (EURCHF); US dollars per unit of Euro (EUR); Canadian dollars per unit of US dollar (CAD); Japanese yen per unit of US dollars (JPY). The asterisk (*) means that the estimate is significantly different from zero at the 5\% level using t-statistics for the mean and the standard error of skew and standard error of excess-kurtosis respectively. Codes for table available in Appendix \ref{AppendixB} Section \ref{apAdes}.    
\end{tablenotes}
\end{threeparttable}
\end{table}
\end{landscape}

Table \ref{tabref:DS1} shows descriptive statistics for the exchange rate data used in the study.  There are daily data for the period between January 1996 and September 2002.  This is the period for which risk-reversal data are available.  The series are of sightly different lengths as there are some gaps in the risk reversal and exchange rate data as a result of holidays and market closures.  From the raw data continuously compounded returns are calculated and multiplied by 100.  The first thing of note is that the mean return does not appear to be statistically different from zero.  This is what would be expected if, as has been identified in many studies, exchange rates follow a random walk. This apparent stability hides some significant changes that are taking place below the surface.  For example, the Japanese yen records an appreciation of 10.66\% on one day.  Most of the exchange rate series appear to be skewed in some way and there is also evidence of kurtosis in many cases.  An asterisk is used to denote statistical significance at the level of 95\% confidence.  A t-test is used for the mean; a \emph{standard error of skew} and a \emph{standard error of kurtosis} are used for the skew and kurtosis respectively.  These are calculated as $SES = \sqrt{\frac{6n(n-1)}{(n-2)(N+1)(N+3)}}$ and $SEK = \sqrt{\frac{n^2-1}{(n-3)(N+5)}}$, with the assumption that this estimate is distributed normally, the skew or kurtosis is considered significant if it is more than two standard errors from zero.     The statistics are generated using the code in Appendix \ref{AppendixA} Section \ref{apAdes}.

\begin{landscape}
\begin{table}[t!]
\begin{threeparttable}
\caption{Descriptive Statistics for Risk Reversals: 2 January 1996 to 6 September 2002}
\begin{tabular}{l | r r  r r r r r r r r}
  \hline
 Statistic &  AUD & EURGBP & EURSEK & GBP & CHF & EURCHF & EURJPY & EUR & CAD & JPY \\
 \hline
Number & 1702 & 1697 & 1719 & 1691 & 1374 & 1339 & 1697 & 1692 & 1701 & 1697 \\ 
  Mean & -0.37 & 0.15 & 0.71 & -0.08 & -0.28 & -0.38 & -0.48 & 0.19 & 0.12 & -0.74 \\ 
  Median & -0.30 & 0.10 & 0.80 & -0.10 & -0.15 & -0.30 & -0.10 & 0.10 & 0.15 & -0.55 \\ 
  Standard Deviation & 0.38 & 0.37 & 0.51 & 0.33 & 0.49 & 0.36 & 1.02 & 0.49 & 0.25 & 1.02 \\ 
 Max & 0.45 & 1.72 & 1.90 & 1.00 & 1.05 & 0.50 & 1.70 & 2.05 & 1.45 & 2.05 \\ 
 Min & -1.70 & -0.80 & -0.30 & -1.35 & -2.05 & -1.80 & -4.03 & -1.15 & -0.65 & -4.00 \\ 
 Skewness & -0.50* & 1.27* & -0.23* & 0.33* & -0.79* & -0.58* & -0.93* & 0.68* & 0.23* & -0.51* \\ 
  Kurtosis & 0.02 & 2.29* & -1.67 & 0.72 & 0.69 & 0.19 & 0.15 & 0.96 & 1.83 & -0.03 \\ 
      \hline
 \label{tabref:DS2}
\end{tabular}
\begin{tablenotes}
\small
\item Daily risk reversal skew.  Risk reversal is the premium of implied volatility for calls relative to the equivalent put (equal distance to the strike price, 25 delta in this case).  A figure above zero means that calls are more expensive (in volatility and price terms) than puts.  A figure below zero means that puts are more expensive. The asterisk (*) means that the estimate is significant at the 5\% level using t-statistics for the mean and the standard error of skew and standard error of excess-kurtosis respectively.   These are calculated as $SES = \sqrt{\frac{6n(n-1)}{(n-2)(N+1)(N+3)}}$ and $SEK = \sqrt{\frac{n^2-1}{(n-3)(N+5)}}$, with the assumption that this estimate is distributed normally, the skew or kurtosis is considered significant if it is more than two standard errors from zero.
\end{tablenotes}
\end{threeparttable}
\end{table}
\end{landscape}

Table \ref{tabref:DS2} gives the descriptive statistics for the risk-reversal data.  The statistics are generated using the R code in Appendix \ref{AppendixB} Section \ref{apAdes}.  For a neutral risk-reversal, a reading of zero, there would be no price preference for calls or puts at the similar delta.  The figures show that, in contrast to the average returns, there are often average skews that are different from zero.  The most significant of these are the positive bias for EUR-GBP and a negative bias for USD-JPY.  The latter may be related to hedging activity by Japanese exporters.  While this could suggest that implied volatility is an imperfect measure of speculative activity, note that extreme readings are based on quantiles of the distribution and therefore an underlying hedging-induced shift in the mean of the distribution should not affect the results.  In other words, if the extreme is denoted by the 90th percentile, it does not matter if the mean is zero and this comes at 0.8 or if the mean is 0.2 and this comes at 1.0.

\subsection{Measuring the proportion of speculators}
Recording speculative activity in an over-the-counter market like foreign exchange is extremely difficult.  However, there are some exchange-traded currency derivative markets in the US, and participants in these markets are required to report their positions to the US regulators.  The Commodity Futures Trading Commission (CFTC) was set up by Congress in 1974 to regulate commodity future and options markets. The CFTC collect weekly information about the open positions held by private entities in the US derivative markets.   A sub-set of this information is released to the public each week as the \emph{Commitment of Traders Report} (CoT).  The weekly release has been in place since September 30, 1992.  Before that the data were released every two weeks.  The CoT is released each Friday and details the positions outstanding the previous Tuesday \citep{cot}.

The CFTC require traders to categorise themselves as being either commercial (C) or non-commercial (NC).  Commercial traders must have some underlying business interest in the security, commodity or instrument that they are trading.  These could be seen as natural hedgers.  Non-commercial accounts are generally considered to be speculators.  See \citep{FuturesSanders} and \citep{FuturesWang} for examples of previous use of this data in this fashion.

The administrative requirements are more onerous for non-commercial accounts but the CFTC have the right, which they use, to re-categorise parties if they are not happy with the self-selection.  Positions are either long or short, meaning that they will receive or deliver the underlying; for non-commercial accounts there are also \emph{spreading} positions which are simultaneous holdings of long and short positions in different maturities of the same contract, taking a position on the spread between two maturity dates.  These are assumed to be speculators.  The CFTC data also report \emph{Open Interest} (OI) which is the record of outstanding open positions.

If it is assumed that futures and cash markets are two separate markets and that most participants have a preference for one or the other and therefore do not frequently switch; if it is also assumed that the nature and level of speculative activity in the futures market is an indicator of what is happening in the broader spot market, it will be possible to use this data, which measures speculative activity in the futures market to assess the intensity and level of speculative activity in the foreign exchange market more generally.  In other words, the futures market is used as a sample of the entire foreign exchange market. This is important because, unlike many other financial markets, it is the cash foreign exchange market that dominates in terms of activity and liquidity.  The futures market is relatively small \citep{BISFX}.

%The operation of derivative exchanges is a competitive business that has evolved over the course of the data collection period.  The two original exchanges in Chicago, the Chicago Mercantile Exchange (CME) and the  Chicago Board of Trade (CBOT) went public in 2000 and 2005 respectively and merged in 2007 to become the CEM Group.  In 2008 the New York Mercantile Exchange (NYMEX) was added,  Most of the successful currency derivative contracts were traded on the International Monetary Market (IMM) which was a sub-division of the original CME.  Many exchange rates and cross rate products were launched by these and other US exchanges, but the five contract which maintained a presence and liquidity through the whole period were the Deutschmark and then the Euro, the Swiss Franc, the British Pound, the Japanese Yen and the Canadian Dollar.  There is a reasonably long series for the Australian Dollar and for the Mexican Peso, but nether of these is complete.  The US dollar index has been active since it was launched in May 1992 but it is not trade weighted or really representative of the US dollar.

It could be argued that this is a biased sample as it is a US market.  Though international financial markets are global in nature, there is some evidence to support this claim.  For example, there is a tendency to have a positive holding of the overseas currency as this is a hedge for exporters.  Nonetheless, as was argued with the risk-reversals, it is the relative rather than absolute position of the signal that will determine the extreme and therefore if the US sample is proportionate to the population, sample signals can still be regarded as measuring an extreme.  If the distribution of the sample itself is different from the population this is a problem, but there is no evidence that this is the case.

The data used run from September 30 1998 when the data started to be released on a weekly basis and continue up to December 31 2008.  This date is shortly before this part of the research was started. The Euro data runs  from March 3 1998 to December 31 2008.  There are two gaps in the Swiss Franc data for September 14 2004 and September 21 2004.  The absence of these data seems to be a result of problems with the CFTC database.  The contracts are for EUR 100,000; GBP 62,500; JPY 12,500,000; CHF 100,000 and CAD 100,000.  For the CoT data, two series are created:  S1 measures the net long non-commercial positions relative to total non-commercial positions and S2 measures the net long non-commercial positions relative to open interest.  Table \ref{tabref:DS3} given an overview of the descriptive statistics for the returns on the weekly foreign exchange rates that are relevant for the futures market data.  There are no significant differences from the daily data that is used in the risk-reversal part of the study recorded in Table \ref{tabref:DS1}.  Once again, the mean return is very close to zero.  This is consistent with the idea that exchange rates follow a random walk.  The distribution of returns are skewed for some pairs and often fat-tailed (identified with an asterisk where statistically significant as either a t-test or a test of the standard error of skew or standard error of kurtosis as calculated above).

 \begin{table}[t]
\begin{threeparttable}
\caption{Descriptive Statistics for Weekly Exchange Rate Returns: 30 September 1998 to 31 December 2008}
\begin{tabular}{l | r r  r r r }
  \hline
 Statistic &  EUR & JPY & GBP & CHF & CAD  \\
  \hline
Number & 522 &  849 & 849 & 847 & 849 \\
Mean & 0.0351 & 0.0337 & -0.0202 & 0.0233 & 0.0018\\
Median & 0.0650 & -0.0800 & 0.0000 & 0.0000 & -0.0200\\
Standard deviation  & 1.4063 & 1.6355 & 1.2755 & 1.5418 & 1.0558\\
Max & 8.7900 & 11.6700 & 5.7200 & 8.3400 & 10.0100\\
Min & -4.2400 & -15.0400 & -6.4100 & -5.7100 & -5.0900\\
Skew  & 0.27* & -0.06 & -0.34* & 0.31* & 0.78*\\
Kurtosis & 2.76*& 12.58* & 2.07 * & 1.82 & 11.59*\\
\hline
 \label{tabref:DS3}
\end{tabular}
\begin{tablenotes}
\small
\item Continuously compounded returns multiplied by 100 for US dollar per unit of Euro (EUR), US dollar per unit of Japanese Yen (JPY), US dollar per unit of UK Sterling (GBP), US dollar per unit of Swiss Franc (CHF) and US dollar per unit of Canadian dollar (CAD) for the period 30 September 1998 to 31 Dec 2008;     An asterisk (*) means that the estimate is significantly different from zero at the 5\% level of statistical significance using t-statistics for the mean and the standard error of skew and standard error of excess-kurtosis respectively.  These are calculated as $SES = \sqrt{\frac{6n(n-1)}{(n-2)(N+1)(N+3)}}$ and $SEK = \sqrt{\frac{n^2-1}{(n-3)(N+5)}}$, with the assumption that this estimate is distributed normally, the skew or kurtosis is considered significant if it is more than two standard errors from zero.
\end{tablenotes}
\end{threeparttable}
\end{table}

Table \ref{tabref:DS4} shows the descriptive statistics for the two measures of speculative sentiment.  The first (S1) is measured as the net non-commercial long and non-commercial short proportionate to the sum of non-commercial positions long and short positions; (S2) is measured as the net non-commercial long and spreading positions proportionate to open interest.  The first captures directional speculation and, as can be seen from a number of minimums, this can reach 100\%.  The second records the weight of speculation relative to everything else in the market.  The former is more sensitive, as can be seen by the range and the standard deviation; the second is usually above zero because there are always some spreading positions that seek to take advantage of calendar spreads and other strategies.

\begin{landscape}
\begin{table}[t]
\begin{threeparttable}
\caption{Descriptive Statistics for S1 (Speculative Sentiment) and S2 (Speculative Weight): 30 September 1998 to 31 December 2008}
\begin{tabular}{l | r r r r r r r  r r r }
  \hline
 Statistic &  \multicolumn{2}{c}{EUR} & \multicolumn{2}{c}{JPY} & \multicolumn{2}{c}{GBP} & \multicolumn{2}{c}{CHF} & \multicolumn{2}{c}{CAD}  \\
 & S1 & S2 & S1 & S2 & S1 & S2 & S1 & S2 & S1 & S2 \\  
\hline
Number & 522 &  522 & 849 & 849 & 849 & 849 & 847 & 847 & 849 & 849\\
Mean &  31.31 & 43.43 & -23.90 & 44.21 & 2.76 & 41.68 & -21.14 & 46.13 & 6.68 & 41.23\\
Median & 42.75 & 39.56 & -33.30 & 43.56 & 11.15 & 38.88 & -37.32 & 42.88 & 7.44 & 39.81\\
Standard deviation  & 41.94 & 16.13 & 51.49 & 12.17 & 60.59 & 16.46 & 60.76 & 15.73 & 55.26 & 15.33\\
Max & 95.67 & 86.06 & 87.43 & 84.50 & 96.15 & 89.45 & 100.00 & 92.49 & 100.00 & 86.63\\
Min &  -100.00 & 11.15 & -99.83 & 12.88 & -100.00 & 9.10 & -100.00 & 14.61 & -90.08 & 11.89\\
Skew  & -0.88* & 0.55* & 0.33* & 0.57* & -0.17 & 0.76* & 0.45* & 0.65* & 0.02 & 0.50*\\
Kurtosis & 0.18 & -0.57 & -1.14 & 0.64 & -1.45 & 0.15 & -1.25 & -0.15 & -1.35 & -0.28\\
\hline
 \label{tabref:DS4}
\end{tabular}
\begin{tablenotes}
\small
\item S1 is the first sentiment indicator, calculated as ratio of net non-commercial long less non-commercial short to the sum of non-commercial long and short positions multiplied by one hundred.  S2 is the second sentiment indicators, calculated as the ratio of net non-commercial long to net non-commercial short plus spreading to open interest.   An asterisk (*) means that the estimate is significantly different from zero at the 5\% level using t-statistics for the mean and the standard error of skew and standard error of excess-kurtosis respectively.  These are calculated as $SES = \sqrt{\frac{6n(n-1)}{(n-2)(N+1)(N+3)}}$ and $SEK = \sqrt{\frac{n^2-1}{(n-3)(N+5)}}$, with the assumption that this estimate is distributed normally, the skew or kurtosis is considered significant if it is more than two standard errors from zero.
\end{tablenotes}
\end{threeparttable}
\end{table}
\end{landscape}

\subsection{Event studies}
The event studies method ???? of how the extreme speculative sentiment and extreme speculative positions in the DeLong model could be used to analyse future returns.  Fuller explanation is provided here. Event studies are used to find the effect of a particular event on the performance of an asset price or security.  It is often suggested that James Dolley was the first to employ the event study when looking at the effect of a stock split on the price of securities.\footnote{A stock split is when the number of shares are doubled and the value of each share is cut in half.  This is usually done to prevent the price or cost of each share rising too far beyond the means of the small investor.}   Dolley took a sample of 174 stocks splits for all common stocks listed on the New York Stock Exchange during the period 1921 to 1931, and using what he called a case study, examined the process of stock splitting.  Price information was taken from \emph{Poor's Manual} and \emph{Moody's Manual} as well as the \emph{New York Times}; a survey was sent to all the companies that were involved and (remarkably) 63 of the 111 questionnaires were returned completed.  The analysis of the price change was a very small part of the study.  Dolly found that of the 95 stock splits that he had been able to compile a time series for, 57 subsequently rose and 26 fell \citep{Dolly1933}.

However, Dolly only looked at the market reaction the day after the stock split.  This was some time after the stock holder meeting that would have ratified the decision of the management. Myer and Bakay argued that, if stocks tend to rise after a stock split, it would be best to buy the stocks when the announcement is made and before the split takes place and that, therefore, the price adjustment is likely to be spread over a number of days.  They decided to measure the price action from a point eight weeks before the formal acceptance of the decision by the stock-holders. They measured the price change between the base date and the meeting and tried to account for other influences on the price by recording the ratio of the stock price to the Standard and Poor Industry average, thereby providing the first attempt to separate the specific returns from those of the market.  An additional measure of the price ratio was also taken eight weeks after the formal meeting.  For the splits that occurred between 1945 and 1946 they found that 62 of the 70 stocks exhibited a positive price performance between the base date and the meeting, during the subsequent time there was some reversal with only 55 of the stocks still showing a net gain between the base and the closure point eight weeks after the announcement meeting \citep{MyersBakay}.

Fama, Fisher and Roll use the event study method to conduct a much more systematic assessment of the effect of stock splits on security returns \citep{FamaFisherJensenRoll}.    Fama later said that the method was secondary to his desire to show off the new dataset that he had helped to construct.  This was one of the first uses of the Center for Research in Security Prices (CRSP) data.  Nonetheless, the Fama study is considered to be the benchmark of Event Studies and it is the framework that is largely followed in the current work.

The event is identified and the time at which the event occurs is marked as time zero (or $t_0$).  Returns to the reference security in a window before and after the event are recorded and compared to the normal or expected returns.  Therefore, 
\begin{equation}
AR_{i \tau} = R_{i \tau} - E[R_{i \tau} | X_{\tau}]
\end{equation}
Where $AR_{i \tau}$ are the abnormal returns for security i at time $\tau$, $R_{i \tau}$ are the actual returns to security i at time $\tau$,  $E[R_{i \tau}]$ are the expected returns for security i at time $\tau$ conditional on the information $X_{\tau}$ and $\tau$ is the time relative to the event.  Estimating the normal or expected returns is one of the key challenges of an event  study.  It is usually tackled in one of two ways:  the security can be assumed to take a mean return or, as in the case of the paper by Fama and the others, a market model can be used.  The original market model just takes a constant relationship between security returns and the market estimated by a linear regression.  Subsequent models have added additional factors.

The mean return is calculated as 
\begin{equation}
\mu_{R, i} = \frac{1}{n}\sum_{j = 1}^{j = n} R_{i j}
\end{equation}

where $R_{i j}$ is the return to security i at time j and n is the number of total number of observations to that 
\begin{equation}
E[R_{i \tau} | X_{\tau}] = \mu_{R, i} + \varepsilon_{i, \tau}
\end{equation} 
with $\varepsilon_{i, \tau}$ being a random variable that is assumed to be normally distributed with a zero mean and a standard deviation of $\sigma_{R, i}$. 
The simple market model is estimated by 
\begin{equation}
R_{i, j} = \alpha + \beta_i R_{m, j} + \varepsilon
\end{equation}

Where the $\beta_i$ is the relationship between the market return $R_{m, j}$ and the return on the security of interest $R_{i, j}$ and $\varepsilon$ is a random variable that represents all the idiosyncratic factors that affect the returns to the security.

The lack of any clear benchmark for currency returns and the fact that theoretically exchange rate returns should be on average equal to the forward rate, have made the case for a mean model.  The seminal event studies have used equities and a market model.  The expected return is zero and any returns that differ from that are identified as abnormal. Theoretically, the forward rate should be used but given the high frequency of the data, the interest rate differential is ignored.   The evidence about returns in Tables \ref{tabref:DS1} and \ref{tabref:DS3} suggests that this is not likely to be a major problem.

\subsection{Analysis of events and abnormal returns}
The aim of the study is to assess the effect of extreme speculative positions on future returns to ascertain the informational content of these extremes.  Therefore, the events that will be used to assess abnormal returns are the extreme readings on sentiment and the extreme presence of speculators.   Any reading above the ninety-ninth percentile or below the first percentile is deemed to be extreme.  For the risk reversal skew, as a secondary test, a reading above the ninety-fifth and below the fifth percentile is also established as an event.   The Event Studies R package was used to manage the data see \citep{eventstudies} for fuller details on the package and its use.  The cumulative abnormal returns for the whole of the event window (CARW) and the cumulative abnormal returns for after the event (CARA) were recorded.  These cumulations are just the summation of the continuously compounded returns over the appropriate period.

  The date of the events, 1st, 5th, 95th or 99th percentile, reading of the risk-reversal skew or the level of speculative positions,  are identified and the exchange rate returns for the window around the event are aligned so that the average for each event and each event day can be calculated.  For example, in Table \ref{tabref:RR1} the analysis of the event that the risk-reversal skew for the EUR vs the US dollar is an extreme low (measured as being equal to the 5th percentile or lower), all 106 cases of this extreme for the period under study are aligned and the average cumulative returns for the whole window (CARW - one day before and after the event in column four) and the post-event window (CARA - the event and the day after the event) are calculated.   A \emph{bootstrap} is conducted.  This means that a sample of size 106 (in this case), with replacement, is drawn from each of the event days and new means are calculated for the CARW and CARA.   If 95\% of the means that are calculated after drawing random samples from the array of windows are above or below zero, it is assumed that the figure is statistically different from zero, and this is represented by an asterisk in the table.

\section{Analysis of the results}
Table \ref{tabref:RR1} shows the result of the event studies that were carried out on extreme risk reversal.  
For each exchange rate the second column distinguishes between extreme high and extreme low, the third  column identifies the number of extreme events that were recorded and the next two are the cumulative average returns for the whole window (CARW) and the cumulative average returns for the period of the event and afterwards (CARA).  The windows were chosen to correspond to one day and one week of trading.  However, Figure \ref{fig:ES1} makes it clear that altering the size of the event window does not change the findings to any significant degree.

Looking down the seventh column (the 4-day CARW) shows the average cumulative return for the period from four days before the extreme event to the four days after the extreme for extreme sentiment highs and lows of each currency pair.   The extreme highs are (in this case) those points when the risk-reversal skew is at or above the 95th percentile.  It should not be surprising to see that the extreme highs are associated with positive returns, which appear in many cases to be significantly different from zero (where significance is identified as more than 95\% of 1000 random bootstrap samples taken from the events being above zero).  
It is clear that, as speculators become more extreme in their opinion, prices are driven in the direction of sentiment.  It shows that the spot price is following the option risk-reversal skew and, as Figure \ref{fig:ES1} reveals, it would be even more pronounced if the event window just ran from the starting point to the event.  The findings here are consistent with the evidence that orders and sentiment drive prices \citep{Evans2002Order}, \citep{FuturesSanders} and \citep{FuturesWang}.

What is more interesting is whether knowledge about the extreme reading at time $t = 0$ provides any information about future returns.  If it is known that speculative sentiment is at the extreme, does this mean that it is more likely that there will be a reversal? Is there an inefficiency?  This information is provided in columns 5, 8, 11 and 14 of Table \ref{tabref:RR1}. A reversal is indicated by a significant negative reading for highs or positive reading for lows.   Looking, for example, at the four-day window in column 8, it is clear that the extreme risk reversal skew does not provide any information about future returns.  Of the eighteen cases using the 95\% th percentile to identify an extreme, ten show a continuation, rather than a reversal, that is not significantly different from zero; one shows a significant continuation; five show reversals that are not significant; only two show significant reversal (EURCHF in each case).  For the more extreme event when the 99\% the percentile is used to measure the extreme, there is an even more mixed picture.  Just over half the cases with the four-day window show a reversal (8 of 18), but only 2 of these are statistically significant.

\begin{landscape}
% multirow package is required.   
%pdflandscape is required to facilitate the landscape position. 
\begin{table}[t]
\begin{threeparttable}
\caption{Event Study: Cumulative Abnormal Returns and Extreme Risk Reversal}
\begin{tabular}{llccccccccccccc}  
 % \hline
& &\multicolumn{6}{c}{5th and 95th percentile} & \multicolumn{1}{c}{} & \multicolumn{6}{c}{1st and 99th percentile}\\
 & &\multicolumn{3}{c}{1 day} & \multicolumn{3}{c}{4 day} & \multicolumn{1}{c}{} & \multicolumn{3}{c}{1 day} & \multicolumn{3}{c}{4 day} \\ 
%\cline{4 - 10} \cline{13 - 19}
& &\multicolumn{1}{ c}{N} & \multicolumn{1}{c}{CARW} & \multicolumn{1}{c}{CARA} & \multicolumn{1}{ c}{N} & \multicolumn{1}{c}{CARW} & \multicolumn{1}{c}{CARA} & \multicolumn{1}{c}{} & 
\multicolumn{1}{ c}{N} & \multicolumn{1}{c}{CARW} & \multicolumn{1}{c}{CARA}  
& \multicolumn{1}{ c}{N} & \multicolumn{1}{c}{CARW} & \multicolumn{1}{c}{CARA}  \\   
\hline
\multirow{2}{*}{EUR} 
& Hi &  87 &  0.4103* &0.1586 &87 &0.9442* & 0.1505 &&22 &0.0309 & -0.1748 &22 & 0.5781*& -0.5387*  \\ 
& Lo & 106 &-0.3313* & -0.0933 &106 &-0.0816* &-0.1991 & & 21&0.3040 &0.1820 &21 &0.7935* & 0.4377  \\
\multirow{2}{*}{JPY}
& Hi & 90 &0.4860* &0.1341 &90 &1.5158* &0.2531 & &20 &0.8227* &0.3078* &20 &2.4337* & 0.8284*  \\ 
& Lo & 90 &-0.3168 &0.0945 &90 &-1.3532* &0.1984 & &17 &-1.6990* &-0.9347* &17 &-3.6751* & -1.2556*  \\
\multirow{2}{*}{GBP}
& Hi & 96 &0.3303* &0.1903* &96 &0.8966* &0.2547 & &21 &0.5252* &0.1774 &21 &1.3131* &0.3284  \\ 
& Lo & 97 &-0.1894 &0.0312 &97 &-0.8031* &0.1500 & &24 &-0.2630 &-0.0866 &24 &-0.6660 &0.1888  \\
\multirow{2}{*}{CHF}
& Hi & 80 &0.0065 &-0.0565 &80 &0.5624* &0.1322 & &18 & -0.0022 & -0.1518 &18 &0.4887 &-0.1768  \\ 
& Lo & 75 &-0.3716* & -0.1125& 75 &-1.0071* &-0.0219 & &18 &0.0064 &0.1704 &18 &-0.8661* &0.7631  \\
\multirow{2}{*}{CAD}
& Hi & 113 &-0.0038 &-0.0489 &113 &0.1872 &-0.1951 & & 24 &-0.1081 & -0.0462 &24 &0.3935* & -0.1592  \\ 
& Lo & 97 &-0.1422* & -0.0571 & 97 & -0.2896* & 0.0035& & 26 & -0.2008 & -0.0146 &26 &0.1185 & 0.4560*  \\
\multirow{2}{*}{AUD}
& Hi & 101 &0.2539* &0.0125 &101 &0.8728* & 0.0875 &  &33 &0.5099* &0.0682 &33 & 1.1501*& 0.1046  \\ 
& Lo & 138 &-0.4384* & -0.1711 & 138 & -1.14935* & -0.2714* & &20 &-0.4334 & 0.0563 &20  & -1.7360* & 0.1914  \\
\multirow{2}{*}{EURJPY}
& Hi & 95 &-0.0504 &-0.1103 &95 &-0.0128 &-0.2579 & &23 &-0.2932 &-0.4188* &23 &-0.8203 & -1.0195  \\ 
& Lo & 97 &-0.4746* & -0.1647 &97 &-1.6757* &-0.4595 & &20 &-1.3630* & -0.6976* & 20 & -3.6849* &-1.1999*  \\
\multirow{2}{*}{EURGBP}
& Hi & 91 & 0.1544 &0.0696 &91 &0.5176* &0.0979 & &28 &0.2048 &0.1195 &28 & 0.4146 & -0.0645  \\ 
& Lo & 122 & 0.0451 & 0.0996 &122 &-0.3097* &0.2903* & & 21 &-0.07523 &-0.0780 &21 & -0.4120 &
-0.1290  \\
\multirow{2}{*}{EURCHF}
& Hi & 77 & 0.1183* & 0.0690 &77 & 0.3444* & 0.0749& & 	31 & 0.1482* &0.0739  &31 & 0.4993* 
& 0.1816*  \\ 
& Lo & 94 &-0.0344 & 0.0194 & 94 & -0.1596 &0.1952* & &24 & -0.3626* &-0.1483 & 24 & -0.7573* 
& 0.0875  \\
\hline
\label{tabref:RR1}
\end{tabular}
\begin{tablenotes}
\small 
\item Where Hi means extreme high in the risk-reversal skew and Lo means extreme low in the risk reversal skew; extreme is calculated as being equal or above the 95th or 99th percentile or equal or below the 5th or 1st percentile respectively; CARW is the cumulative abnormal return for the whole window, before and after the extreme event, and CARA is the cumulative abnormal return for the period after the event, which is the event day and the window; abnormal return is anything that is different from zero; the asterisk denotes significantly different from zero, where statistical significance means that more than 95\% of 1000 random bootstraps from the events are above or below zero respectively.   
\end{tablenotes}
\end{threeparttable}  
\end{table}
\end{landscape}
%This is a section on the methodology that I am writing now (while I remember) so that I can place where necessary.  
%================Event Study Method=====================================
% The dates when the risk reversal skew is at or above the ninety fifth (ninetieth) or at or below the first (tenth percentile are identified and a matrix of returns is created with the rows being equal to plus and minus the size of the event window and the date that the extreme was recorded being equal to zero.  The number of columns is equal to the number of events which, given that we are looking at the extreme one or five percent, will be related to the sample size.  The average of the returns for each event day are then recorded.  As such, with an event window of four days, there is an average return four days before the event window and for three days before the event window, through to four days after the event window.  To get an idea of the precision of this estimate, the returns from each event day are sampled (with replacement) one thousand times and the mean is calculated.  The dispersion of the resulting means provide the confidence intervals for the estimate of the mean returns for each event day that has been made.

The results do not appear to be a function of the percentile that is used to measure the extreme event or the size of the window that is selected. Additional cases were run using percentiles at 80\% and 99.5\%.  For the event window, figures \ref{fig:ES1} and \ref{fig:ES2} show the cumulative abnormal returns for the major exchanges rates relative to the US dollar across a 16 day event window.   Figure \ref{fig:ES1} shows the extreme highs for as a $99^{th}$ percentile and figure \ref{fig:ES2} shows the extreme low for as a $10^{th}$ percentile.  The figures reveal the mean cumulative return (in solid red) and the 95\% confidence intervals (dashed blue).  The only case where there seems to be any evidence of a reversal is for the extreme low risk-reversals on the Canadian dollar.

\begin{figure}[h!]
\graphicspath{{../Figures/}}
\centering
\caption{Event Study:  Extreme (High 99th percentile) Risk Reversal Skew and 16 day event window}
\label{fig:ES1}
\includegraphics[scale=0.8]{RRCum16}
\end{figure}

Again, the most prominent observation is positive association between sentiment and returns.  Intuitively, this is not surprising.  However, it does suggest that foreign exchange prices are not always driven by a random walk. There is clear evidence here of serial correlation in the returns. As sentiment rises (as measured by the increase in the risk-reversal) there is a price appreciation. However, the adjustment is not instantaneous.  This is an example of the sort of trending behaviour that is common in the literature (as reported in Section \ref{secref:empiric}).  It is likely to be at least partially explained by the sort of gradual adjustment to information that was identified as being a consequence of the conservatism heuristic in Section \ref{sec:behavioural} on behavioural finance.

\begin{figure}[h!]
\graphicspath{{../Figures/}}
\centering
\caption{Event Study:  Extreme (Low - 10th percentile) Risk Reversal Skew and 16 day event window}
\label{fig:ES2}
\includegraphics[scale=0.8]{RRCum16a}
\end{figure}

However, it does indicate that the evolution of exchange rates are not a random walk.  There is evidence here that prices gradually respond to information, as would be the case if decision-making is following a \emph{conservatism} heuristic or if there were institutional features.  The positive (or negative) expectations that are evident in the risk reversal skew do not seem to represent a deviation or a mis-pricing that will be reversed.  Rather they seem to be part of the price-discovery process.

The second heuristic that was discussed in Section \ref{sec:behavioural} was representitiveness.  This was the tendency of decision-makers to focus on the immediate and to forget base probabilities.  It would suggest that investors would think too much about the new information that has driven prices towards a new equilibrium in Figures \ref{fig:ES1} and \ref{fig:ES2} not sufficient to the broad background where other currencies may be more or less attractive.  In practical terms, it suggests that there will be overshooting and that this overshooting will cause reversals when it becomes extreme.  There is little evidence here of any reversal.  After the extremes are reached, there seems more likely to be a period of consolidation or a fluctuation around the new equilibrium.  This is what would be expected if speculation were part of the process of getting information into the price. Once this has taken place, there is consolidation.

\begin{landscape}
% multirow package is required.   
%pdflandscape is required to facilitate the landscape position. 
\begin{table}[ht]
\begin{threeparttable}
\caption{Event Study: Cumulative Abnormal Returns and Extreme Speculative Positions}
\begin{tabular}{llccccccccccccc}	
 % \hline
& &\multicolumn{6}{c}{S1 - Net Long per Speculators} & \multicolumn{1}{c}{} & \multicolumn{6}{c}{S2 -Net Long per Open Positions}\\
 & &\multicolumn{3}{c}{2 week} & \multicolumn{3}{c}{4 week} & \multicolumn{1}{c}{} & \multicolumn{3}{c}{2 week} & \multicolumn{3}{c}{4 week} \\ 
%\cline{4 - 10} \cline{13 - 19}
& &\multicolumn{1}{ c}{N} & \multicolumn{1}{c}{CARW} & \multicolumn{1}{c}{CARA} & \multicolumn{1}{ c}{N} & \multicolumn{1}{c}{CARW} & \multicolumn{1}{c}{CARA} & \multicolumn{1}{c}{} & 
\multicolumn{1}{ c}{N} & \multicolumn{1}{c}{CARW} & \multicolumn{1}{c}{CARA}  
& \multicolumn{1}{ c}{N} & \multicolumn{1}{c}{CARW} & \multicolumn{1}{c}{CARA}  \\   
\hline
\multirow{2}{*}{EUR} 
& Hi &  27 &  2.9415* & 1.3163* &27 & 3.8059* & 1.7341* &&27 & 1.0974* & 0.2896 &27 & 1.7667*& 0.0419  \\ 
& Lo & 27 & -2.4258* & -1.2512* & 27 &-3.9235* &-1.9254* & & 27 & -2.5122* & -1.0607* & 27 & -4.4478* & -1.9285* \\
\multirow{2}{*}{JPY}
& Hi & 43 & 1.9749* & 0.7286* & 43 & 2.8628* & 0.7909 & & 43 & 2.2463* & 1.0554* &43 & 3.7125* & 1.4443*  \\ 
& Lo & 43 & -2.2014* & -1.1312* & 43&-3.2488* & -1.5758* & &43 & -2.0605* & -0.9791* &43  & -2.6865* & -0.8318  \\
\multirow{2}{*}{GBP}
& Hi & 43 & 0.7102 & -0.2116 &43 & 1.0172* & -0.7902* & &43 & 1.2916* & 0.2370 &43 &1.6912* &0.0126  \\ 
& Lo & 43 &-0.6979* &0.2902 &43 &-1.5330* &0.1014 & &43 &-1.8993* &-0.1415 &43 &-2.8830* & -0.0768 \\
\multirow{2}{*}{CHF}
& Hi & 43 & 1.4400* & -0.0988 & 43 & 2.8665* & -0.0993 & & 43 &  2.5260* &  0.5074 & 43 & 3.6060* & 0.2158  \\ 
& Lo & 43 & -2.1816* &  -0.9770* & 43 &-3.3030* & -0.09798* & & 43 & -1.2605* & -0.3070& 43 &-0.1.6021* & 0.0670  \\
\multirow{2}{*}{CAD}
& Hi & 43 & 1.8658* & 1.0195* & 43 & 3.5298* & 1.8319* & & 43 & 2.0160* & 0.8788* &43 & 3.0995* & 0.9921*  \\ 
& Lo & 43 &-1.3238* & -0.5548* & 43 & -1.9151* & -0.7859* & & 43 & -1.2462* & -0.4505*  &43 & -1.9390*  & -0.6195*  \\
\hline
\label{tabref:SP1}
\end{tabular}
\begin{tablenotes}
\small 
\item Where Hi means extreme high in the sentient index, either S1 which is the measure of net long non-commercial (speculative) positions per number of speculators) or S2 which is the measure of net long non-commercial (speculative) positions per open interest (total outstanding open positions); high and low are above the 95th percentile or below the 5th percentile respectively; CARW is the cumulative abnormal return for the whole window, before and after the extreme event, and CARA is the cumulative abnormal return for the period after the event, which is the event day and the window; abnormal return is anything that is different from zero; the asterisk denotes significantly different from zero, where statistical significance means that more than 95\% of 1000 random bootstraps from the events are above or below zero respectively.   
\end{tablenotes}	
\end{threeparttable}  
\end{table}
\end{landscape}

Table \ref{tabref:SP1} shows the results from the event studies that were carried out using speculative positions as reported to the US regulators.  The table works in a similar way to Table \ref{tabref:RR1}.  The first column is the exchange rate, defined as units per US dollar.  The table is then split into two parts.  The first part uses the measure S1 which is the net non-commercial (speculative) positions relative to total non-commercial (speculative) positions; the second is S2 as the net non-commercial position relative to open interest or the total outstanding open positions.  Therefore, the first captures the balance of speculative positions while the second captures the balance of speculative positions relative to the whole market.  If speculators become more prevalent or dominant in the market, this will show up in S2 but not S1.  S1 is more sensitive to changes in sentiment and more volatile.\footnote{When the same data are used in the model for international capital flows in ref, it is found that S1 tends to provide a better fit for the module of capital flow and the real trade weighted index.} Each of these sections is broken into the results for a 2 week window and a 4 week window.  Each of these has the same pattern.   There is a row for extreme high and extreme low which represents extreme long positions or extreme short positions; the first  column identifies the number of cases; CARW is the cumulative abnormal return for the whole event window; CARA is the cumulative abnormal return for the period after the event to the end of the window.

\begin{figure}[h!]
\graphicspath{{../Figures/}}
\centering
\caption{Event Study:  Extreme (95th percentile), Non-commercial net per OI (S2), 6 week event window}
\label{fig:ES3}
\includegraphics[scale=0.8]{FPCum6w}
\end{figure}

Once again, there is a very strong association between returns and the movement of speculators.  For the 4-day window and the S1 and S2 measures, every one of the currencies studied, high and low, had a mean cumulative return for the whole of the event window that was in the same direction as sentiment.  When re-sampled a thousand times, more than ninety five percent of the means calculated were greater than zero in all cases but one (GBP).  Once again there is evidence here of momentum and of speculator positions following the movement of prices.

However, there is little evidence that these speculators are driving prices to extremes with their extreme behaviour.  Sixty percent of the S1 cases still show a significant abnormal return in the direction of the extreme 4 weeks after the extreme has been reached; only thirty percent of the cases show a reversal and in only one of those do ninety five percent of the re-sampled means remain above zero.  For S2 it is even more extreme, with ninety percent of the cases showing a continuation and forty percent of them being significant.  Again the evidence is consistent with the gradual adjustment of prices to information.

\begin{figure}[h!]
\graphicspath{{../Figures/}}
\centering
\caption{Event Study:  Extreme (90th percentile), Non-commercial net long (S1), 6 week window}
\label{fig:ES4}
\includegraphics[scale=0.8]{FPCum6wa}
\end{figure}

The R code for all the event window figures is available in Appendix \ref{AppendixB} Section \ref{Asecref:event}.  This merely takes the data that was used to construct Tables \ref{tabref:RR1} and \ref{tabref:SP1} and puts it in graphical form.  The event windows are stretched to give a clear view of the evolution of exchange rates  around the event.  Confidence intervals are added.  Once event objects FX.exc have been created, parameter for exchange rate, extreme measure, quantile, window length and the high or low indicator can be changed for alternative assessments of speculative extremes. Adjustments to these parameters show that these results are not a function of the choices presented in the main text.  They are robust to alternative specifications.

Figure \ref{fig:ES3} shows the performance of the cumulative abnormal returns around a six week event window where the event is the net low (short) speculative positions relative to total activity (open interest) at or below the ($5\%)$ level.  This means that the speculators are inclined to be short the non-domestic (US dollar) currency.  While the Euro continues to move lower after the extreme short positions, the other major currencies seem to reveal a period of consolidation that is consistent with the settlement around an equilibrium after a period where information is being absorbed.  The contrast between the Euro and the other currencies may be associated with the fact that there is a smaller dataset for the Euro due to its more limited lifespan.  Figure \ref{fig:ES4} shows the abnormal returns around extreme high (90th percentile and long position).  There is some tentative evidence of continuation for speculative positions in the Canadian dollar and signs of reversal for UK sterling.  However, the Euro and the Japanese yen move sideways after the event.  There is no clear pattern after the extreme and therefore no information.

\section{Speculation and foreign exchange returns}
This chapter has sought to measure speculation and to use this measure to discover more about the relationship between speculation, information and asset prices.  The measurement of speculation takes two forms:  a measurement of the intensity of belief with the risk-reversal skew, and a measurement of speculative positions.  Neither of these is ideal, in the sense of being a random sample from a population.  However, the option prices reveal the intensity of feeling through the weight of money in a speculative part (options) of a speculative market (foreign exchange).  The commitment of traders report records only speculators in the US futures market.  However, this is an important market, there is no reason to believe that US or speculators in the futures market are systematically different from those in the rest of the world and presumably if there were a major difference there would be an arbitrage opportunity between the two markets.

 there is strong evidence that speculation affects prices and that this effect is gradual rather than instantaneous.  This suggests a market inefficiency that would reward momentum-based trading.  It is no wonder therefore that \citep{TaylorTechnical}, reporting a survey conducted by the Bank of England, found that of the Chief foreign exchange dealers contacted in London in November 1988, 90\% of them attached some weight to \emph{technical analysis}, a method that is fundamentally based on finding trends in asset prices.  The survey found that the importance of technical analysis became greater at shorter horizons.  This has not disappeared.  The importance of momentum trading and technical analysis has not waned with time.  Neely and Wller  show that gains from technical trading rules still generate excess returns and attribute this to behavioural biases \citep{NeelyWeller}. It seems that speculation forms part of the process of price discovery, but this process is not instantaneous as would be expected in the purest form of the EMH.

The notion that speculation is at least partially informed is supported by the fact that there is little evidence from the event studies conducted here that speculation causes overshooting or bubbles.  When speculation is extreme, whether measured by intensity of belief or weight of speculative positions, future returns are as likely to be positive as negative.  There is no informational content as there would be if speculation was pure noise.  In the simulation of the De Long model, extremes are more likely to be followed by reversals.  In the foreign exchange market, that is not the case.  There are some large price crashes in the dataset, as shown by the descriptive statistics of the returns. For example  in Table \ref{tabref:DS1} it is seen that on one day there is a 10.65\% appreciation of the Japanese yen and Table \ref{tabref:DS2} shows a week when the Canadian dollar falls by 10\%.  However, the measures of speculative sentiment or the weight of speculation are not particularly high in either of these cases.  Indeed, an analysis of the top 1\% of positive and negative foreign exchange returns in the dataset reveals that sentiment and positions for this group were very close to the average of the whole set.  This would be consistent with a model where extreme price movement was a function of new information rather than the structure of existing positions.    
%expand.  Could put data in appendix.

The evidence here, that speculation is associated with serial correlation in the returns but cannot be used in a systematic way to determine reversals, has important implications for speculative activity.  It means that the largest returns will be to those that identify the trend at the earliest opportunity.  This means that speculators must be prepared to act as swiftly as possible, probably when there is a high level of uncertainty and a lack of information and where any scrap of additional news is highly valued.   The risk that actions by other speculators, most notable by their effect on price, will be construed to contain information exacerbates the likelihood of feedback effects.  The need to be able to take action before others has reached the most extreme form in the activity of \emph{high frequency trading} (HFT), where decision times are measured in nanoseconds and there is advantage in gaining spacial proximity to exchanges.  HFT is all about getting  a larger share of the inefficiency in pricing information by trading or investing closer to where the information is released.

These findings have implications for speculators, economic modellers and policy makers.  Speculators may jump on the trend, but they do not know by looking at how popular that trend is when to get off.  They need other signs to tell them that the trend has reached its conclusion or equilibrium has been achieved.  For the modellers, there is confirmation of the need  for feedback effects and evidence that reversals cannot be systematically identified from high levels of speculative positions.  It may suggest another trigger for bursting bubbles; it may, for example, suggest that this reversal has to be modelled as stochastic element or pure shock. For policy makers, there is no escape from the requirement to look at economic fundamentals when deciding whether bubbles are building in asset markets.  The presence of speculators is no guarantee that there is excess.  The speculators may be just part of the adjustment to new information.   However, the combination of high level of speculation and divergence from fundamentals is likely to create the greatest risk.

\bibliography{myrefs}

\end{document}

### 2014-09-25 Futures(graph).R

@@ -1,107 +1,0 @@
-# This is the script to create graphs of the futues event window 
-rm(list = ls())
-ls()
-require(eventstudies)
-require(PerformanceAnalytics)
-##################################################################################
-# Set conditions
-par(mfrow = c(2,2))
-# currency is CAD, GBP, JPY, EUR or CHF
-currency = "CAD"
-# window is number of weeks
-wind = as.numeric(6)
-# extreme is high or low
-extreme = "high"
-hi = as.numeric(0.90)
-lo = as.numeric(0.10)
-# cumulative is either whole or after
-cum = "whole lot"
-# Sentiment index is S1 or S2 or S3
-S = "S1"
-##############################################################################
-da <- read.csv("C:/Users/Toshiba/Desktop/Futures.csv", header = TRUE)
-da$Date <- as.Date(da$Date, format = "%d/%m/%Y")
-da$S1 <- (da$NCL-da$NCS)/(da$NCL+da$NCS+da$NCSP)
-da$S2 <- (da$NCL-da$NCS)/da$OI
-da$S3 <- (da$NCL+da$NCS+da$NCSP)/da$OI
-################################################################################
-da.CAD <- subset(da, subset = da$Exchange == "CAD")
-da.CHF <- subset(da, subset = da$Exchange == "CHF")
-da.EUR <- subset(da, subset = da$Exchange == "EUR")
-da.GBP <- subset(da, subset = da$Exchange == "GBP")
-da.JPY <- subset(da, subset = da$Exchange == "JPY")
-###############################################################################
-if(currency == "CAD"){
-FX = da.CAD
-} else if(currency == "CHF"){
-FX = da.CHF
-} else if(currency == "EUR"){
-FX = da.EUR
-} else if(currency == "GBP"){
-FX = da.GBP
-} else if(currency == "JPY"){
-FX = da.JPY
-} else("Error with exchange rate")
-#############################################################################
- if(S == "S1") {
-FX$S <- FX$S1
-} else if(S == "S2"){
-FX$S <- FX$S2
-} else if(S == "S3"){
-FX$S <- FX$S3
-} else("Error with sentiment index")
-##############################################################################
-if(extreme == "high"){
-FX$ex <- as.numeric(FX$S >= quantile(FX$S,  
-         hi, na.rm = TRUE)) 
-} else if(extreme == "low") {
-FX$ex <- as.numeric(FX$S <= quantile(FX$S, 
-         lo, na.rm = TRUE))
[…]

### 2014-09-25 R/Futures.R

@@ -1,4 +1,4 @@
-# The following code draws and prepares the data from the file 
+ # The following code draws and prepares the data from the file 
 # (Futures.csv) and calculates the descriptive statistics of exchange 
 # rates and measures of sentiment.    
 # Read spec position data----------------------------------

### 2014-09-25 R/ReturnExtreme.R

@@ -1,6 +1,6 @@
 # This will try to look at the level speculation before a major price move
 # The following code draws and prepares the data from the file 
-# (Futures.csv) and calculates the descriptive statistics of exchange 
+# (Futures.csv) need to install mystats function
 # rates and measures of sentiment.    
 # Read spec position data----------------------------------
 require(eventstudies)
@@ -20,20 +20,27 @@ da.JPY <- subset(da[,2:6], subset = da$Exchange == "JPY")
 # subset extreme data--------------------------
 head(da.CAD)
 str(da.CAD)
-da.CAD$d <- as.numeric(da.CAD[,2] <= quantile(da.CAD[,2], 0.01, na.rm = TRUE))                   
-da.CAD$u <- as.numeric(da.CAD[,2] >= quantile(da.CAD[,2], 0.99, na.rm = TRUE))
-da.CAD.d <- subset(da.CAD, da.CAD$d == 1)
-da.CAD.u <- subset(da.CAD, da.CAD$u == 1)
-da.CAD.all <- apply(da.CAD, 2, FUN = mystats, na.omit = TRUE)
-da.CAD.up <- apply(da.CAD.u, 2, FUN = mystats, na.omit = TRUE)
-da.CAD.down <- apply(da.CAD.d, 2, FUN = mystats, na.omit = TRUE)
-da.CAD.all
-da.CAD.up
-da.CAD.down
-X <- data.frame(3)
-X[2,1] <- da.CAD.all[2,3]
-X[2,2] <- da.CAD.up[2,3]
-X[2,3] <- da.CAD.down[2,3]
-colnames(X) <- c("Total", "Up", "Down")
-rownames(X) <- "0.05"
+# chose currency (not yet lapply) "CAD", "EUR", "GBP", "CHF", "JPY"))
+j = "GBP"
+a <- paste("da.", j, sep = "")
+X <- matrix(NA, nrow = 2, ncol = 6)
+q = 1
+for(i in c(0.01, 0.05)) { 
+FX.l <- as.numeric(get(a)[,2] <= quantile(get(a)[,2], i, na.rm = TRUE))                   
+FX.g <- as.numeric(get(a)[,2] >= quantile(get(a)[,2], 1-i, na.rm = TRUE))
+FX.d <- subset(get(a), FX.l == 1)
[…]

### 2014-09-25 Knitr/Noise.tex

\documentclass[12pt, a4paper, oneside]{article}\usepackage{graphicx, color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.2, 0.2, 0.2}
\newcommand{\hlnumber}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlfunctioncall}[1]{\textcolor[rgb]{0.501960784313725,0,0.329411764705882}{\textbf{#1}}}%
\newcommand{\hlstring}[1]{\textcolor[rgb]{0.6,0.6,1}{#1}}%
\newcommand{\hlkeyword}[1]{\textcolor[rgb]{0,0,0}{\textbf{#1}}}%
\newcommand{\hlargument}[1]{\textcolor[rgb]{0.690196078431373,0.250980392156863,0.0196078431372549}{#1}}%
\newcommand{\hlcomment}[1]{\textcolor[rgb]{0.180392156862745,0.6,0.341176470588235}{#1}}%
\newcommand{\hlroxygencomment}[1]{\textcolor[rgb]{0.43921568627451,0.47843137254902,0.701960784313725}{#1}}%
\newcommand{\hlformalargs}[1]{\textcolor[rgb]{0.690196078431373,0.250980392156863,0.0196078431372549}{#1}}%
\newcommand{\hleqformalargs}[1]{\textcolor[rgb]{0.690196078431373,0.250980392156863,0.0196078431372549}{#1}}%
\newcommand{\hlassignement}[1]{\textcolor[rgb]{0,0,0}{\textbf{#1}}}%
\newcommand{\hlpackage}[1]{\textcolor[rgb]{0.588235294117647,0.709803921568627,0.145098039215686}{#1}}%
\newcommand{\hlslot}[1]{\textit{#1}}%
\newcommand{\hlsymbol}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlprompt}[1]{\textcolor[rgb]{0.2,0.2,0.2}{#1}}%

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt} % Paper size, default font size and one-sided paper
%\graphicspath{{./Figures/}} % Specifies the directory where pictures are stored
%\usepackage[dcucite]{harvard}
\usepackage{rotating}
\usepackage{amsmath}
\usepackage{setspace}
\usepackage{pdflscape}
\usepackage[flushleft]{threeparttable}
\usepackage{multirow}
\usepackage{listings}
\usepackage[comma, sort&compress]{natbib}% Use the natbib reference package - read up on this to edit the reference style; if you want text (e.g. Smith et al., 2012) for the in-text references (instead of numbers), remove 'numbers' 
\usepackage{graphicx}
%\bibliographystyle{plainnat}
\bibliographystyle{agsm}
\usepackage[colorlinks = true, citecolor = blue, linkcolor = blue]{hyperref}
%\hypersetup{urlcolor=blue, colorlinks=true} % Colors hyperlinks in blue - change to black if annoying
%\renewcommand[\harvardurl]{URL: \url}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}
\title{Noise trader: random walks or forced march?}
\author{Rob Hayward\footnote{University of Brighton Business School, Lewes Road, Brighton, BN2 4AT; Telephone 01273 642586.  rh49@brighton.ac.uk}}
\date{\today}
\maketitle
\begin{abstract}
Noise traders have been characterised as random providers of liquidity, as neutral equivalents of computor-generated random numbers.  hough noise traders may create risk that will deter rational traders from taking advantage of their knowledge of the fundamental value of securities (De-long) and thereby preventing them being driven into extinction as Friedman had asserted (Fridman), their presence  is usually seen as beneficial as they provide liquidity in a neutral fashion.  However, an analysis of noise trading that seeks to measure the weight of noise traders in the market and the strength of their opinin shows that herding is a common feature, highlighting the non-neutral influence that this speculative activity can impose on securities prices. 
\end{abstract}

\section{Introduction}
\section{Literature}
\section{Methods}
\section{Analysis}
\section{Conclusion}
\end{document}

### 2014-11-13 Knitr/Noise.Rnw
Updated the noise document with some deletions and additionas of new reading.  Needs another look and review

@@
 
+\href{http://papers.ssrn.com.ezproxy.liv.ac.uk/sol3/papers.cfm?abstract_id=869827}{Do noise traders move markets?}  There is strong evidence of herding by individual investors.
+THere is also momentum.
+There is reversal:  stocks that were heavily bought one year underperform; stocks that were heavily sold, outperform.
+The opposite in the short run:  heavily bought one week generate large returns the next week.
+
+\href{http://restud.oxfordjournals.org.ezproxy.liv.ac.uk/content/60/1/1.full.pdf+html}{Campbell and Kyle, Smart money, noise trading and stock market behaviour}.
+In response to the evidence that the volatilty of stock prices is too great to be a function of discounted expected future earnings (Shiller (1981),  LeRoy and Porter (1981), this paper assesses whether stocks are influnced by exogenous, serially correlated noise.
+In this model, the noise arises from the interaction of noise traders.
+
 \section{Methods}

### 2014-11-13 Knitr/Noise.tex
Updated the noise document with some deletions and additionas of new reading.  Needs another look and review

@@
-\documentclass[12pt, a4paper, oneside]{article}\usepackage{graphicx, color}
+\documentclass[12pt, a4paper, oneside]{article}\usepackage[]{graphicx}\usepackage[]{color}
 %% maxwidth is the original width if it is less than linewidth
@@
 
-\definecolor{fgcolor}{rgb}{0.2, 0.2, 0.2}
-\newcommand{\hlnumber}[1]{\textcolor[rgb]{0,0,0}{#1}}%
-\newcommand{\hlfunctioncall}[1]{\textcolor[rgb]{0.501960784313725,0,0.329411764705882}{\textbf{#1}}}%
-\newcommand{\hlstring}[1]{\textcolor[rgb]{0.6,0.6,1}{#1}}%
-\newcommand{\hlkeyword}[1]{\textcolor[rgb]{0,0,0}{\textbf{#1}}}%
-\newcommand{\hlargument}[1]{\textcolor[rgb]{0.690196078431373,0.250980392156863,0.0196078431372549}{#1}}%
-\newcommand{\hlcomment}[1]{\textcolor[rgb]{0.180392156862745,0.6,0.341176470588235}{#1}}%
-\newcommand{\hlroxygencomment}[1]{\textcolor[rgb]{0.43921568627451,0.47843137254902,0.701960784313725}{#1}}%
-\newcommand{\hlformalargs}[1]{\textcolor[rgb]{0.690196078431373,0.250980392156863,0.0196078431372549}{#1}}%
-\newcommand{\hleqformalargs}[1]{\textcolor[rgb]{0.690196078431373,0.250980392156863,0.0196078431372549}{#1}}%
-\newcommand{\hlassignement}[1]{\textcolor[rgb]{0,0,0}{\textbf{#1}}}%
-\newcommand{\hlpackage}[1]{\textcolor[rgb]{0.588235294117647,0.709803921568627,0.145098039215686}{#1}}%
-\newcommand{\hlslot}[1]{\textit{#1}}%
-\newcommand{\hlsymbol}[1]{\textcolor[rgb]{0,0,0}{#1}}%
-\newcommand{\hlprompt}[1]{\textcolor[rgb]{0.2,0.2,0.2}{#1}}%
+\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
+\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
+\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
+\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
+\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
+\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
+\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
+\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
+\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
+\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
 
@@
 \section{Literature}
+\subsection{Notes}
+From "How noise trading..." in the other folder. 
+O'Hara and Saar suggest that there is a difference between noise traders and liquidity traders.
+Noise traders exert positive effects on liquidity but weaken price reversals.
+A tax on trading will reduce noise traders but will also reduce the influence of fundamental traders.
+
+Literature.
+Black (1986) highlights the positive effect of noise traders.
+Shliefer, Summers, DeLong etc identify the limits to arbitrage.
+The other writings are Battalio, Hatch and Jennings (1997), Harris and Schultz (1998) and Scheinkman and Xiong (2003).
+Lowenstein and Willard (2006).
+There is considerable debate about the precise role of noise trading in financial markets.
+There are two strands to the literature on noise trading that emerged in the 1980s:  microstricture literature will use the term "noise traders" and "liquidity traders" interchangeably to identify traders without information.
+The motivation of these traders is not identified.
+Examples of this are Glosten and Milgrom (1985) and Kyle (1985).
+It is usually assumed to be hedging or liquidity.
+There is also a \emph{limits-to-arbitrage} literature that uses the term to identify traders that are not movivated by standard explanations of fundamental or liquidity provision.
+This paper seeks to distinguish between those that trade for liquidty or heging and those that trade because of psychological biases.
+Fisher Black ``People who trade on noise
+are willing to trade even though from an objective point of view they would be better off 
+not trading. Perhaps they think the noise they are trading on is information. Or perhaps 
+they just like to trade``.
+
+In theory, traders may trade on mistaken fundamental information.
+Shliefer and Summmer (1990) use momentum.
+However, Kaniel, Saar and Titman (2006) find evidence of contrarian strategies.
+This is associated with short-term speculation.
+Does the activity drive the price away from fundamental value?
+Stiglitz (1990) is one of the arguments in favour of a Tobin tax.
+
+There is a quesiton about whether experimental markets (like this paper) are superior to actual markets where the identification of trader types is difficult.
+Theory - Delong, summers, shleifer etc,  Scheinkman and Xiong (2003).
+Lowenstein and Willard (2006).
+Empirical work:  (see, for example, Garvey and Murphy [2001], Linnainmaa [2003], Barber et.
+al.
+[2004], and Kaniel, Saar, and Titman [2006].
+
+For related papers on securities transaction taxes see Stiglitz [1989], Summers and Summers [1989], 
+Amihud and Mendelson [1993], Schwert and Seguin [1993], Umlauf [1993], Subrahmanyam [1998], Dow 
+and Rahi [2000], Habermeir and Kirilenko [2003], and Hinman [2003].
+
+Interesting difference between supplying and demanding liqudiity.
+Limit orders supply liquidity, market orders demand liqudity.
+The activity of liquidity traders will tend to change through the course of the session, moving from supplying to demanding liquidity.
+Noise traders tend to act in the opposite manner.
+They tend to supply more liquidity as the session progresses.
+Black (1986) noise traders put noise into prices.
+The effect of noise traders on liquidity: price impact of trades is lower when noise traders are present.
+However, this is mostly in the temporary impact on prices rather than the permanent impact.
+Therefore, noise traders tend to reduce the amount of volatility in prices because market orders do not temporarily have an effect and then reverse.
+The presence of noise traders will increase pricing errors.
+However, this depends on fundametal traders have valuable information.
+
+There are a number of different ways that noise traders could behave:  activing as SOES bandits (using the NASDAQ small order execution system to make profits) or by using information about the order book; by providing liqudity like a market-maker as a service to liqudiity-traders as in Grossman Miller (1988);  as irrational noise traders, using trading rules or popular models (Shiller 1984, 1990) positive feedback or contrarian.
+
+Footnote -  There is an empirical evidence that individual investors in various countries trade in a contrarian fashion (e.g., Choe, Kho, and Stulz [1999], Grinblatt and Keloharju [2000, 2001], Jackson [2003], Richards [2005], Kaniel, Saar, Titman [2006]).
+For experimental evidence consistent with contrarian behavior see
+Bloomfield, Tayler, and Zhou (2997)
+
+Can noise traders be viewed as just random noise.
+Dow and Gorton (2006) suggest that this is the case and that this is an outcrop of the rational expectations revolution.
+Informed traders profit by exploiting the diviations from equilibrium.
+This is the beneifts of Grossman-Stiglitz.
+What is the source of this noise?
+What are the insitiutions?
+The importance of noise traders here is that they represent a zero-mean, normally distributed random vaiable.
+This is also the description that is provided in Kyle (1985).
+Liquidity traders trade due to imbalance in the timing of consumption or from portfolio considerations.
+Shleifer, Summers and DeLong have trading activity that is affected by sentiment, models or "fads".
+This is important.
+How the average may not be zero.
+There is a range of behavioural biases that may account for this (see Barberis and Thaler [2003]).The literaure on hetrogenous traders also focuses on noise traders as trend chasers.
+These models (see
+for example Brock and Hommes [1998]; Lux [1995]; Lux and Marchesi [2000]) investigate bubbles and 
+crashes, and feature fundamental traders versus noise traders. In these models, it is the noise traders who 
+induce the aberrant market behavior.
+
+\href{http://papers.ssrn.com.ezproxy.liv.ac.uk/sol3/papers.cfm?abstract_id=869827}{Do noise traders move markets?}  There is strong evidence of herding by individual investors.
+THere is also momentum.
+There is reversal:  stocks that were heavily bought one year underperform; stocks that were heavily sold, outperform.
+The opposite in the short run:  heavily bought one week generate large returns the next week.
+
+\href{http://restud.oxfordjournals.org.ezproxy.liv.ac.uk/content/60/1/1.full.pdf+html}{Campbell and Kyle, Smart money, noise trading and stock market behaviour}.
+In response to the evidence that the volatilty of stock prices is too great to be a function of discounted expected future earnings (Shiller (1981),  LeRoy and Porter (1981), this paper assesses whether stocks are influnced by exogenous, serially correlated noise.
+In this model, the noise arises from the interaction of noise traders.
+
 \section{Methods}

### 2014-11-13 Knitr/Paper.tex
Updated the noise document with some deletions and additionas of new reading.  Needs another look and review

@@
 A unique dataset of risk-reversal skew on option prices is used to measure the weight of speculative sentiment.
+This is suplemented with regulatory data on the futures position of speculative traders in the fugtures market.
 \end{abstract}
@@
 
-There is a fundamental schism in the literature between those that view speculation as uninformed noise activity that can run in waves when affected by common decision-making bias or social influences, and those that view speculation as activity informed by fundamental knowledge that will stabilise markets.
-In the former case, there is a risk that speculative activity will drive prices away from fundamental value as price changes and sentiment feed off each other.
-At its most extreme, speculation in this interpretation is likely to turn into a bubble that may burst in a cascade where the previous run-up in price and sentiment is swiftly reversed.
-However, the alternative view of speculation as an activity that is driven by fundamental knowledge and information would suggest that speculation acts to limit overshooting and divergence from fundamental value, thereby reducing volatility.
+One of the many things that divided John Maynard Keynes and Milton Friedman was their opinion of speculation.
+Keynes main discussion of speculation was in Chapter 13 of \emph{The General Theory} where he identified speuclation as a myopic, sentiment-driven activity that was driven by the desire to ``beat the gun'' or  ``outwit the crowd'' \citep[p.
+101]{Keynes1936}.
+Friedman, however, say speculation as a stabilising force that would ensure that markets swiftly found equilibrium.
+In arguing in favour of flexible exchange rates, Friedman asserts that speculators have a stabilsing influence as their success and domination will depend on correctly buying when prices are below fundamental value and selling when above the correct rate.
+
+In Keynes' view, speculative activity will tend to drive prices away from fundamental value as price changes and sentiment feed off each other.
+At its most extreme, speculation is likely to cause a series of bubbles.
+However, the alternative view regards speculation as an activity that is driven by fundamental knowledge and information and acts to limit overshooting and divergence from fundamental value, thereby reducing volatility.
 This second view says that it is the absence of speculators and the liquidity that they provide that encourages excess price movement and volatility.
-The aim of this chapter is to use the foreign exchange market to assess the relationship between intensity or weight of speculation and the movement of prices.
+
+The aim of this paper is to use the foreign exchange market to assess the relationship between intensity or weight of speculation and the movement of prices.
 If speculators are uninformed, their most extreme sentiment and their greatest weight in market activity should tend to cause overshooting and presage bubble-bursting, reversals and a crash; if speculators are informed, activity will drive prices towards fundamental value and equilibrium where volatility and crash risk should be reduced.
 
-\subsection{Fundamental value}
-Fundamental or intrinsic value is a difficult concept to apply to foreign exchange.
-There is a weight of empirical evidence that standard economic models are not very effective in explaining the behaviour of exchange rates.
-Meese and Rogoff found that in out-of-sample forecasts, a random walk performed as well as the four major economic models of foreign exchange behaviour that were employed \citep{Meese1983Empirical}.
-These findings have been repeated in a number of similar, subsequent studies  \citep{Frankel1984tests},  \citep{Frankel1990Exchange}, \citep{Lyons1995Microstructure} and \citep{Sarno2006}.
-In the previous chapter, the fundamental value of the exchange rate was assessed to be PPP.
-However, even there the real exchange rate was not stable (see Figure \ref{fig:timeseries} and view the variable RTWI to see the evolution of the US real trade weighted index over the period 1985 to 2010).
-Therefore, while the law of one price and purchasing power parity generally determines exchange rates in the long-run for currencies of countries at a similar level of economic development, in the short to medium-term there is still no robust economic model that can consistently explain exchange rate movements, let alone provide a forecast.
-
+From "How noise trading..." in the other folder. 
+O'Hara and Saar suggest that there is a difference between noise traders and liquidity traders.
+Noise traders exert positive effects on liquidity but weaken price reversals.
+A tax on trading will reduce noise traders but will also reduce the influence of fundamental traders.
+
+Literature.
+Black (1986) highlights the positive effect of noise traders.
+Shliefer, Summers, DeLong etc identify the limits to arbitrage.
+The other writings are Battalio, Hatch and Jennings (1997), Harris and Schultz (1998) and Scheinkman and Xiong (2003).
+Lowenstein and Willard (2006).
+There is considerable debate about the precise role of noise trading in financial markets.
+There are two strands to the literature on noise trading that emerged in the 1980s:  microstricture literature will use the term "noise traders" and "liquidity traders" interchangeably to identify traders without information.
+The motivation of these traders is not identified.
+Examples of this are Glosten and Milgrom (1985) and Kyle (1985).
+It is usually assumed to be hedging or liquidity.
+There is also a \emph{limits-to-arbitrage} literature that uses the term to identify traders that are not movivated by standard explanations of fundamental or liquidity provision.
+This paper seeks to distinguish between those that trade for liquidty or heging and those that trade because of psychological biases.
+Fisher Black ``People who trade on noise
+are willing to trade even though from an objective point of view they would be better off 
+not trading. Perhaps they think the noise they are trading on is information. Or perhaps 
+they just like to trade``.
+
+In theory, traders may trade on mistaken fundamental information.
+Shliefer and Summmer (1990) use momentum.
+However, Kaniel, Saar and Titman (2006) find evidence of contrarian strategies.
+This is associated with short-term speculation.
+Does the activity drive the price away from fundamental value?
+Stiglitz (1990) is one of the arguments in favour of a Tobin tax.
+
+There is a quesiton about whether experimental markets (like this paper) are superior to actual markets where the identification of trader types is difficult.
+Theory - Delong, summers, shleifer etc,  Scheinkman and Xiong (2003).
+Lowenstein and Willard (2006).
+Empirical work:  (see, for example, Garvey and Murphy [2001], Linnainmaa [2003], Barber et.
+al.
+[2004], and Kaniel, Saar, and Titman [2006].
+
+For related papers on securities transaction taxes see Stiglitz [1989], Summers and Summers [1989], 
+Amihud and Mendelson [1993], Schwert and Seguin [1993], Umlauf [1993], Subrahmanyam [1998], Dow 
+and Rahi [2000], Habermeir and Kirilenko [2003], and Hinman [2003].
+
+Interesting difference between supplying and demanding liqudiity.
+Limit orders supply liquidity, market orders demand liqudity.
+The activity of liquidity traders will tend to change through the course of the session, moving from supplying to demanding liquidity.
+Noise traders tend to act in the opposite manner.
+They tend to supply more liquidity as the session progresses.
+Black (1986) noise traders put noise into prices.
+The effect of noise traders on liquidity: price impact of trades is lower when noise traders are present.
+However, this is mostly in the temporary impact on prices rather than the permanent impact.
+Therefore, noise traders tend to reduce the amount of volatility in prices because market orders do not temporarily have an effect and then reverse.
+The presence of noise traders will increase pricing errors.
+However, this depends on fundametal traders have valuable information.
+
+There are a number of different ways that noise traders could behave:  activing as SOES bandits (using the NASDAQ small order execution system to make profits) or by using information about the order book; by providing liqudity like a market-maker as a service to liqudiity-traders as in Grossman Miller (1988);  as irrational noise traders, using trading rules or popular models (Shiller 1984, 1990) positive feedback or contrarian.
+
+Footnote -  There is an empirical evidence that individual investors in various countries trade in a contrarian fashion (e.g., Choe, Kho, and Stulz [1999], Grinblatt and Keloharju [2000, 2001], Jackson [2003], Richards [2005], Kaniel, Saar, Titman [2006]).
+For experimental evidence consistent with contrarian behavior see
+Bloomfield, Tayler, and Zhou (2997)
+
+Can noise traders be viewed as just random noise.
+Dow and Gorton (2006) suggest that this is the case and that this is an outcrop of the rational expectations revolution.
+Informed traders profit by exploiting the diviations from equilibrium.
+This is the beneifts of Grossman-Stiglitz.
+What is the source of this noise?
+What are the insitiutions?
+The importance of noise traders here is that they represent a zero-mean, normally distributed random vaiable.
+This is also the description that is provided in Kyle (1985).
+Liquidity traders trade due to imbalance in the timing of consumption or from portfolio considerations.
+Shleifer, Summers and DeLong have trading activity that is affected by sentiment, models or "fads".
+This is important.
+How the average may not be zero.
+There is a range of behavioural biases that may account for this (see Barberis and Thaler [2003]).The literaure on hetrogenous traders also focuses on noise traders as trend chasers.
+These models (see
+for example Brock and Hommes [1998]; Lux [1995]; Lux and Marchesi [2000]) investigate bubbles and 
+crashes, and feature fundamental traders versus noise traders. In these models, it is the noise traders who 
+induce the aberrant market behavior.
+
+\href{http://papers.ssrn.com.ezproxy.liv.ac.uk/sol3/papers.cfm?abstract_id=869827}{Do noise traders move markets?}  There is strong evidence of herding by individual investors.
+THere is also momentum.
+There is reversal:  stocks that were heavily bought one year underperform; stocks that were heavily sold, outperform.
+The opposite in the short run:  heavily bought one week generate large returns the next week.
+
+\href{http://restud.oxfordjournals.org.ezproxy.liv.ac.uk/content/60/1/1.full.pdf+html}{Campbell and Kyle, Smart money, noise trading and stock market behaviour}.
+In response to the evidence that the volatilty of stock prices is too great to be a function of discounted expected future earnings (Shiller (1981),  LeRoy and Porter (1981), this paper assesses whether stocks are influnced by exogenous, serially correlated noise.
+In this model, the noise arises from the interaction of noise traders.
+% Does this now need something on the reason why the foreign exchange market is used?  Could be the Keynes speuclative activity, Fridman paper, liquidity etc.
+
+\subsection{Speculation and noise}
 Noise-trading is the label attached to that trading or investment where decisions seem to have been made with no reference to fundamental information; activity is like a random noise.
-This chapter provides an investigation into the effect of speculation in the foreign exchange market.
-It seeks to assess the relationship between sentiment, speculation, information and exchange rate prices.
 The DeLong-Summers model of noise trading is calibrated against daily and weekly returns for major foreign exchange rates \citep{Delong1990noise}.
@@
 See  \citep{DeBondtOver}, \citep{Chopra1992} and \citep{SteinOptions} amongst others (discussed more fully in Section \ref{secref:empiric} below).
+
 The results of the experiment carried out here suggest that it is not sufficient just to measure the intensity of speculative sentiment or the weight of speculators to determine the divergence from intrinsic value and that the relationship between speculation and price is rather complex.
@@
 
+The difference in the results achieved here compared to that recorded in some other markets may be due to institutional features.
+For example, there is a natural tendency for final investors to hold long positions in equities while the foreign exchange market is much more evenly balanced with selling equivalent to buying when the ratio is reversed.
+The foreign exchange market is also more liquid than the equity market.
+The latest survey of activity from the Bank for International Settlements suggests that average daily activity is in the region of \$4.0trn \citet{BISFX}.
+
 \section{Literature review}  
-In many ways this is a study about how efficient the market  is in processing information, specifically whether the presence of speculators improves or reduces the efficiency with which information is processed.
-As such, there must first be some discussion of information and ideas about the efficiency with which it is incorporated into asset prices.
-
-Bachelier is responsible for the first structured  study of informational efficiency in his 1900 doctoral analysis of the behaviour of French government bond, future and option prices \citep{Bachelier1900}.
-His is apparently the first analysis that suggested that the evolution of securities' prices was a \emph{Martingale}, meaning that past events do not help to predict the future value, that the expected future value is equal to the current value and that the process is a \emph{fair game} for which the probability of winning or losing is equal.
-A \emph{random walk} is a slightly stronger construct as each step is an independent and identically distributed random variable.
-However, a martingale may contain past information about the higher moments.
-So there may be, for example, volatility clustering.
-Brownian motion is named after the Scottish botanist Robert Brown and refers to the random movement of particles suspended in a liquid or gas.
-See \citep{LeRoyMartingales} for an overview of market efficiency, random walks and Martingale processes.
-Bachelier studied a set of bond prices and found that the potential returns of these $19^{th}$ century French government bonds could be represented by a probability distribution where the mean was zero once the gravitation towards parity at maturity and seasonal cycle of `dirty prices' with accrued interest had been accounted for.
-
-\begin{quotation}
-`` The determination of these fluctuations depends on an infinite number of factors; it is, therefore impossible to aspire to mathematical prediction of it.
-Contradictory opinions concerning these changes diverge so much that at the same instant buyers believe in a price increase and sellers in a price decrease."
-\end{quotation}
-
-\citep[p. 17]{Bachelier1900}
-
-Bachelier's ideas received little immediate attention and remained dormant until renewed interest in the process describing securities prices developed in the 1950s.
-However, others were tentatively pursuing similar investigations.
-In the 1930s Working assessed the empirical evidence for random behaviour of stock and commodity prices \citep{Working1934} and U.K.
-statistician Kendall studied the performance of weekly price data from the Dow Jones \citep{Kendall1953}.
-Among those taking an interest in the subject were Roberts, who scorned the \emph{patterns} that were identified by professional investors while making the case for a focus on the change in price of securities rather than the price itself.
-What we now call \emph{returns} were, argued Roberts, much more random than price levels.
-He calibrated a model of random shocks to the price change, using the RAND Corporation's random number text \citep{RAND} and created graphs that were remarkably close to those produced by real stocks.
-Nonetheless, Roberts warned his research would not prevent people from using past data to try to make investment decisions as ``chance behaviour itself provides 'patterns' that invite spurious interpretations'' \citep[p.
-2]{Roberts1959}.
-
-Towards the end of the decade, Osborne made the direct comparison between the price movement of stocks and the activity of small particles through the analysis of prices of a selection of stocks from the American and New York stock exchanges.
-The data were collected from \emph{The Wall Street Journal} by a research company called F.W.
-Stephens for the period from 1924 to 1956.
-The log of the price ratio for a range of different time periods and their distribution was compared to what would have been expected from something evolving with Brownian motion.
-By plotting the ratio over time, Osborne showed that changes in the log of the price ratio was close to being normally distributed.
-Osborne argued that the securities prices jumped around to find the point where buyers and sellers were balanced and overall opinion about the direction of prices was close to zero.
-Therefore, the evolution of prices was a like a sequence of independent random variables that were the consequence of transactions taking place.
-The distribution function for this sequence of random variables was, according to Osborne's study, normal with a zero mean and a variance that increased steadily with time.
-Osborne concluded that the random nature of price changes would ensure that buyers and sellers could only make a standard return on average  \citep{Osborne1959}.
-
-Samuelson provided a theoretical foundation for the idea that securities prices would follow a random walk.
-Starting with a series of simple axioms, he deduced that competitive markets would drive securities prices towards the expected value, where the expected value is defined as the projected mean of the process underlying the evolution of the price.
-This introduced the idea of a \emph{Martingales} or the notion that the expected value of a security or financial asset would be equal to its current value.
-If that is not the case, rational investors in a competitive market will take advantage of the deviation by buying when it is below the expected value and selling when it is above and, as a result of this activity, would drive the price towards the expected level and, in the process, remove the opportunity to make profits.
-Using the backward iteration of expectations, the theory states that a stock price should equal the expected present value of future dividends.
-This work provided the underpinning and stimulus for another wave of empirical examinations but rests, of course, on the assumptions made about the competitive nature of markets and the ability of investors to form expectations about the future \citep{SamuelsonStockPrices}.
-
-Amongst these empirical investigations, Cootner tested Bachelier's ideas on a range of financial markets, most notably the forward market for grain \citep{CootnerSpeculators} and  Fama found that the returns of stocks in the Dow Jones Industrial Average were very close to the benchmark of a random walk \citep{FamaStock}.
-At this point the term \emph{Efficient Market Hypothesis} (EMH) became more widely used.
-Jensen described market efficiency as
-
-\begin{quotation}
-``A market is efficient with respect to information set $\Omega_t$ if it is impossible to make economic profits by trading on the basis of information set $\Omega_t$".  
-\end{quotation}
-\citep[p. 95]{JensenEfficiency}.
-
-More formally,
-
-\begin{equation}
-E[R_{t+1}, Q_{t+1} | \Omega_t] = 0
-\end{equation}
-
-where $E$ is the expectations operator, $Q_{t+1}$ is the risk premium at time $t+1$, $R_{t+1}$ is the return at time $t +1$ and $\Omega_t$ is the information set at time $t$ \citep[p.
-95]{JensenEfficiency}.
-Jensen also said "I believe that there is no other proposition in economics which has more solid empirical evidence supporting it than the Efficient Market Hypothesis" \citep[p.
-93]{JensenEfficiency}.
-Jensen was, however, identifying a number of inconsistencies and anomalies that had been uncovered and was anticipating a revolution that would overthrow the theory.
-He was anticipating the paradigm shift that was to follow.
-
-%this paragraph goes somewhere else. 
-%Among these were positive returns following an earnings release \citep{Ball1978} and \citep{Watts1978};  a successful trading rule for buying closed-end funds trading at a discount to asset value \citep{Thompson1978}; discrepancies between stock and option prices on the NYSE \citep{Galai1978}; and, information in option prices \citep{ChirasManaster}.
-
-\subsection{Failures of market efficiency}
-Questions about the EMH had emerged gradually with doubts about some of the assumptions that Samuelson had used in the construction of his theory and with a rising number of indications that financial asset prices did not evolve in a totally random fashion.
-There were three main strands to the question that arose:   empirical evidence that the price of securities did not follow a random walk made up of a series of independent and identically distributed random variables;  questions about the effects of institutional structures of financial markets on price;  the criticism of VNM expected utility theory as a valid model of investment decision-making.
-
-\subsubsection{Empirical evidence}
-The empirical evidence that has been gathered about the nature of the returns to securities reveals two primary issues.
-The first is that serial correlation or trends are evident and the second is the leptokurtic and skewed distribution of these returns.
-The former is the most damaging to the idea that financial markets can be informationally efficient as it suggests, contrary to the implications of the EMH, that future price changes could be partially foreseen by the movement of past price changes.\footnote{This of course depends on the assumption of \emph{risk neutrality} and the notion that investor decisions will be determined by the first two moments of the distribution of expected returns.
-If the return distribution is skewed or fat tailed or there is serial correlation in the conditional variance of the returns, there may be unexplored opportunities to reduce risk or to increase risk-adjusted returns.
-The third and fourth moments of the distribution of returns can be extremely important in assessing the level of risk.}  The evidence seems to support the notion that prices can evolve in trends, and that there are times when positive price developments are more likely to be followed by further positive developments and that there are other times when positive price developments are more likely to be followed by a negative outcome.
-
+There has been an accumulation of evidence to suggest that financial asset prices do not evolve as a random walk.
+Serial correlation or trends are evident suggesting that, contrary to the efficient market hypthesis, future price changes could be partially foreseen by the movement of past price changes.
 There are a number of studies that find evidence that the reaction of stocks to information is less than instantaneous.
@@
 
-\subsubsection{Normal distribution}
-The indications that the distribution of financial security returns are likely to be leptokurtic and skewed would not necessarily undermine the EMH as long as the distributions were independent.
-However, this evidence that the distribution of returns is not normal, reinforces the view that standard analysis of investment based on the first two moments of the return distribution may not be adequate to understand risk.
-This is something that will be more fully investigated in Chapter \ref{Chapter4} with a study of the carry trade.
-In addition, if there is a pattern to the evolution of the higher moments over time, which appears to be the case, this can also suggest that there may be informational inefficiencies that can be exploited to make abnormal returns.
-
-For example, volatility clustering means that a period when the returns to financial assets exhibit a high variance is much more likely to be followed by another period with high volatility than one where the volatility returns to some average level.
-The creation of autoregressive conditional hetroscedasiticity (ARCH) and general autoregressive  conditional hetroscedasticity (GARCH) models have been designed to understand this process and provide greater precision about the effect that a shock to volatility will have on the system.
-See \citep{Engle1982} for the original model, \citep{Bollerslev1986} for the generalised version, \citep{Engle1987} for an application to time-varying risk premia and \citep[pp.
-657 - 665]{Hamilton} for a fuller overview of the subject.
-
 Lo and MacKinlay directly tested the random walk hypothesis for weekly stock returns for the period between 1962 and 1985.
@@
 
-In 1959 Osborne found that returns were nearly normally distributed.
-However, the quantile plots that he produced showed that there were slightly more extreme readings than would be expected \citep{Osborne1959}.
-This was a finding that was repeated by a number of other studies.
-With computer power lacking and measurement imprecise, these finding of a small deviation from normality were initially ignored.
 Paul Cootner had already found that extreme returns were more likely to be recorded at the daily frequency than at the monthly level.
@@
 This is the distrubiton that Mandelbrot and Fama proposed.
-\subsubsection{Decision-making} 
-As discussed in Section \ref{sec:prob} there are alternative views of the decision-making process.
-Knight, Shackle, Keynes and others had suggested that that not all decisions could be subjected to a quantifiable, probabilisitic approach.
-In addition, Simon updated the idea of \emph{economic man} with \emph{global rationality} to something that was consistent with the actual quality of information and computational capacities of economic agents.
-Simon questioned the model of the \emph{economic man} with stable preferences, clear and voluminous information and computational skill.
-Simon introduced the term \emph{bounded rationality} to express the limits to VNM rational-decision-making: there is only a certain amount of information that can be processed and probably only a certain amount of information that is available  \citep[p.
-99]{Simon1955Behavioural}.
-Simon wanted to use psychological evidence to support his ideas, but found it lacking at that time.
-However, as more experimental work on the psychology of decision-making was carried out, empirical support was provided for many of the themes that Knight, Shackle, Keynes and Simon had already proposed.
-
-%Of course, even those involved in the original construction of the EMH were aware that this was a model rather than a description of the decision-making process.
-For example, back in 1959 Osborne stated that investment decisions would be made "as if" calculating the expected change in the log ratio of prices and
-
-%\begin{quotation}
-%"From what has been said about the anatomy of logical decisions (they need not be consciously logical, but this is the supposition as to how they are reached), we must suppose that for the buyer, his estimate of the expectation value of the change in value $(\Delta log_e P)$ is positive while the sellers estimate of the same quantity is negative"
-%\end{quotation}
-%\citep[p. 153]{Osborne1959}
-
-%What is important is whether this model provides outcomes that are consistent with what is evident.
-Again, it may be the case that in the long-term the standard model of rational decision-making provides a reasonable description of what happens in aggregate but there is a risk that at the individual level and on the path to the equilibrium, the process of groping around with limited information and limited ability to assess possible outcomes can create outcomes that diverge from that of the standard model.
-
-%This needs to find a place in the introduction.   
-%Herbert Simon introduced the term 'bounded rationality' to express the limits to rational-decision-making: there is only a certain amount of information that can be processed and probably only a certain amount of information that is available.
-Simon was focused more on the decision-making process itself than the aggregate outcome of decisions that had been made.
-He questioned the model of the "economic man" with stable preferences, clear and voluminous information and computational skill.
-One of the key ideas that he developed was that of 'satisficing' rather than maximising utility.
-It means that decisions are taken to achieve at least a particular outcome.
-In the field of investment, it may mean that decisions are taken to ensure that a minimum level of return is achieved rather than the maximum.
-This can reduce the amount of options that have to be assessed \citep{Simon1955Behavioural}.
-
-%As we have seen, Knight, Shackle and Keynes %ref for introduction
-%were particularly sceptical about rational decision-making.
-Keynes argued that given the uncertainty that is attached to long-term investment projects, it is likely that investors will tend to use a short-cut to decision making that would be based on the assumption that the valuation or price will remain unchanged unless there is some strong reason to believe that it should be elsewhere.
-Keynes' own statement of the EMH is
-%\begin{quotation}
-%We are assuming, in effect, that the existing market valuation, however, arrived at, is uniquely correct in relation to our existing knowledge of the facts which will influence the yield of the investment, and that it will only change in proportion to changes in this knowledge.
-%\end{quotation}
-%\citep[p. 152]{Keynes1936}.
-
-%However, Keynes questioned this notion, arguing that the existing market valuation was more likely to be influenced by a variety of ephemeral, short-term fluctuations in sentiment and popular opinion with market valuations being based more on ignorance and subject to sudden fluctuations in the mass interpretation of the available information.
-As a result, he felt that "the market would be subject to waves of optimistic and pessimistic sentiment" \citep[p.
-33]{Keynes1936}.
-One important aspect of this is the influence of public opinion and the importance of social pressures.
-Keynes argues that the investment process is like a competition where the winner is the one that can guess what everyone else will choose.\footnote{The third degree "where we devote our intelligence to anticipating what average opinion expects average opinion to be" \citep[p.
-156]{Keynes1936} has been achieved with 'sniffer algorithms' which try to identify the patterns used by other machines so that they can get a head start on opponents strategies.
-See, for example, More or Less, 11 August 2011, BBC 4.}
 
 \subsubsection{Behavioural finance}
-In 2002 Daniel Kahneman shared the Nobel prize in economics for the work that he did with Amos Tversky.
-The Nobel committee cited the application of experiments and surveys to question the assumptions of economic rationality and the model of VNM expected utility and to build a more empirically-based theory of how decisions are made as a justification for the award \citep{Kahneman1979Prospect} and \citep{KahnemanThinking}.
-Kahneman and Tversky present \emph{Prospect Theory} which is an alternative to the VNM expected utility theory.
-The theory was developed from the results of a large number of experiments that Kahneman and Tverskey carried out to ask people what would be their decision given a choice of alternatives.
-Kahneman and Tverskey then compared their results to those that would be expected from a VNM expected utility model.
-
-Kahneman and Tversky developed a \emph{decision-weighting function} $\pi = \pi(p)$ which had a similar effect as the adjustments that Keynes makes to standard probability estimates using $q$ to cover risk aversion and $w$ to cover uncertainty.
-The essential difference between Prospect Theory and the VNM expected utility benchmark are that people underweigh outcomes that are probable compared to certainties, which contributes to relatively high risk-aversion in choices involving sure gains and risk-taking in choices involving sure losses \citep[p.
-263]{Kahneman1979Prospect}.
-In practical terms, for investors, this encourages the phenomenon of selling winners early to lock in gains and holding on to losers for too long, gambling in the face of adverse consequences.
-In addition, the people making decisions in the experiments conducted by Kahneman and Tversky are concerned about changes in wealth rather than the level of wealth itself.
-According to Kahneman and Tverskey, this explains why people are prepared to purchase insurance that will cover losses that are a relatively small proportion of their wealth, losses that they could presumably self-insure without the administration costs of a formal insurance company \citep[p.
-286]{Kahneman1979Prospect}.
-
-%can we cover what Shackle, Knight, Keynes and Simon said here....
-More relevant for this study are the \emph{heuristics} or short-cuts to decision-making that have been identified by Kahneman, Tversky and other psychologists.
+
+Most relevant for this study are the \emph{heuristics} or short-cuts to decision-making that have been identified by Kahneman, Tversky and other psychologists.
 A range of investigations reveal that people deal with the problem of limited information and limited cognitive processing power by using these heuristics  that will economise on time and effort.
@@
 
-There is a certain scepticism about new information and a caution about changing an established opinion.
+With a high level of uncertainty, there is a certain scepticism about new information and a caution about changing an established opinion.
 The Psychologist Edwards describes his findings relative to the benchmark of Bayesian decision-making.
@@
 If there are trends, there are inefficiencies in prices that can be exploited.
-
-Another decision-making short-cut that has been identified by a number of psychological experiments is \emph{representativeness}.
-This is the tendency to estimate the probability of an event by reference to its similarity to the parent population or similarity to the process that generated the parent population.
-Base effects tend to be downgraded or ignored.
-Kahneman presents the example of an accident that takes place in a town that has two taxi companies: 85\% of the cabs are green and 15\% are blue.
-A witness identifies the taxi involved in the accident as being blue and tests of the witness reveal that she is correct in her assertion 80\% of the time.
-What is the probability that the accident was caused by a blue cab?
-Most people put the probability in the range of 80\% to 50\% but it is actually 41\%.
-
-By Bayes' Rule (see Section \ref{secref:ex} for derivation). 
-\begin{equation}
-P(B|I) = \frac{P(I|B)P(B)}{P(I)}
-\end{equation}
-
-where $P(B|I)$ is the probability that the car was blue conditional on it having been identified as blue.  
-\begin{align*}
-P(I|B) &= 0.80 \text{   as there is 80\% accuracy}\\
-P(B) & = 0.15 \text{   15\% blue cabs}\\
-P(I|B)P(B) & = 0.80 \times 0.15\\
- &= 0.12\\
-P(I|G)P(G) &= 0.2 \times 0.85 \\
-&=0.17\\
-P(I)&= P(I|B)P(B) + P(I|G)P(G) \text{   as B and G are the only options}\\
-P(I) &= 0.12 + 0.17\\
-P(B|I) &=0.41\\
-\end{align*}
-\citep{KahnemanCabs}
-
-%Need more here.  Look for some psychology paper.  Is there one in the book, use taxi example?
-There is too much influence from the immediate information about the witness and too little about the bigger picture.
-The information about the proportion of blue and green cars is ignored.
-However, Kahneman and Tverskey reveal that if the thought experiment is changed so that the participants are told that green taxis are involved in 85\% of the accidents, the information is regarded as relevant and it is incorporated into the decision-making process and probability estimates are very close to those of the Bayesian calculation.
-
-Therefore, the representativeness heuristic also means that there is a departure from VNM expected utility outcomes and therefore some systematic biases that could be exploited by informed investors.
-This is consistent with the way that stock analysts over-react to new information \citep{DeBondtAnalysts} and more generally to the sort of myopic focus on the short-term that Keynes spoke about in Chapter Twelve of \emph{The General Theory} (see Section \ref{secref:noise} for fuller details).
-It tends to lead to over-shooting, over-reaction and the creation of bubbles.
 
@@
 
-For the noise-trader utility is
-
-\begin{equation}
-E(U) = c_0 + \lambda_t^i[r + _tp_{t+1} - p_t(1 + r)] - \gamma (\lambda_t^i)^2 \left (_t\sigma^2_{p, t+1} \right ) + \lambda_t^i (\rho_i) 
-\end{equation}
-
-where the only difference from that of the sophisticated trader is the addition of the final term, which adds utility through the noise-traders' misperception of the expected return from holding $\lambda_t^i$ of the risky asset.
-
-Maximising utility with respect to the share of the risky asset in the portfolio of the informed $(\lambda^i_t)$ and noise-traders $(\lambda^n_t)$ respectively,
-
-\begin{subequations}
-\begin{align}
-\frac{\delta (U)}{\delta \lambda} &= 0 \\
-\label{eq:NTM1}
-\lambda_t^i &= \frac{r + _tp_{t+1} - (1 + r)p_t}{2\gamma(_t\sigma^2_{p,t +1})}\\
-\lambda_t^n &= \frac{r + _tp_{t+1} - (1 + r)p_t}{2\gamma(_t\sigma^2_{t +1})} + \frac{\rho_t}{2\gamma(_t\sigma^2_{p,t+1})}
-\label{eq:NTM2}
-\end{align}
-\end{subequations}
-
-with the only difference between the two being the final term of equation \ref{eq:NTM2} which represents the misperception.
-
-The old sell their holdings to the young so that demand of the young must equal unity. Equations \ref{eq:NTM1} and \ref{eq:NTM2} above imply that
-
-\begin{subequations}
-\begin{equation}
-(1 - \mu)\lambda_t^i + \mu \lambda_t^n = 1
-\end{equation}
-
-where $\mu$ is the proportion of noise-traders to total in the market.
-
-by substitution and re-arranging
-
-\begin{equation}
-(1 - \mu) \left [\frac{r + _tp_{t+1} - (1 + r)p_t}{2\gamma(_t\sigma^2_{p,t +1})} \right ] + \mu \left [ \frac{r + _tp_{t+1} - (1 + r)p_t}{2\gamma(_t\sigma^2_{p,t +1})} + \frac{\rho_t}{2\gamma(_t\sigma^2_{p,t+1})} \right ] = 1
-\end{equation}
-\begin{equation}
-1 = \frac{r + _tp_{t+1} - (1 + r)p_t}{2\gamma(_t\sigma^2_{p,t +1})} + \frac{\mu \rho_t}{2\gamma(_t\sigma^2_{p,t +1})}
-\end{equation}
-\begin{equation}
-p_t = \frac{1}{1+r} \left [ r + _tp_{t+1} - 2\gamma(_t\sigma^2_{t+1}) + \mu\rho_t \right ]
-\end{equation}
-\end{subequations}
-
-In words, the price is a function of the discounted value of the future dividend, the expected capital gain and the misperception of speculative noise-traders, less the perceived risk.
-To obtain a more general solution, it is possible to substitute for future expected price, so
-
-\begin{equation}
-p_t = \frac{1}{1+r} \left [ r + \frac{1}{1 + r} \big \{  r + _tp_{t+2} - 2\gamma(_t\sigma^2_{p,t+2}) + \mu\rho_t \ \big \}  - 2\gamma(_t\sigma^2_{p,t+1}) + \mu\rho_t \right ]
-\end{equation}
-
-If this is repeated,  the price becomes the discounted value of the future dividends, the expected capital gain and the general misperception of noise-traders less risk.
+Maximising utility with respect to holdings of the risky asset and iterating backward to solve for the current value,  the price becomes the discounted value of the future dividends, the expected capital gain and the general misperception of noise-traders less risk.
 
@@
 
-De Long and the others use the utility function to compare the outcomes for the two types of investor for different realisations of the parameters.
+The paper makes use of the utility function to compare the outcomes for the two types of investor for different realisations of the parameters.
 They note four effects:  the \emph{Friedman Effect}, which is the tendency of informed investors to push prices back towards fundamental value; \emph{Price Pressure} as the way that speculative noise-traders buy more and drive the price higher as they become more bullish; \emph{Hold More},whereby the noise-traders hold more of the risky asset; \emph{Create Space} is the added risk that is caused by an increase in the variability of noise-trader belief.

### 2014-11-14 Knitr/Paper.tex
Completed the review of eventdoc.  Now needs to be re-read and the conclusions need to be drawn towards the front.  Some similification of the literature.

@@
 \begin{abstract}
-An event study of extreme speculation is used to test the informational content of speculation.
+An event study of extreme speculation is used to assess the Keynesian and Friedman views of speculation.
+This is partly a test of the informational content of speculation.
 Speculative extremes are ranked by weight of activity and intensity of sentiment.
-In either case, the extreme says little about the future, suggesting that speculation is more than just noise and that there is some fundamental informational content.
+In either case, the extreme says little about the future, suggesting that speculation is more than just noise and that there is some informational content that is being passed through to the price.
 A unique dataset of risk-reversal skew on option prices is used to measure the weight of speculative sentiment.
-This is suplemented with regulatory data on the futures position of speculative traders in the fugtures market.
+This is suplemented with regulatory data on the futures position of speculative traders.
 \end{abstract}
@@
 
-One of the many things that divided John Maynard Keynes and Milton Friedman was their opinion of speculation.
-Keynes main discussion of speculation was in Chapter 13 of \emph{The General Theory} where he identified speuclation as a myopic, sentiment-driven activity that was driven by the desire to ``beat the gun'' or  ``outwit the crowd'' \citep[p.
+There are a range of ideas about the way that speculation can affect financial markets.
+Towards one end of the spectrum is the view of Keynes from Chapter 13 of \emph{The General Theory} where he identified speculation as a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd'' \citep[p.
 101]{Keynes1936}.
-Friedman, however, say speculation as a stabilising force that would ensure that markets swiftly found equilibrium.
-In arguing in favour of flexible exchange rates, Friedman asserts that speculators have a stabilsing influence as their success and domination will depend on correctly buying when prices are below fundamental value and selling when above the correct rate.
-
-In Keynes' view, speculative activity will tend to drive prices away from fundamental value as price changes and sentiment feed off each other.
-At its most extreme, speculation is likely to cause a series of bubbles.
-However, the alternative view regards speculation as an activity that is driven by fundamental knowledge and information and acts to limit overshooting and divergence from fundamental value, thereby reducing volatility.
-This second view says that it is the absence of speculators and the liquidity that they provide that encourages excess price movement and volatility.
-
-The aim of this paper is to use the foreign exchange market to assess the relationship between intensity or weight of speculation and the movement of prices.
-If speculators are uninformed, their most extreme sentiment and their greatest weight in market activity should tend to cause overshooting and presage bubble-bursting, reversals and a crash; if speculators are informed, activity will drive prices towards fundamental value and equilibrium where volatility and crash risk should be reduced.
-
+However, towards the other extreme, Milton Friedman argued that speculation was a stabilising force that would ensure that markets swiftly found equilibrium.
+In arguing in favour of flexible exchange rates, Friedman asserted that successful speculators would be those that managed to correctly buy when prices  were below fundamental value and sell above that rate \citep{FriedmanPositive}.
+
+This is a contrast between an idea of speculative activity as a nuisance that, at its most extreme, will likely to cause bubbles and reversal and will discourage fundamental long-term investors from participating in the market.
+It is noise.
+that does not emerge randomly, but feeds off itsef as a feedback of sentiment  encourages  trends and serial correlation in asset prices returns.
+The alternative view regards speculation as an key lubricant of financial market as it provides liquidity and facilitates price discovery.
+
+The aim of this paper is to understand mroe about the nature of speculation in the foreign exchange market by assess the relationship between the intensity or the weight of speculation and the movement of prices.
+If speculators are uninformed and their activity can be regarded as noise, their most extreme sentiment and the time when speculators have the greatest weight in market activity should be those when price reversals are most likely to take place.
+An event study is used to assess the relationship between speculative extremes and foreign exchange returns.
+A unique set of option prices are used to measure the intensity of speculative sentiment.
+US regulatory positions measure the prominence of speculators in the market.
+
+\section{Literature}
 From "How noise trading..." in the other folder. 
-O'Hara and Saar suggest that there is a difference between noise traders and liquidity traders.
+\citet{bloomfield2009noise} suggests that there is a difference between noise traders and liquidity traders.
 Noise traders exert positive effects on liquidity but weaken price reversals.
 A tax on trading will reduce noise traders but will also reduce the influence of fundamental traders.
+Their experiment distinguishes between informed traders, noise traders and liquidity traders.
+\emph{SOES bandits} are those taking advantage of the SOES system to trade quickly and take advantage of small price movements.
+There is a lot of focus on the effect that taxation may have on the different categories of trader.
+There is an experiment that allows much more to be held constant.
+There are three views of noise traders that are tested here:  noise traders are skill experts taking advantage of information in the order book or price movements;  rationally providing liquidity by acting as market-maker; acting irrationally as noise traders.
+Noise traders in this experiement appear to act as irrational contrarian traders.
+Noise traders refuce the efficiency of the market but increase liquidity.
+
+Noise is related to rational expectations.
+A mean zero, normally-distributed random variable influences supply citet{Muth1961Rational}.
+Why do noise traders exist?
+There are liquidity and behavioural explanations.
+There are also a number of models that show noise traders as momentum traders.
+The effect of noise traders on liquidity is unclear.
+They may supply liquidity by providing limit orders and making some returns for providing liquidity; they may increase bid-ask spreads by using market orders to take advantage of (potentially non-existant) informational advantages.
+Noise traders may increase (if they have some priviledged information, they add to the price pressure) or reduce efficiecy (if they increase liquiity and reduce the ratio of informed to uninformed traders).
+
+Are day traders profitable? Are these the face of noise traders?  How does this show up in teh foreign exchange market?
+
+\citet{da2010sum} Seek to measure invetor sentiment and assess how it affects securities prices.
+They create a index of investor sentiment from web searches and use this to predict return reversals.
+The resutls show that sentiment can predict market turns.
+This is a contrast to traditional ways of measuring sentiment:  proxy for sentiment with market-based measures like volume, closed-end fund discount, IPO first-day returns, IPO volume, option implied volatities and mutual fund flows.
+Baker and Wurgaler (2007) have a comprehensive survey.
+This is more than just investors sentiment.
+There is an issue with the measurment of an input (sentiment) with an output (market price).
+Use of surveys like the University of Michigan consumer sentiment survey etc.
+There is a negative contemporaneous correlation between FEARS and stock market returns.
+However, there is a reversal in the relationship after two days.
+This is consistent with ``sentiment induced temporary mispricing'' \citet[p.
+4]{da2010sum}.
+There is a positive correlation between the FEARS index and bond prices.
+However, thsi also reverses in following days.
+There is a relatioknship betwen the FEARS index and daily market volatility (realised volatlity and the VIX index).
+There is also evidence of inflows and outflows from mutual funds.
+This may help to explain the contrast between evidence from the equity market and the bond market.
 
@@
 This paper seeks to distinguish between those that trade for liquidty or heging and those that trade because of psychological biases.
-Fisher Black ``People who trade on noise
-are willing to trade even though from an objective point of view they would be better off 
-not trading. Perhaps they think the noise they are trading on is information. Or perhaps 
-they just like to trade``.
+Fisher Black ``People who trade on noise are willing to trade even though from an objective point of view they would be better off not trading.
+Perhaps they think the noise they are trading on is information.
+Or perhaps they just like to trade``.
 
@@
 There is a range of behavioural biases that may account for this (see Barberis and Thaler [2003]).The literaure on hetrogenous traders also focuses on noise traders as trend chasers.
-These models (see
-for example Brock and Hommes [1998]; Lux [1995]; Lux and Marchesi [2000]) investigate bubbles and 
-crashes, and feature fundamental traders versus noise traders. In these models, it is the noise traders who 
+These models (see for example Brock and Hommes [1998]; Lux [1995]; Lux and Marchesi [2000]) investigate bubbles and crashes, and feature fundamental traders versus noise traders.
+In these models, it is the noise traders who
 induce the aberrant market behavior.
@@
 \subsection{Event studies}
-The event studies method ????
-of how the extreme speculative sentiment and extreme speculative positions in the DeLong model could be used to analyse future returns.
-Fuller explanation is provided here.
-Event studies are used to find the effect of a particular event on the performance of an asset price or security.
-It is often suggested that James Dolley was the first to employ the event study when looking at the effect of a stock split on the price of securities.\footnote{A stock split is when the number of shares are doubled and the value of each share is cut in half.
-This is usually done to prevent the price or cost of each share rising too far beyond the means of the small investor.}   Dolley took a sample of 174 stocks splits for all common stocks listed on the New York Stock Exchange during the period 1921 to 1931, and using what he called a case study, examined the process of stock splitting.
-Price information was taken from \emph{Poor's Manual} and \emph{Moody's Manual} as well as the \emph{New York Times}; a survey was sent to all the companies that were involved and (remarkably) 63 of the 111 questionnaires were returned completed.
-The analysis of the price change was a very small part of the study.
-Dolly found that of the 95 stock splits that he had been able to compile a time series for, 57 subsequently rose and 26 fell \citep{Dolly1933}.
-
+Event studies are used to find the influence of a particular event on the performance of an asset price or security.
+James Dolley is credited with the first use of the method when looking at the effect of a stock split on the price of securities.\footnote{A stock split is when the number of shares are doubled and the value of each share is cut in half.
+This is usually done to prevent the price or cost of each share rising too far beyond the means of the small investor.} \citep{Dolly1933}.
 However, Dolly only looked at the market reaction the day after the stock split.
@@
 They decided to measure the price action from a point eight weeks before the formal acceptance of the decision by the stock-holders.
-They measured the price change between the base date and the meeting and tried to account for other influences on the price by recording the ratio of the stock price to the Standard and Poor Industry average, thereby providing the first attempt to separate the specific returns from those of the market.
-An additional measure of the price ratio was also taken eight weeks after the formal meeting.
-For the splits that occurred between 1945 and 1946 they found that 62 of the 70 stocks exhibited a positive price performance between the base date and the meeting, during the subsequent time there was some reversal with only 55 of the stocks still showing a net gain between the base and the closure point eight weeks after the announcement meeting \citep{MyersBakay}.
+Myer and Barkay measured the price change between the base date and the meeting and tried to account for other influences on the price by recording the ratio of the stock price to the Standard and Poor Industry average, thereby providing the first attempt to separate the specific returns from those of the market \citep{MyersBakay}.
 
 Fama, Fisher and Roll use the event study method to conduct a much more systematic assessment of the effect of stock splits on security returns \citep{FamaFisherJensenRoll}.
-Fama later said that the method was secondary to his desire to show off the new dataset that he had helped to construct.
-This was one of the first uses of the Center for Research in Security Prices (CRSP) data.
-Nonetheless, the Fama study is considered to be the benchmark of Event Studies and it is the framework that is largely followed in the current work.
+the study is considered to be the benchmark of Event Studies and it is the framework that is largely followed in the current work.
 
@@
 Subsequent models have added additional factors.
-
-The mean return is calculated as 
-\begin{equation}
-\mu_{R, i} = \frac{1}{n}\sum_{j = 1}^{j = n} R_{i j}
-\end{equation}
-
-where $R_{i j}$ is the return to security i at time j and n is the number of total number of observations to that 
-\begin{equation}
-E[R_{i \tau} | X_{\tau}] = \mu_{R, i} + \varepsilon_{i, \tau}
-\end{equation} 
-with $\varepsilon_{i, \tau}$ being a random variable that is assumed to be normally distributed with a zero mean and a standard deviation of $\sigma_{R, i}$. 
-The simple market model is estimated by 
-\begin{equation}
-R_{i, j} = \alpha + \beta_i R_{m, j} + \varepsilon
-\end{equation}
-
-Where the $\beta_i$ is the relationship between the market return $R_{m, j}$ and the return on the security of interest $R_{i, j}$ and $\varepsilon$ is a random variable that represents all the idiosyncratic factors that affect the returns to the security.
-
 The lack of any clear benchmark for currency returns and the fact that theoretically exchange rate returns should be on average equal to the forward rate, have made the case for a mean model.
@@
 The evidence about returns in Tables \ref{tabref:DS1} and \ref{tabref:DS3} suggests that this is not likely to be a major problem.
+
+Is more needed here?
 
@@
 
-Again, the most prominent observation is positive association between sentiment and returns.
-Intuitively, this is not surprising.
-However, it does suggest that foreign exchange prices are not always driven by a random walk.
-There is clear evidence here of serial correlation in the returns.
-As sentiment rises (as measured by the increase in the risk-reversal) there is a price appreciation.
-However, the adjustment is not instantaneous.
-This is an example of the sort of trending behaviour that is common in the literature (as reported in Section \ref{secref:empiric}).
-It is likely to be at least partially explained by the sort of gradual adjustment to information that was identified as being a consequence of the conservatism heuristic in Section \ref{sec:behavioural} on behavioural finance.
-
 \begin{figure}[h!]
@@
 Rather they seem to be part of the price-discovery process.
-
-The second heuristic that was discussed in Section \ref{sec:behavioural} was representitiveness.
-This was the tendency of decision-makers to focus on the immediate and to forget base probabilities.
-It would suggest that investors would think too much about the new information that has driven prices towards a new equilibrium in Figures \ref{fig:ES1} and \ref{fig:ES2} not sufficient to the broad background where other currencies may be more or less attractive.
-In practical terms, it suggests that there will be overshooting and that this overshooting will cause reversals when it becomes extreme.
-There is little evidence here of any reversal.
-After the extremes are reached, there seems more likely to be a period of consolidation or a fluctuation around the new equilibrium.
-This is what would be expected if speculation were part of the process of getting information into the price.
-Once this has taken place, there is consolidation.
 
@@
 
-The R code for all the event window figures is available in Appendix \ref{AppendixB} Section \ref{Asecref:event}.
-This merely takes the data that was used to construct Tables \ref{tabref:RR1} and \ref{tabref:SP1} and puts it in graphical form.
-The event windows are stretched to give a clear view of the evolution of exchange rates  around the event.
+The event windows are stretched to give a clear view of the evolution of exchange rates  around the event and to ensure that results are not a function of the size of the event window.
 Confidence intervals are added.
@@
 \section{Speculation and foreign exchange returns}
-This chapter has sought to measure speculation and to use this measure to discover more about the relationship between speculation, information and asset prices.
+This paper has sought to measure speculation and to use this measure to discover more about the relationship between speculation, information and asset prices.
 The measurement of speculation takes two forms:  a measurement of the intensity of belief with the risk-reversal skew, and a measurement of speculative positions.
@@
 
-\bibliography{myrefs}
+\bibliography{../../myrefs}
 

### 2015-05-09 Knitr/Paper.tex
Updated the files. Now need to read through and make sure that it is ready to be sent off. What is the angle?

@@
 A unique dataset of risk-reversal skew on option prices is used to measure the weight of speculative sentiment.
-This is suplemented with regulatory data on the futures position of speculative traders.
+This is supplemented with regulatory data on the futures position of speculative traders.
 \end{abstract}
@@
 It is noise.
-that does not emerge randomly, but feeds off itsef as a feedback of sentiment  encourages  trends and serial correlation in asset prices returns.
+that does not emerge randomly, but feeds off itself as a feedback of sentiment  encourages  trends and serial correlation in asset prices returns.
 The alternative view regards speculation as an key lubricant of financial market as it provides liquidity and facilitates price discovery.
 
-The aim of this paper is to understand mroe about the nature of speculation in the foreign exchange market by assess the relationship between the intensity or the weight of speculation and the movement of prices.
+The aim of this paper is to understand more about the nature of speculation in the foreign exchange market by assess the relationship between the intensity or the weight of speculation and the movement of prices.
 If speculators are uninformed and their activity can be regarded as noise, their most extreme sentiment and the time when speculators have the greatest weight in market activity should be those when price reversals are most likely to take place.
@@
 There are three views of noise traders that are tested here:  noise traders are skill experts taking advantage of information in the order book or price movements;  rationally providing liquidity by acting as market-maker; acting irrationally as noise traders.
-Noise traders in this experiement appear to act as irrational contrarian traders.
-Noise traders refuce the efficiency of the market but increase liquidity.
+Noise traders in this experiment appear to act as irrational contrariety traders.
+Noise traders reduce the efficiency of the market but increase liquidity.
 
 Noise is related to rational expectations.
-A mean zero, normally-distributed random variable influences supply citet{Muth1961Rational}.
+A mean zero, normally-distributed random variable influences supply \citet{Muth1961Rational}.
 Why do noise traders exist?
@@
 The effect of noise traders on liquidity is unclear.
-They may supply liquidity by providing limit orders and making some returns for providing liquidity; they may increase bid-ask spreads by using market orders to take advantage of (potentially non-existant) informational advantages.
-Noise traders may increase (if they have some priviledged information, they add to the price pressure) or reduce efficiecy (if they increase liquiity and reduce the ratio of informed to uninformed traders).
-
-Are day traders profitable? Are these the face of noise traders?  How does this show up in teh foreign exchange market?
-
-\citet{da2010sum} Seek to measure invetor sentiment and assess how it affects securities prices.
+They may supply liquidity by providing limit orders and making some returns for providing liquidity; they may increase bid-ask spreads by using market orders to take advantage of (potentially non-existent) informational advantages.
+Noise traders may increase (if they have some privileged information, they add to the price pressure) or reduce efficiency (if they increase liquidity and reduce the ratio of informed to uninformed traders).
+
+Are day traders profitable? Are these the face of noise traders?  How does this show up in the foreign exchange market?
+
+\citet{da2010sum} Seek to measure investor sentiment and assess how it affects securities prices.
 They create a index of investor sentiment from web searches and use this to predict return reversals.
-The resutls show that sentiment can predict market turns.
-This is a contrast to traditional ways of measuring sentiment:  proxy for sentiment with market-based measures like volume, closed-end fund discount, IPO first-day returns, IPO volume, option implied volatities and mutual fund flows.
+The results show that sentiment can predict market turns.
+This is a contrast to traditional ways of measuring sentiment:  proxy for sentiment with market-based measures like volume, closed-end fund discount, IPO first-day returns, IPO volume, option implied volatilities and mutual fund flows.
 Baker and Wurgaler (2007) have a comprehensive survey.
 This is more than just investors sentiment.
-There is an issue with the measurment of an input (sentiment) with an output (market price).
+There is an issue with the measurement of an input (sentiment) with an output (market price).
 Use of surveys like the University of Michigan consumer sentiment survey etc.
@@
 There is a positive correlation between the FEARS index and bond prices.
-However, thsi also reverses in following days.
-There is a relatioknship betwen the FEARS index and daily market volatility (realised volatlity and the VIX index).
+However, this also reverses in following days.
+There is a relationship between the FEARS index and daily market volatility (realised volatility and the VIX index).
 There is also evidence of inflows and outflows from mutual funds.
@@
 It is usually assumed to be hedging or liquidity.
-There is also a \emph{limits-to-arbitrage} literature that uses the term to identify traders that are not movivated by standard explanations of fundamental or liquidity provision.
-This paper seeks to distinguish between those that trade for liquidty or heging and those that trade because of psychological biases.
+There is also a \emph{limits-to-arbitrage} literature that uses the term to identify traders that are not motivated by standard explanations of fundamental or liquidity provision.
+This paper seeks to distinguish between those that trade for liquidity or hedging and those that trade because of psychological biases.
 Fisher Black ``People who trade on noise are willing to trade even though from an objective point of view they would be better off not trading.
@@
 Shliefer and Summmer (1990) use momentum.
-However, Kaniel, Saar and Titman (2006) find evidence of contrarian strategies.
+However, Kaniel, Saar and Titman (2006) find evidence of contrariety strategies.
 This is associated with short-term speculation.
@@
 
-There is a quesiton about whether experimental markets (like this paper) are superior to actual markets where the identification of trader types is difficult.
+There is a question about whether experimental markets (like this paper) are superior to actual markets where the identification of trader types is difficult.
 Theory - Delong, summers, shleifer etc,  Scheinkman and Xiong (2003).
@@
 
-Interesting difference between supplying and demanding liqudiity.
-Limit orders supply liquidity, market orders demand liqudity.
+Interesting difference between supplying and demanding liquidity.
+Limit orders supply liquidity, market orders demand liquidity.
 The activity of liquidity traders will tend to change through the course of the session, moving from supplying to demanding liquidity.
@@
 The presence of noise traders will increase pricing errors.
-However, this depends on fundametal traders have valuable information.
-
-There are a number of different ways that noise traders could behave:  activing as SOES bandits (using the NASDAQ small order execution system to make profits) or by using information about the order book; by providing liqudity like a market-maker as a service to liqudiity-traders as in Grossman Miller (1988);  as irrational noise traders, using trading rules or popular models (Shiller 1984, 1990) positive feedback or contrarian.
-
-Footnote -  There is an empirical evidence that individual investors in various countries trade in a contrarian fashion (e.g., Choe, Kho, and Stulz [1999], Grinblatt and Keloharju [2000, 2001], Jackson [2003], Richards [2005], Kaniel, Saar, Titman [2006]).
-For experimental evidence consistent with contrarian behavior see
+However, this depends on fundamental traders have valuable information.
+
+There are a number of different ways that noise traders could behave:  activating as SOES bandits (using the NASDAQ small order execution system to make profits) or by using information about the order book; by providing liquidity like a market-maker as a service to liquidity-traders as in Grossman Miller (1988);  as irrational noise traders, using trading rules or popular models (Shiller 1984, 1990) positive feedback or contrariety.
+
+Footnote -  There is an empirical evidence that individual investors in various countries trade in a contrariety fashion (e.g., Choe, Kho, and Stulz [1999], Grinblatt and Keloharju [2000, 2001], Jackson [2003], Richards [2005], Kaniel, Saar, Titman [2006]).
+For experimental evidence consistent with contrariety behaviour see
 Bloomfield, Tayler, and Zhou (2997)
@@
 Dow and Gorton (2006) suggest that this is the case and that this is an outcrop of the rational expectations revolution.
-Informed traders profit by exploiting the diviations from equilibrium.
-This is the beneifts of Grossman-Stiglitz.
+Informed traders profit by exploiting the divinations from equilibrium.
+This is the benefits of Grossman-Stiglitz.
 What is the source of this noise?
-What are the insitiutions?
-The importance of noise traders here is that they represent a zero-mean, normally distributed random vaiable.
+What are the institutions?
+The importance of noise traders here is that they represent a zero-mean, normally distributed random viable.
 This is also the description that is provided in Kyle (1985).
@@
 How the average may not be zero.
-There is a range of behavioural biases that may account for this (see Barberis and Thaler [2003]).The literaure on hetrogenous traders also focuses on noise traders as trend chasers.
+There is a range of behavioural biases that may account for this (see Barberis and Thaler [2003]).The literature on heterogeneous traders also focuses on noise traders as trend chasers.
 These models (see for example Brock and Hommes [1998]; Lux [1995]; Lux and Marchesi [2000]) investigate bubbles and crashes, and feature fundamental traders versus noise traders.
 In these models, it is the noise traders who
-induce the aberrant market behavior.
+induce the aberrant market behaviour.
 
 \href{http://papers.ssrn.com.ezproxy.liv.ac.uk/sol3/papers.cfm?abstract_id=869827}{Do noise traders move markets?}  There is strong evidence of herding by individual investors.
-THere is also momentum.
-There is reversal:  stocks that were heavily bought one year underperform; stocks that were heavily sold, outperform.
+There is also momentum.
+There is reversal:  stocks that were heavily bought one year under perform; stocks that were heavily sold, outperform.
 The opposite in the short run:  heavily bought one week generate large returns the next week.
@@
 \href{http://restud.oxfordjournals.org.ezproxy.liv.ac.uk/content/60/1/1.full.pdf+html}{Campbell and Kyle, Smart money, noise trading and stock market behaviour}.
-In response to the evidence that the volatilty of stock prices is too great to be a function of discounted expected future earnings (Shiller (1981),  LeRoy and Porter (1981), this paper assesses whether stocks are influnced by exogenous, serially correlated noise.
+In response to the evidence that the volatility of stock prices is too great to be a function of discounted expected future earnings (Shiller (1981),  LeRoy and Porter (1981), this paper assesses whether stocks are influenced by exogenous, serially correlated noise.
 In this model, the noise arises from the interaction of noise traders.
@@
 There has been an accumulation of evidence to suggest that financial asset prices do not evolve as a random walk.
-Serial correlation or trends are evident suggesting that, contrary to the efficient market hypthesis, future price changes could be partially foreseen by the movement of past price changes.
+Serial correlation or trends are evident suggesting that, contrary to the efficient market hypothesis, future price changes could be partially foreseen by the movement of past price changes.
 There are a number of studies that find evidence that the reaction of stocks to information is less than instantaneous.
@@
 It is shown that book-to-market and cash-flow ratios have the most positive impact on positive returns.
-Lakonishok, Shleifer and Vishny provide an overview of \emph{contrarian} investment strategies \citep{LakonishokContrarian}.
+Lakonishok, Shleifer and Vishny provide an overview of \emph{contrariety} investment strategies \citep{LakonishokContrarian}.
 This whole area is closely linked to the debate about the relative merits of \emph{value} and \emph{growth} stocks, with ample evidence from academic and practitioners that while growth stocks outperform value in the short-run, in line with the evidence of momentum, there are frequent reversals as over-reactions are reversed.
@@
 \citep{LoMacKinlay1990} argue that one reason for the serial correlation in prices is the asynchronous trading system and the time lag between some stocks reacting to market news.
-In other words, it could be that the underperforming stocks have not yet reacted to the news that the outperforming stocks have already incorporated.
+In other words, it could be that the under performing stocks have not yet reacted to the news that the outperforming stocks have already incorporated.
 However, this idea is rejected by Jagadeesh and Titman as a result of their method of decomposing returns into market and specific factors \citep[p.
@@
 ``An abundance of research has shown that human beings are conservative processors of fallible information.
-Such experiments compare human behavior to the outputs of Bayes's theorem, the formally optimal rule about how opinions (that is, probabilities) should be revised on the basis of new information.
+Such experiments compare human behaviour to the outputs of Bayes's theorem, the formally optimal rule about how opinions (that is, probabilities) should be revised on the basis of new information.
 It turns out that opinion change is very \emph{orderly}, and usually proportional to number calculated from Bayes's theorem  - but it is insufficient in amount.

### 2015-05-10 Knitr/Paper.tex
Made a few changes.  Looks like a complete start is more apprppriate.

@@
 \begin{abstract}
-An event study of extreme speculation is used to assess the Keynesian and Friedman views of speculation.
-This is partly a test of the informational content of speculation.
+An event study of extremes is used to assess the Keynesian and Friedman views of speculation.
 Speculative extremes are ranked by weight of activity and intensity of sentiment.
@@
 A unique dataset of risk-reversal skew on option prices is used to measure the weight of speculative sentiment.
-This is supplemented with regulatory data on the futures position of speculative traders.
+This is supplemented with regulatory data on the weight of speculative positions.
 \end{abstract}
@@
 
-This is a contrast between an idea of speculative activity as a nuisance that, at its most extreme, will likely to cause bubbles and reversal and will discourage fundamental long-term investors from participating in the market.
-It is noise.
-that does not emerge randomly, but feeds off itself as a feedback of sentiment  encourages  trends and serial correlation in asset prices returns.
+This is a contrast between an idea of speculative activity as a nuisance that, at its most extreme, will likely to cause booms and reversals as bubbles expand and burst to discourage fundamental long-term investors from participating.
+It is noise that does not emerge randomly, but feeds off itself as a feedback of sentiment  encourages  trends and serial correlation in asset price returns.
 The alternative view regards speculation as an key lubricant of financial market as it provides liquidity and facilitates price discovery.
-
-The aim of this paper is to understand more about the nature of speculation in the foreign exchange market by assess the relationship between the intensity or the weight of speculation and the movement of prices.
+In this view, speculation is stabilising and its absence will increase volatility and will undermine the process of price discovery.
+
+The aim of this paper is to understand more about the nature of speculation by focusing on the foreign exchange market, using this to assess the relationship between the intensity or the weight of speculation and the movement of prices.
 If speculators are uninformed and their activity can be regarded as noise, their most extreme sentiment and the time when speculators have the greatest weight in market activity should be those when price reversals are most likely to take place.
-An event study is used to assess the relationship between speculative extremes and foreign exchange returns.
 A unique set of option prices are used to measure the intensity of speculative sentiment.
 US regulatory positions measure the prominence of speculators in the market.
+If speculation is stabilising, the largest weight of positions should be associated with stability and extremes of sentiment will drive prices towards equilibrium; if specualtion is destabilising, the largest positions should increase volatilit and extemes of sentiment would increase the risk of violent reversal back towards equilibrium.
 

### 2015-05-30 R/Futures.R
Tidied up the repo ready for the re-write The R code has been put into R.  There used to be a directory that was called Rcode.  Files from that were moved to R.  They had the same name. I think that this was noted.  There was no warning.  I tested one.  THe files exist somewhere else.

@@ -1,4 +1,4 @@
- # The following code draws and prepares the data from the file 
+# The following code draws and prepares the data from the file 
 # (Futures.csv) and calculates the descriptive statistics of exchange 
 # rates and measures of sentiment.    
 # Read spec position data----------------------------------

### 2015-05-30 Rcode/Delong.R

@@ -1,121 +1,0 @@
-# Build the DeLong function made of 'rho' as the actual misperception, rhomu as
-# the average misperception, r as the risk free rate, mu is the weight of 
-# speculators in the total market participants, ra as the parameter for 
-# risk aversion, rhosd is the standard deviation of the misperception. 
-require(zoo)
-require(eventstudies)
-# Function to calculate price under DeLong model-----------------
-DeLong <- function(rho, rhomu, r, mu, ra, rhosd){
-  p = 100 + (mu * (rho - rhomu))/(1 + r) + (mu * rhomu)/r - 
-    ((2 * ra * mu^2 * rhosd^2)/(r * (1 + r)^2))
-  return(p)
-}
-# Set misperception as a random vaiable with a mean of 1 and a 
-# standard deviation of 1
-set.seed(1)
-rho <- rnorm(1000, 4, 2)
-p <- DeLong(rho = rho, rhomu = 4, r = 0.1, mu = 0.5,
-            ra = 0.3, rhosd = 2)
-length(p)
-plot(p, type = 'l', main = "Delong model with misperception as random variable")
-da <- cbind(p, rho)
-head(da)
-pz = as.zoo(da)
-# Create returns
-pz$r <- 100*diff(log(pz$p))
-str(pz)
-head(pz)
-# Create extreme as above 95th percentile
-pz$e = as.numeric(pz$rho >= quantile(pz$rho, 0.95, na.rm = TRUE))
-head(pz)
-eventslist.ex <- data.frame(index(pz$e[pz$e == 1]))
-eventslist.ex$unit <- as.character("r")
-colnames(eventslist.ex) <- c("when", "unit")
-dim(eventslist.ex)
-tail(eventslist.ex)
-# Align events
-a <- phys2eventtime(pz, eventslist.ex, width = 2)
-P.ex <- window(a$z.e, start = 0, end = 1)
-P.ex
-# Calculate descriptive statistics of rho extreme-------------------
-P.ext <- t(P.ex)
-P.ext
-mystats <- function(x, na.omit=FALSE){
-  if (na.omit)
-    x <- x[!is.na(x)]
-  m <- mean(x)
-  s <- sd(x) 
-  t <- m/s
-  n <- length(x)
-  skew <- sum((x-m)^3/s^3)/n
-  ps <- ((6*n*(n-1))/((n-1)*(n+1)*(n+3)))^0.5
[…]

### 2015-05-30 Rcode/Futures.R

@@ -1,123 +1,0 @@
-# The following code draws and prepares the data from the file 
-# (Futures.csv) and calculates the descriptive statistics of exchange 
-# rates and measures of sentiment.    
-# Read spec position data----------------------------------
-require(eventstudies)
-da <- read.csv("./Futures.csv", header = TRUE)
-da$S1 <- (da$NCL-da$NCS)/(da$NCL+da$NCS+da$NCSP)
-da$S2 <- (da$NCL-da$NCS)/da$OI
-da$S3 <- (da$NCL+da$NCS+da$NCSP)/da$OI
-da <- da[,c(2, 3, 5, 19, 20, 21)]
-str(da)
-head(da)
-# Sub set files------------------------------------------
-da.CAD <- subset(da[,2:6], subset = da$Exchange == "CAD")
-da.CHF <- subset(da[,2:6], subset = da$Exchange == "CHF")
-da.EUR <- subset(da[,2:6], subset = da$Exchange == "EUR")
-da.GBP <- subset(da[,2:6], subset = da$Exchange == "GBP")
-da.JPY <- subset(da[,2:6], subset = da$Exchange == "JPY")
-# Calculate descriptive stats---------------------------
-mystats <- function(x, na.omit=FALSE){
-  if (na.omit)
-    x <- x[!is.na(x)]
-  m <- mean(x)
-  s <- sd(x) 
-  t <- m/s
-  med <- median(x)
-  n <- length(x)
-  skew <- sum((x-m)^3/s^3)/n
-  ps <- ((6*n*(n-1))/((n-1)*(n+1)*(n+3)))^0.5
-  kurt <- sum((x-m)^4/s^4)/n - 3
-  pk <- ((n^2-1)/((n-3)*(n + 5)))^0.5
-  max <- max(x)
-  min <- min(x)
-  return(c(n=n, mean=m * 100, t= t, medium = med *100, stdev=s *100, 
-           skew=skew, ps = ps, kurtosis=kurt, pk = pk, max = max *100, min = min * 100))
-}
-# change currency file for each exchange rates (ie da.CHF etc). 
-FX.d <- apply(da.CAD, 2, FUN = mystats, na.omit = TRUE)
-FX.d
-# Create extreme events and conditions----------------------
-# Set the conditions for each run
-# currency is CAD, GBP, JPY, EUR or CHF
-currency = "CHF"
-# window is number of weeks
-wind = as.numeric(2)
[…]

### 2015-05-30 Rcode/RRGraph.R

@@ -1,61 +1,0 @@
-# This will create the graph for paper
-# See event(final) for fuller explanation
-rm(list = ls())
-ls()
-require(eventstudies)
-require(PerformanceAnalytics)
-require(xtable)
-################################################################################
-par(mfrow = c(3,2))
-currency = "AUD"
-extreme = "low"
-# select percentile
-hi = as.numeric(0.95)
-lo = as.numeric(0.05)
-# select window either "whole lot" or "after".
-cum = "whole lot"
-wind = as.numeric(16) 
-###############################################################################
-daz <- read.zoo("C:/Users/Toshiba/Desktop/Event2.csv", header = TRUE, 
-  format = "%d-%b-%y", sep = ",")
-FX.z <-daz[,c(paste(currency), paste(currency, "RR", sep = ""))]
-FX.z$r <- 100*diff(log(FX.z[,1]))
-(eventslist.FXex)
-##################################################################################
-if(extreme == "high"){
-FX.z$ex <- as.numeric(FX.z[,2] >= quantile(FX.z[,2],  
-         hi, na.rm = TRUE)) 
-} else if(extreme == "low") {
-FX.z$ex <- as.numeric(FX.z[,2] <= quantile(FX.z[,2], 
-         lo, na.rm = TRUE))
-} else(print("No Extreme"))
-################################################################################
-eventslist.FXex <- data.frame(index(FX.z$ex[FX.z$ex ==1]))
-eventslist.FXex$unit <- as.character("r")
-colnames(eventslist.FXex) <- c("when", "unit")
-eventslist.FXex$when <-as.Date(eventslist.FXex$when)
-a <- phys2eventtime(FX.z, eventslist.FXex, width = wind)
-################################################################################
-head(a$z.e)
-if(cum == "whole lot"){
-FX.ex <- window(a$z.e, start = wind * -1, end = wind)
-} else if(cum == "after"){
-FX.ex <- window(a$z.e, start = 0, end = wind)
-} else(print("Error"))
-################################################################################
-FX.exc <- remap.cumsum(FX.ex, is.pc = FALSE, base = 0)
-###############################################################################
-# 
-# the bootstrap
-dim = seq(-wind, wind, 1)
-FX.exc.b <- inference.Ecar(FX.exc)
-FX.exc.b
-FX.exc.bm <- matrix(FX.exc.b, ncol = 3, dimnames = list(dim, 
[…]

### 2015-05-31 Knitr/NewPaper.Rnw
Updated some of the NewPaper. Still working on the literature review.

\documentclass[12pt, a4paper, oneside]{article} % Paper size, default font size and one-sided paper
%\graphicspath{{./Figures/}} % Specifies the directory where pictures are stored
%\usepackage[dcucite]{harvard}
\usepackage{rotating}
\usepackage{amsmath}
\usepackage{setspace}
\usepackage{pdflscape}
\usepackage[flushleft]{threeparttable}
\usepackage{multirow}
\usepackage[comma, sort&compress]{natbib}% Use the natbib reference package - read up on this to edit the reference style; if you want text (e.g. Smith et al., 2012) for the in-text references (instead of numbers), remove 'numbers' 
\usepackage{graphicx}
%\bibliographystyle{plainnat}
\bibliographystyle{agsm}
\usepackage[colorlinks = true, citecolor = blue, linkcolor = blue]{hyperref}
%\hypersetup{urlcolor=blue, colorlinks=true} % Colors hyperlinks in blue - change to black if annoying
%\renewcommand[\harvardurl]{URL: \url}
 \usepackage{listings}
  \usepackage{tikz}
 \usetikzlibrary{arrows,positioning}
 \usepackage{color}
 %\graphicspath{{../Pictures/}}
\definecolor{mygrey}{gray}{0.95}
\lstset{backgroundcolor=\color{mygrey}}
\begin{document}
\title{Speculation in foreign exchange: noise or information}
\author{Rob Hayward\footnote{University of Brighton Business School, Lewes Road, Brighton, BN2 4AT; Telephone 01273 642586.  rh49@brighton.ac.uk. }} 
\date{\today}
\maketitle
\begin{abstract}
An event study of extremes is used to assess the Keynesian and Friedman views of speculation.    Speculative extremes are ranked by weight of activity and intensity of sentiment.  In either case, the extreme says little about the future, suggesting that speculation is more than just random noise and supporting the idea there is some informational content that is being passed through to the price. A unique dataset of risk-reversal skew on option prices is used to measure the intensity of speculative sentiment.  This is supplemented with regulatory data on the weight of speculative positions.  
\end{abstract}
\section*{Introduction}
  There are a range of ideas about the way that speculation can affect financial markets.  Towards one end of the spectrum is the view of Keynes from Chapter 13 of \emph{The General Theory} where he identified speculation as a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd'' \citet[p. 101]{Keynes1936}. Here speculative activity is a problem that, at its most extreme, will cause booms, reversals and will discourage fundamental long-term investors from participating. It is noise that does not emerge randomly, but feeds off itself as a feedback of sentiment  encourages  trends and serial correlation in asset prices.

At the other end, Milton Friedman argued that speculation was a stabilising force, providing liquidity and helping to ensure that markets swiftly find equilibrium. In arguing in favour of flexible exchange rates, Friedman asserted that successful speculators would be those that managed to correctly buy when prices  were below fundamental value and sell above that price \citet{FriedmanPositive}. Therefore, the speculative process would favour those able to access or utilise information over those that trading on \emph{noise} and, through a Darwinian process, informed speculation would dominate.  This view regards speculation as an key lubricant of financial market as it provides liquidity and facilitates price discovery. Here, speculation is stabilising and its absence will increase volatility and will undermine the process of price discovery.

The aim of this paper is to understand more about the nature of speculation by investigating the relationship between both the intensity of speculation and the weight of speculataive positions with the movement of prices in the foreign exchange market.   If speculators are uninformed and their activity can be regarded as noise, their most extreme sentiment and the time when speculators have the greatest weight in market activity should be those when price are driven from their fundamnetal value and reversals are most likely to take place; if speculators are informed price should follow a random walk after the extremes.

A unique set of option prices are used to identify the intensity of speculative sentiment.  US regulatory positions measure the prominence of speculators in the market. Speculation does not appear to be random, there is momentum to the rise and fall in sentiment and positions.  However, while currency prices co-move with positions and sentiment, there is no informational content from these positions, supporting the suggestion that speculative actvity has a permanent affect on prices and that is consistent with speculation being part of the process of price discovery. Speculators do not seem to drive prices from fundamental value and their presence does not seem to hamper the incorporation of information into prices.

\section{Literature}
\subsection*{Overview}
  Speculation is an important concept in economics, particular in the area of finance.  However, there is no clear, unambiguous theory of specuation. In a number of fields the term \emph{noise-trading} has been used.   \citet{bloomfield2009noise} suggests that there are three categories of noise-trader: uninformed speculators, market-making liquidity-providers and informed speculators. The first of these groups is similar to the picture of speculation provided by Keynes, the second refers to those agents or institutions paid for providing liqidity and for taking the risk of meeting an informed trader in the market (these are also called market-makets), and the final category is close to the idea of speculation that has been presented by Fridman.

In the \emph{microstrucrture} literature, the term noise-trader is interchanged with \emph{liquidity-trader} to mean those trading with a lack of information and those trading for liquidty reasons.  They do not usually distinguish between institutional and psychological reasons for the lack of attention to fundametal information. Nonetheless, liquidity tends to refer to those trading without reference to fundamental value: they must execute orders.  This may mean that they just execute customer orders at the current price or it may be that they buy and sell with reference to a benchmark index or to fulfil a hedging strategy.  Examples of the microstructure approach are found in \citet{Kyle1985Continuous} and \citet{glosten1985bid}.

%this paragraph must go somewhere else.  I do not know where. 
Nonetheless, the distinction between market-making speculative activity is not precise.\footnote{The implementation of the \emph{Volcker} Rule in the US which will limit the scope of US investment banks to trade on their own account has raised the question of where market-making turns into speculation.}

The other question is whether either of these types uninformed speculative noise-trading can be considered as being just random noise?    \citet[p. 1]{GortonNoise}  say that noise-traders are all those trading in markets for ``non-information-based reasons''. This is a solution to the  \citet{Grossman1980Impossibility} paradox that will allow those traders prepared to pay the cost of acquiring information to take advantage of it. Informed traders, knowing that the current price is below fundamental value, can purchase from noise-traders, who either do not know or do not care. The noise drives prices from their fundamental value and ensures that the cost of acquiring information can be regained and therefore provides the incentive for informed traders to improve the informational efficiency of asset prices.

However, though the idea that noise is a random variable may be a construction that is useful for modelling and it may be justified in analysis of broader themes of microstructure and levels of market efficiency, this exogenous construct may be an over-simplification that is not apprpropriate when looking for more understanding of speculation itself. Institutional and psychological explanations for noise-trading suggest that there are endogenous elements that suggest non-random effects.The modelling of noise is related to the rise of rational expectations in economics.  A mean zero, normally-distributed random variable influences supply in the original version supplied by \citet{Muth1961Rational}.  
\subsection*{The problem of uninformed noise-traders}
The traditional Keynesian view of speculation asserted that it will be damaging because it will make it even harder for informed-traders to overcome undertainty and apply fundamental information about long-term value. This is view finds some supported in Shiller's observation that stock prices are more volatile than the underlying earnings information on which valuation should be based \citet{ShillerVol}. In addition, \citet{Delong1990noise} and \citet{shleifer1997limits} show that constrained horizons can cause noise-trading to limit the ability of informed traders to take advantage of profitable opportunities for fear of further divergence from fundamental value before convergence is achieved.

 Shliefer, Summers, DeLong etc identify the limits to arbitrage. The other writings are Battalio, Hatch and Jennings (1997), Harris and Schultz (1998) and Scheinkman and Xiong (2003).  Lowenstein and Willard (2006)

It would be good to have some more modern ideas here.  Maybe they are just the empirical results. 
\subsection{Interplay between Informed and uninformed}
Fisher Black identifies noise as ``the arbitrary element in expectations'' \citet[p.529]{BlackNoise}. He cites a number of ways that noise affects the trading in financial markets.  He speaks about people trading on noise over information; he says that noise makes financial markets possible but imperfect. The more noise-traders that there are, the more liqiity and the easier to trade; the more noise-traders the higher the level of inefficiency in the market and the more likely that prices will be driven from their fundamental value. He suggest that the noise will limit the scope for complete arbitrage and that it will be difficult to distinguish between noise and information traders.

  Why do noise traders exist?  There are liquidity and behavioural explanations. There are also a number of models that show noise traders as momentum traders.

\subsection{Other empirical work}
The assessmen of these theoreis about speculation and noise-trading is rather limited. In part this is becuase information is limited and it is usually difficult to distiguish between speculators and other investors and, even more so, between informed and uninformed.

Some of the empirical work will identify specuation or noise with momentum trading. 
\citet{da2010sum} seek to measure investor sentiment and assess how it affects securities prices. they create a index of investor sentiment from web searches and use this to predict return reversals.  the results show that sentiment can predict market turns. this is a contrast to traditional ways of measuring sentiment:  proxy for sentiment with market-based measures like volume, closed-end fund discount, ipo first-day returns, ipo volume, option implied volatilities and mutual fund flows. baker and wurgaler (2007) have a comprehensive survey. this is more than just investors sentiment. there is an issue with the measurement of an input (sentiment) with an output (market price). use of surveys like the university of michigan consumer sentiment survey etc.  there is a negative contemporaneous correlation between fears and stock market returns. however, there is a reversal in the relationship after two days.  this is consistent with ``sentiment induced temporary mispricing'' \citet[p. 4]{da2010sum}.  there is a positive correlation between the fears index and bond prices.  however, this also reverses in following days. there is a relationship between the fears index and daily market volatility (realised volatility and the vix index). there is also evidence of inflows and outflows from mutual funds.  this may help to explain the contrast between evidence from the equity market and the bond market.

the effect of noise traders on liquidity is unclear. they may supply liquidity by providing limit orders and making some returns for providing liquidity; they may increase bid-ask spreads by using market orders to take advantage of (potentially non-existent) informational advantages. noise traders may increase (if they have some privileged information, they add to the price pressure) or reduce efficiency (if they increase liquidity and reduce the ratio of informed to uninformed traders). 
are day traders profitable? are these the face of noise traders?  how does this show up in the foreign exchange market?  There is a very recent paper. \href{http://www.sciencedirect.com/science/article/pii/S1042443115000682}{Paper on profitability of fx trading}.

Noise traders exert positive effects on liquidity but weaken price reversals. a tax on trading will reduce noise traders but will also reduce the influence of fundamental traders. their experiment distinguishes between informed traders, noise traders and liquidity traders.  \emph{soes bandits} are those taking advantage of the soes system to trade quickly and take advantage of small price movements. there is a lot of focus on the effect that taxation may have on the different categories of trader. there is an experiment that allows much more to be held constant. there are three views of noise traders that are tested here:  noise traders are skill experts taking advantage of information in the order book or price movements;  rationally providing liquidity by acting as market-maker; acting irrationally as noise traders. noise traders in this experiment appear to act as irrational contrariety traders. noise traders reduce the efficiency of the market but increase liquidity.
in theory, traders may trade on mistaken fundamental information.  shliefer and summmer (1990) use momentum.  however, kaniel, saar and titman (2006) find evidence of contrariety strategies. this is associated with short-term speculation. does the activity drive the price away from fundamental value? stiglitz (1990) is one of the arguments in favour of a tobin tax.

there is a question about whether experimental markets (like this paper) are superior to actual markets where the identification of trader types is difficult.  theory - delong, summers, shleifer etc,  scheinkman and xiong (2003).  lowenstein and willard (2006).  empirical work:  (see, for example, garvey and murphy [2001], linnainmaa [2003], barber et. al. [2004], and kaniel, saar, and titman [2006].

\href{http://papers.ssrn.com.ezproxy.liv.ac.uk/sol3/papers.cfm?abstract_id=869827}{do noise traders move markets?}  there is strong evidence of herding by individual investors. there is also momentum. there is reversal:  stocks that were heavily bought one year under perform; stocks that were heavily sold, outperform. the opposite in the short run:  heavily bought one week generate large returns the next week.

\href{http://restud.oxfordjournals.org.ezproxy.liv.ac.uk/content/60/1/1.full.pdf+html}{campbell and kyle, smart money, noise trading and stock market behaviour}. in response to the evidence that the volatility of stock prices is too great to be a function of discounted expected future earnings (shiller (1981),  leroy and porter (1981), this paper assesses whether stocks are influenced by exogenous, serially correlated noise. in this model, the noise arises from the interaction of noise traders.

\subsection{why this is important}
The nature of speculation and the role of noise-traders has become increasingly imporant since the golbal financial crisis (GFC).  There are a number of comtemporary questions where the role and nature of speculation is at the center of the debate.  Regulations or taxes that reduce the reduce the amount of speculation have generally been considered to be helpful in reducing risk and encouraging more long-term attitudes towards investment.  However, where these measures reduce liquidity, higher risk and an increase in other costs may provide an countervailing force.  For example, the so-called \emph{Volcker Rule} would prevent US inventment banks trading on their own-account.  This, it has been argued, will reduce liquidity in the bond market by causing investment banks to hold lower levels of bond inventory on their books.  This may increase the amount of \emph{price slippage} when final investor adjust their positions.

for related papers on securities transaction taxes see stiglitz [1989], summers and summers [1989], 
amihud and mendelson [1993], schwert and seguin [1993], umlauf [1993], subrahmanyam [1998], dow 
and rahi [2000], habermeir and kirilenko [2003], and hinman [2003].

interesting difference between supplying and demanding liquidity.  limit orders supply liquidity, market orders demand liquidity.  the activity of liquidity traders will tend to change through the course of the session, moving from supplying to demanding liquidity. noise traders tend to act in the opposite manner.  they tend to supply more liquidity as the session progresses. black (1986) noise traders put noise into prices.  the effect of noise traders on liquidity: price impact of trades is lower when noise traders are present.  however, this is mostly in the temporary impact on prices rather than the permanent impact. therefore, noise traders tend to reduce the amount of volatility in prices because market orders do not temporarily have an effect and then reverse.  the presence of noise traders will increase pricing errors. however, this depends on fundamental traders have valuable information.

there are a number of different ways that noise traders could behave:  activating as soes bandits (using the nasdaq small order execution system to make profits) or by using information about the order book; by providing liquidity like a market-maker as a service to liquidity-traders as in grossman miller (1988);  as irrational noise traders, using trading rules or popular models (shiller 1984, 1990) positive feedback or contrariety.

footnote -  there is an empirical evidence that individual investors in various countries trade in a contrariety fashion (e.g., choe, kho, and stulz [1999], grinblatt and keloharju [2000, 2001], jackson [2003], richards [2005], kaniel, saar, titman [2006]). for experimental evidence consistent with contrariety behaviour see 
bloomfield, tayler, and zhou (2997)

 what is the source of this noise?  what are the institutions? the importance of noise traders here is that they represent a zero-mean, normally distributed random viable.  this is also the description that is provided in kyle (1985).  liquidity traders trade due to imbalance in the timing of consumption or from portfolio considerations. shleifer, summers and delong have trading activity that is affected by sentiment, models or "fads".  this is important.  how the average may not be zero. there is a range of behavioural biases that may account for this (see barberis and thaler [2003]).the literature on heterogeneous traders also focuses on noise traders as trend chasers.  these models (see for example brock and hommes [1998]; lux [1995]; lux and marchesi [2000]) investigate bubbles and crashes, and feature fundamental traders versus noise traders. in these models, it is the noise traders who 
induce the aberrant market behaviour.

there is also the issue of speculation in commodities markets.  if this is random, the speculation does not affect the overall price, if this is ,liquidity-driven, it may affect the price level.

fisher black says ```anything that changes the amount of character of noise trading will change the volatility of price." \citet[p. 533]{blacknoise}.

% does this now need something on the reason why the foreign exchange market is used?  could be the keynes speuclative activity, fridman paper, liquidity etc. 
\bibliography{../../../../myrefs}

\end{document}

### 2015-06-01 Knitr/NewPaper.Rnw
Major adjustments to the lit review. Last section is to be completed.

@@
 An event study of extremes is used to assess the Keynesian and Friedman views of speculation.
-Speculative extremes are ranked by weight of activity and intensity of sentiment.
-In either case, the extreme says little about the future, suggesting that speculation is more than just random noise and supporting the idea there is some informational content that is being passed through to the price.
+Speculative extremes are ranked by intensity of sentiment and weight of activity.
+The extreme says little about the future in either case, suggesting that speculation is more than just random noise and supporting the view there is some informational content that is being passed through to the price with speculative activity.
 A unique dataset of risk-reversal skew on option prices is used to measure the intensity of speculative sentiment.
@@
 There are a range of ideas about the way that speculation can affect financial markets.
-Towards one end of the spectrum is the view of Keynes from Chapter 13 of \emph{The General Theory} where he identified speculation as a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd'' \citet[p.
+Towards one end of this spectrum is the view of Keynes from Chapter 13 of \emph{The General Theory} where he identified speculation as a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd'' \citet[p.
 101]{Keynes1936}.
-Here speculative activity is a problem that, at its most extreme, will cause booms, reversals and will discourage fundamental long-term investors from participating.
-It is noise that does not emerge randomly, but feeds off itself as a feedback of sentiment  encourages  trends and serial correlation in asset prices.
+Here speculative activity is a problem that, at its most extreme, will cause booms, reversals and will discourage fundamental, long-term investors from participating.
+It is noise that does not emerge randomly, but feeds off itself as a collective sentiment  encourages  trends and serial correlation in asset prices.
 
-At the other end, Milton Friedman argued that speculation was a stabilising force, providing liquidity and helping to ensure that markets swiftly find equilibrium.
-In arguing in favour of flexible exchange rates, Friedman asserted that successful speculators would be those that managed to correctly buy when prices  were below fundamental value and sell above that price \citet{FriedmanPositive}.
-Therefore, the speculative process would favour those able to access or utilise information over those that trading on \emph{noise} and, through a Darwinian process, informed speculation would dominate.
+Towards the other end, Milton Friedman argued that speculation was a stabilising force, providing liquidity and helping to ensure that markets swiftly find equilibrium.
+In arguing in favour of flexible exchange rates, Friedman asserted that successful speculators would be those who managed to correctly buy when prices  were below fundamental value and sell above that price \citet{FriedmanPositive}.
+Therefore, Friedmand argued, the speculative process would favour those able to access or utilise information over the uninformed and, through a Darwinian process, informed speculation would dominate.
 This view regards speculation as an key lubricant of financial market as it provides liquidity and facilitates price discovery.
-Here, speculation is stabilising and its absence will increase volatility and will undermine the process of price discovery.
 
-The aim of this paper is to understand more about the nature of speculation by investigating the relationship between both the intensity of speculation and the weight of speculataive positions with the movement of prices in the foreign exchange market.
-If speculators are uninformed and their activity can be regarded as noise, their most extreme sentiment and the time when speculators have the greatest weight in market activity should be those when price are driven from their fundamnetal value and reversals are most likely to take place; if speculators are informed price should follow a random walk after the extremes.
+The aim of this paper is to understand more about the nature of speculation by investigating the relationship between both the intensity of speculation and the weight of speculative positions with the movement of prices in the foreign exchange market.
+If speculators are uninformed and their activity can be regarded as noise, the most extreme sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, and when reversals are most likely to take place; if speculators are informed prices should follow a random walk after either of these extremes.
 
@@
 Speculation does not appear to be random, there is momentum to the rise and fall in sentiment and positions.
-However, while currency prices co-move with positions and sentiment, there is no informational content from these positions, supporting the suggestion that speculative actvity has a permanent affect on prices and that is consistent with speculation being part of the process of price discovery.
+However, while currency prices co-move with sentiment and positions, there is no informational content from these positions, meaning that knowing the positions does not imoprove forecasts of future returns, supporting the suggestion that speculative activity has a permanent affect on prices and that is consistent with speculation being part of the process of price discovery.
 Speculators do not seem to drive prices from fundamental value and their presence does not seem to hamper the incorporation of information into prices.
@@
   Speculation is an important concept in economics, particular in the area of finance.
-However, there is no clear, unambiguous theory of specuation.
-In a number of fields the term \emph{noise-trading} has been used.
+However, there is no clear, unambiguous theory of speculation.
+In a number of fields the term \emph{noise-trading} has been used for something like the activity discussed by Keynes and Fridman.
 \citet{bloomfield2009noise} suggests that there are three categories of noise-trader: uninformed speculators, market-making liquidity-providers and informed speculators.
-The first of these groups is similar to the picture of speculation provided by Keynes, the second refers to those agents or institutions paid for providing liqidity and for taking the risk of meeting an informed trader in the market (these are also called market-makets), and the final category is close to the idea of speculation that has been presented by Fridman.
+The first of these groups is similar to the picture of speculation provided by Keynes, the second refers to those agents or institutions paid for providing liquidity and for taking the risk of meeting an informed trader in the market (these are also called market-makers), and the final category is close to the idea of speculation that has been presented by Friedman.
 
-In the \emph{microstrucrture} literature, the term noise-trader is interchanged with \emph{liquidity-trader} to mean those trading with a lack of information and those trading for liquidty reasons.
-They do not usually distinguish between institutional and psychological reasons for the lack of attention to fundametal information.
-Nonetheless, liquidity tends to refer to those trading without reference to fundamental value: they must execute orders.
+In the \emph{microstructure} literature, the term noise-trader is interchanged with \emph{liquidity-trader} to mean those trading with a lack of information and those trading for liquidity reasons.
+They do not usually distinguish between psychological and institutional reasons for the lack of attention to fundamental information.
+Nonetheless, liquidity tends to refer to institutional explanations for trading without reference to fundamental value: these agents trade without reference to value.
 This may mean that they just execute customer orders at the current price or it may be that they buy and sell with reference to a benchmark index or to fulfil a hedging strategy.
 Examples of the microstructure approach are found in \citet{Kyle1985Continuous} and \citet{glosten1985bid}.
+
+The other issue is whether uninformed speculative noise-trading, of either variety, can be considered as just random noise?
+The original microstrucutre version had moise-traders arriving in a random fashion so that the market-maker could infer information from an imbalance of orders \citet{Kyle1985Continuous}.
+However, though the idea that noise is a random variable may be a construction that is useful for modelling and it may be justified in analysis of broader themes of microstructure and levels of market efficiency, this exogenous construct may be an over-simplification that is not apprpropriate when looking for more understanding of speculation itself.
 
@@
 
-The other question is whether either of these types uninformed speculative noise-trading can be considered as being just random noise?
-\citet[p.
-1]{GortonNoise}  say that noise-traders are all those trading in markets for ``non-information-based reasons''.
-This is a solution to the  \citet{Grossman1980Impossibility} paradox that will allow those traders prepared to pay the cost of acquiring information to take advantage of it.
-Informed traders, knowing that the current price is below fundamental value, can purchase from noise-traders, who either do not know or do not care.
-The noise drives prices from their fundamental value and ensures that the cost of acquiring information can be regained and therefore provides the incentive for informed traders to improve the informational efficiency of asset prices.
-
-However, though the idea that noise is a random variable may be a construction that is useful for modelling and it may be justified in analysis of broader themes of microstructure and levels of market efficiency, this exogenous construct may be an over-simplification that is not apprpropriate when looking for more understanding of speculation itself.
-Institutional and psychological explanations for noise-trading suggest that there are endogenous elements that suggest non-random effects.The modelling of noise is related to the rise of rational expectations in economics.
-A mean zero, normally-distributed random variable influences supply in the original version supplied by \citet{Muth1961Rational}.
 \subsection*{The problem of uninformed noise-traders}
-The traditional Keynesian view of speculation asserted that it will be damaging because it will make it even harder for informed-traders to overcome undertainty and apply fundamental information about long-term value.
-This is view finds some supported in Shiller's observation that stock prices are more volatile than the underlying earnings information on which valuation should be based \citet{ShillerVol}.
-In addition, \citet{Delong1990noise} and \citet{shleifer1997limits} show that constrained horizons can cause noise-trading to limit the ability of informed traders to take advantage of profitable opportunities for fear of further divergence from fundamental value before convergence is achieved.
+The traditional Keynesian view of speculation asserted that it would be an additional noise that would make it even harder for informed-traders to overcome uncertainty and apply fundamental information about long-term value.
+This view finds some supported in Shiller's observation that stock prices are more volatile than the underlying earnings information on which valuation should be based \citet{ShillerVol}.
+In addition, \citet{Delong1990noise} and \citet{shleifer1997limits} show that constrained horizons can cause noise-trading to limit the ability of informed traders to take advantage of profitable opportunities for fear of further divergence from fundamental value before realignment is achieved.
 
@@
 \subsection{Interplay between Informed and uninformed}
+\citet[p.
+1]{GortonNoise}  say that noise-traders are all those trading in markets for ``non-information-based reasons'', and that this was originally a solution to the problem of understanding or justifying the trading of securities: if economic agents know the fundamental value of a security how can some be buying and some be selling.
+While will traders be prepared to pay the cost of acquiring information if they are unable to take advantage of it.
+Informed traders, knowing that the current price is below fundamental value, can purchase from noise-traders, who either do not know or do not care.
+Noise-traders can therefore create a gap between the price and the fundamtental value that will encourage informed traders to pay the cost of acquiring information.
+The divergence between this price and fundametnal value is essential to encourage te process of price discovery.
+
 Fisher Black identifies noise as ``the arbitrary element in expectations'' \citet[p.529]{BlackNoise}.
-He cites a number of ways that noise affects the trading in financial markets.
-He speaks about people trading on noise over information; he says that noise makes financial markets possible but imperfect.
-The more noise-traders that there are, the more liqiity and the easier to trade; the more noise-traders the higher the level of inefficiency in the market and the more likely that prices will be driven from their fundamental value.
+He asserts that noise makes financial markets possible but imperfect: the more noise-traders that there are, the more liquidity and the easier it is to trade; the more noise-traders the higher the level of inefficiency in the market and the more likely that prices will be driven from their fundamental value.
 He suggest that the noise will limit the scope for complete arbitrage and that it will be difficult to distinguish between noise and information traders.
@@
 \subsection{Other empirical work}
-The assessmen of these theoreis about speculation and noise-trading is rather limited.
-In part this is becuase information is limited and it is usually difficult to distiguish between speculators and other investors and, even more so, between informed and uninformed.
+The assessment of these theories about speculation and noise-trading is rather limited.
+In part this is because information is limited and it is usually difficult to distinguish between speculators and other investors and, even more so, between informed and uninformed.
 
-Some of the empirical work will identify specuation or noise with momentum trading. 
+One approach to this problem is to use the evidence that noise-traders utilise indicators or techniques that are believed to provide information when they do not.
+\href{http://papers.ssrn.com.ezproxy.liv.ac.uk/sol3/papers.cfm?abstract_id=869827}{do noise traders move markets?}  there is strong evidence of herding by individual investors.
+there is also momentum.
+there is reversal:  stocks that were heavily bought one year under perform; stocks that were heavily sold, outperform.
+the opposite in the short run:  heavily bought one week generate large returns the next week.
+\citet{ShillerFashion} discusses fashionable techniques.
+In this way, speculation noise is associated with sentiment and measures of sentiment are then used as ways of measuring speculation.
+In addition,empirical work will often identify speculation or noise with momentum trading.
+
+This is controversial, there is some empirical evidence that individual investors in various countries trade in a contrariety fashion (e.g., choe, kho, and stulz [1999], grinblatt and keloharju [2000, 2001], jackson [2003], richards [2005], kaniel, saar, titman [2006]).
+for experimental evidence consistent with contrariety behaviour see
+bloomfield, tayler, and zhou (2997)
+
+the literature on heterogeneous traders also focuses on noise traders as trend chasers.
+these models (see for example brock and hommes [1998]; lux [1995]; lux and marchesi [2000]) investigate bubbles and crashes, and feature fundamental traders versus noise traders.
+in these models, it is the noise traders who
+induce the aberrant market behaviour.
+
 \citet{da2010sum} seek to measure investor sentiment and assess how it affects securities prices.
 they create a index of investor sentiment from web searches and use this to predict return reversals.
-the results show that sentiment can predict market turns.
-this is a contrast to traditional ways of measuring sentiment:  proxy for sentiment with market-based measures like volume, closed-end fund discount, ipo first-day returns, ipo volume, option implied volatilities and mutual fund flows.
-baker and wurgaler (2007) have a comprehensive survey.
-this is more than just investors sentiment.
-there is an issue with the measurement of an input (sentiment) with an output (market price).
-use of surveys like the university of michigan consumer sentiment survey etc.
-there is a negative contemporaneous correlation between fears and stock market returns.
+The results show that sentiment can predict market turns.
+There is a negative contemporaneous correlation between fears and stock market returns.
 however, there is a reversal in the relationship after two days.
-this is consistent with ``sentiment induced temporary mispricing'' \citet[p.
+this is consistent with ``sentiment induced temporary mis-pricing'' \citet[p.
 4]{da2010sum}.
@@
 
-the effect of noise traders on liquidity is unclear.
+The effect of noise traders on liquidity is unclear.
 they may supply liquidity by providing limit orders and making some returns for providing liquidity; they may increase bid-ask spreads by using market orders to take advantage of (potentially non-existent) informational advantages.
@@
 
-Noise traders exert positive effects on liquidity but weaken price reversals.
-a tax on trading will reduce noise traders but will also reduce the influence of fundamental traders.
-their experiment distinguishes between informed traders, noise traders and liquidity traders.
-\emph{soes bandits} are those taking advantage of the soes system to trade quickly and take advantage of small price movements.
-there is a lot of focus on the effect that taxation may have on the different categories of trader.
-there is an experiment that allows much more to be held constant.
-there are three views of noise traders that are tested here:  noise traders are skill experts taking advantage of information in the order book or price movements;  rationally providing liquidity by acting as market-maker; acting irrationally as noise traders.
-noise traders in this experiment appear to act as irrational contrariety traders.
-noise traders reduce the efficiency of the market but increase liquidity.
-in theory, traders may trade on mistaken fundamental information.
-shliefer and summmer (1990) use momentum.
-however, kaniel, saar and titman (2006) find evidence of contrariety strategies.
-this is associated with short-term speculation.
-does the activity drive the price away from fundamental value?
-stiglitz (1990) is one of the arguments in favour of a tobin tax.
-
-there is a question about whether experimental markets (like this paper) are superior to actual markets where the identification of trader types is difficult.
+There is a question about whether experimental markets (like this paper) are superior to actual markets where the identification of trader types is difficult.
 theory - delong, summers, shleifer etc,  scheinkman and xiong (2003).
@@
 
-\href{http://papers.ssrn.com.ezproxy.liv.ac.uk/sol3/papers.cfm?abstract_id=869827}{do noise traders move markets?}  there is strong evidence of herding by individual investors.
-there is also momentum.
-there is reversal:  stocks that were heavily bought one year under perform; stocks that were heavily sold, outperform.
-the opposite in the short run:  heavily bought one week generate large returns the next week.
+\subsection{Why this is important}
+  The nature of speculation and the role of noise-traders has become increasingly important since the global financial crisis (GFC).
+There are a number of contemporary questions where the role and nature of speculation is at the centre of the debate.
+Regulations or taxes that reduce the amount of speculation have generally been considered to be beneficial by reducing risk and encouraging more long-term attitudes towards investment.
+However, where these measures reduce liquidity, higher risk and an increase in other costs may provide an countervailing force.
 
-\href{http://restud.oxfordjournals.org.ezproxy.liv.ac.uk/content/60/1/1.full.pdf+html}{campbell and kyle, smart money, noise trading and stock market behaviour}.
-in response to the evidence that the volatility of stock prices is too great to be a function of discounted expected future earnings (shiller (1981),  leroy and porter (1981), this paper assesses whether stocks are influenced by exogenous, serially correlated noise.
-in this model, the noise arises from the interaction of noise traders.
-
-\subsection{why this is important}
-The nature of speculation and the role of noise-traders has become increasingly imporant since the golbal financial crisis (GFC).
-There are a number of comtemporary questions where the role and nature of speculation is at the center of the debate.
-Regulations or taxes that reduce the reduce the amount of speculation have generally been considered to be helpful in reducing risk and encouraging more long-term attitudes towards investment.
-However, where these measures reduce liquidity, higher risk and an increase in other costs may provide an countervailing force.
-For example, the so-called \emph{Volcker Rule} would prevent US inventment banks trading on their own-account.
+For example, the so-called \emph{Volcker Rule} would prevent US investment banks trading on their own-account.
 This, it has been argued, will reduce liquidity in the bond market by causing investment banks to hold lower levels of bond inventory on their books.
 This may increase the amount of \emph{price slippage} when final investor adjust their positions.
+In addition, the debate about a tax on financial transactions, \emph{the Tobin tax} \citet{TobinTax}, has been intended to reduce speculation.
+The surge in commodities prices that was seen prior to the GFC was seen to be partially the consequence of increased speculation in commodities and the creation of exchange traded funds (ETF) focusing on commodities markets.
+This argument depends on the idea that speculators will drive the price away from its fundamental value.
 
@@
 
-interesting difference between supplying and demanding liquidity.
-limit orders supply liquidity, market orders demand liquidity.
-the activity of liquidity traders will tend to change through the course of the session, moving from supplying to demanding liquidity.
-noise traders tend to act in the opposite manner.
-they tend to supply more liquidity as the session progresses.
-black (1986) noise traders put noise into prices.
-the effect of noise traders on liquidity: price impact of trades is lower when noise traders are present.
-however, this is mostly in the temporary impact on prices rather than the permanent impact.
-therefore, noise traders tend to reduce the amount of volatility in prices because market orders do not temporarily have an effect and then reverse.
-the presence of noise traders will increase pricing errors.
-however, this depends on fundamental traders have valuable information.
-
-there are a number of different ways that noise traders could behave:  activating as soes bandits (using the nasdaq small order execution system to make profits) or by using information about the order book; by providing liquidity like a market-maker as a service to liquidity-traders as in grossman miller (1988);  as irrational noise traders, using trading rules or popular models (shiller 1984, 1990) positive feedback or contrariety.
-
-footnote -  there is an empirical evidence that individual investors in various countries trade in a contrariety fashion (e.g., choe, kho, and stulz [1999], grinblatt and keloharju [2000, 2001], jackson [2003], richards [2005], kaniel, saar, titman [2006]).
-for experimental evidence consistent with contrariety behaviour see
-bloomfield, tayler, and zhou (2997)
-
- what is the source of this noise?
-what are the institutions?
-the importance of noise traders here is that they represent a zero-mean, normally distributed random viable.
-this is also the description that is provided in kyle (1985).
-liquidity traders trade due to imbalance in the timing of consumption or from portfolio considerations.
-shleifer, summers and delong have trading activity that is affected by sentiment, models or "fads".
-this is important.
-how the average may not be zero.
-there is a range of behavioural biases that may account for this (see barberis and thaler [2003]).the literature on heterogeneous traders also focuses on noise traders as trend chasers.
-these models (see for example brock and hommes [1998]; lux [1995]; lux and marchesi [2000]) investigate bubbles and crashes, and feature fundamental traders versus noise traders.
-in these models, it is the noise traders who
-induce the aberrant market behaviour.
-
-there is also the issue of speculation in commodities markets.  if this is random, the speculation does not affect the overall price, if this is ,liquidity-driven, it may affect the price level.
-
-fisher black says ```anything that changes the amount of character of noise trading will change the volatility of price." \citet[p. 533]{blacknoise}.
+There is also the issue of speculation in commodities markets.  if this is random, the speculation does not affect the overall price, if this is ,liquidity-driven, it may affect the price level.
 

### 2015-06-02 RRdes.R
Hope that this works.  Deleted some files to allow the merge

@@ -1,80 +1,0 @@
-require(eventstudies) 
-require(PerformanceAnalytics)
-require(xtable)
-# Read in the data------------------------------------------
-daz <- read.zoo("./RR.csv", header = TRUE, 
-                format = "%d-%b-%y", sep = ",")
-summary(daz)
-FX.z <- daz[,c("AUDRR", "EURGBPRR", "EURSEKRR", "GBPRR", "CHFRR", "EURCHFRR", 
-               "EURJPYRR", "EURRR","CADRR", "JPYRR")]
-head(FX.z)
-# Descriptive statistics of RR-------------------------------
-mystats <- function(x, na.omit=FALSE){
-  if (na.omit)
-    x <- x[!is.na(x)]
-  m <- mean(x)
-  md <- median(x)
-  s <- sd(x) 
-  t <- m/s
-  n <- length(x)
-  skew <- sum((x-m)^3/s^3)/n
-  ps <- ((6*n*(n-1))/((n-1)*(n+1)*(n+3)))^0.5
-  kurt <- sum((x-m)^4/s^4)/n - 3
-  pk <- ((n^2-1)/((n-3)*(n + 5)))^0.5
-  max <- max(x)
-  min <- min(x)
-  return(c(n=n, mean=m, t= t, median = md, stdev=s, skew=skew, ses = 2 * ps, 
-           kurtosis=kurt, sek  = 2 * pk, max = max, min = min))
-}
-FX.d <- apply(FX.z, 2, FUN = mystats, na.omit = TRUE)
-FX.d
-# This generates la tex code for table 
-print(xtable(FX.d))
-# Create extreme readings and prepare event files-------------------------
-currency = "EURGBP"
-extreme = "high"
-hi = as.numeric(0.95)
-lo = as.numeric(0.05)
-# select window either "whole lot" or "after".
-cum = "whole lot"
-wind = as.numeric(1) 
-# Read file (as zoo object) if not already done
-daz <- read.zoo("./RR.csv", header = TRUE, 
-                format = "%d-%b-%y", sep = ",")
-FX.z <-daz[,c(paste(currency), paste(currency, "RR", sep = ""))]
-FX.z$r <- 100*diff(log(FX.z[,1]))
-if(extreme == "high"){
-  FX.z$ex <- as.numeric(FX.z[,2] >= quantile(FX.z[,2],  
-                                             hi, na.rm = TRUE)) 
-} else if(extreme == "low") {
[…]

### 2015-06-02 Notes.Rnw

\documentclass{hitec}
\author{RobHayward}
\tite{Notes on Event study and specultion}

\begin{document}
\maketitle
\section*{Introduction}
This is a file to build notes about the event study and speculation.  The first step is to assess the literture. There are two broad strands: the Keynesian view of speculation as something that is de-stbilising and caused by the inability of investors to act on the basis of long-term assessment of fundamnetal value and the Friedman view of speculation as something that provides benefits through the provision of liquidity.  Is speculation de-stbilising or stbilising.

\end{document}

### 2015-06-02 Knitr/NewPaper.Rnw
Work and Home aligned. Literature review progressing.  Please enter the commit message for your changes. Lines starting
Changes on home pc.  Everything seems to be fine. Now completing the last of the lit rev.

@@
 \begin{abstract}
-An event study of extremes is used to assess the Keynesian and Friedman views of speculation.
+An event study is used to assess the Keynesian and Friedman views of speculation.
 Speculative extremes are ranked by intensity of sentiment and weight of activity.
 The extreme says little about the future in either case, suggesting that speculation is more than just random noise and supporting the view there is some informational content that is being passed through to the price with speculative activity.
-A unique dataset of risk-reversal skew on option prices is used to measure the intensity of speculative sentiment.
-This is supplemented with regulatory data on the weight of speculative positions.
+A unique dataset of risk-reversal skew on option prices is used to measure the intensity of speculative sentiment; the weight of positions is based on US regulatory data.
 \end{abstract}
@@
 101]{Keynes1936}.
-Here speculative activity is a problem that, at its most extreme, will cause booms, reversals and will discourage fundamental, long-term investors from participating.
+Here speculative activity is a problem that, at its most extreme, will cause booms, reversals and will discourage fundamental, long-term investment.
 It is noise that does not emerge randomly, but feeds off itself as a collective sentiment  encourages  trends and serial correlation in asset prices.
@@
 Therefore, Friedmand argued, the speculative process would favour those able to access or utilise information over the uninformed and, through a Darwinian process, informed speculation would dominate.
-This view regards speculation as an key lubricant of financial market as it provides liquidity and facilitates price discovery.
+This view regards speculation as an key lubricant of financial market as it provides liquidity and facilitates the process of price discovery.
 
 The aim of this paper is to understand more about the nature of speculation by investigating the relationship between both the intensity of speculation and the weight of speculative positions with the movement of prices in the foreign exchange market.
-If speculators are uninformed and their activity can be regarded as noise, the most extreme sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, and when reversals are most likely to take place; if speculators are informed prices should follow a random walk after either of these extremes.
+If speculators are uninformed and their activity can be regarded as noise, the most extreme sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when reversals are most likely to take place; if speculators are informed prices should follow a random walk after either of these extremes.
 
@@
 Speculation does not appear to be random, there is momentum to the rise and fall in sentiment and positions.
-However, while currency prices co-move with sentiment and positions, there is no informational content from these positions, meaning that knowing the positions does not imoprove forecasts of future returns, supporting the suggestion that speculative activity has a permanent affect on prices and that is consistent with speculation being part of the process of price discovery.
+However, while currency prices co-move with sentiment and positions, there is no informational content from either.
+As such, knowing the positions does not improve forecasts of future returns.
+This supports the suggestion that speculative activity has a permanent affect on prices and that is consistent with speculation being part of the process of price discovery.
 Speculators do not seem to drive prices from fundamental value and their presence does not seem to hamper the incorporation of information into prices.
@@
 They do not usually distinguish between psychological and institutional reasons for the lack of attention to fundamental information.
-Nonetheless, liquidity tends to refer to institutional explanations for trading without reference to fundamental value: these agents trade without reference to value.
+Nonetheless, liquidity tends to refer to institutional explanations for trading without reference to fundamental value.
 This may mean that they just execute customer orders at the current price or it may be that they buy and sell with reference to a benchmark index or to fulfil a hedging strategy.
@@
 However, though the idea that noise is a random variable may be a construction that is useful for modelling and it may be justified in analysis of broader themes of microstructure and levels of market efficiency, this exogenous construct may be an over-simplification that is not apprpropriate when looking for more understanding of speculation itself.
+Institutional and psychological explanations for noise-trading suggest that there are endogenous elements that suggest non-random effects.The modelling of noise is related to the rise of rational expectations in economics.
 
@@
 
- Shliefer, Summers, DeLong etc identify the limits to arbitrage.
+% Shliefer, Summers, DeLong etc identify the limits to arbitrage.
 The other writings are Battalio, Hatch and Jennings (1997), Harris and Schultz (1998) and Scheinkman and Xiong (2003).
@@
 
-It would be good to have some more modern ideas here.  Maybe they are just the empirical results. 
+%It would be good to have some more modern ideas here.  Maybe they are just the empirical results. 
 \subsection{Interplay between Informed and uninformed}
@@
 1]{GortonNoise}  say that noise-traders are all those trading in markets for ``non-information-based reasons'', and that this was originally a solution to the problem of understanding or justifying the trading of securities: if economic agents know the fundamental value of a security how can some be buying and some be selling.
-While will traders be prepared to pay the cost of acquiring information if they are unable to take advantage of it.
+Will traders be prepared to pay the cost of acquiring information if they are unable to take advantage of it.
 Informed traders, knowing that the current price is below fundamental value, can purchase from noise-traders, who either do not know or do not care.
@@
 
-  Why do noise traders exist?  There are liquidity and behavioural explanations. There are also a number of models that show noise traders as momentum traders.
+ % Why do too extra  exist?  There are liquidity and behavioural explanations. There are also a number of models that show noise traders as momentum traders.
 
@@
 One approach to this problem is to use the evidence that noise-traders utilise indicators or techniques that are believed to provide information when they do not.
-\href{http://papers.ssrn.com.ezproxy.liv.ac.uk/sol3/papers.cfm?abstract_id=869827}{do noise traders move markets?}  there is strong evidence of herding by individual investors.
-there is also momentum.
-there is reversal:  stocks that were heavily bought one year under perform; stocks that were heavily sold, outperform.
-the opposite in the short run:  heavily bought one week generate large returns the next week.
 \citet{ShillerFashion} discusses fashionable techniques.
@@
 In addition,empirical work will often identify speculation or noise with momentum trading.
+However, this is controversial as from \citet{Jagadeesh} to the present, there is evidence that momentum trading is profitable citet(that latest paper in citulike).
 

### 2015-06-05 Knitr/NewPaper.Rnw
Literature review is complete The methods will now be covered.

@@
 If speculators are uninformed and their activity can be regarded as noise, the most extreme sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when reversals are most likely to take place; if speculators are informed prices should follow a random walk after either of these extremes.
+    The foreign exchange market is utilised because it is the largest and most liquid market in the world.
+The latest survey taken by the Bank for International Settlements (BIS) in April 2013 estimated average daily turnover at \$5.4trn with average daily spot transactions of \$2.0trn \citet{BISFX2013}.
+In addition, the cost of trading in this markket, when measurd in terms of the size of the bid-ask spread, is extremely low.
+An analysis of the bid-ask spreads for USD-JPY, GBP-USD and EUR-USD for the period January 2001 to 2005 and finds that the average spread for each is 0.342\%, 0.305\% and 0.326\% respectively.
+This would be the cost for a round-trip (buying and selling immediately)\citep{Steely2013}.
 
@@
 However, while currency prices co-move with sentiment and positions, there is no informational content from either.
-As such, knowing the positions does not improve forecasts of future returns.
+As such, knowing the the strength of speculative opinion or the weight of speulative positions does not improve forecasts of future returns.
 This supports the suggestion that speculative activity has a permanent affect on prices and that is consistent with speculation being part of the process of price discovery.
@@
 The first of these groups is similar to the picture of speculation provided by Keynes, the second refers to those agents or institutions paid for providing liquidity and for taking the risk of meeting an informed trader in the market (these are also called market-makers), and the final category is close to the idea of speculation that has been presented by Friedman.
+  % is this the place to state that market-making will not be considered separately?  %this paragraph must go somewhere else.  I do not know where. 
+Nonetheless, the distinction between market-making speculative activity is not precise.\footnote{The implementation of the \emph{Volcker} Rule in the US which will limit the scope of US investment banks to trade on their own account has raised the question of where market-making turns into speculation.}
 
-In the \emph{microstructure} literature, the term noise-trader is interchanged with \emph{liquidity-trader} to mean those trading with a lack of information and those trading for liquidity reasons.
+In the \emph{microstructure} literature, the term noise-trader is often interchanged with \emph{liquidity-trader}.
 They do not usually distinguish between psychological and institutional reasons for the lack of attention to fundamental information.
@@
 Examples of the microstructure approach are found in \citet{Kyle1985Continuous} and \citet{glosten1985bid}.
+See \citep{Lyons2006Microstructure} and \citep{Evans2002Order} for examples of the micrstrucure approach in foreign exchange.
 
@@
 The original microstrucutre version had moise-traders arriving in a random fashion so that the market-maker could infer information from an imbalance of orders \citet{Kyle1985Continuous}.
-However, though the idea that noise is a random variable may be a construction that is useful for modelling and it may be justified in analysis of broader themes of microstructure and levels of market efficiency, this exogenous construct may be an over-simplification that is not apprpropriate when looking for more understanding of speculation itself.
-Institutional and psychological explanations for noise-trading suggest that there are endogenous elements that suggest non-random effects.The modelling of noise is related to the rise of rational expectations in economics.
-
-%this paragraph must go somewhere else.  I do not know where. 
-Nonetheless, the distinction between market-making speculative activity is not precise.\footnote{The implementation of the \emph{Volcker} Rule in the US which will limit the scope of US investment banks to trade on their own account has raised the question of where market-making turns into speculation.}
+However, though the idea that noise is a random variable may be a construction that is useful for modelling and it may be justified in analysis of broader themes of microstructure and levels of market efficiency, this exogenous construct may be an over-simplification that is inapprpropriate when looking for more understanding of speculation itself.
+Institutional and psychological explanations for noise-trading frequently suggest that there are endogenous non-random elements.
+% is this the right places for this.  If so, it must be expanded. If not, it can go somewhere near the interaction noise and informaed traders. That is line 58 currently and mark a.  
+The modelling of noise is related to the rise of rational expectations in economics.
 
@@
 In addition, \citet{Delong1990noise} and \citet{shleifer1997limits} show that constrained horizons can cause noise-trading to limit the ability of informed traders to take advantage of profitable opportunities for fear of further divergence from fundamental value before realignment is achieved.
-
-% Shliefer, Summers, DeLong etc identify the limits to arbitrage.
-The other writings are Battalio, Hatch and Jennings (1997), Harris and Schultz (1998) and Scheinkman and Xiong (2003).
-Lowenstein and Willard (2006)
-
 %It would be good to have some more modern ideas here.  Maybe they are just the empirical results. 
@@
 \citet[p.
-1]{GortonNoise}  say that noise-traders are all those trading in markets for ``non-information-based reasons'', and that this was originally a solution to the problem of understanding or justifying the trading of securities: if economic agents know the fundamental value of a security how can some be buying and some be selling.
-Will traders be prepared to pay the cost of acquiring information if they are unable to take advantage of it.
-Informed traders, knowing that the current price is below fundamental value, can purchase from noise-traders, who either do not know or do not care.
+1]{GortonNoise} categorise noise-traders as all those trading in markets for ``non-information-based reasons''.
+In this case, the construction olves the problem of financial market trading.
+If economic agents know the fundamental value of a security how can some be buying and some be selling?
+There must be different opinions.
+Why Will traders be prepared to pay the cost of acquiring information if they are unable to take advantage of that information?
+However, with noise-traders, the informed can take advantage of their information.
+This creaes a justification for acquiring information.
 Noise-traders can therefore create a gap between the price and the fundamtental value that will encourage informed traders to pay the cost of acquiring information.
@@
 
-Fisher Black identifies noise as ``the arbitrary element in expectations'' \citet[p.529]{BlackNoise}.
-He asserts that noise makes financial markets possible but imperfect: the more noise-traders that there are, the more liquidity and the easier it is to trade; the more noise-traders the higher the level of inefficiency in the market and the more likely that prices will be driven from their fundamental value.
-He suggest that the noise will limit the scope for complete arbitrage and that it will be difficult to distinguish between noise and information traders.
-
- % Why do too extra  exist?  There are liquidity and behavioural explanations. There are also a number of models that show noise traders as momentum traders.
+Fisher Black identified noise as ``the arbitrary element in expectations'' \citet[p.529]{BlackNoise}.
+Therefore, noise makes financial markets possible but imperfect: the more noise-traders that there are, the more liquidity and the easier it is to trade; the more noise-traders the higher the level of inefficiency in the market and the more likely that price will be driven from fundamental value.
 
 \subsection{Other empirical work}
-The assessment of these theories about speculation and noise-trading is rather limited.
+There is little empirical eassessment of these theories about speculation and noise-trading.
 In part this is because information is limited and it is usually difficult to distinguish between speculators and other investors and, even more so, between informed and uninformed.
 
-One approach to this problem is to use the evidence that noise-traders utilise indicators or techniques that are believed to provide information when they do not.
-\citet{ShillerFashion} discusses fashionable techniques.
+One approach has been to assume that noise-traders utilise simple investment strategies that involve momentum or reversals.
+For example, \citet{ShillerFashion} discusses the use of fashionable techniques or fads in investment.
 In this way, speculation noise is associated with sentiment and measures of sentiment are then used as ways of measuring speculation.
-In addition,empirical work will often identify speculation or noise with momentum trading.
-However, this is controversial as from \citet{Jagadeesh} to the present, there is evidence that momentum trading is profitable citet(that latest paper in citulike).
-
-This is controversial, there is some empirical evidence that individual investors in various countries trade in a contrariety fashion (e.g., choe, kho, and stulz [1999], grinblatt and keloharju [2000, 2001], jackson [2003], richards [2005], kaniel, saar, titman [2006]).
-for experimental evidence consistent with contrariety behaviour see
-bloomfield, tayler, and zhou (2997)
-
-the literature on heterogeneous traders also focuses on noise traders as trend chasers.
-these models (see for example brock and hommes [1998]; lux [1995]; lux and marchesi [2000]) investigate bubbles and crashes, and feature fundamental traders versus noise traders.
-in these models, it is the noise traders who
-induce the aberrant market behaviour.
-
-\citet{da2010sum} seek to measure investor sentiment and assess how it affects securities prices.
-they create a index of investor sentiment from web searches and use this to predict return reversals.
-The results show that sentiment can predict market turns.
-There is a negative contemporaneous correlation between fears and stock market returns.
-however, there is a reversal in the relationship after two days.
-this is consistent with ``sentiment induced temporary mis-pricing'' \citet[p.
-4]{da2010sum}.
-there is a positive correlation between the fears index and bond prices.
-however, this also reverses in following days.
-there is a relationship between the fears index and daily market volatility (realised volatility and the vix index).
-there is also evidence of inflows and outflows from mutual funds.
-this may help to explain the contrast between evidence from the equity market and the bond market.
-
-The effect of noise traders on liquidity is unclear.
-they may supply liquidity by providing limit orders and making some returns for providing liquidity; they may increase bid-ask spreads by using market orders to take advantage of (potentially non-existent) informational advantages.
-noise traders may increase (if they have some privileged information, they add to the price pressure) or reduce efficiency (if they increase liquidity and reduce the ratio of informed to uninformed traders).
-are day traders profitable?
-are these the face of noise traders?
-how does this show up in the foreign exchange market?
-There is a very recent paper.
-\href{http://www.sciencedirect.com/science/article/pii/S1042443115000682}{Paper on profitability of fx trading}.
-
-There is a question about whether experimental markets (like this paper) are superior to actual markets where the identification of trader types is difficult.
-theory - delong, summers, shleifer etc,  scheinkman and xiong (2003).
-lowenstein and willard (2006).
-empirical work:  (see, for example, garvey and murphy [2001], linnainmaa [2003], barber et.
-al.
-[2004], and kaniel, saar, and titman [2006].
+Alternatively, speculation can be identified with momentum trading.
+The literature on heterogeneous traders, another solution to the justificcation for trading securities, has very frequently used noise-traders as trend chasers or followers of technical analysis.
+These models (see for example brock and hommes [1998]; lux [1995]; lux and marchesi [2000]) investigate bubbles and crashes, and feature fundamental traders versus noise traders.
 
@@
 
-For example, the so-called \emph{Volcker Rule} would prevent US investment banks trading on their own-account.
+For example, the debate about a tax on financial transactions, \emph{the Tobin tax} \citet{TobinTax}, has been intended to reduce speculation.
+The surge in commodities prices that was seen prior to the GFC was seen to be partially the consequence of increased speculation in commodities and the creation of exchange traded funds (ETF) focusing on commodities markets.
+This argument depends on the idea that speculators will drive the price away from its fundamental value.
+In addition, so-called \emph{Volcker Rule} would prevent US investment banks trading on their own-account.
 This, it has been argued, will reduce liquidity in the bond market by causing investment banks to hold lower levels of bond inventory on their books.
 This may increase the amount of \emph{price slippage} when final investor adjust their positions.
-In addition, the debate about a tax on financial transactions, \emph{the Tobin tax} \citet{TobinTax}, has been intended to reduce speculation.
-The surge in commodities prices that was seen prior to the GFC was seen to be partially the consequence of increased speculation in commodities and the creation of exchange traded funds (ETF) focusing on commodities markets.
-This argument depends on the idea that speculators will drive the price away from its fundamental value.
+In addition,
 
-for related papers on securities transaction taxes see stiglitz [1989], summers and summers [1989], 
-amihud and mendelson [1993], schwert and seguin [1993], umlauf [1993], subrahmanyam [1998], dow 
-and rahi [2000], habermeir and kirilenko [2003], and hinman [2003].
-
-There is also the issue of speculation in commodities markets.  if this is random, the speculation does not affect the overall price, if this is ,liquidity-driven, it may affect the price level.
-
-% does this now need something on the reason why the foreign exchange market is used?  could be the keynes speuclative activity, fridman paper, liquidity etc. 
 \bibliography{../../../../myrefs}

### 2015-06-05 Knitr/NewPaper.Rnw
Updated the first part of the methods. This needs to be looked at and corrected.

@@
 This would be the cost for a round-trip (buying and selling immediately)\citep{Steely2013}.
+The idea of noise trading may be particularly important for the foreign exchange market where it is widely perceived that there are a relatively large number of traders engaged in technical analysis.
+See \citep{FrootChartists} for a survey.
 
@@
 
+\section{Methodology}
+
+The aim of this study is to test whether the noise-trader model provides an appropriate explanation for speculators' activity in the  foreign exchange market .
+This will involve three steps:  first,assess the price implication of the noise-trader modele when speculative sentiment or activity becomes extreme; second, measure speculative activity in the foreign exchange market; finally,  whether extreme speculative activity in the foreign exchange market is consistent with the noise-trader model.
+
+The noise-trader model used in this study is from De Long, Shleifer, Summers and Waldeman \citep{Delong1990noise}.
+This is a derivative of the Samuelson overlapping generations model  with two-period agents making a portfolio choice \citep{Samuelson1958}.
+There are two assets:  a safe asset $(s)$ which pays a set dividend $(d)$ and is in perfectly elastic supply; the risky asset $(u)$ also pays the same dividend but is not in elastic supply.
+There are two types of investor or trader: sophisticated traders $(i)$ who know the fundamental value of the risky assets and noise-traders $(n)$ who misperceive the fundamental value.
+In the absence of noise-traders, the assets are substitutes.
+Each trader chooses their portfolio in the first period $(t)$ to maximise their utility on the basis of their belief about the price of the unsafe asset in the second period $(t + 1)$.
+The representative noise-trader mistakes the price of the risky asset by a random variable
+
+\begin{equation}
+\rho_t \sim N(\rho^*, \sigma_{\rho^2})
+\end{equation}
+
+where $\rho^*$ is the average mistake of the noise-traders and $\sigma_{\rho}^2$ is the variance of the misperception.
+Assuming an exponential utility function or constant absolute risk aversion\footnote{Though \citep{TobinLiquidity} used a quadratic utility function to show that investor preference could be defined by just the mean and the variance of the returns, the increasing absolute risk aversion this implies does not seem to be a suitable property as it implies people become more unwilling to take risk as they become more wealthy (\citep{Hicks1962} and \citep{ArrowRisk}.
+Constant absolute risk aversion is the \citep{PrattCARA}, \citep{ArrowRisk}  measure based on $u(c) = 1 - e^{-\gamma c}$ where the \emph{coefficient of absolute risk aversion} $A(c) = -\frac{u''(c)}{u'(c)} = \gamma$.
+This is a standard device.}, each type of investor chooses the proportion $(\lambda)$ of the risky asset to hold to maximise utility.
+
+\begin{equation}
+E(U) = \bar{w} - \gamma \sigma_{w}^2
+\end{equation}
+
+where $\bar{w}$ is end of period wealth, $\gamma$ is a measure of absolute risk aversion and $\sigma_w^2$ is the variance of wealth.
+Utility is a function of risk and return.
+This depends on holdings of the risky asset.
+For the informed trader this is,
+
+\begin{equation}
+E(U) = c_0 + \lambda_t^i[r + _tp_{t+1} - p_t(1 + r)] - \gamma (\lambda_t^i)^2 \left (_t\sigma^2_{p, t+1} \right ) 
+\end{equation}
+
+where $c_0$ is the endowment, following the notation in the original paper, $_tp_{t+1}$ is the expected price of the risky asset at $t + 1$, conditional on information at time $t$ and $_t\sigma^2_{\rho, t+1}$ is the expected variance of the misperception conditional on information at time $t$.
+
+Maximising utility with respect to holdings of the risky asset and iterating backward to solve for the current value,  the price becomes the discounted value of the future dividends, the expected capital gain and the general misperception of noise-traders less risk.
+
+\begin{equation}
+p_t = 1 + \frac{\mu(\rho_t - \rho^*)}{1 + r} + \frac{ \mu \rho^*}{r} - \frac{2 \gamma}{r}\ \big (_t\sigma^2_{p,t+1} \big )
+\label{eq:NP1}
+\end{equation}
+
+In words, the deviation from the true fundamental value of unity is a result of the discounted value of the noise-traders' misperception of the price $(p_t - \rho^*)$, multiplied by the weight of noise-traders in the total $(\mu)$ plus a perpetuity of the average noise-trader misperception $(\frac{ \mu \rho^*}{r} )$, adjusted for the risk.
+This risk is a function of the volatility of the price which,  from equation \ref{eq:NP1} is a result of the variability of the misperception of speculative noise-traders.
+All but the second term in equation \ref{eq:NP1} are constants, and if the variance of the price one period ahead  is equal to the variance of the misperception,
+
+\begin{equation} 
+ _t\sigma^2_{p,t+1} = \frac{\mu^2 \sigma_{\rho}^2}{1+r^2}  
+\end{equation}
+
+the final price equation becomes.
+
+\begin{equation}
+p_t = 1 + \frac{\mu(\rho_t - \rho^*)}{1 + r} + \frac{ \mu \rho^*}{r} - \frac{2 \gamma \mu^2 \sigma_{\rho}^2}{r(1 + r)^2}
+\label{eq:NP2}
+\end{equation}
+
+The paper makes use of the utility function to compare the outcomes for the two types of investor for different realisations of the parameters.
+They note four effects:  the \emph{Friedman Effect}, which is the tendency of informed investors to push prices back towards fundamental value; \emph{Price Pressure} as the way that speculative noise-traders buy more and drive the price higher as they become more bullish; \emph{Hold More},whereby the noise-traders hold more of the risky asset; \emph{Create Space} is the added risk that is caused by an increase in the variability of noise-trader belief.
+The Friedman Effect and Price Pressure tend to reduce noise-trader utility as they work to create losses or reduced gains for noise-trader positions; Hold More and Create Space can increase noise-trader utility and profits \citep[pp.
+14-15]{Delong1990noise}.
+%Put something on these four in the conclusion. 
+A number of important ideas flow from the model.
+The main findings are that noise-traders will not necessarily disappear as Friedman had suggested \citep{FriedmanPositive}.
+While noise-traders will tend to get less utility than sophisticated traders, this does not necessarily drive them from the market.
+The noise-traders expected to get higher utility due to their misperception.
+The reduction in utility comes mainly from the fact that they are not compensated for taking risk that they are not aware of.
+This is consistent with the type of over-confident, entrepreneurial misperception discussed in Section \ref{sec:prob} in relation to the returns to speculation.
+If the noise-traders become aware of the risk that they are taking they are less likely to undertake the activity.
+
+De Long and the others test a variety of specifications that include modest switching from one style of investment to another and changes in the weight of new traders after previous performance.
+Some result in the disappearance of noise-traders while others suggest that they can become more dominant.
+The model also explains the importance of investment with a long-term horizon as this is the only way that \emph{noise- trader risk} can be overcome.
+If agents live for more than two periods, fundamental investors can make an arbitrage between safe and risky assets, driving the price back to fundamental value.
+Therefore, investors or traders with a reputation that will allow resources and time to consider the long-term would be able to take full advantage of the noise-traders' misperception.
+The model relies crucially on the institutional or psychological myopia.
+% read again and add more.
+
+Noise-traders drive the price in the direction of their mis-perception.
+The second and third terms in Equation \ref{eq:NP2} show the deviation from fundamental value due to the fact that the noise-trader mis-perception is not equal to zero but equal to $p^*$.
+This means that the mis-perception is not random but biased in a positive fashion.
+If the mis-perception is random, we are back to the random walk with noise-traders adding liquidity on each side of the market and the informed traders able to take advantage of any deviation of the price from its fundamental value.
+There is still some risk that the outcome will be different from the fundamental value, because the noise-traders may still become more extreme in their misperception, but on average informed traders will make profit if they act.
+This deviation in the price is what De Long and co-authors call price pressure.
+While there is no fundamental risk in this model, price pressure causes risk for which there is no compensation and which is assumed by the noise-traders.
+This is rather similar to the risk created by speculators in the Keynes models where additional uncertainty, in the Knightian sense, encourages more investors towards self-perpetuating myopic behaviour.
+
+%One of the key assumptions of the model is the two period construction that means that final value is not determined.
+This means that informed investors cannot just wait for the fundamental value to assert itself.
+Therefore, the model is consistent with the sort of myopia that may be caused by impatience \citep{Keynes1936}, funding costs, social pressure and institutional features.
+% references needed.
+%\footnote{An good practical example is that of Tony Dye the Philips and Drew Fund Manager who refused to participate in the technology boom of the late 1990s.
+The relative under-performance of the fund cause a loss of investors and his eventual resignation in March 2000.
+This is the same date as the peak of the US technology index NASDAQ.
+It is also consistent with the words of Citicorp Chief Executive Charles Prince who is reported to have said, "As long as the music's playing, you've got to get up and dance" when asked by the \emph{Financial Times} in July 2007 about the bank's lending activity.}
+
+%is this paragraph necessary?
+%The De Long and others do not go into the details of the misperception that is suffered by the noise traders.
+There are a number of explanation for this deviation of the misperception from zero.
+The first are the behavioural biases that were addressed in section \ref{sec:behavioural}.
+If investors generally are suffering from the \emph{representativeness} heuristic an over-focus on the immediate information that the expense of the bigger picture or the base risk to securities value would cause a systematic deviation in favour or optimism or pessimism.
+This effect can be exacerbated by the overconfidence that has been identified generally in the behavioural economic literature and by Odean specifically for investment markets.
+Systematic bias creates price pressure which feeds back into a successful outcome that is regarded as being the consequence of skill, ability and good judgment and investment talent rather than luck and temporary circumstance.
+%more may be required here Hirshliefer and Luo (2001) on the survival of noise traders.
+%There is also the social pressure and a feedback mechanism that comes from taking an action that others are taking.
+Therefore, if others have the same investment strategy, this can make investors more confident and, in some cases, will cause investors to infer private information from the actions of other investors who are basing investment decisions on the same style of imitation.
+All the investors think that they are in good company and that the knowledge of the others fills the gaps where their own uncertainty begins.
+% need references for this.
+
+\
 \bibliography{../../../../myrefs}

### 2015-06-07 Knitr/NewPaper.Rnw
NewPaper.Rnw has been taken from the previous commit. It did not seem to have been updated. NOw the file has the methods (though these need to be looked at) and it hs the changes to reason for fx etc.

@@
 This would be the cost for a round-trip (buying and selling immediately)\citep{Steely2013}.
-The idea of noise trading may be particularly important for the foreign exchange market where it is widely perceived that there are a relatively large number of traders engaged in technical analysis.
-See \citep{FrootChartists} for a survey.
 
@@
 The divergence between this price and fundametnal value is essential to encourage te process of price discovery.
-
+% What where the reasons for creating the concept of noise-traders?  There are three:  the rtionale for trading, the motivtion or mraket-marking and... 
 Fisher Black identified noise as ``the arbitrary element in expectations'' \citet[p.529]{BlackNoise}.
-Therefore, noise makes financial markets possible but imperfect: the more noise-traders that there are, the more liquidity and the easier it is to trade; the more noise-traders the higher the level of inefficiency in the market and the more likely that price will be driven from fundamental value.
+Therefore, noise makes financial markets possible but imperfect: the more noise-traders that there are, the more liquidity and the easier it is to trade; the more noise-traders, the higher the level of inefficiency in the market and the more likely that price will be driven from fundamental value.
 
@@
 Alternatively, speculation can be identified with momentum trading.
-The literature on heterogeneous traders, another solution to the justificcation for trading securities, has very frequently used noise-traders as trend chasers or followers of technical analysis.
+The literature on heterogeneous traders, another way that the trading securities is explained, has very frequently used noise-traders as trend-chasers or followers of technical analysis.
 These models (see for example brock and hommes [1998]; lux [1995]; lux and marchesi [2000]) investigate bubbles and crashes, and feature fundamental traders versus noise traders.
@@
 This argument depends on the idea that speculators will drive the price away from its fundamental value.
+
 In addition, so-called \emph{Volcker Rule} would prevent US investment banks trading on their own-account.
@@
 \section{Methodology}
-
-The aim of this study is to test whether the noise-trader model provides an appropriate explanation for speculators' activity in the  foreign exchange market .
-This will involve three steps:  first,assess the price implication of the noise-trader modele when speculative sentiment or activity becomes extreme; second, measure speculative activity in the foreign exchange market; finally,  whether extreme speculative activity in the foreign exchange market is consistent with the noise-trader model.
+The aim of this study is to test whether the noise-trader model provides an appropriate explanation for speculator activity in the  foreign exchange market.
+This will be use to gain a fuller understanding of speculative behaviour.
+This will involve three steps:  first,assess the price implication of the noise-trader model;  second, measure speculative activity in the foreign exchange market; finally, assess whether extreme speculative activity in the foreign exchange market is consistent with the noise-trader model.
+If the speculative activity that is seen in the foreign exchange market produces results that are like the noise-trader model, it can be said that speculation is noise trading as Keynes asserted, if not it may be more informed as suggested by Friedman.
+% where it is widely perceived that there are a relatively large number of traders engaged in technical analysis.
+See \citep{FrootChartists} for a survey.
+%this may go in the explanation for why fx is used.
+Use this elsewhere.
+Add more?
 
@@
 \end{equation}
-
-where $\rho^*$ is the average mistake of the noise-traders and $\sigma_{\rho}^2$ is the variance of the misperception.
-Assuming an exponential utility function or constant absolute risk aversion\footnote{Though \citep{TobinLiquidity} used a quadratic utility function to show that investor preference could be defined by just the mean and the variance of the returns, the increasing absolute risk aversion this implies does not seem to be a suitable property as it implies people become more unwilling to take risk as they become more wealthy (\citep{Hicks1962} and \citep{ArrowRisk}.
-Constant absolute risk aversion is the \citep{PrattCARA}, \citep{ArrowRisk}  measure based on $u(c) = 1 - e^{-\gamma c}$ where the \emph{coefficient of absolute risk aversion} $A(c) = -\frac{u''(c)}{u'(c)} = \gamma$.
-This is a standard device.}, each type of investor chooses the proportion $(\lambda)$ of the risky asset to hold to maximise utility.
-
-\begin{equation}
-E(U) = \bar{w} - \gamma \sigma_{w}^2
-\end{equation}
-
-where $\bar{w}$ is end of period wealth, $\gamma$ is a measure of absolute risk aversion and $\sigma_w^2$ is the variance of wealth.
-Utility is a function of risk and return.
-This depends on holdings of the risky asset.
-For the informed trader this is,
-
-\begin{equation}
-E(U) = c_0 + \lambda_t^i[r + _tp_{t+1} - p_t(1 + r)] - \gamma (\lambda_t^i)^2 \left (_t\sigma^2_{p, t+1} \right ) 
-\end{equation}
-
-where $c_0$ is the endowment, following the notation in the original paper, $_tp_{t+1}$ is the expected price of the risky asset at $t + 1$, conditional on information at time $t$ and $_t\sigma^2_{\rho, t+1}$ is the expected variance of the misperception conditional on information at time $t$.
 
@@
 The reduction in utility comes mainly from the fact that they are not compensated for taking risk that they are not aware of.
-This is consistent with the type of over-confident, entrepreneurial misperception discussed in Section \ref{sec:prob} in relation to the returns to speculation.
-If the noise-traders become aware of the risk that they are taking they are less likely to undertake the activity.
+This is consistent with the type of over-confidence that has been found in day-traders \citet{odean1998volume}
 
@@
 
-%One of the key assumptions of the model is the two period construction that means that final value is not determined.
-This means that informed investors cannot just wait for the fundamental value to assert itself.
-Therefore, the model is consistent with the sort of myopia that may be caused by impatience \citep{Keynes1936}, funding costs, social pressure and institutional features.
-% references needed.
-%\footnote{An good practical example is that of Tony Dye the Philips and Drew Fund Manager who refused to participate in the technology boom of the late 1990s.
-The relative under-performance of the fund cause a loss of investors and his eventual resignation in March 2000.
-This is the same date as the peak of the US technology index NASDAQ.
-It is also consistent with the words of Citicorp Chief Executive Charles Prince who is reported to have said, "As long as the music's playing, you've got to get up and dance" when asked by the \emph{Financial Times} in July 2007 about the bank's lending activity.}
-
-%is this paragraph necessary?
-%The De Long and others do not go into the details of the misperception that is suffered by the noise traders.
-There are a number of explanation for this deviation of the misperception from zero.
-The first are the behavioural biases that were addressed in section \ref{sec:behavioural}.
-If investors generally are suffering from the \emph{representativeness} heuristic an over-focus on the immediate information that the expense of the bigger picture or the base risk to securities value would cause a systematic deviation in favour or optimism or pessimism.
-This effect can be exacerbated by the overconfidence that has been identified generally in the behavioural economic literature and by Odean specifically for investment markets.
-Systematic bias creates price pressure which feeds back into a successful outcome that is regarded as being the consequence of skill, ability and good judgment and investment talent rather than luck and temporary circumstance.
-%more may be required here Hirshliefer and Luo (2001) on the survival of noise traders.
-%There is also the social pressure and a feedback mechanism that comes from taking an action that others are taking.
-Therefore, if others have the same investment strategy, this can make investors more confident and, in some cases, will cause investors to infer private information from the actions of other investors who are basing investment decisions on the same style of imitation.
-All the investors think that they are in good company and that the knowledge of the others fills the gaps where their own uncertainty begins.
-% need references for this.
-
-\
 \bibliography{../../../../myrefs}

### 2015-06-10 Knitr/NewPaper.Rnw
Updated the whole NewPaper Now it needs to be read through and corrected.

@@
 \begin{abstract}
-An event study is used to assess the Keynesian and Friedman views of speculation.
-Speculative extremes are ranked by intensity of sentiment and weight of activity.
-The extreme says little about the future in either case, suggesting that speculation is more than just random noise and supporting the view there is some informational content that is being passed through to the price with speculative activity.
+An event study is used to assess the views of Keynes and Friedman on speculation.
+Speculative extremes are ranked by intensity of sentiment and weight of speculative activity.
 A unique dataset of risk-reversal skew on option prices is used to measure the intensity of speculative sentiment; the weight of positions is based on US regulatory data.
+The extremes say little about the future in either case, suggesting that speculation is more than just random noise and supporting the view there is some informational content that is being passed through to the price with speculative activity.
 \end{abstract}
@@
 Here speculative activity is a problem that, at its most extreme, will cause booms, reversals and will discourage fundamental, long-term investment.
-It is noise that does not emerge randomly, but feeds off itself as a collective sentiment  encourages  trends and serial correlation in asset prices.
+It is noise that does not emerge randomly, but feeds off itself as the evolution of collective sentiment encourages trends in asset prices.
 
@@
 The latest survey taken by the Bank for International Settlements (BIS) in April 2013 estimated average daily turnover at \$5.4trn with average daily spot transactions of \$2.0trn \citet{BISFX2013}.
-In addition, the cost of trading in this markket, when measurd in terms of the size of the bid-ask spread, is extremely low.
+The cost of trading in this markket, when measurd in terms of the size of the bid-ask spread, is extremely low.
 An analysis of the bid-ask spreads for USD-JPY, GBP-USD and EUR-USD for the period January 2001 to 2005 and finds that the average spread for each is 0.342\%, 0.305\% and 0.326\% respectively.
@@
 However, there is no clear, unambiguous theory of speculation.
-In a number of fields the term \emph{noise-trading} has been used for something like the activity discussed by Keynes and Fridman.
-\citet{bloomfield2009noise} suggests that there are three categories of noise-trader: uninformed speculators, market-making liquidity-providers and informed speculators.
-The first of these groups is similar to the picture of speculation provided by Keynes, the second refers to those agents or institutions paid for providing liquidity and for taking the risk of meeting an informed trader in the market (these are also called market-makers), and the final category is close to the idea of speculation that has been presented by Friedman.
-  % is this the place to state that market-making will not be considered separately?  %this paragraph must go somewhere else.  I do not know where. 
-Nonetheless, the distinction between market-making speculative activity is not precise.\footnote{The implementation of the \emph{Volcker} Rule in the US which will limit the scope of US investment banks to trade on their own account has raised the question of where market-making turns into speculation.}
+Speculation plays a key role in three importantant and related areas: market-microstructure, the rationale for the trading of financial securities and rational expectations.
 
+One key element of the microstructure literature involves distinguishing between uninformed or noise-traders and those who have infomration.
+The Kyle model has market-makers reciving orders from the informed and the uninformed.
+On the assumption that the uniformed arrive in a random fashion without any reason to trade, the weight of orders implies something about the activity of the informed and causes prices to change in the process of price discovery.
+The bid-ask spread is the payment that the market-maker recieves in return for taking the risk of meeting an informed trader.
+Ultimately, the uninformed traders make payments to the informed traders with the mark-makers taking a cut for the risk and the privision of exchange \citet{Kyle1985Continuous} and \citet{glosten1985bid}.
+Here the speculators may be informed or uninformed.
+
+% does this following paragraph come later?  I think so....Once the three
+% main categories are done. 
 In the \emph{microstructure} literature, the term noise-trader is often interchanged with \emph{liquidity-trader}.
@@
 This may mean that they just execute customer orders at the current price or it may be that they buy and sell with reference to a benchmark index or to fulfil a hedging strategy.
-Examples of the microstructure approach are found in \citet{Kyle1985Continuous} and \citet{glosten1985bid}.
-See \citep{Lyons2006Microstructure} and \citep{Evans2002Order} for examples of the micrstrucure approach in foreign exchange.
 
-The other issue is whether uninformed speculative noise-trading, of either variety, can be considered as just random noise?
+\citet[p.
+1]{GortonNoise} categorise noise-traders as all those trading in markets for ``non-information-based reasons''.
+In this case, uninforrmed speculation provides a rational and an opportunity for the trading of financial securities.
+If all economic agents know the fundamental value of a security how can some be buying and some be selling?
+There must be different opinions.
+Why Will traders be prepared to pay the cost of acquiring information if they are unable to take advantage of that information?
+With speculative noise-traders, the informed can take advantage of their information.
+This creates a justification for acquiring information.
+Speulative noise-traders can therefore create a gap between the price and the fundamtental value that will encourage informed traders to pay the cost of acquiring information.
+The divergence between this price and fundametnal value is essential to encourage te process of price discovery.
+%
+Fisher Black suggested that this combination was essential to the smooth function of financial markets.
+It is ``the arbitrary element in expectations'' \citet[p.529]{BlackNoise}.
+Therefore, noise makes financial markets possible but imperfect: the more noise-traders that there are, the more liquidity and the easier it is to trade; the more noise-traders, the higher the level of inefficiency in the market and the more likely that price will be driven from fundamental value.
+
+  Speculation is also an important part of the argument for why prices should tend towards that consistent with \emph{rational expectations}.
+\citet{muth1961} uses commodity speculation to argue for model-consistent expectation formation.
+In this case, the speculators are informed and tend to reduce the volatility of prices.
+
+% See \citep{Lyons2006Microstructure} and \citep{Evans2002Order} for examples of the micrstrucure approach in foreign exchange.
+
+If there are a mixture of informed and uninformed specualtors, the question becomes whether the uninformed speculative activity can be considered as just random noise?
 The original microstrucutre version had moise-traders arriving in a random fashion so that the market-maker could infer information from an imbalance of orders \citet{Kyle1985Continuous}.
@@
 Institutional and psychological explanations for noise-trading frequently suggest that there are endogenous non-random elements.
-% is this the right places for this.  If so, it must be expanded. If not, it can go somewhere near the interaction noise and informaed traders. That is line 58 currently and mark a.  
-The modelling of noise is related to the rise of rational expectations in economics.
 
@@
 %It would be good to have some more modern ideas here.  Maybe they are just the empirical results. 
-\subsection{Interplay between Informed and uninformed}
-\citet[p.
-1]{GortonNoise} categorise noise-traders as all those trading in markets for ``non-information-based reasons''.
-In this case, the construction olves the problem of financial market trading.
-If economic agents know the fundamental value of a security how can some be buying and some be selling?
-There must be different opinions.
-Why Will traders be prepared to pay the cost of acquiring information if they are unable to take advantage of that information?
-However, with noise-traders, the informed can take advantage of their information.
-This creaes a justification for acquiring information.
-Noise-traders can therefore create a gap between the price and the fundamtental value that will encourage informed traders to pay the cost of acquiring information.
-The divergence between this price and fundametnal value is essential to encourage te process of price discovery.
-% What where the reasons for creating the concept of noise-traders?  There are three:  the rtionale for trading, the motivtion or mraket-marking and... 
-Fisher Black identified noise as ``the arbitrary element in expectations'' \citet[p.529]{BlackNoise}.
-Therefore, noise makes financial markets possible but imperfect: the more noise-traders that there are, the more liquidity and the easier it is to trade; the more noise-traders, the higher the level of inefficiency in the market and the more likely that price will be driven from fundamental value.
-
+%this paragraph is not very good.  It needs to be updated.  Look at the file in citulike. 
 \subsection{Other empirical work}
@@
 One approach has been to assume that noise-traders utilise simple investment strategies that involve momentum or reversals.
-For example, \citet{ShillerFashion} discusses the use of fashionable techniques or fads in investment.
+For example, \citet{  shillerFashion} discusses the use of fashionable techniques or fads in investment.
 In this way, speculation noise is associated with sentiment and measures of sentiment are then used as ways of measuring speculation.
@@
 
-The paper makes use of the utility function to compare the outcomes for the two types of investor for different realisations of the parameters.
-They note four effects:  the \emph{Friedman Effect}, which is the tendency of informed investors to push prices back towards fundamental value; \emph{Price Pressure} as the way that speculative noise-traders buy more and drive the price higher as they become more bullish; \emph{Hold More},whereby the noise-traders hold more of the risky asset; \emph{Create Space} is the added risk that is caused by an increase in the variability of noise-trader belief.
-The Friedman Effect and Price Pressure tend to reduce noise-trader utility as they work to create losses or reduced gains for noise-trader positions; Hold More and Create Space can increase noise-trader utility and profits \citep[pp.
+The following effects are noted:  the \emph{Friedman Effect}, which is the tendency of informed investors to push prices back towards fundamental value; \emph{Price Pressure} as the way that speculative noise-traders buy more and drive the price higher as they become more bullish; \emph{Hold More},whereby the noise-traders hold more of the risky asset; \emph{Create Space} is the added risk that is caused by an increase in the variability of noise-trader belief \citep[pp.
 14-15]{Delong1990noise}.
 %Put something on these four in the conclusion. 
-A number of important ideas flow from the model.
 The main findings are that noise-traders will not necessarily disappear as Friedman had suggested \citep{FriedmanPositive}.
@@
 
-De Long and the others test a variety of specifications that include modest switching from one style of investment to another and changes in the weight of new traders after previous performance.
-Some result in the disappearance of noise-traders while others suggest that they can become more dominant.
-The model also explains the importance of investment with a long-term horizon as this is the only way that \emph{noise- trader risk} can be overcome.
-If agents live for more than two periods, fundamental investors can make an arbitrage between safe and risky assets, driving the price back to fundamental value.
-Therefore, investors or traders with a reputation that will allow resources and time to consider the long-term would be able to take full advantage of the noise-traders' misperception.
-The model relies crucially on the institutional or psychological myopia.
-% read again and add more.
+\subsection{The implications of the noise-trader model}
+The aim of this study is to test the implications of the noise-trader model.
+There are two parameters from the noise-trader price Equation \ref{eq:NP2} that can be used to measure the amount of speculation:  the price misperception $(\rho)$ and the weight of noise-traders in the market $(\mu)$.
+The aim of this section is to  identify the returns that are implied by the noise-trader model when either the intensity of the misperception or the weight of noise-traders is increased to extreme levels so that this can be used as a benchmark against which foreign exchange market activity can be judged.
 
-Noise-traders drive the price in the direction of their mis-perception.
-The second and third terms in Equation \ref{eq:NP2} show the deviation from fundamental value due to the fact that the noise-trader mis-perception is not equal to zero but equal to $p^*$.
-This means that the mis-perception is not random but biased in a positive fashion.
-If the mis-perception is random, we are back to the random walk with noise-traders adding liquidity on each side of the market and the informed traders able to take advantage of any deviation of the price from its fundamental value.
-There is still some risk that the outcome will be different from the fundamental value, because the noise-traders may still become more extreme in their misperception, but on average informed traders will make profit if they act.
-This deviation in the price is what De Long and co-authors call price pressure.
-While there is no fundamental risk in this model, price pressure causes risk for which there is no compensation and which is assumed by the noise-traders.
-This is rather similar to the risk created by speculators in the Keynes models where additional uncertainty, in the Knightian sense, encourages more investors towards self-perpetuating myopic behaviour.
+To understand what the noise-trader model of speculative behaviour means for foreign exchange rates, a \emph{Monte Carlo} simulation is carried out with a variety of mis-perceptions and noise-trader weights.
+Two time series are created.
+The first series is based on a random noise-trade misperception, $(\rho^*)$.
+This is called the Misperception Model (MPM).
+The second is based on a random weight of noise-traders $(\mu)$ and is called the Random Weight Model (RWM).
+The simulation creates a 1000 period price series from Equation \ref{eq:NP2}.
+The dividend paid on the risky asset $(r)$ is posted at 10\%; the risk-aversion parameter $(\gamma)$ is assumed to be 0.3 (in line with that suggested in \citep[p.
+127]{FrobesBehaviour}); starting from a price of 100, the average noise-trader mis-perception $(p^*)$ is a random variable in the first simulation with a mean of 4 and a variance $(\sigma_{\rho^*}^2)$ of 2, and has a fixed value of 4 in the second simulation; the weight of noise-traders $(\mu)$ is 50\% in the first case but is modelled as a two-period moving average of a normal random variable centred at 50\% with a standard deviation of 20\% and equally weighted average of its previous value (starting at 50\%) to allow smooth adjustment, in the second simulation.
+
+A 1000 period prices simulation is created for each model and the returns for these series are calculated as the log change in price.
+Those periods when mis-perception or the weight of speculators is \emph{extreme} are identified as those that are at or above the ninety-fifth percentile.
+These extremes are the \emph{event} for an event study.
+The package \emph{eventstudies} \citep{eventstudies} aligns the defined events (in this case the extreme noise-trader activity) so that descriptive statistics and other analysis can be carried out on the period after the event has occurred.
+Descriptive statistics for the whole series of returns and for those when misperception or weight of noise-traders are presented in Table \ref{tabref:MC1}.
+The process is then repeated for one thousand simulations of each model so that confidence intervals can be established for the estimates of the descriptive statistics.
+
+\subsection{The results of simulation}
+The results of the simulation are recorded in Table \ref{tabref:MC1}.
+It shows that the mean return to an investment strategy in the whole of the sample is very close to zero.
+However, for each of the extreme readings, there are large negative returns to follow.
+This is consistent with the structure of the model: noise-traders drive price above intrinsic value and subsequent returns will be negative; the more extreme the misperception or the weight of uninformed speculative noise-traders in the market, the greater the negative return that is likely to be recorded.
+
+\begin{table}[h!]
+\begin{threeparttable}
+\caption{Results of the Monte Carlo Simulation of the Noise-Trader Model}
+\begin{tabular}{p{6cm} | r r r r}
+ \label{tabref:MC1}
+                            & Base Case  & Extreme (MPM) & Extreme (RWM) \\
+  \hline
+\hline
+Mean                     & -0.00 & -0.90*   & -0.49 \\
+Standard Deviation  & 0.98 &  0.49*     &  0.42*\\
+Skew                      &  -0.07 &  -0.16   &  -0.78*\\
+Kurtosis                  &  -0.14 &  -0.38    &  -0.35\\
+Number                 &  1000        &   50        & 50\\
+
+\end{tabular}
+\begin{tablenotes}
+\small
+\item Descriptive statistics for the De Long Noise-Trader model \citep{Delong1990noise}.
+The Base Case measures the benchmark as returns from prices generated by Equation \ref{eq:NP2} for the whole sample.
+The MPM measures returns after a period of extreme sentiment when the misperception of noise-traders ($\rho$) is equal to or above the 95th percentile of all the misperceptions; the RMW measures the returns following a recording of extreme weight of noise-traders ($\mu$) is equal to or above the 95th percentile of all noise-trader realisations.
+The asterisk denotes significantly different from zero at the 95th confidence interval.
+\end{tablenotes}
+\end{threeparttable}
+\end{table}
+
+% this needs to be completed. 
+\subsection{Measuring the intensity of speculative sentiment}
+The measurement of speculative intensity must identify how strongly speculators believe that the market will move in a particular direction.
+In the absence of a daily survey of speculators, this study uses \emph{risk-reversal skew} as an objective measure of speculative sentiment rather than surveys of opinion.\footnote{There are a number of surveys of investor opinion that are regularly conducted by financial service information companies.
+However, these are disparate and, in many cases, unsystematic, and it is not clear whether they measure speculative sentiment or investor opinion more generally.
+The Wall Street Journal runs an annual survey of investor opinion, but this is focused mostly on US equities.
+The \emph{Economist} magazine also solicits opinion, but in a relatively unstructured and \emph{ad hoc} fashion.
+In the foreign exchange market, there is a broad set of market forecasts that was collected by the financial services company \emph{MMS International} (formerly Money Market Services) which polled a broad range of financial market participants on key exchange rates for one and four week forecast horizons.
+Frankel and Froot found that the median forecast from the MMS survey exhibited trend-following behaviour at the one to three week horizon while there is mean-reversion beyond that point.
+Frankel and Froot also assess the 'Financial Report' that has been produced by \emph{The Economist} magazine and \emph{Currency Forecasters Digest} and find a wide dispersion of opinion and express some concern about the aggregation of such divergent forecasters \citep{FrootFrankelFDB} and \citep{FrankelFroot1987} \label{fnsurvey}.}    This is the relative price of currency put and call options for the same strike price.
+The \emph{Put-Call-Parity} in the Black-Scholes option pricing model assumes that these will be the same \citep{Black1973Options}.
+While the Black-Scholes model is the basis for valuing options, where the price of a non-dividend paying stock trades at a discount to the forward rate equal to the risk free rate, the forward rate in a currency option is determined by the relative interest rate differential between the two currencies.
+The Garman-Kohlhagan adjustment that accounts for the difference in interest payments on the two currencies is used.
+The spot discount is determined therefore by the relative interest rate differential for the appropriate maturity \citep{GarmanKohlhagan}.
+This model allows the market practice of pricing options from the volatility implied by the model.
+In other words, the Garman-Kohlhagan is used to extract the volatility that is implied by a specific option price given the strike and other parameters.
+
+Determining the fundamental value of exchange rates has proved to be extremely difficult.
+As PPP is, at best, a medium-term phenomenon and given the finding by Meese and Rogoff about the superiority of the random walk relative to other economic models and accepting for the moment the EMH, it will be assumed that the forward rate is the best estimate of the future value of the exchange rate.
+Any forecast which deviates from the forward rate will in this case be counted as a misperception and the further away from the random walk the greater the misperception.
+\footnote{The expected  exchange rate should be equal on average to the forward rate once any risk premia has been take into account.
+Though it will be seen that UIP and the assumption that the future exchange rate is on average equal to the interest rate differential does not hold, the fact that daily data is being used here and the attention is on the most extreme divergences from UIP suggests that this will not greatly affect the results.}
+
+This study uses a unique dataset of \emph{risk-reversals} for eleven major currency pairs.
+This is the only period for which the data currently exist.
+The risk-reversal data is very difficult to obtain in an over-the-counter market like foreign exchange where there is no centralised exchange to record and collect information, generally being the property of quoting institutions.
+The risk-reversals are the average prices quoted at 10:00 GMT by major banks on the Thomson-Reuters composite option pages (OPT1)  for the period between January 2 1996 and September 6 2002.
+The risk-reversals are the spread of implied volatility for a 1 month 25 delta at-the-money call option relative to the equivalent put.
+A call option is the right to buy and a put option is the right to sell.
+The delta of an option measures the effect of a change in the spot price on the value of the option.
+This can range from zero for deeply out-of-the-money options to unity for deeply in the money strikes.
+The at-the-money delta is 50 as there is a 50:50 chance of the option being exercised.
+The benchmark for option risk-reversals is 25 delta.
+This is a measure of how far the option is out of the money or how far the strike is from the forward rate.
+
+If the exchange rate trajectory is expected to follow a random walk with drift towards the 1 month forward, implied volatility on the put and the call for the same delta should be equal.
+Any deviation from this equivalence is an indication that a deviation from the random walk with drift is implied in the market price of options.\footnote{If the risk-reversal for a 1-month USD/CAD is quoted at 0.15-0.28\% and the implied volatility is 8.50\%, the market-maker would be willing to buy the 1-month 25 delta USD-put-CAD-call at 8.65\% and sell the USD-call-CAD-put at 8.50\%.
+The dealer pays 0.15\%.
+Alternatively, the dealer would sell a 25 delta call at 8.78\% and buy the USD-call-CAD-put at 8.50\%, earning 0.28\% \citep{Global}.
+}  Therefore, the relationship between implied volatility for calls and implied volatilities for puts for the same delta and maturity can be taken as an indication of how much more expensive calls are relative to the equivalent put or as the market bias towards puts over calls or, in this case, as an indication of market sentiment.
+The larger the absolute level of the risk-reversal skew, the greater the relative price and, in equilibrium, the greater the belief, expressed through the action of market participants or the inventory of market-makers, that there will be a deviation from the forward rate.
+If extreme deviations or speculations are noise they should presage a sharp reversal.
+If speculation is informed, this is part of the price discovery process and extremes may be followed by calm conditions around a new equilibrium.
+
+% The table may be pushed onwards.  Use t (i think). 
+\begin{landscape}
+\begin{table}[t]
+\begin{threeparttable}
+\caption{Descriptive Statistics for Daily Exchange Rate Returns: 2 January 1996 to 6 September 2002}
+\begin{tabular}{l | r r  r r r r r r r r}
+  \hline
+ Statistic &  AUD & EURGBP & EURSEK & GBP & CHF & EURCHF & EURJPY & EUR & CAD & JPY \\
+  \hline
+Number & 1685 & 1675 & 1716 & 1667 & 1674 & 1674 & 1675 & 1668 & 1683 & 1675  \\
+Mean & -0.0174 & -0.0187 & 0.0011 & 0.0025 & 0.0120 & -0.0039 & -0.0072 & -0.0172 & 0.0084 & 0.0076\\
+Median &  0.0000 & -0.0298 & 0.0000& 0.0000 & 0.0114 & 0.0000 & 0.0000 & -0.0253 & 0.0000 & 0.0188\\
+Standard deviation  & 0.7241 & 0.5542 & 0.0438 & 0.0489 & 0.6819 & 0.0277 & 0.8416 & 0.6338 & 0.0363 & 0.8170\\
+Max & 5.8301 & 3.0543 & 2.1633 & 2.2573 & 3.5086 & 1.6477 & 4.0879 & 2.4136 & 2.2170 & 5.1788 \\
+Min & -3.4666 & -2.2858 & -2.2900 & -2.2701 & -3.6525 & -1.8466 & -8.2848 & -2.2901 & -2.9707 & -10.6573 \\
+Skewness  & 0.1206* & 0.2923* & 0.0432 & -0.0452 & -0.1423* & -0.4046* & -0.7388* & 0.1738* & -0.1095 & -1.7237*\\
+Kurtosis & 4.17* & 1.88 & 2.43* & 1.48 & 1.96 & 5.84* & 7.08* & 1.12 & 5.11* & 21.52*\\
+\hline
+ \label{tabref:DS1}
+\end{tabular}
+\begin{tablenotes}
+\small
+\item Continuously compounded returns (multiplied by 100) for period 2 January 1996 to  September 2002 for US dollar per unit of Australian dollar (AUD), UK sterling per unit of Euro (EURGBP), Swedish Crown per unit of EUR (EURSEK); US dollar per unit of UK sterling (GBP);  Swiss franc per unit of US dollar (CHF); Swiss franc per unit of Euro (EURCHF); US dollars per unit of Euro (EUR); Canadian dollars per unit of US dollar (CAD); Japanese yen per unit of US dollars (JPY).
+The asterisk (*) means that the estimate is significantly different from zero at the 5\% level using t-statistics for the mean and the standard error of skew and standard error of excess-kurtosis respectively.
+\end{tablenotes}
+\end{threeparttable}
+\end{table}
+\end{landscape}
+
+Table \ref{tabref:DS1} shows descriptive statistics for the exchange rate data used in the study.
+There are daily data for the period between January 1996 and September 2002.
+This is the period for which risk-reversal data are available.
+The series are of sightly different lengths as there are some gaps in the risk reversal and exchange rate data as a result of holidays and market closures.
+From the raw data continuously compounded returns are calculated and multiplied by 100.
+The first thing of note is that the mean return does not appear to be statistically different from zero.
+This is what would be expected if, as has been identified in many studies, exchange rates follow a random walk.
+This apparent stability hides some significant changes that are taking place below the surface.
+For example, the Japanese yen records an appreciation of 10.66\% on one day.
+Most of the exchange rate series appear to be skewed in some way and there is also evidence of kurtosis in many cases.
+An asterisk is used to denote statistical significance at the level of 95\% confidence.
+A t-test is used for the mean; a \emph{standard error of skew} and a \emph{standard error of kurtosis} are used for the skew and kurtosis respectively.
+These are calculated as $SES = \sqrt{\frac{6n(n-1)}{(n-2)(N+1)(N+3)}}$ and $SEK = \sqrt{\frac{n^2-1}{(n-3)(N+5)}}$, with the assumption that this estimate is distributed normally, the skew or kurtosis is considered significant if it is more than two standard errors from zero.
+
+\begin{landscape}
+\begin{table}[t!]
+\begin{threeparttable}
+\caption{Descriptive Statistics for Risk Reversals: 2 January 1996 to 6 September 2002}
+\begin{tabular}{l | r r  r r r r r r r r}
+  \hline
+ Statistic &  AUD & EURGBP & EURSEK & GBP & CHF & EURCHF & EURJPY & EUR & CAD & JPY \\
+ \hline
+Number & 1702 & 1697 & 1719 & 1691 & 1374 & 1339 & 1697 & 1692 & 1701 & 1697 \\ 
+  Mean & -0.37 & 0.15 & 0.71 & -0.08 & -0.28 & -0.38 & -0.48 & 0.19 & 0.12 & -0.74 \\ 
+  Median & -0.30 & 0.10 & 0.80 & -0.10 & -0.15 & -0.30 & -0.10 & 0.10 & 0.15 & -0.55 \\ 
+  Standard Deviation & 0.38 & 0.37 & 0.51 & 0.33 & 0.49 & 0.36 & 1.02 & 0.49 & 0.25 & 1.02 \\ 
+ Max & 0.45 & 1.72 & 1.90 & 1.00 & 1.05 & 0.50 & 1.70 & 2.05 & 1.45 & 2.05 \\ 
+ Min & -1.70 & -0.80 & -0.30 & -1.35 & -2.05 & -1.80 & -4.03 & -1.15 & -0.65 & -4.00 \\ 
+ Skewness & -0.50* & 1.27* & -0.23* & 0.33* & -0.79* & -0.58* & -0.93* & 0.68* & 0.23* & -0.51* \\ 
+  Kurtosis & 0.02 & 2.29* & -1.67 & 0.72 & 0.69 & 0.19 & 0.15 & 0.96 & 1.83 & -0.03 \\ 
+      \hline
+ \label{tabref:DS2}
+\end{tabular}
+\begin{tablenotes}
+\small
+\item Daily risk reversal skew.
+Risk reversal is the premium of implied volatility for calls relative to the equivalent put (equal distance to the strike price, 25 delta in this case).
+A figure above zero means that calls are more expensive (in volatility and price terms) than puts.
+A figure below zero means that puts are more expensive.
+The asterisk (*) means that the estimate is significant at the 5\% level using t-statistics for the mean and the standard error of skew and standard error of excess-kurtosis respectively.
+These are calculated as $SES = \sqrt{\frac{6n(n-1)}{(n-2)(N+1)(N+3)}}$ and $SEK = \sqrt{\frac{n^2-1}{(n-3)(N+5)}}$, with the assumption that this estimate is distributed normally, the skew or kurtosis is considered significant if it is more than two standard errors from zero.
+\end{tablenotes}
+\end{threeparttable}
+\end{table}
+\end{landscape}
+
+Table \ref{tabref:DS2} gives the descriptive statistics for the risk-reversal data.
+For a neutral risk-reversal, a reading of zero, there would be no price preference for calls or puts at the similar delta.
+The figures show that, in contrast to the average returns, there are often average skews that are different from zero.
+The most significant of these are the positive bias for EUR-GBP and a negative bias for USD-JPY.
+The latter may be related to hedging activity by Japanese exporters.
+While this could suggest that implied volatility is an imperfect measure of speculative activity, note that extreme readings are based on quantiles of the distribution and therefore an underlying hedging-induced shift in the mean of the distribution should not affect the results.
+In other words, if the extreme is denoted by the 90th percentile, it does not matter if the mean is zero and this comes at 0.8 or if the mean is 0.2 and this comes at 1.0.
+\subsection{Measuring the intensity of speculative sentiment}
+The measurement of speculative intensity must identify how strongly speculators believe that the market will move in a particular direction.
+In the absence of a daily survey of speculators, this study uses \emph{risk-reversal skew} as an objective measure of speculative sentiment rather than surveys of opinion.\footnote{There are a number of surveys of investor opinion that are regularly conducted by financial service information companies.
+However, these are disparate and, in many cases, unsystematic, and it is not clear whether they measure speculative sentiment or investor opinion more generally.
+The Wall Street Journal runs an annual survey of investor opinion, but this is focused mostly on US equities.
+The \emph{Economist} magazine also solicits opinion, but in a relatively unstructured and \emph{ad hoc} fashion.
+In the foreign exchange market, there is a broad set of market forecasts that was collected by the financial services company \emph{MMS International} (formerly Money Market Services) which polled a broad range of financial market participants on key exchange rates for one and four week forecast horizons.
+Frankel and Froot found that the median forecast from the MMS survey exhibited trend-following behaviour at the one to three week horizon while there is mean-reversion beyond that point.
+Frankel and Froot also assess the 'Financial Report' that has been produced by \emph{The Economist} magazine and \emph{Currency Forecasters Digest} and find a wide dispersion of opinion and express some concern about the aggregation of such divergent forecasters \citep{FrootFrankelFDB} and \citep{FrankelFroot1987} \label{fnsurvey}.}    This is the relative price of currency put and call options for the same strike price.
+The \emph{Put-Call-Parity} in the Black-Scholes option pricing model assumes that these will be the same \citep{Black1973Options}.
+While the Black-Scholes model is the basis for valuing options, where the price of a non-dividend paying stock trades at a discount to the forward rate equal to the risk free rate, the forward rate in a currency option is determined by the relative interest rate differential between the two currencies.
+The Garman-Kohlhagan adjustment that accounts for the difference in interest payments on the two currencies is used.
+The spot discount is determined therefore by the relative interest rate differential for the appropriate maturity \citep{GarmanKohlhagan}.
+This model allows the market practice of pricing options from the volatility implied by the model.
+In other words, the Garman-Kohlhagan is used to extract the volatility that is implied by a specific option price given the strike and other parameters.
+
+Determining the fundamental value of exchange rates has proved to be extremely difficult.
+As PPP is, at best, a medium-term phenomenon and given the finding by Meese and Rogoff about the superiority of the random walk relative to other economic models and accepting for the moment the EMH, it will be assumed that the forward rate is the best estimate of the future value of the exchange rate.
+Any forecast which deviates from the forward rate will in this case be counted as a misperception and the further away from the random walk the greater the misperception.
+\footnote{The expected  exchange rate should be equal on average to the forward rate once any risk premia has been take into account.
+Though it will be seen that UIP and the assumption that the future exchange rate is on average equal to the interest rate differential does not hold, the fact that daily data is being used here and the attention is on the most extreme divergences from UIP suggests that this will not greatly affect the results.}
+
+This study uses a unique dataset of \emph{risk-reversals} for eleven major currency pairs.
+This is the only period for which the data currently exist.
+The risk-reversal data is very difficult to obtain in an over-the-counter market like foreign exchange where there is no centralised exchange to record and collect information, generally being the property of quoting institutions.
+The risk-reversals are the average prices quoted at 10:00 GMT by major banks on the Thomson-Reuters composite option pages (OPT1)  for the period between January 2 1996 and September 6 2002.
+The risk-reversals are the spread of implied volatility for a 1 month 25 delta at-the-money call option relative to the equivalent put.
+A call option is the right to buy and a put option is the right to sell.
+The delta of an option measures the effect of a change in the spot price on the value of the option.
+This can range from zero for deeply out-of-the-money options to unity for deeply in the money strikes.
+The at-the-money delta is 50 as there is a 50:50 chance of the option being exercised.
+The benchmark for option risk-reversals is 25 delta.
+This is a measure of how far the option is out of the money or how far the strike is from the forward rate.
+
+If the exchange rate trajectory is expected to follow a random walk with drift towards the 1 month forward, implied volatility on the put and the call for the same delta should be equal.
+Any deviation from this equivalence is an indication that a deviation from the random walk with drift is implied in the market price of options.\footnote{If the risk-reversal for a 1-month USD/CAD is quoted at 0.15-0.28\% and the implied volatility is 8.50\%, the market-maker would be willing to buy the 1-month 25 delta USD-put-CAD-call at 8.65\% and sell the USD-call-CAD-put at 8.50\%.
+The dealer pays 0.15\%.
+Alternatively, the dealer would sell a 25 delta call at 8.78\% and buy the USD-call-CAD-put at 8.50\%, earning 0.28\% \citep{Global}.
+}  Therefore, the relationship between implied volatility for calls and implied volatilities for puts for the same delta and maturity can be taken as an indication of how much more expensive calls are relative to the equivalent put or as the market bias towards puts over calls or, in this case, as an indication of market sentiment.
+The larger the absolute level of the risk-reversal skew, the greater the relative price and, in equilibrium, the greater the belief, expressed through the action of market participants or the inventory of market-makers, that there will be a deviation from the forward rate.
+If extreme deviations or speculations are noise they should presage a sharp reversal.
+If speculation is informed, this is part of the price discovery process and extremes may be followed by calm conditions around a new equilibrium.
+
+% The table may be pushed onwards.  Use t (i think). 
+\begin{landscape}
+\begin{table}[t]
+\begin{threeparttable}
+\caption{Descriptive Statistics for Daily Exchange Rate Returns: 2 January 1996 to 6 September 2002}
+\begin{tabular}{l | r r  r r r r r r r r}
+  \hline
+ Statistic &  AUD & EURGBP & EURSEK & GBP & CHF & EURCHF & EURJPY & EUR & CAD & JPY \\
+  \hline
+Number & 1685 & 1675 & 1716 & 1667 & 1674 & 1674 & 1675 & 1668 & 1683 & 1675  \\
+Mean & -0.0174 & -0.0187 & 0.0011 & 0.0025 & 0.0120 & -0.0039 & -0.0072 & -0.0172 & 0.0084 & 0.0076\\
+Median &  0.0000 & -0.0298 & 0.0000& 0.0000 & 0.0114 & 0.0000 & 0.0000 & -0.0253 & 0.0000 & 0.0188\\
+Standard deviation  & 0.7241 & 0.5542 & 0.0438 & 0.0489 & 0.6819 & 0.0277 & 0.8416 & 0.6338 & 0.0363 & 0.8170\\
+Max & 5.8301 & 3.0543 & 2.1633 & 2.2573 & 3.5086 & 1.6477 & 4.0879 & 2.4136 & 2.2170 & 5.1788 \\
+Min & -3.4666 & -2.2858 & -2.2900 & -2.2701 & -3.6525 & -1.8466 & -8.2848 & -2.2901 & -2.9707 & -10.6573 \\
+Skewness  & 0.1206* & 0.2923* & 0.0432 & -0.0452 & -0.1423* & -0.4046* & -0.7388* & 0.1738* & -0.1095 & -1.7237*\\
+Kurtosis & 4.17* & 1.88 & 2.43* & 1.48 & 1.96 & 5.84* & 7.08* & 1.12 & 5.11* & 21.52*\\
+\hline
+ \label{tabref:DS1}
+\end{tabular}
+\begin{tablenotes}
+\small
+\item Continuously compounded returns (multiplied by 100) for period 2 January 1996 to  September 2002 for US dollar per unit of Australian dollar (AUD), UK sterling per unit of Euro (EURGBP), Swedish Crown per unit of EUR (EURSEK); US dollar per unit of UK sterling (GBP);  Swiss franc per unit of US dollar (CHF); Swiss franc per unit of Euro (EURCHF); US dollars per unit of Euro (EUR); Canadian dollars per unit of US dollar (CAD); Japanese yen per unit of US dollars (JPY).
+The asterisk (*) means that the estimate is significantly different from zero at the 5\% level using t-statistics for the mean and the standard error of skew and standard error of excess-kurtosis respectively.
+\end{tablenotes}
+\end{threeparttable}
+\end{table}
+\end{landscape}
+
+Table \ref{tabref:DS1} shows descriptive statistics for the exchange rate data used in the study.
+There are daily data for the period between January 1996 and September 2002.
+This is the period for which risk-reversal data are available.
+The series are of sightly different lengths as there are some gaps in the risk reversal and exchange rate data as a result of holidays and market closures.
+From the raw data continuously compounded returns are calculated and multiplied by 100.
+The first thing of note is that the mean return does not appear to be statistically different from zero.
+This is what would be expected if, as has been identified in many studies, exchange rates follow a random walk.
+This apparent stability hides some significant changes that are taking place below the surface.
+For example, the Japanese yen records an appreciation of 10.66\% on one day.
+Most of the exchange rate series appear to be skewed in some way and there is also evidence of kurtosis in many cases.
+An asterisk is used to denote statistical significance at the level of 95\% confidence.
+A t-test is used for the mean; a \emph{standard error of skew} and a \emph{standard error of kurtosis} are used for the skew and kurtosis respectively.
+These are calculated as $SES = \sqrt{\frac{6n(n-1)}{(n-2)(N+1)(N+3)}}$ and $SEK = \sqrt{\frac{n^2-1}{(n-3)(N+5)}}$, with the assumption that this estimate is distributed normally, the skew or kurtosis is considered significant if it is more than two standard errors from zero.
+
+\begin{landscape}
+\begin{table}[t!]
+\begin{threeparttable}
+\caption{Descriptive Statistics for Risk Reversals: 2 January 1996 to 6 September 2002}
+\begin{tabular}{l | r r  r r r r r r r r}
+  \hline
+ Statistic &  AUD & EURGBP & EURSEK & GBP & CHF & EURCHF & EURJPY & EUR & CAD & JPY \\
+ \hline
+Number & 1702 & 1697 & 1719 & 1691 & 1374 & 1339 & 1697 & 1692 & 1701 & 1697 \\ 
+  Mean & -0.37 & 0.15 & 0.71 & -0.08 & -0.28 & -0.38 & -0.48 & 0.19 & 0.12 & -0.74 \\ 
+  Median & -0.30 & 0.10 & 0.80 & -0.10 & -0.15 & -0.30 & -0.10 & 0.10 & 0.15 & -0.55 \\ 
+  Standard Deviation & 0.38 & 0.37 & 0.51 & 0.33 & 0.49 & 0.36 & 1.02 & 0.49 & 0.25 & 1.02 \\ 
+ Max & 0.45 & 1.72 & 1.90 & 1.00 & 1.05 & 0.50 & 1.70 & 2.05 & 1.45 & 2.05 \\ 
+ Min & -1.70 & -0.80 & -0.30 & -1.35 & -2.05 & -1.80 & -4.03 & -1.15 & -0.65 & -4.00 \\ 
+ Skewness & -0.50* & 1.27* & -0.23* & 0.33* & -0.79* & -0.58* & -0.93* & 0.68* & 0.23* & -0.51* \\ 
+  Kurtosis & 0.02 & 2.29* & -1.67 & 0.72 & 0.69 & 0.19 & 0.15 & 0.96 & 1.83 & -0.03 \\ 
+      \hline
+ \label{tabref:DS2}
+\end{tabular}
+\begin{tablenotes}
+\small
+\item Daily risk reversal skew.
+Risk reversal is the premium of implied volatility for calls relative to the equivalent put (equal distance to the strike price, 25 delta in this case).
+A figure above zero means that calls are more expensive (in volatility and price terms) than puts.
+A figure below zero means that puts are more expensive.
+The asterisk (*) means that the estimate is significant at the 5\% level using t-statistics for the mean and the standard error of skew and standard error of excess-kurtosis respectively.
+These are calculated as $SES = \sqrt{\frac{6n(n-1)}{(n-2)(N+1)(N+3)}}$ and $SEK = \sqrt{\frac{n^2-1}{(n-3)(N+5)}}$, with the assumption that this estimate is distributed normally, the skew or kurtosis is considered significant if it is more than two standard errors from zero.
+\end{tablenotes}
+\end{threeparttable}
+\end{table}
+\end{landscape}
+
+Table \ref{tabref:DS2} gives the descriptive statistics for the risk-reversal data.
+For a neutral risk-reversal, a reading of zero, there would be no price preference for calls or puts at the similar delta.
+The figures show that, in contrast to the average returns, there are often average skews that are different from zero.
+The most significant of these are the positive bias for EUR-GBP and a negative bias for USD-JPY.
+The latter may be related to hedging activity by Japanese exporters.
+While this could suggest that implied volatility is an imperfect measure of speculative activity, note that extreme readings are based on quantiles of the distribution and therefore an underlying hedging-induced shift in the mean of the distribution should not affect the results.
+In other words, if the extreme is denoted by the 90th percentile, it does not matter if the mean is zero and this comes at 0.8 or if the mean is 0.2 and this comes at 1.0.
+
+\subsection{Measuring the proportion of speculators}
+Recording speculative activity in an over-the-counter market like foreign exchange is extremely difficult.
+However, there are some exchange-traded currency derivative markets in the US, and participants in these markets are required to report their positions to the US regulators.
+The Commodity Futures Trading Commission (CFTC) was set up by Congress in 1974 to regulate commodity future and options markets.
+The CFTC collect weekly information about the open positions held by private entities in the US derivative markets.
+A sub-set of this information is released to the public each week as the \emph{Commitment of Traders Report} (CoT).
+The weekly release has been in place since September 30, 1992.
+Before that the data were released every two weeks.
+The CoT is released each Friday and details the positions outstanding the previous Tuesday \citep{cot}.
+
+The CFTC require traders to categorise themselves as being either commercial (C) or non-commercial (NC).
+Commercial traders must have some underlying business interest in the security, commodity or instrument that they are trading.
+These could be seen as natural hedgers.
+Non-commercial accounts are generally considered to be speculators.
+See \citep{FuturesSanders} and \citep{FuturesWang} for examples of previous use of this data in this fashion.
+
+The administrative requirements are more onerous for non-commercial accounts but the CFTC have the right, which they use, to re-categorise parties if they are not happy with the self-selection.
+Positions are either long or short, meaning that they will receive or deliver the underlying; for non-commercial accounts there are also \emph{spreading} positions which are simultaneous holdings of long and short positions in different maturities of the same contract, taking a position on the spread between two maturity dates.
+These are assumed to be speculators.
+The CFTC data also report \emph{Open Interest} (OI) which is the record of outstanding open positions.
+
+If it is assumed that futures and cash markets are two separate markets and that most participants have a preference for one or the other and therefore do not frequently switch; if it is also assumed that the nature and level of speculative activity in the futures market is an indicator of what is happening in the broader spot market, it will be possible to use this data, which measures speculative activity in the futures market to assess the intensity and level of speculative activity in the foreign exchange market more generally.
+In other words, the futures market is used as a sample of the entire foreign exchange market.
+This is important because, unlike many other financial markets, it is the cash foreign exchange market that dominates in terms of activity and liquidity.
+The futures market is relatively small \citep{BISFX}.
+
+%The operation of derivative exchanges is a competitive business that has evolved over the course of the data collection period.
+The two original exchanges in Chicago, the Chicago Mercantile Exchange (CME) and the  Chicago Board of Trade (CBOT) went public in 2000 and 2005 respectively and merged in 2007 to become the CEM Group.
+In 2008 the New York Mercantile Exchange (NYMEX) was added,  Most of the successful currency derivative contracts were traded on the International Monetary Market (IMM) which was a sub-division of the original CME.
+Many exchange rates and cross rate products were launched by these and other US exchanges, but the five contract which maintained a presence and liquidity through the whole period were the Deutschmark and then the Euro, the Swiss Franc, the British Pound, the Japanese Yen and the Canadian Dollar.
+There is a reasonably long series for the Australian Dollar and for the Mexican Peso, but nether of these is complete.
+The US dollar index has been active since it was launched in May 1992 but it is not trade weighted or really representative of the US dollar.
+
+It could be argued that this is a biased sample as it is a US market.
+Though international financial markets are global in nature, there is some evidence to support this claim.
+For example, there is a tendency to have a positive holding of the overseas currency as this is a hedge for exporters.
+Nonetheless, as was argued with the risk-reversals, it is the relative rather than absolute position of the signal that will determine the extreme and therefore if the US sample is proportionate to the population, sample signals can still be regarded as measuring an extreme.
+If the distribution of the sample itself is different from the population this is a problem, but there is no evidence that this is the case.
+
+The data used run from September 30 1998 when the data started to be released on a weekly basis and continue up to December 31 2008.
+This date is shortly before this part of the research was started.
+The Euro data runs  from March 3 1998 to December 31 2008.
+There are two gaps in the Swiss Franc data for September 14 2004 and September 21 2004.
+The absence of these data seems to be a result of problems with the CFTC database.
+The contracts are for EUR 100,000; GBP 62,500; JPY 12,500,000; CHF 100,000 and CAD 100,000.
+For the CoT data, two series are created:  S1 measures the net long non-commercial positions relative to total non-commercial positions and S2 measures the net long non-commercial positions relative to open interest.
+Table \ref{tabref:DS3} given an overview of the descriptive statistics for the returns on the weekly foreign exchange rates that are relevant for the futures market data.
+There are no significant differences from the daily data that is used in the risk-reversal part of the study recorded in Table \ref{tabref:DS1}.
+Once again, the mean return is very close to zero.
+This is consistent with the idea that exchange rates follow a random walk.
+The distribution of returns are skewed for some pairs and often fat-tailed (identified with an asterisk where statistically significant as either a t-test or a test of the standard error of skew or standard error of kurtosis as calculated above).
+
+ \begin{table}[t]
+\begin{threeparttable}
+\caption{Descriptive Statistics for Weekly Exchange Rate Returns: 30 September 1998 to 31 December 2008}
+\begin{tabular}{l | r r  r r r }
+  \hline
+ Statistic &  EUR & JPY & GBP & CHF & CAD  \\
+  \hline
+Number & 522 &  849 & 849 & 847 & 849 \\
+Mean & 0.0351 & 0.0337 & -0.0202 & 0.0233 & 0.0018\\
+Median & 0.0650 & -0.0800 & 0.0000 & 0.0000 & -0.0200\\
+Standard deviation  & 1.4063 & 1.6355 & 1.2755 & 1.5418 & 1.0558\\
+Max & 8.7900 & 11.6700 & 5.7200 & 8.3400 & 10.0100\\
+Min & -4.2400 & -15.0400 & -6.4100 & -5.7100 & -5.0900\\
+Skew  & 0.27* & -0.06 & -0.34* & 0.31* & 0.78*\\
+Kurtosis & 2.76*& 12.58* & 2.07 * & 1.82 & 11.59*\\
+\hline
+ \label{tabref:DS3}
+\end{tabular}
+\begin{tablenotes}
+\small
+\item Continuously compounded returns multiplied by 100 for US dollar per unit of Euro (EUR), US dollar per unit of Japanese Yen (JPY), US dollar per unit of UK Sterling (GBP), US dollar per unit of Swiss Franc (CHF) and US dollar per unit of Canadian dollar (CAD) for the period 30 September 1998 to 31 Dec 2008;     An asterisk (*) means that the estimate is significantly different from zero at the 5\% level of statistical significance using t-statistics for the mean and the standard error of skew and standard error of excess-kurtosis respectively.
+These are calculated as $SES = \sqrt{\frac{6n(n-1)}{(n-2)(N+1)(N+3)}}$ and $SEK = \sqrt{\frac{n^2-1}{(n-3)(N+5)}}$, with the assumption that this estimate is distributed normally, the skew or kurtosis is considered significant if it is more than two standard errors from zero.
+\end{tablenotes}
+\end{threeparttable}
+\end{table}
+
+Table \ref{tabref:DS4} shows the descriptive statistics for the two measures of speculative sentiment.
+The first (S1) is measured as the net non-commercial long and non-commercial short proportionate to the sum of non-commercial positions long and short positions; (S2) is measured as the net non-commercial long and spreading positions proportionate to open interest.
+The first captures directional speculation and, as can be seen from a number of minimums, this can reach 100\%.
+The second records the weight of speculation relative to everything else in the market.
+The former is more sensitive, as can be seen by the range and the standard deviation; the second is usually above zero because there are always some spreading positions that seek to take advantage of calendar spreads and other strategies.
+
+\begin{landscape}
+\begin{table}[t]
+\begin{threeparttable}
+\caption{Descriptive Statistics for S1 (Speculative Sentiment) and S2 (Speculative Weight): 30 September 1998 to 31 December 2008}
+\begin{tabular}{l | r r r r r r r  r r r }
+  \hline
+ Statistic &  \multicolumn{2}{c}{EUR} & \multicolumn{2}{c}{JPY} & \multicolumn{2}{c}{GBP} & \multicolumn{2}{c}{CHF} & \multicolumn{2}{c}{CAD}  \\
+ & S1 & S2 & S1 & S2 & S1 & S2 & S1 & S2 & S1 & S2 \\  
+\hline
+Number & 522 &  522 & 849 & 849 & 849 & 849 & 847 & 847 & 849 & 849\\
+Mean &  31.31 & 43.43 & -23.90 & 44.21 & 2.76 & 41.68 & -21.14 & 46.13 & 6.68 & 41.23\\
+Median & 42.75 & 39.56 & -33.30 & 43.56 & 11.15 & 38.88 & -37.32 & 42.88 & 7.44 & 39.81\\
+Standard deviation  & 41.94 & 16.13 & 51.49 & 12.17 & 60.59 & 16.46 & 60.76 & 15.73 & 55.26 & 15.33\\
+Max & 95.67 & 86.06 & 87.43 & 84.50 & 96.15 & 89.45 & 100.00 & 92.49 & 100.00 & 86.63\\
+Min &  -100.00 & 11.15 & -99.83 & 12.88 & -100.00 & 9.10 & -100.00 & 14.61 & -90.08 & 11.89\\
+Skew  & -0.88* & 0.55* & 0.33* & 0.57* & -0.17 & 0.76* & 0.45* & 0.65* & 0.02 & 0.50*\\
+Kurtosis & 0.18 & -0.57 & -1.14 & 0.64 & -1.45 & 0.15 & -1.25 & -0.15 & -1.35 & -0.28\\
+\hline
+ \label{tabref:DS4}
+\end{tabular}
+\begin{tablenotes}
+\small
+\item S1 is the first sentiment indicator, calculated as ratio of net non-commercial long less non-commercial short to the sum of non-commercial long and short positions multiplied by one hundred.
+S2 is the second sentiment indicators, calculated as the ratio of net non-commercial long to net non-commercial short plus spreading to open interest.
+An asterisk (*) means that the estimate is significantly different from zero at the 5\% level using t-statistics for the mean and the standard error of skew and standard error of excess-kurtosis respectively.
+These are calculated as $SES = \sqrt{\frac{6n(n-1)}{(n-2)(N+1)(N+3)}}$ and $SEK = \sqrt{\frac{n^2-1}{(n-3)(N+5)}}$, with the assumption that this estimate is distributed normally, the skew or kurtosis is considered significant if it is more than two standard errors from zero.
+\end{tablenotes}
+\end{threeparttable}
+\end{table}
+\end{landscape}
+
+\subsection{Event studies}
+Event studies are used to find the influence of a particular event on the performance of an asset price or security.
+James Dolley is credited with the first use of the method when looking at the effect of a stock split on the price of securities.\footnote{A stock split is when the number of shares are doubled and the value of each share is cut in half.
+This is usually done to prevent the price or cost of each share rising too far beyond the means of the small investor.} \citep{Dolly1933}.
+However, Dolly only looked at the market reaction the day after the stock split.
+This was some time after the stock holder meeting that would have ratified the decision of the management.
+Myer and Bakay argued that, if stocks tend to rise after a stock split, it would be best to buy the stocks when the announcement is made and before the split takes place and that, therefore, the price adjustment is likely to be spread over a number of days.
+They decided to measure the price action from a point eight weeks before the formal acceptance of the decision by the stock-holders.
+Myer and Barkay measured the price change between the base date and the meeting and tried to account for other influences on the price by recording the ratio of the stock price to the Standard and Poor Industry average, thereby providing the first attempt to separate the specific returns from those of the market \citep{MyersBakay}.
+
+Fama, Fisher and Roll use the event study method to conduct a much more systematic assessment of the effect of stock splits on security returns \citep{FamaFisherJensenRoll}.
+the study is considered to be the benchmark of Event Studies and it is the framework that is largely followed in the current work.
+
+The event is identified and the time at which the event occurs is marked as time zero (or $t_0$).
+Returns to the reference security in a window before and after the event are recorded and compared to the normal or expected returns.
+Therefore,
+\begin{equation}
+AR_{i \tau} = R_{i \tau} - E[R_{i \tau} | X_{\tau}]
+\end{equation}
+Where $AR_{i \tau}$ are the abnormal returns for security i at time $\tau$, $R_{i \tau}$ are the actual returns to security i at time $\tau$,  $E[R_{i \tau}]$ are the expected returns for security i at time $\tau$ conditional on the information $X_{\tau}$ and $\tau$ is the time relative to the event.
+Estimating the normal or expected returns is one of the key challenges of an event  study.
+It is usually tackled in one of two ways:  the security can be assumed to take a mean return or, as in the case of the paper by Fama and the others, a market model can be used.
+The original market model just takes a constant relationship between security returns and the market estimated by a linear regression.
+Subsequent models have added additional factors.
+The lack of any clear benchmark for currency returns and the fact that theoretically exchange rate returns should be on average equal to the forward rate, have made the case for a mean model.
+The seminal event studies have used equities and a market model.
+The expected return is zero and any returns that differ from that are identified as abnormal.
+Theoretically, the forward rate should be used but given the high frequency of the data, the interest rate differential is ignored.
+The evidence about returns in Tables \ref{tabref:DS1} and \ref{tabref:DS3} suggests that this is not likely to be a major problem.
+
+Is more needed here?
+
+\subsection{Analysis of events and abnormal returns}
+The aim of the study is to assess the effect of extreme speculative positions on future returns to ascertain the informational content of these extremes.
+Therefore, the events that will be used to assess abnormal returns are the extreme readings on sentiment and the extreme presence of speculators.
+Any reading above the ninety-ninth percentile or below the first percentile is deemed to be extreme.
+For the risk reversal skew, as a secondary test, a reading above the ninety-fifth and below the fifth percentile is also established as an event.
+The Event Studies R package was used to manage the data see \citep{eventstudies} for fuller details on the package and its use.
+The cumulative abnormal returns for the whole of the event window (CARW) and the cumulative abnormal returns for after the event (CARA) were recorded.
+These cumulations are just the summation of the continuously compounded returns over the appropriate period.
+
+  The date of the events, 1st, 5th, 95th or 99th percentile, reading of the risk-reversal skew or the level of speculative positions,  are identified and the exchange rate returns for the window around the event are aligned so that the average for each event and each event day can be calculated.
+For example, in Table \ref{tabref:RR1} the analysis of the event that the risk-reversal skew for the EUR vs the US dollar is an extreme low (measured as being equal to the 5th percentile or lower), all 106 cases of this extreme for the period under study are aligned and the average cumulative returns for the whole window (CARW - one day before and after the event in column four) and the post-event window (CARA - the event and the day after the event) are calculated.
+A \emph{bootstrap} is conducted.
+This means that a sample of size 106 (in this case), with replacement, is drawn from each of the event days and new means are calculated for the CARW and CARA.
+If 95\% of the means that are calculated after drawing random samples from the array of windows are above or below zero, it is assumed that the figure is statistically different from zero, and this is represented by an asterisk in the table.
+
+\section{Analysis of the results}
+Table \ref{tabref:RR1} shows the result of the event studies that were carried out on extreme risk reversal.  
+For each exchange rate the second column distinguishes between extreme high and extreme low, the third  column identifies the number of extreme events that were recorded and the next two are the cumulative average returns for the whole window (CARW) and the cumulative average returns for the period of the event and afterwards (CARA).
+The windows were chosen to correspond to one day and one week of trading.
+However, Figure \ref{fig:ES1} makes it clear that altering the size of the event window does not change the findings to any significant degree.
+
+Looking down the seventh column (the 4-day CARW) shows the average cumulative return for the period from four days before the extreme event to the four days after the extreme for extreme sentiment highs and lows of each currency pair.
+The extreme highs are (in this case) those points when the risk-reversal skew is at or above the 95th percentile.
+It should not be surprising to see that the extreme highs are associated with positive returns, which appear in many cases to be significantly different from zero (where significance is identified as more than 95\% of 1000 random bootstrap samples taken from the events being above zero).
+It is clear that, as speculators become more extreme in their opinion, prices are driven in the direction of sentiment.
+It shows that the spot price is following the option risk-reversal skew and, as Figure \ref{fig:ES1} reveals, it would be even more pronounced if the event window just ran from the starting point to the event.
+The findings here are consistent with the evidence that orders and sentiment drive prices \citep{Evans2002Order}, \citep{FuturesSanders} and \citep{FuturesWang}.
+
+What is more interesting is whether knowledge about the extreme reading at time $t = 0$ provides any information about future returns.
+If it is known that speculative sentiment is at the extreme, does this mean that it is more likely that there will be a reversal?
+Is there an inefficiency?
+This information is provided in columns 5, 8, 11 and 14 of Table \ref{tabref:RR1}.
+A reversal is indicated by a significant negative reading for highs or positive reading for lows.
+Looking, for example, at the four-day window in column 8, it is clear that the extreme risk reversal skew does not provide any information about future returns.
+Of the eighteen cases using the 95\% th percentile to identify an extreme, ten show a continuation, rather than a reversal, that is not significantly different from zero; one shows a significant continuation; five show reversals that are not significant; only two show significant reversal (EURCHF in each case).
+For the more extreme event when the 99\% the percentile is used to measure the extreme, there is an even more mixed picture.
+Just over half the cases with the four-day window show a reversal (8 of 18), but only 2 of these are statistically significant.
+
+\begin{landscape}
+% multirow package is required.   
+%pdflandscape is required to facilitate the landscape position. 
+\begin{table}[t]
+\begin{threeparttable}
+\caption{Event Study: Cumulative Abnormal Returns and Extreme Risk Reversal}
+\begin{tabular}{llccccccccccccc}  
+ % \hline
+& &\multicolumn{6}{c}{5th and 95th percentile} & \multicolumn{1}{c}{} & \multicolumn{6}{c}{1st and 99th percentile}\\
+ & &\multicolumn{3}{c}{1 day} & \multicolumn{3}{c}{4 day} & \multicolumn{1}{c}{} & \multicolumn{3}{c}{1 day} & \multicolumn{3}{c}{4 day} \\ 
+%\cline{4 - 10} \cline{13 - 19}
+& &\multicolumn{1}{ c}{N} & \multicolumn{1}{c}{CARW} & \multicolumn{1}{c}{CARA} & \multicolumn{1}{ c}{N} & \multicolumn{1}{c}{CARW} & \multicolumn{1}{c}{CARA} & \multicolumn{1}{c}{} & 
+\multicolumn{1}{ c}{N} & \multicolumn{1}{c}{CARW} & \multicolumn{1}{c}{CARA}  
+& \multicolumn{1}{ c}{N} & \multicolumn{1}{c}{CARW} & \multicolumn{1}{c}{CARA}  \\   
+\hline
+\multirow{2}{*}{EUR} 
+& Hi &  87 &  0.4103* &0.1586 &87 &0.9442* & 0.1505 &&22 &0.0309 & -0.1748 &22 & 0.5781*& -0.5387*  \\ 
+& Lo & 106 &-0.3313* & -0.0933 &106 &-0.0816* &-0.1991 & & 21&0.3040 &0.1820 &21 &0.7935* & 0.4377  \\
+\multirow{2}{*}{JPY}
+& Hi & 90 &0.4860* &0.1341 &90 &1.5158* &0.2531 & &20 &0.8227* &0.3078* &20 &2.4337* & 0.8284*  \\ 
+& Lo & 90 &-0.3168 &0.0945 &90 &-1.3532* &0.1984 & &17 &-1.6990* &-0.9347* &17 &-3.6751* & -1.2556*  \\
+\multirow{2}{*}{GBP}
+& Hi & 96 &0.3303* &0.1903* &96 &0.8966* &0.2547 & &21 &0.5252* &0.1774 &21 &1.3131* &0.3284  \\ 
+& Lo & 97 &-0.1894 &0.0312 &97 &-0.8031* &0.1500 & &24 &-0.2630 &-0.0866 &24 &-0.6660 &0.1888  \\
+\multirow{2}{*}{CHF}
+& Hi & 80 &0.0065 &-0.0565 &80 &0.5624* &0.1322 & &18 & -0.0022 & -0.1518 &18 &0.4887 &-0.1768  \\ 
+& Lo & 75 &-0.3716* & -0.1125& 75 &-1.0071* &-0.0219 & &18 &0.0064 &0.1704 &18 &-0.8661* &0.7631  \\
+\multirow{2}{*}{CAD}
+& Hi & 113 &-0.0038 &-0.0489 &113 &0.1872 &-0.1951 & & 24 &-0.1081 & -0.0462 &24 &0.3935* & -0.1592  \\ 
+& Lo & 97 &-0.1422* & -0.0571 & 97 & -0.2896* & 0.0035& & 26 & -0.2008 & -0.0146 &26 &0.1185 & 0.4560*  \\
+\multirow{2}{*}{AUD}
+& Hi & 101 &0.2539* &0.0125 &101 &0.8728* & 0.0875 &  &33 &0.5099* &0.0682 &33 & 1.1501*& 0.1046  \\ 
+& Lo & 138 &-0.4384* & -0.1711 & 138 & -1.14935* & -0.2714* & &20 &-0.4334 & 0.0563 &20  & -1.7360* & 0.1914  \\
+\multirow{2}{*}{EURJPY}
+& Hi & 95 &-0.0504 &-0.1103 &95 &-0.0128 &-0.2579 & &23 &-0.2932 &-0.4188* &23 &-0.8203 & -1.0195  \\ 
+& Lo & 97 &-0.4746* & -0.1647 &97 &-1.6757* &-0.4595 & &20 &-1.3630* & -0.6976* & 20 & -3.6849* &-1.1999*  \\
+\multirow{2}{*}{EURGBP}
+& Hi & 91 & 0.1544 &0.0696 &91 &0.5176* &0.0979 & &28 &0.2048 &0.1195 &28 & 0.4146 & -0.0645  \\ 
+& Lo & 122 & 0.0451 & 0.0996 &122 &-0.3097* &0.2903* & & 21 &-0.07523 &-0.0780 &21 & -0.4120 &
+-0.1290  \\
+\multirow{2}{*}{EURCHF}
+& Hi & 77 & 0.1183* & 0.0690 &77 & 0.3444* & 0.0749& & 	31 & 0.1482* &0.0739  &31 & 0.4993* 
+& 0.1816*  \\ 
+& Lo & 94 &-0.0344 & 0.0194 & 94 & -0.1596 &0.1952* & &24 & -0.3626* &-0.1483 & 24 & -0.7573* 
+& 0.0875  \\
+\hline
+\label{tabref:RR1}
+\end{tabular}
+\begin{tablenotes}
+\small 
+\item Where Hi means extreme high in the risk-reversal skew and Lo means extreme low in the risk reversal skew; extreme is calculated as being equal or above the 95th or 99th percentile or equal or below the 5th or 1st percentile respectively; CARW is the cumulative abnormal return for the whole window, before and after the extreme event, and CARA is the cumulative abnormal return for the period after the event, which is the event day and the window; abnormal return is anything that is different from zero; the asterisk denotes significantly different from zero, where statistical significance means that more than 95\% of 1000 random bootstraps from the events are above or below zero respectively.
+\end{tablenotes}
+\end{threeparttable}  
+\end{table}
+\end{landscape}
+%This is a section on the methodology that I am writing now (while I remember) so that I can place where necessary.  
+%================Event Study Method=====================================
+% The dates when the risk reversal skew is at or above the ninety fifth (ninetieth) or at or below the first (tenth percentile are identified and a matrix of returns is created with the rows being equal to plus and minus the size of the event window and the date that the extreme was recorded being equal to zero.
+The number of columns is equal to the number of events which, given that we are looking at the extreme one or five percent, will be related to the sample size.
+The average of the returns for each event day are then recorded.
+As such, with an event window of four days, there is an average return four days before the event window and for three days before the event window, through to four days after the event window.
+To get an idea of the precision of this estimate, the returns from each event day are sampled (with replacement) one thousand times and the mean is calculated.
+The dispersion of the resulting means provide the confidence intervals for the estimate of the mean returns for each event day that has been made.
+
+The results do not appear to be a function of the percentile that is used to measure the extreme event or the size of the window that is selected.
+Additional cases were run using percentiles at 80\% and 99.5\%.
+For the event window, figures \ref{fig:ES1} and \ref{fig:ES2} show the cumulative abnormal returns for the major exchanges rates relative to the US dollar across a 16 day event window.
+Figure \ref{fig:ES1} shows the extreme highs for as a $99^{th}$ percentile and figure \ref{fig:ES2} shows the extreme low for as a $10^{th}$ percentile.
+The figures reveal the mean cumulative return (in solid red) and the 95\% confidence intervals (dashed blue).
+The only case where there seems to be any evidence of a reversal is for the extreme low risk-reversals on the Canadian dollar.
+
+\begin{figure}[h!]
+\graphicspath{{../Figures/}}
+\centering
+\caption{Event Study:  Extreme (High 99th percentile) Risk Reversal Skew and 16 day event window}
+\label{fig:ES1}
+\includegraphics[scale=0.8]{RRCum16}
+\end{figure}
+
+\begin{figure}[h!]
+\graphicspath{{../Figures/}}
+\centering
+\caption{Event Study:  Extreme (Low - 10th percentile) Risk Reversal Skew and 16 day event window}
+\label{fig:ES2}
+\includegraphics[scale=0.8]{RRCum16a}
+\end{figure}
+
+However, it does indicate that the evolution of exchange rates are not a random walk.
+There is evidence here that prices gradually respond to information, as would be the case if decision-making is following a \emph{conservatism} heuristic or if there were institutional features.
+The positive (or negative) expectations that are evident in the risk reversal skew do not seem to represent a deviation or a mis-pricing that will be reversed.
+Rather they seem to be part of the price-discovery process.
+
+\begin{landscape}
+% multirow package is required.   
+%pdflandscape is required to facilitate the landscape position. 
+\begin{table}[ht]
+\begin{threeparttable}
+\caption{Event Study: Cumulative Abnormal Returns and Extreme Speculative Positions}
+\begin{tabular}{llccccccccccccc}	
+ % \hline
+& &\multicolumn{6}{c}{S1 - Net Long per Speculators} & \multicolumn{1}{c}{} & \multicolumn{6}{c}{S2 -Net Long per Open Positions}\\
+ & &\multicolumn{3}{c}{2 week} & \multicolumn{3}{c}{4 week} & \multicolumn{1}{c}{} & \multicolumn{3}{c}{2 week} & \multicolumn{3}{c}{4 week} \\ 
+%\cline{4 - 10} \cline{13 - 19}
+& &\multicolumn{1}{ c}{N} & \multicolumn{1}{c}{CARW} & \multicolumn{1}{c}{CARA} & \multicolumn{1}{ c}{N} & \multicolumn{1}{c}{CARW} & \multicolumn{1}{c}{CARA} & \multicolumn{1}{c}{} & 
+\multicolumn{1}{ c}{N} & \multicolumn{1}{c}{CARW} & \multicolumn{1}{c}{CARA}  
+& \multicolumn{1}{ c}{N} & \multicolumn{1}{c}{CARW} & \multicolumn{1}{c}{CARA}  \\   
+\hline
+\multirow{2}{*}{EUR} 
+& Hi &  27 &  2.9415* & 1.3163* &27 & 3.8059* & 1.7341* &&27 & 1.0974* & 0.2896 &27 & 1.7667*& 0.0419  \\ 
+& Lo & 27 & -2.4258* & -1.2512* & 27 &-3.9235* &-1.9254* & & 27 & -2.5122* & -1.0607* & 27 & -4.4478* & -1.9285* \\
+\multirow{2}{*}{JPY}
+& Hi & 43 & 1.9749* & 0.7286* & 43 & 2.8628* & 0.7909 & & 43 & 2.2463* & 1.0554* &43 & 3.7125* & 1.4443*  \\ 
+& Lo & 43 & -2.2014* & -1.1312* & 43&-3.2488* & -1.5758* & &43 & -2.0605* & -0.9791* &43  & -2.6865* & -0.8318  \\
+\multirow{2}{*}{GBP}
+& Hi & 43 & 0.7102 & -0.2116 &43 & 1.0172* & -0.7902* & &43 & 1.2916* & 0.2370 &43 &1.6912* &0.0126  \\ 
+& Lo & 43 &-0.6979* &0.2902 &43 &-1.5330* &0.1014 & &43 &-1.8993* &-0.1415 &43 &-2.8830* & -0.0768 \\
+\multirow{2}{*}{CHF}
+& Hi & 43 & 1.4400* & -0.0988 & 43 & 2.8665* & -0.0993 & & 43 &  2.5260* &  0.5074 & 43 & 3.6060* & 0.2158  \\ 
+& Lo & 43 & -2.1816* &  -0.9770* & 43 &-3.3030* & -0.09798* & & 43 & -1.2605* & -0.3070& 43 &-0.1.6021* & 0.0670  \\
+\multirow{2}{*}{CAD}
+& Hi & 43 & 1.8658* & 1.0195* & 43 & 3.5298* & 1.8319* & & 43 & 2.0160* & 0.8788* &43 & 3.0995* & 0.9921*  \\ 
+& Lo & 43 &-1.3238* & -0.5548* & 43 & -1.9151* & -0.7859* & & 43 & -1.2462* & -0.4505*  &43 & -1.9390*  & -0.6195*  \\
+\hline
+\label{tabref:SP1}
+\end{tabular}
+\begin{tablenotes}
+\small 
+\item Where Hi means extreme high in the sentient index, either S1 which is the measure of net long non-commercial (speculative) positions per number of speculators) or S2 which is the measure of net long non-commercial (speculative) positions per open interest (total outstanding open positions); high and low are above the 95th percentile or below the 5th percentile respectively; CARW is the cumulative abnormal return for the whole window, before and after the extreme event, and CARA is the cumulative abnormal return for the period after the event, which is the event day and the window; abnormal return is anything that is different from zero; the asterisk denotes significantly different from zero, where statistical significance means that more than 95\% of 1000 random bootstraps from the events are above or below zero respectively.
+\end{tablenotes}	
+\end{threeparttable}  
+\end{table}
+\end{landscape}
+
+Table \ref{tabref:SP1} shows the results from the event studies that were carried out using speculative positions as reported to the US regulators.
+The table works in a similar way to Table \ref{tabref:RR1}.
+The first column is the exchange rate, defined as units per US dollar.
+The table is then split into two parts.
+The first part uses the measure S1 which is the net non-commercial (speculative) positions relative to total non-commercial (speculative) positions; the second is S2 as the net non-commercial position relative to open interest or the total outstanding open positions.
+Therefore, the first captures the balance of speculative positions while the second captures the balance of speculative positions relative to the whole market.
+If speculators become more prevalent or dominant in the market, this will show up in S2 but not S1.
+S1 is more sensitive to changes in sentiment and more volatile.
+Each of these sections is broken into the results for a 2 week window and a 4 week window.
+Each of these has the same pattern.
+There is a row for extreme high and extreme low which represents extreme long positions or extreme short positions; the first  column identifies the number of cases; CARW is the cumulative abnormal return for the whole event window; CARA is the cumulative abnormal return for the period after the event to the end of the window.
+
+\begin{figure}[h!]
+\graphicspath{{../Figures/}}
+\centering
+\caption{Event Study:  Extreme (95th percentile), Non-commercial net per OI (S2), 6 week event window}
+\label{fig:ES3}
+\includegraphics[scale=0.8]{FPCum6w}
+\end{figure}
+
+Once again, there is a very strong association between returns and the movement of speculators.
+For the 4-day window and the S1 and S2 measures, every one of the currencies studied, high and low, had a mean cumulative return for the whole of the event window that was in the same direction as sentiment.
+When re-sampled a thousand times, more than ninety five percent of the means calculated were greater than zero in all cases but one (GBP).
+Once again there is evidence here of momentum and of speculator positions following the movement of prices.
+
+However, there is little evidence that these speculators are driving prices to extremes with their extreme behaviour.
+Sixty percent of the S1 cases still show a significant abnormal return in the direction of the extreme 4 weeks after the extreme has been reached; only thirty percent of the cases show a reversal and in only one of those do ninety five percent of the re-sampled means remain above zero.
+For S2 it is even more extreme, with ninety percent of the cases showing a continuation and forty percent of them being significant.
+Again the evidence is consistent with the gradual adjustment of prices to information.
+
+\begin{figure}[h!]
+\graphicspath{{../Figures/}}
+\centering
+\caption{Event Study:  Extreme (90th percentile), Non-commercial net long (S1), 6 week window}
+\label{fig:ES4}
+\includegraphics[scale=0.8]{FPCum6wa}
+\end{figure}
+
+The event windows are stretched to give a clear view of the evolution of exchange rates  around the event and to ensure that results are not a function of the size of the event window.
+Confidence intervals are added.
+Once event objects FX.exc have been created, parameter for exchange rate, extreme measure, quantile, window length and the high or low indicator can be changed for alternative assessments of speculative extremes.
+Adjustments to these parameters show that these results are not a function of the choices presented in the main text.
+They are robust to alternative specifications.
+
+Figure \ref{fig:ES3} shows the performance of the cumulative abnormal returns around a six week event window where the event is the net low (short) speculative positions relative to total activity (open interest) at or below the ($5\%)$ level.
+This means that the speculators are inclined to be short the non-domestic (US dollar) currency.
+While the Euro continues to move lower after the extreme short positions, the other major currencies seem to reveal a period of consolidation that is consistent with the settlement around an equilibrium after a period where information is being absorbed.
+The contrast between the Euro and the other currencies may be associated with the fact that there is a smaller dataset for the Euro due to its more limited lifespan.
+Figure \ref{fig:ES4} shows the abnormal returns around extreme high (90th percentile and long position).
+There is some tentative evidence of continuation for speculative positions in the Canadian dollar and signs of reversal for UK sterling.
+However, the Euro and the Japanese yen move sideways after the event.
+There is no clear pattern after the extreme and therefore no information.
+
+\section{Speculation and foreign exchange returns}
+This paper has sought to measure speculation and to use this measure to discover more about the relationship between speculation, information and asset prices.
+The measurement of speculation takes two forms:  a measurement of the intensity of belief with the risk-reversal skew, and a measurement of speculative positions.
+Neither of these is ideal, in the sense of being a random sample from a population.
+However, the option prices reveal the intensity of feeling through the weight of money in a speculative part (options) of a speculative market (foreign exchange).
+The commitment of traders report records only speculators in the US futures market.
+However, this is an important market, there is no reason to believe that US or speculators in the futures market are systematically different from those in the rest of the world and presumably if there were a major difference there would be an arbitrage opportunity between the two markets.
+
+ there is strong evidence that speculation affects prices and that this effect is gradual rather than instantaneous.
+This suggests a market inefficiency that would reward momentum-based trading.
+It is no wonder therefore that \citep{TaylorTechnical}, reporting a survey conducted by the Bank of England, found that of the Chief foreign exchange dealers contacted in London in November 1988, 90\% of them attached some weight to \emph{technical analysis}, a method that is fundamentally based on finding trends in asset prices.
+The survey found that the importance of technical analysis became greater at shorter horizons.
+This has not disappeared.
+The importance of momentum trading and technical analysis has not waned with time.
+Neely and Wller  show that gains from technical trading rules still generate excess returns and attribute this to behavioural biases \citep{NeelyWeller}.
+It seems that speculation forms part of the process of price discovery, but this process is not instantaneous as would be expected in the purest form of the EMH.
+
+The notion that speculation is at least partially informed is supported by the fact that there is little evidence from the event studies conducted here that speculation causes overshooting or bubbles.
+When speculation is extreme, whether measured by intensity of belief or weight of speculative positions, future returns are as likely to be positive as negative.
+There is no informational content as there would be if speculation was pure noise.
+In the simulation of the De Long model, extremes are more likely to be followed by reversals.
+In the foreign exchange market, that is not the case.
+There are some large price crashes in the dataset, as shown by the descriptive statistics of the returns.
+For example  in Table \ref{tabref:DS1} it is seen that on one day there is a 10.65\% appreciation of the Japanese yen and Table \ref{tabref:DS2} shows a week when the Canadian dollar falls by 10\%.
+However, the measures of speculative sentiment or the weight of speculation are not particularly high in either of these cases.
+Indeed, an analysis of the top 1\% of positive and negative foreign exchange returns in the dataset reveals that sentiment and positions for this group were very close to the average of the whole set.
+This would be consistent with a model where extreme price movement was a function of new information rather than the structure of existing positions.
+%expand.  Could put data in appendix.
+
+The evidence here, that speculation is associated with serial correlation in the returns but cannot be used in a systematic way to determine reversals, has important implications for speculative activity.
+It means that the largest returns will be to those that identify the trend at the earliest opportunity.
+This means that speculators must be prepared to act as swiftly as possible, probably when there is a high level of uncertainty and a lack of information and where any scrap of additional news is highly valued.
+The risk that actions by other speculators, most notable by their effect on price, will be construed to contain information exacerbates the likelihood of feedback effects.
+The need to be able to take action before others has reached the most extreme form in the activity of \emph{high frequency trading} (HFT), where decision times are measured in nanoseconds and there is advantage in gaining spacial proximity to exchanges.
+HFT is all about getting  a larger share of the inefficiency in pricing information by trading or investing closer to where the information is released.
+
+These findings have implications for speculators, economic modellers and policy makers.
+Speculators may jump on the trend, but they do not know by looking at how popular that trend is when to get off.
+They need other signs to tell them that the trend has reached its conclusion or equilibrium has been achieved.
+For the modellers, there is confirmation of the need  for feedback effects and evidence that reversals cannot be systematically identified from high levels of speculative positions.
+It may suggest another trigger for bursting bubbles; it may, for example, suggest that this reversal has to be modelled as stochastic element or pure shock.
+For policy makers, there is no escape from the requirement to look at economic fundamentals when deciding whether bubbles are building in asset markets.
+The presence of speculators is no guarantee that there is excess.
+The speculators may be just part of the adjustment to new information.
+However, the combination of high level of speculation and divergence from fundamentals is likely to create the greatest risk.
 

### 2015-06-10 Knitr/NewPaper.Rnw

@@
   Speculation is also an important part of the argument for why prices should tend towards that consistent with \emph{rational expectations}.
-\citet{muth1961} uses commodity speculation to argue for model-consistent expectation formation.
+\citet{muth1961Rational} uses commodity speculation to argue for model-consistent expectation formation.
 In this case, the speculators are informed and tend to reduce the volatility of prices.
@@
 % this needs to be completed.
+
 \subsection{Measuring the intensity of speculative sentiment}
@@
 Frankel and Froot found that the median forecast from the MMS survey exhibited trend-following behaviour at the one to three week horizon while there is mean-reversion beyond that point.
-Frankel and Froot also assess the 'Financial Report' that has been produced by \emph{The Economist} magazine and \emph{Currency Forecasters Digest} and find a wide dispersion of opinion and express some concern about the aggregation of such divergent forecasters \citep{FrootFrankelFDB} and \citep{FrankelFroot1987} \label{fnsurvey}.}    This is the relative price of currency put and call options for the same strike price.
-The \emph{Put-Call-Parity} in the Black-Scholes option pricing model assumes that these will be the same \citep{Black1973Options}.
-While the Black-Scholes model is the basis for valuing options, where the price of a non-dividend paying stock trades at a discount to the forward rate equal to the risk free rate, the forward rate in a currency option is determined by the relative interest rate differential between the two currencies.
-The Garman-Kohlhagan adjustment that accounts for the difference in interest payments on the two currencies is used.
-The spot discount is determined therefore by the relative interest rate differential for the appropriate maturity \citep{GarmanKohlhagan}.
-This model allows the market practice of pricing options from the volatility implied by the model.
-In other words, the Garman-Kohlhagan is used to extract the volatility that is implied by a specific option price given the strike and other parameters.
-
-Determining the fundamental value of exchange rates has proved to be extremely difficult.
-As PPP is, at best, a medium-term phenomenon and given the finding by Meese and Rogoff about the superiority of the random walk relative to other economic models and accepting for the moment the EMH, it will be assumed that the forward rate is the best estimate of the future value of the exchange rate.
-Any forecast which deviates from the forward rate will in this case be counted as a misperception and the further away from the random walk the greater the misperception.
-\footnote{The expected  exchange rate should be equal on average to the forward rate once any risk premia has been take into account.
-Though it will be seen that UIP and the assumption that the future exchange rate is on average equal to the interest rate differential does not hold, the fact that daily data is being used here and the attention is on the most extreme divergences from UIP suggests that this will not greatly affect the results.}
-
-This study uses a unique dataset of \emph{risk-reversals} for eleven major currency pairs.
-This is the only period for which the data currently exist.
-The risk-reversal data is very difficult to obtain in an over-the-counter market like foreign exchange where there is no centralised exchange to record and collect information, generally being the property of quoting institutions.
-The risk-reversals are the average prices quoted at 10:00 GMT by major banks on the Thomson-Reuters composite option pages (OPT1)  for the period between January 2 1996 and September 6 2002.
-The risk-reversals are the spread of implied volatility for a 1 month 25 delta at-the-money call option relative to the equivalent put.
-A call option is the right to buy and a put option is the right to sell.
-The delta of an option measures the effect of a change in the spot price on the value of the option.
-This can range from zero for deeply out-of-the-money options to unity for deeply in the money strikes.
-The at-the-money delta is 50 as there is a 50:50 chance of the option being exercised.
-The benchmark for option risk-reversals is 25 delta.
-This is a measure of how far the option is out of the money or how far the strike is from the forward rate.
-
-If the exchange rate trajectory is expected to follow a random walk with drift towards the 1 month forward, implied volatility on the put and the call for the same delta should be equal.
-Any deviation from this equivalence is an indication that a deviation from the random walk with drift is implied in the market price of options.\footnote{If the risk-reversal for a 1-month USD/CAD is quoted at 0.15-0.28\% and the implied volatility is 8.50\%, the market-maker would be willing to buy the 1-month 25 delta USD-put-CAD-call at 8.65\% and sell the USD-call-CAD-put at 8.50\%.
-The dealer pays 0.15\%.
-Alternatively, the dealer would sell a 25 delta call at 8.78\% and buy the USD-call-CAD-put at 8.50\%, earning 0.28\% \citep{Global}.
-}  Therefore, the relationship between implied volatility for calls and implied volatilities for puts for the same delta and maturity can be taken as an indication of how much more expensive calls are relative to the equivalent put or as the market bias towards puts over calls or, in this case, as an indication of market sentiment.
-The larger the absolute level of the risk-reversal skew, the greater the relative price and, in equilibrium, the greater the belief, expressed through the action of market participants or the inventory of market-makers, that there will be a deviation from the forward rate.
-If extreme deviations or speculations are noise they should presage a sharp reversal.
-If speculation is informed, this is part of the price discovery process and extremes may be followed by calm conditions around a new equilibrium.
-
-% The table may be pushed onwards.  Use t (i think). 
-\begin{landscape}
-\begin{table}[t]
-\begin{threeparttable}
-\caption{Descriptive Statistics for Daily Exchange Rate Returns: 2 January 1996 to 6 September 2002}
-\begin{tabular}{l | r r  r r r r r r r r}
-  \hline
- Statistic &  AUD & EURGBP & EURSEK & GBP & CHF & EURCHF & EURJPY & EUR & CAD & JPY \\
-  \hline
-Number & 1685 & 1675 & 1716 & 1667 & 1674 & 1674 & 1675 & 1668 & 1683 & 1675  \\
-Mean & -0.0174 & -0.0187 & 0.0011 & 0.0025 & 0.0120 & -0.0039 & -0.0072 & -0.0172 & 0.0084 & 0.0076\\
-Median &  0.0000 & -0.0298 & 0.0000& 0.0000 & 0.0114 & 0.0000 & 0.0000 & -0.0253 & 0.0000 & 0.0188\\
-Standard deviation  & 0.7241 & 0.5542 & 0.0438 & 0.0489 & 0.6819 & 0.0277 & 0.8416 & 0.6338 & 0.0363 & 0.8170\\
-Max & 5.8301 & 3.0543 & 2.1633 & 2.2573 & 3.5086 & 1.6477 & 4.0879 & 2.4136 & 2.2170 & 5.1788 \\
-Min & -3.4666 & -2.2858 & -2.2900 & -2.2701 & -3.6525 & -1.8466 & -8.2848 & -2.2901 & -2.9707 & -10.6573 \\
-Skewness  & 0.1206* & 0.2923* & 0.0432 & -0.0452 & -0.1423* & -0.4046* & -0.7388* & 0.1738* & -0.1095 & -1.7237*\\
-Kurtosis & 4.17* & 1.88 & 2.43* & 1.48 & 1.96 & 5.84* & 7.08* & 1.12 & 5.11* & 21.52*\\
-\hline
- \label{tabref:DS1}
-\end{tabular}
-\begin{tablenotes}
-\small
-\item Continuously compounded returns (multiplied by 100) for period 2 January 1996 to  September 2002 for US dollar per unit of Australian dollar (AUD), UK sterling per unit of Euro (EURGBP), Swedish Crown per unit of EUR (EURSEK); US dollar per unit of UK sterling (GBP);  Swiss franc per unit of US dollar (CHF); Swiss franc per unit of Euro (EURCHF); US dollars per unit of Euro (EUR); Canadian dollars per unit of US dollar (CAD); Japanese yen per unit of US dollars (JPY).
-The asterisk (*) means that the estimate is significantly different from zero at the 5\% level using t-statistics for the mean and the standard error of skew and standard error of excess-kurtosis respectively.
-\end{tablenotes}
-\end{threeparttable}
-\end{table}
-\end{landscape}
-
-Table \ref{tabref:DS1} shows descriptive statistics for the exchange rate data used in the study.
-There are daily data for the period between January 1996 and September 2002.
-This is the period for which risk-reversal data are available.
-The series are of sightly different lengths as there are some gaps in the risk reversal and exchange rate data as a result of holidays and market closures.
-From the raw data continuously compounded returns are calculated and multiplied by 100.
-The first thing of note is that the mean return does not appear to be statistically different from zero.
-This is what would be expected if, as has been identified in many studies, exchange rates follow a random walk.
-This apparent stability hides some significant changes that are taking place below the surface.
-For example, the Japanese yen records an appreciation of 10.66\% on one day.
-Most of the exchange rate series appear to be skewed in some way and there is also evidence of kurtosis in many cases.
-An asterisk is used to denote statistical significance at the level of 95\% confidence.
-A t-test is used for the mean; a \emph{standard error of skew} and a \emph{standard error of kurtosis} are used for the skew and kurtosis respectively.
-These are calculated as $SES = \sqrt{\frac{6n(n-1)}{(n-2)(N+1)(N+3)}}$ and $SEK = \sqrt{\frac{n^2-1}{(n-3)(N+5)}}$, with the assumption that this estimate is distributed normally, the skew or kurtosis is considered significant if it is more than two standard errors from zero.
-
-\begin{landscape}
-\begin{table}[t!]
-\begin{threeparttable}
-\caption{Descriptive Statistics for Risk Reversals: 2 January 1996 to 6 September 2002}
-\begin{tabular}{l | r r  r r r r r r r r}
-  \hline
- Statistic &  AUD & EURGBP & EURSEK & GBP & CHF & EURCHF & EURJPY & EUR & CAD & JPY \\
- \hline
-Number & 1702 & 1697 & 1719 & 1691 & 1374 & 1339 & 1697 & 1692 & 1701 & 1697 \\ 
-  Mean & -0.37 & 0.15 & 0.71 & -0.08 & -0.28 & -0.38 & -0.48 & 0.19 & 0.12 & -0.74 \\ 
-  Median & -0.30 & 0.10 & 0.80 & -0.10 & -0.15 & -0.30 & -0.10 & 0.10 & 0.15 & -0.55 \\ 
-  Standard Deviation & 0.38 & 0.37 & 0.51 & 0.33 & 0.49 & 0.36 & 1.02 & 0.49 & 0.25 & 1.02 \\ 
- Max & 0.45 & 1.72 & 1.90 & 1.00 & 1.05 & 0.50 & 1.70 & 2.05 & 1.45 & 2.05 \\ 
- Min & -1.70 & -0.80 & -0.30 & -1.35 & -2.05 & -1.80 & -4.03 & -1.15 & -0.65 & -4.00 \\ 
- Skewness & -0.50* & 1.27* & -0.23* & 0.33* & -0.79* & -0.58* & -0.93* & 0.68* & 0.23* & -0.51* \\ 
-  Kurtosis & 0.02 & 2.29* & -1.67 & 0.72 & 0.69 & 0.19 & 0.15 & 0.96 & 1.83 & -0.03 \\ 
-      \hline
- \label{tabref:DS2}
-\end{tabular}
-\begin{tablenotes}
-\small
-\item Daily risk reversal skew.
-Risk reversal is the premium of implied volatility for calls relative to the equivalent put (equal distance to the strike price, 25 delta in this case).
-A figure above zero means that calls are more expensive (in volatility and price terms) than puts.
-A figure below zero means that puts are more expensive.
-The asterisk (*) means that the estimate is significant at the 5\% level using t-statistics for the mean and the standard error of skew and standard error of excess-kurtosis respectively.
-These are calculated as $SES = \sqrt{\frac{6n(n-1)}{(n-2)(N+1)(N+3)}}$ and $SEK = \sqrt{\frac{n^2-1}{(n-3)(N+5)}}$, with the assumption that this estimate is distributed normally, the skew or kurtosis is considered significant if it is more than two standard errors from zero.
-\end{tablenotes}
-\end{threeparttable}
-\end{table}
-\end{landscape}
-
-Table \ref{tabref:DS2} gives the descriptive statistics for the risk-reversal data.
-For a neutral risk-reversal, a reading of zero, there would be no price preference for calls or puts at the similar delta.
-The figures show that, in contrast to the average returns, there are often average skews that are different from zero.
-The most significant of these are the positive bias for EUR-GBP and a negative bias for USD-JPY.
-The latter may be related to hedging activity by Japanese exporters.
-While this could suggest that implied volatility is an imperfect measure of speculative activity, note that extreme readings are based on quantiles of the distribution and therefore an underlying hedging-induced shift in the mean of the distribution should not affect the results.
-In other words, if the extreme is denoted by the 90th percentile, it does not matter if the mean is zero and this comes at 0.8 or if the mean is 0.2 and this comes at 1.0.
-\subsection{Measuring the intensity of speculative sentiment}
-The measurement of speculative intensity must identify how strongly speculators believe that the market will move in a particular direction.
-In the absence of a daily survey of speculators, this study uses \emph{risk-reversal skew} as an objective measure of speculative sentiment rather than surveys of opinion.\footnote{There are a number of surveys of investor opinion that are regularly conducted by financial service information companies.
-However, these are disparate and, in many cases, unsystematic, and it is not clear whether they measure speculative sentiment or investor opinion more generally.
-The Wall Street Journal runs an annual survey of investor opinion, but this is focused mostly on US equities.
-The \emph{Economist} magazine also solicits opinion, but in a relatively unstructured and \emph{ad hoc} fashion.
-In the foreign exchange market, there is a broad set of market forecasts that was collected by the financial services company \emph{MMS International} (formerly Money Market Services) which polled a broad range of financial market participants on key exchange rates for one and four week forecast horizons.
-Frankel and Froot found that the median forecast from the MMS survey exhibited trend-following behaviour at the one to three week horizon while there is mean-reversion beyond that point.
-Frankel and Froot also assess the 'Financial Report' that has been produced by \emph{The Economist} magazine and \emph{Currency Forecasters Digest} and find a wide dispersion of opinion and express some concern about the aggregation of such divergent forecasters \citep{FrootFrankelFDB} and \citep{FrankelFroot1987} \label{fnsurvey}.}    This is the relative price of currency put and call options for the same strike price.
+Frankel and Froot also assess the 'Financial Report' that has been produced by \emph{The Economist} magazine and \emph{Currency Forecasters Digest} and find a wide dispersion of opinion and express some concern about the aggregation of such divergent forecasters \citep{FrootFrankelFDB} and \citep{FrankelFroot1987}.}    This is the relative price of currency put and call options for the same strike price.
 The \emph{Put-Call-Parity} in the Black-Scholes option pricing model assumes that these will be the same \citep{Black1973Options}.

### 2015-06-15 Knitr/NewPaper.Rnw
Updated most of the paper. The final couple of sections need to be completed. The whole paper could be made more dramatic with the key event charts mroe towards the front.

@@
 \usepackage{pdflscape}
+\usepackage{appendix}
 \usepackage[flushleft]{threeparttable}
@@
 Towards the other end, Milton Friedman argued that speculation was a stabilising force, providing liquidity and helping to ensure that markets swiftly find equilibrium.
-In arguing in favour of flexible exchange rates, Friedman asserted that successful speculators would be those who managed to correctly buy when prices  were below fundamental value and sell above that price \citet{FriedmanPositive}.
+In arguing in favour of flexible exchange rates,  \citet{FriedmanPositive} asserted that successful speculators would be those who managed to correctly buy when prices  were below fundamental value and sell above that price.
 Therefore, Friedmand argued, the speculative process would favour those able to access or utilise information over the uninformed and, through a Darwinian process, informed speculation would dominate.
@@
 The latest survey taken by the Bank for International Settlements (BIS) in April 2013 estimated average daily turnover at \$5.4trn with average daily spot transactions of \$2.0trn \citet{BISFX2013}.
-The cost of trading in this markket, when measurd in terms of the size of the bid-ask spread, is extremely low.
+The cost of trading in this market, when measurd in terms of the size of the bid-ask spread, is extremely low.
 An analysis of the bid-ask spreads for USD-JPY, GBP-USD and EUR-USD for the period January 2001 to 2005 and finds that the average spread for each is 0.342\%, 0.305\% and 0.326\% respectively.
@@
 One key element of the microstructure literature involves distinguishing between uninformed or noise-traders and those who have infomration.
-The Kyle model has market-makers reciving orders from the informed and the uninformed.
+The \citet{Kyle1985Continuous} model has market-makers reciving orders from the informed and the uninformed.
 On the assumption that the uniformed arrive in a random fashion without any reason to trade, the weight of orders implies something about the activity of the informed and causes prices to change in the process of price discovery.
 The bid-ask spread is the payment that the market-maker recieves in return for taking the risk of meeting an informed trader.
-Ultimately, the uninformed traders make payments to the informed traders with the mark-makers taking a cut for the risk and the privision of exchange \citet{Kyle1985Continuous} and \citet{glosten1985bid}.
-Here the speculators may be informed or uninformed.
+Ultimately, the uninformed traders make payments to the informed traders with the mark-makers taking a cut for the risk and the privision of exchange.
+See also \citet{glosten1985bid}.
 
@@
 Nonetheless, liquidity tends to refer to institutional explanations for trading without reference to fundamental value.
-This may mean that they just execute customer orders at the current price or it may be that they buy and sell with reference to a benchmark index or to fulfil a hedging strategy.
+This may mean that economic agents just execute customer orders at the current price or it may be that they buy and sell with reference to a benchmark index or to fulfil a hedging strategy.
 
@@
 Speulative noise-traders can therefore create a gap between the price and the fundamtental value that will encourage informed traders to pay the cost of acquiring information.
-The divergence between this price and fundametnal value is essential to encourage te process of price discovery.
+The divergence between this price and fundametnal value is thereby essential to encourage te process of price discovery.
 %
-Fisher Black suggested that this combination was essential to the smooth function of financial markets.
-It is ``the arbitrary element in expectations'' \citet[p.529]{BlackNoise}.
+ \citet[p.529]{BlackNoise} suggested that this combination was essential to the smooth function of financial markets.
+It is ``the arbitrary element in expectations''.
 Therefore, noise makes financial markets possible but imperfect: the more noise-traders that there are, the more liquidity and the easier it is to trade; the more noise-traders, the higher the level of inefficiency in the market and the more likely that price will be driven from fundamental value.
@@
   Speculation is also an important part of the argument for why prices should tend towards that consistent with \emph{rational expectations}.
-\citet{muth1961Rational} uses commodity speculation to argue for model-consistent expectation formation.
+\citet{muth1961Rational} uses commodity speculation as the means to drive prices towards the \emph{rationa-expectations-equilibrium} (RER).
 In this case, the speculators are informed and tend to reduce the volatility of prices.
+This is a reiteration of the Friedman idea of speuclation but runs immediately into the \citet{milgrom1982information} problem of finding counterparties willing to make the other side of the transaction.
 
@@
 
-\subsection*{The problem of uninformed noise-traders}
-The traditional Keynesian view of speculation asserted that it would be an additional noise that would make it even harder for informed-traders to overcome uncertainty and apply fundamental information about long-term value.
-This view finds some supported in Shiller's observation that stock prices are more volatile than the underlying earnings information on which valuation should be based \citet{ShillerVol}.
-In addition, \citet{Delong1990noise} and \citet{shleifer1997limits} show that constrained horizons can cause noise-trading to limit the ability of informed traders to take advantage of profitable opportunities for fear of further divergence from fundamental value before realignment is achieved.
 %It would be good to have some more modern ideas here.  Maybe they are just the empirical results. 
@@
 \subsection{Why this is important}
+The traditional Keynesian view of speculation asserted that it would be an additional noise that would make it even harder for informed-traders to overcome uncertainty and apply fundamental information about long-term value.
+This view finds some supported in Shiller's observation that stock prices are more volatile than the underlying earnings information on which valuation should be based \citet{ShillerVol}.
+In addition, \citet{Delong1990noise} and \citet{shleifer1997limits} show that constrained horizons can cause noise-trading to limit the ability of informed traders to take advantage of profitable opportunities for fear of further divergence from fundamental value before realignment is achieved.
+
 The nature of speculation and the role of noise-traders has become increasingly important since the global financial crisis (GFC).
@@
 This will be use to gain a fuller understanding of speculative behaviour.
-This will involve three steps:  first,assess the price implication of the noise-trader model;  second, measure speculative activity in the foreign exchange market; finally, assess whether extreme speculative activity in the foreign exchange market is consistent with the noise-trader model.
+This will involve three steps:  first, assess the price implication of the noise-trader model;  second, measure speculative activity in the foreign exchange market; finally, assess whether extreme speculative activity in the foreign exchange market is consistent with the noise-trader model.
 If the speculative activity that is seen in the foreign exchange market produces results that are like the noise-trader model, it can be said that speculation is noise trading as Keynes asserted, if not it may be more informed as suggested by Friedman.
@@
 
-The noise-trader model used in this study is from De Long, Shleifer, Summers and Waldeman \citep{Delong1990noise}.
-This is a derivative of the Samuelson overlapping generations model  with two-period agents making a portfolio choice \citep{Samuelson1958}.
+The \citep{Delong1990noise} noise-trader model used in this study.
+This is a derivative of the \citep{Samuelson1958} overlapping generations model  with two-period agents making a portfolio choice.
 There are two assets:  a safe asset $(s)$ which pays a set dividend $(d)$ and is in perfectly elastic supply; the risky asset $(u)$ also pays the same dividend but is not in elastic supply.
@@
 \begin{equation}
-p_t = 1 + \frac{\mu(\rho_t - \rho^*)}{1 + r} + \frac{ \mu \rho^*}{r} - \frac{2 \gamma \mu^2 \sigma_{\rho}^2}{r(1 + r)^2}
+p_t = 1 + \frac{\mu(\rho_t - \rho^*)}{1 + r} + \frac{ \mu \rho^*}{r} - \frac{2 \gamma \mu^2 \sigma_{\rho}^2}{r(1 + r^2)}
 \label{eq:NP2}
@@
 The reduction in utility comes mainly from the fact that they are not compensated for taking risk that they are not aware of.
-This is consistent with the type of over-confidence that has been found in day-traders \citet{odean1998volume}
+This is consistent with the type of over-confidence that has been found by \citet{odean1998volume}in day-traders.
 
 \subsection{The implications of the noise-trader model}
-The aim of this study is to test the implications of the noise-trader model.
-There are two parameters from the noise-trader price Equation \ref{eq:NP2} that can be used to measure the amount of speculation:  the price misperception $(\rho)$ and the weight of noise-traders in the market $(\mu)$.
-The aim of this section is to  identify the returns that are implied by the noise-trader model when either the intensity of the misperception or the weight of noise-traders is increased to extreme levels so that this can be used as a benchmark against which foreign exchange market activity can be judged.
+There are two parameters from the noise-trader price equation \ref{eq:NP2} that can be used to measure the amount of speculation:  the price misperception $(\rho)$ and the weight of noise-traders in the market $(\mu)$.
+the aim of this section is to  identify the returns that are implied by the noise-trader model when either the intensity of the misperception or the weight of noise-traders is increased to extreme levels so that this can be used as a benchmark against which foreign exchange market activity can be judged.
 
@@
 These extremes are the \emph{event} for an event study.
-The package \emph{eventstudies} \citep{eventstudies} aligns the defined events (in this case the extreme noise-trader activity) so that descriptive statistics and other analysis can be carried out on the period after the event has occurred.
+The R package \emph{eventstudies} createed by \citet{eventstudies} is used for the data manipulation.
 Descriptive statistics for the whole series of returns and for those when misperception or weight of noise-traders are presented in Table \ref{tabref:MC1}.
@@
 \caption{Results of the Monte Carlo Simulation of the Noise-Trader Model}
-\begin{tabular}{p{6cm} | r r r r}
+\begin{tabular}{p{4cm} | r r r r}
  \label{tabref:MC1}
@@
 The measurement of speculative intensity must identify how strongly speculators believe that the market will move in a particular direction.
-In the absence of a daily survey of speculators, this study uses \emph{risk-reversal skew} as an objective measure of speculative sentiment rather than surveys of opinion.\footnote{There are a number of surveys of investor opinion that are regularly conducted by financial service information companies.
+In the absence of a daily survey of speculators, this study uses \emph{risk-reversal skew} as an objective measure of speculative sentiment rather than surveys of opinion.%\footnote{There are a number of surveys of investor opinion that are regularly conducted by financial service information companies.
 However, these are disparate and, in many cases, unsystematic, and it is not clear whether they measure speculative sentiment or investor opinion more generally.
@@
 Frankel and Froot found that the median forecast from the MMS survey exhibited trend-following behaviour at the one to three week horizon while there is mean-reversion beyond that point.
-Frankel and Froot also assess the 'Financial Report' that has been produced by \emph{The Economist} magazine and \emph{Currency Forecasters Digest} and find a wide dispersion of opinion and express some concern about the aggregation of such divergent forecasters \citep{FrootFrankelFDB} and \citep{FrankelFroot1987}.}    This is the relative price of currency put and call options for the same strike price.
+Frankel and Froot also assess the 'Financial Report' that has been produced by \emph{The Economist} magazine and \emph{Currency Forecasters Digest} and find a wide dispersion of opinion and express some concern about the aggregation of such divergent forecasters \citep{FrootFrankelFDB} and \citep{FrankelFroot1987}.}
+This is the relative price of currency put and call options for the same strike price.
 The \emph{Put-Call-Parity} in the Black-Scholes option pricing model assumes that these will be the same \citep{Black1973Options}.
 While the Black-Scholes model is the basis for valuing options, where the price of a non-dividend paying stock trades at a discount to the forward rate equal to the risk free rate, the forward rate in a currency option is determined by the relative interest rate differential between the two currencies.
-The Garman-Kohlhagan adjustment that accounts for the difference in interest payments on the two currencies is used.
-The spot discount is determined therefore by the relative interest rate differential for the appropriate maturity \citep{GarmanKohlhagan}.
+The \citep{GarmanKohlhagan} adjustment that accounts for the difference in interest payments on the two currencies is used.
+The spot discount is determined therefore by the relative interest rate differential for the appropriate maturity.
 This model allows the market practice of pricing options from the volatility implied by the model.
 In other words, the Garman-Kohlhagan is used to extract the volatility that is implied by a specific option price given the strike and other parameters.
-
-Determining the fundamental value of exchange rates has proved to be extremely difficult.
-As PPP is, at best, a medium-term phenomenon and given the finding by Meese and Rogoff about the superiority of the random walk relative to other economic models and accepting for the moment the EMH, it will be assumed that the forward rate is the best estimate of the future value of the exchange rate.
-Any forecast which deviates from the forward rate will in this case be counted as a misperception and the further away from the random walk the greater the misperception.
-\footnote{The expected  exchange rate should be equal on average to the forward rate once any risk premia has been take into account.
-Though it will be seen that UIP and the assumption that the future exchange rate is on average equal to the interest rate differential does not hold, the fact that daily data is being used here and the attention is on the most extreme divergences from UIP suggests that this will not greatly affect the results.}
 
@@
 % The table may be pushed onwards.  Use t (i think).
-\begin{landscape}
-\begin{table}[t]
-\begin{threeparttable}
-\caption{Descriptive Statistics for Daily Exchange Rate Returns: 2 January 1996 to 6 September 2002}
-\begin{tabular}{l | r r  r r r r r r r r}
-  \hline
- Statistic &  AUD & EURGBP & EURSEK & GBP & CHF & EURCHF & EURJPY & EUR & CAD & JPY \\
-  \hline
-Number & 1685 & 1675 & 1716 & 1667 & 1674 & 1674 & 1675 & 1668 & 1683 & 1675  \\
-Mean & -0.0174 & -0.0187 & 0.0011 & 0.0025 & 0.0120 & -0.0039 & -0.0072 & -0.0172 & 0.0084 & 0.0076\\
-Median &  0.0000 & -0.0298 & 0.0000& 0.0000 & 0.0114 & 0.0000 & 0.0000 & -0.0253 & 0.0000 & 0.0188\\
-Standard deviation  & 0.7241 & 0.5542 & 0.0438 & 0.0489 & 0.6819 & 0.0277 & 0.8416 & 0.6338 & 0.0363 & 0.8170\\
-Max & 5.8301 & 3.0543 & 2.1633 & 2.2573 & 3.5086 & 1.6477 & 4.0879 & 2.4136 & 2.2170 & 5.1788 \\
-Min & -3.4666 & -2.2858 & -2.2900 & -2.2701 & -3.6525 & -1.8466 & -8.2848 & -2.2901 & -2.9707 & -10.6573 \\
-Skewness  & 0.1206* & 0.2923* & 0.0432 & -0.0452 & -0.1423* & -0.4046* & -0.7388* & 0.1738* & -0.1095 & -1.7237*\\
-Kurtosis & 4.17* & 1.88 & 2.43* & 1.48 & 1.96 & 5.84* & 7.08* & 1.12 & 5.11* & 21.52*\\
-\hline
- \label{tabref:DS1}
-\end{tabular}
-\begin{tablenotes}
-\small
-\item Continuously compounded returns (multiplied by 100) for period 2 January 1996 to  September 2002 for US dollar per unit of Australian dollar (AUD), UK sterling per unit of Euro (EURGBP), Swedish Crown per unit of EUR (EURSEK); US dollar per unit of UK sterling (GBP);  Swiss franc per unit of US dollar (CHF); Swiss franc per unit of Euro (EURCHF); US dollars per unit of Euro (EUR); Canadian dollars per unit of US dollar (CAD); Japanese yen per unit of US dollars (JPY).
-The asterisk (*) means that the estimate is significantly different from zero at the 5\% level using t-statistics for the mean and the standard error of skew and standard error of excess-kurtosis respectively.
-\end{tablenotes}
-\end{threeparttable}
-\end{table}
-\end{landscape}
 
-Table \ref{tabref:DS1} shows descriptive statistics for the exchange rate data used in the study.
+Table \ref{tabref:DS1} in Appendix \ref{apref:1} shows descriptive statistics for the exchange rate data used in the study.
 There are daily data for the period between January 1996 and September 2002.
@@
 The latter may be related to hedging activity by Japanese exporters.
-While this could suggest that implied volatility is an imperfect measure of speculative activity, note that extreme readings are based on quantiles of the distribution and therefore an underlying hedging-induced shift in the mean of the distribution should not affect the results.
-In other words, if the extreme is denoted by the 90th percentile, it does not matter if the mean is zero and this comes at 0.8 or if the mean is 0.2 and this comes at 1.0.
+While this could suggest that implied volatility is an imperfect measure of speculative activity, note that extreme readings are based on quantiles of the distribution and therefore a discrete adjustment of this distribution to reflect the hedging power should not change the measurement of extremes.
 
@@
 Non-commercial accounts are generally considered to be speculators.
-See \citep{FuturesSanders} and \citep{FuturesWang} for examples of previous use of this data in this fashion.
+\citep{FuturesSanders} and \citep{FuturesWang} have examples of previous use of this data in this fashion.
 
@@
 In other words, the futures market is used as a sample of the entire foreign exchange market.
-This is important because, unlike many other financial markets, it is the cash foreign exchange market that dominates in terms of activity and liquidity.
-The futures market is relatively small \citep{BISFX}.
 
@@
 The US dollar index has been active since it was launched in May 1992 but it is not trade weighted or really representative of the US dollar.
-
-It could be argued that this is a biased sample as it is a US market.
-Though international financial markets are global in nature, there is some evidence to support this claim.
-For example, there is a tendency to have a positive holding of the overseas currency as this is a hedge for exporters.
-Nonetheless, as was argued with the risk-reversals, it is the relative rather than absolute position of the signal that will determine the extreme and therefore if the US sample is proportionate to the population, sample signals can still be regarded as measuring an extreme.
-If the distribution of the sample itself is different from the population this is a problem, but there is no evidence that this is the case.
 
@@
 \caption{Descriptive Statistics for Weekly Exchange Rate Returns: 30 September 1998 to 31 December 2008}
-\begin{tabular}{l | r r  r r r }
+\begin{tabular}{{4cm} | r r  r r r }
   \hline
@@
 \end{landscape}
-
-\subsection{Event studies}
-Event studies are used to find the influence of a particular event on the performance of an asset price or security.
-James Dolley is credited with the first use of the method when looking at the effect of a stock split on the price of securities.\footnote{A stock split is when the number of shares are doubled and the value of each share is cut in half.
-This is usually done to prevent the price or cost of each share rising too far beyond the means of the small investor.} \citep{Dolly1933}.
-However, Dolly only looked at the market reaction the day after the stock split.
-This was some time after the stock holder meeting that would have ratified the decision of the management.
-Myer and Bakay argued that, if stocks tend to rise after a stock split, it would be best to buy the stocks when the announcement is made and before the split takes place and that, therefore, the price adjustment is likely to be spread over a number of days.
-They decided to measure the price action from a point eight weeks before the formal acceptance of the decision by the stock-holders.
-Myer and Barkay measured the price change between the base date and the meeting and tried to account for other influences on the price by recording the ratio of the stock price to the Standard and Poor Industry average, thereby providing the first attempt to separate the specific returns from those of the market \citep{MyersBakay}.
-
-Fama, Fisher and Roll use the event study method to conduct a much more systematic assessment of the effect of stock splits on security returns \citep{FamaFisherJensenRoll}.
-the study is considered to be the benchmark of Event Studies and it is the framework that is largely followed in the current work.
-
-The event is identified and the time at which the event occurs is marked as time zero (or $t_0$).
-Returns to the reference security in a window before and after the event are recorded and compared to the normal or expected returns.
-Therefore,
-\begin{equation}
-AR_{i \tau} = R_{i \tau} - E[R_{i \tau} | X_{\tau}]
-\end{equation}
-Where $AR_{i \tau}$ are the abnormal returns for security i at time $\tau$, $R_{i \tau}$ are the actual returns to security i at time $\tau$,  $E[R_{i \tau}]$ are the expected returns for security i at time $\tau$ conditional on the information $X_{\tau}$ and $\tau$ is the time relative to the event.
-Estimating the normal or expected returns is one of the key challenges of an event  study.
-It is usually tackled in one of two ways:  the security can be assumed to take a mean return or, as in the case of the paper by Fama and the others, a market model can be used.
-The original market model just takes a constant relationship between security returns and the market estimated by a linear regression.
-Subsequent models have added additional factors.
-The lack of any clear benchmark for currency returns and the fact that theoretically exchange rate returns should be on average equal to the forward rate, have made the case for a mean model.
-The seminal event studies have used equities and a market model.
-The expected return is zero and any returns that differ from that are identified as abnormal.
-Theoretically, the forward rate should be used but given the high frequency of the data, the interest rate differential is ignored.
-The evidence about returns in Tables \ref{tabref:DS1} and \ref{tabref:DS3} suggests that this is not likely to be a major problem.
-
-Is more needed here?
 
@@
 
+\begin{appendices}
+\section{Appendix 1}\label{appref:1}
+\begin{landscape}
+\begin{table}[t]
+\begin{threeparttable}
+\caption{Descriptive Statistics for Daily Exchange Rate Returns: 2 January 1996 to 6 September 2002}
+\begin{tabular}{l | r r  r r r r r r r r}
+  \hline
+ Statistic &  AUD & EURGBP & EURSEK & GBP & CHF & EURCHF & EURJPY & EUR & CAD & JPY \\
+  \hline
+Number & 1685 & 1675 & 1716 & 1667 & 1674 & 1674 & 1675 & 1668 & 1683 & 1675  \\
+Mean & -0.0174 & -0.0187 & 0.0011 & 0.0025 & 0.0120 & -0.0039 & -0.0072 & -0.0172 & 0.0084 & 0.0076\\
+Median &  0.0000 & -0.0298 & 0.0000& 0.0000 & 0.0114 & 0.0000 & 0.0000 & -0.0253 & 0.0000 & 0.0188\\
+Standard deviation  & 0.7241 & 0.5542 & 0.0438 & 0.0489 & 0.6819 & 0.0277 & 0.8416 & 0.6338 & 0.0363 & 0.8170\\
+Max & 5.8301 & 3.0543 & 2.1633 & 2.2573 & 3.5086 & 1.6477 & 4.0879 & 2.4136 & 2.2170 & 5.1788 \\
+Min & -3.4666 & -2.2858 & -2.2900 & -2.2701 & -3.6525 & -1.8466 & -8.2848 & -2.2901 & -2.9707 & -10.6573 \\
+Skewness  & 0.1206* & 0.2923* & 0.0432 & -0.0452 & -0.1423* & -0.4046* & -0.7388* & 0.1738* & -0.1095 & -1.7237*\\
+Kurtosis & 4.17* & 1.88 & 2.43* & 1.48 & 1.96 & 5.84* & 7.08* & 1.12 & 5.11* & 21.52*\\
+\hline
+ \label{tabref:DS1}
+\end{tabular}
+\begin{tablenotes}
+\small
+\item Continuously compounded returns (multiplied by 100) for period 2 January 1996 to  September 2002 for US dollar per unit of Australian dollar (AUD), UK sterling per unit of Euro (EURGBP), Swedish Crown per unit of EUR (EURSEK); US dollar per unit of UK sterling (GBP);  Swiss franc per unit of US dollar (CHF); Swiss franc per unit of Euro (EURCHF); US dollars per unit of Euro (EUR); Canadian dollars per unit of US dollar (CAD); Japanese yen per unit of US dollars (JPY).
+The asterisk (*) means that the estimate is significantly different from zero at the 5\% level using t-statistics for the mean and the standard error of skew and standard error of excess-kurtosis respectively.
+\end{tablenotes}
+\end{threeparttable}
+\end{table}
+\end{landscape}
+\end{appendices}
+
 \end{document}

### 2015-07-11 Knitr/NewPaper.Rnw
Updated part of the NewPaper. This does not fully work.  The pdf does not come up.  It is not clear what is wrong. A pdf is created.  Need to update something on momentum trading and on the information that some traders have.

@@
 There are a range of ideas about the way that speculation can affect financial markets.
-Towards one end of this spectrum is the view of Keynes from Chapter 13 of \emph{The General Theory} where he identified speculation as a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd'' \citet[p.
+Towards one end of this spectrum is the view of Keynes from Chapter 12 of \emph{The General Theory} where he identified speculation as a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd'' \citet[p.
 101]{Keynes1936}.
-Here speculative activity is a problem that, at its most extreme, will cause booms, reversals and will discourage fundamental, long-term investment.
-It is noise that does not emerge randomly, but feeds off itself as the evolution of collective sentiment encourages trends in asset prices.
+Here speculation is a problem that, at its most extreme, will cause booms, reversals and will discourage fundamental, long-term investment.
+It is noise that does not emerge randomly, but feeds off itself in an evolution of collective sentiment that encourages trends in asset prices.
 
-Towards the other end, Milton Friedman argued that speculation was a stabilising force, providing liquidity and helping to ensure that markets swiftly find equilibrium.
+Towards the other end of the spectrum, Milton Friedman argued that speculation was a stabilising force, providing liquidity and helping to ensure that markets swiftly find equilibrium.
 In arguing in favour of flexible exchange rates,  \citet{FriedmanPositive} asserted that successful speculators would be those who managed to correctly buy when prices  were below fundamental value and sell above that price.
 Therefore, Friedmand argued, the speculative process would favour those able to access or utilise information over the uninformed and, through a Darwinian process, informed speculation would dominate.
-This view regards speculation as an key lubricant of financial market as it provides liquidity and facilitates the process of price discovery.
+This view regards speculation as an key lubricant of financial markets as it provides liquidity and facilitates the process of price discovery.
 
@@
 If speculators are uninformed and their activity can be regarded as noise, the most extreme sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when reversals are most likely to take place; if speculators are informed prices should follow a random walk after either of these extremes.
-    The foreign exchange market is utilised because it is the largest and most liquid market in the world.
-The latest survey taken by the Bank for International Settlements (BIS) in April 2013 estimated average daily turnover at \$5.4trn with average daily spot transactions of \$2.0trn \citet{BISFX2013}.
+    The foreign exchange market is utilised because it is the largest and most liquid financial market in the world.
+The latest survey taken by the Bank for International Settlements (BIS) in April 2013 estimated average daily turnover at \$5.4trn with average daily spot transactions of \$2.0trn See \citet{BISFX2013}.
 The cost of trading in this market, when measurd in terms of the size of the bid-ask spread, is extremely low.
-An analysis of the bid-ask spreads for USD-JPY, GBP-USD and EUR-USD for the period January 2001 to 2005 and finds that the average spread for each is 0.342\%, 0.305\% and 0.326\% respectively.
-This would be the cost for a round-trip (buying and selling immediately)\citep{Steely2013}.
+An analysis by \citet{Steely2013} of the bid-ask spreads for USD-JPY, GBP-USD and EUR-USD for the period January 2001 to 2005 and finds that the average spread for each is 0.342\%, 0.305\% and 0.326\% respectively.
+This would be the cost for a round-trip (buying and selling immediately).
 
@@
 Speculation does not appear to be random, there is momentum to the rise and fall in sentiment and positions.
-However, while currency prices co-move with sentiment and positions, there is no informational content from either.
-As such, knowing the the strength of speculative opinion or the weight of speulative positions does not improve forecasts of future returns.
+However, while currency prices co-move with sentiment and positions, there is no informational content from either: knowing the the strength of speculative opinion or the weight of speulative positions does not improve forecasts of future returns.
 This supports the suggestion that speculative activity has a permanent affect on prices and that is consistent with speculation being part of the process of price discovery.
@@
 However, there is no clear, unambiguous theory of speculation.
-Speculation plays a key role in three importantant and related areas: market-microstructure, the rationale for the trading of financial securities and rational expectations.
+Speculation plays a key role in three important and related areas: market-microstructure, the rationale for the trading of financial securities and rational expectations.
 
@@
 There must be different opinions.
-Why Will traders be prepared to pay the cost of acquiring information if they are unable to take advantage of that information?
+Why will traders be prepared to pay the cost of acquiring information if they are unable to take advantage of that information?
 With speculative noise-traders, the informed can take advantage of their information.
 This creates a justification for acquiring information.
-Speulative noise-traders can therefore create a gap between the price and the fundamtental value that will encourage informed traders to pay the cost of acquiring information.
+Speculative noise-traders can therefore create a gap between the price and the fundamtental value that will encourage informed traders to pay the cost of acquiring information.
 The divergence between this price and fundametnal value is thereby essential to encourage te process of price discovery.
@@
 
-  Speculation is also an important part of the argument for why prices should tend towards that consistent with \emph{rational expectations}.
-\citet{muth1961Rational} uses commodity speculation as the means to drive prices towards the \emph{rationa-expectations-equilibrium} (RER).
-In this case, the speculators are informed and tend to reduce the volatility of prices.
-This is a reiteration of the Friedman idea of speuclation but runs immediately into the \citet{milgrom1982information} problem of finding counterparties willing to make the other side of the transaction.
+  Speculation is also an important part of the argument for why prices should tend towards a level that consistent with \emph{the rational expectations equilibrium} (RER).
+This is the level that is consistent with rational expectations.
+In the original version of rational expectations,  \citet{muth1961Rational} uses commodity speculation as the means by which prices are driven towards the level where expectations are model-consistent.
+These are of Friedman-type speculators as they base their expectations on their knowledge of the underlying model and they are the means by which the components of the model tend towards equilibrium values.
+% See \citet gc{Lyons2006Microstructure} and \citep{Evans2002Order} for examples of the micrstrucure approach in foreign exchange.
 
-% See \citep{Lyons2006Microstructure} and \citep{Evans2002Order} for examples of the micrstrucure approach in foreign exchange.
-
-If there are a mixture of informed and uninformed specualtors, the question becomes whether the uninformed speculative activity can be considered as just random noise?
-The original microstrucutre version had moise-traders arriving in a random fashion so that the market-maker could infer information from an imbalance of orders \citet{Kyle1985Continuous}.
 However, though the idea that noise is a random variable may be a construction that is useful for modelling and it may be justified in analysis of broader themes of microstructure and levels of market efficiency, this exogenous construct may be an over-simplification that is inapprpropriate when looking for more understanding of speculation itself.
@@
 
+The institutional explanations are assciated with the allocation of credit for investment and trading purposes:  as asset values decline exchanges, banks and other providers of liqudiity require larger margins and deeper haircuts; investors liquidate their fund holdings; the atttraction of hedging increases.
+See \citet{BrunnermeierLiquidity} and \citet{Gorton} for an overview of the institutional literature.
+Psychological explanations revolve around the common panic, the group-thining and the tendency to read private information from the actions of other investors: the so-called \emph{wisdom of crowds} depends on independent evaluation but as Keynes noted in his discussion of the \emph{beauty contest}, it is generally more important to identify what other investors will believe than acertain underlying, fundamental value \citet[p.101]{Keynes1936}
+
+it has also been suggested that speculative technieques can change overtime.
+For example, \citet{ShillerFashion} discusses the use of fashionable techniques or fads in investment.
+In this way, speculation noise is associated with sentiment and measures of sentiment are then used as ways of measuring speculation.
 %It would be good to have some more modern ideas here.  Maybe they are just the empirical results. 
 %this paragraph is not very good.  It needs to be updated.  Look at the file in citulike. 
-\subsection{Other empirical work}
-There is little empirical eassessment of these theories about speculation and noise-trading.
-In part this is because information is limited and it is usually difficult to distinguish between speculators and other investors and, even more so, between informed and uninformed.
+\subsection{Empirical work}
+Empirical assessment of speuclation has been limited by the scarcity of data.
+It is not very easy to distinguish between speculators and other types of investors.
+It is even harder to distinguish those that may be noise-traders from those assumed to be informed.
+One approach has been to assume that noise-traders utilise simple investment strategies and to use tests of these strategies as a means of testing speculation.
+For example speculation can be identified with momentum trading.
+It is also possible to assume that certain investors, such as investment banksa and hedge funds, are informed while other trade without inforation.
+The latter group may include day-traders or private individuals.
 
-One approach has been to assume that noise-traders utilise simple investment strategies that involve momentum or reversals.
-For example, \citet{  shillerFashion} discusses the use of fashionable techniques or fads in investment.
-In this way, speculation noise is associated with sentiment and measures of sentiment are then used as ways of measuring speculation.
-Alternatively, speculation can be identified with momentum trading.
-The literature on heterogeneous traders, another way that the trading securities is explained, has very frequently used noise-traders as trend-chasers or followers of technical analysis.
+%The literature on heterogeneous traders, another way that the trading securities is explained, has very frequently used noise-traders as trend-chasers or followers of technical analysis.
 These models (see for example brock and hommes [1998]; lux [1995]; lux and marchesi [2000]) investigate bubbles and crashes, and feature fundamental traders versus noise traders.
@@
 \subsection{Why this is important}
-The traditional Keynesian view of speculation asserted that it would be an additional noise that would make it even harder for informed-traders to overcome uncertainty and apply fundamental information about long-term value.
-This view finds some supported in Shiller's observation that stock prices are more volatile than the underlying earnings information on which valuation should be based \citet{ShillerVol}.
+The Keynesian view of speculation is that it is an addition to market noise that makes it even harder for informed-traders to overcome uncertainty and apply fundamental information about long-term value.
+This description provides an explanation for Shiller's observation that stock prices are more volatile than the underlying earnings information on which valuation should be based \citet{ShillerVol}.
 In addition, \citet{Delong1990noise} and \citet{shleifer1997limits} show that constrained horizons can cause noise-trading to limit the ability of informed traders to take advantage of profitable opportunities for fear of further divergence from fundamental value before realignment is achieved.
@@
 
-For example, the debate about a tax on financial transactions, \emph{the Tobin tax} \citet{TobinTax}, has been intended to reduce speculation.
+The debate about a tax on financial transactions, initiated by \citet{TobinTax} and now called the \emph{the Tobin tax} , has been intended to reduce speculation.
 The surge in commodities prices that was seen prior to the GFC was seen to be partially the consequence of increased speculation in commodities and the creation of exchange traded funds (ETF) focusing on commodities markets.
-This argument depends on the idea that speculators will drive the price away from its fundamental value.
+The argument for a tax on finanical transactions depends on the idea that speculators will drive the price away from its fundamental value.
 
@@
 This will be use to gain a fuller understanding of speculative behaviour.
-This will involve three steps:  first, assess the price implication of the noise-trader model;  second, measure speculative activity in the foreign exchange market; finally, assess whether extreme speculative activity in the foreign exchange market is consistent with the noise-trader model.
+The experiement will involve three steps:  assess the price implication of the noise-trader model;  measure speculative activity in the foreign exchange market; finally, assess whether extreme speculative activity in the foreign exchange market is consistent with the noise-trader model.
 If the speculative activity that is seen in the foreign exchange market produces results that are like the noise-trader model, it can be said that speculation is noise trading as Keynes asserted, if not it may be more informed as suggested by Friedman.
 % where it is widely perceived that there are a relatively large number of traders engaged in technical analysis.
-See \citep{FrootChartists} for a survey.
+See \citet gc{FrootChartists} for a survey.
 %this may go in the explanation for why fx is used.
@@
 
-The \citep{Delong1990noise} noise-trader model used in this study.
-This is a derivative of the \citep{Samuelson1958} overlapping generations model  with two-period agents making a portfolio choice.
+The \citet{Delong1990noise} noise-trader model used in this study.
+This is a derivative of the \citet{Samuelson1958} overlapping generations model  with two-period agents making a portfolio choice.
 There are two assets:  a safe asset $(s)$ which pays a set dividend $(d)$ and is in perfectly elastic supply; the risky asset $(u)$ also pays the same dividend but is not in elastic supply.
@@
 
-In words, the deviation from the true fundamental value of unity is a result of the discounted value of the noise-traders' misperception of the price $(p_t - \rho^*)$, multiplied by the weight of noise-traders in the total $(\mu)$ plus a perpetuity of the average noise-trader misperception $(\frac{ \mu \rho^*}{r} )$, adjusted for the risk.
-This risk is a function of the volatility of the price which,  from equation \ref{eq:NP1} is a result of the variability of the misperception of speculative noise-traders.
+The deviation from the true fundamental value of unity is a result of the discounted value of the noise-traders' misperception of the price $(p_t - \rho^*)$, multiplied by the weight of noise-traders in the total $(\mu)$ plus a perpetuity of the average noise-trader misperception $(\frac{ \mu \rho^*}{r} )$, adjusted for the risk.
+This risk is a function of the volatility of the price which,  from Equation \ref{eq:NP1} is a consequence of the variability of the misperception of speculative noise-traders.
 All but the second term in equation \ref{eq:NP1} are constants, and if the variance of the price one period ahead  is equal to the variance of the misperception,
@@
 
-The following effects are noted:  the \emph{Friedman Effect}, which is the tendency of informed investors to push prices back towards fundamental value; \emph{Price Pressure} as the way that speculative noise-traders buy more and drive the price higher as they become more bullish; \emph{Hold More},whereby the noise-traders hold more of the risky asset; \emph{Create Space} is the added risk that is caused by an increase in the variability of noise-trader belief \citep[pp.
-14-15]{Delong1990noise}.
+The following four effects are noted by \citet[pp.
+14-15]{Delong1990noise}:  the \emph{Friedman Effect}, as informed speculators move prices back towards fundamental value; \emph{Price Pressure} as the speculative noise-traders drive prices towards their misperceived vaue; \emph{Hold More} as noise-traders misperceive the value of the risky asset and tend to hold more of it; \emph{Create Space} is the noise-trader risk caused by the difference between their perception and the actual value, it is an added risk that is caused by the variability of noise-trader belief as noise-traders could drive the price even further from its fundamental value.
+In this model, tHe more numerous the noise-traders, the more volatile prices.
 %Put something on these four in the conclusion. 
-The main findings are that noise-traders will not necessarily disappear as Friedman had suggested \citep{FriedmanPositive}.
+The main findings are that noise-traders will not necessarily disappear as Friedman had suggested \citet{FriedmanPositive}.
 While noise-traders will tend to get less utility than sophisticated traders, this does not necessarily drive them from the market.
@@
 The simulation creates a 1000 period price series from Equation \ref{eq:NP2}.
-The dividend paid on the risky asset $(r)$ is posted at 10\%; the risk-aversion parameter $(\gamma)$ is assumed to be 0.3 (in line with that suggested in \citep[p.
+The dividend paid on the risky asset $(r)$ is posted at 10\%; the risk-aversion parameter $(\gamma)$ is assumed to be 0.3 (in line with that suggested in \citet[p.
 127]{FrobesBehaviour}); starting from a price of 100, the average noise-trader mis-perception $(p^*)$ is a random variable in the first simulation with a mean of 4 and a variance $(\sigma_{\rho^*}^2)$ of 2, and has a fixed value of 4 in the second simulation; the weight of noise-traders $(\mu)$ is 50\% in the first case but is modelled as a two-period moving average of a normal random variable centred at 50\% with a standard deviation of 20\% and equally weighted average of its previous value (starting at 50\%) to allow smooth adjustment, in the second simulation.
+% it is possible to have other specifications. For example, the noise-traders oculd become more prominent if their strategy is more successful,
 
@@
 The R package \emph{eventstudies} createed by \citet{eventstudies} is used for the data manipulation.
-Descriptive statistics for the whole series of returns and for those when misperception or weight of noise-traders are presented in Table \ref{tabref:MC1}.
+Descriptive statistics for the whole series of returns and for those when either the extent of their misperception or the weight of noise-traders is extreme are presented in Table \ref{tabref:MC1}.
 The process is then repeated for one thousand simulations of each model so that confidence intervals can be established for the estimates of the descriptive statistics.
@@
 \small
-\item Descriptive statistics for the De Long Noise-Trader model \citep{Delong1990noise}.
+\item Descriptive statistics for the De Long Noise-Trader model \citet{Delong1990noise}.
 The Base Case measures the benchmark as returns from prices generated by Equation \ref{eq:NP2} for the whole sample.
@@
 Frankel and Froot found that the median forecast from the MMS survey exhibited trend-following behaviour at the one to three week horizon while there is mean-reversion beyond that point.
-Frankel and Froot also assess the 'Financial Report' that has been produced by \emph{The Economist} magazine and \emph{Currency Forecasters Digest} and find a wide dispersion of opinion and express some concern about the aggregation of such divergent forecasters \citep{FrootFrankelFDB} and \citep{FrankelFroot1987}.}
-This is the relative price of currency put and call options for the same strike price.
-The \emph{Put-Call-Parity} in the Black-Scholes option pricing model assumes that these will be the same \citep{Black1973Options}.
+Frankel and Froot also assess the 'Financial Report' that has been produced by \emph{The Economist} magazine and \emph{Currency Forecasters Digest} and find a wide dispersion of opinion and express some concern about the aggregation of such divergent forecasters \citet gc{FrootFrankelFDB} and \citep{FrankelFroot1987}.}
+Risk-reversal skew is the relative price of currency put and call options for the same strike price.
+The \emph{Put-Call-Parity} in the  \citet{Black1973Options} option pricing model assumes that these will be the same.
 While the Black-Scholes model is the basis for valuing options, where the price of a non-dividend paying stock trades at a discount to the forward rate equal to the risk free rate, the forward rate in a currency option is determined by the relative interest rate differential between the two currencies.
-The \citep{GarmanKohlhagan} adjustment that accounts for the difference in interest payments on the two currencies is used.
+The \citet{GarmanKohlhagan} adjustment that accounts for the difference in interest payments on the two currencies is used.
 The spot discount is determined therefore by the relative interest rate differential for the appropriate maturity.
@@
 The dealer pays 0.15\%.
-Alternatively, the dealer would sell a 25 delta call at 8.78\% and buy the USD-call-CAD-put at 8.50\%, earning 0.28\% \citep{Global}.
+Alternatively, the dealer would sell a 25 delta call at 8.78\% and buy the USD-call-CAD-put at 8.50\%, earning 0.28\% \citet{Global}.
 }  Therefore, the relationship between implied volatility for calls and implied volatilities for puts for the same delta and maturity can be taken as an indication of how much more expensive calls are relative to the equivalent put or as the market bias towards puts over calls or, in this case, as an indication of market sentiment.
@@
 If extreme deviations or speculations are noise they should presage a sharp reversal.
-If speculation is informed, this is part of the price discovery process and extremes may be followed by calm conditions around a new equilibrium.
+If speculation is informed, this is part of the price discovery process and extremes may be followed by a random walk.
 
@@
 This is what would be expected if, as has been identified in many studies, exchange rates follow a random walk.
-This apparent stability hides some significant changes that are taking place below the surface.
-For example, the Japanese yen records an appreciation of 10.66\% on one day.
-Most of the exchange rate series appear to be skewed in some way and there is also evidence of kurtosis in many cases.
-An asterisk is used to denote statistical significance at the level of 95\% confidence.
-A t-test is used for the mean; a \emph{standard error of skew} and a \emph{standard error of kurtosis} are used for the skew and kurtosis respectively.
-These are calculated as $SES = \sqrt{\frac{6n(n-1)}{(n-2)(N+1)(N+3)}}$ and $SEK = \sqrt{\frac{n^2-1}{(n-3)(N+5)}}$, with the assumption that this estimate is distributed normally, the skew or kurtosis is considered significant if it is more than two standard errors from zero.
-
 \begin{landscape}
@@
 The latter may be related to hedging activity by Japanese exporters.
-While this could suggest that implied volatility is an imperfect measure of speculative activity, note that extreme readings are based on quantiles of the distribution and therefore a discrete adjustment of this distribution to reflect the hedging power should not change the measurement of extremes.
+While this could suggest that implied volatility is an imperfect measure of speculative activity, note that extreme readings are based on quantiles of the distribution and therefore a discrete adjustment of this distribution to reflect the hedging power should not change the timing of signals measured by extremes.
 
@@
 Before that the data were released every two weeks.
-The CoT is released each Friday and details the positions outstanding the previous Tuesday \citep{cot}.
+The CoT is released each Friday and details the positions outstanding the previous Tuesday.
+See \citet{cot}.
 
@@
 Non-commercial accounts are generally considered to be speculators.
-\citep{FuturesSanders} and \citep{FuturesWang} have examples of previous use of this data in this fashion.
-
-The administrative requirements are more onerous for non-commercial accounts but the CFTC have the right, which they use, to re-categorise parties if they are not happy with the self-selection.
-Positions are either long or short, meaning that they will receive or deliver the underlying; for non-commercial accounts there are also \emph{spreading} positions which are simultaneous holdings of long and short positions in different maturities of the same contract, taking a position on the spread between two maturity dates.
-These are assumed to be speculators.
-The CFTC data also report \emph{Open Interest} (OI) which is the record of outstanding open positions.
-
-If it is assumed that futures and cash markets are two separate markets and that most participants have a preference for one or the other and therefore do not frequently switch; if it is also assumed that the nature and level of speculative activity in the futures market is an indicator of what is happening in the broader spot market, it will be possible to use this data, which measures speculative activity in the futures market to assess the intensity and level of speculative activity in the foreign exchange market more generally.
-In other words, the futures market is used as a sample of the entire foreign exchange market.
+See \citet{FuturesSanders} and \citet{FuturesWang} for examples of previous use of this data in this fashion.
 
@@
 The contracts are for EUR 100,000; GBP 62,500; JPY 12,500,000; CHF 100,000 and CAD 100,000.
-For the CoT data, two series are created:  S1 measures the net long non-commercial positions relative to total non-commercial positions and S2 measures the net long non-commercial positions relative to open interest.
-Table \ref{tabref:DS3} given an overview of the descriptive statistics for the returns on the weekly foreign exchange rates that are relevant for the futures market data.
-There are no significant differences from the daily data that is used in the risk-reversal part of the study recorded in Table \ref{tabref:DS1}.
-Once again, the mean return is very close to zero.
-This is consistent with the idea that exchange rates follow a random walk.
-The distribution of returns are skewed for some pairs and often fat-tailed (identified with an asterisk where statistically significant as either a t-test or a test of the standard error of skew or standard error of kurtosis as calculated above).
-
- \begin{table}[t]
-\begin{threeparttable}
-\caption{Descriptive Statistics for Weekly Exchange Rate Returns: 30 September 1998 to 31 December 2008}
-\begin{tabular}{{4cm} | r r  r r r }
-  \hline
- Statistic &  EUR & JPY & GBP & CHF & CAD  \\
-  \hline
-Number & 522 &  849 & 849 & 847 & 849 \\
-Mean & 0.0351 & 0.0337 & -0.0202 & 0.0233 & 0.0018\\
-Median & 0.0650 & -0.0800 & 0.0000 & 0.0000 & -0.0200\\
-Standard deviation  & 1.4063 & 1.6355 & 1.2755 & 1.5418 & 1.0558\\
-Max & 8.7900 & 11.6700 & 5.7200 & 8.3400 & 10.0100\\
-Min & -4.2400 & -15.0400 & -6.4100 & -5.7100 & -5.0900\\
-Skew  & 0.27* & -0.06 & -0.34* & 0.31* & 0.78*\\
-Kurtosis & 2.76*& 12.58* & 2.07 * & 1.82 & 11.59*\\
-\hline
- \label{tabref:DS3}
-\end{tabular}
-\begin{tablenotes}
-\small
-\item Continuously compounded returns multiplied by 100 for US dollar per unit of Euro (EUR), US dollar per unit of Japanese Yen (JPY), US dollar per unit of UK Sterling (GBP), US dollar per unit of Swiss Franc (CHF) and US dollar per unit of Canadian dollar (CAD) for the period 30 September 1998 to 31 Dec 2008;     An asterisk (*) means that the estimate is significantly different from zero at the 5\% level of statistical significance using t-statistics for the mean and the standard error of skew and standard error of excess-kurtosis respectively.
-These are calculated as $SES = \sqrt{\frac{6n(n-1)}{(n-2)(N+1)(N+3)}}$ and $SEK = \sqrt{\frac{n^2-1}{(n-3)(N+5)}}$, with the assumption that this estimate is distributed normally, the skew or kurtosis is considered significant if it is more than two standard errors from zero.
-\end{tablenotes}
-\end{threeparttable}
-\end{table}
 
@@
 For the risk reversal skew, as a secondary test, a reading above the ninety-fifth and below the fifth percentile is also established as an event.
-The Event Studies R package was used to manage the data see \citep{eventstudies} for fuller details on the package and its use.
+The Event Studies R package was used to manage the data see \citet{eventstudies} for fuller details on the package and its use.
 The cumulative abnormal returns for the whole of the event window (CARW) and the cumulative abnormal returns for after the event (CARA) were recorded.
@@
 It shows that the spot price is following the option risk-reversal skew and, as Figure \ref{fig:ES1} reveals, it would be even more pronounced if the event window just ran from the starting point to the event.
-The findings here are consistent with the evidence that orders and sentiment drive prices \citep{Evans2002Order}, \citep{FuturesSanders} and \citep{FuturesWang}.
+The findings here are consistent with the evidence that orders and sentiment drive prices \citet{Evans2002Order}, \citet{FuturesSanders} and \citet{FuturesWang}.
 
@@
 This suggests a market inefficiency that would reward momentum-based trading.
-It is no wonder therefore that \citep{TaylorTechnical}, reporting a survey conducted by the Bank of England, found that of the Chief foreign exchange dealers contacted in London in November 1988, 90\% of them attached some weight to \emph{technical analysis}, a method that is fundamentally based on finding trends in asset prices.
+It is no wonder therefore that \citet{TaylorTechnical}, reporting a survey conducted by the Bank of England, found that of the Chief foreign exchange dealers contacted in London in November 1988, 90\% of them attached some weight to \emph{technical analysis}, a method that is fundamentally based on finding trends in asset prices.
 The survey found that the importance of technical analysis became greater at shorter horizons.
@@
 The importance of momentum trading and technical analysis has not waned with time.
-Neely and Wller  show that gains from technical trading rules still generate excess returns and attribute this to behavioural biases \citep{NeelyWeller}.
+Neely and Wller  show that gains from technical trading rules still generate excess returns and attribute this to behavioural biases \citet{NeelyWeller}.
 It seems that speculation forms part of the process of price discovery, but this process is not instantaneous as would be expected in the purest form of the EMH.
@@
 \begin{appendices}
-\section{Appendix 1}\label{appref:1}
+\section{Appendix 1}\label{apref:1}
 \begin{landscape}
@@
 \end{landscape}
+\section{Appendix 2}
+ \begin{table}[t]
+\begin{threeparttable}
+\caption{Descriptive Statistics for Weekly Exchange Rate Returns: 30 September 1998 to 31 December 2008}
+\begin{tabular}{{4cm} | r r  r r r }
+  \hline
+ Statistic &  EUR & JPY & GBP & CHF & CAD  \\
+  \hline
+Number & 522 &  849 & 849 & 847 & 849 \\
+Mean & 0.0351 & 0.0337 & -0.0202 & 0.0233 & 0.0018\\
+Median & 0.0650 & -0.0800 & 0.0000 & 0.0000 & -0.0200\\
+Standard deviation  & 1.4063 & 1.6355 & 1.2755 & 1.5418 & 1.0558\\
+Max & 8.7900 & 11.6700 & 5.7200 & 8.3400 & 10.0100\\
+Min & -4.2400 & -15.0400 & -6.4100 & -5.7100 & -5.0900\\
+Skew  & 0.27* & -0.06 & -0.34* & 0.31* & 0.78*\\
+Kurtosis & 2.76*& 12.58* & 2.07 * & 1.82 & 11.59*\\
+\hline
+ \label{tabref:DS3}
+\end{tabular}
+\begin{tablenotes}
+\small
+\item Continuously compounded returns multiplied by 100 for US dollar per unit of Euro (EUR), US dollar per unit of Japanese Yen (JPY), US dollar per unit of UK Sterling (GBP), US dollar per unit of Swiss Franc (CHF) and US dollar per unit of Canadian dollar (CAD) for the period 30 September 1998 to 31 Dec 2008;     An asterisk (*) means that the estimate is significantly different from zero at the 5\% level of statistical significance using t-statistics for the mean and the standard error of skew and standard error of excess-kurtosis respectively.
+These are calculated as $SES = \sqrt{\frac{6n(n-1)}{(n-2)(N+1)(N+3)}}$ and $SEK = \sqrt{\frac{n^2-1}{(n-3)(N+5)}}$, with the assumption that this estimate is distributed normally, the skew or kurtosis is considered significant if it is more than two standard errors from zero.
+\end{tablenotes}
+\end{threeparttable}
+\end{table}
+
 \end{appendices}

### 2015-07-12 Knitr/NewPaper.Rnw
NewPaper updated. Now try to use sideways table rather than landscape.  The appendices prevent the internal pdf version from poping up.

@@
 There are a number of contemporary questions where the role and nature of speculation is at the centre of the debate.
-Regulations or taxes that reduce the amount of speculation have generally been considered to be beneficial by reducing risk and encouraging more long-term attitudes towards investment.
+Regulations or taxes that reduce the amount of speculation have generally been considered to be beneficial as they reduce risk and encourage more long-term attitudes towards investment.
 However, where these measures reduce liquidity, higher risk and an increase in other costs may provide an countervailing force.
 
-The debate about a tax on financial transactions, initiated by \citet{TobinTax} and now called the \emph{the Tobin tax} , has been intended to reduce speculation.
+The debate about a tax on financial transactions, initiated by \citet{TobinTax} and now called the \emph{the Tobin tax}, has been intended to reduce speculation.
 The surge in commodities prices that was seen prior to the GFC was seen to be partially the consequence of increased speculation in commodities and the creation of exchange traded funds (ETF) focusing on commodities markets.
@@
 
-In addition, so-called \emph{Volcker Rule} would prevent US investment banks trading on their own-account.
+%In addition, so-called \emph{Volcker Rule} would prevent US investment banks trading on their own-account.
 This, it has been argued, will reduce liquidity in the bond market by causing investment banks to hold lower levels of bond inventory on their books.
@@
 In this model, tHe more numerous the noise-traders, the more volatile prices.
-%Put something on these four in the conclusion. 
+
+The Friedman-effect and the Price-pressure are the effects that are studied here.
+Their combination suggests that when the weight of noise-traders is high or the times when they have particularly stong opinions will tend to be those when there is the greatest risk of reversals.
+The hold-more effect suggests that the weight of speculative involvement in the market will tend to increase as a trend.
+Create-space effect would suggest that when the weight of noise-traders is high, there is incresaed risk.
+%this can be assessed.
+
 The main findings are that noise-traders will not necessarily disappear as Friedman had suggested \citet{FriedmanPositive}.
@@
 However, the combination of high level of speculation and divergence from fundamentals is likely to create the greatest risk.
-
 \bibliography{../../../../myrefs}

### 2015-07-13 Knitr/NewPaper.Rnw
Read through complete.  One more read and then send. Appendix still does not work correctly.

@@
 In arguing in favour of flexible exchange rates,  \citet{FriedmanPositive} asserted that successful speculators would be those who managed to correctly buy when prices  were below fundamental value and sell above that price.
-Therefore, Friedmand argued, the speculative process would favour those able to access or utilise information over the uninformed and, through a Darwinian process, informed speculation would dominate.
+Therefore, Friedman argued, the speculative process would favour those able to access or utilise information over the uninformed and, through a Darwinian process, informed speculation would dominate.
 This view regards speculation as an key lubricant of financial markets as it provides liquidity and facilitates the process of price discovery.
@@
 If speculators are uninformed and their activity can be regarded as noise, the most extreme sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when reversals are most likely to take place; if speculators are informed prices should follow a random walk after either of these extremes.
+
 The foreign exchange market is utilised because it is the largest and most liquid financial market in the world.
 The latest survey taken by the Bank for International Settlements (BIS) in April 2013 estimated average daily turnover at \$5.4trn with average daily spot transactions of \$2.0trn See \citet{BISFX2013}.
-The cost of trading in this market, when measurd in terms of the size of the bid-ask spread, is extremely low.
+The cost of trading in this market, when measured in terms of the size of the bid-ask spread, is extremely low.
 An analysis by \citet{Steely2013} of the bid-ask spreads for USD-JPY, GBP-USD and EUR-USD for the period January 2001 to 2005 and finds that the average spread for each is 0.342\%, 0.305\% and 0.326\% respectively.
@@
 Speculation does not appear to be random, there is momentum to the rise and fall in sentiment and positions.
-However, while currency prices co-move with sentiment and positions, there is no informational content from either: knowing the the strength of speculative opinion or the weight of speulative positions does not improve forecasts of future returns.
+However, while currency prices co-move with sentiment and positions, there is no informational content from either: knowing the the strength of speculative opinion or the weight of speculative positions does not improve forecasts of future returns.
 This supports the suggestion that speculative activity has a permanent affect on prices and that is consistent with speculation being part of the process of price discovery.
@@
 
-One key element of the microstructure literature involves distinguishing between uninformed or noise-traders and those who have infomration.
-The \citet{Kyle1985Continuous} model has market-makers reciving orders from the informed and the uninformed.
+One key element of the microstructure literature involves distinguishing between uninformed or noise-traders and those who have information.
+The \citet{Kyle1985Continuous} model has market-makers receiving orders from the informed and the uninformed.
 On the assumption that the uniformed arrive in a random fashion without any reason to trade, the weight of orders implies something about the activity of the informed and causes prices to change in the process of price discovery.
-The bid-ask spread is the payment that the market-maker recieves in return for taking the risk of meeting an informed trader.
-Ultimately, the uninformed traders make payments to the informed traders with the mark-makers taking a cut for the risk and the privision of exchange.
+The bid-ask spread is the payment that the market-maker receives in return for taking the risk of meeting an informed trader.
+Ultimately, the uninformed traders make payments to the informed traders with the mark-makers taking a cut for the risk and the provision of exchange.
 See also \citet{glosten1985bid}.
@@
 1]{GortonNoise} categorise noise-traders as all those trading in markets for ``non-information-based reasons''.
-In this case, uninforrmed speculation provides a rational and an opportunity for the trading of financial securities.
+In this case, uninformed speculation provides a rational and an opportunity for the trading of financial securities.
 If all economic agents know the fundamental value of a security how can some be buying and some be selling?
@@
 This creates a justification for acquiring information.
-Speculative noise-traders can therefore create a gap between the price and the fundamtental value that will encourage informed traders to pay the cost of acquiring information.
-The divergence between this price and fundametnal value is thereby essential to encourage te process of price discovery.
+Speculative noise-traders can therefore create a gap between the price and the fundamental value that will encourage informed traders to pay the cost of acquiring information.
+The divergence between this price and fundamental value is thereby essential to encourage the process of price discovery.
 %
@@
 
-However, though the idea that noise is a random variable may be a construction that is useful for modelling and it may be justified in analysis of broader themes of microstructure and levels of market efficiency, this exogenous construct may be an over-simplification that is inapprpropriate when looking for more understanding of speculation itself.
+However, though the idea that noise is a random variable may be a construction that is useful for modelling and it may be justified in analysis of broader themes of microstructure and levels of market efficiency, this exogenous construct may be an over-simplification that is inappropriate when looking for more understanding of speculation itself.
 Institutional and psychological explanations for noise-trading frequently suggest that there are endogenous non-random elements.
 
-The institutional explanations are assciated with the allocation of credit for investment and trading purposes:  as asset values decline exchanges, banks and other providers of liqudiity require larger margins and deeper haircuts; investors liquidate their fund holdings; the atttraction of hedging increases.
+The institutional explanations are associated with the allocation of credit for investment and trading purposes:  as asset values decline exchanges, banks and other providers of liquidity require larger margins and deeper haircuts; investors liquidate their fund holdings; the attraction of hedging increases.
 See \citet{BrunnermeierLiquidity} and \citet{Gorton} for an overview of the institutional literature.
-Psychological explanations revolve around the common panic, the group-thining and the tendency to read private information from the actions of other investors: the so-called \emph{wisdom of crowds} depends on independent evaluation but as Keynes noted in his discussion of the \emph{beauty contest}, it is generally more important to identify what other investors will believe than acertain underlying, fundamental value \citet[p.101]{Keynes1936}
+Psychological explanations revolve around the common panic, the group-thinking and the tendency to read private information from the actions of other investors: the so-called \emph{wisdom of crowds} depends on independent evaluation but as Keynes noted in his discussion of the \emph{beauty contest}, it is generally more important to identify what other investors will believe than actual underlying, fundamental value \citet[p.101]{Keynes1936}
 
-it has also been suggested that speculative technieques can change overtime.
+it has also been suggested that speculative techniques can change overtime.
 For example, \citet{ShillerFashion} discusses the use of fashionable techniques or fads in investment.
@@
 \subsection{Empirical work}
-Empirical assessment of speuclation has been limited by the scarcity of data.
+Empirical assessment of speculation has been limited by the scarcity of data.
 It is not very easy to distinguish between speculators and other types of investors.
@@
 For example speculation can be identified with momentum trading.
-It is also possible to assume that certain investors, such as investment banksa and hedge funds, are informed while other trade without inforation.
+It is also possible to assume that certain investors, such as investment banks and hedge funds, are informed while other trade without information.
 The latter group may include day-traders or private individuals.
@@
 The surge in commodities prices that was seen prior to the GFC was seen to be partially the consequence of increased speculation in commodities and the creation of exchange traded funds (ETF) focusing on commodities markets.
-The argument for a tax on finanical transactions depends on the idea that speculators will drive the price away from its fundamental value.
+The argument for a tax on financial transactions depends on the idea that speculators will drive the price away from its fundamental value.
 
@@
 This will be use to gain a fuller understanding of speculative behaviour.
-The experiement will involve three steps:  assess the price implication of the noise-trader model;  measure speculative activity in the foreign exchange market; finally, assess whether extreme speculative activity in the foreign exchange market is consistent with the noise-trader model.
+The experiment will involve three steps:  assess the price implication of the noise-trader model;  measure speculative activity in the foreign exchange market; finally, assess whether extreme speculative activity in the foreign exchange market is consistent with the noise-trader model.
 If the speculative activity that is seen in the foreign exchange market produces results that are like the noise-trader model, it can be said that speculation is noise trading as Keynes asserted, if not it may be more informed as suggested by Friedman.
@@
 The following four effects are noted by \citet[pp.
-14-15]{Delong1990noise}:  the \emph{Friedman Effect}, as informed speculators move prices back towards fundamental value; \emph{Price Pressure} as the speculative noise-traders drive prices towards their misperceived vaue; \emph{Hold More} as noise-traders misperceive the value of the risky asset and tend to hold more of it; \emph{Create Space} is the noise-trader risk caused by the difference between their perception and the actual value, it is an added risk that is caused by the variability of noise-trader belief as noise-traders could drive the price even further from its fundamental value.
-In this model, tHe more numerous the noise-traders, the more volatile prices.
+14-15]{Delong1990noise}:  the \emph{Friedman Effect}, as informed speculators move prices back towards fundamental value; \emph{Price Pressure} as the speculative noise-traders drive prices towards their misperceived value; \emph{Hold More} as noise-traders misperceive the value of the risky asset and tend to hold more of it; \emph{Create Space} is the noise-trader risk caused by the difference between their perception and the actual value, it is an added risk that is caused by the variability of noise-trader belief as noise-traders could drive the price even further from its fundamental value.
+In this model, the more numerous the noise-traders, the more volatile prices.
 
 The Friedman-effect and the Price-pressure are the effects that are studied here.
-Their combination suggests that when the weight of noise-traders is high or the times when they have particularly stong opinions will tend to be those when there is the greatest risk of reversals.
+Their combination suggests that when the weight of noise-traders is high or the times when they have particularly strong opinions will tend to be those when there is the greatest risk of reversals.
 The hold-more effect suggests that the weight of speculative involvement in the market will tend to increase as a trend.
-Create-space effect would suggest that when the weight of noise-traders is high, there is incresaed risk.
+Create-space effect would suggest that when the weight of noise-traders is high, there is increased risk.
 %this can be assessed.
@@
 
-To understand what the noise-trader model of speculative behaviour means for foreign exchange rates, a \emph{Monte Carlo} simulation is carried out with a variety of mis-perceptions and noise-trader weights.
+To understand what the noise-trader model of speculative behaviour means for foreign exchange rates, a \emph{Monte Carlo} simulation is carried out with a variety of misperception-perceptions and noise-trader weights.
 Two time series are created.
@@
 The dividend paid on the risky asset $(r)$ is posted at 10\%; the risk-aversion parameter $(\gamma)$ is assumed to be 0.3 (in line with that suggested in \citet[p.
-127]{FrobesBehaviour}); starting from a price of 100, the average noise-trader mis-perception $(p^*)$ is a random variable in the first simulation with a mean of 4 and a variance $(\sigma_{\rho^*}^2)$ of 2, and has a fixed value of 4 in the second simulation; the weight of noise-traders $(\mu)$ is 50\% in the first case but is modelled as a two-period moving average of a normal random variable centred at 50\% with a standard deviation of 20\% and equally weighted average of its previous value (starting at 50\%) to allow smooth adjustment, in the second simulation.
+127]{FrobesBehaviour}); starting from a price of 100, the average noise-trader misconception $(p^*)$ is a random variable in the first simulation with a mean of 4 and a variance $(\sigma_{\rho^*}^2)$ of 2, and has a fixed value of 4 in the second simulation; the weight of noise-traders $(\mu)$ is 50\% in the first case but is modelled as a two-period moving average of a normal random variable centred at 50\% with a standard deviation of 20\% and equally weighted average of its previous value (starting at 50\%) to allow smooth adjustment, in the second simulation.
 % it is possible to have other specifications. For example, the noise-traders oculd become more prominent if their strategy is more successful,
@@
 A 1000 period prices simulation is created for each model and the returns for these series are calculated as the log change in price.
-Those periods when mis-perception or the weight of speculators is \emph{extreme} are identified as those that are at or above the ninety-fifth percentile.
+Those periods when misperception-perception or the weight of speculators is \emph{extreme} are identified as those that are at or above the ninety-fifth percentile.
 These extremes are the \emph{event} for an event study.
-The R package \emph{eventstudies} createed by \citet{eventstudies} is used for the data manipulation.
+The R package \emph{eventstudies} created by \citet{eventstudies} is used for the data manipulation.
 Descriptive statistics for the whole series of returns and for those when either the extent of their misperception or the weight of noise-traders is extreme are presented in Table \ref{tabref:MC1}.
@@
 This is what would be expected if, as has been identified in many studies, exchange rates follow a random walk.
-\begin{landscape}
-\begin{table}[t!]
+\begin{sidewaystable}
 \begin{threeparttable}
@@
 \end{threeparttable}
-\end{table}
-\end{landscape}
+\end{sidewaystable}
 
@@
 
-\begin{landscape}
-\begin{table}[t]
+\begin{sidewaystable}
 \begin{threeparttable}
@@
 For the risk reversal skew, as a secondary test, a reading above the ninety-fifth and below the fifth percentile is also established as an event.
-The Event Studies R package was used to manage the data see \citet{eventstudies} for fuller details on the package and its use.
+The \citet{eventstudies} R package was used to manage the data.
 The cumulative abnormal returns for the whole of the event window (CARW) and the cumulative abnormal returns for after the event (CARA) were recorded.
@@
 What is more interesting is whether knowledge about the extreme reading at time $t = 0$ provides any information about future returns.
-If it is known that speculative sentiment is at the extreme, does this mean that it is more likely that there will be a reversal?
-Is there an inefficiency?
+The Friedman-effect would suggest a reversal after an extreme.
 This information is provided in columns 5, 8, 11 and 14 of Table \ref{tabref:RR1}.
@@
 
-\begin{landscape}
+\begin{sidewaystable}
 % multirow package is required.   
 %pdflandscape is required to facilitate the landscape position. 
-\begin{table}[t]
 \begin{threeparttable}
@@
 \end{threeparttable}  
-\end{table}
-\end{landscape}
+\end{sidewaystable}
 %This is a section on the methodology that I am writing now (while I remember) so that I can place where necessary.  
@@
 
-\begin{landscape}
+\begin{sidewaystable}
 % multirow package is required.   
 %pdflandscape is required to facilitate the landscape position. 
-\begin{table}[ht]
 \begin{threeparttable}
@@
 \end{threeparttable}  
-\end{table}
-\end{landscape}
+\end{sidewaystable}
 
@@
 Once again there is evidence here of momentum and of speculator positions following the movement of prices.
+This consistent with both the gradual absorption of information into the price and with the feedback of price into additional speculation.
 
-However, there is little evidence that these speculators are driving prices to extremes with their extreme behaviour.
+However, there is little evidence that these speculators are driving prices to extreme deviations from fundamentals with their extreme behaviour.
 Sixty percent of the S1 cases still show a significant abnormal return in the direction of the extreme 4 weeks after the extreme has been reached; only thirty percent of the cases show a reversal and in only one of those do ninety five percent of the re-sampled means remain above zero.
-For S2 it is even more extreme, with ninety percent of the cases showing a continuation and forty percent of them being significant.
+For S2 there is even less ambiguity, with ninety percent of the cases showing a continuation and forty percent of them being significant.
 Again the evidence is consistent with the gradual adjustment of prices to information.
@@
 The event windows are stretched to give a clear view of the evolution of exchange rates  around the event and to ensure that results are not a function of the size of the event window.
-Confidence intervals are added.
-Once event objects FX.exc have been created, parameter for exchange rate, extreme measure, quantile, window length and the high or low indicator can be changed for alternative assessments of speculative extremes.
+Confidence intervals are added\footnote{It is possible to make changes to the estimate of the \emph{expected exchange rate}, the extreme measurement, quantile, window length and the high or low indicator can be changed for alternative assessments of speculative extremes.
 Adjustments to these parameters show that these results are not a function of the choices presented in the main text.
-They are robust to alternative specifications.
+They are robust to alternative specifications.}.
 
@@
 This means that the speculators are inclined to be short the non-domestic (US dollar) currency.
-While the Euro continues to move lower after the extreme short positions, the other major currencies seem to reveal a period of consolidation that is consistent with the settlement around an equilibrium after a period where information is being absorbed.
-The contrast between the Euro and the other currencies may be associated with the fact that there is a smaller dataset for the Euro due to its more limited lifespan.
+While the Euro continues to move lower after the extreme short positions, the other major currencies move broadly sideways, consistent with the settlement around an equilibrium after a period where information is being absorbed.
+This is also consistent with market efficiency where a random walk would be expected after an extreme.
+
 Figure \ref{fig:ES4} shows the abnormal returns around extreme high (90th percentile and long position).
@@
 The measurement of speculation takes two forms:  a measurement of the intensity of belief with the risk-reversal skew, and a measurement of speculative positions.
-Neither of these is ideal, in the sense of being a random sample from a population.
-However, the option prices reveal the intensity of feeling through the weight of money in a speculative part (options) of a speculative market (foreign exchange).
+The option prices reveal the intensity of feeling through the weight of money in a speculative part (options) of a speculative market (foreign exchange).
 The commitment of traders report records only speculators in the US futures market.
@@
 
- there is strong evidence that speculation affects prices and that this effect is gradual rather than instantaneous.
-This suggests a market inefficiency that would reward momentum-based trading.
+There is strong evidence that speculation affects prices and that this effect is gradual rather than instantaneous.
+This consistent with a market inefficiency that would reward momentum-based trading.
 It is no wonder therefore that \citet{TaylorTechnical}, reporting a survey conducted by the Bank of England, found that of the Chief foreign exchange dealers contacted in London in November 1988, 90\% of them attached some weight to \emph{technical analysis}, a method that is fundamentally based on finding trends in asset prices.
@@
 The importance of momentum trading and technical analysis has not waned with time.
-Neely and Wller  show that gains from technical trading rules still generate excess returns and attribute this to behavioural biases \citet{NeelyWeller}.
+Neely and Weller  show that gains from technical trading rules still generate excess returns and attribute this to behavioural biases \citet{NeelyWeller}.
 It seems that speculation forms part of the process of price discovery, but this process is not instantaneous as would be expected in the purest form of the EMH.
@@
 However, the combination of high level of speculation and divergence from fundamentals is likely to create the greatest risk.
+
 \bibliography{../../../../myrefs}
+\begin{appendices}
 
-\begin{appendices}
 \section{Appendix 1}\label{apref:1}
-\begin{landscape}
-\begin{table}[t]
+
+\begin{sidewaystable}
 \begin{threeparttable}
 \caption{Descriptive Statistics for Daily Exchange Rate Returns: 2 January 1996 to 6 September 2002}
+ \label{tabref:DS1}
 \begin{tabular}{l | r r  r r r r r r r r}
@@
 \hline
- \label{tabref:DS1}
 \end{tabular}
@@
 \end{threeparttable}
-\end{table}
-\end{landscape}
+\end{sidewaystable}
+
 \section{Appendix 2}
- \begin{table}[t]
+
+\begin{sidewaystable}
 \begin{threeparttable}
@@
 \end{threeparttable}
-\end{table}
+\end{sidewaystable}
 

### 2015-07-13 Knitr/NewPaper.Rnw
Completed writing NewPaper.Rnw There is still a problem with the appendix.

@@
 Speculation does not appear to be random, there is momentum to the rise and fall in sentiment and positions.
-However, while currency prices co-move with sentiment and positions, there is no informational content from either: knowing the the strength of speculative opinion or the weight of speculative positions does not improve forecasts of future returns.
+However, while currency prices co-move with sentiment and positions, there is no informational content from either: knowing  the strength of speculative opinion or the weight of speculative positions does not improve forecasts of future returns.
 This supports the suggestion that speculative activity has a permanent affect on prices and that is consistent with speculation being part of the process of price discovery.
-Speculators do not seem to drive prices from fundamental value and their presence does not seem to hamper the incorporation of information into prices.
+Speculators do not seem to drive prices from fundamental value and their presence seems to help rather than hinder the incorporation of information into prices.
 
@@
 
-% does this following paragraph come later?  I think so....Once the three
-% main categories are done. 
 In the \emph{microstructure} literature, the term noise-trader is often interchanged with \emph{liquidity-trader}.
-They do not usually distinguish between psychological and institutional reasons for the lack of attention to fundamental information.
+They do not usually distinguish between institutional and psychological reasons for the lack of attention to fundamental information.
 Nonetheless, liquidity tends to refer to institutional explanations for trading without reference to fundamental value.
@@
 Therefore, noise makes financial markets possible but imperfect: the more noise-traders that there are, the more liquidity and the easier it is to trade; the more noise-traders, the higher the level of inefficiency in the market and the more likely that price will be driven from fundamental value.
+This begs the question: are speculators trading on noise or are they informed?
 
@@
 In the original version of rational expectations,  \citet{muth1961Rational} uses commodity speculation as the means by which prices are driven towards the level where expectations are model-consistent.
-These are of Friedman-type speculators as they base their expectations on their knowledge of the underlying model and they are the means by which the components of the model tend towards equilibrium values.
-% See \citet gc{Lyons2006Microstructure} and \citep{Evans2002Order} for examples of the micrstrucure approach in foreign exchange.
+These are of Friedman-type speculators as they base their expectations on their knowledge of the underlying model and they are the mechanisem through which equilibrium is achieved.
 
-However, though the idea that noise is a random variable may be a construction that is useful for modelling and it may be justified in analysis of broader themes of microstructure and levels of market efficiency, this exogenous construct may be an over-simplification that is inappropriate when looking for more understanding of speculation itself.
-Institutional and psychological explanations for noise-trading frequently suggest that there are endogenous non-random elements.
+However, though equating noise and noise-trading as a random variable may be a construction that is useful for modelling and it may be justified in analysis of broader themes of microstructure and levels of market efficiency, it may also be an over-simplification that is inappropriate when looking for more understanding of speculation itself.
+Institutional and psychological explanations for noise-trading frequently suggest that there are endogenous non-random elements to speuclative activity.
 
@@
 See \citet{BrunnermeierLiquidity} and \citet{Gorton} for an overview of the institutional literature.
-Psychological explanations revolve around the common panic, the group-thinking and the tendency to read private information from the actions of other investors: the so-called \emph{wisdom of crowds} depends on independent evaluation but as Keynes noted in his discussion of the \emph{beauty contest}, it is generally more important to identify what other investors will believe than actual underlying, fundamental value \citet[p.101]{Keynes1936}
+Psychological explanations revolve around the common panic, the group-thinking and the tendency to read private information from the actions of other investors: the so-called \emph{wisdom of crowds} depends on independent evaluation but as Keynes noted in his discussion of the \emph{beauty contest}, it is generally more important to identify what other investors will believe than actual underlying, fundamental value \citet[p.101]{Keynes1936}.
 
-it has also been suggested that speculative techniques can change overtime.
-For example, \citet{ShillerFashion} discusses the use of fashionable techniques or fads in investment.
-In this way, speculation noise is associated with sentiment and measures of sentiment are then used as ways of measuring speculation.
-%It would be good to have some more modern ideas here.  Maybe they are just the empirical results. 
-%this paragraph is not very good.  It needs to be updated.  Look at the file in citulike. 
 \subsection{Empirical work}
@@
 The latter group may include day-traders or private individuals.
+It has also been suggested that speculative techniques can change over time.
+For example, \citet{ShillerFashion} discusses the use of fashionable techniques or fads in investment.
+In this way, speculation noise is associated with sentiment and measures of sentiment are then used as ways of measuring speculation.
 
-%The literature on heterogeneous traders, another way that the trading securities is explained, has very frequently used noise-traders as trend-chasers or followers of technical analysis.
+%Noise stragegies may be successful for a while.
+Momentum trades tend to be successful for certain ties.
+Also discussion the literature on heterogeneous traders, another way that the trading securities is explained, has very frequently used noise-traders as trend-chasers or followers of technical analysis.
 These models (see for example brock and hommes [1998]; lux [1995]; lux and marchesi [2000]) investigate bubbles and crashes, and feature fundamental traders versus noise traders.
@@
 The Keynesian view of speculation is that it is an addition to market noise that makes it even harder for informed-traders to overcome uncertainty and apply fundamental information about long-term value.
+As such, it exacerbates the tendency towards putting too much emphasis on the immediate and works against longer-term and strategic thinking.
 This description provides an explanation for Shiller's observation that stock prices are more volatile than the underlying earnings information on which valuation should be based \citet{ShillerVol}.
@@
 There are a number of contemporary questions where the role and nature of speculation is at the centre of the debate.
-Regulations or taxes that reduce the amount of speculation have generally been considered to be beneficial as they reduce risk and encourage more long-term attitudes towards investment.
+For example, regulations or taxes that reduce the amount of speculation have generally been considered to be beneficial as they reduce risk and encourage more long-term attitudes towards investment.
 However, where these measures reduce liquidity, higher risk and an increase in other costs may provide an countervailing force.
@@
 
-%In addition, so-called \emph{Volcker Rule} would prevent US investment banks trading on their own-account.
-This, it has been argued, will reduce liquidity in the bond market by causing investment banks to hold lower levels of bond inventory on their books.
-This may increase the amount of \emph{price slippage} when final investor adjust their positions.
-In addition,
+%volcker rule?
 
@@
 The aim of this study is to test whether the noise-trader model provides an appropriate explanation for speculator activity in the foreign exchange market.
-This will be use to gain a fuller understanding of speculative behaviour.
-The experiment will involve three steps:  assess the price implication of the noise-trader model;  measure speculative activity in the foreign exchange market; finally, assess whether extreme speculative activity in the foreign exchange market is consistent with the noise-trader model.
+This will be used to gain a fuller understanding of speculative behaviour.
+The analysis will involve three steps:  assessement of the price implication of the noise-trader model;  measurement of speculative activity in the foreign exchange market; finally, assessement of whether extreme speculative market activity is consistent with the noise-trader model.
 If the speculative activity that is seen in the foreign exchange market produces results that are like the noise-trader model, it can be said that speculation is noise trading as Keynes asserted, if not it may be more informed as suggested by Friedman.
-% where it is widely perceived that there are a relatively large number of traders engaged in technical analysis.
-See \citet gc{FrootChartists} for a survey.
-%this may go in the explanation for why fx is used.
-Use this elsewhere.
-Add more?
 
 The \citet{Delong1990noise} noise-trader model used in this study.
-This is a derivative of the \citet{Samuelson1958} overlapping generations model  with two-period agents making a portfolio choice.
-There are two assets:  a safe asset $(s)$ which pays a set dividend $(d)$ and is in perfectly elastic supply; the risky asset $(u)$ also pays the same dividend but is not in elastic supply.
-There are two types of investor or trader: sophisticated traders $(i)$ who know the fundamental value of the risky assets and noise-traders $(n)$ who misperceive the fundamental value.
+This is a derivative of the \citet{Samuelson1958} overlapping generations model with two-period agents making a portfolio choice between two assets: a safe asset $(s)$ which pays a set dividend $(d)$ and is in perfectly elastic supply; a risky asset $(u)$ also pays the same dividend but is not in elastic supply.
+There are two types of investor: sophisticated traders $(i)$ who know the fundamental value of the risky assets and noise-traders $(n)$ who misperceive the fundamental value.
 In the absence of noise-traders, the assets are substitutes.
@@
 
-the final price equation becomes.
+the final price equation becomes:
 
 \begin{equation}
-p_t = 1 + \frac{\mu(\rho_t - \rho^*)}{1 + r} + \frac{ \mu \rho^*}{r} - \frac{2 \gamma \mu^2 \sigma_{\rho}^2}{r(1 + r^2)}
+p_t = 1 + \frac{\mu(\rho_t - \rho^*)}{1 + r} + \frac{ \mu \rho^*}{r} - \frac{2 \gamma \mu^2 \sigma_{\rho}^2}{r(1 + r^2)}.
 \label{eq:NP2}
@@
 The Friedman-effect and the Price-pressure are the effects that are studied here.
-Their combination suggests that when the weight of noise-traders is high or the times when they have particularly strong opinions will tend to be those when there is the greatest risk of reversals.
+Their combination suggests that when the weight of noise-traders is high or the times when they have particularly strong opinions will tend to be those when there is the greatest risk of reversal.
 The hold-more effect suggests that the weight of speculative involvement in the market will tend to increase as a trend.
@@
 
-The main findings are that noise-traders will not necessarily disappear as Friedman had suggested \citet{FriedmanPositive}.
-While noise-traders will tend to get less utility than sophisticated traders, this does not necessarily drive them from the market.
-The noise-traders expected to get higher utility due to their misperception.
-The reduction in utility comes mainly from the fact that they are not compensated for taking risk that they are not aware of.
-This is consistent with the type of over-confidence that has been found by \citet{odean1998volume}in day-traders.
-
 \subsection{The implications of the noise-trader model}
 There are two parameters from the noise-trader price equation \ref{eq:NP2} that can be used to measure the amount of speculation:  the price misperception $(\rho)$ and the weight of noise-traders in the market $(\mu)$.
-the aim of this section is to  identify the returns that are implied by the noise-trader model when either the intensity of the misperception or the weight of noise-traders is increased to extreme levels so that this can be used as a benchmark against which foreign exchange market activity can be judged.
+The aim of this section is to  identify the returns that are implied by the noise-trader model when either the intensity of the misperception or the weight of noise-traders is increased to extreme levels so that this can be used as a benchmark against which foreign exchange market activity can be judged.
 
@@
 
-% this needs to be completed.
-
 \subsection{Measuring the intensity of speculative sentiment}
@@
 
-% The table may be pushed onwards.  Use t (i think).
-
 Table \ref{tabref:DS1} in Appendix \ref{apref:1} shows descriptive statistics for the exchange rate data used in the study.
@@
 The first thing of note is that the mean return does not appear to be statistically different from zero.
-This is what would be expected if, as has been identified in many studies, exchange rates follow a random walk.
+This is what would be expected if, as has been identified in many studies, exchange rates follow something close to a random walk.
 \begin{sidewaystable}
@@
 
-%The operation of derivative exchanges is a competitive business that has evolved over the course of the data collection period.
-The two original exchanges in Chicago, the Chicago Mercantile Exchange (CME) and the  Chicago Board of Trade (CBOT) went public in 2000 and 2005 respectively and merged in 2007 to become the CEM Group.
-In 2008 the New York Mercantile Exchange (NYMEX) was added,  Most of the successful currency derivative contracts were traded on the International Monetary Market (IMM) which was a sub-division of the original CME.
-Many exchange rates and cross rate products were launched by these and other US exchanges, but the five contract which maintained a presence and liquidity through the whole period were the Deutschmark and then the Euro, the Swiss Franc, the British Pound, the Japanese Yen and the Canadian Dollar.
-There is a reasonably long series for the Australian Dollar and for the Mexican Peso, but nether of these is complete.
-The US dollar index has been active since it was launched in May 1992 but it is not trade weighted or really representative of the US dollar.
-
 The data used run from September 30 1998 when the data started to be released on a weekly basis and continue up to December 31 2008.
-This date is shortly before this part of the research was started.
 The Euro data runs from March 3 1998 to December 31 2008.
@@
 It should not be surprising to see that the extreme highs are associated with positive returns, which appear in many cases to be significantly different from zero (where significance is identified as more than 95\% of 1000 random bootstrap samples taken from the events being above zero).
+
 It is clear that, as speculators become more extreme in their opinion, prices are driven in the direction of sentiment.
@@
 \end{sidewaystable}
-%This is a section on the methodology that I am writing now (while I remember) so that I can place where necessary.  
-%================Event Study Method=====================================
-% The dates when the risk reversal skew is at or above the ninety fifth (ninetieth) or at or below the first (tenth percentile are identified and a matrix of returns is created with the rows being equal to plus and minus the size of the event window and the date that the extreme was recorded being equal to zero.
-The number of columns is equal to the number of events which, given that we are looking at the extreme one or five percent, will be related to the sample size.
-The average of the returns for each event day are then recorded.
-As such, with an event window of four days, there is an average return four days before the event window and for three days before the event window, through to four days after the event window.
-To get an idea of the precision of this estimate, the returns from each event day are sampled (with replacement) one thousand times and the mean is calculated.
-The dispersion of the resulting means provide the confidence intervals for the estimate of the mean returns for each event day that has been made.
 
@@
 However, it does indicate that the evolution of exchange rates are not a random walk.
-There is evidence here that prices gradually respond to information, as would be the case if decision-making is following a \emph{conservatism} heuristic or if there were institutional features.
+There is evidence here that prices gradually respond to information, as would be the case if there were institutional or psychological features that caused momentum or gradual adjustment.
 The positive (or negative) expectations that are evident in the risk reversal skew do not seem to represent a deviation or a mis-pricing that will be reversed.
@@
 The importance of momentum trading and technical analysis has not waned with time.
-Neely and Weller  show that gains from technical trading rules still generate excess returns and attribute this to behavioural biases \citet{NeelyWeller}.
+The more up-to-date CFTC data show the same pattern.
+In addition,  \citet{NeelyWeller} show that gains from technical trading rules still generate excess returns and attribute this to behavioural biases.
 It seems that speculation forms part of the process of price discovery, but this process is not instantaneous as would be expected in the purest form of the EMH.
@@
 This would be consistent with a model where extreme price movement was a function of new information rather than the structure of existing positions.
-%expand.  Could put data in appendix.
-
-The evidence here, that speculation is associated with serial correlation in the returns but cannot be used in a systematic way to determine reversals, has important implications for speculative activity.
-It means that the largest returns will be to those that identify the trend at the earliest opportunity.
-This means that speculators must be prepared to act as swiftly as possible, probably when there is a high level of uncertainty and a lack of information and where any scrap of additional news is highly valued.
-The risk that actions by other speculators, most notable by their effect on price, will be construed to contain information exacerbates the likelihood of feedback effects.
-The need to be able to take action before others has reached the most extreme form in the activity of \emph{high frequency trading} (HFT), where decision times are measured in nanoseconds and there is advantage in gaining spacial proximity to exchanges.
-HFT is all about getting  a larger share of the inefficiency in pricing information by trading or investing closer to where the information is released.
 
@@
 \begin{sidewaystable}
+
 \begin{threeparttable}
@@
 
-\begin{sidewaystable}
 \begin{threeparttable}
@@
 \end{threeparttable}
-\end{sidewaystable}
 

### 2015-07-16 Knitr/NewPaper.Rnw
This paper has been completed NewPaper.Rnw.  However, there is some issue with the appendix. I think that is to do with the table.

@@
 There are a range of ideas about the way that speculation can affect financial markets.
-Towards one end of this spectrum is the view of Keynes from Chapter 12 of \emph{The General Theory} where he identified speculation as a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd'' \citet[p.
-101]{Keynes1936}.
+Towards one end of this spectrum is the view of  \citet[p.
+101]{Keynes1936} in Chapter 12 of \emph{The General Theory} where he identified speculation as a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd''.
 Here speculation is a problem that, at its most extreme, will cause booms, reversals and will discourage fundamental, long-term investment.
@@
 The bid-ask spread is the payment that the market-maker receives in return for taking the risk of meeting an informed trader.
-Ultimately, the uninformed traders make payments to the informed traders with the mark-makers taking a cut for the risk and the provision of exchange.
+Ultimately, the uninformed traders make payments to the informed traders with the market-makers taking a cut for the risk and the provision of exchange.
 See also \citet{glosten1985bid}.
@@
 In the \emph{microstructure} literature, the term noise-trader is often interchanged with \emph{liquidity-trader}.
-They do not usually distinguish between institutional and psychological reasons for the lack of attention to fundamental information.
+There may be institutional and psychological reasons for the lack of attention to fundamental information.
 Nonetheless, liquidity tends to refer to institutional explanations for trading without reference to fundamental value.
 This may mean that economic agents just execute customer orders at the current price or it may be that they buy and sell with reference to a benchmark index or to fulfil a hedging strategy.
+Noise is more psychological.
 
@@
 In this case, uninformed speculation provides a rational and an opportunity for the trading of financial securities.
-If all economic agents know the fundamental value of a security how can some be buying and some be selling?
 There must be different opinions.
 Why will traders be prepared to pay the cost of acquiring information if they are unable to take advantage of that information?
-With speculative noise-traders, the informed can take advantage of their information.
+With speculative noise-traders, prepared to take the other side of the transaction, the informed can take advantage of their information.
 This creates a justification for acquiring information.
 Speculative noise-traders can therefore create a gap between the price and the fundamental value that will encourage informed traders to pay the cost of acquiring information.
-The divergence between this price and fundamental value is thereby essential to encourage the process of price discovery.
-%
 \citet[p.529]{BlackNoise} suggested that this combination was essential to the smooth function of financial markets.
@@
 In the original version of rational expectations,  \citet{muth1961Rational} uses commodity speculation as the means by which prices are driven towards the level where expectations are model-consistent.
-These are of Friedman-type speculators as they base their expectations on their knowledge of the underlying model and they are the mechanisem through which equilibrium is achieved.
+These are of Friedman-type speculators as they base their expectations on their knowledge of the underlying model and they are the mechanism through which equilibrium is achieved.
 
 However, though equating noise and noise-trading as a random variable may be a construction that is useful for modelling and it may be justified in analysis of broader themes of microstructure and levels of market efficiency, it may also be an over-simplification that is inappropriate when looking for more understanding of speculation itself.
-Institutional and psychological explanations for noise-trading frequently suggest that there are endogenous non-random elements to speuclative activity.
+Institutional and psychological explanations for noise-trading frequently suggest that there are endogenous non-random elements to speculative activity.
 
@@
 See \citet{BrunnermeierLiquidity} and \citet{Gorton} for an overview of the institutional literature.
-Psychological explanations revolve around the common panic, the group-thinking and the tendency to read private information from the actions of other investors: the so-called \emph{wisdom of crowds} depends on independent evaluation but as Keynes noted in his discussion of the \emph{beauty contest}, it is generally more important to identify what other investors will believe than actual underlying, fundamental value \citet[p.101]{Keynes1936}.
+Psychological explanations revolve around the common panic, the group-thinking and the tendency to read private information into the actions of other investors: the so-called \emph{wisdom of crowds} depends on independent evaluation but as Keynes noted in his discussion of the \emph{beauty contest}, it is generally more important to identify what other investors will believe than to form an independent view of the actual underlying, fundamental value \citet[p.101]{Keynes1936}.
 
@@
 For example, \citet{ShillerFashion} discusses the use of fashionable techniques or fads in investment.
-In this way, speculation noise is associated with sentiment and measures of sentiment are then used as ways of measuring speculation.
+In this way, speculation can be be associated with sentiment and measures of sentiment are then used as ways of measuring speculation.
 
@@
 As such, it exacerbates the tendency towards putting too much emphasis on the immediate and works against longer-term and strategic thinking.
-This description provides an explanation for Shiller's observation that stock prices are more volatile than the underlying earnings information on which valuation should be based \citet{ShillerVol}.
+This description provides an explanation for the observation from \citet{ShillerVol} that stock prices are more volatile than the underlying earnings information on which valuation should be based.
 In addition, \citet{Delong1990noise} and \citet{shleifer1997limits} show that constrained horizons can cause noise-trading to limit the ability of informed traders to take advantage of profitable opportunities for fear of further divergence from fundamental value before realignment is achieved.
@@
 There are a number of contemporary questions where the role and nature of speculation is at the centre of the debate.
-For example, regulations or taxes that reduce the amount of speculation have generally been considered to be beneficial as they reduce risk and encourage more long-term attitudes towards investment.
+For example, regulations or taxes that reduce the amount of speculation have generally been considered to be beneficial as they limit risk-taking and encourage more long-term attitudes towards investment.
 However, where these measures reduce liquidity, higher risk and an increase in other costs may provide an countervailing force.
@@
 The aim of this study is to test whether the noise-trader model provides an appropriate explanation for speculator activity in the foreign exchange market.
-This will be used to gain a fuller understanding of speculative behaviour.
-The analysis will involve three steps:  assessement of the price implication of the noise-trader model;  measurement of speculative activity in the foreign exchange market; finally, assessement of whether extreme speculative market activity is consistent with the noise-trader model.
-If the speculative activity that is seen in the foreign exchange market produces results that are like the noise-trader model, it can be said that speculation is noise trading as Keynes asserted, if not it may be more informed as suggested by Friedman.
+The analysis will involve three steps:  assessment of the price implication of the noise-trader model;  measurement of speculative activity in the foreign exchange market; finally, assessment of whether extreme speculative market activity is consistent with the underlying model that is used.
+If the speculative activity in the foreign exchange market produces results that are like the noise-trader model, it can be said that speculation is closer to the Keynesian view than that of Friedman.
 
 The \citet{Delong1990noise} noise-trader model used in this study.
-This is a derivative of the \citet{Samuelson1958} overlapping generations model with two-period agents making a portfolio choice between two assets: a safe asset $(s)$ which pays a set dividend $(d)$ and is in perfectly elastic supply; a risky asset $(u)$ also pays the same dividend but is not in elastic supply.
+This is a derivative of the \citet{Samuelson1958} overlapping generations model with two-period agents making a portfolio choice between two assets: a safe asset $(s)$ which pays a set dividend $(d)$ and is in perfectly elastic supply; a risky asset $(u)$ which also pays the same dividend but is not in elastic supply.
 There are two types of investor: sophisticated traders $(i)$ who know the fundamental value of the risky assets and noise-traders $(n)$ who misperceive the fundamental value.
@@
 
-Maximising utility with respect to holdings of the risky asset and iterating backward to solve for the current value,  the price becomes the discounted value of the future dividends, the expected capital gain and the general misperception of noise-traders less risk.
+Maximising utility with respect to holdings of the sage and risky assets for the informed and uninformed traders and iterating backward to solve for the current value,  the price becomes the discounted value of the future dividends, the expected capital gain and the general misperception of noise-traders less risk.
 
@@
 The following four effects are noted by \citet[pp.
-14-15]{Delong1990noise}:  the \emph{Friedman Effect}, as informed speculators move prices back towards fundamental value; \emph{Price Pressure} as the speculative noise-traders drive prices towards their misperceived value; \emph{Hold More} as noise-traders misperceive the value of the risky asset and tend to hold more of it; \emph{Create Space} is the noise-trader risk caused by the difference between their perception and the actual value, it is an added risk that is caused by the variability of noise-trader belief as noise-traders could drive the price even further from its fundamental value.
+14-15]{Delong1990noise}:  the \emph{Friedman Effect}, as informed speculators move prices back towards fundamental value; \emph{Price Pressure} is the \emph{noise-trader risk} as the speculative noise-traders drive prices towards their misperceived value; \emph{Hold More} as noise-traders misperceive the value of the risky asset and tend to hold more of it; \emph{Create Space} is the difference between noise-trader perception and the actual value, it creates the opportunity for informed traders to profit from their information.
 In this model, the more numerous the noise-traders, the more volatile prices.
@@
 The hold-more effect suggests that the weight of speculative involvement in the market will tend to increase as a trend.
-Create-space effect would suggest that when the weight of noise-traders is high, there is increased risk.
-%this can be assessed.
+Create-space effect would suggest that when the weight of noise-traders is high and extreme, there is greater incentive and opportunity for the informed to take advantage of the divergence between fundamental value and price.
 
@@
 These extremes are the \emph{event} for an event study.
-The R package \emph{eventstudies} created by \citet{eventstudies} is used for the data manipulation.
 Descriptive statistics for the whole series of returns and for those when either the extent of their misperception or the weight of noise-traders is extreme are presented in Table \ref{tabref:MC1}.
@@
 Risk-reversal skew is the relative price of currency put and call options for the same strike price.
-The \emph{Put-Call-Parity} in the  \citet{Black1973Options} option pricing model assumes that these will be the same.
-While the Black-Scholes model is the basis for valuing options, where the price of a non-dividend paying stock trades at a discount to the forward rate equal to the risk free rate, the forward rate in a currency option is determined by the relative interest rate differential between the two currencies.
+The \emph{Put-Call-Parity} in the  \citet{Black1973Options} option pricing model assumes that these will be the same, in practice they are not\footnote{While the Black-Scholes model is the basis for valuing options, where the price of a non-dividend paying stock trades at a discount to the forward rate equal to the risk free rate, the forward rate in a currency option is determined by the relative interest rate differential between the two currencies.
 The \citet{GarmanKohlhagan} adjustment that accounts for the difference in interest payments on the two currencies is used.
@@
 This model allows the market practice of pricing options from the volatility implied by the model.
-In other words, the Garman-Kohlhagan is used to extract the volatility that is implied by a specific option price given the strike and other parameters.
+In other words, the Garman-Kohlhagan is used to extract the volatility that is implied by a specific option price given the strike and other parameters}.
 
 This study uses a unique dataset of \emph{risk-reversals} for eleven major currency pairs.
-This is the only period for which the data currently exist.
+It appears that this is the only period for which the data currently exist.
 The risk-reversal data is very difficult to obtain in an over-the-counter market like foreign exchange where there is no centralised exchange to record and collect information, generally being the property of quoting institutions.
@@
 Alternatively, the dealer would sell a 25 delta call at 8.78\% and buy the USD-call-CAD-put at 8.50\%, earning 0.28\% \citet{Global}.
-}  Therefore, the relationship between implied volatility for calls and implied volatilities for puts for the same delta and maturity can be taken as an indication of how much more expensive calls are relative to the equivalent put or as the market bias towards puts over calls or, in this case, as an indication of market sentiment.
-The larger the absolute level of the risk-reversal skew, the greater the relative price and, in equilibrium, the greater the belief, expressed through the action of market participants or the inventory of market-makers, that there will be a deviation from the forward rate.
-If extreme deviations or speculations are noise they should presage a sharp reversal.
-If speculation is informed, this is part of the price discovery process and extremes may be followed by a random walk.
+}  Therefore, the relationship between implied volatility for calls and implied volatilities for puts for the same delta and maturity is an indication of how much more expensive calls are relative to the equivalent put or as the market bias towards puts over calls and as an indication of market sentiment.
+The larger the absolute level of the risk-reversal skew, the greater the relative price and, in equilibrium, the more intense the belief, expressed through the action of market participants or in the inventory of market-makers, that there will be a deviation from the forward rate.
+If sentiment is based   on noise rather than information, extremes should presage a sharp reversal; if speculative sentiment is informed, this is part of the price discovery process and extremes may be followed by a random walk.
 
@@
 Table \ref{tabref:DS2} gives the descriptive statistics for the risk-reversal data.
-For a neutral risk-reversal, a reading of zero, there would be no price preference for calls or puts at the similar delta.
+For a neutral risk-reversal, a reading of zero, there would be no price preference for calls or puts for the same delta.
 The figures show that, in contrast to the average returns, there are often average skews that are different from zero.
@@
 For the risk reversal skew, as a secondary test, a reading above the ninety-fifth and below the fifth percentile is also established as an event.
-The \citet{eventstudies} R package was used to manage the data.
+The  R package \emph{eventstudies} \citet{eventstudies} was used to manage the data.
 The cumulative abnormal returns for the whole of the event window (CARW) and the cumulative abnormal returns for after the event (CARA) were recorded.
@@
 This means that a sample of size 106 (in this case), with replacement, is drawn from each of the event days and new means are calculated for the CARW and CARA.
-If 95\% of the means that are calculated after drawing random samples from the array of windows are above or below zero, it is assumed that the figure is statistically different from zero, and this is represented by an asterisk in the table.
+If 95\% of the means that are calculated after drawing random samples from the array of windows are above or below zero, it is asserted that the figure is statistically different from zero, and this is represented by an asterisk in the table.
 
@@
 What is more interesting is whether knowledge about the extreme reading at time $t = 0$ provides any information about future returns.
-The Friedman-effect would suggest a reversal after an extreme.
+If this speculation is noise, Friedman's informed traders should take advantage of that and a reversal should be seen.
 This information is provided in columns 5, 8, 11 and 14 of Table \ref{tabref:RR1}.
@@
 Of the eighteen cases using the 95\% th percentile to identify an extreme, ten show a continuation, rather than a reversal, that is not significantly different from zero; one shows a significant continuation; five show reversals that are not significant; only two show significant reversal (EURCHF in each case).
-For the more extreme event when the 99\% the percentile is used to measure the extreme, there is an even more mixed picture.
+For the more extreme event when the 99\% the percentile is used to measure the extreme, there is even more ambiguity.
 Just over half the cases with the four-day window show a reversal (8 of 18), but only 2 of these are statistically significant.
@@
 Additional cases were run using percentiles at 80\% and 99.5\%.
-For the event window, figures \ref{fig:ES1} and \ref{fig:ES2} show the cumulative abnormal returns for the major exchanges rates relative to the US dollar across a 16 day event window.
+For the event window, Figures \ref{fig:ES1} and \ref{fig:ES2} show the cumulative abnormal returns for the major exchanges rates relative to the US dollar across a 16 day event window.
 Figure \ref{fig:ES1} shows the extreme highs for as a $99^{th}$ percentile and figure \ref{fig:ES2} shows the extreme low for as a $10^{th}$ percentile.
@@
 The event windows are stretched to give a clear view of the evolution of exchange rates  around the event and to ensure that results are not a function of the size of the event window.
-Confidence intervals are added\footnote{It is possible to make changes to the estimate of the \emph{expected exchange rate}, the extreme measurement, quantile, window length and the high or low indicator can be changed for alternative assessments of speculative extremes.
+Confidence intervals are added\footnote{It is possible to make changes to the estimate of the \emph{expected exchange rate}, the extreme measurement, quantile, window length and the high or low indicator and for alternative assessments of speculative extremes to be used.
 Adjustments to these parameters show that these results are not a function of the choices presented in the main text.
@@
 The commitment of traders report records only speculators in the US futures market.
-However, this is an important market, there is no reason to believe that US or speculators in the futures market are systematically different from those in the rest of the world and presumably if there were a major difference there would be an arbitrage opportunity between the two markets.
+However, this is an important market, there is no reason to believe that US speculators in the futures market are systematically different from those in the rest of the world and presumably if there were a major difference there would be an arbitrage opportunity between the two markets.
 
@@
 The survey found that the importance of technical analysis became greater at shorter horizons.
-This has not disappeared.
 The importance of momentum trading and technical analysis has not waned with time.
@@
 In addition,  \citet{NeelyWeller} show that gains from technical trading rules still generate excess returns and attribute this to behavioural biases.
-It seems that speculation forms part of the process of price discovery, but this process is not instantaneous as would be expected in the purest form of the EMH.
+It seems that speculation forms part of the process of price discovery, but this process is not instantaneous as would be expected if the market were informationally efficient.
 
@@
 The speculators may be just part of the adjustment to new information.
-However, the combination of high level of speculation and divergence from fundamentals is likely to create the greatest risk.
 

### 2015-09-08 Knitr/NewPaper.Rnw
Extreme2.R is the latest on extremes This needs to be developed so that there are distributions. The paper has been re-structured.

@@
 \begin{abstract}
-An event study is used to assess the views of Keynes and Friedman on speculation.
-Speculative extremes are ranked by intensity of sentiment and weight of speculative activity.
-A unique dataset of risk-reversal skew on option prices is used to measure the intensity of speculative sentiment; the weight of positions is based on US regulatory data.
-The extremes say little about the future in either case, suggesting that speculation is more than just random noise and supporting the view there is some informational content that is being passed through to the price with speculative activity.
+  If speculation is just noise, price movements associated with extremes of speuclative sentiment or heavy involvement of speculative activity should be reveversed; if speculation is informed, associated price movement is more likely to be sustained.
+A unique set of risk-reversals on currency options is used to measure speculative sentiment and regulatory positions measure speculative activity.
+An event study shows that extremes do not increase the probability of a price reversal, supoprting the view that speculation is part of the price disovery process.
 \end{abstract}
 \section*{Introduction}
-There are a range of ideas about the way that speculation can affect financial markets.
-Towards one end of this spectrum is the view of  \citet[p.
-101]{Keynes1936} in Chapter 12 of \emph{The General Theory} where he identified speculation as a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd''.
-Here speculation is a problem that, at its most extreme, will cause booms, reversals and will discourage fundamental, long-term investment.
-It is noise that does not emerge randomly, but feeds off itself in an evolution of collective sentiment that encourages trends in asset prices.
+Ideas about the way that speculation affects financial market span a broad range.
+Towards one end is the Keynesian view that speculation is a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd'' \citet[p.
+101]{Keynes1936}.
+Speculation is a problem that will cause booms, reversals and will discourage fundamental, long-term investment; It is noise that emerges from the interaction of collective sentiment to encourages trends in asset prices.
 
-Towards the other end of the spectrum, Milton Friedman argued that speculation was a stabilising force, providing liquidity and helping to ensure that markets swiftly find equilibrium.
-In arguing in favour of flexible exchange rates,  \citet{FriedmanPositive} asserted that successful speculators would be those who managed to correctly buy when prices  were below fundamental value and sell above that price.
-Therefore, Friedman argued, the speculative process would favour those able to access or utilise information over the uninformed and, through a Darwinian process, informed speculation would dominate.
+Towards the other end of the spectrum, Friedman argued that speculation was a stabilising force, providing liquidity and helping to ensure that markets swiftly find equilibrium.
+\citet{FriedmanPositive} asserted that successful speculators would be those who managed to correctly buy when prices  were below fundamental value and sell above that price.
+Therefore, Friedman argued, the speculative process would reward those able to access or utilise information relative to the uninformed and, through a Darwinian process, informed speculation would dominate.
 This view regards speculation as an key lubricant of financial markets as it provides liquidity and facilitates the process of price discovery.
 
-The aim of this paper is to understand more about the nature of speculation by investigating the relationship between both the intensity of speculation and the weight of speculative positions with the movement of prices in the foreign exchange market.
-If speculators are uninformed and their activity can be regarded as noise, the most extreme sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when reversals are most likely to take place; if speculators are informed prices should follow a random walk after either of these extremes.
-
-The foreign exchange market is utilised because it is the largest and most liquid financial market in the world.
+The aim of this paper is to understand more about the nature of speculation by investigating the relationship between the intensity of speculation and the weight of speculative positions relative to the movement of prices in the foreign exchange market\footnote{The foreign exchange market is utilised because it is the largest and most liquid financial market in the world.
 The latest survey taken by the Bank for International Settlements (BIS) in April 2013 estimated average daily turnover at \$5.4trn with average daily spot transactions of \$2.0trn See \citet{BISFX2013}.
@@
 An analysis by \citet{Steely2013} of the bid-ask spreads for USD-JPY, GBP-USD and EUR-USD for the period January 2001 to 2005 and finds that the average spread for each is 0.342\%, 0.305\% and 0.326\% respectively.
-This would be the cost for a round-trip (buying and selling immediately).
+This would be the cost for a round-trip (buying and selling immediately).}.
+If speculators are uninformed and their activity can be regarded as noise, the most extreme sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when reversals are most likely to take place; if speculators are informed prices should follow a random walk after either of these extremes.
 
-A unique set of option prices are used to identify the intensity of speculative sentiment.
-US regulatory positions measure the prominence of speculators in the market.
-Speculation does not appear to be random, there is momentum to the rise and fall in sentiment and positions.
-However, while currency prices co-move with sentiment and positions, there is no informational content from either: knowing  the strength of speculative opinion or the weight of speculative positions does not improve forecasts of future returns.
-This supports the suggestion that speculative activity has a permanent affect on prices and that is consistent with speculation being part of the process of price discovery.
-Speculators do not seem to drive prices from fundamental value and their presence seems to help rather than hinder the incorporation of information into prices.
+This paper contributes to the debate about the nature of speculation in a number of ways.
+A unique set of option prices are used to identify the intensity of speculative sentiment; an event study method is used to establish the effect of extremes of speulative sentiment and speculative positions; these speculative extremes are compared to a common noise-trader model.
+Speculative extrmes do not provide information about the future; the evolution of foreign exchange prices after an extreme can be chracterised as a random walk; this suggests that speculation is part of the process of price discovery.
 
-\section{Literature}
-\subsection*{Overview}
-  Speculation is an important concept in economics, particular in the area of finance.
+For example, Figure \ref{fig:ES1} show the evolution of a selection of foreign exchange prices after an exrtreme reading of speculative sentiment.
+Sentiment is measured by an extreme in the price of currency call options relative to the equivalent put.
+It is clear that the foreign exchange rate has a contemporaneous movement with the sentiment but that after reading the extreme, there is no clear direction.
+The lack of clear direction indicates lack of information.
+As such, extreme speuclative positions do presage reversales, indicating that they are not associated with divergence from equilibrium.
+
+\begin{figure}
+\graphicspath{{../Figures/}}
+\centering
+\includegraphics[scale=0.6]{RRCum16}
+\caption{Event Study:  Extreme (High 99th percentile) Risk Reversal Skew and 16 day event window: Risk reversals are the ratio of implied volatility on calls to the equivalent put; extreme high is the hghest 99th percentile for the whole range, cumulative returns are the sum of log exchange rate returns.
+The event day is the day of the extreme risk reversal reading as a proxy for extreme speculative sentiment.
+If speculation is uninformed, extremes shoudl be followed by reversals; if specuation is informed, extremes are more likely to be followed by a random walk.}
+\label{fig:ES1}
+\end{figure}
+% This description provides an explanation for the observation from \citet{ShillerVol} that stock prices are more volatile than the underlying earnings information on which valuation should be based.
+In addition, \citet{Delong1990noise} and \citet{shleifer1997limits} show that constrained horizons can cause noise-trading to limit the ability of informed traders to take advantage of profitable opportunities for fear of further divergence from fundamental value before realignment is achieved.
+
+The nature of speculation is central to many contemporary questions about international financial markets.
+For example, regulations or taxes that reduce the amount of speculation may be beneficial if they limit risk-taking and encourage more long-term attitudes towards investment.
+If speculation is largely uninformed it will add to the level of noise and maker it even harder for informed-traders to overcome uncertainty and apply fundamental information about long-term value.
+It will increase the tendency towards activity that is focused on the short-term.
+However, if speculation is informed and part of the process of price discovery, regulatory measures may reduce liquidity and undermine the pricess of price discovery.
+Increased volatilty, heightened risk and reduced transparency together with other costs may provide an countervailing force.
+
+The rest of the paper proceeds as follows:  Section Two analyses the distinction between informed and uninformed speuclators; Section Three discusses the measurement of speculation and the testing of extreme positions; Section Four reviews the evidence; Section Five concludes with a discussion of policy implications.
+
+\section{Informed and uninformed traders}
+Speculation plays an  important role in the areas of market-microstructure, the rationale for the trading of financial securities and rational expectations.
 However, there is no clear, unambiguous theory of speculation.
-Speculation plays a key role in three important and related areas: market-microstructure, the rationale for the trading of financial securities and rational expectations.
+While speculation may be viwed as being informed or uninformed, those trading without information are often categorised as being either \emph{noise-traders} or \emph{liquidity traders}.
 
-One key element of the microstructure literature involves distinguishing between uninformed or noise-traders and those who have information.
-The \citet{Kyle1985Continuous} model has market-makers receiving orders from the informed and the uninformed.
-On the assumption that the uniformed arrive in a random fashion without any reason to trade, the weight of orders implies something about the activity of the informed and causes prices to change in the process of price discovery.
-The bid-ask spread is the payment that the market-maker receives in return for taking the risk of meeting an informed trader.
-Ultimately, the uninformed traders make payments to the informed traders with the market-makers taking a cut for the risk and the provision of exchange.
+Noise-trading is associated with psychogical biases and restricted information while liquidity-trading is about institutional constraints.
+Therefore, noise is associated with behavioural issues, over-confidence and the use of irrational or purely fashionable investment tools.
+See \citet{DeBondtAnalysts} for an analysis of over-reaction and the \emph{represetntativeness heuristic}, \citet{OdeanOC} for a study of over-confidence and  \citet{ShillerFashion} for an assessment of investment fads.
+Market-participants do not make decisions independently of those made by others.
+Information is inferred from the activities of others.
+\citet[p.101]{Keynes1936} said investment decisions could be based on the expecttions of others' valuations.
+institutional constraints may mean that economic agents just execute customer orders at the current price or it may be that they buy and sell with reference to a benchmark index or to fulfil a hedging strategy.
+The institutional explanations are associated with the allocation of credit for investment and trading purposes:  as asset values decline exchanges, banks and other providers of liquidity require larger margins and deeper haircuts; investors liquidate their fund holdings; the attraction of hedging increases.
+See \citet{BrunnermeierLiquidity} and \citet{Gorton} for some discussion of the way that margin-calls and collaeral constraints can force investment decisions.
+However, the terms are not fixed and  \citet[p.
+1]{GortonNoise} categorise noise-traders as all those trading in markets for ``non-information-based reasons''.
+
+An important part of microstructure research is the distinction between informed and uninformed traders.
+The focus of the \citet{Kyle1985Continuous} model is a market-maker receiving orders from the informed and the uninformed.
+On the assumption that uniformed orders arrive in a random fashion, the weight of orders implies something about the knowledge of the informed and should led to adjustment of the bid-ask spread as part of the process of price discovery.
+This bid-ask spread is the payment that the market-maker receives from the uninformed in return for taking the risk of meeting an informed trader.
 See also \citet{glosten1985bid}.
 
-In the \emph{microstructure} literature, the term noise-trader is often interchanged with \emph{liquidity-trader}.
-There may be institutional and psychological reasons for the lack of attention to fundamental information.
-Nonetheless, liquidity tends to refer to institutional explanations for trading without reference to fundamental value.
-This may mean that economic agents just execute customer orders at the current price or it may be that they buy and sell with reference to a benchmark index or to fulfil a hedging strategy.
-Noise is more psychological.
+As such, Uninformed activity provides a rationale and the opportunity for trading.
+Unless there is a difference of opinion about value, there is no reason to trade.
+Uninformed speculation can create space between price and fundamental value that can be exploited by informed speculators.
+Indeed, this space can encourage informed traders to pay the cost of acquiring information.
+See, for example \citet{Grossman1980Impossibility}.
 
-\citet[p.
-1]{GortonNoise} categorise noise-traders as all those trading in markets for ``non-information-based reasons''.
-In this case, uninformed speculation provides a rational and an opportunity for the trading of financial securities.
-There must be different opinions.
-Why will traders be prepared to pay the cost of acquiring information if they are unable to take advantage of that information?
-With speculative noise-traders, prepared to take the other side of the transaction, the informed can take advantage of their information.
-This creates a justification for acquiring information.
-Speculative noise-traders can therefore create a gap between the price and the fundamental value that will encourage informed traders to pay the cost of acquiring information.
-\citet[p.529]{BlackNoise} suggested that this combination was essential to the smooth function of financial markets.
-It is ``the arbitrary element in expectations''.
+\citet[p.529]{BlackNoise} suggested that this interaction of informed and uninformed activity was essential to the smooth functioning of financial markets.
+Noise, he argued, was the  ``the arbitrary element in expectations''.
 Therefore, noise makes financial markets possible but imperfect: the more noise-traders that there are, the more liquidity and the easier it is to trade; the more noise-traders, the higher the level of inefficiency in the market and the more likely that price will be driven from fundamental value.
-This begs the question: are speculators trading on noise or are they informed?
 
-  Speculation is also an important part of the argument for why prices should tend towards a level that consistent with \emph{the rational expectations equilibrium} (RER).
-This is the level that is consistent with rational expectations.
-In the original version of rational expectations,  \citet{muth1961Rational} uses commodity speculation as the means by which prices are driven towards the level where expectations are model-consistent.
-These are of Friedman-type speculators as they base their expectations on their knowledge of the underlying model and they are the mechanism through which equilibrium is achieved.
+\citet{EasleyPIN} use informaton as an explanation for the effect of large orders on stock prices.
+The \emph{order effect} may be a consequence of the need to adjust inventories when a market-maker recieves a large order; it may also be a function of the market-maker knowledge that large orders are more likely to informed traders seeking to take advantage of their informational advantage.
+\citet{EasleyPIN2} develop a technique to estimate the probability that trading is informed.
+The find that the high liquidity stocks have the highest level of information and the most frequent arrival rate of informed traders but that this is more than offset by the high quantity of uninformed traders.
+Low liquidity stock, by contrast, have fewer information events but informed traders have a much larger effect on prices.
 
-However, though equating noise and noise-trading as a random variable may be a construction that is useful for modelling and it may be justified in analysis of broader themes of microstructure and levels of market efficiency, it may also be an over-simplification that is inappropriate when looking for more understanding of speculation itself.
+Informed speculation is also used as the theoetical justification for the tendency of Rational Expectations Equilibrium (REE) to prevail in the long-run.
+Those economic agents able to determine the fundamental value or discern the underlying economic model will be able to speculate or sell forecasts that will tend to drive prices towards the equilibrium \citet{muth1961Rational}.
+This tends to leave uninformed speculation as a pure noise that can be characterlised as white-noise random variable.
+While this construction may be justified for the analysis of microstructure and market efficiency, it is an over-simplification that is inappropriate when seeking more understanding of speculation itself.
 Institutional and psychological explanations for noise-trading frequently suggest that there are endogenous non-random elements to speculative activity.
 
-  The institutional explanations are associated with the allocation of credit for investment and trading purposes:  as asset values decline exchanges, banks and other providers of liquidity require larger margins and deeper haircuts; investors liquidate their fund holdings; the attraction of hedging increases.
-See \citet{BrunnermeierLiquidity} and \citet{Gorton} for an overview of the institutional literature.
-Psychological explanations revolve around the common panic, the group-thinking and the tendency to read private information into the actions of other investors: the so-called \emph{wisdom of crowds} depends on independent evaluation but as Keynes noted in his discussion of the \emph{beauty contest}, it is generally more important to identify what other investors will believe than to form an independent view of the actual underlying, fundamental value \citet[p.101]{Keynes1936}.
+However, a large proportion of uninformed traders in the market can undermine this process.
+\citet{Delong1990noise} and \citet{shleifer1997limits} have presented models where \emph{noise-trader risk} can prevent the arbitrate of informed speculators from driving prices back towards the fundamental value.
+The risk that noise-traders will drive prices even further away from equilibrium before the convegence can limit the activity of informed speculators when they are risk averse.\footnote{This three-period, overlapping generation model with informed and uininformed traders has a price equation \begin{equation}
+p_t = 1 + \frac{\mu(\rho_t - \rho^*)}{1 + r} + \frac{ \mu \rho^*}{r} - \frac{2 \gamma \mu^2 \sigma_{\rho}^2}{r(1 + r^2)}.
+\label{eq:NP2}
+\end{equation} 
+ where $p_t$ is the price; $\mu$ is the share of uninformed traders; $\rho$ is the price mis-perception of the uninformed traders; $\gamma$ is the level of risk aversion; $\sigma^2_{p,t+1}$ is the variance of the misperception.}
+
+The De-Long noise-trader model also provides a useful benchmark against which to assess the activity of speculators.
+The following four effects are noted by \citet[pp.
+14-15]{Delong1990noise}:  the \emph{Friedman Effect}, as informed speculators move prices back towards fundamental value; \emph{Price Pressure} is the \emph{noise-trader risk} as the speculative noise-traders drive prices towards their misperceived value; \emph{Hold More} as noise-traders misperceive the value of the risky asset and tend to hold more of it; \emph{Create Space} is the difference between noise-trader perception and the actual value, it creates the opportunity for informed traders to profit from their information.
+In this model, the more numerous the noise-traders, the more volatile prices.
+
+The Friedman-effect is the activity of informed speulators.
+They take advantage of the space that has been created by the price pressure of the uninformed traders to profit from this mis-pricing.
+Uniformed traders tend to hold more when their opinions are strong.
+This is the noise-trader risk that can prevent informed traders from fulfuillig their arbitrage role.
+If uninformed traders are very prominent or have strong opinions, prices should show the greatest divergence from fundamental value and the risk of a reversal shoould be at its greatest.
 
@@
 It is not very easy to distinguish between speculators and other types of investors.
-It is even harder to distinguish those that may be noise-traders from those assumed to be informed.
+It is even harder to distinguish between the informed and uninformed.
 One approach has been to assume that noise-traders utilise simple investment strategies and to use tests of these strategies as a means of testing speculation.
-For example speculation can be identified with momentum trading.
-It is also possible to assume that certain investors, such as investment banks and hedge funds, are informed while other trade without information.
-The latter group may include day-traders or private individuals.
-It has also been suggested that speculative techniques can change over time.
+Common strategies that have been used are momentum (ref), contrarian (ref) and carry (ref).
+It is also possible to assume that certain investors usually have more information than others.
+For example, hedge funds, investment banks and other professionals can be assuemd to be informed while retail or day traders may be uninformed.
+
+%It has also been suggested that speculative techniques can change over time.
 For example, \citet{ShillerFashion} discusses the use of fashionable techniques or fads in investment.
@@
 Momentum trades tend to be successful for certain ties.
-Also discussion the literature on heterogeneous traders, another way that the trading securities is explained, has very frequently used noise-traders as trend-chasers or followers of technical analysis.
+Also discssion the literature on heterogeneous traders, another way that the trading securities is explained, has very frequently used noise-traders as trend-chasers or followers of technical analysis.
 These models (see for example brock and hommes [1998]; lux [1995]; lux and marchesi [2000]) investigate bubbles and crashes, and feature fundamental traders versus noise traders.
+\citet{EasleyPIN2} propose a model to try to measure informed market activity based on their assertion that the bid-ask spread is a function of the level of informed trading.
+As informed traders are most likely to want to take advantage of their informational advantage, large orders are assumed to be informed.\footnote{With market-makers adopting Bayesian reasoning to update their fundamental valuation by using order flow to estimate the probability that there has been an information event and if there has whether it is good news or bad news, the authors seek to operationalise the following model:
+\begin{subequations}
+\begin{equation}
+b(t) = E[V_i|t] - \frac{\mu P_b(t)}{\varepsilon + \mu P_b(t)}(E[V-t|-\underline{V}])
+\end{equation}
+\begin{equation}
+a(t) = E[V_i|t] + \frac{\mu P_g(t)}{\varepsilon + \mu P_g(t)}(\overline{V}E+ [V-t|])
+\end{equation}
+\end{subequations}
+  where $b(t)$ is the bid price at time $t$; $V$ is the value of the security; $\mu$ is the share of informed traders in the market; $P_g(t)$ is the probability that there has been some good news and $P_b(t)$ is the probability that there has been some bad news; $\varepsilon$ is the share of uninformed traders in the market; $\underline{V}$ is the value if there is bad news and $\overline{V}$ is the value if there is good news.}
 
-\subsection{Why this is important}
-The Keynesian view of speculation is that it is an addition to market noise that makes it even harder for informed-traders to overcome uncertainty and apply fundamental information about long-term value.
-As such, it exacerbates the tendency towards putting too much emphasis on the immediate and works against longer-term and strategic thinking.
-This description provides an explanation for the observation from \citet{ShillerVol} that stock prices are more volatile than the underlying earnings information on which valuation should be based.
-In addition, \citet{Delong1990noise} and \citet{shleifer1997limits} show that constrained horizons can cause noise-trading to limit the ability of informed traders to take advantage of profitable opportunities for fear of further divergence from fundamental value before realignment is achieved.
+However, identifying large orders and making an assessment of the level of informed trading based on that is much harder to achieve in an over-the-counter market like foreign exchange.
+While electronic order books have become more widely used in recent years, much of the activity still takes place through bilateral agreements.
+Information about the nature of transactions including size and origin of orders is not widely available\footnote{\citet{Lyons1995Microstructure} and \citet{Evans2002Order} have used the initiation of orders as an explanatory variable in an equation to explain the evolution of major exchange rates.
+They find that $\$$ 1 billion of net US dollar purchase moves the deutsche mark price of the US unit by 0.5 percent.
+They argue that the orders cystalise a range of information that is used to value currencies.}    \citet{LeeandReady} seek to use price action to infer the nature of the order that has been placed into the market.
+Their tick test has four classifications: the uptic, the downtick, the zero-uptick and the zero-downtick.
+A trade is an uptick if the price is higher than the previous trade.
+If the price is the same as the previous trade, if the previous trade was classified as an uptick, this trade is a zero-uptick.
+\section{Methodology}
+The aim of this study is to assess whether speculation in the foreign exchange market is informed or uninformed.
+Speculation must be identified and then extreme speculation will then be compared with what would be expected if it were informed or uninformed.
+If informed and part of the price-discovery process, extreme speculation would be a function of perceived divergence from equilibrium or new information.
+The following price performance should be a random walk.
+If speculation is uninformed, extremes should cause a divergence from fundamental value and a price reversal should be anticiapted.
 
-The nature of speculation and the role of noise-traders has become increasingly important since the global financial crisis (GFC).
-There are a number of contemporary questions where the role and nature of speculation is at the centre of the debate.
-For example, regulations or taxes that reduce the amount of speculation have generally been considered to be beneficial as they limit risk-taking and encourage more long-term attitudes towards investment.
-However, where these measures reduce liquidity, higher risk and an increase in other costs may provide an countervailing force.
-
-The debate about a tax on financial transactions, initiated by \citet{TobinTax} and now called the \emph{the Tobin tax}, has been intended to reduce speculation.
-The surge in commodities prices that was seen prior to the GFC was seen to be partially the consequence of increased speculation in commodities and the creation of exchange traded funds (ETF) focusing on commodities markets.
-The argument for a tax on financial transactions depends on the idea that speculators will drive the price away from its fundamental value.
-
-%volcker rule?
-
-\section{Methodology}
-The aim of this study is to test whether the noise-trader model provides an appropriate explanation for speculator activity in the foreign exchange market.
-The analysis will involve three steps:  assessment of the price implication of the noise-trader model;  measurement of speculative activity in the foreign exchange market; finally, assessment of whether extreme speculative market activity is consistent with the underlying model that is used.
-If the speculative activity in the foreign exchange market produces results that are like the noise-trader model, it can be said that speculation is closer to the Keynesian view than that of Friedman.
-
-The \citet{Delong1990noise} noise-trader model used in this study.
-This is a derivative of the \citet{Samuelson1958} overlapping generations model with two-period agents making a portfolio choice between two assets: a safe asset $(s)$ which pays a set dividend $(d)$ and is in perfectly elastic supply; a risky asset $(u)$ which also pays the same dividend but is not in elastic supply.
-There are two types of investor: sophisticated traders $(i)$ who know the fundamental value of the risky assets and noise-traders $(n)$ who misperceive the fundamental value.
-In the absence of noise-traders, the assets are substitutes.
-Each trader chooses their portfolio in the first period $(t)$ to maximise their utility on the basis of their belief about the price of the unsafe asset in the second period $(t + 1)$.
-The representative noise-trader mistakes the price of the risky asset by a random variable
-
-\begin{equation}
-\rho_t \sim N(\rho^*, \sigma_{\rho^2})
-\end{equation}
-
-Maximising utility with respect to holdings of the sage and risky assets for the informed and uninformed traders and iterating backward to solve for the current value,  the price becomes the discounted value of the future dividends, the expected capital gain and the general misperception of noise-traders less risk.
-
-\begin{equation}
-p_t = 1 + \frac{\mu(\rho_t - \rho^*)}{1 + r} + \frac{ \mu \rho^*}{r} - \frac{2 \gamma}{r}\ \big (_t\sigma^2_{p,t+1} \big )
-\label{eq:NP1}
-\end{equation}
-
-The deviation from the true fundamental value of unity is a result of the discounted value of the noise-traders' misperception of the price $(p_t - \rho^*)$, multiplied by the weight of noise-traders in the total $(\mu)$ plus a perpetuity of the average noise-trader misperception $(\frac{ \mu \rho^*}{r} )$, adjusted for the risk.
-This risk is a function of the volatility of the price which,  from Equation \ref{eq:NP1} is a consequence of the variability of the misperception of speculative noise-traders.
-All but the second term in equation \ref{eq:NP1} are constants, and if the variance of the price one period ahead  is equal to the variance of the misperception,
-
-\begin{equation} 
- _t\sigma^2_{p,t+1} = \frac{\mu^2 \sigma_{\rho}^2}{1+r^2}  
-\end{equation}
-
-the final price equation becomes:
-
-\begin{equation}
-p_t = 1 + \frac{\mu(\rho_t - \rho^*)}{1 + r} + \frac{ \mu \rho^*}{r} - \frac{2 \gamma \mu^2 \sigma_{\rho}^2}{r(1 + r^2)}.
-\label{eq:NP2}
-\end{equation}
-
-The following four effects are noted by \citet[pp.
-14-15]{Delong1990noise}:  the \emph{Friedman Effect}, as informed speculators move prices back towards fundamental value; \emph{Price Pressure} is the \emph{noise-trader risk} as the speculative noise-traders drive prices towards their misperceived value; \emph{Hold More} as noise-traders misperceive the value of the risky asset and tend to hold more of it; \emph{Create Space} is the difference between noise-trader perception and the actual value, it creates the opportunity for informed traders to profit from their information.
-In this model, the more numerous the noise-traders, the more volatile prices.
-
-The Friedman-effect and the Price-pressure are the effects that are studied here.
-Their combination suggests that when the weight of noise-traders is high or the times when they have particularly strong opinions will tend to be those when there is the greatest risk of reversal.
-The hold-more effect suggests that the weight of speculative involvement in the market will tend to increase as a trend.
-Create-space effect would suggest that when the weight of noise-traders is high and extreme, there is greater incentive and opportunity for the informed to take advantage of the divergence between fundamental value and price.
-
-\subsection{The implications of the noise-trader model}
-There are two parameters from the noise-trader price equation \ref{eq:NP2} that can be used to measure the amount of speculation:  the price misperception $(\rho)$ and the weight of noise-traders in the market $(\mu)$.
-The aim of this section is to  identify the returns that are implied by the noise-trader model when either the intensity of the misperception or the weight of noise-traders is increased to extreme levels so that this can be used as a benchmark against which foreign exchange market activity can be judged.
-
-To understand what the noise-trader model of speculative behaviour means for foreign exchange rates, a \emph{Monte Carlo} simulation is carried out with a variety of misperception-perceptions and noise-trader weights.
-Two time series are created.
-The first series is based on a random noise-trade misperception, $(\rho^*)$.
-This is called the Misperception Model (MPM).
-The second is based on a random weight of noise-traders $(\mu)$ and is called the Random Weight Model (RWM).
-The simulation creates a 1000 period price series from Equation \ref{eq:NP2}.
-The dividend paid on the risky asset $(r)$ is posted at 10\%; the risk-aversion parameter $(\gamma)$ is assumed to be 0.3 (in line with that suggested in \citet[p.
-127]{FrobesBehaviour}); starting from a price of 100, the average noise-trader misconception $(p^*)$ is a random variable in the first simulation with a mean of 4 and a variance $(\sigma_{\rho^*}^2)$ of 2, and has a fixed value of 4 in the second simulation; the weight of noise-traders $(\mu)$ is 50\% in the first case but is modelled as a two-period moving average of a normal random variable centred at 50\% with a standard deviation of 20\% and equally weighted average of its previous value (starting at 50\%) to allow smooth adjustment, in the second simulation.
-% it is possible to have other specifications. For example, the noise-traders oculd become more prominent if their strategy is more successful,
-
-A 1000 period prices simulation is created for each model and the returns for these series are calculated as the log change in price.
-Those periods when misperception-perception or the weight of speculators is \emph{extreme} are identified as those that are at or above the ninety-fifth percentile.
-These extremes are the \emph{event} for an event study.
-Descriptive statistics for the whole series of returns and for those when either the extent of their misperception or the weight of noise-traders is extreme are presented in Table \ref{tabref:MC1}.
-The process is then repeated for one thousand simulations of each model so that confidence intervals can be established for the estimates of the descriptive statistics.
-
-\subsection{The results of simulation}
-The results of the simulation are recorded in Table \ref{tabref:MC1}.
-It shows that the mean return to an investment strategy in the whole of the sample is very close to zero.
-However, for each of the extreme readings, there are large negative returns to follow.
-This is consistent with the structure of the model: noise-traders drive price above intrinsic value and subsequent returns will be negative; the more extreme the misperception or the weight of uninformed speculative noise-traders in the market, the greater the negative return that is likely to be recorded.
-
-\begin{table}[h!]
-\begin{threeparttable}
-\caption{Results of the Monte Carlo Simulation of the Noise-Trader Model}
-\begin{tabular}{p{4cm} | r r r r}
- \label{tabref:MC1}
-                            & Base Case  & Extreme (MPM) & Extreme (RWM) \\
-  \hline
-\hline
-Mean                     & -0.00 & -0.90*   & -0.49 \\
-Standard Deviation  & 0.98 &  0.49*     &  0.42*\\
-Skew                      &  -0.07 &  -0.16   &  -0.78*\\
-Kurtosis                  &  -0.14 &  -0.38    &  -0.35\\
-Number                 &  1000        &   50        & 50\\
-
-\end{tabular}
-\begin{tablenotes}
-\small
-\item Descriptive statistics for the De Long Noise-Trader model \citet{Delong1990noise}.
-The Base Case measures the benchmark as returns from prices generated by Equation \ref{eq:NP2} for the whole sample.
-The MPM measures returns after a period of extreme sentiment when the misperception of noise-traders ($\rho$) is equal to or above the 95th percentile of all the misperceptions; the RMW measures the returns following a recording of extreme weight of noise-traders ($\mu$) is equal to or above the 95th percentile of all noise-trader realisations.
-The asterisk denotes significantly different from zero at the 95th confidence interval.
-\end{tablenotes}
-\end{threeparttable}
-\end{table}
+Returning to the \citet[pp.
+14-15]{Delong1990noise} model summarised by Equation \ref{eq:NP2}, there are two parameters that can be used to measure speculation:  the price misperception $(\rho)$ and the weight of noise-traders in the market $(\mu)$.
+If either of these reach an extreme, prices should be expected to reverse\footnote{See \citet{Hayward2013} for full details of simulations of this model.}
 
@@
 Frankel and Froot also assess the 'Financial Report' that has been produced by \emph{The Economist} magazine and \emph{Currency Forecasters Digest} and find a wide dispersion of opinion and express some concern about the aggregation of such divergent forecasters \citet gc{FrootFrankelFDB} and \citep{FrankelFroot1987}.}
+
 Risk-reversal skew is the relative price of currency put and call options for the same strike price.
@@
 If sentiment is based   on noise rather than information, extremes should presage a sharp reversal; if speculative sentiment is informed, this is part of the price discovery process and extremes may be followed by a random walk.
-
-Table \ref{tabref:DS1} in Appendix \ref{apref:1} shows descriptive statistics for the exchange rate data used in the study.
-There are daily data for the period between January 1996 and September 2002.
-This is the period for which risk-reversal data are available.
-The series are of sightly different lengths as there are some gaps in the risk reversal and exchange rate data as a result of holidays and market closures.
-From the raw data continuously compounded returns are calculated and multiplied by 100.
-The first thing of note is that the mean return does not appear to be statistically different from zero.
-This is what would be expected if, as has been identified in many studies, exchange rates follow something close to a random walk.
-\begin{sidewaystable}
-\begin{threeparttable}
-\caption{Descriptive Statistics for Risk Reversals: 2 January 1996 to 6 September 2002}
-\begin{tabular}{l | r r  r r r r r r r r}
-  \hline
- Statistic &  AUD & EURGBP & EURSEK & GBP & CHF & EURCHF & EURJPY & EUR & CAD & JPY \\
- \hline
-Number & 1702 & 1697 & 1719 & 1691 & 1374 & 1339 & 1697 & 1692 & 1701 & 1697 \\ 
-  Mean & -0.37 & 0.15 & 0.71 & -0.08 & -0.28 & -0.38 & -0.48 & 0.19 & 0.12 & -0.74 \\ 
-  Median & -0.30 & 0.10 & 0.80 & -0.10 & -0.15 & -0.30 & -0.10 & 0.10 & 0.15 & -0.55 \\ 
-  Standard Deviation & 0.38 & 0.37 & 0.51 & 0.33 & 0.49 & 0.36 & 1.02 & 0.49 & 0.25 & 1.02 \\ 
- Max & 0.45 & 1.72 & 1.90 & 1.00 & 1.05 & 0.50 & 1.70 & 2.05 & 1.45 & 2.05 \\ 
- Min & -1.70 & -0.80 & -0.30 & -1.35 & -2.05 & -1.80 & -4.03 & -1.15 & -0.65 & -4.00 \\ 
- Skewness & -0.50* & 1.27* & -0.23* & 0.33* & -0.79* & -0.58* & -0.93* & 0.68* & 0.23* & -0.51* \\ 
-  Kurtosis & 0.02 & 2.29* & -1.67 & 0.72 & 0.69 & 0.19 & 0.15 & 0.96 & 1.83 & -0.03 \\ 
-      \hline
- \label{tabref:DS2}
-\end{tabular}
-\begin{tablenotes}
-\small
-\item Daily risk reversal skew.
-Risk reversal is the premium of implied volatility for calls relative to the equivalent put (equal distance to the strike price, 25 delta in this case).
-A figure above zero means that calls are more expensive (in volatility and price terms) than puts.
-A figure below zero means that puts are more expensive.
-The asterisk (*) means that the estimate is significant at the 5\% level using t-statistics for the mean and the standard error of skew and standard error of excess-kurtosis respectively.
-These are calculated as $SES = \sqrt{\frac{6n(n-1)}{(n-2)(N+1)(N+3)}}$ and $SEK = \sqrt{\frac{n^2-1}{(n-3)(N+5)}}$, with the assumption that this estimate is distributed normally, the skew or kurtosis is considered significant if it is more than two standard errors from zero.
-\end{tablenotes}
-\end{threeparttable}
-\end{sidewaystable}
-
-Table \ref{tabref:DS2} gives the descriptive statistics for the risk-reversal data.
-For a neutral risk-reversal, a reading of zero, there would be no price preference for calls or puts for the same delta.
-The figures show that, in contrast to the average returns, there are often average skews that are different from zero.
-The most significant of these are the positive bias for EUR-GBP and a negative bias for USD-JPY.
-The latter may be related to hedging activity by Japanese exporters.
-While this could suggest that implied volatility is an imperfect measure of speculative activity, note that extreme readings are based on quantiles of the distribution and therefore a discrete adjustment of this distribution to reflect the hedging power should not change the timing of signals measured by extremes.
 
@@
 The former is more sensitive, as can be seen by the range and the standard deviation; the second is usually above zero because there are always some spreading positions that seek to take advantage of calendar spreads and other strategies.
-
-\begin{sidewaystable}
-\begin{threeparttable}
-\caption{Descriptive Statistics for S1 (Speculative Sentiment) and S2 (Speculative Weight): 30 September 1998 to 31 December 2008}
-\begin{tabular}{l | r r r r r r r  r r r }
-  \hline
- Statistic &  \multicolumn{2}{c}{EUR} & \multicolumn{2}{c}{JPY} & \multicolumn{2}{c}{GBP} & \multicolumn{2}{c}{CHF} & \multicolumn{2}{c}{CAD}  \\
- & S1 & S2 & S1 & S2 & S1 & S2 & S1 & S2 & S1 & S2 \\  
-\hline
-Number & 522 &  522 & 849 & 849 & 849 & 849 & 847 & 847 & 849 & 849\\
-Mean &  31.31 & 43.43 & -23.90 & 44.21 & 2.76 & 41.68 & -21.14 & 46.13 & 6.68 & 41.23\\
-Median & 42.75 & 39.56 & -33.30 & 43.56 & 11.15 & 38.88 & -37.32 & 42.88 & 7.44 & 39.81\\
-Standard deviation  & 41.94 & 16.13 & 51.49 & 12.17 & 60.59 & 16.46 & 60.76 & 15.73 & 55.26 & 15.33\\
-Max & 95.67 & 86.06 & 87.43 & 84.50 & 96.15 & 89.45 & 100.00 & 92.49 & 100.00 & 86.63\\
-Min &  -100.00 & 11.15 & -99.83 & 12.88 & -100.00 & 9.10 & -100.00 & 14.61 & -90.08 & 11.89\\
-Skew  & -0.88* & 0.55* & 0.33* & 0.57* & -0.17 & 0.76* & 0.45* & 0.65* & 0.02 & 0.50*\\
-Kurtosis & 0.18 & -0.57 & -1.14 & 0.64 & -1.45 & 0.15 & -1.25 & -0.15 & -1.35 & -0.28\\
-\hline
- \label{tabref:DS4}
-\end{tabular}
-\begin{tablenotes}
-\small
-\item S1 is the first sentiment indicator, calculated as ratio of net non-commercial long less non-commercial short to the sum of non-commercial long and short positions multiplied by one hundred.
-S2 is the second sentiment indicators, calculated as the ratio of net non-commercial long to net non-commercial short plus spreading to open interest.
-An asterisk (*) means that the estimate is significantly different from zero at the 5\% level using t-statistics for the mean and the standard error of skew and standard error of excess-kurtosis respectively.
-These are calculated as $SES = \sqrt{\frac{6n(n-1)}{(n-2)(N+1)(N+3)}}$ and $SEK = \sqrt{\frac{n^2-1}{(n-3)(N+5)}}$, with the assumption that this estimate is distributed normally, the skew or kurtosis is considered significant if it is more than two standard errors from zero.
-\end{tablenotes}
-\end{threeparttable}
-\end{sidewaystable}
 
@@
 \caption{Event Study:  Extreme (High 99th percentile) Risk Reversal Skew and 16 day event window}
-\label{fig:ES1}
+\label{fig:ES5}
 \includegraphics[scale=0.8]{RRCum16}
 \end{figure}
+% the reference is ES5 to avoid conflict with the earlier version.
 
@@
 This paper has sought to measure speculation and to use this measure to discover more about the relationship between speculation, information and asset prices.  
-The measurement of speculation takes two forms:  a measurement of the intensity of belief with the risk-reversal skew, and a measurement of speculative positions.
-The option prices reveal the intensity of feeling through the weight of money in a speculative part (options) of a speculative market (foreign exchange).
-The commitment of traders report records only speculators in the US futures market.
-However, this is an important market, there is no reason to believe that US speculators in the futures market are systematically different from those in the rest of the world and presumably if there were a major difference there would be an arbitrage opportunity between the two markets.
+Whether measured as intensity of speculative sentiment or as the weight of speculator action, it is clear that extreme speculation does not provide information about future returns.
+This means that specualtion is more likely to be informed than uninformed.
+If it were uninformed it woudl be much more likely that these extremes would be followe by reversals as informed market participants took advantage of the space created by this uninformed speculative activity.
 
-There is strong evidence that speculation affects prices and that this effect is gradual rather than instantaneous.
-This consistent with a market inefficiency that would reward momentum-based trading.
-It is no wonder therefore that \citet{TaylorTechnical}, reporting a survey conducted by the Bank of England, found that of the Chief foreign exchange dealers contacted in London in November 1988, 90\% of them attached some weight to \emph{technical analysis}, a method that is fundamentally based on finding trends in asset prices.
-The survey found that the importance of technical analysis became greater at shorter horizons.
-The importance of momentum trading and technical analysis has not waned with time.
-The more up-to-date CFTC data show the same pattern.
-In addition,  \citet{NeelyWeller} show that gains from technical trading rules still generate excess returns and attribute this to behavioural biases.
-It seems that speculation forms part of the process of price discovery, but this process is not instantaneous as would be expected if the market were informationally efficient.
+The possibility that noise-trader risk limits the activity of informed traders is limited by the large windows that are employed when constructing figures for the event study.
+Informed traders may fear the uninformed in the short-run and this may delay the chance of a reversal but over the medium-term reversals would be expected.
+This event study uses 20 days and 6 weeks for the daily and weekly event windows.
+
+The nature speculation is central to a number of contemporary policy questions.
+Amongst these are the design and regulation of financial markets, the surge in commodity prices and the kimits put on the activities of financial institutions.
+The debate about a tax on financial transactions, initiated by \citet{TobinTax} and now called the \emph{the Tobin tax}, has been intended to reduce speculation.
+The argument for a tax on financial transactions depends on the idea that speculators will drive the price away from its fundamental value.
+However, if speuclation is part of the process of price discovery, reducing its activity may incure costs that offset some of the benefits that are expected.
+
+The reduction in speculation in the US government bond market that has been seen since the introduction of new costs and restrictions under the array of ????
+regulations has been seen by some to reduce liquidity and increase volatiolity, rising risk and imposing additional costs for final investors.
+
+The surge in commodities prices that was seen prior to the GFC was seen to be partially the consequence of increased speculation in commodities and the creation of exchange traded funds (ETF) focusing on commodities markets.
+Even if speculators are on both side of the market and not driving prices in one direction, they may still increase the amount of volatility, incurring costs for producers and consumers of underlying materials.
 
@@
 
-These findings have implications for speculators, economic modellers and policy makers.
-Speculators may jump on the trend, but they do not know by looking at how popular that trend is when to get off.
-They need other signs to tell them that the trend has reached its conclusion or equilibrium has been achieved.
-For the modellers, there is confirmation of the need  for feedback effects and evidence that reversals cannot be systematically identified from high levels of speculative positions.
-It may suggest another trigger for bursting bubbles; it may, for example, suggest that this reversal has to be modelled as stochastic element or pure shock.
-For policy makers, there is no escape from the requirement to look at economic fundamentals when deciding whether bubbles are building in asset markets.
-The presence of speculators is no guarantee that there is excess.
-The speculators may be just part of the adjustment to new information.
-
 \bibliography{../../../../myrefs}
@@
 \begin{sidewaystable}
-
 \begin{threeparttable}
-\caption{Descriptive Statistics for Daily Exchange Rate Returns: 2 January 1996 to 6 September 2002}
- \label{tabref:DS1}
+\caption{Appendix 1: Descriptive Statistics for Risk Reversals: 2 January 1996 to 6 September 2002}
 \begin{tabular}{l | r r  r r r r r r r r}
@@
  \hline
-Number & 1685 & 1675 & 1716 & 1667 & 1674 & 1674 & 1675 & 1668 & 1683 & 1675  \\
-Mean & -0.0174 & -0.0187 & 0.0011 & 0.0025 & 0.0120 & -0.0039 & -0.0072 & -0.0172 & 0.0084 & 0.0076\\
-Median &  0.0000 & -0.0298 & 0.0000& 0.0000 & 0.0114 & 0.0000 & 0.0000 & -0.0253 & 0.0000 & 0.0188\\
-Standard deviation  & 0.7241 & 0.5542 & 0.0438 & 0.0489 & 0.6819 & 0.0277 & 0.8416 & 0.6338 & 0.0363 & 0.8170\\
-Max & 5.8301 & 3.0543 & 2.1633 & 2.2573 & 3.5086 & 1.6477 & 4.0879 & 2.4136 & 2.2170 & 5.1788 \\
-Min & -3.4666 & -2.2858 & -2.2900 & -2.2701 & -3.6525 & -1.8466 & -8.2848 & -2.2901 & -2.9707 & -10.6573 \\
-Skewness  & 0.1206* & 0.2923* & 0.0432 & -0.0452 & -0.1423* & -0.4046* & -0.7388* & 0.1738* & -0.1095 & -1.7237*\\
-Kurtosis & 4.17* & 1.88 & 2.43* & 1.48 & 1.96 & 5.84* & 7.08* & 1.12 & 5.11* & 21.52*\\
+Number & 1702 & 1697 & 1719 & 1691 & 1374 & 1339 & 1697 & 1692 & 1701 & 1697 \\ 
+  Mean & -0.37 & 0.15 & 0.71 & -0.08 & -0.28 & -0.38 & -0.48 & 0.19 & 0.12 & -0.74 \\ 
+  Median & -0.30 & 0.10 & 0.80 & -0.10 & -0.15 & -0.30 & -0.10 & 0.10 & 0.15 & -0.55 \\ 
+  Standard Deviation & 0.38 & 0.37 & 0.51 & 0.33 & 0.49 & 0.36 & 1.02 & 0.49 & 0.25 & 1.02 \\ 
+ Max & 0.45 & 1.72 & 1.90 & 1.00 & 1.05 & 0.50 & 1.70 & 2.05 & 1.45 & 2.05 \\ 
+ Min & -1.70 & -0.80 & -0.30 & -1.35 & -2.05 & -1.80 & -4.03 & -1.15 & -0.65 & -4.00 \\ 
+ Skewness & -0.50* & 1.27* & -0.23* & 0.33* & -0.79* & -0.58* & -0.93* & 0.68* & 0.23* & -0.51* \\ 
+  Kurtosis & 0.02 & 2.29* & -1.67 & 0.72 & 0.69 & 0.19 & 0.15 & 0.96 & 1.83 & -0.03 \\ 
       \hline
+ \label{tabref:DS2}
 \end{tabular}
@@
 \small
-\item Continuously compounded returns (multiplied by 100) for period 2 January 1996 to  September 2002 for US dollar per unit of Australian dollar (AUD), UK sterling per unit of Euro (EURGBP), Swedish Crown per unit of EUR (EURSEK); US dollar per unit of UK sterling (GBP);  Swiss franc per unit of US dollar (CHF); Swiss franc per unit of Euro (EURCHF); US dollars per unit of Euro (EUR); Canadian dollars per unit of US dollar (CAD); Japanese yen per unit of US dollars (JPY).
-The asterisk (*) means that the estimate is significantly different from zero at the 5\% level using t-statistics for the mean and the standard error of skew and standard error of excess-kurtosis respectively.
+\item Daily risk reversal skew.
+Risk reversal is the premium of implied volatility for calls relative to the equivalent put (equal distance to the strike price, 25 delta in this case).
+A figure above zero means that calls are more expensive (in volatility and price terms) than puts.
+A figure below zero means that puts are more expensive.
+The asterisk (*) means that the estimate is significant at the 5\% level using t-statistics for the mean and the standard error of skew and standard error of excess-kurtosis respectively.
+These are calculated as $SES = \sqrt{\frac{6n(n-1)}{(n-2)(N+1)(N+3)}}$ and $SEK = \sqrt{\frac{n^2-1}{(n-3)(N+5)}}$, with the assumption that this estimate is distributed normally, the skew or kurtosis is considered significant if it is more than two standard errors from zero.
 \end{tablenotes}
@@
 \section{Appendix 2}
-
+\begin{sidewaystable}
 \begin{threeparttable}
-\caption{Descriptive Statistics for Weekly Exchange Rate Returns: 30 September 1998 to 31 December 2008}
-\begin{tabular}{{4cm} | r r  r r r }
+\caption{Appendix 2: Descriptive Statistics for S1 (Speculative Sentiment) and S2 (Speculative Weight): 30 September 1998 to 31 December 2008}
+\begin{tabular}{l | r r r r r r r  r r r }
   \hline
- Statistic &  EUR & JPY & GBP & CHF & CAD  \\
+ Statistic &  \multicolumn{2}{c}{EUR} & \multicolumn{2}{c}{JPY} & \multicolumn{2}{c}{GBP} & \multicolumn{2}{c}{CHF} & \multicolumn{2}{c}{CAD}  \\
+ & S1 & S2 & S1 & S2 & S1 & S2 & S1 & S2 & S1 & S2 \\  
 \hline
-Number & 522 &  849 & 849 & 847 & 849 \\
-Mean & 0.0351 & 0.0337 & -0.0202 & 0.0233 & 0.0018\\
-Median & 0.0650 & -0.0800 & 0.0000 & 0.0000 & -0.0200\\
-Standard deviation  & 1.4063 & 1.6355 & 1.2755 & 1.5418 & 1.0558\\
-Max & 8.7900 & 11.6700 & 5.7200 & 8.3400 & 10.0100\\
-Min & -4.2400 & -15.0400 & -6.4100 & -5.7100 & -5.0900\\
-Skew  & 0.27* & -0.06 & -0.34* & 0.31* & 0.78*\\
-Kurtosis & 2.76*& 12.58* & 2.07 * & 1.82 & 11.59*\\
+Number & 522 &  522 & 849 & 849 & 849 & 849 & 847 & 847 & 849 & 849\\
+Mean &  31.31 & 43.43 & -23.90 & 44.21 & 2.76 & 41.68 & -21.14 & 46.13 & 6.68 & 41.23\\
+Median & 42.75 & 39.56 & -33.30 & 43.56 & 11.15 & 38.88 & -37.32 & 42.88 & 7.44 & 39.81\\
+Standard deviation  & 41.94 & 16.13 & 51.49 & 12.17 & 60.59 & 16.46 & 60.76 & 15.73 & 55.26 & 15.33\\
+Max & 95.67 & 86.06 & 87.43 & 84.50 & 96.15 & 89.45 & 100.00 & 92.49 & 100.00 & 86.63\\
+Min &  -100.00 & 11.15 & -99.83 & 12.88 & -100.00 & 9.10 & -100.00 & 14.61 & -90.08 & 11.89\\
+Skew  & -0.88* & 0.55* & 0.33* & 0.57* & -0.17 & 0.76* & 0.45* & 0.65* & 0.02 & 0.50*\\
+Kurtosis & 0.18 & -0.57 & -1.14 & 0.64 & -1.45 & 0.15 & -1.25 & -0.15 & -1.35 & -0.28\\
 \hline
- \label{tabref:DS3}
+ \label{tabref:DS4}
 \end{tabular}
@@
 \small
-\item Continuously compounded returns multiplied by 100 for US dollar per unit of Euro (EUR), US dollar per unit of Japanese Yen (JPY), US dollar per unit of UK Sterling (GBP), US dollar per unit of Swiss Franc (CHF) and US dollar per unit of Canadian dollar (CAD) for the period 30 September 1998 to 31 Dec 2008;     An asterisk (*) means that the estimate is significantly different from zero at the 5\% level of statistical significance using t-statistics for the mean and the standard error of skew and standard error of excess-kurtosis respectively.
+\item S1 is the first sentiment indicator, calculated as ratio of net non-commercial long less non-commercial short to the sum of non-commercial long and short positions multiplied by one hundred.
+S2 is the second sentiment indicators, calculated as the ratio of net non-commercial long to net non-commercial short plus spreading to open interest.
+An asterisk (*) means that the estimate is significantly different from zero at the 5\% level using t-statistics for the mean and the standard error of skew and standard error of excess-kurtosis respectively.
 These are calculated as $SES = \sqrt{\frac{6n(n-1)}{(n-2)(N+1)(N+3)}}$ and $SEK = \sqrt{\frac{n^2-1}{(n-3)(N+5)}}$, with the assumption that this estimate is distributed normally, the skew or kurtosis is considered significant if it is more than two standard errors from zero.
@@
 \end{threeparttable}
+\end{sidewaystable}
 

### 2015-09-08 R/CompareDist.R

@@ -1,0 +1,22 @@
+# This comes from http://memosisland.blogspot.co.uk/2015/08/practical-kullback-leibler-kl.html
+# K-L divergence (Kullback-leiber)
+# Measures the association between two random variables
+set.seed(3951824)
+# Table of counts (contingency)
+X <- sample(letters[1:3], 100, 1)
+Y <- sample(letters[4:5], 100, 1)
+t  <- table(X,Y)
+t
+tp  <- t/100 # proportions
+tn  <- tp/sum(tp)     # normalized, joints
+p_x <- rowSums(tn)    # marginals
+p_y <- colSums(tn)
+
+P <- tn 
+Q <- p_x %o% p_y 
+
+# P(X, Y)   : bin frequency: P_i
+# P(X) P(Y) : bin frequency, Q_i 
+mi <- sum(P*log(P/Q))
+library(entropy)
+mi.empirical(t) == mi

### 2015-09-08 R/Extreme.R

@@ -1,0 +1,98 @@
+# measure distributions
+# http://www.r-bloggers.com/
+# practical-kullback-leibler-kl-divergence-discrete-case/
+require(eventstudies) 
+require(PerformanceAnalytics)
+require(xtable)
+require(dplyr)
+# Read in the data------------------------------------------
+daz <- read.zoo("Data/RR.csv", header = TRUE, 
+                format = "%d-%b-%y", sep = ",")
+das <- read.csv("Data/RR.csv", header = TRUE)
+das[,1] <- as.Date(das[,1], format = "%d-%b-%y")
+head(daz)
+str(das)
+#------------------------------------------------------------------
+# use dplyr
+riskrev <- paste(currency, "RR", sep = "")
+da <- select(das, Date, EUR, EURRR) %>% 
+  mutate(EUR.R = log(EUR) - log(lag(EUR))) %>% 
+  head()
+da$exh <- 0
+da$exh[,da$EURRR >= quantile(da$EURRR, 0.95)]
+quantile(da$EURRR, 0.05)
+da$EURRR >= quantile(da$EURRR, 0.95)
+head(da)
+currencylist <- colnames(daz)[c(2, 4, 6, 8, 10, 12, 14, 16, 18, 20)]
+currencylist
+highlow = c("high","low")
+hist(da$EURRR)
+# Create extreme readings and prepare event files-------------------------
+fxlist = list()
+extreme = "high"
+hi = as.numeric(0.95)
+lo = as.numeric(0.05)
+# select window either "whole lot" or "after".
+cum = "whole lot"
+wind = as.numeric(1) 
+# Read file (as zoo object) if not already done
+currency = "GBP"
+fxret <- function(currency){
+title <- paste(currency, ".r")
+dat <- daz[,c(paste(currency), paste(currency, "RR", sep = ""))]
+dat[,3] <- 100*diff(log(dat[,1]))
+colnames(dat)[3] <- title
+fxlist[currency] <- dat
+fxlist$currency <- dat
+}
+head(dat)
+str(fxlist)
+str(dat)
+a <- fxret("EUR")
+dat[-1,currency] <- 100*diff(log(dat[,1]))
+title
+dat$title <- 100
+head(FX.z)
+for(extreme in highlow)
+if(extreme == "high"){
+  FX.z$ex <- as.numeric(FX.z[,2] >= quantile(FX.z[,2],  
[…]

### 2015-09-08 R/Extreme2.R

@@ -1,0 +1,81 @@
+require(eventstudies) 
+require(PerformanceAnalytics)
+require(xtable)
+# Read in the data------------------------------------------
+daz <- read.zoo("Data/RR.csv", header = TRUE, 
+                format = "%d-%b-%y", sep = ",")
+summary(daz)
+FX.z <- daz[,c("AUDRR", "EURGBPRR", "EURSEKRR", "GBPRR", "CHFRR", "EURCHFRR", 
+               "EURJPYRR", "EURRR","CADRR", "JPYRR")]
+head(FX.z)
+# Create extreme readings and prepare event files-------------------------
+currency = "CAD"
+extreme = "high"
+hi = as.numeric(0.95)
+lo = as.numeric(0.05)
+# select window either "whole lot" or "after".
+cum = "whole lot"
+wind = as.numeric(10) 
+# Read file (as zoo object) if not already done
+daz <- read.zoo("Data/RR.csv", header = TRUE, 
+                format = "%d-%b-%y", sep = ",")
+FX.z <-daz[,c(paste(currency), paste(currency, "RR", sep = ""))]
+FX.z$r <- 100*diff(log(FX.z[,1]))
+# THis has been changed from the original so that the dataframe has a high and low
+# column to show the high and low extreme. 
+exlist <- c("high", "low")
+for(ex in exlist){
+if(ex == "high"){
+  FX.z$exh <- as.numeric(FX.z[,2] >= quantile(FX.z[,2],  
+                                             hi, na.rm = TRUE)) 
+} else if(ex == "low") {
+  FX.z$exl <- as.numeric(FX.z[,2] <= quantile(FX.z[,2], 
+                                             lo, na.rm = TRUE))
+} else(print("No Extreme"))
+}
+head(FX.z)
+
+
+# Create event list-----------------------------------------
+eventslist.FXex <- data.frame(index(FX.z$exh[FX.z$exh ==1]))
+eventslist.FXex$unit <- as.character("r")
+colnames(eventslist.FXex) <- c("when", "unit")
+eventslist.FXex$when <-as.Date(eventslist.FXex$when)
+# Align returns with extreme events-----------------------
+a <- phys2eventtime(FX.z, eventslist.FXex, width = wind)
+str(a)
[…]

### 2015-09-09 Knitr/NewPaper.Rnw
Extreme.R now has the comparison of extreme and regular distribution This needs to be turned into a function that produces graphs and table.

@@
 \begin{abstract}
-  If speculation is just noise, price movements associated with extremes of speuclative sentiment or heavy involvement of speculative activity should be reveversed; if speculation is informed, associated price movement is more likely to be sustained.
-A unique set of risk-reversals on currency options is used to measure speculative sentiment and regulatory positions measure speculative activity.
-An event study shows that extremes do not increase the probability of a price reversal, supoprting the view that speculation is part of the price disovery process.
+ Uninformed speculation based on noise should lead to price reversals; informed speculation is sustainable.
+A unique set of option prices measure speculative sentiment while regulatory positions determine the weight of speculative activity.
+An event study shows that extremes of speculative sentiment or speculative activity do not increase the probability of price reversal.
+Price activity after the extreme is close to a random walk.
+This supports the view that speculation is part of the price-discovery process.
 \end{abstract}
@@
 This paper contributes to the debate about the nature of speculation in a number of ways.
-A unique set of option prices are used to identify the intensity of speculative sentiment; an event study method is used to establish the effect of extremes of speulative sentiment and speculative positions; these speculative extremes are compared to a common noise-trader model.
-Speculative extrmes do not provide information about the future; the evolution of foreign exchange prices after an extreme can be chracterised as a random walk; this suggests that speculation is part of the process of price discovery.
+A unique set of option prices are used to identify the intensity of speculative sentiment; an event study method is used to establish the effect of extremes of speculative sentiment and speculative positions; these speculative extremes are compared to a common noise-trader model.
+Speculative extremes do not provide information about the future; the evolution of foreign exchange prices after an extreme can be characterised as a random walk; this suggests that speculation is part of the process of price discovery.
 
-For example, Figure \ref{fig:ES1} show the evolution of a selection of foreign exchange prices after an exrtreme reading of speculative sentiment.
+For example, Figure \ref{fig:ES1} show the evolution of a selection of foreign exchange prices after an extreme reading of speculative sentiment.
 Sentiment is measured by an extreme in the price of currency call options relative to the equivalent put.
@@
 The lack of clear direction indicates lack of information.
-As such, extreme speuclative positions do presage reversales, indicating that they are not associated with divergence from equilibrium.
+As such, extreme speculative positions do presage reversals, indicating that they are not associated with divergence from equilibrium.
 
@@
 \includegraphics[scale=0.6]{RRCum16}
-\caption{Event Study:  Extreme (High 99th percentile) Risk Reversal Skew and 16 day event window: Risk reversals are the ratio of implied volatility on calls to the equivalent put; extreme high is the hghest 99th percentile for the whole range, cumulative returns are the sum of log exchange rate returns.
+\caption{Event Study:  Extreme (High 99th percentile) Risk Reversal Skew and 16 day event window: Risk reversals are the ratio of implied volatility on calls to the equivalent put; extreme high is the highest 99th percentile for the whole range, cumulative returns are the sum of log exchange rate returns.
 The event day is the day of the extreme risk reversal reading as a proxy for extreme speculative sentiment.
-If speculation is uninformed, extremes shoudl be followed by reversals; if specuation is informed, extremes are more likely to be followed by a random walk.}
+If speculation is uninformed, extremes should be followed by reversals; if speculation is informed, extremes are more likely to be followed by a random walk.}
 \label{fig:ES1}
@@
 It will increase the tendency towards activity that is focused on the short-term.
-However, if speculation is informed and part of the process of price discovery, regulatory measures may reduce liquidity and undermine the pricess of price discovery.
-Increased volatilty, heightened risk and reduced transparency together with other costs may provide an countervailing force.
+However, if speculation is informed and part of the process of price discovery, regulatory measures may reduce liquidity and undermine the prices of price discovery.
+Increased volatility, heightened risk and reduced transparency together with other costs may provide an countervailing force.
 
-The rest of the paper proceeds as follows:  Section Two analyses the distinction between informed and uninformed speuclators; Section Three discusses the measurement of speculation and the testing of extreme positions; Section Four reviews the evidence; Section Five concludes with a discussion of policy implications.
+The rest of the paper proceeds as follows:  Section Two analyses the distinction between informed and uninformed speculators; Section Three discusses the measurement of speculation and the testing of extreme positions; Section Four reviews the evidence; Section Five concludes with a discussion of policy implications.
 
@@
 However, there is no clear, unambiguous theory of speculation.
-While speculation may be viwed as being informed or uninformed, those trading without information are often categorised as being either \emph{noise-traders} or \emph{liquidity traders}.
+While speculation may be vowed as being informed or uninformed, those trading without information are often categorised as being either \emph{noise-traders} or \emph{liquidity traders}.
 
-Noise-trading is associated with psychogical biases and restricted information while liquidity-trading is about institutional constraints.
+Noise-trading is associated with psychological biases and restricted information while liquidity-trading is about institutional constraints.
 Therefore, noise is associated with behavioural issues, over-confidence and the use of irrational or purely fashionable investment tools.
-See \citet{DeBondtAnalysts} for an analysis of over-reaction and the \emph{represetntativeness heuristic}, \citet{OdeanOC} for a study of over-confidence and  \citet{ShillerFashion} for an assessment of investment fads.
+See \citet{DeBondtAnalysts} for an analysis of over-reaction and the \emph{representativeness heuristic}, \citet{OdeanOC} for a study of over-confidence and  \citet{ShillerFashion} for an assessment of investment fads.
 Market-participants do not make decisions independently of those made by others.
 Information is inferred from the activities of others.
-\citet[p.101]{Keynes1936} said investment decisions could be based on the expecttions of others' valuations.
+\citet[p.101]{Keynes1936} said investment decisions could be based on the expectations of others' valuations.
 institutional constraints may mean that economic agents just execute customer orders at the current price or it may be that they buy and sell with reference to a benchmark index or to fulfil a hedging strategy.
 The institutional explanations are associated with the allocation of credit for investment and trading purposes:  as asset values decline exchanges, banks and other providers of liquidity require larger margins and deeper haircuts; investors liquidate their fund holdings; the attraction of hedging increases.
-See \citet{BrunnermeierLiquidity} and \citet{Gorton} for some discussion of the way that margin-calls and collaeral constraints can force investment decisions.
+See \citet{BrunnermeierLiquidity} and \citet{Gorton} for some discussion of the way that margin-calls and collateral constraints can force investment decisions.
 However, the terms are not fixed and  \citet[p.
@@
 
-\citet{EasleyPIN} use informaton as an explanation for the effect of large orders on stock prices.
-The \emph{order effect} may be a consequence of the need to adjust inventories when a market-maker recieves a large order; it may also be a function of the market-maker knowledge that large orders are more likely to informed traders seeking to take advantage of their informational advantage.
+\citet{EasleyPIN} use information as an explanation for the effect of large orders on stock prices.
+The \emph{order effect} may be a consequence of the need to adjust inventories when a market-maker receives a large order; it may also be a function of the market-maker knowledge that large orders are more likely to informed traders seeking to take advantage of their informational advantage.
 \citet{EasleyPIN2} develop a technique to estimate the probability that trading is informed.
@@
 
-Informed speculation is also used as the theoetical justification for the tendency of Rational Expectations Equilibrium (REE) to prevail in the long-run.
+Informed speculation is also used as the theoretical justification for the tendency of Rational Expectations Equilibrium (REE) to prevail in the long-run.
 Those economic agents able to determine the fundamental value or discern the underlying economic model will be able to speculate or sell forecasts that will tend to drive prices towards the equilibrium \citet{muth1961Rational}.
-This tends to leave uninformed speculation as a pure noise that can be characterlised as white-noise random variable.
+This tends to leave uninformed speculation as a pure noise that can be characterised as white-noise random variable.
 While this construction may be justified for the analysis of microstructure and market efficiency, it is an over-simplification that is inappropriate when seeking more understanding of speculation itself.
@@
 \citet{Delong1990noise} and \citet{shleifer1997limits} have presented models where \emph{noise-trader risk} can prevent the arbitrate of informed speculators from driving prices back towards the fundamental value.
-The risk that noise-traders will drive prices even further away from equilibrium before the convegence can limit the activity of informed speculators when they are risk averse.\footnote{This three-period, overlapping generation model with informed and uininformed traders has a price equation \begin{equation}
+The risk that noise-traders will drive prices even further away from equilibrium before the convergence can limit the activity of informed speculators when they are risk averse.\footnote{This three-period, overlapping generation model with informed and uninformed traders has a price equation \begin{equation}
 p_t = 1 + \frac{\mu(\rho_t - \rho^*)}{1 + r} + \frac{ \mu \rho^*}{r} - \frac{2 \gamma \mu^2 \sigma_{\rho}^2}{r(1 + r^2)}.
@@
 
-The Friedman-effect is the activity of informed speulators.
+The Friedman-effect is the activity of informed speculators.
 They take advantage of the space that has been created by the price pressure of the uninformed traders to profit from this mis-pricing.
 Uniformed traders tend to hold more when their opinions are strong.
-This is the noise-trader risk that can prevent informed traders from fulfuillig their arbitrage role.
-If uninformed traders are very prominent or have strong opinions, prices should show the greatest divergence from fundamental value and the risk of a reversal shoould be at its greatest.
+This is the noise-trader risk that can prevent informed traders from fulfilled their arbitrage role.
+If uninformed traders are very prominent or have strong opinions, prices should show the greatest divergence from fundamental value and the risk of a reversal should be at its greatest.
 
@@
 It is also possible to assume that certain investors usually have more information than others.
-For example, hedge funds, investment banks and other professionals can be assuemd to be informed while retail or day traders may be uninformed.
+For example, hedge funds, investment banks and other professionals can be assumed to be informed while retail or day traders may be uninformed.
 
@@
 Information about the nature of transactions including size and origin of orders is not widely available\footnote{\citet{Lyons1995Microstructure} and \citet{Evans2002Order} have used the initiation of orders as an explanatory variable in an equation to explain the evolution of major exchange rates.
-They find that $\$$ 1 billion of net US dollar purchase moves the deutsche mark price of the US unit by 0.5 percent.
-They argue that the orders cystalise a range of information that is used to value currencies.}    \citet{LeeandReady} seek to use price action to infer the nature of the order that has been placed into the market.
-Their tick test has four classifications: the uptic, the downtick, the zero-uptick and the zero-downtick.
-A trade is an uptick if the price is higher than the previous trade.
-If the price is the same as the previous trade, if the previous trade was classified as an uptick, this trade is a zero-uptick.
+They find that $\$$ 1 billion of net US dollar purchase moves the Deutsche mark price of the US unit by 0.5 percent.
+They argue that the orders crystallise a range of information that is used to value currencies.}    \citet{LeeandReady} seek to use price action to infer the nature of the order that has been placed into the market.
+Their tick test has four classifications: the up-tick, the down-tick, the zero-up-tick and the zero-down-tick.
+A trade is an up-tick if the price is higher than the previous trade.
+If the price is the same as the previous trade, if the previous trade was classified as an up-tick, this trade is a zero-up-tick.
 \section{Methodology}
@@
 The following price performance should be a random walk.
-If speculation is uninformed, extremes should cause a divergence from fundamental value and a price reversal should be anticiapted.
+If speculation is uninformed, extremes should cause a divergence from fundamental value and a price reversal should be anticipated.
 
@@
 Whether measured as intensity of speculative sentiment or as the weight of speculator action, it is clear that extreme speculation does not provide information about future returns.
-This means that specualtion is more likely to be informed than uninformed.
-If it were uninformed it woudl be much more likely that these extremes would be followe by reversals as informed market participants took advantage of the space created by this uninformed speculative activity.
+This means that speculation is more likely to be informed than uninformed.
+If it were uninformed it would be much more likely that these extremes would be follow by reversals as informed market participants took advantage of the space created by this uninformed speculative activity.
 
@@
 The nature speculation is central to a number of contemporary policy questions.
-Amongst these are the design and regulation of financial markets, the surge in commodity prices and the kimits put on the activities of financial institutions.
+Amongst these are the design and regulation of financial markets, the surge in commodity prices and the limits put on the activities of financial institutions.
 The debate about a tax on financial transactions, initiated by \citet{TobinTax} and now called the \emph{the Tobin tax}, has been intended to reduce speculation.
 The argument for a tax on financial transactions depends on the idea that speculators will drive the price away from its fundamental value.
-However, if speuclation is part of the process of price discovery, reducing its activity may incure costs that offset some of the benefits that are expected.
+However, if speculation is part of the process of price discovery, reducing its activity may incur costs that offset some of the benefits that are expected.
 
 The reduction in speculation in the US government bond market that has been seen since the introduction of new costs and restrictions under the array of ????
-regulations has been seen by some to reduce liquidity and increase volatiolity, rising risk and imposing additional costs for final investors.
+regulations has been seen by some to reduce liquidity and increase volatility, rising risk and imposing additional costs for final investors.
 

### 2015-09-09 R/Extreme.R

@@ -1,90 +1,56 @@
-# measure distributions
-# http://www.r-bloggers.com/
-# practical-kullback-leibler-kl-divergence-discrete-case/
 require(eventstudies) 
 require(PerformanceAnalytics)
 require(xtable)
-require(dplyr)
-# Read in the data------------------------------------------
-daz <- read.zoo("Data/RR.csv", header = TRUE, 
-                format = "%d-%b-%y", sep = ",")
-das <- read.csv("Data/RR.csv", header = TRUE)
-das[,1] <- as.Date(das[,1], format = "%d-%b-%y")
-head(daz)
-str(das)
-#------------------------------------------------------------------
-# use dplyr
-riskrev <- paste(currency, "RR", sep = "")
-da <- select(das, Date, EUR, EURRR) %>% 
-  mutate(EUR.R = log(EUR) - log(lag(EUR))) %>% 
-  head()
-da$exh <- 0
-da$exh[,da$EURRR >= quantile(da$EURRR, 0.95)]
-quantile(da$EURRR, 0.05)
-da$EURRR >= quantile(da$EURRR, 0.95)
-head(da)
-currencylist <- colnames(daz)[c(2, 4, 6, 8, 10, 12, 14, 16, 18, 20)]
-currencylist
-highlow = c("high","low")
-hist(da$EURRR)
 # Create extreme readings and prepare event files-------------------------
-fxlist = list()
+currency = "GBP"
 extreme = "high"
 hi = as.numeric(0.95)
 lo = as.numeric(0.05)
 # select window either "whole lot" or "after".
 cum = "whole lot"
-wind = as.numeric(1) 
-# Read file (as zoo object) if not already done
-currency = "GBP"
-fxret <- function(currency){
-title <- paste(currency, ".r")
-dat <- daz[,c(paste(currency), paste(currency, "RR", sep = ""))]
-dat[,3] <- 100*diff(log(dat[,1]))
-colnames(dat)[3] <- title
-fxlist[currency] <- dat
-fxlist$currency <- dat
-}
-head(dat)
-str(fxlist)
-str(dat)
-a <- fxret("EUR")
-dat[-1,currency] <- 100*diff(log(dat[,1]))
-title
-dat$title <- 100
-head(FX.z)
-for(extreme in highlow)
-if(extreme == "high"){
[…]

### 2015-09-09 R/Extreme.R
Updated Extreme2 to create functions.

@@ -2,7 +2,7 @@ require(eventstudies)
 require(PerformanceAnalytics)
 require(xtable)
 # Create extreme readings and prepare event files-------------------------
-currency = "GBP"
+currency = "EUR"
 extreme = "high"
 hi = as.numeric(0.95)
 lo = as.numeric(0.05)
@@ -63,11 +63,7 @@ FX.exwc.b
 FX.exac.b <- inference.Ecar(FX.exac, to.plot = TRUE)
 FX.exac.b
 # create charts for comparison of extreme and regular returns. 
-FX.exac # This is the list of returns. 
-hist(FX.exacm[11,], breaks = 10, prob = TRUE)
-mean(FX.exacm[11,], na.rm = TRUE)
-sd(FX.exacm[11,])
-mean(FX.exacm[11,])/sd(FX.exacm[11,])
+ 
 par(mfrow = c(2,1))
 plot(density(100*FX.exacm[2,]), xlim = c(-2.0, 2.0), main = "Extreme Density",
      lwd = 2, col = "red")
@@ -78,3 +74,15 @@ plot(density(sample(FX.z[-1,3], 113), na.rm = TRUE), xlim = c(-2.0, 2.0),
      main = "Regular Density", lwd = 2, col = "red")
 abline(v = mean(FX.z[-1, 3], na.rm = TRUE))
 text(x = .22, y = 0.4, labels = "Mean")
+# Mean and standard deviation of return day after extreme
+me <- round(mean(FX.exacm[1,], na.rm = TRUE),4)
+sde <- round(sd(FX.exacm[11,]), 4)
+tex <- round(mean(FX.exacm[11,])/sd(FX.exacm[11,]),4)
+# Mean and standard deviation of return (full sample)
+mr <- round(mean(FX.z[-1,3], na.rm = TRUE),4)
+sdr <- round(sd(FX.z[-1,3], na.rm = TRUE), 4)
+tr <- round(mean(FX.z[-1,3], na.rm = TRUE)/sd(FX.z[-1,3], na.rm = TRUE),4)
+df <- data.frame(ex = c(me, sde, tex), r = c(mr, sdr, tr))
+rownames(df) <- c("Mean", "SD", "T-stat")
+df
+

### 2015-09-09 R/Extreme2.R

@@ -1,24 +1,18 @@
+# Load required packages
 require(eventstudies) 
 require(PerformanceAnalytics)
 require(xtable)
-# Read in the data------------------------------------------
+# Load data-------------------------
 daz <- read.zoo("Data/RR.csv", header = TRUE, 
                 format = "%d-%b-%y", sep = ",")
-summary(daz)
-FX.z <- daz[,c("AUDRR", "EURGBPRR", "EURSEKRR", "GBPRR", "CHFRR", "EURCHFRR", 
-               "EURJPYRR", "EURRR","CADRR", "JPYRR")]
-head(FX.z)
-# Create extreme readings and prepare event files-------------------------
-currency = "CAD"
-extreme = "high"
+# Not now needed, Part of function
+# currency = "EUR"
+# extreme = "high"
+# Currencydf is a dataframe to store currency, returns and extremes
+currencydf <- function(currency){
 hi = as.numeric(0.95)
 lo = as.numeric(0.05)
-# select window either "whole lot" or "after".
-cum = "whole lot"
-wind = as.numeric(10) 
-# Read file (as zoo object) if not already done
-daz <- read.zoo("Data/RR.csv", header = TRUE, 
-                format = "%d-%b-%y", sep = ",")
+# are these needed?
 FX.z <-daz[,c(paste(currency), paste(currency, "RR", sep = ""))]
 FX.z$r <- 100*diff(log(FX.z[,1]))
 # THis has been changed from the original so that the dataframe has a high and low
@@ -33,23 +27,41 @@ if(ex == "high"){
                                              lo, na.rm = TRUE))
 } else(print("No Extreme"))
 }
-head(FX.z)
-
-
-# Create event list-----------------------------------------
-eventslist.FXex <- data.frame(index(FX.z$exh[FX.z$exh ==1]))
+return(FX.z)
+}
+EUR <- currencydf("EUR")
+head(EUR)
+# Create zoo object for events----------------------------------------
+# Events must be stored in particular format.
+# Requires currencydf zoo object, wind for size of window and hilo for high or low extreme.
[…]

### 2015-09-10 R/Extreme2.R
Extreem2.R updated The file contains three functions to create the currency zoo object, manipulate the zoo object to be compatible with the eventstudy package and to make a statistical and graphical comparison of the extreme and regular currency returns.

@@ -33,7 +33,8 @@ EUR <- currencydf("EUR")
 head(EUR)
 # Create zoo object for events----------------------------------------
 # Events must be stored in particular format.
-# Requires currencydf zoo object, wind for size of window and hilo for high or low extreme.
+# Requires currencydf zoo object, wind for size of window and hilo 
+#for high or low extreme.
 eventzoo <- function(currencydf, wind, hilo){
   if(hilo == "hi"){
     ho = 4
@@ -47,61 +48,53 @@ eventslist.FXex$unit <- as.character("r")
 colnames(eventslist.FXex) <- c("when", "unit")
 eventslist.FXex$when <-as.Date(eventslist.FXex$when)
 # Align returns with extreme events-----------------------
-a <- phys2eventtime(FX.z, eventslist.FXex, width = wind)
+a <- phys2eventtime(currencydf, eventslist.FXex, width = wind)
 # Create windows (exw - whole, exa - after, cumsum - cumulative sum)
 FX.exw <- window(a$z.e, start = wind * -1, end = wind)
 FX.exa <- window(a$z.e, start = 0, end = wind)
 FX.exwc <- remap.cumsum(FX.exw, is.pc = TRUE, base = 0)
 FX.exac <- remap.cumsum(FX.exa, is.pc = TRUE, base = 0)
-g <- list(FX.exw = FX.exw, FX.exwc = FX.exwc, FX.exac = FX.exac, hilo = hilo)
-return(g)
-}
-t <- eventzoo(currencydf = EUR, wind = 2, hilo = "low")
-hilo = "low"
-head(currencydf$exl)
-str(t)
-currencydf = EUR
-head(currencydf)
-# window size-----------------------------------------
 dims = seq(-wind, wind, 1)
-dim(eventslist.FXex)
 FX.exwcm <- as.matrix(FX.exwc)
 FX.exacm <- as.matrix(FX.exac)
-dim(FX.exacm)
-hist(FX.exacm[10,], prob = TRUE, breaks = 10, main = "Distribution of extreme")
-lines(density(FX.exacm[10,]), col = "red")
 MRW <- mean(FX.exwcm[(wind*2 + 1),])
 MRA <- mean(FX.exacm[(wind + 1),])
-# Mean return whole
-MRW
[…]

### 2015-09-15 Knitr/NewPaper.Rnw
Updated the NewPaper and Extreme2.R The paper is ready for Andros, the R code will loop thorugh the extremes and create charts of the distribution post extreme. There is also a comparison of extreme move and the level of speculative sentiment.  Table produced.

@@
 \begin{abstract}
- Uninformed speculation based on noise should lead to price reversals; informed speculation is sustainable.
+If prices are driven by speculation based on noise rather than information, extreme levels of speculative sentiment and speculative activity should lead to price reversals.
+Informed speculation is sustainable as it is part of the price discovery process.
 A unique set of option prices measure speculative sentiment while regulatory positions determine the weight of speculative activity.
-An event study shows that extremes of speculative sentiment or speculative activity do not increase the probability of price reversal.
+An event study shows that extremes of speculative sentiment or speculative activity do not increase the probability of price reversal, though variance-measured risk increases.
 Price activity after the extreme is close to a random walk.
-This supports the view that speculation is part of the price-discovery process.
+This supports the view that speculation is part of the price-discovery process rather than noise.
 \end{abstract}
@@
 Ideas about the way that speculation affects financial market span a broad range.
-Towards one end is the Keynesian view that speculation is a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd'' \citet[p.
+Towards one end is the view of \citet[p.
+101]{Keynes1936} that speculation is a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd'' \citet[p.
 101]{Keynes1936}.
-Speculation is a problem that will cause booms, reversals and will discourage fundamental, long-term investment; It is noise that emerges from the interaction of collective sentiment to encourages trends in asset prices.
+Speculation is a problem that will cause booms, reversals and will discourage fundamental, long-term investment.
+It is noise that emerges from the interaction of collective sentiment to encourages trends in asset prices.
 
-Towards the other end of the spectrum, Friedman argued that speculation was a stabilising force, providing liquidity and helping to ensure that markets swiftly find equilibrium.
+Towards the other end of the spectrum is the idea that speculation is a stabilising force, providing liquidity and helping to ensure that markets swiftly find equilibrium.
 \citet{FriedmanPositive} asserted that successful speculators would be those who managed to correctly buy when prices  were below fundamental value and sell above that price.
@@
 A unique set of option prices are used to identify the intensity of speculative sentiment; an event study method is used to establish the effect of extremes of speculative sentiment and speculative positions; these speculative extremes are compared to a common noise-trader model.
-Speculative extremes do not provide information about the future; the evolution of foreign exchange prices after an extreme can be characterised as a random walk; this suggests that speculation is part of the process of price discovery.
+Speculative extremes do not provide information about the future; the evolution of foreign exchange prices after an extreme can be characterised as a random walk.
+However, extremes are followed by an increase in the variance of exchange rate returns.
+This suggests that information from informed speculation contributes to the process of price discovery.
 
-For example, Figure \ref{fig:ES1} show the evolution of a selection of foreign exchange prices after an extreme reading of speculative sentiment.
-Sentiment is measured by an extreme in the price of currency call options relative to the equivalent put.
-It is clear that the foreign exchange rate has a contemporaneous movement with the sentiment but that after reading the extreme, there is no clear direction.
-The lack of clear direction indicates lack of information.
-As such, extreme speculative positions do presage reversals, indicating that they are not associated with divergence from equilibrium.
+Figure \ref{fig:ES1} shows the evolution of a selection of foreign exchange prices after speculative sentiment is extreme.
+An event study is conducted with the speculative extreme being the event $(t_0)$,  Sentiment is measured by the price of currency call options relative to the equivalent put.
+The foreign exchange rate has a contemporaneous movement with the sentiment.
+This suggests that either the price is adjusting to new information or that uninformed speculators are driving the price in the direction of their irrational belief.
+However, when sentiment reaches an extreme, there is no clear direction: prices are as likely to go up as they are to go down.
+As such, extreme speculative positions do not presage reversals, indicating that they are not associated with divergence from equilibrium.
 
@@
 \centering
-\includegraphics[scale=0.6]{RRCum16}
+\includegraphics[scale=0.8]{RRCum16}
 \caption{Event Study:  Extreme (High 99th percentile) Risk Reversal Skew and 16 day event window: Risk reversals are the ratio of implied volatility on calls to the equivalent put; extreme high is the highest 99th percentile for the whole range, cumulative returns are the sum of log exchange rate returns.
@@
 For example, regulations or taxes that reduce the amount of speculation may be beneficial if they limit risk-taking and encourage more long-term attitudes towards investment.
-If speculation is largely uninformed it will add to the level of noise and maker it even harder for informed-traders to overcome uncertainty and apply fundamental information about long-term value.
-It will increase the tendency towards activity that is focused on the short-term.
-However, if speculation is informed and part of the process of price discovery, regulatory measures may reduce liquidity and undermine the prices of price discovery.
-Increased volatility, heightened risk and reduced transparency together with other costs may provide an countervailing force.
+If speculation is largely uninformed it will add to the level of noise and make it even harder for informed investors to overcome uncertainty and apply fundamental analysis about long-term value; will increase the tendency towards activity that is focused on the short-term.
+However, if speculation is informed and part of the process of price discovery, regulatory measures to reduce its influence may also reduce liquidity and undermine the prices of price discovery.
+Increased volatility, heightened risk and reduced transparency together with other costs may provide a counteraction to any positive benefits.
 
-The rest of the paper proceeds as follows:  Section Two analyses the distinction between informed and uninformed speculators; Section Three discusses the measurement of speculation and the testing of extreme positions; Section Four reviews the evidence; Section Five concludes with a discussion of policy implications.
+The rest of the paper proceeds as follows:  Section Two discusses the distinction between informed and uninformed speculators; Section Three talks about the measurement of speculation and the testing of extreme positions; Section Four reviews the evidence; Section Five concludes with a discussion of policy implications.
 
 \section{Informed and uninformed traders}
-Speculation plays an  important role in the areas of market-microstructure, the rationale for the trading of financial securities and rational expectations.
+Speculation plays an  important role in the areas of market-microstructure, the rationale for the trading of financial securities as well as providing the explanation for rational expectations are supposed to operate.
 However, there is no clear, unambiguous theory of speculation.
-While speculation may be vowed as being informed or uninformed, those trading without information are often categorised as being either \emph{noise-traders} or \emph{liquidity traders}.
-
-Noise-trading is associated with psychological biases and restricted information while liquidity-trading is about institutional constraints.
-Therefore, noise is associated with behavioural issues, over-confidence and the use of irrational or purely fashionable investment tools.
-See \citet{DeBondtAnalysts} for an analysis of over-reaction and the \emph{representativeness heuristic}, \citet{OdeanOC} for a study of over-confidence and  \citet{ShillerFashion} for an assessment of investment fads.
-Market-participants do not make decisions independently of those made by others.
-Information is inferred from the activities of others.
-\citet[p.101]{Keynes1936} said investment decisions could be based on the expectations of others' valuations.
-institutional constraints may mean that economic agents just execute customer orders at the current price or it may be that they buy and sell with reference to a benchmark index or to fulfil a hedging strategy.
-The institutional explanations are associated with the allocation of credit for investment and trading purposes:  as asset values decline exchanges, banks and other providers of liquidity require larger margins and deeper haircuts; investors liquidate their fund holdings; the attraction of hedging increases.
-See \citet{BrunnermeierLiquidity} and \citet{Gorton} for some discussion of the way that margin-calls and collateral constraints can force investment decisions.
+Speculation is frequently associated with the term \emph{trading}, but trading is broader as it encompasses market-making.
+A distinction made between the short-term orientation of speculation and trading and the more long-term activity of investment.
+Speculation and trading can be categorised as being informed or uninformed.
+Those trading without information are also called \emph{noise-traders} or \emph{liquidity traders}.
 However, the terms are not fixed and  \citet[p.
 1]{GortonNoise} categorise noise-traders as all those trading in markets for ``non-information-based reasons''.
-
-An important part of microstructure research is the distinction between informed and uninformed traders.
+%these two paragraphs above may go later and just a very short coverage of the distinction between noise-traders and liquidity traders. 
+The distinction between informed and uninformed traders is a central component of microstructure research.
 The focus of the \citet{Kyle1985Continuous} model is a market-maker receiving orders from the informed and the uninformed.
@@
 This bid-ask spread is the payment that the market-maker receives from the uninformed in return for taking the risk of meeting an informed trader.
-See also \citet{glosten1985bid}.
+In a similar way \citet{glosten1985bid}describes a model where the bid-ask spread is a function of adverse selection, being caused by the risk that orders come from informed traders.
 
-As such, Uninformed activity provides a rationale and the opportunity for trading.
+Uninformed activity provides a rationale and the opportunity for trading.
 Unless there is a difference of opinion about value, there is no reason to trade.
 Uninformed speculation can create space between price and fundamental value that can be exploited by informed speculators.
-Indeed, this space can encourage informed traders to pay the cost of acquiring information.
-See, for example \citet{Grossman1980Impossibility}.
+This space can encourage informed traders to pay the cost of acquiring information.
+\citet{Grossman1980Impossibility} argue that there is equilibrium between the cost of obtaining information and the benefits that are expected to come from the use of the information.
 
@@
 \citet{EasleyPIN} use information as an explanation for the effect of large orders on stock prices.
-The \emph{order effect} may be a consequence of the need to adjust inventories when a market-maker receives a large order; it may also be a function of the market-maker knowledge that large orders are more likely to informed traders seeking to take advantage of their informational advantage.
+The \emph{order effect} may be a consequence of the need to adjust inventories when a market-maker receives a large order; that effect may also be a function of the market-maker knowledge that large orders are more likely come from to informed traders seeking to realise their informational advantage.
 \citet{EasleyPIN2} develop a technique to estimate the probability that trading is informed.
-The find that the high liquidity stocks have the highest level of information and the most frequent arrival rate of informed traders but that this is more than offset by the high quantity of uninformed traders.
+The find that the high liquidity stocks have the highest level of information and the most frequent arrival rate of informed traders but that price changes are limited by the volume of uniformed traders.
 Low liquidity stock, by contrast, have fewer information events but informed traders have a much larger effect on prices.
 
-Informed speculation is also used as the theoretical justification for the tendency of Rational Expectations Equilibrium (REE) to prevail in the long-run.
-Those economic agents able to determine the fundamental value or discern the underlying economic model will be able to speculate or sell forecasts that will tend to drive prices towards the equilibrium \citet{muth1961Rational}.
-This tends to leave uninformed speculation as a pure noise that can be characterised as white-noise random variable.
+Informed speculation is also used as the explanation for the tendency of Rational Expectations Equilibrium (REE) to prevail in the long-run.
+Those economic agents able to determine the fundamental value or discern the underlying economic model will be able to speculate and drive prices towards equilibrium.
+This tends to leave uninformed speculation as a pure noise that can be characterised as white-noise random variable\footnote{See \citet{muth1961Rational} for the original adjustment to a cobweb model.}.
 While this construction may be justified for the analysis of microstructure and market efficiency, it is an over-simplification that is inappropriate when seeking more understanding of speculation itself.
 Institutional and psychological explanations for noise-trading frequently suggest that there are endogenous non-random elements to speculative activity.
+Noise-trading is associated with psychological biases and restricted information while liquidity-trading is about institutional constraints.
+Therefore, noise is associated with behavioural issues, over-confidence and the use of irrational or purely fashionable investment tools.
+See \citet{DeBondtAnalysts} for an analysis of over-reaction and the \emph{representativeness heuristic}, \citet{OdeanOC} for a study of over-confidence and  \citet{ShillerFashion} for an assessment of investment fads.
+Traders without information may not make decisions independently of others.
+Information may be inferred from the buying and selling conducted by other traders\footnote{See for example \citet[p.101]{Keynes1936} and the beauty contest.}.
 
-However, a large proportion of uninformed traders in the market can undermine this process.
+Institutional constraints may mean that economic agents just execute customer orders at the current price or it may be that they buy and sell with reference to a benchmark index or to fulfil a hedging strategy.
+The institutional explanations are associated with the allocation of credit for investment and trading purposes:  as asset values decline exchanges, banks and other providers of liquidity require larger margins and deeper haircuts; investors liquidate their fund holdings; the attraction of hedging increases.
+See \citet{BrunnermeierLiquidity} and \citet{Gorton} for some discussion of the way that margin-calls and collateral constraints can force investment decisions.
+
+\citet{FriedmanPositive} said that the informed market participants would drive the uninformed from the market.
+However, a large proportion of uninformed traders in a market can undermine this process.
 \citet{Delong1990noise} and \citet{shleifer1997limits} have presented models where \emph{noise-trader risk} can prevent the arbitrate of informed speculators from driving prices back towards the fundamental value.
-The risk that noise-traders will drive prices even further away from equilibrium before the convergence can limit the activity of informed speculators when they are risk averse.\footnote{This three-period, overlapping generation model with informed and uninformed traders has a price equation \begin{equation}
+Noise-trader risk is the risk that noise-traders will drive prices even further away from equilibrium before convergence.
+This can limit the activity of informed speculators when they are risk averse.\footnote{This three-period, overlapping generation model with informed and uninformed traders has a price equation \begin{equation}
 p_t = 1 + \frac{\mu(\rho_t - \rho^*)}{1 + r} + \frac{ \mu \rho^*}{r} - \frac{2 \gamma \mu^2 \sigma_{\rho}^2}{r(1 + r^2)}.
@@
 \end{equation} 
- where $p_t$ is the price; $\mu$ is the share of uninformed traders; $\rho$ is the price mis-perception of the uninformed traders; $\gamma$ is the level of risk aversion; $\sigma^2_{p,t+1}$ is the variance of the misperception.}
+ where $p_t$ is the price; $\mu$ is the share of uninformed traders; $\rho^*$ is the average misperception of uninformed traders; $\rho_t$ is the price mis-perception of the uninformed traders at time t; $\gamma$ is the level of risk aversion; $\sigma^2_{p,t+1}$ is the variance of the misperception.}
 
@@
 One approach has been to assume that noise-traders utilise simple investment strategies and to use tests of these strategies as a means of testing speculation.
-Common strategies that have been used are momentum (ref), contrarian (ref) and carry (ref).
+Amongst the strategies that have been used in this way are momentum, contrarian and carry trades.
+However, as these techniques can be profitable, it is not clear that these can be identified as the domain of the uniformed.
 It is also possible to assume that certain investors usually have more information than others.
@@
 
-%It has also been suggested that speculative techniques can change over time.
-For example, \citet{ShillerFashion} discusses the use of fashionable techniques or fads in investment.
-In this way, speculation can be be associated with sentiment and measures of sentiment are then used as ways of measuring speculation.
-
-%Noise stragegies may be successful for a while.
-Momentum trades tend to be successful for certain ties.
-Also discssion the literature on heterogeneous traders, another way that the trading securities is explained, has very frequently used noise-traders as trend-chasers or followers of technical analysis.
-These models (see for example brock and hommes [1998]; lux [1995]; lux and marchesi [2000]) investigate bubbles and crashes, and feature fundamental traders versus noise traders.
-\citet{EasleyPIN2} propose a model to try to measure informed market activity based on their assertion that the bid-ask spread is a function of the level of informed trading.
-As informed traders are most likely to want to take advantage of their informational advantage, large orders are assumed to be informed.\footnote{With market-makers adopting Bayesian reasoning to update their fundamental valuation by using order flow to estimate the probability that there has been an information event and if there has whether it is good news or bad news, the authors seek to operationalise the following model:
+\citet{EasleyPIN2} propose a model to try to measure informed market activity based on the assertion that the bid-ask spread is a function of the level of informed trading.
+Large orders are assumed to be informed as informed traders are most likely to want to swift use of their informational advantage.\footnote{With market-makers using Bayes' Rule to update their fundamental valuation by using order flow to estimate the joint probability that there has been an information event and that it has been good news or bad news, the authors seek to operationalise the following model:
 \begin{subequations}
@@
 
-However, identifying large orders and making an assessment of the level of informed trading based on that is much harder to achieve in an over-the-counter market like foreign exchange.
-While electronic order books have become more widely used in recent years, much of the activity still takes place through bilateral agreements.
-Information about the nature of transactions including size and origin of orders is not widely available\footnote{\citet{Lyons1995Microstructure} and \citet{Evans2002Order} have used the initiation of orders as an explanatory variable in an equation to explain the evolution of major exchange rates.
+However, identifying large orders and making an assessment of the level of informed trading is much harder in an over-the-counter market like foreign exchange.
+While electronic order books have become more widely used, much of the market-making still takes place through bilateral agreements and information about the nature of transactions including size and origin of orders is not widely available.\footnote{\citet{Lyons1995Microstructure} and \citet{Evans2002Order} have gained access to proprietary data and have used the initiation of orders as an explanatory variable in an equation to explain the evolution of major exchange rates.
 They find that $\$$ 1 billion of net US dollar purchase moves the Deutsche mark price of the US unit by 0.5 percent.
 They argue that the orders crystallise a range of information that is used to value currencies.}  \citet{LeeandReady} seek to use price action to infer the nature of the order that has been placed into the market.
-Their tick test has four classifications: the up-tick, the down-tick, the zero-up-tick and the zero-down-tick.
+The \emph{tick-test} has four classifications: the up-tick, the down-tick, the zero-up-tick and the zero-down-tick.
 A trade is an up-tick if the price is higher than the previous trade.
 If the price is the same as the previous trade, if the previous trade was classified as an up-tick, this trade is a zero-up-tick.
-\section{Methodology}
+Under this system, the up-ticks are assumed to be buy orders and the down-ticks are sell orders.
+Up-ticks are classified as buy orders and down-ticks are classified as sell orders.
+These estimates of the market activity can be used to estimate the level of informed trading.
+\section{Measuring speculation and the event study}
 The aim of this study is to assess whether speculation in the foreign exchange market is informed or uninformed.
 Speculation must be identified and then extreme speculation will then be compared with what would be expected if it were informed or uninformed.
-If informed and part of the price-discovery process, extreme speculation would be a function of perceived divergence from equilibrium or new information.
-The following price performance should be a random walk.
+If informed and part of the price-discovery process, extreme speculation should drive the price towards a new equilibrium; the following price performance should be a random walk.
 If speculation is uninformed, extremes should cause a divergence from fundamental value and a price reversal should be anticipated.
 
-Returning to the \citet[pp.
-14-15]{Delong1990noise} model summarised by Equation \ref{eq:NP2}, there are two parameters that can be used to measure speculation:  the price misperception $(\rho)$ and the weight of noise-traders in the market $(\mu)$.
-If either of these reach an extreme, prices should be expected to reverse\footnote{See \citet{Hayward2013} for full details of simulations of this model.}
+The \citet[pp.
+14-15]{Delong1990noise} model is summarised by Equation \ref{eq:NP2}.
+There are two parameters that can be used to measure speculation:  the price misperception $(\rho)$ and the weight of noise-traders in the market $(\mu)$.
+If speculation is uninformed and either of these parameters reach an extreme, the model indicates that prices should subsequently reverse.\footnote{See \citet{Hayward2013} for full details of simulations of this model.} Therefore, if misperception or weight of speculation can be measured, this assertion can be tested.
 
 \subsection{Measuring the intensity of speculative sentiment}
-The measurement of speculative intensity must identify how strongly speculators believe that the market will move in a particular direction.
-In the absence of a daily survey of speculators, this study uses \emph{risk-reversal skew} as an objective measure of speculative sentiment rather than surveys of opinion.%\footnote{There are a number of surveys of investor opinion that are regularly conducted by financial service information companies.
-However, these are disparate and, in many cases, unsystematic, and it is not clear whether they measure speculative sentiment or investor opinion more generally.
-The Wall Street Journal runs an annual survey of investor opinion, but this is focused mostly on US equities.
-The \emph{Economist} magazine also solicits opinion, but in a relatively unstructured and \emph{ad hoc} fashion.
-In the foreign exchange market, there is a broad set of market forecasts that was collected by the financial services company \emph{MMS International} (formerly Money Market Services) which polled a broad range of financial market participants on key exchange rates for one and four week forecast horizons.
-Frankel and Froot found that the median forecast from the MMS survey exhibited trend-following behaviour at the one to three week horizon while there is mean-reversion beyond that point.
-Frankel and Froot also assess the 'Financial Report' that has been produced by \emph{The Economist} magazine and \emph{Currency Forecasters Digest} and find a wide dispersion of opinion and express some concern about the aggregation of such divergent forecasters \citet gc{FrootFrankelFDB} and \citep{FrankelFroot1987}.}
-
-Risk-reversal skew is the relative price of currency put and call options for the same strike price.
-The \emph{Put-Call-Parity} in the  \citet{Black1973Options} option pricing model assumes that these will be the same, in practice they are not\footnote{While the Black-Scholes model is the basis for valuing options, where the price of a non-dividend paying stock trades at a discount to the forward rate equal to the risk free rate, the forward rate in a currency option is determined by the relative interest rate differential between the two currencies.
+The measurement of speculative intensity must identify how far speculators believe that the market will move in a particular direction.
+In the De-Long model, this would mean that for Equation \ref{eq:NP2}, it is the extreme deviation of $\rho$ from one.
+In the absence of a daily survey of speculators, this study uses a unique series of \emph{risk-reversal skew} as an objective measure of speculative sentiment.\footnote{It appears that this is the only period for which the data currently exist.
+The risk-reversal data is very difficult to obtain in an over-the-counter market like foreign exchange where there is no centralised exchange to record and collect information, generally being the property of quoting institutions.
+The risk-reversals are the average prices quoted at 10:00 GMT by major banks on the Thomson-Reuters composite option pages (OPT1)  for the period between January 2 1996 and September 6 2002.
+The risk-reversals are the spread of implied volatility for a 1 month 25 delta at-the-money call option relative to the equivalent put.
+Risk-reversal skew is the relative price of currency put and call options for the same strike price.}  The \emph{Put-Call-Parity} in the  \citet{Black1973Options} option pricing model assumes that these will be the same, in practice they are not\footnote{While the Black-Scholes model is the basis for valuing options, where the price of a non-dividend paying stock trades at a discount to the forward rate equal to the risk free rate, the forward rate in a currency option is determined by the relative interest rate differential between the two currencies.
 The \citet{GarmanKohlhagan} adjustment that accounts for the difference in interest payments on the two currencies is used.
@@
 This model allows the market practice of pricing options from the volatility implied by the model.
-In other words, the Garman-Kohlhagan is used to extract the volatility that is implied by a specific option price given the strike and other parameters}.
-
-This study uses a unique dataset of \emph{risk-reversals} for eleven major currency pairs.
-It appears that this is the only period for which the data currently exist.
-The risk-reversal data is very difficult to obtain in an over-the-counter market like foreign exchange where there is no centralised exchange to record and collect information, generally being the property of quoting institutions.
-The risk-reversals are the average prices quoted at 10:00 GMT by major banks on the Thomson-Reuters composite option pages (OPT1)  for the period between January 2 1996 and September 6 2002.
-The risk-reversals are the spread of implied volatility for a 1 month 25 delta at-the-money call option relative to the equivalent put.
-A call option is the right to buy and a put option is the right to sell.
+In other words, the Garman-Kohlhagan is used to extract the volatility that is implied by a specific option price given the strike and other parameters.}  A call option is the right to buy and a put option is the right to sell.
 The delta of an option measures the effect of a change in the spot price on the value of the option.
@@
 
-If the exchange rate trajectory is expected to follow a random walk with drift towards the 1 month forward, implied volatility on the put and the call for the same delta should be equal.
+If the exchange rate is expected to follow a random walk with drift towards the 1 month forward, implied volatility on the put and the call for the same delta should be equal.
 Any deviation from this equivalence is an indication that a deviation from the random walk with drift is implied in the market price of options.\footnote{If the risk-reversal for a 1-month USD/CAD is quoted at 0.15-0.28\% and the implied volatility is 8.50\%, the market-maker would be willing to buy the 1-month 25 delta USD-put-CAD-call at 8.65\% and sell the USD-call-CAD-put at 8.50\%.
@@
 }  Therefore, the relationship between implied volatility for calls and implied volatilities for puts for the same delta and maturity is an indication of how much more expensive calls are relative to the equivalent put or as the market bias towards puts over calls and as an indication of market sentiment.
-The larger the absolute level of the risk-reversal skew, the greater the relative price and, in equilibrium, the more intense the belief, expressed through the action of market participants or in the inventory of market-makers, that there will be a deviation from the forward rate.
-If sentiment is based   on noise rather than information, extremes should presage a sharp reversal; if speculative sentiment is informed, this is part of the price discovery process and extremes may be followed by a random walk.
+The larger the absolute level of the risk-reversal skew, the greater the relative price and, if the deviation is based on information, subsequent price action should be zero on average; if the deviation is based on noise, there should be a good chance of a reversal.
 
@@
 However, there are some exchange-traded currency derivative markets in the US, and participants in these markets are required to report their positions to the US regulators.
-The Commodity Futures Trading Commission (CFTC) was set up by Congress in 1974 to regulate commodity future and options markets.
-The CFTC collect weekly information about the open positions held by private entities in the US derivative markets.
-A sub-set of this information is released to the public each week as the \emph{Commitment of Traders Report} (CoT).
-The weekly release has been in place since September 30, 1992.
-Before that the data were released every two weeks.
-The CoT is released each Friday and details the positions outstanding the previous Tuesday.
-See \citet{cot}.
-
-The CFTC require traders to categorise themselves as being either commercial (C) or non-commercial (NC).
+The Commodity Futures Trading Commission (CFTC) collect weekly information about the open positions held by private entities in the US derivative markets.
+A sub-set of this information is released to the public each week as the \emph{Commitment of Traders Report} (CoT).\footnote{The \citet{cot} require traders to categorise themselves as being either commercial (C) or non-commercial (NC).
 Commercial traders must have some underlying business interest in the security, commodity or instrument that they are trading.
@@
 The absence of these data seems to be a result of problems with the CFTC database.
-The contracts are for EUR 100,000; GBP 62,500; JPY 12,500,000; CHF 100,000 and CAD 100,000.
-
-Table \ref{tabref:DS4} shows the descriptive statistics for the two measures of speculative sentiment.
-The first (S1) is measured as the net non-commercial long and non-commercial short proportionate to the sum of non-commercial positions long and short positions; (S2) is measured as the net non-commercial long and spreading positions proportionate to open interest.
-The first captures directional speculation and, as can be seen from a number of minimums, this can reach 100\%.
-The second records the weight of speculation relative to everything else in the market.
-The former is more sensitive, as can be seen by the range and the standard deviation; the second is usually above zero because there are always some spreading positions that seek to take advantage of calendar spreads and other strategies.
+The contracts are for EUR 100,000; GBP 62,500; JPY 12,500,000; CHF 100,000 and CAD 100,000.}
 
 \subsection{Analysis of events and abnormal returns}
-The aim of the study is to assess the effect of extreme speculative positions on future returns to ascertain the informational content of these extremes.
-Therefore, the events that will be used to assess abnormal returns are the extreme readings on sentiment and the extreme presence of speculators.
-Any reading above the ninety-ninth percentile or below the first percentile is deemed to be extreme.
-For the risk reversal skew, as a secondary test, a reading above the ninety-fifth and below the fifth percentile is also established as an event.
-The  R package \emph{eventstudies} \citet{eventstudies} was used to manage the data.
-The cumulative abnormal returns for the whole of the event window (CARW) and the cumulative abnormal returns for after the event (CARA) were recorded.
-These cumulations are just the summation of the continuously compounded returns over the appropriate period.
-
-The date of the events, 1st, 5th, 95th or 99th percentile, reading of the risk-reversal skew or the level of speculative positions,  are identified and the exchange rate returns for the window around the event are aligned so that the average for each event and each event day can be calculated.
-For example, in Table \ref{tabref:RR1} the analysis of the event that the risk-reversal skew for the EUR vs the US dollar is an extreme low (measured as being equal to the 5th percentile or lower), all 106 cases of this extreme for the period under study are aligned and the average cumulative returns for the whole window (CARW - one day before and after the event in column four) and the post-event window (CARA - the event and the day after the event) are calculated.
-A \emph{bootstrap} is conducted.
-This means that a sample of size 106 (in this case), with replacement, is drawn from each of the event days and new means are calculated for the CARW and CARA.
-If 95\% of the means that are calculated after drawing random samples from the array of windows are above or below zero, it is asserted that the figure is statistically different from zero, and this is represented by an asterisk in the table.
+An \emph{event study} is used to assess the effect of extreme speculation on future returns.
+See \citet{Dolly1933} for the original version of the event study method and \citet{FamaFisherJensenRoll} for an overview of the method used in this study.
+The R package \emph{eventstudies} created by \citet{eventstudies} is used to manage the data.
+Extreme measures of speculative sentiment and activity are the event and the cumulative abnormal returns for the whole of the event window (CARW) and the cumulative abnormal returns for after the event (CARA) are the exchange rates to be assessed.
 
 \section{Analysis of the results}
-Table \ref{tabref:RR1} shows the result of the event studies that were carried out on extreme risk reversal.
+\subsection{Results for extreme speculative sentiment}
+Table \ref{tabref:RR1} records the result of the event studies that were carried out on extreme risk reversal.
 For each exchange rate the second column distinguishes between extreme high and extreme low, the third  column identifies the number of extreme events that were recorded and the next two are the cumulative average returns for the whole window (CARW) and the cumulative average returns for the period of the event and afterwards (CARA).
@@
 
+As speculators become more extreme in their opinion, exchange rates are driven in the direction of sentiment.
 Looking down the seventh column (the 4-day CARW) shows the average cumulative return for the period from four days before the extreme event to the four days after the extreme for extreme sentiment highs and lows of each currency pair.
 The extreme highs are (in this case) those points when the risk-reversal skew is at or above the 95th percentile.
-It should not be surprising to see that the extreme highs are associated with positive returns, which appear in many cases to be significantly different from zero (where significance is identified as more than 95\% of 1000 random bootstrap samples taken from the events being above zero).
-
-It is clear that, as speculators become more extreme in their opinion, prices are driven in the direction of sentiment.
-It shows that the spot price is following the option risk-reversal skew and, as Figure \ref{fig:ES1} reveals, it would be even more pronounced if the event window just ran from the starting point to the event.
-The findings here are consistent with the evidence that orders and sentiment drive prices \citet{Evans2002Order}, \citet{FuturesSanders} and \citet{FuturesWang}.
-
-What is more interesting is whether knowledge about the extreme reading at time $t = 0$ provides any information about future returns.
-If this speculation is noise, Friedman's informed traders should take advantage of that and a reversal should be seen.
-This information is provided in columns 5, 8, 11 and 14 of Table \ref{tabref:RR1}.
-A reversal is indicated by a significant negative reading for highs or positive reading for lows.
-Looking, for example, at the four-day window in column 8, it is clear that the extreme risk reversal skew does not provide any information about future returns.
-Of the eighteen cases using the 95\% th percentile to identify an extreme, ten show a continuation, rather than a reversal, that is not significantly different from zero; one shows a significant continuation; five show reversals that are not significant; only two show significant reversal (EURCHF in each case).
-For the more extreme event when the 99\% the percentile is used to measure the extreme, there is even more ambiguity.
-Just over half the cases with the four-day window show a reversal (8 of 18), but only 2 of these are statistically significant.
+For the whole event window (CARW), extreme highs (labelled "Hi") are associated with positive returns and the extreme lows (labelled "Lo") are associated with negative returns.
+In most cases these positive and negative returns appear to be statistically significant.
+Significance is identified as those cases where more than 95\% of 1000 random bootstrap samples taken from the events being above zero.
+Mean values of the cumulative returns are calculated from samples of the event days and the distribution of these means are used to construct the confidence intervals.
+Returns that are statistically different from zero are identified by an asterisk.
 
@@
 
-The results do not appear to be a function of the percentile that is used to measure the extreme event or the size of the window that is selected.
+That sentiment drives price is shows that speculators have influence on the market.
+This is consistent with the evidence of previous research from the likes of \citet{Evans2002Order}, \citet{FuturesSanders} and \citet{FuturesWang} that orders and sentiment drive prices.
+What is more interesting is whether knowledge about the extreme reading at time $t = 0$ provides any information about future returns.
+If this speculation is noise, informed traders should take advantage of the space that has been created between the price and the fundamental value and a reversal should be seen.
+Whether this is the case is recorded in columns 5, 8, 11 and 14 of Table \ref{tabref:RR1}.
+A reversal is indicated by a significant negative reading for highs or positive reading for lows.
+Looking, for example, at the four-day window in column 8, it is clear that the extreme risk reversal skew does not provide any information about future returns.
+Of the eighteen cases using the 95\% th percentile to identify an extreme, ten show a continuation rather than a reversal that is not significantly different from zero; one shows a significant continuation; five show reversals that are not significant; only two show significant reversal (EURCHF in each case).
+For the more extreme event when the 99\% the percentile is used to measure the extreme, there is even more ambiguity.
+Just over half the cases with the four-day window show a reversal (8 of 18), but only 2 of these are statistically significant.
+Exchange rate prices are as likely to go up as down after an extreme and there does not appear to be any information about the future from knowledge of these extremes.
+\begin{figure}
+\graphicspath{{../Figures/}}
+\centering
+\includegraphics[scale=0.8]{RRCum16}
+\caption{Event Study: Extreme (High 99th percentile) Risk Reversal Skew and 16 day event window: Risk reversals are the ratio of implied volatility on calls to the equivalent put; extreme high is the highest 99th percentile for the whole range, cumulative returns are the sum of log exchange rate returns.
+The event day is the day of the extreme risk reversal reading as a proxy for extreme speculative sentiment.
+If speculation is uninformed, extremes should be followed by reversals; if speculation is informed, extremes are more likely to be followed by a random walk.}
+\label{fig:ES5}
+\end{figure}
+
+These results do not appear to be a function of the percentile that is used to measure the extreme event or the size of the window that is selected.
 Additional cases were run using percentiles at 80\% and 99.5\%.
@@
 The figures reveal the mean cumulative return (in solid red) and the 95\% confidence intervals (dashed blue).
+The confidence intervals are created with the bootstrap as outlined above.
 The only case where there seems to be any evidence of a reversal is for the extreme low risk-reversals on the Canadian dollar.
 
-\begin{figure}[h!]
+\begin{figure}
 \graphicspath{{../Figures/}}
 \centering
-\caption{Event Study:  Extreme (High 99th percentile) Risk Reversal Skew and 16 day event window}
-\label{fig:ES5}
-\includegraphics[scale=0.8]{RRCum16}
-\end{figure}
-% the reference is ES5 to avoid conflict with the earlier version.
-
-\begin{figure}[h!]
-\graphicspath{{../Figures/}}
-\centering
-\caption{Event Study:  Extreme (Low - 10th percentile) Risk Reversal Skew and 16 day event window}
+\includegraphics[scale=0.8]{RRCum16a}
+\caption{Event Study:  Extreme (Low - 10th percentile) Risk Reversal Skew and 16 day event window: Risk reversals are the ratio of implied volatility on calls to the equivalent put; extreme high is the highest 99th percentile for the whole range, cumulative returns are the sum of log exchange rate returns.
+The event day is the day of the extreme risk reversal reading as a proxy for extreme speculative sentiment.
+If speculation is uninformed, extremes should be followed by reversals; if speculation is informed, extremes are more likely to be followed by a random walk.}
 \label{fig:ES2}
-\includegraphics[scale=0.8]{RRCum16a}
 \end{figure}
 
-However, it does indicate that the evolution of exchange rates are not a random walk.
-There is evidence here that prices gradually respond to information, as would be the case if there were institutional or psychological features that caused momentum or gradual adjustment.
-The positive (or negative) expectations that are evident in the risk reversal skew do not seem to represent a deviation or a mis-pricing that will be reversed.
-Rather they seem to be part of the price-discovery process.
+Though it appears that the speculation in examples is informed, it is clear from figures \ref{fig:ES1} and \ref{fig:ES2} that exchange rates do not react instantly to the news.
+There is a gradual evolution.
+This is consistent with the evidence of serial correlation in asset prices that has been identified.
+See, for example, \citet{Jagadeesh}.
+It appears that information is gradually absorbed into the price.\footnote{It may be that exchange rate traders take time to adjust to the information that comes from order flow or it may be evidence of the \emph{conservatism heuristic} that suggests that people are slow to change their opinion to new information.}
+
+It is also possible to look at the distribution exchange rate returns after these extremes and to compare them to the distribution of the whole sample.
+If speculation is informed, we expect to see a distribution of return that is largely symmetrical around zero and very similar to the distribution for the whole sample.
+The distribution of return is very similar but, but much broader than that of the whole distribution.
+Conventionally, this would suggest that there is greater risk.
+It may be that the increase in information adds to the level of uncertainty in the market.
+% Add the table with the descriptive statistics for returns that follow an extreme compared to the whole sample. 
+\subsection{Results for extreme speculative positions}
+Table \ref{tabref:SP1} shows the results from the event studies that were carried out using speculative positions as reported to US regulators.
+The table works in a similar way to Table \ref{tabref:RR1}.
+The first column is the exchange rate, defined as units per US dollar.
+The table is then split into two parts.
+The first part uses the S1 measure, the net non-commercial (speculative) positions relative to total non-commercial (speculative) positions; the second is S2 as the net non-commercial position relative to open interest or the total outstanding open positions.
+Therefore, the first captures the balance of speculative positions while the second captures the balance of speculative positions relative to the whole market.
+If speculators become more prevalent or dominant in the market, this will show up in S2 but not S1.
+S1 is more sensitive to changes in sentiment and more volatile.
+Each of these sections is broken into the results for a 2 week window and a 4 week window.
+There is a row for extreme high and extreme low which represents extreme long positions or extreme short positions; the first  column identifies the number of cases; CARW is the cumulative abnormal return for the whole event window; CARA is the cumulative abnormal return for the period after the event to the end of the window.
 
@@
 
-Table \ref{tabref:SP1} shows the results from the event studies that were carried out using speculative positions as reported to the US regulators.
-The table works in a similar way to Table \ref{tabref:RR1}.
-The first column is the exchange rate, defined as units per US dollar.
-The table is then split into two parts.
-The first part uses the measure S1 which is the net non-commercial (speculative) positions relative to total non-commercial (speculative) positions; the second is S2 as the net non-commercial position relative to open interest or the total outstanding open positions.
-Therefore, the first captures the balance of speculative positions while the second captures the balance of speculative positions relative to the whole market.
-If speculators become more prevalent or dominant in the market, this will show up in S2 but not S1.
-S1 is more sensitive to changes in sentiment and more volatile.
-Each of these sections is broken into the results for a 2 week window and a 4 week window.
-Each of these has the same pattern.
-There is a row for extreme high and extreme low which represents extreme long positions or extreme short positions; the first  column identifies the number of cases; CARW is the cumulative abnormal return for the whole event window; CARA is the cumulative abnormal return for the period after the event to the end of the window.
+Once again, there is a very strong association between returns and the activity of speculators.
+For the 4-day window and the S1 and S2 measures, every one of the currencies studied, high and low, had a mean cumulative return for the whole of the event window that was in the same direction as sentiment.
+When re-sampled a thousand times, more than ninety five percent of the means calculated were greater than zero in all cases but one (GBP).
+There is evidence here of momentum and of prices following the movement of speculator positions.
+This consistent with both the gradual absorption of information into the price and with the feedback of price into additional speculation.
 
-\begin{figure}[h!]
+\begin{figure}
 \graphicspath{{../Figures/}}
 \centering
-\caption{Event Study:  Extreme (95th percentile), Non-commercial net per OI (S2), 6 week event window}
+\includegraphics[scale=0.8]{FPCum6w}
+\caption{Event Study:  Extreme (95th percentile), Non-commercial net short per OI (S2), 6 week event window.
+Non-commercial net per OI (S2) are the net speculative positions (as reported to US regulator the CFTC) relative to all open positions; extreme low is the highest 95th percentile for the whole range, cumulative returns are the sum of log exchange rate returns.
+The event day is the day of the extreme risk reversal reading as a measure of extreme speculative activity.
+If speculation is uninformed, extremes should be followed by reversals; if speculation is informed, extremes are more likely to be followed by a random walk.}
 \label{fig:ES3}
-\includegraphics[scale=0.8]{FPCum6w}
 \end{figure}
 
-Once again, there is a very strong association between returns and the movement of speculators.
-For the 4-day window and the S1 and S2 measures, every one of the currencies studied, high and low, had a mean cumulative return for the whole of the event window that was in the same direction as sentiment.
-When re-sampled a thousand times, more than ninety five percent of the means calculated were greater than zero in all cases but one (GBP).
-Once again there is evidence here of momentum and of speculator positions following the movement of prices.
-This consistent with both the gradual absorption of information into the price and with the feedback of price into additional speculation.
+\begin{figure}
+\graphicspath{{../Figures/}}
+\centering
+\includegraphics[scale=0.8]{FPCum6wa}
+\caption{Event Study:  Extreme (90th percentile), Non-commercial net long (S1), 6 week window.
+Non-commercial net per OI (S2) are the net speculative positions (as reported to US regulator the CFTC) relative to all open positions; extreme high is the highest 95th percentile for the whole range, cumulative returns are the sum of log exchange rate returns.
+The event day is the day of the extreme risk reversal reading as a measure of extreme speculative activity.
+If speculation is uninformed, extremes should be followed by reversals; if speculation is informed, extremes are more likely to be followed by a random walk.}
+\label{fig:ES4}
+\end{figure}
 
-However, there is little evidence that these speculators are driving prices to extreme deviations from fundamentals with their extreme behaviour.
+There is little evidence that these self-reported speculators are driving prices away from fundamentals with their activity.
 Sixty percent of the S1 cases still show a significant abnormal return in the direction of the extreme 4 weeks after the extreme has been reached; only thirty percent of the cases show a reversal and in only one of those do ninety five percent of the re-sampled means remain above zero.
 For S2 there is even less ambiguity, with ninety percent of the cases showing a continuation and forty percent of them being significant.
-Again the evidence is consistent with the gradual adjustment of prices to information.
+The evidence here with speculative positions is the same as that of speculative sentiment.
+Prices move in the direction of speculative activity but speculative extremes do not provide information about the future.
+The price action of the event study is consistent with speculation being informed rather than uninformed.
 
-\begin{figure}[h!]
-\graphicspath{{../Figures/}}
-\centering
-\caption{Event Study:  Extreme (90th percentile), Non-commercial net long (S1), 6 week window}
-\label{fig:ES4}
-\includegraphics[scale=0.8]{FPCum6wa}
-\end{figure}
-
-The event windows are stretched to give a clear view of the evolution of exchange rates  around the event and to ensure that results are not a function of the size of the event window.
-Confidence intervals are added\footnote{It is possible to make changes to the estimate of the \emph{expected exchange rate}, the extreme measurement, quantile, window length and the high or low indicator and for alternative assessments of speculative extremes to be used.
-Adjustments to these parameters show that these results are not a function of the choices presented in the main text.
-They are robust to alternative specifications.}.
-
-Figure \ref{fig:ES3} shows the performance of the cumulative abnormal returns around a six week event window where the event is the net low (short) speculative positions relative to total activity (open interest) at or below the ($5\%)$ level.
-This means that the speculators are inclined to be short the non-domestic (US dollar) currency.
-While the Euro continues to move lower after the extreme short positions, the other major currencies move broadly sideways, consistent with the settlement around an equilibrium after a period where information is being absorbed.
-This is also consistent with market efficiency where a random walk would be expected after an extreme.
-
-Figure \ref{fig:ES4} shows the abnormal returns around extreme high (90th percentile and long position).
-There is some tentative evidence of continuation for speculative positions in the Canadian dollar and signs of reversal for UK sterling.
-However, the Euro and the Japanese yen move sideways after the event.
-There is no clear pattern after the extreme and therefore no information.
+Figures \ref{fig:ES3} and \ref{fig:ES4} indicate that these results are not a function of the event window that has been chosen.
+Looking at the 95 percentile confidence intervals (once again constructed by bootstrap re-sampling) indicates that in all the cases, there is a large amount of uncertainty about the future direction well after the extreme reading of speculative activity.
+It does not appear that informed traders are constrained in the short-run from driving prices back to equilibrium.
+It appears that the informed traders are driving prices to a new level.
 
 \section{Speculation and foreign exchange returns}
-This paper has sought to measure speculation and to use this measure to discover more about the relationship between speculation, information and asset prices.  
-Whether measured as intensity of speculative sentiment or as the weight of speculator action, it is clear that extreme speculation does not provide information about future returns.
-This means that speculation is more likely to be informed than uninformed.
-If it were uninformed it would be much more likely that these extremes would be follow by reversals as informed market participants took advantage of the space created by this uninformed speculative activity.
-
-The possibility that noise-trader risk limits the activity of informed traders is limited by the large windows that are employed when constructing figures for the event study.
-Informed traders may fear the uninformed in the short-run and this may delay the chance of a reversal but over the medium-term reversals would be expected.
-This event study uses 20 days and 6 weeks for the daily and weekly event windows.
-
-The nature speculation is central to a number of contemporary policy questions.
+The nature speculative activity in financial markets is central to a number of contemporary policy questions.
 Amongst these are the design and regulation of financial markets, the surge in commodity prices and the limits put on the activities of financial institutions.
 The debate about a tax on financial transactions, initiated by \citet{TobinTax} and now called the \emph{the Tobin tax}, has been intended to reduce speculation.
-The argument for a tax on financial transactions depends on the idea that speculators will drive the price away from its fundamental value.
-However, if speculation is part of the process of price discovery, reducing its activity may incur costs that offset some of the benefits that are expected.
+This is based on the belief that speculation is uninformed, that it drive prices away from their fundamental value and as a consequence  increases the volatility of financial asset prices.
+The argument for a tax on financial transactions can be presented as a tax on an externality with the level of volatility of financial markets being the public cost.
 
-The reduction in speculation in the US government bond market that has been seen since the introduction of new costs and restrictions under the array of ????
-regulations has been seen by some to reduce liquidity and increase volatility, rising risk and imposing additional costs for final investors.
+The increase in volatility that is evident when speculative activity is high seems to provide some support to this argument.
+However, it is not clear that the direction of causality is from speculation to volatility.
+An alternative view would suggest that an increase in information raises the amount of informed speculation and that this increase in information naturally increase the volatility of financial assets.
+If speculation is part of the process of price discovery, reducing its activity may reduce liquidity and hinder the process of price discovery.
+In that case unanticipated costs may offset some of the benefits that are expected from the tax on speculation.
+
+Regulatory changes in the US government bond market give some hints at the way that reduce speculation can affect the smooth functioning of markets.\footnote{Increased capital requirements and the so-called \emph{Volcker Rule} that would prevent investment banks from trading on their own account have increased the cost of trading US government bonds and reduced the ability of some key players to participate in the market.}  Though bid-ask spreads have remained rather stable, the \emph{market-depth} and the \emph{price impact} of large trades has deteriorated in the last two years.
+There have also been a number of occasions where very large price changes and an evaporation of liquidity have affected the market.
 
@@
 
-The notion that speculation is at least partially informed is supported by the fact that there is little evidence from the event studies conducted here that speculation causes overshooting or bubbles.
-When speculation is extreme, whether measured by intensity of belief or weight of speculative positions, future returns are as likely to be positive as negative.
-There is no informational content as there would be if speculation was pure noise.
-In the simulation of the De Long model, extremes are more likely to be followed by reversals.
-In the foreign exchange market, that is not the case.
-There are some large price crashes in the dataset, as shown by the descriptive statistics of the returns.
-For example  in Table \ref{tabref:DS1} it is seen that on one day there is a 10.65\% appreciation of the Japanese yen and Table \ref{tabref:DS2} shows a week when the Canadian dollar falls by 10\%.
-However, the measures of speculative sentiment or the weight of speculation are not particularly high in either of these cases.
-Indeed, an analysis of the top 1\% of positive and negative foreign exchange returns in the dataset reveals that sentiment and positions for this group were very close to the average of the whole set.
-This would be consistent with a model where extreme price movement was a function of new information rather than the structure of existing positions.
+This paper has sought to measure speculation and to use this measure to discover more about the relationship between speculation, information and asset prices.  
+Whether measured as speculative sentiment or as the weight of speculator action, it is clear that extreme speculation does not provide information about future returns.
+This means that speculation is more likely to be informed than uninformed.
+If it were uninformed it would be much more likely that these extremes would be follow by reversals as informed market participants took advantage of the space created by this uninformed speculative activity.
 
-\bibliography{../../../../myrefs}
-\begin{appendices}
-
-\section{Appendix 1}\label{apref:1}
-
-\begin{sidewaystable}
-\begin{threeparttable}
-\caption{Appendix 1: Descriptive Statistics for Risk Reversals: 2 January 1996 to 6 September 2002}
-\begin{tabular}{l | r r  r r r r r r r r}
-  \hline
- Statistic &  AUD & EURGBP & EURSEK & GBP & CHF & EURCHF & EURJPY & EUR & CAD & JPY \\
- \hline
-Number & 1702 & 1697 & 1719 & 1691 & 1374 & 1339 & 1697 & 1692 & 1701 & 1697 \\ 
-  Mean & -0.37 & 0.15 & 0.71 & -0.08 & -0.28 & -0.38 & -0.48 & 0.19 & 0.12 & -0.74 \\ 
-  Median & -0.30 & 0.10 & 0.80 & -0.10 & -0.15 & -0.30 & -0.10 & 0.10 & 0.15 & -0.55 \\ 
-  Standard Deviation & 0.38 & 0.37 & 0.51 & 0.33 & 0.49 & 0.36 & 1.02 & 0.49 & 0.25 & 1.02 \\ 
- Max & 0.45 & 1.72 & 1.90 & 1.00 & 1.05 & 0.50 & 1.70 & 2.05 & 1.45 & 2.05 \\ 
- Min & -1.70 & -0.80 & -0.30 & -1.35 & -2.05 & -1.80 & -4.03 & -1.15 & -0.65 & -4.00 \\ 
- Skewness & -0.50* & 1.27* & -0.23* & 0.33* & -0.79* & -0.58* & -0.93* & 0.68* & 0.23* & -0.51* \\ 
-  Kurtosis & 0.02 & 2.29* & -1.67 & 0.72 & 0.69 & 0.19 & 0.15 & 0.96 & 1.83 & -0.03 \\ 
-      \hline
- \label{tabref:DS2}
-\end{tabular}
-\begin{tablenotes}
-\small
-\item Daily risk reversal skew.
-Risk reversal is the premium of implied volatility for calls relative to the equivalent put (equal distance to the strike price, 25 delta in this case).
-A figure above zero means that calls are more expensive (in volatility and price terms) than puts.
-A figure below zero means that puts are more expensive.
-The asterisk (*) means that the estimate is significant at the 5\% level using t-statistics for the mean and the standard error of skew and standard error of excess-kurtosis respectively.
-These are calculated as $SES = \sqrt{\frac{6n(n-1)}{(n-2)(N+1)(N+3)}}$ and $SEK = \sqrt{\frac{n^2-1}{(n-3)(N+5)}}$, with the assumption that this estimate is distributed normally, the skew or kurtosis is considered significant if it is more than two standard errors from zero.
-\end{tablenotes}
-\end{threeparttable}
-\end{sidewaystable}
-
-\section{Appendix 2}
-\begin{sidewaystable}
-\begin{threeparttable}
-\caption{Appendix 2: Descriptive Statistics for S1 (Speculative Sentiment) and S2 (Speculative Weight): 30 September 1998 to 31 December 2008}
-\begin{tabular}{l | r r r r r r r  r r r }
-  \hline
- Statistic &  \multicolumn{2}{c}{EUR} & \multicolumn{2}{c}{JPY} & \multicolumn{2}{c}{GBP} & \multicolumn{2}{c}{CHF} & \multicolumn{2}{c}{CAD}  \\
- & S1 & S2 & S1 & S2 & S1 & S2 & S1 & S2 & S1 & S2 \\  
-\hline
-Number & 522 &  522 & 849 & 849 & 849 & 849 & 847 & 847 & 849 & 849\\
-Mean &  31.31 & 43.43 & -23.90 & 44.21 & 2.76 & 41.68 & -21.14 & 46.13 & 6.68 & 41.23\\
-Median & 42.75 & 39.56 & -33.30 & 43.56 & 11.15 & 38.88 & -37.32 & 42.88 & 7.44 & 39.81\\
-Standard deviation  & 41.94 & 16.13 & 51.49 & 12.17 & 60.59 & 16.46 & 60.76 & 15.73 & 55.26 & 15.33\\
-Max & 95.67 & 86.06 & 87.43 & 84.50 & 96.15 & 89.45 & 100.00 & 92.49 & 100.00 & 86.63\\
-Min &  -100.00 & 11.15 & -99.83 & 12.88 & -100.00 & 9.10 & -100.00 & 14.61 & -90.08 & 11.89\\
-Skew  & -0.88* & 0.55* & 0.33* & 0.57* & -0.17 & 0.76* & 0.45* & 0.65* & 0.02 & 0.50*\\
-Kurtosis & 0.18 & -0.57 & -1.14 & 0.64 & -1.45 & 0.15 & -1.25 & -0.15 & -1.35 & -0.28\\
-\hline
- \label{tabref:DS4}
-\end{tabular}
-\begin{tablenotes}
-\small
-\item S1 is the first sentiment indicator, calculated as ratio of net non-commercial long less non-commercial short to the sum of non-commercial long and short positions multiplied by one hundred.
-S2 is the second sentiment indicators, calculated as the ratio of net non-commercial long to net non-commercial short plus spreading to open interest.
-An asterisk (*) means that the estimate is significantly different from zero at the 5\% level using t-statistics for the mean and the standard error of skew and standard error of excess-kurtosis respectively.
-These are calculated as $SES = \sqrt{\frac{6n(n-1)}{(n-2)(N+1)(N+3)}}$ and $SEK = \sqrt{\frac{n^2-1}{(n-3)(N+5)}}$, with the assumption that this estimate is distributed normally, the skew or kurtosis is considered significant if it is more than two standard errors from zero.
-\end{tablenotes}
-\end{threeparttable}
-\end{sidewaystable}
-
-\end{appendices}
+\bibliography{../../myrefs}
 

### 2015-09-15 R/Extreme2.R

@@ -12,7 +12,7 @@ daz <- read.zoo("Data/RR.csv", header = TRUE,
 currencydf <- function(currency){
 hi = as.numeric(0.95)
 lo = as.numeric(0.05)
-# are these needed?
+# these set the high and low quantoles.  Can be changed or programmed.
 FX.z <-daz[,c(paste(currency), paste(currency, "RR", sep = ""))]
 FX.z$r <- 100*diff(log(FX.z[,1]))
 # THis has been changed from the original so that the dataframe has a high and low
@@ -29,57 +29,85 @@ if(ex == "high"){
 }
 return(FX.z)
 }
-EUR <- currencydf("EUR")
-head(EUR)
+GBP <- currencydf("GBP")
+head(GBP)
+#-------------------------------------------------------------
+# Create the currencydf files for each currency and put them in a list
+clist <- list()
+currencylist <- colnames(daz)[seq(2, 20, 2)]
+for(i in currencylist){
+# this assign idea comes frmom 
+#http://librestats.com/2013/12/16/rules-for-naming-objects-in-r/
+clist[i] <- list(assign(x = i, value = currencydf(i)))
+}
 # Create zoo object for events----------------------------------------
 # Events must be stored in particular format.
-# Requires currencydf zoo object, wind for size of window and hilo 
+# Requires  fxdf zoo object created with currencydf function
+# and wind for size of window and hilo for whether extreme high or extreme low 
 #for high or low extreme.
-eventzoo <- function(currencydf, wind, hilo){
+eventzoo <- function(fxdf, wind, hilo){
   if(hilo == "hi"){
     ho = 4
-  } else if(hilo == "low"){
-    ho = 5
+  } else if(hilo == "lo"){
+    ho = 5 
   } else {
     print("No Extreme")
   }
-eventslist.FXex <- data.frame(index(currencydf[,ho][currencydf[,ho] ==1]))
+# THe next line gives the data (taking the index) for the extremes.
+eventslist.FXex <- data.frame(index(fxdf[,ho][fxdf[,ho] ==1]))
 eventslist.FXex$unit <- as.character("r")
[…]

### 2015-09-15 R/Test.R

@@ -1,0 +1,79 @@
+#o Create zoo object for events----------------------------------------
+# Events must be stored in particular format.
+# Requires fxdf which is a zoo object compatible with event studies package.
+# and wind for size of window and hilo for whether extreme high or extreme low 
+#for high or low extreme.
+eventzoo <- function(fxdf, wind, hilo){
+  if(hilo == "hi"){
+    ho = 4
+  } else if(hilo == "lo"){
+    ho = 5 
+  } else {
+    print("No Extreme")
+  }
+# THe next line gives the data (taking the index) for the extremes.
+eventslist.FXex <- data.frame(index(fxdf[,ho][fxdf[,ho] ==1]))
+eventslist.FXex$unit <- as.character("r")
+colnames(eventslist.FXex) <- c("when", "unit")
+eventslist.FXex$when <-as.Date(eventslist.FXex$when)
+# Align returns with extreme events-----------------------
+a <- phys2eventtime(fxdf, eventslist.FXex, width = wind)
+# Create windows (exw - whole, exa - after, cumsum - cumulative sum)
+FX.exw <- window(a$z.e, start = wind * -1, end = wind)
+FX.exa <- window(a$z.e, start = 0, end = wind)
+FX.exwc <- remap.cumsum(FX.exw, is.pc = TRUE, base = 0)
+FX.exac <- remap.cumsum(FX.exa, is.pc = TRUE, base = 0)
+dims = seq(-wind, wind, 1)
+FX.exwcm <- as.matrix(FX.exwc)
+FX.exacm <- as.matrix(FX.exac)
+MRW <- mean(FX.exwcm[(wind*2 + 1),])
+MRA <- mean(FX.exacm[(wind + 1),])
+FX.exwc.b <- inference.Ecar(FX.exwc, to.plot = FALSE)
+FX.exac.b <- inference.Ecar(FX.exac, to.plot = FALSE)
+g <- list(MRW = MRW, MRA = MRA, hilo = hilo, wind = wind, FX.exacm = FX.exacm)
+return(g)
+}
+EURSEK2lo <- eventzoo(fxdf = EURSEK, wind = 2, hilo = "lo")
+index(fxdf[, ho][fxdf[, ho] == 1])
+for(i in currencylist){
+title <- paste(i, 2, "hi", sep = "")
+  eventlist[[title]] <- assign(x = title, value = eventzoo(clist[[i]], 
+                                                   wind = 2, hilo = "hi"))
+}
+
[…]

### 2015-09-15 R/Extreme2.R
Updated some of the Extreme2.R with figures and table.

@@ -93,20 +93,30 @@ eventlist <- lapply(clist, FUN = eventzoo, wind = 2, hilo = "lo")
 # create charts for comparison of extreme and regular returns. 
 # this requires the currencydf and currdf (best to take from list)
 twochart <- function(currencydf, currdf){ 
-par(mfrow = c(2,1))
-title1 <- paste("Extreme", colnames(clist[[1]])[1], sep = " ")
-title2 <- paste("Regular", colnames(clist[[1]])[1], sep = " ")
-plot(density(100*currdf$currdf[2,]), xlim = c(-2.0, 2.0), main = title1,
+par(mfrow = c(2,3))
+title1 <- paste("Extreme", colnames(clist[[i]])[1], sep = " ")
+title2 <- paste("Regular", colnames(clist[[i]])[1], sep = " ")
+plot(density(100*eventlist[[i]]$FX.exacm[2,]), xlim = c(-2.0, 2.0), main = title1,
      lwd = 2, col = "red")
-abline(v = mean(currdf$currdf[2,], na.rm = TRUE))
-text(x = mean(currdf$currdf[2, ], na.rm = TRUE) + 0.2, y = 0.2, labels = "Mean")
+abline(v = mean(eventlist[[i]]$FX.exacm[2,], na.rm = TRUE))
+text(x = mean(eventlist[[i]]$FX.exacm[2,], na.rm = TRUE) + 0.2, y = 0.2, labels = "Mean")
 # need to sample 
-plot(density(sample(currencydf[-1,3], 113), na.rm = TRUE), xlim = c(-2.0, 2.0), 
+plot(density(sample(clist[[i]]$r,113), na.rm = TRUE), xlim = c(-2.0, 2.0), 
      main = title2, lwd = 2, col = "red")
-abline(v = mean(currencydf[-1, 3], na.rm = TRUE))
-text(x = mean(currencydf[-1,3], na.rm = TRUE) + 0.2, y = 0.4, labels = "Mean")
+abline(v = mean(sample(clist[[i]]$r, 113), na.rm = TRUE))
+text(x = mean(sample(clist[[i]]$r, 113), na.rm = TRUE) + 0.2, y = 0.4, labels = "Mean")
 }
-twochart(clist[[10]], eventlist[[10]])
+twochart(clist[[3]], eventlist[[3]])
+#------------------------------------------------
+# This will create the pdf figures for the distribution
+# The graphs are not put on the same page
[…]

### 2015-09-17 Knitr/NewPaper.Rnw
Updated some of the Extreme2.R with figures and table.
Updated Extreme2.R and NewPaper.Rnw Still needs the extreme-regular table to be sorted and the conclusion.

@@
 \section*{Introduction}
-Ideas about the way that speculation affects financial market span a broad range.
+Ideas about the way that speculation affects financial markets span a broad range.
 Towards one end is the view of \citet[p.
-101]{Keynes1936} that speculation is a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd'' \citet[p.
-101]{Keynes1936}.
+101]{Keynes1936} that speculation is a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd''.
 Speculation is a problem that will cause booms, reversals and will discourage fundamental, long-term investment.
-It is noise that emerges from the interaction of collective sentiment to encourages trends in asset prices.
+It is noise that emerges from the interaction of collective sentiment to create trends in asset prices.
+% This description provides an explanation for the observation from \citet{ShillerVol} that stock prices are more volatile than the underlying earnings information on which valuation should be based.
 
@@
 
-The aim of this paper is to understand more about the nature of speculation by investigating the relationship between the intensity of speculation and the weight of speculative positions relative to the movement of prices in the foreign exchange market\footnote{The foreign exchange market is utilised because it is the largest and most liquid financial market in the world.
+The aim of this paper is to understand more about the nature of speculation by investigating the relationship between the intensity of speculation and the weight of speculative positions relative to the movement of prices in the foreign exchange market\footnote{The foreign exchange market is chosen for this study because it is the largest and most liquid financial market in the world.
 The latest survey taken by the Bank for International Settlements (BIS) in April 2013 estimated average daily turnover at \$5.4trn with average daily spot transactions of \$2.0trn See \citet{BISFX2013}.
@@
 If speculators are uninformed and their activity can be regarded as noise, the most extreme sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when reversals are most likely to take place; if speculators are informed prices should follow a random walk after either of these extremes.
+\begin{figure}
+\graphicspath{{../Figures/}}
+\centering
+\includegraphics[scale=0.8]{RRCum16}
+\caption{Event Study:  Extreme (High $99^{th}$ percentile) Risk Reversal Skew and 16 day event window: Risk reversals are the ratio of implied volatility on calls to the equivalent put; extreme high is at or above the $99^{th}$ percentile for the whole range, cumulative returns are the sum of log exchange rate returns.
+The event day is the day of the extreme risk reversal reading as a proxy for extreme speculative sentiment.
+If speculation is uninformed, extremes should be followed by reversals; if speculation is informed, extremes are more likely to be followed by a random walk.}
+\label{fig:ES1}
+\end{figure}
 
 This paper contributes to the debate about the nature of speculation in a number of ways.
+It provides clear evidence that speculation is more informed than uninformed.
 A unique set of option prices are used to identify the intensity of speculative sentiment; an event study method is used to establish the effect of extremes of speculative sentiment and speculative positions; these speculative extremes are compared to a common noise-trader model.
@@
 Figure \ref{fig:ES1} shows the evolution of a selection of foreign exchange prices after speculative sentiment is extreme.
-An event study is conducted with the speculative extreme being the event $(t_0)$,  Sentiment is measured by the price of currency call options relative to the equivalent put.
+An event study is conducted with the speculative extreme being the event $(t_0)$, sentiment is measured by the price of currency call options relative to the equivalent put.
 The foreign exchange rate has a contemporaneous movement with the sentiment.
@@
 
-\begin{figure}
-\graphicspath{{../Figures/}}
-\centering
-\includegraphics[scale=0.8]{RRCum16}
-\caption{Event Study:  Extreme (High 99th percentile) Risk Reversal Skew and 16 day event window: Risk reversals are the ratio of implied volatility on calls to the equivalent put; extreme high is the highest 99th percentile for the whole range, cumulative returns are the sum of log exchange rate returns.
-The event day is the day of the extreme risk reversal reading as a proxy for extreme speculative sentiment.
-If speculation is uninformed, extremes should be followed by reversals; if speculation is informed, extremes are more likely to be followed by a random walk.}
-\label{fig:ES1}
-\end{figure}
-% This description provides an explanation for the observation from \citet{ShillerVol} that stock prices are more volatile than the underlying earnings information on which valuation should be based.
-In addition, \citet{Delong1990noise} and \citet{shleifer1997limits} show that constrained horizons can cause noise-trading to limit the ability of informed traders to take advantage of profitable opportunities for fear of further divergence from fundamental value before realignment is achieved.
+%In addition, \citet{Delong1990noise} and \citet{shleifer1997limits} show that constrained horizons can cause noise-trading to limit the ability of informed traders to take advantage of profitable opportunities for fear of further divergence from fundamental value before realignment is achieved.
 
@@
 For example, regulations or taxes that reduce the amount of speculation may be beneficial if they limit risk-taking and encourage more long-term attitudes towards investment.
-If speculation is largely uninformed it will add to the level of noise and make it even harder for informed investors to overcome uncertainty and apply fundamental analysis about long-term value; will increase the tendency towards activity that is focused on the short-term.
-However, if speculation is informed and part of the process of price discovery, regulatory measures to reduce its influence may also reduce liquidity and undermine the prices of price discovery.
-Increased volatility, heightened risk and reduced transparency together with other costs may provide a counteraction to any positive benefits.
+If speculation is largely uninformed it will add to the level of noise and make it even harder for informed investors to overcome uncertainty and apply fundamental analysis about long-term value; uninformed speculators will increase the tendency towards activity that is focused on the short-term.
+However, if speculation is informed and part of the process of price discovery, regulatory measures to reduce its influence may also reduce liquidity and undermine the process of price discovery.
+In this case, increased volatility, heightened risk and reduced transparency together with other costs may provide a counteraction to any benefits.
 
-The rest of the paper proceeds as follows:  Section Two discusses the distinction between informed and uninformed speculators; Section Three talks about the measurement of speculation and the testing of extreme positions; Section Four reviews the evidence; Section Five concludes with a discussion of policy implications.
+The rest of the paper proceeds as follows:  Section Two discusses the distinction between informed and uninformed speculators; Section Three talks about the measurement of speculation and the event study method; Section Four reviews the evidence; Section Five concludes with a discussion of policy implications.
 
-\section{Informed and uninformed traders}
-Speculation plays an  important role in the areas of market-microstructure, the rationale for the trading of financial securities as well as providing the explanation for rational expectations are supposed to operate.
+\section{Informed and uninformed speculation}
+Speculation plays an  important role in the areas of market-microstructure, the rationale for the trading of financial securities as well as providing the mechanism for the operation of rational expectations.
 However, there is no clear, unambiguous theory of speculation.
@@
 This bid-ask spread is the payment that the market-maker receives from the uninformed in return for taking the risk of meeting an informed trader.
-In a similar way \citet{glosten1985bid}describes a model where the bid-ask spread is a function of adverse selection, being caused by the risk that orders come from informed traders.
+In a similar way \citet{glosten1985bid} describes a model where the bid-ask spread is a function of adverse selection, being caused by the risk that orders come from informed traders.
 
@@
 \citet{EasleyPIN2} develop a technique to estimate the probability that trading is informed.
-The find that the high liquidity stocks have the highest level of information and the most frequent arrival rate of informed traders but that price changes are limited by the volume of uniformed traders.
-Low liquidity stock, by contrast, have fewer information events but informed traders have a much larger effect on prices.
-
-Informed speculation is also used as the explanation for the tendency of Rational Expectations Equilibrium (REE) to prevail in the long-run.
-Those economic agents able to determine the fundamental value or discern the underlying economic model will be able to speculate and drive prices towards equilibrium.
-This tends to leave uninformed speculation as a pure noise that can be characterised as white-noise random variable\footnote{See \citet{muth1961Rational} for the original adjustment to a cobweb model.}.
-While this construction may be justified for the analysis of microstructure and market efficiency, it is an over-simplification that is inappropriate when seeking more understanding of speculation itself.
-Institutional and psychological explanations for noise-trading frequently suggest that there are endogenous non-random elements to speculative activity.
-Noise-trading is associated with psychological biases and restricted information while liquidity-trading is about institutional constraints.
-Therefore, noise is associated with behavioural issues, over-confidence and the use of irrational or purely fashionable investment tools.
-See \citet{DeBondtAnalysts} for an analysis of over-reaction and the \emph{representativeness heuristic}, \citet{OdeanOC} for a study of over-confidence and  \citet{ShillerFashion} for an assessment of investment fads.
-Traders without information may not make decisions independently of others.
-Information may be inferred from the buying and selling conducted by other traders\footnote{See for example \citet[p.101]{Keynes1936} and the beauty contest.}.
-
-Institutional constraints may mean that economic agents just execute customer orders at the current price or it may be that they buy and sell with reference to a benchmark index or to fulfil a hedging strategy.
-The institutional explanations are associated with the allocation of credit for investment and trading purposes:  as asset values decline exchanges, banks and other providers of liquidity require larger margins and deeper haircuts; investors liquidate their fund holdings; the attraction of hedging increases.
-See \citet{BrunnermeierLiquidity} and \citet{Gorton} for some discussion of the way that margin-calls and collateral constraints can force investment decisions.
-
-\citet{FriedmanPositive} said that the informed market participants would drive the uninformed from the market.
-However, a large proportion of uninformed traders in a market can undermine this process.
-\citet{Delong1990noise} and \citet{shleifer1997limits} have presented models where \emph{noise-trader risk} can prevent the arbitrate of informed speculators from driving prices back towards the fundamental value.
-Noise-trader risk is the risk that noise-traders will drive prices even further away from equilibrium before convergence.
-This can limit the activity of informed speculators when they are risk averse.\footnote{This three-period, overlapping generation model with informed and uninformed traders has a price equation \begin{equation}
-p_t = 1 + \frac{\mu(\rho_t - \rho^*)}{1 + r} + \frac{ \mu \rho^*}{r} - \frac{2 \gamma \mu^2 \sigma_{\rho}^2}{r(1 + r^2)}.
-\label{eq:NP2}
-\end{equation} 
- where $p_t$ is the price; $\mu$ is the share of uninformed traders; $\rho^*$ is the average misperception of uninformed traders; $\rho_t$ is the price mis-perception of the uninformed traders at time t; $\gamma$ is the level of risk aversion; $\sigma^2_{p,t+1}$ is the variance of the misperception.}
-
-The De-Long noise-trader model also provides a useful benchmark against which to assess the activity of speculators.
-The following four effects are noted by \citet[pp.
-14-15]{Delong1990noise}:  the \emph{Friedman Effect}, as informed speculators move prices back towards fundamental value; \emph{Price Pressure} is the \emph{noise-trader risk} as the speculative noise-traders drive prices towards their misperceived value; \emph{Hold More} as noise-traders misperceive the value of the risky asset and tend to hold more of it; \emph{Create Space} is the difference between noise-trader perception and the actual value, it creates the opportunity for informed traders to profit from their information.
-In this model, the more numerous the noise-traders, the more volatile prices.
-
-The Friedman-effect is the activity of informed speculators.
-They take advantage of the space that has been created by the price pressure of the uninformed traders to profit from this mis-pricing.
-Uniformed traders tend to hold more when their opinions are strong.
-This is the noise-trader risk that can prevent informed traders from fulfilled their arbitrage role.
-If uninformed traders are very prominent or have strong opinions, prices should show the greatest divergence from fundamental value and the risk of a reversal should be at its greatest.
-
-\subsection{Empirical work}
-Empirical assessment of speculation has been limited by the scarcity of data.
-It is not very easy to distinguish between speculators and other types of investors.
-It is even harder to distinguish between the informed and uninformed.
-One approach has been to assume that noise-traders utilise simple investment strategies and to use tests of these strategies as a means of testing speculation.
-Amongst the strategies that have been used in this way are momentum, contrarian and carry trades.
-However, as these techniques can be profitable, it is not clear that these can be identified as the domain of the uniformed.
-It is also possible to assume that certain investors usually have more information than others.
-For example, hedge funds, investment banks and other professionals can be assumed to be informed while retail or day traders may be uninformed.
+They find that the high liquidity stocks have the highest level of information and the most frequent arrival rate of informed traders but that price changes are limited by the volume of uniformed traders.
+Low liquidity stocks by contrast, have fewer information events but informed traders have a much larger effect on prices.
 
@@
 Under this system, the up-ticks are assumed to be buy orders and the down-ticks are sell orders.
-Up-ticks are classified as buy orders and down-ticks are classified as sell orders.
 These estimates of the market activity can be used to estimate the level of informed trading.
+
+Informed speculation is also used as the explanation for the tendency of Rational Expectations Equilibrium (REE) to prevail in the long-run.
+Those economic agents able to determine the fundamental value or discern the underlying economic model will be able to speculate and drive prices towards equilibrium.
+This tends to leave uninformed speculation as a pure noise that can be characterised as white-noise random variable.\footnote{See \citet{muth1961Rational} for the original adjustment to a cobweb model.} While this construction may be justified for the analysis of microstructure and market efficiency, it is an over-simplification that is inappropriate when seeking more understanding of speculation itself.
+Institutional and psychological explanations for noise-trading frequently suggest that there are endogenous non-random elements to speculative activity.
+
+Noise-trading caused by psychological biases and restricted information is associated with behavioural issues, over-confidence and the use of irrational or purely fashionable investment tools.
+See \citet{DeBondtAnalysts} for an analysis of over-reaction and the \emph{representativeness heuristic}, \citet{OdeanOC} for a study of over-confidence and  \citet{ShillerFashion} for an assessment of investment fads.
+Traders without information may not make decisions independently of others.
+Information may be inferred from the buying and selling conducted by other traders.\footnote{This is an example of the beauty contest that has been presented by \citet[p.101]{Keynes1936}.}
+
+Liquidity trading and institutional constraints may mean that economic agents just execute customer orders at the current price or that they buy and sell with reference to a benchmark index or to fulfil a hedging strategy.
+The institutional explanations are associated with the allocation of credit for investment and trading purposes:  as asset values decline exchanges, banks and other providers of liquidity require larger margins and deeper haircuts; investors liquidate their fund holdings; the attraction of hedging increases.
+See \citet{BrunnermeierLiquidity} and \citet{Gorton} for some discussion of the way that margin-calls and collateral constraints can force investment decisions.
+
+\citet{FriedmanPositive} said that the informed market participants would drive the uninformed from the market.
+However, a large proportion of uninformed traders in a market can undermine this process.
+\citet{Delong1990noise} and \citet{shleifer1997limits} have presented models where the risk that noise-traders will drive prices even further away from equilibrium before convergence can limit the activity of informed speculators when they are risk averse.
+This can be called \emph{noise-trader risk}.\footnote{This three-period, overlapping generation model with informed and uninformed traders has a price equation \begin{equation}
+p_t = 1 + \frac{\mu(\rho_t - \rho^*)}{1 + r} + \frac{ \mu \rho^*}{r} - \frac{2 \gamma \mu^2 \sigma_{\rho}^2}{r(1 + r^2)}.
+\label{eq:NP2}
+\end{equation} 
+ where $p_t$ is the price; $\mu$ is the share of uninformed traders; $\rho^*$ is the average misperception of uninformed traders; $\rho_t$ is the price mis-perception of the uninformed traders at time t; $\gamma$ is the level of risk aversion; $\sigma^2_{p,t+1}$ is the variance of the misperception.}
+
+The De-Long noise-trader model also provides a useful benchmark against which to assess the activity of speculators.
+The following four effects are noted by \citet[pp.
+14-15]{Delong1990noise}:  the \emph{Friedman Effect}, is the original arbitrage from the misperception of the noise-traders; \emph{Price Pressure} is the \emph{noise-trader risk} that prices may diverge further before returning to equilibrium; \emph{Hold More} means that noise-traders become more involved in the market as their views become more extreme; \emph{Create Space} is the difference between noise-trader perception and the actual value, it creates the opportunity for informed traders to profit from their information.
+The aim of this paper is to identify the Friedman-effect.
+
 \section{Measuring speculation and the event study}
@@
 There are two parameters that can be used to measure speculation:  the price misperception $(\rho)$ and the weight of noise-traders in the market $(\mu)$.
-If speculation is uninformed and either of these parameters reach an extreme, the model indicates that prices should subsequently reverse.\footnote{See \citet{Hayward2013} for full details of simulations of this model.} Therefore, if misperception or weight of speculation can be measured, this assertion can be tested.
+If speculation is uninformed and either of these parameters reach an extreme, the model indicates that prices should subsequently reverse.\footnote{See \citet{Hayward2013} for full details of a Monte-Carlo simulations of this model when the speculative misperception $(\rho)$ and the weight of speculation $(\mu)$ parameters are adjusted.} Therefore, if misperception or weight of speculation can be measured, this assertion can be tested.
 
@@
 The measurement of speculative intensity must identify how far speculators believe that the market will move in a particular direction.
-In the De-Long model, this would mean that for Equation \ref{eq:NP2}, it is the extreme deviation of $\rho$ from one.
+In the De-Long model, this would mean that for Equation \ref{eq:NP2}, it is the absolute value of $\rho - 1$.
 In the absence of a daily survey of speculators, this study uses a unique series of \emph{risk-reversal skew} as an objective measure of speculative sentiment.\footnote{It appears that this is the only period for which the data currently exist.
@@
 The risk-reversals are the spread of implied volatility for a 1 month 25 delta at-the-money call option relative to the equivalent put.
-Risk-reversal skew is the relative price of currency put and call options for the same strike price.}  The \emph{Put-Call-Parity} in the  \citet{Black1973Options} option pricing model assumes that these will be the same, in practice they are not\footnote{While the Black-Scholes model is the basis for valuing options, where the price of a non-dividend paying stock trades at a discount to the forward rate equal to the risk free rate, the forward rate in a currency option is determined by the relative interest rate differential between the two currencies.
+Risk-reversal skew is the relative price of currency put and call options for the same strike price.} The risk-reversal skew is the relationship between a call option and the equivalent put.
+The \emph{Put-Call-Parity} in the  \citet{Black1973Options} option pricing model assumes that these will be the same, in practice they are not\footnote{While the Black-Scholes model is the basis for valuing options, where the price of a non-dividend paying stock trades at a discount to the forward rate equal to the risk free rate, the forward rate in a currency option is determined by the relative interest rate differential between the two currencies.
 The \citet{GarmanKohlhagan} adjustment that accounts for the difference in interest payments on the two currencies is used.
@@
 }  Therefore, the relationship between implied volatility for calls and implied volatilities for puts for the same delta and maturity is an indication of how much more expensive calls are relative to the equivalent put or as the market bias towards puts over calls and as an indication of market sentiment.
-The larger the absolute level of the risk-reversal skew, the greater the relative price and, if the deviation is based on information, subsequent price action should be zero on average; if the deviation is based on noise, there should be a good chance of a reversal.
+The larger the absolute level of the risk-reversal skew, the greater the relative price and, if the deviation is based on information, subsequent price action should be zero on average; if the deviation is based on noise, this movement should be reversed.
 
@@
 
-\subsection{Analysis of events and abnormal returns}
+\section{Analysis of the results}
 An \emph{event study} is used to assess the effect of extreme speculation on future returns.
@@
 
-\section{Analysis of the results}
-\subsection{Results for extreme speculative sentiment}
 Table \ref{tabref:RR1} records the result of the event studies that were carried out on extreme risk reversal.
@@
 Looking down the seventh column (the 4-day CARW) shows the average cumulative return for the period from four days before the extreme event to the four days after the extreme for extreme sentiment highs and lows of each currency pair.
-The extreme highs are (in this case) those points when the risk-reversal skew is at or above the 95th percentile.
+The extreme highs are (in this case) those points when the risk-reversal skew is at or above the $95^{th}$ percentile.
 For the whole event window (CARW), extreme highs (labelled "Hi") are associated with positive returns and the extreme lows (labelled "Lo") are associated with negative returns.
@@
  % \hline
-& &\multicolumn{6}{c}{5th and 95th percentile} & \multicolumn{1}{c}{} & \multicolumn{6}{c}{1st and 99th percentile}\\
+& &\multicolumn{6}{c}{$5^{th}$ and $95^{th}$ percentile} & \multicolumn{1}{c}{} & \multicolumn{6}{c}{$1^{st}$ and $99^{th}$ percentile}\\
  & &\multicolumn{3}{c}{1 day} & \multicolumn{3}{c}{4 day} & \multicolumn{1}{c}{} & \multicolumn{3}{c}{1 day} & \multicolumn{3}{c}{4 day} \\ 
@@
 \small 
-\item Where Hi means extreme high in the risk-reversal skew and Lo means extreme low in the risk reversal skew; extreme is calculated as being equal or above the 95th or 99th percentile or equal or below the 5th or 1st percentile respectively; CARW is the cumulative abnormal return for the whole window, before and after the extreme event, and CARA is the cumulative abnormal return for the period after the event, which is the event day and the window; abnormal return is anything that is different from zero; the asterisk denotes significantly different from zero, where statistical significance means that more than 95\% of 1000 random bootstraps from the events are above or below zero respectively.
+\item Where Hi means extreme high in the risk-reversal skew and Lo means extreme low in the risk reversal skew; extreme is calculated as being equal or above the $95^{th}$ or $99^{th}$ percentile or equal or below the $5^{th}$ or $1^{st}$ percentile respectively; CARW is the cumulative abnormal return for the whole window, before and after the extreme event, and CARA is the cumulative abnormal return for the period after the event, which is the event day and the window; abnormal return is anything that is different from zero; the asterisk denotes significantly different from zero, where statistical significance means that more than 95\% of 1000 random bootstraps from the events are above or below zero respectively.
 \end{tablenotes}
@@
 
-That sentiment drives price is shows that speculators have influence on the market.
-This is consistent with the evidence of previous research from the likes of \citet{Evans2002Order}, \citet{FuturesSanders} and \citet{FuturesWang} that orders and sentiment drive prices.
+That sentiment drives price is evidence that speculators have influence on the market.
+This is consistent with the evidence of previous research from the likes of \citet{Evans2002Order}, \citet{FuturesSanders} and \citet{FuturesWang}.
 What is more interesting is whether knowledge about the extreme reading at time $t = 0$ provides any information about future returns.
@@
 Looking, for example, at the four-day window in column 8, it is clear that the extreme risk reversal skew does not provide any information about future returns.
-Of the eighteen cases using the 95\% th percentile to identify an extreme, ten show a continuation rather than a reversal that is not significantly different from zero; one shows a significant continuation; five show reversals that are not significant; only two show significant reversal (EURCHF in each case).
-For the more extreme event when the 99\% the percentile is used to measure the extreme, there is even more ambiguity.
+Of the eighteen cases using the $95^{th}$ percentile to identify an extreme, ten show a continuation rather than a reversal that is not significantly different from zero; one shows a significant continuation; five show reversals that are not significant; only two show significant reversal (EURCHF in each case).
+For the more extreme event when the $99^{th}$ percentile is used to measure the extreme, there is even more ambiguity.
 Just over half the cases with the four-day window show a reversal (8 of 18), but only 2 of these are statistically significant.
@@
 \includegraphics[scale=0.8]{RRCum16}
-\caption{Event Study: Extreme (High 99th percentile) Risk Reversal Skew and 16 day event window: Risk reversals are the ratio of implied volatility on calls to the equivalent put; extreme high is the highest 99th percentile for the whole range, cumulative returns are the sum of log exchange rate returns.
+\caption{Event Study: Extreme (High $99{th}$ percentile) Risk Reversal Skew and 16 day event window: Risk reversals are the ratio of implied volatility on calls to the equivalent put; extreme high is at or above the $99^{th}$ percentile for the whole range, cumulative returns are the sum of log exchange rate returns.
 The event day is the day of the extreme risk reversal reading as a proxy for extreme speculative sentiment.
@@
 For the event window, Figures \ref{fig:ES1} and \ref{fig:ES2} show the cumulative abnormal returns for the major exchanges rates relative to the US dollar across a 16 day event window.
-Figure \ref{fig:ES1} shows the extreme highs for as a $99^{th}$ percentile and figure \ref{fig:ES2} shows the extreme low for as a $10^{th}$ percentile.
+Figure \ref{fig:ES1} shows the extreme highs for as a $99^{th}$ percentile and figure \ref{fig:ES2} shows the extreme low for the $10^{th}$ percentile.
 The figures reveal the mean cumulative return (in solid red) and the 95\% confidence intervals (dashed blue).
@@
 \includegraphics[scale=0.8]{RRCum16a}
-\caption{Event Study:  Extreme (Low - 10th percentile) Risk Reversal Skew and 16 day event window: Risk reversals are the ratio of implied volatility on calls to the equivalent put; extreme high is the highest 99th percentile for the whole range, cumulative returns are the sum of log exchange rate returns.
+\caption{Event Study:  Extreme (Low - $10^{th}$ percentile) Risk Reversal Skew and 16 day event window: Risk reversals are the ratio of implied volatility on calls to the equivalent put; extreme low is at or below the $10^{th}$ percentile for the whole range, cumulative returns are the sum of log exchange rate returns.
 The event day is the day of the extreme risk reversal reading as a proxy for extreme speculative sentiment.
@@
 If speculation is informed, we expect to see a distribution of return that is largely symmetrical around zero and very similar to the distribution for the whole sample.
-The distribution of return is very similar but, but much broader than that of the whole distribution.
+The results are presented in Table \ref{tabref:D}.
+They show that the distribution of returns is very similar but, but much broader than that of the whole distribution.
 Conventionally, this would suggest that there is greater risk.
@@
 % Add the table with the descriptive statistics for returns that follow an extreme compared to the whole sample. 
+\begin{table}[h]
+\begin{threeparttable}
+\begin{tabular}{p{3cm} rrllll}
+& \multicolumn{4}{c}{Extreme Distribution}&\multicolumn{2}{c}{Regular Distribution}\\
+& \multicolumn{2}{c}{High}&\multicolumn{2}{c}{Low}&
+\multicolumn{2}{l}{}\\
+ & Mean & Std & Mean & Std & Mean & Std \\
+\hline
+AUD & -0.0174 & -0.0017 &  &  & 0.7241 & 0.0106 \\ 
+  EURGBP & -0.0187 & 0.0010 &  &  & 0.5542 & 0.0090 \\ 
+  EURSEK & 0.0011 & 0.0014 &  &  & 0.4384 & 0.0061 \\ 
+  GBP & 0.0025 & 0.0003 &  &  & 0.4892 & 0.0093 \\ 
+  CHF & 0.0120 & -0.0011 &  &  & 0.6819 & 0.0135 \\ 
+  EURCHF & -0.0039 & 0.0002 &  &  & 0.2769 & 0.0054 \\ 
+  EURJPY & -0.0072 & -0.0016 &  &  & 0.8416 & 0.0154 \\ 
+  EUR & -0.0172 & -0.0009 &  &  & 0.6338 & 0.0094 \\ 
+  CAD & 0.0084 & -0.0006 &  &  & 0.3632 & 0.0052 \\ 
+  JPY & 0.0076 & 0.0009 &  &  & 0.8170 & 0.0139 \\ 
+\caption{Comparison of extreme and regular return distribution}
+\label{tabref:D}
+\end{tabular}
+\begin{tablenotes}
+\small 
+\item This table shows a comparison of the mean return for the whole of the sample compared to the mean return for the day after an extreme has been reached.
+It is clear that there is no significant different between the mean returns, arguing against the idea that speculation is uninformed.
+However, the standard deviation of the returns is much broader once there has been extreme speculation.
+This may suggest increased risk or increased uncertainty.
+\hline
+\end{tablenotes}
+\end{threeparttable}  
+\end{table}
+
 \subsection{Results for extreme speculative positions}
@@
 \small 
-\item Where Hi means extreme high in the sentient index, either S1 which is the measure of net long non-commercial (speculative) positions per number of speculators) or S2 which is the measure of net long non-commercial (speculative) positions per open interest (total outstanding open positions); high and low are above the 95th percentile or below the 5th percentile respectively; CARW is the cumulative abnormal return for the whole window, before and after the extreme event, and CARA is the cumulative abnormal return for the period after the event, which is the event day and the window; abnormal return is anything that is different from zero; the asterisk denotes significantly different from zero, where statistical significance means that more than 95\% of 1000 random bootstraps from the events are above or below zero respectively.
+\item Where Hi means extreme high in the sentient index, either S1 which is the measure of net long non-commercial (speculative) positions per number of speculators) or S2 which is the measure of net long non-commercial (speculative) positions per open interest (total outstanding open positions); high and low are above the $95^{th}$ percentile or below the $5^{th}$ percentile respectively; CARW is the cumulative abnormal return for the whole window, before and after the extreme event, and CARA is the cumulative abnormal return for the period after the event, which is the event day and the window; abnormal return is anything that is different from zero; the asterisk denotes significantly different from zero, where statistical significance means that more than 95\% of 1000 random bootstraps from the events are above or below zero respectively.
 \end{tablenotes}	
@@
 \includegraphics[scale=0.8]{FPCum6w}
-\caption{Event Study:  Extreme (95th percentile), Non-commercial net short per OI (S2), 6 week event window.
-Non-commercial net per OI (S2) are the net speculative positions (as reported to US regulator the CFTC) relative to all open positions; extreme low is the highest 95th percentile for the whole range, cumulative returns are the sum of log exchange rate returns.
+\caption{Event Study:  Extreme low ($5^{th}$ percentile), Non-commercial net short per OI (S2), 6 week event window.
+Non-commercial net per OI (S2) are the net speculative positions (as reported to US regulator the CFTC) relative to all open positions; extreme low is at or below the lowest $5^{th}$ percentile for the whole range, cumulative returns are the sum of log exchange rate returns.
 The event day is the day of the extreme risk reversal reading as a measure of extreme speculative activity.
@@
 \includegraphics[scale=0.8]{FPCum6wa}
-\caption{Event Study:  Extreme (90th percentile), Non-commercial net long (S1), 6 week window.
-Non-commercial net per OI (S2) are the net speculative positions (as reported to US regulator the CFTC) relative to all open positions; extreme high is the highest 95th percentile for the whole range, cumulative returns are the sum of log exchange rate returns.
+\caption{Event Study:  Extreme ($90^{th}$ percentile), Non-commercial net long (S1), 6 week window.
+Non-commercial net per OI (S2) are the net speculative positions (as reported to US regulator the CFTC) relative to all open positions; extreme high is at or above the $90^{th}$ percentile for the whole range, cumulative returns are the sum of log exchange rate returns.
 The event day is the day of the extreme risk reversal reading as a measure of extreme speculative activity.
@@
 There is little evidence that these self-reported speculators are driving prices away from fundamentals with their activity.
-Sixty percent of the S1 cases still show a significant abnormal return in the direction of the extreme 4 weeks after the extreme has been reached; only thirty percent of the cases show a reversal and in only one of those do ninety five percent of the re-sampled means remain above zero.
-For S2 there is even less ambiguity, with ninety percent of the cases showing a continuation and forty percent of them being significant.
+Sixty percent of the S1 cases still show a significant abnormal return in the direction of the extreme 4 weeks after the extreme has been reached; only thirty percent of the cases show a reversal and in only one of those do ninety five percent of the means calculated after re-sampling remain above zero.
+For S2 there is even less ambiguity, with ninety percent of the cases showing a continuation and forty percent of them showing significant continuation.
 The evidence here with speculative positions is the same as that of speculative sentiment.
@@
 Figures \ref{fig:ES3} and \ref{fig:ES4} indicate that these results are not a function of the event window that has been chosen.
-Looking at the 95 percentile confidence intervals (once again constructed by bootstrap re-sampling) indicates that in all the cases, there is a large amount of uncertainty about the future direction well after the extreme reading of speculative activity.
+Looking at the $95^{th}$ percentile confidence intervals (once again constructed by bootstrap re-sampling) indicates that in all the cases, there is a large amount of uncertainty about the future direction well after the extreme reading of speculative activity.
 It does not appear that informed traders are constrained in the short-run from driving prices back to equilibrium.
@@
 The nature speculative activity in financial markets is central to a number of contemporary policy questions.
-Amongst these are the design and regulation of financial markets, the surge in commodity prices and the limits put on the activities of financial institutions.
+Amongst these are the design and regulation of financial markets, the response to the surge in commodity prices that took place between the period 2002 and 2007 and the limits that should be put on the activities of financial institutions.
+This paper provides some evidence to assist in policy making in these areas.
+
 The debate about a tax on financial transactions, initiated by \citet{TobinTax} and now called the \emph{the Tobin tax}, has been intended to reduce speculation.
-This is based on the belief that speculation is uninformed, that it drive prices away from their fundamental value and as a consequence  increases the volatility of financial asset prices.
-The argument for a tax on financial transactions can be presented as a tax on an externality with the level of volatility of financial markets being the public cost.
+This is based on the belief that if speculation is uninformed, that it drive prices away from their fundamental value and, as a consequence, increases the volatility of financial asset prices.
+The argument for a tax on financial transactions can be presented as a tax on an externalities with the level of volatility of financial markets being the public cost.
 
-The increase in volatility that is evident when speculative activity is high seems to provide some support to this argument.
+The increase in volatility that is evident when speculative activity is high that has been identified in this work seems to provide some support to this argument.
 However, it is not clear that the direction of causality is from speculation to volatility.
 An alternative view would suggest that an increase in information raises the amount of informed speculation and that this increase in information naturally increase the volatility of financial assets.
+Causation runs from information to volatility with speculation between.
 If speculation is part of the process of price discovery, reducing its activity may reduce liquidity and hinder the process of price discovery.
 In that case unanticipated costs may offset some of the benefits that are expected from the tax on speculation.
+Volatility may even increase.
 
-Regulatory changes in the US government bond market give some hints at the way that reduce speculation can affect the smooth functioning of markets.\footnote{Increased capital requirements and the so-called \emph{Volcker Rule} that would prevent investment banks from trading on their own account have increased the cost of trading US government bonds and reduced the ability of some key players to participate in the market.}  Though bid-ask spreads have remained rather stable, the \emph{market-depth} and the \emph{price impact} of large trades has deteriorated in the last two years.
-There have also been a number of occasions where very large price changes and an evaporation of liquidity have affected the market.
+Regulatory changes in the US government bond market give some hints at the way that reduce speculation can affect the smooth functioning of markets.
+Increased capital requirements and the so-called \emph{Volcker Rule} that would prevent investment banks from trading on their own account have increased the cost of trading US government bonds and reduced the ability of some key players to participate in the market.
+Though bid-ask spreads have remained stable, the New York Federal Reserve reports that \emph{market-depth} and the \emph{price impact} of large trades has deteriorated in the last two years.
+This is regarded as one of the reason that volatility has increased at times of tension and as liquidty has evaporated and large price changes have been recorded.
 
-The surge in commodities prices that was seen prior to the GFC was seen to be partially the consequence of increased speculation in commodities and the creation of exchange traded funds (ETF) focusing on commodities markets.
-Even if speculators are on both side of the market and not driving prices in one direction, they may still increase the amount of volatility, incurring costs for producers and consumers of underlying materials.
-
-This paper has sought to measure speculation and to use this measure to discover more about the relationship between speculation, information and asset prices.  
-Whether measured as speculative sentiment or as the weight of speculator action, it is clear that extreme speculation does not provide information about future returns.
-This means that speculation is more likely to be informed than uninformed.
-If it were uninformed it would be much more likely that these extremes would be follow by reversals as informed market participants took advantage of the space created by this uninformed speculative activity.
-
+The increased presence of high frequency trading (HFT) has also been seen by some as an unwanted development that has increased the amount of speculation and volatility.
+Again, the results of this study suggest that there should be some caution in removing speculative elements from the market.
 \bibliography{../../myrefs}

### 2015-09-17 R/Extreme2.R

@@ -1,4 +1,4 @@
-# Load required packages
+  # Load required packages
 require(eventstudies) 
 require(PerformanceAnalytics)
 require(xtable)
@@ -88,12 +88,12 @@ str(EUR2lo)
 # Use lapply to apply function (currencydf to the currdf list of zoo objects)
 # Can the high and low each be added to the components of the list and can the 
 # the optional output be selected in the function? 
-eventlist <- lapply(clist, FUN = eventzoo, wind = 2, hilo = "lo")
+eventlist <- lapply(clist, FUN = eventzoo, wind = 2, hilo = "hi")
 #-------------------------------------------------------------
 # create charts for comparison of extreme and regular returns. 
 # this requires the currencydf and currdf (best to take from list)
 twochart <- function(currencydf, currdf){ 
-par(mfrow = c(2,3))
+par(mfrow = c(2,1))
 title1 <- paste("Extreme", colnames(clist[[i]])[1], sep = " ")
 title2 <- paste("Regular", colnames(clist[[i]])[1], sep = " ")
 plot(density(100*eventlist[[i]]$FX.exacm[2,]), xlim = c(-2.0, 2.0), main = title1,
@@ -106,11 +106,12 @@ plot(density(sample(clist[[i]]$r,113), na.rm = TRUE), xlim = c(-2.0, 2.0),
 abline(v = mean(sample(clist[[i]]$r, 113), na.rm = TRUE))
 text(x = mean(sample(clist[[i]]$r, 113), na.rm = TRUE) + 0.2, y = 0.4, labels = "Mean")
 }
-twochart(clist[[3]], eventlist[[3]])
+twochart(clist[["GBP"]], eventlist[["GBP"]])
 #------------------------------------------------
 # This will create the pdf figures for the distribution
 # The graphs are not put on the same page
-pdf("Dist.pdf", paper= "a4r", width = 9, title = "Distribution of returns")
+pdf("Figures/Extreme/Dist.pdf", paper= "a4r", width = 9, 
+    title = "Distribution of returns")
 par(mfcol=c(3,1), oma = c(0,0,1,0))
 for(i in c("EUR", "GBP", "JPY")){
   twochart(clist[[i]], eventlist[[i]])
@@ -119,21 +120,20 @@ dev.off()
 twochart(clist[[3]], eventlist[[3]])
[…]

### 2015-09-17 R/Test.R

@@ -77,3 +77,15 @@ somefunction(x, data = myData)
 A = c(1,2)
 somefunction(A)
 somefunction(A, data = myData)
+
+df <- matrix(NA, ncol = 6, nrow = 10)
+rownames(df) <- currencylist
+for(i in currencylist){
+df[i,1] <- round(mean(clist[[i]]$r, na.rm = TRUE), 4)
+df[i,2] <- round(mean(eventlist[[i]]$FX.exacm[2,], na.rm = TRUE), 4)
+df[i,5] <- round(sd(clist[[i]]$r, na.rm = TRUE), 4)
+df[i,6] <- round(sd(eventlist[[i]]$FX.exacm[2,], na.rm = TRUE), 4)
+}
+df
+dfx <- xtable(df, digits = 4)
+dfx

### 2015-09-17 Knitr/NewPaper.Rnw
Updated Table 2 Code to create the Table 2 in NewPaper.Rnw This creates eventlist.hi and eventlist.lo for the high and low events.

@@
 \begin{threeparttable}
-\begin{tabular}{p{3cm} rrllll}
+\begin{tabular}{rrrrrrr}
 & \multicolumn{4}{c}{Extreme Distribution}&\multicolumn{2}{c}{Regular Distribution}\\
@@
 \hline
-AUD & -0.0174 & -0.0017 &  &  & 0.7241 & 0.0106 \\ 
-  EURGBP & -0.0187 & 0.0010 &  &  & 0.5542 & 0.0090 \\ 
-  EURSEK & 0.0011 & 0.0014 &  &  & 0.4384 & 0.0061 \\ 
-  GBP & 0.0025 & 0.0003 &  &  & 0.4892 & 0.0093 \\ 
-  CHF & 0.0120 & -0.0011 &  &  & 0.6819 & 0.0135 \\ 
-  EURCHF & -0.0039 & 0.0002 &  &  & 0.2769 & 0.0054 \\ 
-  EURJPY & -0.0072 & -0.0016 &  &  & 0.8416 & 0.0154 \\ 
-  EUR & -0.0172 & -0.0009 &  &  & 0.6338 & 0.0094 \\ 
-  CAD & 0.0084 & -0.0006 &  &  & 0.3632 & 0.0052 \\ 
-  JPY & 0.0076 & 0.0009 &  &  & 0.8170 & 0.0139 \\ 
+AUD & 0.0125 & 0.8312 & -0.1711 & 1.0588 & -0.0174 & 0.7241 \\ 
+  EURGBP & 0.0696 & 0.7582 & 0.0996 & 0.9028 & -0.0187 & 0.5542 \\ 
+  EURSEK & 0.0033 & 0.6572 & 0.1373 & 0.6059 & 0.0011 & 0.4384 \\ 
+  GBP & 0.1909 & 0.8861 & 0.0312 & 0.9314 & 0.0025 & 0.4892 \\ 
+  CHF & -0.0565 & 0.9577 & -0.1125 & 1.3452 & 0.0120 & 0.6819 \\ 
+  EURCHF & 0.0690 & 0.4198 & 0.0194 & 0.5445 & -0.0039 & 0.2769 \\ 
+  EURJPY & -0.1103 & 1.2729 & -0.1647 & 1.5424 & -0.0072 & 0.8416 \\ 
+  EUR & 0.1586 & 1.0942 & -0.0933 & 0.9411 & -0.0172 & 0.6338 \\ 
+  CAD & -0.0489 & 0.8006 & -0.0571 & 0.5236 & 0.0084 & 0.3632 \\ 
+  JPY & 0.1341 & 1.0090 & 0.0946 & 1.3921 & 0.0076 & 0.8170 \\ 
+   \hline
 \caption{Comparison of extreme and regular return distribution}
@@
 This may suggest increased risk or increased uncertainty.
-\hline
 \end{tablenotes}
@@
 The nature speculative activity in financial markets is central to a number of contemporary policy questions.
-Amongst these are the design and regulation of financial markets, the response to the surge in commodity prices that took place between the period 2002 and 2007 and the limits that should be put on the activities of financial institutions.
+Amongst these are the design and regulation of financial markets, the response to the surge in commodity prices that took place between 2002 and 2007 and the limits that should be put on the activities of financial institutions.
 This paper provides some evidence to assist in policy making in these areas.
@@
 The debate about a tax on financial transactions, initiated by \citet{TobinTax} and now called the \emph{the Tobin tax}, has been intended to reduce speculation.
-This is based on the belief that if speculation is uninformed, that it drive prices away from their fundamental value and, as a consequence, increases the volatility of financial asset prices.
+This is based on the belief that if speculation is uninformed, that it drives prices away from their fundamental value and, as a consequence, increases the volatility of financial asset prices.
 The argument for a tax on financial transactions can be presented as a tax on an externalities with the level of volatility of financial markets being the public cost.
@@
 
-Regulatory changes in the US government bond market give some hints at the way that reduce speculation can affect the smooth functioning of markets.
-Increased capital requirements and the so-called \emph{Volcker Rule} that would prevent investment banks from trading on their own account have increased the cost of trading US government bonds and reduced the ability of some key players to participate in the market.
+If speculation is more informed than uninformed, reducing the amount of speculation will reduce the ratio of information to noise and reduce liquidity.
+
+Regulatory changes in the US government bond market give some idea about way that a reduction in speculation can affect the smooth functioning of financial markets.
+Increased capital requirements and the so-called \emph{Volcker Rule} that would prevent investment banks from trading on their own account have raised the cost of trading US government bonds and reduced the ability of some key players to participate in the market.
 Though bid-ask spreads have remained stable, the New York Federal Reserve reports that \emph{market-depth} and the \emph{price impact} of large trades has deteriorated in the last two years.
-This is regarded as one of the reason that volatility has increased at times of tension and as liquidty has evaporated and large price changes have been recorded.
-
+\footnote{See New York Fed citation for an assessment of the evolution of bid-ask spreads, price impact and the size of trades.
+}  This is regarded as one of the reason that volatility has increased at times of tension and as liquidty has evaporated and large price changes have been recorded.
+%@misc{NTFedbonds, 
+%  author = {Tobias Adrian and Machael Fleming and Daniel Stackman and Erik Vogt},
+%  title = {Has {US} Treasury Market Liquidity Deteriorated?},
+%  year = {2015},
+%  month = {August},
+%  URL = {http://libertystreeteconomics.newyorkfed.org/2015/08/has-us-treasury-market-liquidity-deteriorated.html},
+%  publisher = {Federal Researve Bank of New York},
+%  journal = {Liberty Street Economics}
+%}
 The increased presence of high frequency trading (HFT) has also been seen by some as an unwanted development that has increased the amount of speculation and volatility.

### 2015-09-17 R/Extreme2.R

@@ -88,7 +88,9 @@ str(EUR2lo)
 # Use lapply to apply function (currencydf to the currdf list of zoo objects)
 # Can the high and low each be added to the components of the list and can the 
 # the optional output be selected in the function? 
-eventlist <- lapply(clist, FUN = eventzoo, wind = 2, hilo = "hi")
+# eventlist.lo is for extreme low and event.list is for extreme high
+eventlist.hi <- lapply(clist, FUN = eventzoo, wind = 2, hilo = "hi")
+eventlist.lo <- lapply(clist, FUN = eventzoo, wind = 2, hilo = "lo")
 #-------------------------------------------------------------
 # create charts for comparison of extreme and regular returns. 
 # this requires the currencydf and currdf (best to take from list)
@@ -106,7 +108,7 @@ plot(density(sample(clist[[i]]$r,113), na.rm = TRUE), xlim = c(-2.0, 2.0),
 abline(v = mean(sample(clist[[i]]$r, 113), na.rm = TRUE))
 text(x = mean(sample(clist[[i]]$r, 113), na.rm = TRUE) + 0.2, y = 0.4, labels = "Mean")
 }
-twochart(clist[["GBP"]], eventlist[["GBP"]])
+twochart(clist[[i]], eventlist[[i]])
 #------------------------------------------------
 # This will create the pdf figures for the distribution
 # The graphs are not put on the same page
@@ -124,10 +126,12 @@ twochart(clist[[3]], eventlist[[3]])
 df <- matrix(NA, ncol = 6, nrow = 10)
 rownames(df) <- currencylist
 for(i in currencylist){
-df[i,1] <- round(mean(clist[[i]]$r, na.rm = TRUE), 4)
-df[i,5] <- round(mean(eventlist[[i]]$FX.exacm[2,], na.rm = TRUE), 4)
-df[i,2] <- round(sd(clist[[i]]$r, na.rm = TRUE), 4)
-df[i,6] <- round(sd(eventlist[[i]]$FX.exacm[2,], na.rm = TRUE), 4)
+df[i, 1] <- round(mean(eventlist.hi[[i]]$FX.exacm[2,], na.rm = TRUE)*100, 4)
+df[i, 2] <- round(sd(eventlist.hi[[i]]$FX.exacm[2,], na.rm = TRUE)*100, 4)
[…]

### 2015-09-18 Knitr/NewPaper.Rnw
Updated paper.  This is now ready for Andros.

@@
 Though bid-ask spreads have remained stable, the New York Federal Reserve reports that \emph{market-depth} and the \emph{price impact} of large trades has deteriorated in the last two years.
-\footnote{See New York Fed citation for an assessment of the evolution of bid-ask spreads, price impact and the size of trades.
-}  This is regarded as one of the reason that volatility has increased at times of tension and as liquidty has evaporated and large price changes have been recorded.
-%@misc{NTFedbonds, 
-%  author = {Tobias Adrian and Machael Fleming and Daniel Stackman and Erik Vogt},
-%  title = {Has {US} Treasury Market Liquidity Deteriorated?},
-%  year = {2015},
-%  month = {August},
-%  URL = {http://libertystreeteconomics.newyorkfed.org/2015/08/has-us-treasury-market-liquidity-deteriorated.html},
-%  publisher = {Federal Researve Bank of New York},
-%  journal = {Liberty Street Economics}
-%}
+\footnote{See \citet{NYFedbonds} for an assessment of the evolution of bid-ask spreads, price impact and the size of trades.
+}  This is regarded as one of the reason that volatility has increased at times of tension and as liquidty has evaporated and large price changes, like the one see on 15 October 2014 have been recorded.
+\footnote{The US 10-year note yield fell nearly 30bp in 30 miunutes, falling from 2.2\% to about 1.9\%.} A report by the NY Fed, the US TReasury and the CFTC about this very large adjustment in US bond prices concluded that a combination of the unwinding of a very large large short position, a decline in order book depth and changes in the provision of liquidity contributed to the action \citet{USbondflash}.
+
 The increased presence of high frequency trading (HFT) has also been seen by some as an unwanted development that has increased the amount of speculation and volatility.

### 2015-10-13 Knitr/NewPaper.Rnw
Corrected paper.  Now need to write conclusion

@@
 Price activity after the extreme is close to a random walk.
-This supports the view that speculation is part of the price-discovery process rather than noise.
+This supports the view that speculation is informed and part of the process of price-discovery rather than uninformed and noise.
 \end{abstract}
@@
 Therefore, Friedman argued, the speculative process would reward those able to access or utilise information relative to the uninformed and, through a Darwinian process, informed speculation would dominate.
-This view regards speculation as an key lubricant of financial markets as it provides liquidity and facilitates the process of price discovery.
+This view regards speculation as an key lubricant of financial markets as it provides liquidity and facilitates the process of price discovery whereby information is assimilated into price.
 
@@
 This would be the cost for a round-trip (buying and selling immediately).}.
-If speculators are uninformed and their activity can be regarded as noise, the most extreme sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when reversals are most likely to take place; if speculators are informed prices should follow a random walk after either of these extremes.
-\begin{figure}
-\graphicspath{{../Figures/}}
-\centering
-\includegraphics[scale=0.8]{RRCum16}
-\caption{Event Study:  Extreme (High $99^{th}$ percentile) Risk Reversal Skew and 16 day event window: Risk reversals are the ratio of implied volatility on calls to the equivalent put; extreme high is at or above the $99^{th}$ percentile for the whole range, cumulative returns are the sum of log exchange rate returns.
-The event day is the day of the extreme risk reversal reading as a proxy for extreme speculative sentiment.
-If speculation is uninformed, extremes should be followed by reversals; if speculation is informed, extremes are more likely to be followed by a random walk.}
-\label{fig:ES1}
-\end{figure}
+If speculators are uninformed and their activity can be regarded as noise, the most extreme speculative sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when reversals are most likely to take place; if speculators are informed prices should follow a random walk after either of these extremes.
 
-This paper contributes to the debate about the nature of speculation in a number of ways.
-It provides clear evidence that speculation is more informed than uninformed.
-A unique set of option prices are used to identify the intensity of speculative sentiment; an event study method is used to establish the effect of extremes of speculative sentiment and speculative positions; these speculative extremes are compared to a common noise-trader model.
-Speculative extremes do not provide information about the future; the evolution of foreign exchange prices after an extreme can be characterised as a random walk.
-However, extremes are followed by an increase in the variance of exchange rate returns.
-This suggests that information from informed speculation contributes to the process of price discovery.
-
-Figure \ref{fig:ES1} shows the evolution of a selection of foreign exchange prices after speculative sentiment is extreme.
-An event study is conducted with the speculative extreme being the event $(t_0)$, sentiment is measured by the price of currency call options relative to the equivalent put.
-The foreign exchange rate has a contemporaneous movement with the sentiment.
-This suggests that either the price is adjusting to new information or that uninformed speculators are driving the price in the direction of their irrational belief.
-However, when sentiment reaches an extreme, there is no clear direction: prices are as likely to go up as they are to go down.
-As such, extreme speculative positions do not presage reversals, indicating that they are not associated with divergence from equilibrium.
-
-%In addition, \citet{Delong1990noise} and \citet{shleifer1997limits} show that constrained horizons can cause noise-trading to limit the ability of informed traders to take advantage of profitable opportunities for fear of further divergence from fundamental value before realignment is achieved.
-
-The nature of speculation is central to many contemporary questions about international financial markets.
-For example, regulations or taxes that reduce the amount of speculation may be beneficial if they limit risk-taking and encourage more long-term attitudes towards investment.
-If speculation is largely uninformed it will add to the level of noise and make it even harder for informed investors to overcome uncertainty and apply fundamental analysis about long-term value; uninformed speculators will increase the tendency towards activity that is focused on the short-term.
-However, if speculation is informed and part of the process of price discovery, regulatory measures to reduce its influence may also reduce liquidity and undermine the process of price discovery.
-In this case, increased volatility, heightened risk and reduced transparency together with other costs may provide a counteraction to any benefits.
+This paper contributes to the debate about the nature of speculation by providing clear evidence that speculation is more informed than uninformed.
+A unique set of option prices are used to identify the intensity of speculative sentiment; an event study method is used to assess the price action around extremes of speculative sentiment and speculative positions; these speculative extremes are compared to a common noise-trader model.
 
@@
 \section{Informed and uninformed speculation}
-Speculation plays an  important role in the areas of market-microstructure, the rationale for the trading of financial securities as well as providing the mechanism for the operation of rational expectations.
+Speculation plays an important part in several areas of financial economics.
 However, there is no clear, unambiguous theory of speculation.
@@
 Speculation and trading can be categorised as being informed or uninformed.
-Those trading without information are also called \emph{noise-traders} or \emph{liquidity traders}.
+Those trading without information are frequently called \emph{noise-traders} or \emph{liquidity traders}.
 However, the terms are not fixed and  \citet[p.
 1]{GortonNoise} categorise noise-traders as all those trading in markets for ``non-information-based reasons''.
-%these two paragraphs above may go later and just a very short coverage of the distinction between noise-traders and liquidity traders. 
+
 The distinction between informed and uninformed traders is a central component of microstructure research.
-The focus of the \citet{Kyle1985Continuous} model is a market-maker receiving orders from the informed and the uninformed.
+The focus of the \citet{Kyle1985Continuous} model of market-making is the interaction of orders from the informed and the uninformed.
 On the assumption that uniformed orders arrive in a random fashion, the weight of orders implies something about the knowledge of the informed and should led to adjustment of the bid-ask spread as part of the process of price discovery.
-This bid-ask spread is the payment that the market-maker receives from the uninformed in return for taking the risk of meeting an informed trader.
 In a similar way \citet{glosten1985bid} describes a model where the bid-ask spread is a function of adverse selection, being caused by the risk that orders come from informed traders.
+
+\citet[p.529]{BlackNoise} suggested that the interaction of informed and uninformed activity was essential to the smooth functioning of financial markets.
+Noise, he argued, was the  ``the arbitrary element in expectations''.
+Therefore, noise makes financial markets possible but imperfect: the more noise-traders that there are, the more liquidity and the easier it is to trade; the more noise-traders, the higher the level of inefficiency in the market and the more likely that price will be driven from fundamental value.
 
@@
 \citet{Grossman1980Impossibility} argue that there is equilibrium between the cost of obtaining information and the benefits that are expected to come from the use of the information.
-
-\citet[p.529]{BlackNoise} suggested that this interaction of informed and uninformed activity was essential to the smooth functioning of financial markets.
-Noise, he argued, was the  ``the arbitrary element in expectations''.
-Therefore, noise makes financial markets possible but imperfect: the more noise-traders that there are, the more liquidity and the easier it is to trade; the more noise-traders, the higher the level of inefficiency in the market and the more likely that price will be driven from fundamental value.
 
@@
 \citet{EasleyPIN2} propose a model to try to measure informed market activity based on the assertion that the bid-ask spread is a function of the level of informed trading.
-Large orders are assumed to be informed as informed traders are most likely to want to swift use of their informational advantage.\footnote{With market-makers using Bayes' Rule to update their fundamental valuation by using order flow to estimate the joint probability that there has been an information event and that it has been good news or bad news, the authors seek to operationalise the following model:
+Large orders are assumed to be informed as informed traders are most likely to want to swift use of their informational advantage.\footnote{Market-makers use Bayes' Rule to update their fundamental valuation by using order flow to estimate the joint probability that there has been an information event and that it has been good news or bad news, the authors seek to operationalise the following model:
 \begin{subequations}
@@
 Those economic agents able to determine the fundamental value or discern the underlying economic model will be able to speculate and drive prices towards equilibrium.
-This tends to leave uninformed speculation as a pure noise that can be characterised as white-noise random variable.\footnote{See \citet{muth1961Rational} for the original adjustment to a cobweb model.} While this construction may be justified for the analysis of microstructure and market efficiency, it is an over-simplification that is inappropriate when seeking more understanding of speculation itself.
+This leaves uninformed speculation as a pure noise that can be characterised as white-noise random variable.\footnote{See \citet{muth1961Rational} for the original adjustment to a cobweb model.} While this construction may be justified for the analysis of microstructure and market efficiency, it is an over-simplification that is inappropriate when seeking more understanding of speculation itself.
 Institutional and psychological explanations for noise-trading frequently suggest that there are endogenous non-random elements to speculative activity.
@@
 Traders without information may not make decisions independently of others.
-Information may be inferred from the buying and selling conducted by other traders.\footnote{This is an example of the beauty contest that has been presented by \citet[p.101]{Keynes1936}.}
+Information may be inferred (perhaps incorrectly) from the buying and selling conducted by other traders, causing positive feedback and trends.\footnote{This is an example of the beauty contest that has been presented by \citet[p.101]{Keynes1936}.}
 
-Liquidity trading and institutional constraints may mean that economic agents just execute customer orders at the current price or that they buy and sell with reference to a benchmark index or to fulfil a hedging strategy.
-The institutional explanations are associated with the allocation of credit for investment and trading purposes:  as asset values decline exchanges, banks and other providers of liquidity require larger margins and deeper haircuts; investors liquidate their fund holdings; the attraction of hedging increases.
+Liquidity trading suggests that institutional constraints may ensure that economic agents just execute customer orders at the current price or that they buy and sell with reference to a benchmark index or to fulfil a hedging strategy.
+The institutional explanations are also associated with the allocation of credit for investment and trading purposes:  as asset values decline exchanges, banks and other providers of liquidity require larger margins and deeper haircuts; investors liquidate their fund holdings; the attraction of hedging increases.
 See \citet{BrunnermeierLiquidity} and \citet{Gorton} for some discussion of the way that margin-calls and collateral constraints can force investment decisions.
@@
 14-15]{Delong1990noise}:  the \emph{Friedman Effect}, is the original arbitrage from the misperception of the noise-traders; \emph{Price Pressure} is the \emph{noise-trader risk} that prices may diverge further before returning to equilibrium; \emph{Hold More} means that noise-traders become more involved in the market as their views become more extreme; \emph{Create Space} is the difference between noise-trader perception and the actual value, it creates the opportunity for informed traders to profit from their information.
-The aim of this paper is to identify the Friedman-effect.
+The aim of this paper is to establish whether the Friedman-effect can be identified.
 
 \section{Measuring speculation and the event study}
-The aim of this study is to assess whether speculation in the foreign exchange market is informed or uninformed.
-Speculation must be identified and then extreme speculation will then be compared with what would be expected if it were informed or uninformed.
-If informed and part of the price-discovery process, extreme speculation should drive the price towards a new equilibrium; the following price performance should be a random walk.
-If speculation is uninformed, extremes should cause a divergence from fundamental value and a price reversal should be anticipated.
+There are two steps: identify speculation and assess what happens when it becomes extreme.
+If speculation is uninformed, extremes should cause a divergence from fundamental value and a price reversal should be anticipated; if informed and part of the price-discovery process, extreme speculation should drive the price towards a new equilibrium and the following price performance should be a random walk.
 
 The \citet[pp.
-14-15]{Delong1990noise} model is summarised by Equation \ref{eq:NP2}.
-There are two parameters that can be used to measure speculation:  the price misperception $(\rho)$ and the weight of noise-traders in the market $(\mu)$.
-If speculation is uninformed and either of these parameters reach an extreme, the model indicates that prices should subsequently reverse.\footnote{See \citet{Hayward2013} for full details of a Monte-Carlo simulations of this model when the speculative misperception $(\rho)$ and the weight of speculation $(\mu)$ parameters are adjusted.} Therefore, if misperception or weight of speculation can be measured, this assertion can be tested.
+14-15]{Delong1990noise} presented a widely-used model to understand the interaction of informed and uninformed traders.
+It is summarised by Equation \ref{eq:NP2}.
+There are two parameters that determine the size of the reversal or Friedman-effect that is to be expected:  the price misperception $(\rho)$ and the weight of noise-traders in the market $(\mu)$.
+If speculation is uninformed and either of these parameters reach an extreme, prices should subsequently reverse.
+The following paragraphs show how these can be measured.
+\footnote{See \citet{Hayward2013} for full details of a Monte-Carlo simulations of this model when the speculative misperception $(\rho)$ and the weight of speculation $(\mu)$ parameters are adjusted.
+As the expected deviation from fundamental value and the weight of speculators in the market becomes more extreme, this reversal becomes more likely.} This reversal is the Friedman-effect discussed above.
 
-\subsection{Measuring the intensity of speculative sentiment}
 The measurement of speculative intensity must identify how far speculators believe that the market will move in a particular direction.
@@
 
-\subsection{Measuring the proportion of speculators}
 Recording speculative activity in an over-the-counter market like foreign exchange is extremely difficult.
@@
 The contracts are for EUR 100,000; GBP 62,500; JPY 12,500,000; CHF 100,000 and CAD 100,000.}
-
-\section{Analysis of the results}
 An \emph{event study} is used to assess the effect of extreme speculation on future returns.
@@
 
+\section{Analysis of results}
   Table \ref{tabref:RR1} records the result of the event studies that were carried out on extreme risk reversal.
-For each exchange rate the second column distinguishes between extreme high and extreme low, the third  column identifies the number of extreme events that were recorded and the next two are the cumulative average returns for the whole window (CARW) and the cumulative average returns for the period of the event and afterwards (CARA).
+For each exchange rate the second column distinguishes between extreme high and extreme low, the third  column identifies the number of extreme events that were recorded and the following two are the cumulative average returns for the whole event window (CARW) and the cumulative average returns for the event and the period after (CARA).
 The windows were chosen to correspond to one day and one week of trading.
-However, Figure \ref{fig:ES1} makes it clear that altering the size of the event window does not change the findings to any significant degree.
-
-As speculators become more extreme in their opinion, exchange rates are driven in the direction of sentiment.
-Looking down the seventh column (the 4-day CARW) shows the average cumulative return for the period from four days before the extreme event to the four days after the extreme for extreme sentiment highs and lows of each currency pair.
-The extreme highs are (in this case) those points when the risk-reversal skew is at or above the $95^{th}$ percentile.
-For the whole event window (CARW), extreme highs (labelled "Hi") are associated with positive returns and the extreme lows (labelled "Lo") are associated with negative returns.
-In most cases these positive and negative returns appear to be statistically significant.
-Significance is identified as those cases where more than 95\% of 1000 random bootstrap samples taken from the events being above zero.
-Mean values of the cumulative returns are calculated from samples of the event days and the distribution of these means are used to construct the confidence intervals.
-Returns that are statistically different from zero are identified by an asterisk.
-
+However, Figure \ref{fig:ES1}, which shows the extreme high for a 16 day event window, makes it clear that altering the size of the event window does not change the findings to any significant degree.
 \begin{sidewaystable}
@@
 
-That sentiment drives price is evidence that speculators have influence on the market.
-This is consistent with the evidence of previous research from the likes of \citet{Evans2002Order}, \citet{FuturesSanders} and \citet{FuturesWang}.
-What is more interesting is whether knowledge about the extreme reading at time $t = 0$ provides any information about future returns.
-If this speculation is noise, informed traders should take advantage of the space that has been created between the price and the fundamental value and a reversal should be seen.
+As speculators become more extreme in their opinion, exchange rates are driven in the direction of sentiment.
+Looking down the seventh column (the 4-day CARW) reveals the average cumulative return for the period from four days before the extreme speculative sentiment to the four days after the event.
+The extreme highs are (in this case) those points when the risk-reversal skew is at or above the $95^{th}$ percentile.
+For the whole event window (CARW), extreme highs (labelled "Hi") are associated with positive returns and the extreme lows (labelled "Lo") are associated with negative returns.
+In most cases these are significantly different from zero.
+Significance is identified as those cases where more than 95\% of 1000 means calculated from random bootstrap samples taken from the events are above zero.
+Returns that are statistically different from zero are identified by an asterisk.
+
+That sentiment drives price appears evident from the covariance of speculative sentiment with exchange rate prices.
+This finding is consistent with the evidence of previous research from the likes of \citet{Evans2002Order}, \citet{FuturesSanders} and \citet{FuturesWang}.
+If speculation is not informed, the extreme is caused by herding or institutional features; if speculation is informed, this suggests that there is information that the equilibrium is very different from the current price.
+
+What is interesting is whether knowledge about the extreme reading at time $t = 0$ provides any information about future returns.
+If this speculation is noise, informed traders, such as real money or long-term fundamental accounts, should take advantage of the space that has been created between the price and the fundamental value and a reversal should be seen.
 Whether this is the case is recorded in columns 5, 8, 11 and 14 of Table \ref{tabref:RR1}.
@@
 Looking, for example, at the four-day window in column 8, it is clear that the extreme risk reversal skew does not provide any information about future returns.
-Of the eighteen cases using the $95^{th}$ percentile to identify an extreme, ten show a continuation rather than a reversal that is not significantly different from zero; one shows a significant continuation; five show reversals that are not significant; only two show significant reversal (EURCHF in each case).
+Of the eighteen cases using the $5^{th}$ and $95^{th}$ percentiles to identify extremely negative and extremely positive sentiment respectively, ten show a continuation rather than a reversal that is not significantly different from zero; one shows a significant continuation; five show reversals that are not significant; only two show significant reversal (EURCHF in each case).
 For the more extreme event when the $99^{th}$ percentile is used to measure the extreme, there is even more ambiguity.
@@
 Exchange rate prices are as likely to go up as down after an extreme and there does not appear to be any information about the future from knowledge of these extremes.
+Extremes seem to be followed by a random walk as would be the case if the extreme were the result of information being absorbed into the price.
 \begin{figure}
@@
 
-Though it appears that the speculation in examples is informed, it is clear from figures \ref{fig:ES1} and \ref{fig:ES2} that exchange rates do not react instantly to the news.
-There is a gradual evolution.
+While there is some evidence that speculation is informed, it is clear from figures \ref{fig:ES1} and \ref{fig:ES2} that exchange rates do not react instantly to news.
+There is a gradual absorption of information and a gradual evolution of prices.
 This is consistent with the evidence of serial correlation in asset prices that has been identified.
 See, for example, \citet{Jagadeesh}.
-It appears that information is gradually absorbed into the price.\footnote{It may be that exchange rate traders take time to adjust to the information that comes from order flow or it may be evidence of the \emph{conservatism heuristic} that suggests that people are slow to change their opinion to new information.}
 
-It is also possible to look at the distribution exchange rate returns after these extremes and to compare them to the distribution of the whole sample.
-If speculation is informed, we expect to see a distribution of return that is largely symmetrical around zero and very similar to the distribution for the whole sample.
-The results are presented in Table \ref{tabref:D}.
-They show that the distribution of returns is very similar but, but much broader than that of the whole distribution.
-Conventionally, this would suggest that there is greater risk.
-It may be that the increase in information adds to the level of uncertainty in the market.
+Two other tests have been carried out to understand more about speculative extremes: the distribution of exchange rate returns after the extreme are compared to the distribution of returns for the whole sample; the level of speculative sentiment is identified when large positive and negative returns are recorded.
+The distribution data is provided in Table \ref{tabref:D}.
+The most striking feature is the broader standard deviation that is recorded.
 % Add the table with the descriptive statistics for returns that follow an extreme compared to the whole sample. 
@@
 
-\subsection{Results for extreme speculative positions}
+\begin{figure}
+\graphicspath{{../Figures/}}
+\centering
+\includegraphics[scale=0.8]{RRCum16}
+\caption{Event Study:  Extreme (High $99^{th}$ percentile) Risk Reversal Skew and 16 day event window: Risk reversals are the ratio of implied volatility on calls to the equivalent put; extreme high is at or above the $99^{th}$ percentile for the whole range, cumulative returns are the sum of log exchange rate returns.
+The event day is the day of the extreme risk reversal reading as a proxy for extreme speculative sentiment.
+If speculation is uninformed, extremes should be followed by reversals; if speculation is informed, extremes are more likely to be followed by a random walk.}
+\label{fig:ES1}
+\end{figure}
+Figure \ref{fig:ES1} shows the evolution of a selection of foreign exchange prices after speculative sentiment is extreme.
+An event study is conducted with the speculative extreme being the event $(t_0)$, sentiment is measured by the price of currency call options relative to the equivalent put.
+The foreign exchange rate has a contemporaneous movement with the sentiment.
+This suggests that either the price is adjusting to new information or that uninformed speculators are driving the price in the direction of their irrational belief.
+However, when sentiment reaches an extreme, there is no clear direction: prices are as likely to go up as they are to go down.
+As such, extreme speculative positions do not presage reversals, indicating that they are not associated with divergence from equilibrium.
 Table \ref{tabref:SP1} shows the results from the event studies that were carried out using speculative positions as reported to US regulators.
@@
 \small 
-\item Where Hi means extreme high in the sentient index, either S1 which is the measure of net long non-commercial (speculative) positions per number of speculators) or S2 which is the measure of net long non-commercial (speculative) positions per open interest (total outstanding open positions); high and low are above the $95^{th}$ percentile or below the $5^{th}$ percentile respectively; CARW is the cumulative abnormal return for the whole window, before and after the extreme event, and CARA is the cumulative abnormal return for the period after the event, which is the event day and the window; abnormal return is anything that is different from zero; the asterisk denotes significantly different from zero, where statistical significance means that more than 95\% of 1000 random bootstraps from the events are above or below zero respectively.
+\item Where Hi means extreme high in the sentient index, either S1 which is the measure of net long non-commercial (speculative) positions per total of speculators) or S2 which is the measure of net long non-commercial (speculative) positions per open interest (total outstanding open positions); high and low are above the $95^{th}$ percentile or below the $5^{th}$ percentile respectively; CARW is the cumulative abnormal return for the whole window, before and after the extreme event, and CARA is the cumulative abnormal return for the period after the event, which is the event day and the window; abnormal return is anything that is different from zero; the asterisk denotes significantly different from zero, where statistical significance is that more than 95\% of 1000 means calculated from random bootstraps from the extreme readings are above or below zero respectively.
 \end{tablenotes}	
@@
 Non-commercial net per OI (S2) are the net speculative positions (as reported to US regulator the CFTC) relative to all open positions; extreme low is at or below the lowest $5^{th}$ percentile for the whole range, cumulative returns are the sum of log exchange rate returns.
-The event day is the day of the extreme risk reversal reading as a measure of extreme speculative activity.
+The event day is the day of the extreme reading as a measure of extreme speculative activity.
 If speculation is uninformed, extremes should be followed by reversals; if speculation is informed, extremes are more likely to be followed by a random walk.}
@@
 Non-commercial net per OI (S2) are the net speculative positions (as reported to US regulator the CFTC) relative to all open positions; extreme high is at or above the $90^{th}$ percentile for the whole range, cumulative returns are the sum of log exchange rate returns.
-The event day is the day of the extreme risk reversal reading as a measure of extreme speculative activity.
+The event day is the day of the extreme reading as a measure of extreme speculative activity.
 If speculation is uninformed, extremes should be followed by reversals; if speculation is informed, extremes are more likely to be followed by a random walk.}
@@
 The price action of the event study is consistent with speculation being informed rather than uninformed.
-
 Figures \ref{fig:ES3} and \ref{fig:ES4} indicate that these results are not a function of the event window that has been chosen.
-Looking at the $95^{th}$ percentile confidence intervals (once again constructed by bootstrap re-sampling) indicates that in all the cases, there is a large amount of uncertainty about the future direction well after the extreme reading of speculative activity.
-It does not appear that informed traders are constrained in the short-run from driving prices back to equilibrium.
-It appears that the informed traders are driving prices to a new level.
 
 \section{Speculation and foreign exchange returns}
+The nature of speculation is central to many contemporary questions about international financial markets.
+For example, regulations or taxes that reduce the amount of speculation may be beneficial if they limit risk-taking and encourage more long-term attitudes towards investment.
+If speculation is largely uninformed it will add to the level of noise and make it even harder for informed investors to overcome uncertainty and apply fundamental analysis about long-term value; uninformed speculators will increase the tendency towards activity that is focused on the short-term.
+However, if speculation is informed and part of the process of price discovery, regulatory measures to reduce its influence may also reduce liquidity and undermine the process of price discovery.
+In this case, increased volatility, heightened risk and reduced transparency together with other costs may provide a counteraction to any benefits.
+
  The nature speculative activity in financial markets is central to a number of contemporary policy questions.
@@
 \footnote{See \citet{NYFedbonds} for an assessment of the evolution of bid-ask spreads, price impact and the size of trades.
-}  This is regarded as one of the reason that volatility has increased at times of tension and as liquidty has evaporated and large price changes, like the one see on 15 October 2014 have been recorded.
-\footnote{The US 10-year note yield fell nearly 30bp in 30 miunutes, falling from 2.2\% to about 1.9\%.} A report by the NY Fed, the US TReasury and the CFTC about this very large adjustment in US bond prices concluded that a combination of the unwinding of a very large large short position, a decline in order book depth and changes in the provision of liquidity contributed to the action \citet{USbondflash}.
+}  This is regarded as one of the reason that volatility has increased at times of tension and as liquidity has evaporated and large price changes, like the one see on 15 October 2014 have been recorded.
+\footnote{The US 10-year note yield fell nearly 30bp in 30 minutes, falling from 2.2\% to about 1.9\%.} A report by the NY Fed, the US Treasury and the CFTC about this very large adjustment in US bond prices concluded that a combination of the unwinding of a very large large short position, a decline in order book depth and changes in the provision of liquidity contributed to the action \citet{USbondflash}.
 
@@
 Again, the results of this study suggest that there should be some caution in removing speculative elements from the market.
+
+%this may be added.
+It appears that information is gradually absorbed into the price.\footnote{It may be that exchange rate traders take time to adjust to the information that comes from order flow or it may be evidence of the \emph{conservatism heuristic} that suggests that people are slow to change their opinion to new information.}
+
 \bibliography{../../myrefs}

### 2015-10-13 R/Extreme2.R

@@ -94,24 +94,25 @@ eventlist.lo <- lapply(clist, FUN = eventzoo, wind = 2, hilo = "lo")
 #-------------------------------------------------------------
 # create charts for comparison of extreme and regular returns. 
 # this requires the currencydf and currdf (best to take from list)
-twochart <- function(currencydf, currdf){ 
+twochart <- function(clist, eventlist){ 
 par(mfrow = c(2,1))
 title1 <- paste("Extreme", colnames(clist[[i]])[1], sep = " ")
 title2 <- paste("Regular", colnames(clist[[i]])[1], sep = " ")
-plot(density(100*eventlist[[i]]$FX.exacm[2,]), xlim = c(-2.0, 2.0), main = title1,
-     lwd = 2, col = "red")
+plot(density(100*eventlist[[i]]$FX.exacm[2,]), main = title1,
+     lwd = 2, col = "red", xlab = "Return")
 abline(v = mean(eventlist[[i]]$FX.exacm[2,], na.rm = TRUE))
 text(x = mean(eventlist[[i]]$FX.exacm[2,], na.rm = TRUE) + 0.2, y = 0.2, labels = "Mean")
 # need to sample 
 plot(density(sample(clist[[i]]$r,113), na.rm = TRUE), xlim = c(-2.0, 2.0), 
-     main = title2, lwd = 2, col = "red")
+     main = title2, lwd = 2, col = "red", xlab = "Return")
 abline(v = mean(sample(clist[[i]]$r, 113), na.rm = TRUE))
 text(x = mean(sample(clist[[i]]$r, 113), na.rm = TRUE) + 0.2, y = 0.4, labels = "Mean")
 }
-twochart(clist[[i]], eventlist[[i]])
+twochart(clist, eventlist.hi)
 #------------------------------------------------
 # This will create the pdf figures for the distribution
 # The graphs are not put on the same page
+# NOT WORKING
 pdf("Figures/Extreme/Dist.pdf", paper= "a4r", width = 9, 
     title = "Distribution of returns")
 par(mfcol=c(3,1), oma = c(0,0,1,0))
@@ -136,6 +137,9 @@ df[i, 6] <- round(sd(clist[[i]]$r, na.rm = TRUE), 4)
 df
 dfx <- xtable(df, digits = 4)
 dfx
+a <- boot(as.matrix(eventlist.hi[[i]]$FX.exacm[2,]), sd, R = 100, 
+          na.rm = TRUE)
+a
 
[…]

### 2015-10-20 Knitr/ElsPaper.tex
Created the ElsPaper.tex Need to check the key words and the way that the author is presented.

--- 2015-10-13 Knitr/NewPaper.Rnw
@@
-\documentclass[12pt, a4paper, oneside]{article} % Paper size, default font size and one-sided paper
+\documentclass[preprint,12pt,authoryear]{elsarticle}
 %\graphicspath{{./Figures/}} % Specifies the directory where pictures are stored
 %\usepackage[dcucite]{harvard}
+% \documentclass[final,1p,times,authoryear]{elsarticle}
+%\documentclass[final,1p,times,twocolumn,authoryear]{elsarticle}
+%\documentclass[final,3p,times,authoryear]{elsarticle}
+%\documentclass[final,3p,times,twocolumn,authoryear]{elsarticle}
+%\documentclass[final,5p,times,authoryear]{elsarticle}
+%\documentclass[final,5p,times,twocolumn,authoryear]{elsarticle}
 \usepackage{rotating}
@@
 \usepackage{multirow}
-\usepackage[comma, sort&compress]{natbib}% Use the natbib reference package - read up on this to edit the reference style; if you want text (e.g.
+%\usepackage[comma, sort&compress]{natbib}% Use the natbib reference package - read up on this to edit the reference style; if you want text (e.g.
 Smith et al., 2012) for the in-text references (instead of numbers), remove 'numbers'
@@
 \begin{document}
-\title{Speculation in foreign exchange: noise or information}
-\author{Rob Hayward\footnote{University of Brighton Business School, Lewes Road, Brighton, BN2 4AT; Telephone 01273 642586.  rh49@brighton.ac.uk. }} 
-\date{\today}
-\maketitle
+\begin{frontmatter}
+\title{Speculation in foreign exchange: noise or information?}
+\author{Rob Hayward\footnote{University of Brighton Business School, Lewes Road, Brighton, BN2 4AT; Telephone 01273 642586.  rh49@brighton.ac.uk.}}
+
 \begin{abstract}
-If prices are driven by speculation based on noise rather than information, extreme levels of speculative sentiment and speculative activity should lead to price reversals.
+If foreign exchange prices are driven by speculation based on noise rather than information, extreme levels of speculative sentiment and speculative activity should lead to price reversals.
 Informed speculation is sustainable as it is part of the price discovery process.
@@
 Price activity after the extreme is close to a random walk.
-This supports the view that speculation is informed and part of the process of price-discovery rather than uninformed and noise.
+This supports the view that speculation is informed and part of the process of price-discovery rather than uninformed noise.
 \end{abstract}
+\begin{keyword}
+JEL classification: G12, G14, G15; information, noise-trading, event studies.
+\end{keyword}
+
+\end{frontmatter}
 \section*{Introduction}
@@
 The cost of trading in this market, when measured in terms of the size of the bid-ask spread, is extremely low.
-An analysis by \citet{Steely2013} of the bid-ask spreads for USD-JPY, GBP-USD and EUR-USD for the period January 2001 to 2005 and finds that the average spread for each is 0.342\%, 0.305\% and 0.326\% respectively.
+An analysis by \citet{Steely2013} of the bid-ask spreads for USD-JPY, GBP-USD and EUR-USD for the period January 2001 to 2005 finds that the average spread for each is 0.342\%, 0.305\% and 0.326\% respectively.
 This would be the cost for a round-trip (buying and selling immediately).}.
-If speculators are uninformed and their activity can be regarded as noise, the most extreme speculative sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when reversals are most likely to take place; if speculators are informed prices should follow a random walk after either of these extremes.
+If speculators are uninformed and their activity can be regarded as noise, the most extreme speculative sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when reversals are most likely to take place; if speculators are informed prices should follow a random walk after extremes.
 
@@
 
-The rest of the paper proceeds as follows:  Section Two discusses the distinction between informed and uninformed speculators; Section Three talks about the measurement of speculation and the event study method; Section Four reviews the evidence; Section Five concludes with a discussion of policy implications.
+The rest of the paper proceeds as follows:  Section Two discusses the distinction between informed and uninformed speculators; Section Three talks about the measurement of speculation and the event study method; Section Four reviews the evidence; Section Five concludes.
 
@@
 1]{GortonNoise} categorise noise-traders as all those trading in markets for ``non-information-based reasons''.
+See \citet{Ramiah201589} for an overview of the noise-trader literature.
 
@@
 Traders without information may not make decisions independently of others.
-Information may be inferred (perhaps incorrectly) from the buying and selling conducted by other traders, causing positive feedback and trends.\footnote{This is an example of the beauty contest that has been presented by \citet[p.101]{Keynes1936}.}
+Information may be inferred (perhaps incorrectly) from the buying and selling conducted by other traders, causing positive feedback and trends.\footnote{This is an example of the beauty contest that has been presented by \citet[p.101]{Keynes1936}.
+The aim is to determine not the underlying value but what others perceive to be the underlying value.}
 
@@
 However, a large proportion of uninformed traders in a market can undermine this process.
-\citet{Delong1990noise} and \citet{shleifer1997limits} have presented models where the risk that noise-traders will drive prices even further away from equilibrium before convergence can limit the activity of informed speculators when they are risk averse.
+\citet{Delong1990noise} and \citet{shleifer1997limits} have presented models where the risk that noise-traders will move prices even further away from equilibrium before convergence can limit the activity of informed speculators when they are risk averse.
 This can be called \emph{noise-trader risk}.\footnote{This three-period, overlapping generation model with informed and uninformed traders has a price equation \begin{equation}
@@
 There are two steps: identify speculation and assess what happens when it becomes extreme.
-If speculation is uninformed, extremes should cause a divergence from fundamental value and a price reversal should be anticipated; if informed and part of the price-discovery process, extreme speculation should drive the price towards a new equilibrium and the following price performance should be a random walk.
+If speculation is uninformed, extremes should cause a divergence from fundamental value and a price reversal should be anticipated; if informed and part of the price-discovery process, extreme speculation achieves a new equilibrium and the following price performance should be a random walk.
 
@@
 If speculation is uninformed and either of these parameters reach an extreme, prices should subsequently reverse.
+This reversal is the Friedman-effect discussed above.
 The following paragraphs show how these can be measured.
 \footnote{See \citet{Hayward2013} for full details of a Monte-Carlo simulations of this model when the speculative misperception $(\rho)$ and the weight of speculation $(\mu)$ parameters are adjusted.
-As the expected deviation from fundamental value and the weight of speculators in the market becomes more extreme, this reversal becomes more likely.} This reversal is the Friedman-effect discussed above.
+As the expected deviation from fundamental value and the weight of speculators in the market becomes more extreme, this reversal becomes more likely.}
 
@@
 In the De-Long model, this would mean that for Equation \ref{eq:NP2}, it is the absolute value of $\rho - 1$.
-In the absence of a daily survey of speculators, this study uses a unique series of \emph{risk-reversal skew} as an objective measure of speculative sentiment.\footnote{It appears that this is the only period for which the data currently exist.
+In the absence of a daily survey of speculators, this study uses a unique series of option\emph{risk-reversal skew} as an objective measure of speculative sentiment.\footnote{It appears that this is the only period for which the data currently exist.
 The risk-reversal data is very difficult to obtain in an over-the-counter market like foreign exchange where there is no centralised exchange to record and collect information, generally being the property of quoting institutions.
@@
 This model allows the market practice of pricing options from the volatility implied by the model.
-In other words, the Garman-Kohlhagan is used to extract the volatility that is implied by a specific option price given the strike and other parameters.}  A call option is the right to buy and a put option is the right to sell.
+In other words, the Garman-Kohlhagan is used to extract the volatility that is implied by a specific option price given the strike and other parameters.}  %A call option is the right to buy and a put option is the right to sell.
 The delta of an option measures the effect of a change in the spot price on the value of the option.
@@
 The contracts are for EUR 100,000; GBP 62,500; JPY 12,500,000; CHF 100,000 and CAD 100,000.}
+
 An \emph{event study} is used to assess the effect of extreme speculation on future returns.
@@
 The windows were chosen to correspond to one day and one week of trading.
-However, Figure \ref{fig:ES1}, which shows the extreme high for a 16 day event window, makes it clear that altering the size of the event window does not change the findings to any significant degree.
+However, Figure \ref{fig:ES1}, which shows the extreme high for a 16 day event window, and it becomes clear that altering the size of the event window does not change the findings to any significant degree.
 \begin{sidewaystable}
@@
 \small 
-\item Where Hi means extreme high in the risk-reversal skew and Lo means extreme low in the risk reversal skew; extreme is calculated as being equal or above the $95^{th}$ or $99^{th}$ percentile or equal or below the $5^{th}$ or $1^{st}$ percentile respectively; CARW is the cumulative abnormal return for the whole window, before and after the extreme event, and CARA is the cumulative abnormal return for the period after the event, which is the event day and the window; abnormal return is anything that is different from zero; the asterisk denotes significantly different from zero, where statistical significance means that more than 95\% of 1000 random bootstraps from the events are above or below zero respectively.
+\item Where Hi means extreme high in the risk-reversal skew and Lo means extreme low in the risk reversal skew; extreme is calculated as being equal or above the $95^{th}$ or $99^{th}$ percentile or equal or below the $5^{th}$ or $1^{st}$ percentile respectively; CARW is the cumulative abnormal return for the whole window, before and after the extreme event, and CARA is the cumulative abnormal return for the period after the event, which is the event day and the window; abnormal return is anything that is different from zero; the asterisk denotes significantly different from zero, where statistical significance is more than 95\% of 1000 means calculated from random bootstrap samples from the events are above or below zero respectively.
 \end{tablenotes}
@@
 That sentiment drives price appears evident from the covariance of speculative sentiment with exchange rate prices.
-This finding is consistent with the evidence of previous research from the likes of \citet{Evans2002Order}, \citet{FuturesSanders} and \citet{FuturesWang}.
-If speculation is not informed, the extreme is caused by herding or institutional features; if speculation is informed, this suggests that there is information that the equilibrium is very different from the current price.
+This finding is consistent with the evidence of previous research that has used order-flow or sentiment from the likes of \citet{Evans2002Order}, \citet{FuturesSanders} and \citet{FuturesWang}.
+If speculation is not informed, the extreme is caused by herding or institutional features; if speculation is informed, it is facilitating the absorption of this information into the price.
 
-What is interesting is whether knowledge about the extreme reading at time $t = 0$ provides any information about future returns.
+Does knowledge about the extreme at event time $t = 0$ provide information about the future returns?
 If this speculation is noise, informed traders, such as real money or long-term fundamental accounts, should take advantage of the space that has been created between the price and the fundamental value and a reversal should be seen.
@@
 Looking, for example, at the four-day window in column 8, it is clear that the extreme risk reversal skew does not provide any information about future returns.
-Of the eighteen cases using the $5^{th}$ and $95^{th}$ percentiles to identify extremely negative and extremely positive sentiment respectively, ten show a continuation rather than a reversal that is not significantly different from zero; one shows a significant continuation; five show reversals that are not significant; only two show significant reversal (EURCHF in each case).
-For the more extreme event when the $99^{th}$ percentile is used to measure the extreme, there is even more ambiguity.
+Of the eighteen cases using the $5^{th}$ and $95^{th}$ percentiles to identify extremely negative and extremely positive sentiment respectively, ten show a continuation (rather than a reversal) that is not significantly different from zero; one shows a significant continuation; five show reversals that are not significant; only two show significant reversal - EURCHF in each case.
+For the more extreme event when the $99^{th}$ percentile is used, there is even less clarity.
 Just over half the cases with the four-day window show a reversal (8 of 18), but only 2 of these are statistically significant.
@@
 \caption{Event Study: Extreme (High $99{th}$ percentile) Risk Reversal Skew and 16 day event window: Risk reversals are the ratio of implied volatility on calls to the equivalent put; extreme high is at or above the $99^{th}$ percentile for the whole range, cumulative returns are the sum of log exchange rate returns.
-The event day is the day of the extreme risk reversal reading as a proxy for extreme speculative sentiment.
-If speculation is uninformed, extremes should be followed by reversals; if speculation is informed, extremes are more likely to be followed by a random walk.}
+The event day is the day of the extreme risk reversal reading.
+If speculation is uninformed, extremes should be followed by reversals; if speculation is informed, extremes are more likely to be followed by a random walk.
+The solid line is the cumulative mean return from the start of the event window; the dashed lines show the 95\% confidence intervals constructed from 1000 random bootstrap samples of appropriate event window day.}
 \label{fig:ES5}
@@
 \includegraphics[scale=0.8]{RRCum16a}
-\caption{Event Study:  Extreme (Low - $10^{th}$ percentile) Risk Reversal Skew and 16 day event window: Risk reversals are the ratio of implied volatility on calls to the equivalent put; extreme low is at or below the $10^{th}$ percentile for the whole range, cumulative returns are the sum of log exchange rate returns.
-The event day is the day of the extreme risk reversal reading as a proxy for extreme speculative sentiment.
-If speculation is uninformed, extremes should be followed by reversals; if speculation is informed, extremes are more likely to be followed by a random walk.}
+\caption{Event Study: Extreme (Low $10{th}$ percentile) Risk Reversal Skew and 16 day event window: Risk reversals are the ratio of implied volatility on calls to the equivalent put; extreme low is at or below the $10^{th}$ percentile for the whole range, cumulative returns are the sum of log exchange rate returns.
+The event day is the day of the extreme risk reversal reading.
+If speculation is uninformed, extremes should be followed by reversals; if speculation is informed, extremes are more likely to be followed by a random walk.
+The solid line is the cumulative mean return from the start of the event window; the dashed lines show the 95\% confidence intervals constructed from 1000 random bootstrap samples of appropriate event window day.}
 \label{fig:ES2}
@@
 There is a gradual absorption of information and a gradual evolution of prices.
-This is consistent with the evidence of serial correlation in asset prices that has been identified.
-See, for example, \citet{Jagadeesh}.
-
-Two other tests have been carried out to understand more about speculative extremes: the distribution of exchange rate returns after the extreme are compared to the distribution of returns for the whole sample; the level of speculative sentiment is identified when large positive and negative returns are recorded.
-The distribution data is provided in Table \ref{tabref:D}.
-The most striking feature is the broader standard deviation that is recorded.
-% Add the table with the descriptive statistics for returns that follow an extreme compared to the whole sample. 
-\begin{table}[h]
-\begin{threeparttable}
-\begin{tabular}{rrrrrrr}
-& \multicolumn{4}{c}{Extreme Distribution}&\multicolumn{2}{c}{Regular Distribution}\\
-& \multicolumn{2}{c}{High}&\multicolumn{2}{c}{Low}&
-\multicolumn{2}{l}{}\\
- & Mean & Std & Mean & Std & Mean & Std \\
-\hline
-AUD & 0.0125 & 0.8312 & -0.1711 & 1.0588 & -0.0174 & 0.7241 \\ 
-  EURGBP & 0.0696 & 0.7582 & 0.0996 & 0.9028 & -0.0187 & 0.5542 \\ 
-  EURSEK & 0.0033 & 0.6572 & 0.1373 & 0.6059 & 0.0011 & 0.4384 \\ 
-  GBP & 0.1909 & 0.8861 & 0.0312 & 0.9314 & 0.0025 & 0.4892 \\ 
-  CHF & -0.0565 & 0.9577 & -0.1125 & 1.3452 & 0.0120 & 0.6819 \\ 
-  EURCHF & 0.0690 & 0.4198 & 0.0194 & 0.5445 & -0.0039 & 0.2769 \\ 
-  EURJPY & -0.1103 & 1.2729 & -0.1647 & 1.5424 & -0.0072 & 0.8416 \\ 
-  EUR & 0.1586 & 1.0942 & -0.0933 & 0.9411 & -0.0172 & 0.6338 \\ 
-  CAD & -0.0489 & 0.8006 & -0.0571 & 0.5236 & 0.0084 & 0.3632 \\ 
-  JPY & 0.1341 & 1.0090 & 0.0946 & 1.3921 & 0.0076 & 0.8170 \\ 
-   \hline
-\caption{Comparison of extreme and regular return distribution}
-\label{tabref:D}
-\end{tabular}
-\begin{tablenotes}
-\small 
-\item This table shows a comparison of the mean return for the whole of the sample compared to the mean return for the day after an extreme has been reached.
-It is clear that there is no significant different between the mean returns, arguing against the idea that speculation is uninformed.
-However, the standard deviation of the returns is much broader once there has been extreme speculation.
-This may suggest increased risk or increased uncertainty.
-\end{tablenotes}
-\end{threeparttable}  
-\end{table}
+Serial correlation in asset prices is a consistent observation.
+See \citet{Jagadeesh}.
+In this case the serial correlation is associated with extreme speculation that appears to be informed.
 
@@
 \includegraphics[scale=0.8]{RRCum16}
-\caption{Event Study:  Extreme (High $99^{th}$ percentile) Risk Reversal Skew and 16 day event window: Risk reversals are the ratio of implied volatility on calls to the equivalent put; extreme high is at or above the $99^{th}$ percentile for the whole range, cumulative returns are the sum of log exchange rate returns.
-The event day is the day of the extreme risk reversal reading as a proxy for extreme speculative sentiment.
-If speculation is uninformed, extremes should be followed by reversals; if speculation is informed, extremes are more likely to be followed by a random walk.}
+\caption{Event Study: Extreme (High $99{th}$ percentile) Risk Reversal Skew and 16 day event window: Risk reversals are the ratio of implied volatility on calls to the equivalent put; extreme high is at or above the $99^{th}$ percentile for the whole range, cumulative returns are the sum of log exchange rate returns.
+The event day is the day of the extreme risk reversal reading.
+If speculation is uninformed, extremes should be followed by reversals; if speculation is informed, extremes are more likely to be followed by a random walk.
+The solid line is the cumulative mean return from the start of the event window; the dashed lines show the 95\% confidence intervals constructed from 1000 random bootstrap samples of appropriate event window day.}
 \label{fig:ES1}
 \end{figure}
-Figure \ref{fig:ES1} shows the evolution of a selection of foreign exchange prices after speculative sentiment is extreme.
-An event study is conducted with the speculative extreme being the event $(t_0)$, sentiment is measured by the price of currency call options relative to the equivalent put.
-The foreign exchange rate has a contemporaneous movement with the sentiment.
-This suggests that either the price is adjusting to new information or that uninformed speculators are driving the price in the direction of their irrational belief.
-However, when sentiment reaches an extreme, there is no clear direction: prices are as likely to go up as they are to go down.
-As such, extreme speculative positions do not presage reversals, indicating that they are not associated with divergence from equilibrium.
+
 Table \ref{tabref:SP1} shows the results from the event studies that were carried out using speculative positions as reported to US regulators.
@@
 \small 
-\item Where Hi means extreme high in the sentient index, either S1 which is the measure of net long non-commercial (speculative) positions per total of speculators) or S2 which is the measure of net long non-commercial (speculative) positions per open interest (total outstanding open positions); high and low are above the $95^{th}$ percentile or below the $5^{th}$ percentile respectively; CARW is the cumulative abnormal return for the whole window, before and after the extreme event, and CARA is the cumulative abnormal return for the period after the event, which is the event day and the window; abnormal return is anything that is different from zero; the asterisk denotes significantly different from zero, where statistical significance is that more than 95\% of 1000 means calculated from random bootstraps from the extreme readings are above or below zero respectively.
+\item Where Hi means extreme high, either S1 which is the measure of net long non-commercial (speculative) positions per total of speculators) or S2 which is the measure of net long non-commercial (speculative) positions per open interest (total outstanding open positions); high and low are above the $95^{th}$ percentile or below the $5^{th}$ percentile respectively; CARW is the cumulative abnormal return for the whole window, before and after the extreme event, and CARA is the cumulative abnormal return for the period after the event, which is the event day and the window; abnormal return is anything that is different from zero; the asterisk denotes significantly different from zero, where statistical significance is that more than 95\% of 1000 means calculated from random bootstraps from the extreme readings are above or below zero.
 \end{tablenotes}	
@@
 There is evidence here of momentum and of prices following the movement of speculator positions.
-This consistent with both the gradual absorption of information into the price and with the feedback of price into additional speculation.
 
@@
 \includegraphics[scale=0.8]{FPCum6w}
-\caption{Event Study:  Extreme low ($5^{th}$ percentile), Non-commercial net short per OI (S2), 6 week event window.
-Non-commercial net per OI (S2) are the net speculative positions (as reported to US regulator the CFTC) relative to all open positions; extreme low is at or below the lowest $5^{th}$ percentile for the whole range, cumulative returns are the sum of log exchange rate returns.
-The event day is the day of the extreme reading as a measure of extreme speculative activity.
-If speculation is uninformed, extremes should be followed by reversals; if speculation is informed, extremes are more likely to be followed by a random walk.}
+\caption{Event Study: Extreme (Low $5{th}$ percentile) Risk Reversal Skew and 16 day event window: Risk reversals are the ratio of implied volatility on calls to the equivalent put; extreme low is at or above the $5^{th}$ percentile for the whole range, cumulative returns are the sum of log exchange rate returns.
+The event day is the day of the extreme risk reversal reading.
+If speculation is uninformed, extremes should be followed by reversals; if speculation is informed, extremes are more likely to be followed by a random walk.
+The solid line is the cumulative mean return from the start of the event window; the dashed lines show the 95\% confidence intervals constructed from 1000 random bootstrap samples of appropriate event window day.}
 \label{fig:ES3}
@@
 Sixty percent of the S1 cases still show a significant abnormal return in the direction of the extreme 4 weeks after the extreme has been reached; only thirty percent of the cases show a reversal and in only one of those do ninety five percent of the means calculated after re-sampling remain above zero.
-For S2 there is even less ambiguity, with ninety percent of the cases showing a continuation and forty percent of them showing significant continuation.
+For S2 there is even less clarity, with ninety percent of the cases showing a continuation and forty percent of them showing significant continuation.
 The evidence here with speculative positions is the same as that of speculative sentiment.
@@
 The price action of the event study is consistent with speculation being informed rather than uninformed.
-Figures \ref{fig:ES3} and \ref{fig:ES4} indicate that these results are not a function of the event window that has been chosen.
 
 \section{Speculation and foreign exchange returns}
-The nature of speculation is central to many contemporary questions about international financial markets.
-For example, regulations or taxes that reduce the amount of speculation may be beneficial if they limit risk-taking and encourage more long-term attitudes towards investment.
-If speculation is largely uninformed it will add to the level of noise and make it even harder for informed investors to overcome uncertainty and apply fundamental analysis about long-term value; uninformed speculators will increase the tendency towards activity that is focused on the short-term.
-However, if speculation is informed and part of the process of price discovery, regulatory measures to reduce its influence may also reduce liquidity and undermine the process of price discovery.
-In this case, increased volatility, heightened risk and reduced transparency together with other costs may provide a counteraction to any benefits.
+If foreign exchange speculation is uninformed noise, extreme speculation, whether measured by sentiment or weight of activity, will coincide with deviations from fundamental value; if speculation is informed, speculation is part of the process of price discovery and extremes will tell you nothing about the future as it is most likely to be a random walk.
+An event study showed that speculation is associated with price movement: exchange rates move in the direction of speculative sentiment and activity.
+However, once extreme, prices are as likely to continue as to reverse.
 
- The nature speculative activity in financial markets is central to a number of contemporary policy questions.
-Amongst these are the design and regulation of financial markets, the response to the surge in commodity prices that took place between 2002 and 2007 and the limits that should be put on the activities of financial institutions.
-This paper provides some evidence to assist in policy making in these areas.
+%Limitations of the study.  Can it be broadened? 
+The results that are recorded are not a function of the design of the event study.
+The event window and the quantile that is chosen to measure the extreme can be changed.
+Alternative measures produced very similar results.
+Figures \ref{fig:ES:1}, \ref{fig:ES:2}, \ref{fig:ES:3} and \ref{fig:ES:4} have much wider windows and do not find any alternative to the finding that extreme speculation is followed by a random walk.
+Alternative quantiles from 1\% and 99\% to 20\% and 80\% have also been assessed.
 
-The debate about a tax on financial transactions, initiated by \citet{TobinTax} and now called the \emph{the Tobin tax}, has been intended to reduce speculation.
-This is based on the belief that if speculation is uninformed, that it drives prices away from their fundamental value and, as a consequence, increases the volatility of financial asset prices.
-The argument for a tax on financial transactions can be presented as a tax on an externalities with the level of volatility of financial markets being the public cost.
+The covariance of speculative sentiment and speculative positions with the return indicates that this activity influences prices.
+It is not random.
+These results show that foreign exchange speculation is an important part of the price discovery process.
+The random walk after the extreme is indicative of foreign exchange prices absorbing most of the information that is available.
 
-The increase in volatility that is evident when speculative activity is high that has been identified in this work seems to provide some support to this argument.
-However, it is not clear that the direction of causality is from speculation to volatility.
-An alternative view would suggest that an increase in information raises the amount of informed speculation and that this increase in information naturally increase the volatility of financial assets.
-Causation runs from information to volatility with speculation between.
-If speculation is part of the process of price discovery, reducing its activity may reduce liquidity and hinder the process of price discovery.
-In that case unanticipated costs may offset some of the benefits that are expected from the tax on speculation.
-Volatility may even increase.
+It is encouraging that these two very different measures of speculative activity produce very similar results and that the results hold over different time frames: one wee for the option data and one week for the regulatory position data.
+It is less clear whether the results would be reproduced across markets for other financial assets.
+These results do not rule out uninformed speculation.
+The Friedman effect from the Noise-trader model depends on there being \emph{noise-trader risk} or some limits to arbitrage.
+The effect is not evident in this study of foreign exchange.
+However, other markets may have alternative institutional features that many not mean that these results can be extrapolated without questions being asked.
 
-If speculation is more informed than uninformed, reducing the amount of speculation will reduce the ratio of information to noise and reduce liquidity.
+%what are the implications of broadening?  Why is this interesting. 
+There are a number of very important policy debate that depend on the nature of speculation.
+The \emph{Tobin Tax} would put a charge against financial transactions, particularly in foreign exchange.
+Part of the argument depends on speculation being uninformed as the reduced activity that is caused by the tax would, it is argued, limit volatility and allow more long-term fundamental evaluations to become more prominent.
+However, if speculation is informed and part of the price discovery process, this tax may reduce liquidity and increase volatility.
 
-Regulatory changes in the US government bond market give some idea about way that a reduction in speculation can affect the smooth functioning of financial markets.
-Increased capital requirements and the so-called \emph{Volcker Rule} that would prevent investment banks from trading on their own account have raised the cost of trading US government bonds and reduced the ability of some key players to participate in the market.
-Though bid-ask spreads have remained stable, the New York Federal Reserve reports that \emph{market-depth} and the \emph{price impact} of large trades has deteriorated in the last two years.
-\footnote{See \citet{NYFedbonds} for an assessment of the evolution of bid-ask spreads, price impact and the size of trades.
-}  This is regarded as one of the reason that volatility has increased at times of tension and as liquidity has evaporated and large price changes, like the one see on 15 October 2014 have been recorded.
-\footnote{The US 10-year note yield fell nearly 30bp in 30 minutes, falling from 2.2\% to about 1.9\%.} A report by the NY Fed, the US Treasury and the CFTC about this very large adjustment in US bond prices concluded that a combination of the unwinding of a very large large short position, a decline in order book depth and changes in the provision of liquidity contributed to the action \citet{USbondflash}.
-
-The increased presence of high frequency trading (HFT) has also been seen by some as an unwanted development that has increased the amount of speculation and volatility.
-Again, the results of this study suggest that there should be some caution in removing speculative elements from the market.
-
-%this may be added.
-It appears that information is gradually absorbed into the price.\footnote{It may be that exchange rate traders take time to adjust to the information that comes from order flow or it may be evidence of the \emph{conservatism heuristic} that suggests that people are slow to change their opinion to new information.}
-
+%Future work.  Policy implications. 
 \bibliography{../../myrefs}

### 2015-10-24 Knitr/ElsPaper.tex
Paper ready Maybe one last look,  Ready to send to Andros.

@@
 \usepackage{pdflscape}
-\usepackage{appendix}
 \usepackage[flushleft]{threeparttable}
 \usepackage{multirow}
-%\usepackage[comma, sort&compress]{natbib}% Use the natbib reference package - read up on this to edit the reference style; if you want text (e.g.
-Smith et al., 2012) for the in-text references (instead of numbers), remove 'numbers'
-\usepackage{graphicx}
-%\bibliographystyle{plainnat}
 \bibliographystyle{agsm}
@@
 %\renewcommand[\harvardurl]{URL: \url}
- \usepackage{listings}
-  \usepackage{tikz}
- \usetikzlibrary{arrows,positioning}
- \usepackage{color}
- %\graphicspath{{../Pictures/}}
-\definecolor{mygrey}{gray}{0.95}
-\lstset{backgroundcolor=\color{mygrey}}
 \begin{document}
@@
 It is noise that emerges from the interaction of collective sentiment to create trends in asset prices.
-% This description provides an explanation for the observation from \citet{ShillerVol} that stock prices are more volatile than the underlying earnings information on which valuation should be based.
 
-Towards the other end of the spectrum is the idea that speculation is a stabilising force, providing liquidity and helping to ensure that markets swiftly find equilibrium.
+At the other end of the spectrum is the idea that speculation is a stabilising force, providing liquidity and helping to ensure that markets swiftly find equilibrium.
 \citet{FriedmanPositive} asserted that successful speculators would be those who managed to correctly buy when prices  were below fundamental value and sell above that price.
@@
 This paper contributes to the debate about the nature of speculation by providing clear evidence that speculation is more informed than uninformed.
-A unique set of option prices are used to identify the intensity of speculative sentiment; an event study method is used to assess the price action around extremes of speculative sentiment and speculative positions; these speculative extremes are compared to a common noise-trader model.
+A unique set of option prices are used to identify the intensity of speculative sentiment; an event study method investigates the price action around extremes of speculative sentiment and speculative positions; these speculative extremes are compared to a common noise-trader model.
 
@@
 1]{GortonNoise} categorise noise-traders as all those trading in markets for ``non-information-based reasons''.
-See \citet{Ramiah201589} for an overview of the noise-trader literature.
+See \citet{Ramiah201589} for a recent review of the noise-trader literature.
 
@@
 Noise-trading caused by psychological biases and restricted information is associated with behavioural issues, over-confidence and the use of irrational or purely fashionable investment tools.
-See \citet{DeBondtAnalysts} for an analysis of over-reaction and the \emph{representativeness heuristic}, \citet{OdeanOC} for a study of over-confidence and  \citet{ShillerFashion} for an assessment of investment fads.
+See \citet{DeBondtOver} for an analysis of over-reaction and the \emph{representativeness heuristic}, \citet{OdeanOC} for a study of over-confidence and  \citet{ShillerFashion} for an assessment of investment fads.
 Traders without information may not make decisions independently of others.
@@
 There are two parameters that determine the size of the reversal or Friedman-effect that is to be expected:  the price misperception $(\rho)$ and the weight of noise-traders in the market $(\mu)$.
-If speculation is uninformed and either of these parameters reach an extreme, prices should subsequently reverse.
-This reversal is the Friedman-effect discussed above.
+If speculation is uninformed and either of these parameters reach an extreme, prices should subsequently reverse.\footnote{See \citet{Hayward2013} for full details of a Monte-Carlo simulations of this model when the speculative misperception $(\rho)$ and the weight of speculation $(\mu)$ parameters are adjusted.
+As the expected deviation from fundamental value and the weight of speculators in the market becomes more extreme, this reversal becomes more likely.}  This reversal is the Friedman-effect discussed above.
 The following paragraphs show how these can be measured.
-\footnote{See \citet{Hayward2013} for full details of a Monte-Carlo simulations of this model when the speculative misperception $(\rho)$ and the weight of speculation $(\mu)$ parameters are adjusted.
-As the expected deviation from fundamental value and the weight of speculators in the market becomes more extreme, this reversal becomes more likely.}
 
@@
 Risk-reversal skew is the relative price of currency put and call options for the same strike price.} The risk-reversal skew is the relationship between a call option and the equivalent put.
-The \emph{Put-Call-Parity} in the  \citet{Black1973Options} option pricing model assumes that these will be the same, in practice they are not\footnote{While the Black-Scholes model is the basis for valuing options, where the price of a non-dividend paying stock trades at a discount to the forward rate equal to the risk free rate, the forward rate in a currency option is determined by the relative interest rate differential between the two currencies.
+The \emph{Put-Call-Parity} in the  \citet{Black1973Options} option pricing model assumes that these will be the same, in practice they are not.\footnote{While the Black-Scholes model is the basis for valuing options where the price of a non-dividend paying stock trades at a discount to the forward rate equal to the risk free rate, the forward rate in a currency option is determined by the relative interest rate differential between the two currencies.
 The \citet{GarmanKohlhagan} adjustment that accounts for the difference in interest payments on the two currencies is used.
@@
 This model allows the market practice of pricing options from the volatility implied by the model.
-In other words, the Garman-Kohlhagan is used to extract the volatility that is implied by a specific option price given the strike and other parameters.}  %A call option is the right to buy and a put option is the right to sell.
-The delta of an option measures the effect of a change in the spot price on the value of the option.
-This can range from zero for deeply out-of-the-money options to unity for deeply in the money strikes.
-The at-the-money delta is 50 as there is a 50:50 chance of the option being exercised.
-The benchmark for option risk-reversals is 25 delta.
-This is a measure of how far the option is out of the money or how far the strike is from the forward rate.
+In other words, the Garman-Kohlhagan is used to extract the volatility that is implied by a specific foreign exchange option price given the strike and other parameters.}
 
 If the exchange rate is expected to follow a random walk with drift towards the 1 month forward, implied volatility on the put and the call for the same delta should be equal.
-Any deviation from this equivalence is an indication that a deviation from the random walk with drift is implied in the market price of options.\footnote{If the risk-reversal for a 1-month USD/CAD is quoted at 0.15-0.28\% and the implied volatility is 8.50\%, the market-maker would be willing to buy the 1-month 25 delta USD-put-CAD-call at 8.65\% and sell the USD-call-CAD-put at 8.50\%.
+Any deviation from this equivalence is an indication that a divergence from the random walk with drift is implied in the market price of options.\footnote{If the risk-reversal for a 1-month USD/CAD is quoted at 0.15-0.28\% and the implied volatility is 8.50\%, the market-maker would be willing to buy the 1-month 25 delta USD-put-CAD-call at 8.65\% and sell the USD-call-CAD-put at 8.50\%.
 The dealer pays 0.15\%.
-Alternatively, the dealer would sell a 25 delta call at 8.78\% and buy the USD-call-CAD-put at 8.50\%, earning 0.28\% \citet{Global}.
+Alternatively, the dealer would sell a 25 delta call at 8.78\% and buy the USD-call-CAD-put at 8.50\%, earning 0.28\%.
+See \citet{Global} for a technical review.
 }  Therefore, the relationship between implied volatility for calls and implied volatilities for puts for the same delta and maturity is an indication of how much more expensive calls are relative to the equivalent put or as the market bias towards puts over calls and as an indication of market sentiment.
 The larger the absolute level of the risk-reversal skew, the greater the relative price and, if the deviation is based on information, subsequent price action should be zero on average; if the deviation is based on noise, this movement should be reversed.
+See \citet{Yan2011} and \citet{FengZhangFriesen} for some analysis of implied volatility in the equity market.
 
@@
 
-An \emph{event study} is used to assess the effect of extreme speculation on future returns.
+The effect of extreme speculation on future returns is analysed with an event study.
 See \citet{Dolly1933} for the original version of the event study method and \citet{FamaFisherJensenRoll} for an overview of the method used in this study.
+However, expected returns are identified as being zero rather than established through a \emph{market-model}.
 The R package \emph{eventstudies} created by \citet{eventstudies} is used to manage the data.
@@
 The windows were chosen to correspond to one day and one week of trading.
-However, Figure \ref{fig:ES1}, which shows the extreme high for a 16 day event window, and it becomes clear that altering the size of the event window does not change the findings to any significant degree.
 \begin{sidewaystable}
@@
 Looking down the seventh column (the 4-day CARW) reveals the average cumulative return for the period from four days before the extreme speculative sentiment to the four days after the event.
-The extreme highs are (in this case) those points when the risk-reversal skew is at or above the $95^{th}$ percentile.
+The extreme highs are those points when the risk-reversal skew is at or above the $95^{th}$ percentile.
 For the whole event window (CARW), extreme highs (labelled "Hi") are associated with positive returns and the extreme lows (labelled "Lo") are associated with negative returns.
-In most cases these are significantly different from zero.
+In most instances these are significantly different from zero.
 Significance is identified as those cases where more than 95\% of 1000 means calculated from random bootstrap samples taken from the events are above zero.
@@
 
-These results do not appear to be a function of the percentile that is used to measure the extreme event or the size of the window that is selected.
-Additional cases were run using percentiles at 80\% and 99.5\%.
-For the event window, Figures \ref{fig:ES1} and \ref{fig:ES2} show the cumulative abnormal returns for the major exchanges rates relative to the US dollar across a 16 day event window.
-Figure \ref{fig:ES1} shows the extreme highs for as a $99^{th}$ percentile and figure \ref{fig:ES2} shows the extreme low for the $10^{th}$ percentile.
-The figures reveal the mean cumulative return (in solid red) and the 95\% confidence intervals (dashed blue).
-The confidence intervals are created with the bootstrap as outlined above.
-The only case where there seems to be any evidence of a reversal is for the extreme low risk-reversals on the Canadian dollar.
-
 \begin{figure}
@@
 \end{figure}
-
-While there is some evidence that speculation is informed, it is clear from figures \ref{fig:ES1} and \ref{fig:ES2} that exchange rates do not react instantly to news.
-There is a gradual absorption of information and a gradual evolution of prices.
-Serial correlation in asset prices is a consistent observation.
-See \citet{Jagadeesh}.
-In this case the serial correlation is associated with extreme speculation that appears to be informed.
 
@@
 \section{Speculation and foreign exchange returns}
-If foreign exchange speculation is uninformed noise, extreme speculation, whether measured by sentiment or weight of activity, will coincide with deviations from fundamental value; if speculation is informed, speculation is part of the process of price discovery and extremes will tell you nothing about the future as it is most likely to be a random walk.
+If foreign exchange speculation is uninformed noise, extreme speculation, whether measured by sentiment or weight of activity, will coincide with deviations from fundamental value; if speculation is informed, speculation is part of the process of price discovery and extremes will provide no information and future prices will follow a random walk.
 An event study showed that speculation is associated with price movement: exchange rates move in the direction of speculative sentiment and activity.
@@
 
-%Limitations of the study.  Can it be broadened? 
 The results that are recorded are not a function of the design of the event study.
@@
 Alternative measures produced very similar results.
-Figures \ref{fig:ES:1}, \ref{fig:ES:2}, \ref{fig:ES:3} and \ref{fig:ES:4} have much wider windows and do not find any alternative to the finding that extreme speculation is followed by a random walk.
 Alternative quantiles from 1\% and 99\% to 20\% and 80\% have also been assessed.
+Figures \ref{fig:ES1}, \ref{fig:ES2}, \ref{fig:ES3} and \ref{fig:ES4} have much wider windows and do not find any alternative to the view that extreme speculation is followed by a random walk.
 
@@
 It is not random.
-These results show that foreign exchange speculation is an important part of the price discovery process.
+These results show that foreign exchange speculation, whether measured using either sentiment or the weight of positions, is an important part of the price discovery process.
 The random walk after the extreme is indicative of foreign exchange prices absorbing most of the information that is available.
 
-It is encouraging that these two very different measures of speculative activity produce very similar results and that the results hold over different time frames: one wee for the option data and one week for the regulatory position data.
+It is encouraging that these two very different measures of speculative activity produce very similar results and that the results hold over different time frames: one week for the option data and one week for the regulatory position data.
 It is less clear whether the results would be reproduced across markets for other financial assets.
@@
 The effect is not evident in this study of foreign exchange.
-However, other markets may have alternative institutional features that many not mean that these results can be extrapolated without questions being asked.
+However, this is a large and very liquid market, other asset classes may have alternative institutional features that many not mean that these results can be extrapolated without important institutional questons being asked.
 
-%what are the implications of broadening?  Why is this interesting. 
-There are a number of very important policy debate that depend on the nature of speculation.
 The \emph{Tobin Tax} would put a charge against financial transactions, particularly in foreign exchange.
@@
 
-%Future work.  Policy implications. 
 \bibliography{../../myrefs}

### 2015-10-25 Knitr/ElsPaper.tex
Updated ElsPaper.tex Ready to be assessed by Andros.  Conclusion has been improved.

@@
 \begin{keyword}
-JEL classification: G12, G14, G15; information, noise-trading, event studies.
+JEL classification: G12, G14, G15; event studies, information, noise-trading, speculation.
 \end{keyword}
@@
 
-The aim of this paper is to understand more about the nature of speculation by investigating the relationship between the intensity of speculation and the weight of speculative positions relative to the movement of prices in the foreign exchange market\footnote{The foreign exchange market is chosen for this study because it is the largest and most liquid financial market in the world.
+The aim of this paper is to understand more about the nature of speculation by investigating the relationship between the intensity of speculation and the weight of speculative positions relative to the movement of prices in the foreign exchange market\footnote{\label{FX}The foreign exchange market is chosen for this study because it is the largest and most liquid financial market in the world.
 The latest survey taken by the Bank for International Settlements (BIS) in April 2013 estimated average daily turnover at \$5.4trn with average daily spot transactions of \$2.0trn See \citet{BISFX2013}.
@@
 \section{Speculation and foreign exchange returns}
-If foreign exchange speculation is uninformed noise, extreme speculation, whether measured by sentiment or weight of activity, will coincide with deviations from fundamental value; if speculation is informed, speculation is part of the process of price discovery and extremes will provide no information and future prices will follow a random walk.
+If foreign exchange speculation is uninformed noise, extreme speculation, whether measured by sentiment or weight of activity, will coincide with deviations from fundamental value; if speculation is informed, speculation is part of the process of price discovery and extremes will provide no information about future prices.
 An event study showed that speculation is associated with price movement: exchange rates move in the direction of speculative sentiment and activity.
 However, once extreme, prices are as likely to continue as to reverse.
-
-The results that are recorded are not a function of the design of the event study.
-The event window and the quantile that is chosen to measure the extreme can be changed.
-Alternative measures produced very similar results.
-Alternative quantiles from 1\% and 99\% to 20\% and 80\% have also been assessed.
-Figures \ref{fig:ES1}, \ref{fig:ES2}, \ref{fig:ES3} and \ref{fig:ES4} have much wider windows and do not find any alternative to the view that extreme speculation is followed by a random walk.
-
-The covariance of speculative sentiment and speculative positions with the return indicates that this activity influences prices.
-It is not random.
-These results show that foreign exchange speculation, whether measured using either sentiment or the weight of positions, is an important part of the price discovery process.
+Speculative sentiment and speculative activity influence foreign exchange prices as indicated by the covariance of speculative measures with returns.
+These results show that foreign exchange speculation is an important part of the price discovery process.
 The random walk after the extreme is indicative of foreign exchange prices absorbing most of the information that is available.
 
-It is encouraging that these two very different measures of speculative activity produce very similar results and that the results hold over different time frames: one week for the option data and one week for the regulatory position data.
+The results are not a function of the design of the event study.
+The event window and the quantiles that are chosen to measure the extreme can be changed.
+Alternative measures produced very similar results.
+Alternative quantiles from 1\% and 99\% to 20\% and 80\% have been assessed; Figures \ref{fig:ES1}, \ref{fig:ES2}, \ref{fig:ES3} and \ref{fig:ES4} have much wider windows and do not find any alternative to the view that extreme speculation is followed by a random walk.
+
+It is encouraging that two very different measures of speculative activity produce very similar results and that the results hold over different time frames: one week for the option data and one week for the regulatory position data.
 It is less clear whether the results would be reproduced across markets for other financial assets.
-These results do not rule out uninformed speculation.
-The Friedman effect from the Noise-trader model depends on there being \emph{noise-trader risk} or some limits to arbitrage.
-The effect is not evident in this study of foreign exchange.
-However, this is a large and very liquid market, other asset classes may have alternative institutional features that many not mean that these results can be extrapolated without important institutional questons being asked.
+The results do not rule out uninformed speculation.
+The Friedman effect from the Noise-trader model, whereby the informed speculators take advantage of the space created by price pressure from the uniformed, depends on there being \emph{noise-trader risk} or some limits to arbitrage.
+The effect is not evident in this study.
+However, as discussed in Footnote \ref{FX}, the foreign exchange market is a particularly large and liquid market, other asset classes may have alternative institutional features that mean a different relationship or balance between the informed and uninformed and therefore different results.
 
-The \emph{Tobin Tax} would put a charge against financial transactions, particularly in foreign exchange.
+This is important because there is no unambiguous view of speculative activity and because speculation is an important issues in a number of policy debates.
+Most notably, the \emph{Tobin Tax} would put a charge against financial transactions, particularly in foreign exchange.
 Part of the argument depends on speculation being uninformed as the reduced activity that is caused by the tax would, it is argued, limit volatility and allow more long-term fundamental evaluations to become more prominent.
-However, if speculation is informed and part of the price discovery process, this tax may reduce liquidity and increase volatility.
+However, if speculation is informed and part of the price discovery process, this tax may have the opposite effect to what is envisaged.
+These results suggest that it could reduce liquidity and increase volatility.
 

### 2015-10-26 Knitr/ElsPaper.tex
Updated ELspaper.tex Taken out some of the footnotes and included a heading for references.

@@
 \citet{Delong1990noise} and \citet{shleifer1997limits} have presented models where the risk that noise-traders will move prices even further away from equilibrium before convergence can limit the activity of informed speculators when they are risk averse.
-This can be called \emph{noise-trader risk}.\footnote{This three-period, overlapping generation model with informed and uninformed traders has a price equation \begin{equation}
+This can be called \emph{noise-trader risk}.
+
+This three-period, overlapping generation model with informed and uninformed traders has a price equation \begin{equation}
 p_t = 1 + \frac{\mu(\rho_t - \rho^*)}{1 + r} + \frac{ \mu \rho^*}{r} - \frac{2 \gamma \mu^2 \sigma_{\rho}^2}{r(1 + r^2)}.
@@
 \end{equation} 
- where $p_t$ is the price; $\mu$ is the share of uninformed traders; $\rho^*$ is the average misperception of uninformed traders; $\rho_t$ is the price mis-perception of the uninformed traders at time t; $\gamma$ is the level of risk aversion; $\sigma^2_{p,t+1}$ is the variance of the misperception.}
+ where $p_t$ is the price; $\mu$ is the share of uninformed traders; $\rho^*$ is the average misperception of uninformed traders; $\rho_t$ is the price mis-perception of the uninformed traders at time t; $\gamma$ is the level of risk aversion; $\sigma^2_{p,t+1}$ is the variance of the misperception.
 
@@
 Risk-reversal skew is the relative price of currency put and call options for the same strike price.} The risk-reversal skew is the relationship between a call option and the equivalent put.
-The \emph{Put-Call-Parity} in the  \citet{Black1973Options} option pricing model assumes that these will be the same, in practice they are not.\footnote{While the Black-Scholes model is the basis for valuing options where the price of a non-dividend paying stock trades at a discount to the forward rate equal to the risk free rate, the forward rate in a currency option is determined by the relative interest rate differential between the two currencies.
+The \emph{Put-Call-Parity} in the  \citet{Black1973Options} option pricing model assumes that these will be the same, in practice they are not.
+While the Black-Scholes model is the basis for valuing options where the price of a non-dividend paying stock trades at a discount to the forward rate equal to the risk free rate, the forward rate in a currency option is determined by the relative interest rate differential between the two currencies.
 The \citet{GarmanKohlhagan} adjustment that accounts for the difference in interest payments on the two currencies is used.
@@
 This model allows the market practice of pricing options from the volatility implied by the model.
-In other words, the Garman-Kohlhagan is used to extract the volatility that is implied by a specific foreign exchange option price given the strike and other parameters.}
+In other words, the Garman-Kohlhagan is used to extract the volatility that is implied by a specific foreign exchange option price given the strike and other parameters.
 
@@
 
+\section*{References}
 \bibliography{../../myrefs}

### 2015-11-02 Knitr/ElsPaperFront.tex
Updated ElsPaper The main paper has been broken into Front and Main for the assessment.

--- 2015-10-24 Knitr/ElsPaper.tex
@@
 \usepackage{amsmath}
-\usepackage{setspace}
+%\usepackage{setspace}
 \usepackage{pdflscape}
@@
 \usepackage[colorlinks = true, citecolor = blue, linkcolor = blue]{hyperref}
-%\hypersetup{urlcolor=blue, colorlinks=true} % Colors hyperlinks in blue - change to black if annoying
-%\renewcommand[\harvardurl]{URL: \url}
 \begin{document}
@@
 \title{Speculation in foreign exchange: noise or information?}
-\author{Rob Hayward\footnote{University of Brighton Business School, Lewes Road, Brighton, BN2 4AT; Telephone 01273 642586.  rh49@brighton.ac.uk.}}
+\author{Rob Hayward\footnote{Brighton Business School, Lewes Road, Brighton, BN2 4AT; Telephone 01273 642586.  rh49@brighton.ac.uk.}}
 
 \begin{abstract}
-If foreign exchange prices are driven by speculation based on noise rather than information, extreme levels of speculative sentiment and speculative activity should lead to price reversals.
+If foreign exchange prices are driven by speculation that is based on noise rather than information, extreme levels of speculative sentiment and speculative activity should lead to price reversals.
 Informed speculation is sustainable as it is part of the price discovery process.
 A unique set of option prices measure speculative sentiment while regulatory positions determine the weight of speculative activity.
-An event study shows that extremes of speculative sentiment or speculative activity do not increase the probability of price reversal, though variance-measured risk increases.
+An event study shows that extremes of speculative sentiment or speculative activity do not increase the probability of price reversal.
 Price activity after the extreme is close to a random walk.
@@
 \begin{keyword}
-JEL classification: G12, G14, G15; information, noise-trading, event studies.
+JEL classification: D8; F3; G1. Event study; Information; Noise-trading; Speculation.
 \end{keyword}
@@
 \end{frontmatter}
-\section*{Introduction}
-Ideas about the way that speculation affects financial markets span a broad range.
-Towards one end is the view of \citet[p.
-101]{Keynes1936} that speculation is a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd''.
-Speculation is a problem that will cause booms, reversals and will discourage fundamental, long-term investment.
-It is noise that emerges from the interaction of collective sentiment to create trends in asset prices.
-
-At the other end of the spectrum is the idea that speculation is a stabilising force, providing liquidity and helping to ensure that markets swiftly find equilibrium.
-\citet{FriedmanPositive} asserted that successful speculators would be those who managed to correctly buy when prices  were below fundamental value and sell above that price.
-Therefore, Friedman argued, the speculative process would reward those able to access or utilise information relative to the uninformed and, through a Darwinian process, informed speculation would dominate.
-This view regards speculation as an key lubricant of financial markets as it provides liquidity and facilitates the process of price discovery whereby information is assimilated into price.
-
-The aim of this paper is to understand more about the nature of speculation by investigating the relationship between the intensity of speculation and the weight of speculative positions relative to the movement of prices in the foreign exchange market\footnote{The foreign exchange market is chosen for this study because it is the largest and most liquid financial market in the world.
-The latest survey taken by the Bank for International Settlements (BIS) in April 2013 estimated average daily turnover at \$5.4trn with average daily spot transactions of \$2.0trn See \citet{BISFX2013}.
-The cost of trading in this market, when measured in terms of the size of the bid-ask spread, is extremely low.
-An analysis by \citet{Steely2013} of the bid-ask spreads for USD-JPY, GBP-USD and EUR-USD for the period January 2001 to 2005 finds that the average spread for each is 0.342\%, 0.305\% and 0.326\% respectively.
-This would be the cost for a round-trip (buying and selling immediately).}.
-If speculators are uninformed and their activity can be regarded as noise, the most extreme speculative sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when reversals are most likely to take place; if speculators are informed prices should follow a random walk after extremes.
-
-This paper contributes to the debate about the nature of speculation by providing clear evidence that speculation is more informed than uninformed.
-A unique set of option prices are used to identify the intensity of speculative sentiment; an event study method investigates the price action around extremes of speculative sentiment and speculative positions; these speculative extremes are compared to a common noise-trader model.
-
-The rest of the paper proceeds as follows:  Section Two discusses the distinction between informed and uninformed speculators; Section Three talks about the measurement of speculation and the event study method; Section Four reviews the evidence; Section Five concludes.
-
-\section{Informed and uninformed speculation}
-Speculation plays an important part in several areas of financial economics.
-However, there is no clear, unambiguous theory of speculation.
-Speculation is frequently associated with the term \emph{trading}, but trading is broader as it encompasses market-making.
-A distinction made between the short-term orientation of speculation and trading and the more long-term activity of investment.
-Speculation and trading can be categorised as being informed or uninformed.
-Those trading without information are frequently called \emph{noise-traders} or \emph{liquidity traders}.
-However, the terms are not fixed and  \citet[p.
-1]{GortonNoise} categorise noise-traders as all those trading in markets for ``non-information-based reasons''.
-See \citet{Ramiah201589} for a recent review of the noise-trader literature.
-
-The distinction between informed and uninformed traders is a central component of microstructure research.
-The focus of the \citet{Kyle1985Continuous} model of market-making is the interaction of orders from the informed and the uninformed.
-On the assumption that uniformed orders arrive in a random fashion, the weight of orders implies something about the knowledge of the informed and should led to adjustment of the bid-ask spread as part of the process of price discovery.
-In a similar way \citet{glosten1985bid} describes a model where the bid-ask spread is a function of adverse selection, being caused by the risk that orders come from informed traders.
-
-\citet[p.529]{BlackNoise} suggested that the interaction of informed and uninformed activity was essential to the smooth functioning of financial markets.
-Noise, he argued, was the  ``the arbitrary element in expectations''.
-Therefore, noise makes financial markets possible but imperfect: the more noise-traders that there are, the more liquidity and the easier it is to trade; the more noise-traders, the higher the level of inefficiency in the market and the more likely that price will be driven from fundamental value.
-
-Uninformed activity provides a rationale and the opportunity for trading.
-Unless there is a difference of opinion about value, there is no reason to trade.
-Uninformed speculation can create space between price and fundamental value that can be exploited by informed speculators.
-This space can encourage informed traders to pay the cost of acquiring information.
-\citet{Grossman1980Impossibility} argue that there is equilibrium between the cost of obtaining information and the benefits that are expected to come from the use of the information.
-
-\citet{EasleyPIN} use information as an explanation for the effect of large orders on stock prices.
-The \emph{order effect} may be a consequence of the need to adjust inventories when a market-maker receives a large order; that effect may also be a function of the market-maker knowledge that large orders are more likely come from to informed traders seeking to realise their informational advantage.
-\citet{EasleyPIN2} develop a technique to estimate the probability that trading is informed.
-They find that the high liquidity stocks have the highest level of information and the most frequent arrival rate of informed traders but that price changes are limited by the volume of uniformed traders.
-Low liquidity stocks by contrast, have fewer information events but informed traders have a much larger effect on prices.
-
-\citet{EasleyPIN2} propose a model to try to measure informed market activity based on the assertion that the bid-ask spread is a function of the level of informed trading.
-Large orders are assumed to be informed as informed traders are most likely to want to swift use of their informational advantage.\footnote{Market-makers use Bayes' Rule to update their fundamental valuation by using order flow to estimate the joint probability that there has been an information event and that it has been good news or bad news, the authors seek to operationalise the following model:
-\begin{subequations}
-\begin{equation}
-b(t) = E[V_i|t] - \frac{\mu P_b(t)}{\varepsilon + \mu P_b(t)}(E[V-t|-\underline{V}])
-\end{equation}
-\begin{equation}
-a(t) = E[V_i|t] + \frac{\mu P_g(t)}{\varepsilon + \mu P_g(t)}(\overline{V}E+ [V-t|])
-\end{equation}
-\end{subequations}
-  where $b(t)$ is the bid price at time $t$; $V$ is the value of the security; $\mu$ is the share of informed traders in the market; $P_g(t)$ is the probability that there has been some good news and $P_b(t)$ is the probability that there has been some bad news; $\varepsilon$ is the share of uninformed traders in the market; $\underline{V}$ is the value if there is bad news and $\overline{V}$ is the value if there is good news.}
-
-However, identifying large orders and making an assessment of the level of informed trading is much harder in an over-the-counter market like foreign exchange.
-While electronic order books have become more widely used, much of the market-making still takes place through bilateral agreements and information about the nature of transactions including size and origin of orders is not widely available.\footnote{\citet{Lyons1995Microstructure} and \citet{Evans2002Order} have gained access to proprietary data and have used the initiation of orders as an explanatory variable in an equation to explain the evolution of major exchange rates.
-They find that $\$$ 1 billion of net US dollar purchase moves the Deutsche mark price of the US unit by 0.5 percent.
-They argue that the orders crystallise a range of information that is used to value currencies.}  \citet{LeeandReady} seek to use price action to infer the nature of the order that has been placed into the market.
-The \emph{tick-test} has four classifications: the up-tick, the down-tick, the zero-up-tick and the zero-down-tick.
-A trade is an up-tick if the price is higher than the previous trade.
-If the price is the same as the previous trade, if the previous trade was classified as an up-tick, this trade is a zero-up-tick.
-Under this system, the up-ticks are assumed to be buy orders and the down-ticks are sell orders.
-These estimates of the market activity can be used to estimate the level of informed trading.
-
-Informed speculation is also used as the explanation for the tendency of Rational Expectations Equilibrium (REE) to prevail in the long-run.
-Those economic agents able to determine the fundamental value or discern the underlying economic model will be able to speculate and drive prices towards equilibrium.
-This leaves uninformed speculation as a pure noise that can be characterised as white-noise random variable.\footnote{See \citet{muth1961Rational} for the original adjustment to a cobweb model.} While this construction may be justified for the analysis of microstructure and market efficiency, it is an over-simplification that is inappropriate when seeking more understanding of speculation itself.
-Institutional and psychological explanations for noise-trading frequently suggest that there are endogenous non-random elements to speculative activity.
-
-Noise-trading caused by psychological biases and restricted information is associated with behavioural issues, over-confidence and the use of irrational or purely fashionable investment tools.
-See \citet{DeBondtOver} for an analysis of over-reaction and the \emph{representativeness heuristic}, \citet{OdeanOC} for a study of over-confidence and  \citet{ShillerFashion} for an assessment of investment fads.
-Traders without information may not make decisions independently of others.
-Information may be inferred (perhaps incorrectly) from the buying and selling conducted by other traders, causing positive feedback and trends.\footnote{This is an example of the beauty contest that has been presented by \citet[p.101]{Keynes1936}.
-The aim is to determine not the underlying value but what others perceive to be the underlying value.}
-
-Liquidity trading suggests that institutional constraints may ensure that economic agents just execute customer orders at the current price or that they buy and sell with reference to a benchmark index or to fulfil a hedging strategy.
-The institutional explanations are also associated with the allocation of credit for investment and trading purposes:  as asset values decline exchanges, banks and other providers of liquidity require larger margins and deeper haircuts; investors liquidate their fund holdings; the attraction of hedging increases.
-See \citet{BrunnermeierLiquidity} and \citet{Gorton} for some discussion of the way that margin-calls and collateral constraints can force investment decisions.
-
-\citet{FriedmanPositive} said that the informed market participants would drive the uninformed from the market.
-However, a large proportion of uninformed traders in a market can undermine this process.
-\citet{Delong1990noise} and \citet{shleifer1997limits} have presented models where the risk that noise-traders will move prices even further away from equilibrium before convergence can limit the activity of informed speculators when they are risk averse.
-This can be called \emph{noise-trader risk}.\footnote{This three-period, overlapping generation model with informed and uninformed traders has a price equation \begin{equation}
-p_t = 1 + \frac{\mu(\rho_t - \rho^*)}{1 + r} + \frac{ \mu \rho^*}{r} - \frac{2 \gamma \mu^2 \sigma_{\rho}^2}{r(1 + r^2)}.
-\label{eq:NP2}
-\end{equation} 
- where $p_t$ is the price; $\mu$ is the share of uninformed traders; $\rho^*$ is the average misperception of uninformed traders; $\rho_t$ is the price mis-perception of the uninformed traders at time t; $\gamma$ is the level of risk aversion; $\sigma^2_{p,t+1}$ is the variance of the misperception.}
-
-  The De-Long noise-trader model also provides a useful benchmark against which to assess the activity of speculators.
-The following four effects are noted by \citet[pp.
-14-15]{Delong1990noise}:  the \emph{Friedman Effect}, is the original arbitrage from the misperception of the noise-traders; \emph{Price Pressure} is the \emph{noise-trader risk} that prices may diverge further before returning to equilibrium; \emph{Hold More} means that noise-traders become more involved in the market as their views become more extreme; \emph{Create Space} is the difference between noise-trader perception and the actual value, it creates the opportunity for informed traders to profit from their information.
-The aim of this paper is to establish whether the Friedman-effect can be identified.
-
-\section{Measuring speculation and the event study}
-There are two steps: identify speculation and assess what happens when it becomes extreme.
-If speculation is uninformed, extremes should cause a divergence from fundamental value and a price reversal should be anticipated; if informed and part of the price-discovery process, extreme speculation achieves a new equilibrium and the following price performance should be a random walk.
-
-The \citet[pp.
-14-15]{Delong1990noise} presented a widely-used model to understand the interaction of informed and uninformed traders.
-It is summarised by Equation \ref{eq:NP2}.
-There are two parameters that determine the size of the reversal or Friedman-effect that is to be expected:  the price misperception $(\rho)$ and the weight of noise-traders in the market $(\mu)$.
-If speculation is uninformed and either of these parameters reach an extreme, prices should subsequently reverse.\footnote{See \citet{Hayward2013} for full details of a Monte-Carlo simulations of this model when the speculative misperception $(\rho)$ and the weight of speculation $(\mu)$ parameters are adjusted.
-As the expected deviation from fundamental value and the weight of speculators in the market becomes more extreme, this reversal becomes more likely.}  This reversal is the Friedman-effect discussed above.
-The following paragraphs show how these can be measured.
-
-The measurement of speculative intensity must identify how far speculators believe that the market will move in a particular direction.
-In the De-Long model, this would mean that for Equation \ref{eq:NP2}, it is the absolute value of $\rho - 1$.
-In the absence of a daily survey of speculators, this study uses a unique series of option\emph{risk-reversal skew} as an objective measure of speculative sentiment.\footnote{It appears that this is the only period for which the data currently exist.
-The risk-reversal data is very difficult to obtain in an over-the-counter market like foreign exchange where there is no centralised exchange to record and collect information, generally being the property of quoting institutions.
-The risk-reversals are the average prices quoted at 10:00 GMT by major banks on the Thomson-Reuters composite option pages (OPT1)  for the period between January 2 1996 and September 6 2002.
-The risk-reversals are the spread of implied volatility for a 1 month 25 delta at-the-money call option relative to the equivalent put.
-Risk-reversal skew is the relative price of currency put and call options for the same strike price.} The risk-reversal skew is the relationship between a call option and the equivalent put.
-The \emph{Put-Call-Parity} in the  \citet{Black1973Options} option pricing model assumes that these will be the same, in practice they are not.\footnote{While the Black-Scholes model is the basis for valuing options where the price of a non-dividend paying stock trades at a discount to the forward rate equal to the risk free rate, the forward rate in a currency option is determined by the relative interest rate differential between the two currencies.
-The \citet{GarmanKohlhagan} adjustment that accounts for the difference in interest payments on the two currencies is used.
-The spot discount is determined therefore by the relative interest rate differential for the appropriate maturity.
-This model allows the market practice of pricing options from the volatility implied by the model.
-In other words, the Garman-Kohlhagan is used to extract the volatility that is implied by a specific foreign exchange option price given the strike and other parameters.}
-
-If the exchange rate is expected to follow a random walk with drift towards the 1 month forward, implied volatility on the put and the call for the same delta should be equal.
-Any deviation from this equivalence is an indication that a divergence from the random walk with drift is implied in the market price of options.\footnote{If the risk-reversal for a 1-month USD/CAD is quoted at 0.15-0.28\% and the implied volatility is 8.50\%, the market-maker would be willing to buy the 1-month 25 delta USD-put-CAD-call at 8.65\% and sell the USD-call-CAD-put at 8.50\%.
-The dealer pays 0.15\%.
-Alternatively, the dealer would sell a 25 delta call at 8.78\% and buy the USD-call-CAD-put at 8.50\%, earning 0.28\%.
-See \citet{Global} for a technical review.
-}  Therefore, the relationship between implied volatility for calls and implied volatilities for puts for the same delta and maturity is an indication of how much more expensive calls are relative to the equivalent put or as the market bias towards puts over calls and as an indication of market sentiment.
-The larger the absolute level of the risk-reversal skew, the greater the relative price and, if the deviation is based on information, subsequent price action should be zero on average; if the deviation is based on noise, this movement should be reversed.
-See \citet{Yan2011} and \citet{FengZhangFriesen} for some analysis of implied volatility in the equity market.
-
-Recording speculative activity in an over-the-counter market like foreign exchange is extremely difficult.
-However, there are some exchange-traded currency derivative markets in the US, and participants in these markets are required to report their positions to the US regulators.
-The Commodity Futures Trading Commission (CFTC) collect weekly information about the open positions held by private entities in the US derivative markets.
-A sub-set of this information is released to the public each week as the \emph{Commitment of Traders Report} (CoT).\footnote{The \citet{cot} require traders to categorise themselves as being either commercial (C) or non-commercial (NC).
-Commercial traders must have some underlying business interest in the security, commodity or instrument that they are trading.
-These could be seen as natural hedgers.
-Non-commercial accounts are generally considered to be speculators.
-See \citet{FuturesSanders} and \citet{FuturesWang} for examples of previous use of this data in this fashion.
-
-The data used run from September 30 1998 when the data started to be released on a weekly basis and continue up to December 31 2008.
-The Euro data runs from March 3 1998 to December 31 2008.
-There are two gaps in the Swiss Franc data for September 14 2004 and September 21 2004.
-The absence of these data seems to be a result of problems with the CFTC database.
-The contracts are for EUR 100,000; GBP 62,500; JPY 12,500,000; CHF 100,000 and CAD 100,000.}
-
-The effect of extreme speculation on future returns is analysed with an event study.
-See \citet{Dolly1933} for the original version of the event study method and \citet{FamaFisherJensenRoll} for an overview of the method used in this study.
-However, expected returns are identified as being zero rather than established through a \emph{market-model}.
-The R package \emph{eventstudies} created by \citet{eventstudies} is used to manage the data.
-Extreme measures of speculative sentiment and activity are the event and the cumulative abnormal returns for the whole of the event window (CARW) and the cumulative abnormal returns for after the event (CARA) are the exchange rates to be assessed.
-
-\section{Analysis of results}
-  Table \ref{tabref:RR1} records the result of the event studies that were carried out on extreme risk reversal.
-For each exchange rate the second column distinguishes between extreme high and extreme low, the third  column identifies the number of extreme events that were recorded and the following two are the cumulative average returns for the whole event window (CARW) and the cumulative average returns for the event and the period after (CARA).
-The windows were chosen to correspond to one day and one week of trading.
-\begin{sidewaystable}
-% multirow package is required.   
-%pdflandscape is required to facilitate the landscape position. 
-\begin{threeparttable}
-\caption{Event Study: Cumulative Abnormal Returns and Extreme Risk Reversal}
-\begin{tabular}{llccccccccccccc}  
- % \hline
-& &\multicolumn{6}{c}{$5^{th}$ and $95^{th}$ percentile} & \multicolumn{1}{c}{} & \multicolumn{6}{c}{$1^{st}$ and $99^{th}$ percentile}\\
- & &\multicolumn{3}{c}{1 day} & \multicolumn{3}{c}{4 day} & \multicolumn{1}{c}{} & \multicolumn{3}{c}{1 day} & \multicolumn{3}{c}{4 day} \\ 
-%\cline{4 - 10} \cline{13 - 19}
-& &\multicolumn{1}{ c}{N} & \multicolumn{1}{c}{CARW} & \multicolumn{1}{c}{CARA} & \multicolumn{1}{ c}{N} & \multicolumn{1}{c}{CARW} & \multicolumn{1}{c}{CARA} & \multicolumn{1}{c}{} & 
-\multicolumn{1}{ c}{N} & \multicolumn{1}{c}{CARW} & \multicolumn{1}{c}{CARA}  
-& \multicolumn{1}{ c}{N} & \multicolumn{1}{c}{CARW} & \multicolumn{1}{c}{CARA}  \\   
-\hline
-\multirow{2}{*}{EUR} 
-& Hi &  87 &  0.4103* &0.1586 &87 &0.9442* & 0.1505 &&22 &0.0309 & -0.1748 &22 & 0.5781*& -0.5387*  \\ 
-& Lo & 106 &-0.3313* & -0.0933 &106 &-0.0816* &-0.1991 & & 21&0.3040 &0.1820 &21 &0.7935* & 0.4377  \\
-\multirow{2}{*}{JPY}
-& Hi & 90 &0.4860* &0.1341 &90 &1.5158* &0.2531 & &20 &0.8227* &0.3078* &20 &2.4337* & 0.8284*  \\ 
-& Lo & 90 &-0.3168 &0.0945 &90 &-1.3532* &0.1984 & &17 &-1.6990* &-0.9347* &17 &-3.6751* & -1.2556*  \\
-\multirow{2}{*}{GBP}
-& Hi & 96 &0.3303* &0.1903* &96 &0.8966* &0.2547 & &21 &0.5252* &0.1774 &21 &1.3131* &0.3284  \\ 
-& Lo & 97 &-0.1894 &0.0312 &97 &-0.8031* &0.1500 & &24 &-0.2630 &-0.0866 &24 &-0.6660 &0.1888  \\
-\multirow{2}{*}{CHF}
-& Hi & 80 &0.0065 &-0.0565 &80 &0.5624* &0.1322 & &18 & -0.0022 & -0.1518 &18 &0.4887 &-0.1768  \\ 
-& Lo & 75 &-0.3716* & -0.1125& 75 &-1.0071* &-0.0219 & &18 &0.0064 &0.1704 &18 &-0.8661* &0.7631  \\
-\multirow{2}{*}{CAD}
-& Hi & 113 &-0.0038 &-0.0489 &113 &0.1872 &-0.1951 & & 24 &-0.1081 & -0.0462 &24 &0.3935* & -0.1592  \\ 
-& Lo & 97 &-0.1422* & -0.0571 & 97 & -0.2896* & 0.0035& & 26 & -0.2008 & -0.0146 &26 &0.1185 & 0.4560*  \\
-\multirow{2}{*}{AUD}
-& Hi & 101 &0.2539* &0.0125 &101 &0.8728* & 0.0875 &  &33 &0.5099* &0.0682 &33 & 1.1501*& 0.1046  \\ 
-& Lo & 138 &-0.4384* & -0.1711 & 138 & -1.14935* & -0.2714* & &20 &-0.4334 & 0.0563 &20  & -1.7360* & 0.1914  \\
-\multirow{2}{*}{EURJPY}
-& Hi & 95 &-0.0504 &-0.1103 &95 &-0.0128 &-0.2579 & &23 &-0.2932 &-0.4188* &23 &-0.8203 & -1.0195  \\ 
-& Lo & 97 &-0.4746* & -0.1647 &97 &-1.6757* &-0.4595 & &20 &-1.3630* & -0.6976* & 20 & -3.6849* &-1.1999*  \\
-\multirow{2}{*}{EURGBP}
-& Hi & 91 & 0.1544 &0.0696 &91 &0.5176* &0.0979 & &28 &0.2048 &0.1195 &28 & 0.4146 & -0.0645  \\ 
-& Lo & 122 & 0.0451 & 0.0996 &122 &-0.3097* &0.2903* & & 21 &-0.07523 &-0.0780 &21 & -0.4120 &
--0.1290  \\
-\multirow{2}{*}{EURCHF}
-& Hi & 77 & 0.1183* & 0.0690 &77 & 0.3444* & 0.0749& & 	31 & 0.1482* &0.0739  &31 & 0.4993* 
-& 0.1816*  \\ 
-& Lo & 94 &-0.0344 & 0.0194 & 94 & -0.1596 &0.1952* & &24 & -0.3626* &-0.1483 & 24 & -0.7573* 
-& 0.0875  \\
-\hline
-\label{tabref:RR1}
-\end{tabular}
-\begin{tablenotes}
-\small 
-\item Where Hi means extreme high in the risk-reversal skew and Lo means extreme low in the risk reversal skew; extreme is calculated as being equal or above the $95^{th}$ or $99^{th}$ percentile or equal or below the $5^{th}$ or $1^{st}$ percentile respectively; CARW is the cumulative abnormal return for the whole window, before and after the extreme event, and CARA is the cumulative abnormal return for the period after the event, which is the event day and the window; abnormal return is anything that is different from zero; the asterisk denotes significantly different from zero, where statistical significance is more than 95\% of 1000 means calculated from random bootstrap samples from the events are above or below zero respectively.
-\end{tablenotes}
-\end{threeparttable}  
-\end{sidewaystable}
-
-As speculators become more extreme in their opinion, exchange rates are driven in the direction of sentiment.
-Looking down the seventh column (the 4-day CARW) reveals the average cumulative return for the period from four days before the extreme speculative sentiment to the four days after the event.
-The extreme highs are those points when the risk-reversal skew is at or above the $95^{th}$ percentile.
-For the whole event window (CARW), extreme highs (labelled "Hi") are associated with positive returns and the extreme lows (labelled "Lo") are associated with negative returns.
-In most instances these are significantly different from zero.
-Significance is identified as those cases where more than 95\% of 1000 means calculated from random bootstrap samples taken from the events are above zero.
-Returns that are statistically different from zero are identified by an asterisk.
-
-That sentiment drives price appears evident from the covariance of speculative sentiment with exchange rate prices.
-This finding is consistent with the evidence of previous research that has used order-flow or sentiment from the likes of \citet{Evans2002Order}, \citet{FuturesSanders} and \citet{FuturesWang}.
-If speculation is not informed, the extreme is caused by herding or institutional features; if speculation is informed, it is facilitating the absorption of this information into the price.
-
-Does knowledge about the extreme at event time $t = 0$ provide information about the future returns?
-If this speculation is noise, informed traders, such as real money or long-term fundamental accounts, should take advantage of the space that has been created between the price and the fundamental value and a reversal should be seen.
-Whether this is the case is recorded in columns 5, 8, 11 and 14 of Table \ref{tabref:RR1}.
-A reversal is indicated by a significant negative reading for highs or positive reading for lows.
-Looking, for example, at the four-day window in column 8, it is clear that the extreme risk reversal skew does not provide any information about future returns.
-Of the eighteen cases using the $5^{th}$ and $95^{th}$ percentiles to identify extremely negative and extremely positive sentiment respectively, ten show a continuation (rather than a reversal) that is not significantly different from zero; one shows a significant continuation; five show reversals that are not significant; only two show significant reversal - EURCHF in each case.
-For the more extreme event when the $99^{th}$ percentile is used, there is even less clarity.
-Just over half the cases with the four-day window show a reversal (8 of 18), but only 2 of these are statistically significant.
-Exchange rate prices are as likely to go up as down after an extreme and there does not appear to be any information about the future from knowledge of these extremes.
-Extremes seem to be followed by a random walk as would be the case if the extreme were the result of information being absorbed into the price.
-\begin{figure}
-\graphicspath{{../Figures/}}
-\centering
-\includegraphics[scale=0.8]{RRCum16}
-\caption{Event Study: Extreme (High $99{th}$ percentile) Risk Reversal Skew and 16 day event window: Risk reversals are the ratio of implied volatility on calls to the equivalent put; extreme high is at or above the $99^{th}$ percentile for the whole range, cumulative returns are the sum of log exchange rate returns.
-The event day is the day of the extreme risk reversal reading.
-If speculation is uninformed, extremes should be followed by reversals; if speculation is informed, extremes are more likely to be followed by a random walk.
-The solid line is the cumulative mean return from the start of the event window; the dashed lines show the 95\% confidence intervals constructed from 1000 random bootstrap samples of appropriate event window day.}
-\label{fig:ES5}
-\end{figure}
-
-\begin{figure}
-\graphicspath{{../Figures/}}
-\centering
-\includegraphics[scale=0.8]{RRCum16a}
-\caption{Event Study: Extreme (Low $10{th}$ percentile) Risk Reversal Skew and 16 day event window: Risk reversals are the ratio of implied volatility on calls to the equivalent put; extreme low is at or below the $10^{th}$ percentile for the whole range, cumulative returns are the sum of log exchange rate returns.
-The event day is the day of the extreme risk reversal reading.
-If speculation is uninformed, extremes should be followed by reversals; if speculation is informed, extremes are more likely to be followed by a random walk.
-The solid line is the cumulative mean return from the start of the event window; the dashed lines show the 95\% confidence intervals constructed from 1000 random bootstrap samples of appropriate event window day.}
-\label{fig:ES2}
-\end{figure}
-
-\begin{figure}
-\graphicspath{{../Figures/}}
-\centering
-\includegraphics[scale=0.8]{RRCum16}
-\caption{Event Study: Extreme (High $99{th}$ percentile) Risk Reversal Skew and 16 day event window: Risk reversals are the ratio of implied volatility on calls to the equivalent put; extreme high is at or above the $99^{th}$ percentile for the whole range, cumulative returns are the sum of log exchange rate returns.
-The event day is the day of the extreme risk reversal reading.
-If speculation is uninformed, extremes should be followed by reversals; if speculation is informed, extremes are more likely to be followed by a random walk.
-The solid line is the cumulative mean return from the start of the event window; the dashed lines show the 95\% confidence intervals constructed from 1000 random bootstrap samples of appropriate event window day.}
-\label{fig:ES1}
-\end{figure}
-
-Table \ref{tabref:SP1} shows the results from the event studies that were carried out using speculative positions as reported to US regulators.
-The table works in a similar way to Table \ref{tabref:RR1}.
-The first column is the exchange rate, defined as units per US dollar.
-The table is then split into two parts.
-The first part uses the S1 measure, the net non-commercial (speculative) positions relative to total non-commercial (speculative) positions; the second is S2 as the net non-commercial position relative to open interest or the total outstanding open positions.
-Therefore, the first captures the balance of speculative positions while the second captures the balance of speculative positions relative to the whole market.
-If speculators become more prevalent or dominant in the market, this will show up in S2 but not S1.
-S1 is more sensitive to changes in sentiment and more volatile.
-Each of these sections is broken into the results for a 2 week window and a 4 week window.
-There is a row for extreme high and extreme low which represents extreme long positions or extreme short positions; the first  column identifies the number of cases; CARW is the cumulative abnormal return for the whole event window; CARA is the cumulative abnormal return for the period after the event to the end of the window.
-
-\begin{sidewaystable}
-% multirow package is required.   
-%pdflandscape is required to facilitate the landscape position. 
-\begin{threeparttable}
-\caption{Event Study: Cumulative Abnormal Returns and Extreme Speculative Positions}
-\begin{tabular}{llccccccccccccc}	
- % \hline
-& &\multicolumn{6}{c}{S1 - Net Long per Speculators} & \multicolumn{1}{c}{} & \multicolumn{6}{c}{S2 -Net Long per Open Positions}\\
- & &\multicolumn{3}{c}{2 week} & \multicolumn{3}{c}{4 week} & \multicolumn{1}{c}{} & \multicolumn{3}{c}{2 week} & \multicolumn{3}{c}{4 week} \\ 
-%\cline{4 - 10} \cline{13 - 19}
-& &\multicolumn{1}{ c}{N} & \multicolumn{1}{c}{CARW} & \multicolumn{1}{c}{CARA} & \multicolumn{1}{ c}{N} & \multicolumn{1}{c}{CARW} & \multicolumn{1}{c}{CARA} & \multicolumn{1}{c}{} & 
-\multicolumn{1}{ c}{N} & \multicolumn{1}{c}{CARW} & \multicolumn{1}{c}{CARA}  
-& \multicolumn{1}{ c}{N} & \multicolumn{1}{c}{CARW} & \multicolumn{1}{c}{CARA}  \\   
-\hline
-\multirow{2}{*}{EUR} 
-& Hi &  27 &  2.9415* & 1.3163* &27 & 3.8059* & 1.7341* &&27 & 1.0974* & 0.2896 &27 & 1.7667*& 0.0419  \\ 
-& Lo & 27 & -2.4258* & -1.2512* & 27 &-3.9235* &-1.9254* & & 27 & -2.5122* & -1.0607* & 27 & -4.4478* & -1.9285* \\
-\multirow{2}{*}{JPY}
-& Hi & 43 & 1.9749* & 0.7286* & 43 & 2.8628* & 0.7909 & & 43 & 2.2463* & 1.0554* &43 & 3.7125* & 1.4443*  \\ 
-& Lo & 43 & -2.2014* & -1.1312* & 43&-3.2488* & -1.5758* & &43 & -2.0605* & -0.9791* &43  & -2.6865* & -0.8318  \\
-\multirow{2}{*}{GBP}
-& Hi & 43 & 0.7102 & -0.2116 &43 & 1.0172* & -0.7902* & &43 & 1.2916* & 0.2370 &43 &1.6912* &0.0126  \\ 
-& Lo & 43 &-0.6979* &0.2902 &43 &-1.5330* &0.1014 & &43 &-1.8993* &-0.1415 &43 &-2.8830* & -0.0768 \\
-\multirow{2}{*}{CHF}
-& Hi & 43 & 1.4400* & -0.0988 & 43 & 2.8665* & -0.0993 & & 43 &  2.5260* &  0.5074 & 43 & 3.6060* & 0.2158  \\ 
-& Lo & 43 & -2.1816* &  -0.9770* & 43 &-3.3030* & -0.09798* & & 43 & -1.2605* & -0.3070& 43 &-0.1.6021* & 0.0670  \\
-\multirow{2}{*}{CAD}
-& Hi & 43 & 1.8658* & 1.0195* & 43 & 3.5298* & 1.8319* & & 43 & 2.0160* & 0.8788* &43 & 3.0995* & 0.9921*  \\ 
-& Lo & 43 &-1.3238* & -0.5548* & 43 & -1.9151* & -0.7859* & & 43 & -1.2462* & -0.4505*  &43 & -1.9390*  & -0.6195*  \\
-\hline
-\label{tabref:SP1}
-\end{tabular}
-\begin{tablenotes}
-\small 
-\item Where Hi means extreme high, either S1 which is the measure of net long non-commercial (speculative) positions per total of speculators) or S2 which is the measure of net long non-commercial (speculative) positions per open interest (total outstanding open positions); high and low are above the $95^{th}$ percentile or below the $5^{th}$ percentile respectively; CARW is the cumulative abnormal return for the whole window, before and after the extreme event, and CARA is the cumulative abnormal return for the period after the event, which is the event day and the window; abnormal return is anything that is different from zero; the asterisk denotes significantly different from zero, where statistical significance is that more than 95\% of 1000 means calculated from random bootstraps from the extreme readings are above or below zero.
-\end{tablenotes}	
-\end{threeparttable}  
-\end{sidewaystable}
-
-Once again, there is a very strong association between returns and the activity of speculators.
-For the 4-day window and the S1 and S2 measures, every one of the currencies studied, high and low, had a mean cumulative return for the whole of the event window that was in the same direction as sentiment.
-When re-sampled a thousand times, more than ninety five percent of the means calculated were greater than zero in all cases but one (GBP).
-There is evidence here of momentum and of prices following the movement of speculator positions.
-
-\begin{figure}
-\graphicspath{{../Figures/}}
-\centering
-\includegraphics[scale=0.8]{FPCum6w}
-\caption{Event Study: Extreme (Low $5{th}$ percentile) Risk Reversal Skew and 16 day event window: Risk reversals are the ratio of implied volatility on calls to the equivalent put; extreme low is at or above the $5^{th}$ percentile for the whole range, cumulative returns are the sum of log exchange rate returns.
-The event day is the day of the extreme risk reversal reading.
-If speculation is uninformed, extremes should be followed by reversals; if speculation is informed, extremes are more likely to be followed by a random walk.
-The solid line is the cumulative mean return from the start of the event window; the dashed lines show the 95\% confidence intervals constructed from 1000 random bootstrap samples of appropriate event window day.}
-\label{fig:ES3}
-\end{figure}
-
-\begin{figure}
-\graphicspath{{../Figures/}}
-\centering
-\includegraphics[scale=0.8]{FPCum6wa}
-\caption{Event Study:  Extreme ($90^{th}$ percentile), Non-commercial net long (S1), 6 week window.
-Non-commercial net per OI (S2) are the net speculative positions (as reported to US regulator the CFTC) relative to all open positions; extreme high is at or above the $90^{th}$ percentile for the whole range, cumulative returns are the sum of log exchange rate returns.
-The event day is the day of the extreme reading as a measure of extreme speculative activity.
-If speculation is uninformed, extremes should be followed by reversals; if speculation is informed, extremes are more likely to be followed by a random walk.}
-\label{fig:ES4}
-\end{figure}
-
-There is little evidence that these self-reported speculators are driving prices away from fundamentals with their activity.
-Sixty percent of the S1 cases still show a significant abnormal return in the direction of the extreme 4 weeks after the extreme has been reached; only thirty percent of the cases show a reversal and in only one of those do ninety five percent of the means calculated after re-sampling remain above zero.
-For S2 there is even less clarity, with ninety percent of the cases showing a continuation and forty percent of them showing significant continuation.
-The evidence here with speculative positions is the same as that of speculative sentiment.
-Prices move in the direction of speculative activity but speculative extremes do not provide information about the future.
-The price action of the event study is consistent with speculation being informed rather than uninformed.
-
-\section{Speculation and foreign exchange returns}
-If foreign exchange speculation is uninformed noise, extreme speculation, whether measured by sentiment or weight of activity, will coincide with deviations from fundamental value; if speculation is informed, speculation is part of the process of price discovery and extremes will provide no information and future prices will follow a random walk.
-An event study showed that speculation is associated with price movement: exchange rates move in the direction of speculative sentiment and activity.
-However, once extreme, prices are as likely to continue as to reverse.
-
-The results that are recorded are not a function of the design of the event study.
-The event window and the quantile that is chosen to measure the extreme can be changed.
-Alternative measures produced very similar results.
-Alternative quantiles from 1\% and 99\% to 20\% and 80\% have also been assessed.
-Figures \ref{fig:ES1}, \ref{fig:ES2}, \ref{fig:ES3} and \ref{fig:ES4} have much wider windows and do not find any alternative to the view that extreme speculation is followed by a random walk.
-
-The covariance of speculative sentiment and speculative positions with the return indicates that this activity influences prices.
-It is not random.
-These results show that foreign exchange speculation, whether measured using either sentiment or the weight of positions, is an important part of the price discovery process.
-The random walk after the extreme is indicative of foreign exchange prices absorbing most of the information that is available.
-
-It is encouraging that these two very different measures of speculative activity produce very similar results and that the results hold over different time frames: one week for the option data and one week for the regulatory position data.
-It is less clear whether the results would be reproduced across markets for other financial assets.
-These results do not rule out uninformed speculation.
-The Friedman effect from the Noise-trader model depends on there being \emph{noise-trader risk} or some limits to arbitrage.
-The effect is not evident in this study of foreign exchange.
-However, this is a large and very liquid market, other asset classes may have alternative institutional features that many not mean that these results can be extrapolated without important institutional questons being asked.
-
-The \emph{Tobin Tax} would put a charge against financial transactions, particularly in foreign exchange.
-Part of the argument depends on speculation being uninformed as the reduced activity that is caused by the tax would, it is argued, limit volatility and allow more long-term fundamental evaluations to become more prominent.
-However, if speculation is informed and part of the price discovery process, this tax may reduce liquidity and increase volatility.
-
-\bibliography{../../myrefs}
 

### 2015-11-02 Knitr/ElsPaperMain.tex
Updated ElsPaper The main paper has been broken into Front and Main for the assessment.

--- 2015-10-26 Knitr/ElsPaper.tex
@@
 \usepackage[colorlinks = true, citecolor = blue, linkcolor = blue]{hyperref}
-%\hypersetup{urlcolor=blue, colorlinks=true} % Colors hyperlinks in blue - change to black if annoying
-%\renewcommand[\harvardurl]{URL: \url}
 \begin{document}
-\begin{frontmatter}
-\title{Speculation in foreign exchange: noise or information?}
-\author{Rob Hayward\footnote{University of Brighton Business School, Lewes Road, Brighton, BN2 4AT; Telephone 01273 642586.  rh49@brighton.ac.uk.}}
 
-\begin{abstract}
-If foreign exchange prices are driven by speculation based on noise rather than information, extreme levels of speculative sentiment and speculative activity should lead to price reversals.
-Informed speculation is sustainable as it is part of the price discovery process.
-A unique set of option prices measure speculative sentiment while regulatory positions determine the weight of speculative activity.
-An event study shows that extremes of speculative sentiment or speculative activity do not increase the probability of price reversal, though variance-measured risk increases.
-Price activity after the extreme is close to a random walk.
-This supports the view that speculation is informed and part of the process of price-discovery rather than uninformed noise.
-\end{abstract}
-\begin{keyword}
-JEL classification: G12, G14, G15; event studies, information, noise-trading, speculation.
-\end{keyword}
-
-\end{frontmatter}
 \section*{Introduction}
@@
 The effect is not evident in this study.
-However, as discussed in Footnote \ref{FX}, the foreign exchange market is a particularly large and liquid market, other asset classes may have alternative institutional features that mean a different relationship or balance between the informed and uninformed and therefore different results.
+However, the foreign exchange market is a particularly large and liquid market, other asset classes may have alternative institutional features that mean a different relationship or balance between the informed and uninformed and therefore different results.
 
@@
 Most notably, the \emph{Tobin Tax} would put a charge against financial transactions, particularly in foreign exchange.
-Part of the argument depends on speculation being uninformed as the reduced activity that is caused by the tax would, it is argued, limit volatility and allow more long-term fundamental evaluations to become more prominent.
+Part of the argument depends on speculation being uninformed as the reduced activity that is caused by the tax would, it is argued, limit volatility and allow more long-term fundamental evaluations to become prominent.
 However, if speculation is informed and part of the price discovery process, this tax may have the opposite effect to what is envisaged.

### 2017-09-18 Knitr/ElsPaper.tex
Returning to paper I am returning to the paper after some time.  I want to trh to revive this and do something with it.  I have the suggestion for another source from Andros and I have some ideas from the reviewer.

@@
 \begin{keyword}
-JEL classification: G12, G14, G15; event studies, information, noise-trading, speculation.
+JEL classification: D8; F3; G1. Event study; Information; Noise-trading; Speculation.
 \end{keyword}

### 2017-11-26 Knitr/FinancialEconomics.tex
Updated some comments. I need to print.

--- 2015-11-02 Knitr/ElsPaperMain.tex
@@
 \section*{Introduction}
-Ideas about the way that speculation affects financial markets span a broad range.
-Towards one end is the view of \citet[p.
-101]{Keynes1936} that speculation is a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd''.
-Speculation is a problem that will cause booms, reversals and will discourage fundamental, long-term investment.
-It is noise that emerges from the interaction of collective sentiment to create trends in asset prices.
+Does speculation facilitate price discovery or instability?
+If it is orice discovery, it is beneficial and should be encouraged; if it is instability, welfare is enhanced by its reduction.
+This paper seeks to distinguish between these two characteristics by analysing those times when specualtion in the foreign exchange market is most extreme.
+A series of event studies are conducted on the extremes of speculative activity and speculative sentiment.
+The finding that spculative extremes do not provide information about subsequent returns implies that specualtion is part of the process of price discovery and that efforts to reduce it would leave to a reduction in information efficiency.
 
-At the other end of the spectrum is the idea that speculation is a stabilising force, providing liquidity and helping to ensure that markets swiftly find equilibrium.
-\citet{FriedmanPositive} asserted that successful speculators would be those who managed to correctly buy when prices  were below fundamental value and sell above that price.
+If markets are effective, speculation is a stabilising force, providing liquidity and helping to ensure that markets swiftly find equilibrium.
+i\citet{KeynesHedge} and \citet{HicksHedge} emphasise the role that speculation plays in providing liquidity and say that there is a return to this activity.
+Speculators provide a service for a fee that is determined by the difference between the price that they buy and the price at which they expect to sell.
+Informed speculators will tend to drive out the uninformed noise-traders.
+\citet{FriedmanPositive} asserted that successful speculators would be those buying below intrinsic value and selling above that rate.
 Therefore, Friedman argued, the speculative process would reward those able to access or utilise information relative to the uninformed and, through a Darwinian process, informed speculation would dominate.
@@
 
-The aim of this paper is to understand more about the nature of speculation by investigating the relationship between the intensity of speculation and the weight of speculative positions relative to the movement of prices in the foreign exchange market\footnote{\label{FX}The foreign exchange market is chosen for this study because it is the largest and most liquid financial market in the world.
+However, the later Keynes regarded speculation as a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd'\citet[p.
+101]{Keynes1936}.
+Speculation was a problem that would cause booms, reversals and would discourage fundamental, long-term investment.
+Speculation would emerge from the evolution of collective sentiment and could create trends in financial asset prices.
+
+This paper distinguishes between the two concepts of speculation by investigating the relationship between the intensity of speculation and the weight of speculative positions relative to the movement of prices in the foreign exchange market\footnote{\label{FX}The foreign exchange market is chosen for this study because it is the largest and most liquid financial market in the world.
 The latest survey taken by the Bank for International Settlements (BIS) in April 2013 estimated average daily turnover at \$5.4trn with average daily spot transactions of \$2.0trn See \citet{BISFX2013}.
@@
 This would be the cost for a round-trip (buying and selling immediately).}.
-If speculators are uninformed and their activity can be regarded as noise, the most extreme speculative sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when reversals are most likely to take place; if speculators are informed prices should follow a random walk after extremes.
+If speculators are uninformed and their activity can be regarded as noise, the most extreme speculative sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when reversals are most likely to take place; if speculators are informed their activity drives exchange rates towards equlibrium and prices should follow a random walk after these extremes.
 
-This paper contributes to the debate about the nature of speculation by providing clear evidence that speculation is more informed than uninformed.
-A unique set of option prices are used to identify the intensity of speculative sentiment; an event study method investigates the price action around extremes of speculative sentiment and speculative positions; these speculative extremes are compared to a common noise-trader model.
+The price action after periods of extreme speculative sentiment and extreme speculative positions are comparded to the results of a noise-trader model.
+Noise-trader reversals in price are not evident.
+However, there is evidence that speculation in the foreign exchange market is informed.
+The paper contributes to the understanding of speculation and uses a unique set of option prices to identify the intensity of speculative sentiment
 
@@
 However, there is no clear, unambiguous theory of speculation.
-Speculation is frequently associated with the term \emph{trading}, but trading is broader as it encompasses market-making.
-A distinction made between the short-term orientation of speculation and trading and the more long-term activity of investment.
+Speculation is frequently associated with the term \emph{trading}, but trading is broader as it encompasses market-making and financial investment.
 Speculation and trading can be categorised as being informed or uninformed.
@@
 
-The distinction between informed and uninformed traders is a central component of microstructure research.
+The comparison of informed and uninformed traders is a central component of microstructure research.
 The focus of the \citet{Kyle1985Continuous} model of market-making is the interaction of orders from the informed and the uninformed.
-On the assumption that uniformed orders arrive in a random fashion, the weight of orders implies something about the knowledge of the informed and should led to adjustment of the bid-ask spread as part of the process of price discovery.
-In a similar way \citet{glosten1985bid} describes a model where the bid-ask spread is a function of adverse selection, being caused by the risk that orders come from informed traders.
+On the assumption that uniformed orders arrive in a random fashion, the weight of orders reveals the direction of private information and should led to adjustment of the bid-ask spread as part of the process of price discovery.
+Similarly,  \citet{glosten1985bid} describes a model where the bid-ask spread is a function of adverse selection, being caused by the risk that orders come from informed traders.
 
@@
 
-\citet{EasleyPIN2} propose a model to try to measure informed market activity based on the assertion that the bid-ask spread is a function of the level of informed trading.
-Large orders are assumed to be informed as informed traders are most likely to want to swift use of their informational advantage.\footnote{Market-makers use Bayes' Rule to update their fundamental valuation by using order flow to estimate the joint probability that there has been an information event and that it has been good news or bad news, the authors seek to operationalise the following model:
+\citet{EasleyPIN2} propose to measure informed market activity with the bid-ask spread.
+Large orders are assumed to be informed as participants seek to explit their informational advantage.\footnote{Market-makers use Bayes' Rule to update their fundamental valuation by using order flow to estimate the joint probability that there has been an information event and that it has been good news or bad news, the authors seek to operationalise the following model:
 \begin{subequations}
@@
 
-Informed speculation is also used as the explanation for the tendency of Rational Expectations Equilibrium (REE) to prevail in the long-run.
-Those economic agents able to determine the fundamental value or discern the underlying economic model will be able to speculate and drive prices towards equilibrium.
-This leaves uninformed speculation as a pure noise that can be characterised as white-noise random variable.\footnote{See \citet{muth1961Rational} for the original adjustment to a cobweb model.} While this construction may be justified for the analysis of microstructure and market efficiency, it is an over-simplification that is inappropriate when seeking more understanding of speculation itself.
+Informed speculation is also central to the mechanism proposed by \citet{muth1961Rational} to facilitate \emph{rational expectations}.
+Those economic agents able to determine fundamental value will be able to speculate and drive prices towards equilibrium.
+While this construction may be justified for the analysis of microstructure and market efficiency, it is an over-simplification that is inappropriate when seeking more understanding of speculation itself.
 Institutional and psychological explanations for noise-trading frequently suggest that there are endogenous non-random elements to speculative activity.
@@
 
-This three-period, overlapping generation model with informed and uninformed traders has a price equation \begin{equation}
-p_t = 1 + \frac{\mu(\rho_t - \rho^*)}{1 + r} + \frac{ \mu \rho^*}{r} - \frac{2 \gamma \mu^2 \sigma_{\rho}^2}{r(1 + r^2)}.
-\label{eq:NP2}
-\end{equation} 
- where $p_t$ is the price; $\mu$ is the share of uninformed traders; $\rho^*$ is the average misperception of uninformed traders; $\rho_t$ is the price mis-perception of the uninformed traders at time t; $\gamma$ is the level of risk aversion; $\sigma^2_{p,t+1}$ is the variance of the misperception.
-
-  The De-Long noise-trader model also provides a useful benchmark against which to assess the activity of speculators.
-The following four effects are noted by \citet[pp.
-14-15]{Delong1990noise}:  the \emph{Friedman Effect}, is the original arbitrage from the misperception of the noise-traders; \emph{Price Pressure} is the \emph{noise-trader risk} that prices may diverge further before returning to equilibrium; \emph{Hold More} means that noise-traders become more involved in the market as their views become more extreme; \emph{Create Space} is the difference between noise-trader perception and the actual value, it creates the opportunity for informed traders to profit from their information.
-The aim of this paper is to establish whether the Friedman-effect can be identified.
-
-\section{Measuring speculation and the event study}
-There are two steps: identify speculation and assess what happens when it becomes extreme.
-If speculation is uninformed, extremes should cause a divergence from fundamental value and a price reversal should be anticipated; if informed and part of the price-discovery process, extreme speculation achieves a new equilibrium and the following price performance should be a random walk.
-
-The \citet[pp.
-14-15]{Delong1990noise} presented a widely-used model to understand the interaction of informed and uninformed traders.
-It is summarised by Equation \ref{eq:NP2}.
-There are two parameters that determine the size of the reversal or Friedman-effect that is to be expected:  the price misperception $(\rho)$ and the weight of noise-traders in the market $(\mu)$.
-If speculation is uninformed and either of these parameters reach an extreme, prices should subsequently reverse.\footnote{See \citet{Hayward2013} for full details of a Monte-Carlo simulations of this model when the speculative misperception $(\rho)$ and the weight of speculation $(\mu)$ parameters are adjusted.
-As the expected deviation from fundamental value and the weight of speculators in the market becomes more extreme, this reversal becomes more likely.}  This reversal is the Friedman-effect discussed above.
-The following paragraphs show how these can be measured.
-
+\section{Measuring speculation}
 The measurement of speculative intensity must identify how far speculators believe that the market will move in a particular direction.

### 2017-11-27 Knitr/FinancialEconomics.tex
This is ready to go to Pascal I think.  Look at conclusion.,

@@
 \begin{document}
+\maketitle
+\begin{abstract}
+Does speculation facilitate price discovery or instability?
+If it is price discovery, it is beneficial and should be encouraged; if it is instability, welfare is enhanced by its reduction.
+This paper seeks to distinguish between these two characteristics by analysing those times when speculation in the foreign exchange market is most extreme.
+A series of event studies are conducted on the extremes of speculative activity and speculative sentiment.
+If speculation is noise,  extreme sentiment and extreme positions should lead to overshooting and increase risk of subsequent reversals.
+The finding that speculative extremes do not provide information about subsequent returns implies that speculation is part of the process of price discovery and that efforts to reduce it would leave to a reduction in information efficiency.
 
-\section*{Introduction}
-Does speculation facilitate price discovery or instability?
-If it is orice discovery, it is beneficial and should be encouraged; if it is instability, welfare is enhanced by its reduction.
-This paper seeks to distinguish between these two characteristics by analysing those times when specualtion in the foreign exchange market is most extreme.
-A series of event studies are conducted on the extremes of speculative activity and speculative sentiment.
-The finding that spculative extremes do not provide information about subsequent returns implies that specualtion is part of the process of price discovery and that efforts to reduce it would leave to a reduction in information efficiency.
+\end{abstract}
 
-If markets are effective, speculation is a stabilising force, providing liquidity and helping to ensure that markets swiftly find equilibrium.
-i\citet{KeynesHedge} and \citet{HicksHedge} emphasise the role that speculation plays in providing liquidity and say that there is a return to this activity.
-Speculators provide a service for a fee that is determined by the difference between the price that they buy and the price at which they expect to sell.
-Informed speculators will tend to drive out the uninformed noise-traders.
-\citet{FriedmanPositive} asserted that successful speculators would be those buying below intrinsic value and selling above that rate.
-Therefore, Friedman argued, the speculative process would reward those able to access or utilise information relative to the uninformed and, through a Darwinian process, informed speculation would dominate.
-This view regards speculation as an key lubricant of financial markets as it provides liquidity and facilitates the process of price discovery whereby information is assimilated into price.
-
+\section{Introduction}
+If speculation is based on information it is a stabilising force, providing liquidity and helping to ensure that markets swiftly find equilibrium.
+\citet{KeynesHedge} and \citet{HicksHedge} emphasised the role that speculation played in providing liquidity.
+Speculators provide a service for a fee that is determined by the return on their activity.
+Informed speculators will tend to drive out the uninformed noise-traders \citet{FriedmanPositive}.
+Successful speculators would be those buying below intrinsic value and selling above that rate.
 However, the later Keynes regarded speculation as a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd'\citet[p.
 101]{Keynes1936}.
-Speculation was a problem that would cause booms, reversals and would discourage fundamental, long-term investment.
-Speculation would emerge from the evolution of collective sentiment and could create trends in financial asset prices.
 
-This paper distinguishes between the two concepts of speculation by investigating the relationship between the intensity of speculation and the weight of speculative positions relative to the movement of prices in the foreign exchange market\footnote{\label{FX}The foreign exchange market is chosen for this study because it is the largest and most liquid financial market in the world.
+This paper evaluates the relative importance of these two concepts of speculation by investigating the relationship between the intensity of speculation and the weight of speculative positions relative to subsequent movement of foreign exchange prices.
+\footnote{\label{FX}The foreign exchange market is chosen for this study because it is the largest and most liquid financial market in the world.
 The latest survey taken by the Bank for International Settlements (BIS) in April 2013 estimated average daily turnover at \$5.4trn with average daily spot transactions of \$2.0trn See \citet{BISFX2013}.
@@
 This would be the cost for a round-trip (buying and selling immediately).}.
-If speculators are uninformed and their activity can be regarded as noise, the most extreme speculative sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when reversals are most likely to take place; if speculators are informed their activity drives exchange rates towards equlibrium and prices should follow a random walk after these extremes.
-
-The price action after periods of extreme speculative sentiment and extreme speculative positions are comparded to the results of a noise-trader model.
-Noise-trader reversals in price are not evident.
-However, there is evidence that speculation in the foreign exchange market is informed.
-The paper contributes to the understanding of speculation and uses a unique set of option prices to identify the intensity of speculative sentiment
+If speculators are uninformed and their activity can be regarded as noise, the most extreme speculative sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when reversals are most likely to take place; if speculators are informed their activity drives exchange rates towards equilibrium and prices should follow a random walk after these extremes.
 
@@
 
-\citet[p.529]{BlackNoise} suggested that the interaction of informed and uninformed activity was essential to the smooth functioning of financial markets.
+\citet[p.529]{BlackNoise} argued that the interaction of informed and uninformed activity was essential to the smooth functioning of financial markets.
 Noise, he argued, was the  ``the arbitrary element in expectations''.
@@
 Uninformed activity provides a rationale and the opportunity for trading.
-Unless there is a difference of opinion about value, there is no reason to trade.
+Unless there is a difference of opinion about value, there is no reason to exchange.
 Uninformed speculation can create space between price and fundamental value that can be exploited by informed speculators.
@@
 
-\citet{EasleyPIN} use information as an explanation for the effect of large orders on stock prices.
-The \emph{order effect} may be a consequence of the need to adjust inventories when a market-maker receives a large order; that effect may also be a function of the market-maker knowledge that large orders are more likely come from to informed traders seeking to realise their informational advantage.
-\citet{EasleyPIN2} develop a technique to estimate the probability that trading is informed.
-They find that the high liquidity stocks have the highest level of information and the most frequent arrival rate of informed traders but that price changes are limited by the volume of uniformed traders.
-Low liquidity stocks by contrast, have fewer information events but informed traders have a much larger effect on prices.
-
-\citet{EasleyPIN2} propose to measure informed market activity with the bid-ask spread.
-Large orders are assumed to be informed as participants seek to explit their informational advantage.\footnote{Market-makers use Bayes' Rule to update their fundamental valuation by using order flow to estimate the joint probability that there has been an information event and that it has been good news or bad news, the authors seek to operationalise the following model:
-\begin{subequations}
-\begin{equation}
-b(t) = E[V_i|t] - \frac{\mu P_b(t)}{\varepsilon + \mu P_b(t)}(E[V-t|-\underline{V}])
-\end{equation}
-\begin{equation}
-a(t) = E[V_i|t] + \frac{\mu P_g(t)}{\varepsilon + \mu P_g(t)}(\overline{V}E+ [V-t|])
-\end{equation}
-\end{subequations}
-  where $b(t)$ is the bid price at time $t$; $V$ is the value of the security; $\mu$ is the share of informed traders in the market; $P_g(t)$ is the probability that there has been some good news and $P_b(t)$ is the probability that there has been some bad news; $\varepsilon$ is the share of uninformed traders in the market; $\underline{V}$ is the value if there is bad news and $\overline{V}$ is the value if there is good news.}
-
-However, identifying large orders and making an assessment of the level of informed trading is much harder in an over-the-counter market like foreign exchange.
-While electronic order books have become more widely used, much of the market-making still takes place through bilateral agreements and information about the nature of transactions including size and origin of orders is not widely available.\footnote{\citet{Lyons1995Microstructure} and \citet{Evans2002Order} have gained access to proprietary data and have used the initiation of orders as an explanatory variable in an equation to explain the evolution of major exchange rates.
-They find that $\$$ 1 billion of net US dollar purchase moves the Deutsche mark price of the US unit by 0.5 percent.
-They argue that the orders crystallise a range of information that is used to value currencies.}  \citet{LeeandReady} seek to use price action to infer the nature of the order that has been placed into the market.
-The \emph{tick-test} has four classifications: the up-tick, the down-tick, the zero-up-tick and the zero-down-tick.
-A trade is an up-tick if the price is higher than the previous trade.
-If the price is the same as the previous trade, if the previous trade was classified as an up-tick, this trade is a zero-up-tick.
-Under this system, the up-ticks are assumed to be buy orders and the down-ticks are sell orders.
-These estimates of the market activity can be used to estimate the level of informed trading.
-
-Informed speculation is also central to the mechanism proposed by \citet{muth1961Rational} to facilitate \emph{rational expectations}.
-Those economic agents able to determine fundamental value will be able to speculate and drive prices towards equilibrium.
-While this construction may be justified for the analysis of microstructure and market efficiency, it is an over-simplification that is inappropriate when seeking more understanding of speculation itself.
-Institutional and psychological explanations for noise-trading frequently suggest that there are endogenous non-random elements to speculative activity.
-
-Noise-trading caused by psychological biases and restricted information is associated with behavioural issues, over-confidence and the use of irrational or purely fashionable investment tools.
-See \citet{DeBondtOver} for an analysis of over-reaction and the \emph{representativeness heuristic}, \citet{OdeanOC} for a study of over-confidence and  \citet{ShillerFashion} for an assessment of investment fads.
-Traders without information may not make decisions independently of others.
-Information may be inferred (perhaps incorrectly) from the buying and selling conducted by other traders, causing positive feedback and trends.\footnote{This is an example of the beauty contest that has been presented by \citet[p.101]{Keynes1936}.
-The aim is to determine not the underlying value but what others perceive to be the underlying value.}
-
-Liquidity trading suggests that institutional constraints may ensure that economic agents just execute customer orders at the current price or that they buy and sell with reference to a benchmark index or to fulfil a hedging strategy.
-The institutional explanations are also associated with the allocation of credit for investment and trading purposes:  as asset values decline exchanges, banks and other providers of liquidity require larger margins and deeper haircuts; investors liquidate their fund holdings; the attraction of hedging increases.
-See \citet{BrunnermeierLiquidity} and \citet{Gorton} for some discussion of the way that margin-calls and collateral constraints can force investment decisions.
-
-\citet{FriedmanPositive} said that the informed market participants would drive the uninformed from the market.
-However, a large proportion of uninformed traders in a market can undermine this process.
-\citet{Delong1990noise} and \citet{shleifer1997limits} have presented models where the risk that noise-traders will move prices even further away from equilibrium before convergence can limit the activity of informed speculators when they are risk averse.
-This can be called \emph{noise-trader risk}.
-
 \section{Measuring speculation}
-The measurement of speculative intensity must identify how far speculators believe that the market will move in a particular direction.
-In the De-Long model, this would mean that for Equation \ref{eq:NP2}, it is the absolute value of $\rho - 1$.
-In the absence of a daily survey of speculators, this study uses a unique series of option\emph{risk-reversal skew} as an objective measure of speculative sentiment.\footnote{It appears that this is the only period for which the data currently exist.
+This study uses a unique series of option\emph{risk-reversal skew} as an objective measure of speculative sentiment.\footnote{It appears that this is the only period for which the data currently exist.
 The risk-reversal data is very difficult to obtain in an over-the-counter market like foreign exchange where there is no centralised exchange to record and collect information, generally being the property of quoting institutions.
@@
 Risk-reversal skew is the relative price of currency put and call options for the same strike price.} The risk-reversal skew is the relationship between a call option and the equivalent put.
-The \emph{Put-Call-Parity} in the  \citet{Black1973Options} option pricing model assumes that these will be the same, in practice they are not.
-While the Black-Scholes model is the basis for valuing options where the price of a non-dividend paying stock trades at a discount to the forward rate equal to the risk free rate, the forward rate in a currency option is determined by the relative interest rate differential between the two currencies.
-The \citet{GarmanKohlhagan} adjustment that accounts for the difference in interest payments on the two currencies is used.
-The spot discount is determined therefore by the relative interest rate differential for the appropriate maturity.
-This model allows the market practice of pricing options from the volatility implied by the model.
-In other words, the Garman-Kohlhagan is used to extract the volatility that is implied by a specific foreign exchange option price given the strike and other parameters.
 
@@
 See \citet{Dolly1933} for the original version of the event study method and \citet{FamaFisherJensenRoll} for an overview of the method used in this study.
-However, expected returns are identified as being zero rather than established through a \emph{market-model}.
-The R package \emph{eventstudies} created by \citet{eventstudies} is used to manage the data.
-Extreme measures of speculative sentiment and activity are the event and the cumulative abnormal returns for the whole of the event window (CARW) and the cumulative abnormal returns for after the event (CARA) are the exchange rates to be assessed.
 
@@
   Table \ref{tabref:RR1} records the result of the event studies that were carried out on extreme risk reversal.
-For each exchange rate the second column distinguishes between extreme high and extreme low, the third  column identifies the number of extreme events that were recorded and the following two are the cumulative average returns for the whole event window (CARW) and the cumulative average returns for the event and the period after (CARA).
+For each exchange rate the second column distinguishes between extreme high and extreme low risk-reversals, the third  column identifies the number of extreme events that were recorded while the following two columns show the cumulative average returns for the whole event window (CARW) and the cumulative average returns for the event and the period after (CARA).
 The windows were chosen to correspond to one day and one week of trading.
@@
 In most instances these are significantly different from zero.
-Significance is identified as those cases where more than 95\% of 1000 means calculated from random bootstrap samples taken from the events are above zero.
-Returns that are statistically different from zero are identified by an asterisk.
 

### 2017-11-30 Knitr/FinancialEconomics.tex
Updated type to article.
Updated and corrected.

@@
-\documentclass[preprint,12pt,authoryear]{elsarticle}
+\documentclass{article}
 %\graphicspath{{./Figures/}} % Specifies the directory where pictures are stored
@@
 \usepackage[flushleft]{threeparttable}
+\usepackage[comma, sort&compress]{natbib}% Use the natbib reference package - read up on this to edit the reference style; if you want text (e.g.
+Smith et al., 2012) for the in-text references (instead of numbers), remove 'numbers'
 \usepackage{multirow}
@@
 \usepackage[colorlinks = true, citecolor = blue, linkcolor = blue]{hyperref}
+\title{Foreign Exchange Speculation: An event study}
+\author{Rob Hayward}
+\date{Dec 2017}
 \begin{document}
@@
 If speculation is noise,  extreme sentiment and extreme positions should lead to overshooting and increase risk of subsequent reversals.
-The finding that speculative extremes do not provide information about subsequent returns implies that speculation is part of the process of price discovery and that efforts to reduce it would leave to a reduction in information efficiency.
+The finding that speculative extremes do not provide information about subsequent returns implies that speculation is part of the process of price discovery and that efforts to reduce it would leave to a reduction in information efficiency of financial markets.
 
@@
 If speculation is based on information it is a stabilising force, providing liquidity and helping to ensure that markets swiftly find equilibrium.
-\citet{KeynesHedge} and \citet{HicksHedge} emphasised the role that speculation played in providing liquidity.
-Speculators provide a service for a fee that is determined by the return on their activity.
-Informed speculators will tend to drive out the uninformed noise-traders \citet{FriedmanPositive}.
-Successful speculators would be those buying below intrinsic value and selling above that rate.
-However, the later Keynes regarded speculation as a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd'\citet[p.
-101]{Keynes1936}.
+\citet{KeynesHedge} and \citet{HicksHedge} for example emphasised the role that speculation plays in facilitating financial market transactions.
+Speculators offer a service for a fee that is determined by the return on their activity.
+While speculators may be trading on information or noise, informed speculators will tend to drive out the uninformed noise-traders according to \citet{FriedmanPositive}.
+Nonetheless, the later Keynes regarded speculation as a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd'\citet[p.
+101]{Keynes1936} and \citet{Delong1990noise} provides a model of cases where speculative noise provides a constrain on informed traders benefiting from their knowledge.
 
-This paper evaluates the relative importance of these two concepts of speculation by investigating the relationship between the intensity of speculation and the weight of speculative positions relative to subsequent movement of foreign exchange prices.
-\footnote{\label{FX}The foreign exchange market is chosen for this study because it is the largest and most liquid financial market in the world.
+This paper evaluates the relative importance of these two concepts of speculation by investigating the relationship between the intensity of speculation and the weight of speculative positions relative to subsequent movement of foreign exchange prices.\footnote{\label{FX}The foreign exchange market is chosen for this study because it is the largest and most liquid financial market in the world.
 The latest survey taken by the Bank for International Settlements (BIS) in April 2013 estimated average daily turnover at \$5.4trn with average daily spot transactions of \$2.0trn See \citet{BISFX2013}.
@@
 
-The rest of the paper proceeds as follows:  Section Two discusses the distinction between informed and uninformed speculators; Section Three talks about the measurement of speculation and the event study method; Section Four reviews the evidence; Section Five concludes.
+The rest of the paper proceeds as follows:  Section Two presents the framework of informed and uninformed speculation; Section Three talks about the measurement of speculation and the event study method; Section Four reviews the evidence; Section Five concludes.
 
@@
 
-\citet[p.529]{BlackNoise} argued that the interaction of informed and uninformed activity was essential to the smooth functioning of financial markets.
+\citet[p.529]{BlackNoise} asserted that the interaction of informed and uninformed activity was essential to the smooth functioning of financial markets.
 Noise, he argued, was the  ``the arbitrary element in expectations''.
@@
 Uninformed speculation can create space between price and fundamental value that can be exploited by informed speculators.
-This space can encourage informed traders to pay the cost of acquiring information.
-\citet{Grossman1980Impossibility} argue that there is equilibrium between the cost of obtaining information and the benefits that are expected to come from the use of the information.
+This space can encourage informed traders to pay the cost of acquiring information in the \citet{Grossman1980Impossibility} model.
+There is equilibrium between the cost of obtaining information and the benefits that are expected to come from the use of the information.
 
@@
 Risk-reversal skew is the relative price of currency put and call options for the same strike price.} The risk-reversal skew is the relationship between a call option and the equivalent put.
-
 If the exchange rate is expected to follow a random walk with drift towards the 1 month forward, implied volatility on the put and the call for the same delta should be equal.
@@
 The effect of extreme speculation on future returns is analysed with an event study.
+The events here are the extremes of speculative sentiment and speculative positions.
 See \citet{Dolly1933} for the original version of the event study method and \citet{FamaFisherJensenRoll} for an overview of the method used in this study.
@@
 
-\section*{References}
 \bibliography{../../myrefs}

### 2017-12-06 MDPI/MDPI_template/FinStudies.tex
This is the Financial Studies update

--- 2017-11-30 Knitr/FinancialEconomics.tex
@@
-\documentclass{article}
-%\graphicspath{{./Figures/}} % Specifies the directory where pictures are stored
-%\usepackage[dcucite]{harvard}
-% \documentclass[final,1p,times,authoryear]{elsarticle}
-%\documentclass[final,1p,times,twocolumn,authoryear]{elsarticle}
-%\documentclass[final,3p,times,authoryear]{elsarticle}
-%\documentclass[final,3p,times,twocolumn,authoryear]{elsarticle}
-%\documentclass[final,5p,times,authoryear]{elsarticle}
-%\documentclass[final,5p,times,twocolumn,authoryear]{elsarticle}
+%  LaTeX support: latex@mdpi.com 
+%  In case you need support, please attach all files that are necessary for compiling as well as the log file, and specify the details of your LaTeX setup (which operating system and LaTeX version / tools you are using).
+
+% You need to save the "mdpi.cls" and "mdpi.bst" files into the same folder as this template file.
+
+%=================================================================
+\documentclass[ijfs,article,submit,oneauthor,pdftex,10pt,a4paper]{mdpi} 
+%
+%--------------------
+% Class Options:
+%--------------------
+% journal
+%----------
+% Choose between the following MDPI journals:
+%ijfs, 
+%------
+% article
+%---------
+% The default type of manuscript is article, but can be replaced by: 
+% abstract, addendum, article, benchmark, book, bookreview, briefreport, casereport, changes, comment, commentary, communication, conceptpaper, correction, conferenceproceedings, conferencereport, expressionofconcern, meetingreport, creative, datadescriptor, discussion, editorial, essay, erratum, hypothesis, interestingimages, letter, meetingreport, newbookreceived, opinion, obituary, projectreport, reply, reprint, retraction, review, perspective, protocol, shortnote, supfile, technicalnote, viewpoint
+% supfile = supplementary materials
+% protocol: If you are preparing a "Protocol" paper, please refer to http://www.mdpi.com/journal/mps/instructions for details on its expected structure and content.
+%----------
+%submit
+%----------
+% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted.
+This will only make changes to the frontpage (e.g.
+the logo of the journal will get visible), the headings, and the copyright information.
+Also, line numbering will be removed.
+Journal info and pagination for accepted papers will also be assigned by the Editorial Office.
+%------------------
+% moreauthors
+%------------------
+% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.
+%---------
+% pdftex
+%---------
+% The option pdftex is for use with pdfLaTeX. If eps figures are used, remove the option pdftex and use LaTeX and dvi2pdf.
+
+%=================================================================
+\firstpage{1} 
+\makeatletter 
+\setcounter{page}{\@firstpage} 
+\makeatother 
+\articlenumber{x}
+\doinum{10.3390/------}
+\pubvolume{xx}
+\pubyear{2017}
+\copyrightyear{2017}
+\externaleditor{Academic Editor: name}
+\history{Received: date; Accepted: date; Published: date}
+
+%------------------------------------------------------------------
+% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
+%\pdfoutput=1
+
+%=================================================================
+% Add packages and commands here.
+The following packages are loaded in our class file: fontenc, calc, indentfirst, fancyhdr, graphicx, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, amsthm, hyphenat, natbib, hyperref, footmisc, geometry, caption, url, mdframed, tabto, soul, multirow, microtype, tikz
 \usepackage{rotating}
-\usepackage{amsmath}
-%\usepackage{setspace}
 \usepackage{pdflscape}
 \usepackage[flushleft]{threeparttable}
-\usepackage[comma, sort&compress]{natbib}% Use the natbib reference package - read up on this to edit the reference style; if you want text (e.g.
+%\usepackage[comma, sort&compress]{natbib}% Use the natbib reference package - read up on this to edit the reference style; if you want text (e.g.
 Smith et al., 2012) for the in-text references (instead of numbers), remove 'numbers'
-\usepackage{multirow}
-\bibliographystyle{agsm}
-\usepackage[colorlinks = true, citecolor = blue, linkcolor = blue]{hyperref}
-\title{Foreign Exchange Speculation: An event study}
-\author{Rob Hayward}
-\date{Dec 2017}
-\begin{document}
-\maketitle
-\begin{abstract}
-Does speculation facilitate price discovery or instability?
+%=================================================================
+%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition
+%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).
+
+%=================================================================
+% Full title of the paper (Capitalized)
+\Title{Foreign Exchange Speculation: An Event Study}
+
+% Author Orchid ID: enter ID or remove command
+%\newcommand{\orcidauthorA}{0000-0000-000-000X} % Add \orcidA{} behind the author's name
+%\newcommand{\orcidauthorB}{0000-0000-000-000X} % Add \orcidB{} behind the author's name
+
+% Authors, for the paper (add full first names)
+\Author{Rob Hayward $^{1,\dagger,\ddagger}$}
+
+% Authors, for metadata in PDF
+\AuthorNames{Rob Hayward}
+
+% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
+\address{$^{1}$ \quad University of Brighton; rh49@brighton.ac.uk}
+
+% Contact information of the corresponding author
+\corres{Correspondence: rh49@brighton.ac.uk; Tel.: +44-1273-642-586}
+
+% Current address and/or shared authorship
+\firstnote{Current address: Lewes Road, Brighton BN2 4AT} 
+%\secondnote{These authors contributed equally to this work.}
+% The commands \thirdnote{} till \eighthnote{} are available for further notes
+
+% Simple summary
+%\simplesumm{}
+
+% Abstract (Do not insert blank lines, i.e. \\) 
+\abstract{Does speculation facilitate price discovery or instability?
 If it is price discovery, it is beneficial and should be encouraged; if it is instability, welfare is enhanced by its reduction.
@@
 If speculation is noise,  extreme sentiment and extreme positions should lead to overshooting and increase risk of subsequent reversals.
-The finding that speculative extremes do not provide information about subsequent returns implies that speculation is part of the process of price discovery and that efforts to reduce it would leave to a reduction in information efficiency of financial markets.
+The finding that speculative extremes do not provide information about subsequent returns implies that speculation is part of the process of price discovery and that efforts to reduce it would reduce the informational efficiency of financial markets.}
 
-\end{abstract}
+% Keywords
+\keyword{Foreign Exchange ; Speculation; Event-study.}
 
+% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
+%\PACS{J0101}
+%\MSC{}
+%\JEL{}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Only for the journal Applied Sciences:
+%\featuredapplication{Authors are encouraged to provide a concise description of the specific application or a potential application of the work. This section is not mandatory.}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Only for the journal Data:
+%\dataset{DOI number or link to the deposited data set in cases where the data set is published or set to be published separately.
+If the data set is submitted and will be published as a supplement to this paper in the journal Data, this field will be filled by the editors of the journal.
+In this case, please make sure to submit the data set as a supplement when entering your manuscript into our manuscript editorial system.}
+
+%\datasetlicense{license under which the data set is made available (CC0, CC-BY, CC-BY-SA, CC-BY-NC, etc.)}
+
+%\setcounter{secnumdepth}{4}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{document}
 \section{Introduction}
@@
 
-\bibliography{../../myrefs}
+%=====================================
+% References, variant B: external bibliography
+%=====================================
+\externalbibliography{yes}
+\bibliography{myrefs}
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% optional
+%\sampleavailability{Samples of the compounds ...... are available from the authors.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \end{document}

### 2017-12-06 MDPI/MDPI_template/template.tex
This is the Financial Studies update

%  LaTeX support: latex@mdpi.com 
%  In case you need support, please attach all files that are necessary for compiling as well as the log file, and specify the details of your LaTeX setup (which operating system and LaTeX version / tools you are using).

% You need to save the "mdpi.cls" and "mdpi.bst" files into the same folder as this template file.

%=================================================================
\documentclass[journal,article,submit,moreauthors,pdftex,10pt,a4paper]{mdpi} 
%
%--------------------
% Class Options:
%--------------------
% journal
%----------
% Choose between the following MDPI journals:
% actuators, addictions, admsci, aerospace, agriculture, agronomy, algorithms, animals, antibiotics, antibodies, antioxidants, applsci, arts, asi, atmosphere, atoms, axioms, batteries, bdcc, behavsci, beverages, bioengineering, biology, biomedicines, biomimetics, biomolecules, biosensors, brainsci, buildings, carbon, cancers, catalysts, cells, ceramics, challenges, chemengineering, chemosensors, children, chromatography, climate, coatings, colloids, computation, computers, condensedmatter, cosmetics, cryptography, crystals, cybersecurity, data, dentistry, designs, diagnostics, diseases, diversity, econometrics, economies, education, electrochemistry, electronics, energies, entropy, environments, epigenomes, est, fermentation, fibers, fire, fishes, fluids, foods, forests, fractalfract, futureinternet, galaxies, games, gastrointestdisord, gels, genealogy, genes, geosciences, geriatrics, hazardousmatters, healthcare, heritage, highthroughput, horticulturae, humanities, hydrology, informatics, information, infrastructures, inorganics, insects, instruments, ijerph, ijfs, ijms, ijgi, ijtpp, inventions, j, jcdd, jcm, jcs, jdb, jfb, jfmk, jimaging, jof, jintelligence, jlpea, jmmp, jmse, jpm, jrfm, jsan, land, languages, laws, life, literature, logistics, lubricants, machines, magnetochemistry, make, marinedrugs, materials, mathematics, mca, mti, medsci, medicines, membranes, metabolites, metals, microarrays, micromachines, microorganisms, minerals, modelling, molbank, molecules, mps, nanomaterials, ncrna, neonatalscreening, nitrogen, nutrients, ohbm, particles, pathogens, pharmaceuticals, pharmaceutics, pharmacy, philosophies, photonics, plants, plasma, polymers, preprints, proceedings, processes, proteomes, publications, quaternary, qubs, recycling, religions, remotesensing, resources, risks, robotics, safety, scipharm, sensors, separations, sexes, sinusitis, socsci, societies, soils, sports, standards, surgeries, sustainability, symmetry, systems, technologies, toxics, toxins, tropicalmed, universe, urbansci, vaccines, vetsci, vibration, viruses, vision, water, wem
%---------
% article
%---------
% The default type of manuscript is article, but can be replaced by: 
% abstract, addendum, article, benchmark, book, bookreview, briefreport, casereport, changes, comment, commentary, communication, conceptpaper, correction, conferenceproceedings, conferencereport, expressionofconcern, meetingreport, creative, datadescriptor, discussion, editorial, essay, erratum, hypothesis, interestingimages, letter, meetingreport, newbookreceived, opinion, obituary, projectreport, reply, reprint, retraction, review, perspective, protocol, shortnote, supfile, technicalnote, viewpoint
% supfile = supplementary materials
% protocol: If you are preparing a "Protocol" paper, please refer to http://www.mdpi.com/journal/mps/instructions for details on its expected structure and content.
%----------
% submit
%----------
% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted. This will only make changes to the frontpage (e.g. the logo of the journal will get visible), the headings, and the copyright information. Also, line numbering will be removed. Journal info and pagination for accepted papers will also be assigned by the Editorial Office.
%------------------
% moreauthors
%------------------
% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.
%---------
% pdftex
%---------
% The option pdftex is for use with pdfLaTeX. If eps figures are used, remove the option pdftex and use LaTeX and dvi2pdf.

%=================================================================
\firstpage{1} 
\makeatletter 
\setcounter{page}{\@firstpage} 
\makeatother 
\articlenumber{x}
\doinum{10.3390/------}
\pubvolume{xx}
\pubyear{2017}
\copyrightyear{2017}
\externaleditor{Academic Editor: name}
\history{Received: date; Accepted: date; Published: date}

%------------------------------------------------------------------
% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
%\pdfoutput=1

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, calc, indentfirst, fancyhdr, graphicx, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, amsthm, hyphenat, natbib, hyperref, footmisc, geometry, caption, url, mdframed, tabto, soul, multirow, microtype, tikz

%=================================================================
%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

%=================================================================
% Full title of the paper (Capitalized)
\Title{Title}

% Author Orchid ID: enter ID or remove command
\newcommand{\orcidauthorA}{0000-0000-000-000X} % Add \orcidA{} behind the author's name
%\newcommand{\orcidauthorB}{0000-0000-000-000X} % Add \orcidB{} behind the author's name

% Authors, for the paper (add full first names)
\Author{Firstname Lastname $^{1,\dagger,\ddagger}$\orcidA{}, Firstname Lastname $^{1,\ddagger}$ and Firstname Lastname $^{2,}$*}

% Authors, for metadata in PDF
\AuthorNames{Firstname Lastname, Firstname Lastname and Firstname Lastname}

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad Affiliation 1; e-mail@e-mail.com\\
$^{2}$ \quad Affiliation 2; e-mail@e-mail.com}

% Contact information of the corresponding author
\corres{Correspondence: e-mail@e-mail.com; Tel.: +x-xxx-xxx-xxxx}

% Current address and/or shared authorship
\firstnote{Current address: Affiliation 3} 
\secondnote{These authors contributed equally to this work.}
% The commands \thirdnote{} till \eighthnote{} are available for further notes

% Simple summary
%\simplesumm{}

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{A single paragraph of about 200 words maximum. For research articles, abstracts should give a pertinent overview of the work. We strongly encourage authors to use the following style of structured abstracts, but without headings: 1) Background: Place the question addressed in a broad context and highlight the purpose of the study; 2) Methods: Describe briefly the main methods or treatments applied; 3) Results: Summarize the article's main findings; and 4) Conclusion: Indicate the main conclusions or interpretations. The abstract should be an objective representation of the article, it must not contain results which are not presented and substantiated in the main text and should not exaggerate the main conclusions.}

% Keywords
\keyword{keyword 1; keyword 2; keyword 3 (list three to ten pertinent keywords specific to the article, yet reasonably common within the subject discipline.)}

% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
%\PACS{J0101}
%\MSC{}
%\JEL{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Applied Sciences:
%\featuredapplication{Authors are encouraged to provide a concise description of the specific application or a potential application of the work. This section is not mandatory.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Data:
%\dataset{DOI number or link to the deposited data set in cases where the data set is published or set to be published separately. If the data set is submitted and will be published as a supplement to this paper in the journal Data, this field will be filled by the editors of the journal. In this case, please make sure to submit the data set as a supplement when entering your manuscript into our manuscript editorial system.}

%\datasetlicense{license under which the data set is made available (CC0, CC-BY, CC-BY-SA, CC-BY-NC, etc.)}

%\setcounter{secnumdepth}{4}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Only for the journal Gels: Please place the Experimental Section after the Conclusions

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\setcounter{section}{-1} %% Remove this when starting to work on the template.
\section{How to Use this Template}
The template details the sections that can be used in a manuscript. Note that the order and names of article sections may differ from the requirements of the journal (e.g. the positioning of the Materials and Methods section). Please check the instructions for authors page of the journal to verify the correct order and names. For any questions, please contact the editorial office of the journal or support@mdpi.com. For LaTeX related questions please contact Janine Daum at latex-support@mdpi.com.

\section{Introduction}

The introduction should briefly place the study in a broad context and highlight why it is important. It should define the purpose of the work and its significance. The current state of the research field should be reviewed carefully and key publications cited. Please highlight controversial and diverging hypotheses when necessary. Finally, briefly mention the main aim of the work and highlight the principal conclusions. As far as possible, please keep the introduction comprehensible to scientists outside your particular field of research. Citing a journal paper \cite{ref-journal}. And now citing a book reference \cite{ref-book}. Please use the command \citep{ref-journal} for the following MDPI journals, which use author-date citation: Administrative Sciences, Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, JRFM, Laws, Religions, Risks, Social Sciences.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}

This section may be divided by subheadings. It should provide a concise and precise description of the experimental results, their interpretation as well as the experimental conclusions that can be drawn.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Subsection}

\subsubsection{Subsubsection}

Bulleted lists look like this:
\begin{itemize}[leftmargin=*,labelsep=5.8mm]
\item	First bullet
\item	Second bullet
\item	Third bullet
\end{itemize}

Numbered lists can be added as follows:
\begin{enumerate}[leftmargin=*,labelsep=4.9mm]
\item	First item 
\item	Second item
\item	Third item
\end{enumerate}

The text continues here.

\subsection{Figures, Tables and Schemes}

All figures and tables should be cited in the main text as Figure 1, Table 1, etc.

\begin{figure}[H]
\centering
\includegraphics[width=3 cm]{logo-mdpi}
\caption{This is a figure, Schemes follow the same formatting. If there are multiple panels, they should be listed as: (\textbf{a}) Description of what is contained in the first panel. (\textbf{b}) Description of what is contained in the second panel. Figures should be placed in the main text near to the first time they are cited. A caption on a single line should be centered.}
\end{figure}

\begin{table}[H]
\caption{This is a table caption. Tables should be placed in the main text near to the first time they are cited.}
\centering
%% \tablesize{} %% You can specify the fontsize here, e.g.  \tablesize{\footnotesize}. If commented out \small will be used.
\begin{tabular}{ccc}
\toprule
\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
\midrule
entry 1		& data			& data\\
entry 2		& data			& data\\
\bottomrule
\end{tabular}
\end{table}

\subsection{Formatting of Mathematical Components}

This is an example of an equation:

\begin{equation}
\mathbb{S}
\end{equation}

%% If the documentclass option "submit" is chosen, please insert a blank line before and after any math environment (equation and eqnarray environments). This ensures correct linenumbering. The blank line should be removed when the documentclass option is changed to "accept" because the text following an equation should not be a new paragraph. 
Please punctuate equations as regular text. Theorem-type environments (including propositions, lemmas, corollaries etc.) can be formatted as follows:
%% Example of a theorem:
\begin{Theorem}
Example text of a theorem.
\end{Theorem}

The text continues here. Proofs must be formatted as follows:

%% Example of a proof:
\begin{proof}[Proof of Theorem 1]
Text of the proof. Note that the phrase `of Theorem 1' is optional if it is clear which theorem is being referred to.
\end{proof}
The text continues here.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}

Authors should discuss the results and how they can be interpreted in perspective of previous studies and of the working hypotheses. The findings and their implications should be discussed in the broadest context possible. Future research directions may also be highlighted.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Materials and Methods}

Materials and Methods should be described with sufficient details to allow others to replicate and build on published results. Please note that publication of your manuscript implicates that you must make all materials, data, computer code, and protocols associated with the publication available to readers. Please disclose at the submission stage any restrictions on the availability of materials or information. New methods and protocols should be described in detail while well-established methods can be briefly described and appropriately cited.

Research manuscripts reporting large datasets that are deposited in a publicly available database should specify where the data have been deposited and provide the relevant accession numbers. If the accession numbers have not yet been obtained at the time of submission, please state that they will be provided during review. They must be provided prior to publication.

Interventionary studies involving animals or humans, and other studies require ethical approval must list the authority that provided approval and the corresponding ethical approval code.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}

This section is not mandatory, but can be added to the manuscript if the discussion is unusually long or complex.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Patents}
This section is not mandatory, but may be added if there are patents resulting from the work reported in this manuscript.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{6pt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
\supplementary{The following are available online at www.mdpi.com/link, Figure S1: title, Table S1: title, Video S1: title.}

% Only for the journal Methods and Protocols:
% If you wish to submit a video article, please do so with any other supplementary material.
% \supplementary{The following are available at www.mdpi.com/link: Figure S1: title, Table S1: title, Video S1: title. A supporting video article is available at doi: link.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\acknowledgments{All sources of funding of the study should be disclosed. Please clearly indicate grants that you have received in support of your research work. Clearly state if you received funds for covering the costs to publish in open access.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\authorcontributions{For research articles with several authors, a short paragraph specifying their individual contributions must be provided. The following statements should be used ``X.X. and Y.Y. conceived and designed the experiments; X.X. performed the experiments; X.X. and Y.Y. analyzed the data; W.W. contributed reagents/materials/analysis tools; Y.Y. wrote the paper.'' Authorship must be limited to those who have contributed substantially to the work reported.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\conflictsofinterest{Declare conflicts of interest or state ``The authors declare no conflict of interest.'' Authors must identify and declare any personal circumstances or interest that may be perceived as inappropriately influencing the representation or interpretation of reported research results. Any role of the funding sponsors in the design of the study; in the collection, analyses or interpretation of data; in the writing of the manuscript, or in the decision to publish the results must be declared in this section. If there is no role, please state ``The founding sponsors had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript, and in the decision to publish the results''.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
\abbreviations{The following abbreviations are used in this manuscript:\\

\noindent 
\begin{tabular}{@{}ll}
MDPI & Multidisciplinary Digital Publishing Institute\\
DOAJ & Directory of open access journals\\
TLA & Three letter acronym\\
LD & linear dichroism
\end{tabular}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
\appendixtitles{no} %Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
\appendixsections{multiple} %Leave argument "multiple" if there are multiple sections. Then a counter is printed ("Appendix A"). If there is only one appendix section then change the argument to "one" and no counter is printed ("Appendix").
\appendix
\section{}
\subsection{}
The appendix is an optional section that can contain details and data supplemental to the main text. For example, explanations of experimental details that would disrupt the flow of the main text, but nonetheless remain crucial to understanding and reproducing the research shown; figures of replicates for experiments of which representative data is shown in the main text can be added here if brief, or as Supplementary data. Mathematical proofs of results not central to the paper can be added as an appendix.

\section{}
All appendix sections must be cited in the main text. In the appendixes, Figures, Tables, etc. should be labeled starting with `A', e.g., Figure A1, Figure A2, etc. 
This is the reference - \cite{BernnakeGertlerAgency}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Citations and References in Supplementary files are permitted provided that they also appear in the reference list here.

%=====================================
% References, variant A: internal bibliography
%=====================================
%\reftitle{References}
%\begin{thebibliography}{999}
% Reference 1
%\bibitem[Author1(year)]{ref-journal}
%Author1, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142-149, DOI.
% Reference 2
%\bibitem[Author2(year)]{ref-book}
%Author2, L. The title of the cited contribution. In {\em The Book Title}; Editor1, F., Editor2, A., Eds.; Publishing House: City, Country, 2007; pp. 32-58, ISBN.
%\end{thebibliography}

% The following MDPI journals use author-date citation: Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, JRFM, Laws, Religions, Risks, Social Sciences. For those journals, please follow the formatting guidelines on http://www.mdpi.com/authors/references
% To cite two works by the same author: \citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a}, \citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
% To cite two works by the same author with specific pages: \citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p. 328; \citeyear{ref-journal-3b}, p.475). This produces: Wong (1999, p. 328; 2000, p. 475)

%=====================================
% References, variant B: external bibliography
%=====================================
\externalbibliography{yes}
\bibliography{myrefs}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
\sampleavailability{Samples of the compounds ...... are available from the authors.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}

### 2017-12-08 MDPI/MDPI_template/FinStudies.tex
Updated the files for the publication.

@@
 If speculation is based on information it is a stabilising force, providing liquidity and helping to ensure that markets swiftly find equilibrium.
-\citet{KeynesHedge} and \citet{HicksHedge} for example emphasised the role that speculation plays in facilitating financial market transactions.
+\citet{KeynesHedge} and \citet{HicksHedge} for example emphasise the role that speculation plays in facilitating financial market transactions.
 Speculators offer a service for a fee that is determined by the return on their activity.
 While speculators may be trading on information or noise, informed speculators will tend to drive out the uninformed noise-traders according to \citet{FriedmanPositive}.
-Nonetheless, the later Keynes regarded speculation as a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd'\citet[p.
-101]{Keynes1936} and \citet{Delong1990noise} provides a model of cases where speculative noise provides a constrain on informed traders benefiting from their knowledge.
+However, the later \citet[p.
+101]{Keynes1936} regarded speculation as a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd' and \citet{Delong1990noise} provides a model of cases where speculative noise provides a constrain on informed traders benefiting from their knowledge.
+Now speculation is an uninformed and destabilising force.
 
 This paper evaluates the relative importance of these two concepts of speculation by investigating the relationship between the intensity of speculation and the weight of speculative positions relative to subsequent movement of foreign exchange prices.\footnote{\label{FX}The foreign exchange market is chosen for this study because it is the largest and most liquid financial market in the world.
-The latest survey taken by the Bank for International Settlements (BIS) in April 2013 estimated average daily turnover at \$5.4trn with average daily spot transactions of \$2.0trn See \citet{BISFX2013}.
+The latest survey taken by the Bank for International Settlements (BIS) in April 2016 estimated average daily turnover at \$5.1trn with average daily spot transactions of \$1.7trn See \citet{BISFX2016}.
 The cost of trading in this market, when measured in terms of the size of the bid-ask spread, is extremely low.
 An analysis by \citet{Steely2013} of the bid-ask spreads for USD-JPY, GBP-USD and EUR-USD for the period January 2001 to 2005 finds that the average spread for each is 0.342\%, 0.305\% and 0.326\% respectively.
-This would be the cost for a round-trip (buying and selling immediately).}.
-If speculators are uninformed and their activity can be regarded as noise, the most extreme speculative sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when reversals are most likely to take place; if speculators are informed their activity drives exchange rates towards equilibrium and prices should follow a random walk after these extremes.
+This would be the cost for a round-trip (buying and selling immediately).}  If speculators are informed their activity drives exchange rates towards equilibrium and prices should follow a random walk after these extremes; if speculators are uninformed and their activity can be regarded as noise, the most extreme speculative sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when reversals are most likely to take place.
 
@@
 However, there is no clear, unambiguous theory of speculation.
-Speculation is frequently associated with the term \emph{trading}, but trading is broader as it encompasses market-making and financial investment.
+Speculation is frequently associated with the term trading and the provision of liquidity, but trading is broader as it encompasses financial investment while liquidity provision is really market-making.
 Speculation and trading can be categorised as being informed or uninformed.
@@
 The comparison of informed and uninformed traders is a central component of microstructure research.
-The focus of the \citet{Kyle1985Continuous} model of market-making is the interaction of orders from the informed and the uninformed.
-On the assumption that uniformed orders arrive in a random fashion, the weight of orders reveals the direction of private information and should led to adjustment of the bid-ask spread as part of the process of price discovery.
-Similarly,  \citet{glosten1985bid} describes a model where the bid-ask spread is a function of adverse selection, being caused by the risk that orders come from informed traders.
+At the heart of the \citet{Kyle1985Continuous} model of market-making is the interaction of orders from the informed and the uninformed.
+On the assumption that uniformed orders arrive in a random fashion, the weight of orders reveals the direction of private information and should encourage adjustment of the bid-ask spread as part of the process of price discovery.
+\citet{glosten1985bid} presents a model where the bid-ask spread is a function of adverse selection, being caused by the risk that orders come from informed traders.
 
@@
 \section{Measuring speculation}
-This study uses a unique series of option\emph{risk-reversal skew} as an objective measure of speculative sentiment.\footnote{It appears that this is the only period for which the data currently exist.
+This study uses a unique series of option \emph{risk-reversal skews} to measure of speculative sentiment\footnote{It appears that this is the only period for which the data currently exist.
 The risk-reversal data is very difficult to obtain in an over-the-counter market like foreign exchange where there is no centralised exchange to record and collect information, generally being the property of quoting institutions.
@@
 The risk-reversals are the spread of implied volatility for a 1 month 25 delta at-the-money call option relative to the equivalent put.
-Risk-reversal skew is the relative price of currency put and call options for the same strike price.} The risk-reversal skew is the relationship between a call option and the equivalent put.
+Risk-reversal skew is the relative price of currency put and call options for the same strike price.}, as well as CFTC data on speculative positions.
+The risk-reversal skew is the relationship between a call option and the equivalent put.
 If the exchange rate is expected to follow a random walk with drift towards the 1 month forward, implied volatility on the put and the call for the same delta should be equal.
@@
 See \citet{Global} for a technical review.
-}  Therefore, the relationship between implied volatility for calls and implied volatilities for puts for the same delta and maturity is an indication of how much more expensive calls are relative to the equivalent put or as the market bias towards puts over calls and as an indication of market sentiment.
+}  Therefore, the relationship between implied volatility for calls and implied volatilities for puts for the same delta and maturity strike is an indication of how much more expensive calls are relative to the equivalent put or as the market bias towards puts over calls and as an indication of market sentiment or related market positioning.
 The larger the absolute level of the risk-reversal skew, the greater the relative price and, if the deviation is based on information, subsequent price action should be zero on average; if the deviation is based on noise, this movement should be reversed.
@@
 
-The effect of extreme speculation on future returns is analysed with an event study.
-The events here are the extremes of speculative sentiment and speculative positions.
+An event study will analyse the performance of security of asset prices around an event.
+In this case the event is the extreme of speculative sentiment or extreme speculative activity.
 See \citet{Dolly1933} for the original version of the event study method and \citet{FamaFisherJensenRoll} for an overview of the method used in this study.
@@
 Extremes seem to be followed by a random walk as would be the case if the extreme were the result of information being absorbed into the price.
+Figures \ref{fig:ES5} and \ref{fig:ES5} show details of a selection of the event studies.
+Figure \ref{fig:ES5} is the extreme high ($99^{th}$ percentile) with a 16 day window.
+There is no evidence of a reversal in any of the currency units.
+Figure \ref{fig:ES5} is the extreme low ($10^{th}$ percentile) with a 16 day event window.
+Only the CAD shows any evidence of reversal.
+
 \begin{figure}
@@
 \label{fig:ES2}
-\end{figure}
-
-\begin{figure}
-\graphicspath{{../Figures/}}
-\centering
-\includegraphics[scale=0.8]{RRCum16}
-\caption{Event Study: Extreme (High $99{th}$ percentile) Risk Reversal Skew and 16 day event window: Risk reversals are the ratio of implied volatility on calls to the equivalent put; extreme high is at or above the $99^{th}$ percentile for the whole range, cumulative returns are the sum of log exchange rate returns.
-The event day is the day of the extreme risk reversal reading.
-If speculation is uninformed, extremes should be followed by reversals; if speculation is informed, extremes are more likely to be followed by a random walk.
-The solid line is the cumulative mean return from the start of the event window; the dashed lines show the 95\% confidence intervals constructed from 1000 random bootstrap samples of appropriate event window day.}
-\label{fig:ES1}
 \end{figure}
@@
 The price action of the event study is consistent with speculation being informed rather than uninformed.
+Figures \ref{fig:ES3} and \ref{fig:ES4} show a selection of event studies that were carried out.
+Figure \ref{fig:ES3} is the extreme low ($10^{th}$ percentile) with a 13 week window.
+There is no evidence of a reversal in any of the currency units.
+Figure \ref{fig:ES4} is the extreme high ($90^{th}$ percentile) with a 13 week event window.
+Only the GBP shows any evidence of reversal.
+It is not statistically significant.
 

### 2018-01-10 R/Futures.R
Updated response and some assessment of high and low spec.

@@ -3,7 +3,7 @@
 # rates and measures of sentiment.    
 # Read spec position data----------------------------------
 require(eventstudies)
-da <- read.csv("./Futures.csv", header = TRUE)
+da <- read.csv("./Data/Futures.csv", header = TRUE)
 da$S1 <- (da$NCL-da$NCS)/(da$NCL+da$NCS+da$NCSP)
 da$S2 <- (da$NCL-da$NCS)/da$OI
 da$S3 <- (da$NCL+da$NCS+da$NCSP)/da$OI

### 2018-01-10 R/comparison.R

@@ -1,0 +1,29 @@
+#Comparison.R
+#This will compare the extreme readings for each of the methods
+da <- read.csv("./Data/Futures.csv", header = TRUE, stringsAsFactors = FALSE)
+da$Date <- as.Date(da$Date, format = "%d/%m/%Y")
+da$S1 <- (da$NCL-da$NCS)/(da$NCL+da$NCS+da$NCSP)
+da$S2 <- (da$NCL-da$NCS)/da$OI
+da$S3 <- (da$NCL+da$NCS+da$NCSP)/da$OI
+da <- da[,c(1, 2, 3, 5, 19, 20, 21)]
+str(da)
+head(da)
+# Sub set files------------------------------------------
+CADfut <- subset(da, subset = da$Exchange == "CAD")
+CHFfut <- subset(da, subset = da$Exchange == "CHF")
+EURfut <- subset(da, subset = da$Exchange == "EUR")
+GBPfut <- subset(da, subset = da$Exchange == "GBP")
+JPYfut <- subset(da, subset = da$Exchange == "JPY") 
+# Rank to find extremes
+CADS1lo <- head(CADfut[order(CADfut$S1), ], n = 10)
+CADS1hi <- tail(CADfut[order(CADfut$S1), ], n = 10)
+# plot===================
+plot(CADfut$Date, CADfut$S1, type = 'l')
+#====Now compare to risk reversal
+da2 <- read.csv('./Data/RR.csv', stringsAsFactors = FALSE)
+head(da2)
+head(da2[order(da2$EURGBPRR), c(1, 4, 5)], n = 20)
+tail(da2[order(da2$EURGBPRR), c(1, 4, 5)], n = 150)
+# this last one does not seem to work. 
+da3 <- subset(da2, select = c(da2$Date, da2$EURGBP, da2$EURGBPRR))
+head(da3)

### 2018-01-15 MDPI/MDPI_template/FinStudies.tex
Added the ReviewComments and changed some references. I have added the changes package to track the changes.
Updated some comments and the FinStudies text with changes.

@@
 \usepackage[flushleft]{threeparttable}
+\usepackage{changes}
 %\usepackage[comma, sort&compress]{natbib}% Use the natbib reference package - read up on this to edit the reference style; if you want text (e.g.
@@
 Now speculation is an uninformed and destabilising force.
+\added[remark=1]{We have therefore two concepts of speculation: one that is random and disruptive and one that is informed and beneicial in its provison of liqidity.
+
+The debate about the relative merit of these two concepts of speculation continues to the current era where there has been much discussion over the role that speculation played in the boom and bust of commodities prices in the period between 2006 and 2007.
+A common story that has encouraged congressional investigation says that wild speculation drove oil and other commodities prices well above their fundamental value causing harship for commodity consumers and a subsequent collapse that was equally disruptive for the producders.
+See \citet{} and \citet{} for fuller details.
+However, \citet{} argues that this is not the case and this is backed by \citet{}}
 
@@
 An analysis by \citet{Steely2013} of the bid-ask spreads for USD-JPY, GBP-USD and EUR-USD for the period January 2001 to 2005 finds that the average spread for each is 0.342\%, 0.305\% and 0.326\% respectively.
-This would be the cost for a round-trip (buying and selling immediately).}  If speculators are informed their activity drives exchange rates towards equilibrium and prices should follow a random walk after these extremes; if speculators are uninformed and their activity can be regarded as noise, the most extreme speculative sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when reversals are most likely to take place.
+This would be the cost for a round-trip (buying and selling immediately).}  \added[remark=2]{The intenisty of speculation is measured with the use of option risk-reversal skew.
+As these become more extreme it signals that expectations of buyers and sellers in the foreign exchange options market have diverged significantly from the forward rate.
+Weight of speculative positions is measured by the net long or short of speculative accounts in the futures market.
+If the movement of prices towards equilibrium, \emph{price discovery} is facilitated by order flow from informed traders (as has been asserted theoetically by \citet{Glosten1985Bid} and \citet{Kyle1985Continuous} while being demonstrted empirically by \citet{Lyons1995Microstructure} and \citet{Evans2002Order} and \citet{rime2010exchange}, speculative sentiment and positions would be part of the price discovery process if informed and would disript this process if not.
+Therefore,}if speculators are informed their activity drives exchange rates towards equilibrium and prices should follow a random walk after these extremes \added[remark=3]{of speculative sentiment or position}; if speculators are uninformed and their activity can be regarded as noise, the most extreme speculative sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when reversals are most likely to take place.
 
@@
 
+\replaced{More recently the debate has move towards the way that noise-trading speculators may affect welfare and how that could be contained.
+\citet{Tobin1982} and \citet{Tobvin} developed the argument for the eponymous transaction tax that would seek to limit the influence of exchange rate speculation.
+\citet{ERW} use a series of fundamentals to show how speculation casued unwarranted devaluations during the period of the European Exchange Rate Mechanism (ERM).
+\citet{Tobin3} use this experience to re-state the call for a tax to reduce speculation.
+However, if speculation is informed and part of the process of price discovery this tax can be counterproductive as it will reduce liquidity and may cause evengreater volatility.
+This is a point made by \citet{Mayer} and \citet{McKinnon}.
+
+The call for a tax on speculative activity in financial markets arose again in the middle of the first decade of the new century as a result of surge in commodities prices that was experienced in the period 2005 to 2007.
+Focused mostly on oil but encompassing many other energy, metal and food commodities, one argument attributed the rise in prices again to speculative actvity.
+See..}.
+
 \section{Measuring speculation}
-This study uses a unique series of option \emph{risk-reversal skews} to measure of speculative sentiment\footnote{It appears that this is the only period for which the data currently exist.
-The risk-reversal data is very difficult to obtain in an over-the-counter market like foreign exchange where there is no centralised exchange to record and collect information, generally being the property of quoting institutions.
-The risk-reversals are the average prices quoted at 10:00 GMT by major banks on the Thomson-Reuters composite option pages (OPT1)  for the period between January 2 1996 and September 6 2002.
+\replaced{This study uses a unique series of option}{A set of} \emph{risk-reversal skews}\added{are used} to measure \delete{of} speculative sentiment\footnote{\deleted{It appears that this is the only period for which the data currently exist.
+The risk-reversal data is very difficult to obtain in an over-the-counter market like foreign exchange where there is no centralised exchange to record and collect information, generally being the property of quoting institutions.}  The risk-reversals are the average prices quoted at 10:00 GMT by major banks on the Thomson-Reuters composite option pages (OPT1)  for the period between January 2 1996 and September 6 2002.
 The risk-reversals are the spread of implied volatility for a 1 month 25 delta at-the-money call option relative to the equivalent put.
-Risk-reversal skew is the relative price of currency put and call options for the same strike price.}, as well as CFTC data on speculative positions.
-The risk-reversal skew is the relationship between a call option and the equivalent put.
-If the exchange rate is expected to follow a random walk with drift towards the 1 month forward, implied volatility on the put and the call for the same delta should be equal.
-Any deviation from this equivalence is an indication that a divergence from the random walk with drift is implied in the market price of options.\footnote{If the risk-reversal for a 1-month USD/CAD is quoted at 0.15-0.28\% and the implied volatility is 8.50\%, the market-maker would be willing to buy the 1-month 25 delta USD-put-CAD-call at 8.65\% and sell the USD-call-CAD-put at 8.50\%.
+Risk-reversal skew is the relative price of currency put and call options for the same strike price.}\delete{, as well as CFTC data on speculative positions.} The risk-reversal skew is the relationship between a call option and the equivalent put.
+If the exchange rate is expected to follow a random walk with drift towards the 1 month forward \added[remark=4]{the same maturity as the option}, implied volatility on the put and the call for the same delta should be equal.
+Any deviation from this equivalence is an indication that a divergence from the random walk with drift is implied in the market price of options.
+
+\added{It is common practice to back market expectations out of information drawn from option prices.
+\footnote{If the risk-reversal for a 1-month USD/CAD is quoted at 0.15-0.28\% and the implied volatility is 8.50\%, the market-maker would be willing to buy the 1-month 25 delta USD-put-CAD-call at 8.65\% and sell the USD-call-CAD-put at 8.50\%.
 The dealer pays 0.15\%.
@@
 \includegraphics[scale=0.8]{FPCum6w}
-\caption{Event Study: Extreme (Low $5{th}$ percentile) Risk Reversal Skew and 16 day event window: Risk reversals are the ratio of implied volatility on calls to the equivalent put; extreme low is at or above the $5^{th}$ percentile for the whole range, cumulative returns are the sum of log exchange rate returns.
-The event day is the day of the extreme risk reversal reading.
-If speculation is uninformed, extremes should be followed by reversals; if speculation is informed, extremes are more likely to be followed by a random walk.
-The solid line is the cumulative mean return from the start of the event window; the dashed lines show the 95\% confidence intervals constructed from 1000 random bootstrap samples of appropriate event window day.}
-\label{fig:ES3}
-\end{figure}
+\caption{Event Study:  Extreme ($90^{th}$ percentile), Non-commercial net long (S1), 6 week window.
+Non-commercial net per OI (S2) are the net speculative positions (as reported to US regulator the CFTC) relative to all open positions; extreme high is at or above the $90^{th}$ percentile for the whole range, cumulative returns are the sum of log exchange rate returns.
+The event day is the day of the extreme reading as a measure of extreme speculative activity.
+If speculation is uninformed, extremes should be followed by reversals; if speculation is informed, extremes are more likely to be followed by a random walk.}
 
@@
 If speculation is uninformed, extremes should be followed by reversals; if speculation is informed, extremes are more likely to be followed by a random walk.}
+
 \label{fig:ES4}
@@
 An event study showed that speculation is associated with price movement: exchange rates move in the direction of speculative sentiment and activity.
-However, once extreme, prices are as likely to continue as to reverse.
+However, once extreme \added{speculative sentiment or extreme speculative positions are established}, prices are as likely to continue as to reverse.
 Speculative sentiment and speculative activity influence foreign exchange prices as indicated by the covariance of speculative measures with returns.
@@
 It is less clear whether the results would be reproduced across markets for other financial assets.
-The results do not rule out uninformed speculation.
+\added{The foreign exchange market has a particular microstructure that is dominated by over-the-counter transactions in the spot market and this may influence the results.
+Further reseach might look at risk-reversals in financial markets that are traded on electronic exchanges to see if these results are replicated.
+The data on speculative positions are from the excahnge traded futures market.} \deleted{The results do not rule out uninformed speculation.
 The Friedman effect from the Noise-trader model, whereby the informed speculators take advantage of the space created by price pressure from the uniformed, depends on there being \emph{noise-trader risk} or some limits to arbitrage.
-The effect is not evident in this study.
-However, the foreign exchange market is a particularly large and liquid market, other asset classes may have alternative institutional features that mean a different relationship or balance between the informed and uninformed and therefore different results.
+The effect is not evident in this study.}  However, the foreign exchange market is a particularly large and liquid market, other asset classes may have alternative institutional features that mean a different relationship or balance between the informed and uninformed and therefore different results.
 
-This is important because there is no unambiguous view of speculative activity and because speculation is an important issues in a number of policy debates.
+\added{It is possible to dispute the link between risk reversal and market sentiment.
+It is not only the direct change in expectations about the future that affect  risk-reversals.
+The weight of positions from hedging activity could be influential.
+Therefore future research might also investigate the link between these figures and other measures of short-term sentiment if they exist.}
+\deleted{This is important because there is no unambiguous view of speculative activity and because speculation is an important issues in a number of policy debates.
 Most notably, the \emph{Tobin Tax} would put a charge against financial transactions, particularly in foreign exchange.
@@
 However, if speculation is informed and part of the price discovery process, this tax may have the opposite effect to what is envisaged.
-These results suggest that it could reduce liquidity and increase volatility.
+These results suggest that it could reduce liquidity and increase volatility.}
 

### 2018-01-15 MDPI/MDPI_template/FinStudies.tex
The pdf can be created.

@@
 Now speculation is an uninformed and destabilising force.
-\added[remark=1]{We have therefore two concepts of speculation: one that is random and disruptive and one that is informed and beneicial in its provison of liqidity.
+\added{We have therefore two concepts of speculation: one that is random and disruptive and one that is informed and beneicial in its provison of liqidity.
 
@@
 \section{Measuring speculation}
-\replaced{This study uses a unique series of option}{A set of} \emph{risk-reversal skews}\added{are used} to measure \delete{of} speculative sentiment\footnote{\deleted{It appears that this is the only period for which the data currently exist.
+\replaced{This study uses a unique series of option}{A set of} \emph{risk-reversal skews}\added{are used} to measure \deleted{of} speculative sentiment\footnote{\deleted{It appears that this is the only period for which the data currently exist.
 The risk-reversal data is very difficult to obtain in an over-the-counter market like foreign exchange where there is no centralised exchange to record and collect information, generally being the property of quoting institutions.}  The risk-reversals are the average prices quoted at 10:00 GMT by major banks on the Thomson-Reuters composite option pages (OPT1)  for the period between January 2 1996 and September 6 2002.
 The risk-reversals are the spread of implied volatility for a 1 month 25 delta at-the-money call option relative to the equivalent put.
-Risk-reversal skew is the relative price of currency put and call options for the same strike price.}\delete{, as well as CFTC data on speculative positions.} The risk-reversal skew is the relationship between a call option and the equivalent put.
+Risk-reversal skew is the relative price of currency put and call options for the same strike price.}\deleted{, as well as CFTC data on speculative positions.} The risk-reversal skew is the relationship between a call option and the equivalent put.
 If the exchange rate is expected to follow a random walk with drift towards the 1 month forward \added[remark=4]{the same maturity as the option}, implied volatility on the put and the call for the same delta should be equal.
@@
 
-\added{It is common practice to back market expectations out of information drawn from option prices.
-\footnote{If the risk-reversal for a 1-month USD/CAD is quoted at 0.15-0.28\% and the implied volatility is 8.50\%, the market-maker would be willing to buy the 1-month 25 delta USD-put-CAD-call at 8.65\% and sell the USD-call-CAD-put at 8.50\%.
+\added{It is common practice to back market expectations out of information drawn from option prices.}  \footnote{If the risk-reversal for a 1-month USD/CAD is quoted at 0.15-0.28\% and the implied volatility is 8.50\%, the market-maker would be willing to buy the 1-month 25 delta USD-put-CAD-call at 8.65\% and sell the USD-call-CAD-put at 8.50\%.
 The dealer pays 0.15\%.
@@
 If speculation is uninformed, extremes should be followed by reversals; if speculation is informed, extremes are more likely to be followed by a random walk.}
+\label{fig:ES3}
+\end{figure}
 
@@
 If speculation is uninformed, extremes should be followed by reversals; if speculation is informed, extremes are more likely to be followed by a random walk.}
-
 \label{fig:ES4}

### 2018-01-16 MDPI/MDPI_template/FinStudies.tex
Updated some parts of the introduction

--- 2018-01-15 MDPI/MDPI_template/FinStudies.tex
@@
 \usepackage{changes}
+\setremarkmarkup{(#2)}
 %\usepackage[comma, sort&compress]{natbib}% Use the natbib reference package - read up on this to edit the reference style; if you want text (e.g.
@@
 \section{Introduction}
-If speculation is based on information it is a stabilising force, providing liquidity and helping to ensure that markets swiftly find equilibrium.
-\citet{KeynesHedge} and \citet{HicksHedge} for example emphasise the role that speculation plays in facilitating financial market transactions.
+\added{Speculation can be stabilising or destabilising.} If speculation is based on information it is a stabilising force, providing liquidity and helping to ensure that markets swiftly find equilibrium.
+\citet{KeynesHedge} and \citet{HicksHedge} \deleted{for example} emphasise the role that speculation plays in facilitating financial market transactions \added{by providing liquidity that allows hedgers and other market participants to swiftly find counterparties with the need for large prices concessions.}.
 Speculators offer a service for a fee that is determined by the return on their activity.
-While speculators may be trading on information or noise, informed speculators will tend to drive out the uninformed noise-traders according to \citet{FriedmanPositive}.
-However, the later \citet[p.
-101]{Keynes1936} regarded speculation as a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd' and \citet{Delong1990noise} provides a model of cases where speculative noise provides a constrain on informed traders benefiting from their knowledge.
+\replaced{While speculators may be trading on information or noise, informed speculators will tend to drive out the uninformed noise-traders according to \citet{FriedmanPositive}.}{\citet{FriedmanPositive} argued that even where there are a mixture of infomed and unnformed speculators, the informed woudl tend to buy when prices were below intrinsic value and sell when above, making profits, stabilising the market and driving the loss-making uninformed speculators out of teh market.} \deleted{However, the later} \citet[p.
+101]{Keynes1936} \added{also had a later, more negative view of speculation.
+Here he} regarded speculation as a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd'\added{.
+Uninformed speculators may dominate} and \citet{Delong1990noise} provides a model of cases where speculative noise provides a constraint on informed traders benefiting from their knowledge.
 Now speculation is an uninformed and destabilising force.
-\added[remark=1]{We have therefore two concepts of speculation: one that is random and disruptive and one that is informed and beneicial in its provison of liqidity.
+\added{We have therefore two concepts of speculation: one that is random and disruptive and one that is informed and beneicial in its provison of liqidity.
 
@@
 A common story that has encouraged congressional investigation says that wild speculation drove oil and other commodities prices well above their fundamental value causing harship for commodity consumers and a subsequent collapse that was equally disruptive for the producders.
-See \citet{} and \citet{} for fuller details.
-However, \citet{} argues that this is not the case and this is backed by \citet{}}
+\citet{PindyckSpec} show a broad movement of commodities prices not linked to fundaments that they argue is caused by speculation.
+\citet{TangSpec} suggest that this speculative disruption was exacerbated by speculation on commodity indices.
+However, \citet{ScottSpec} do not find a link between speculative positions and excess price movements in commodities.
+\citet{Jacksspec} documents the history of attacks on speculation following price volatility.}
 
@@
 An analysis by \citet{Steely2013} of the bid-ask spreads for USD-JPY, GBP-USD and EUR-USD for the period January 2001 to 2005 finds that the average spread for each is 0.342\%, 0.305\% and 0.326\% respectively.
-This would be the cost for a round-trip (buying and selling immediately).}  \added[remark=2]{The intenisty of speculation is measured with the use of option risk-reversal skew.
+This would be the cost for a round-trip (buying and selling immediately).}  \added{The intenisty of speculation is measured with the use of option risk-reversal skew.
 As these become more extreme it signals that expectations of buyers and sellers in the foreign exchange options market have diverged significantly from the forward rate.
@@
 If the movement of prices towards equilibrium, \emph{price discovery} is facilitated by order flow from informed traders (as has been asserted theoetically by \citet{Glosten1985Bid} and \citet{Kyle1985Continuous} while being demonstrted empirically by \citet{Lyons1995Microstructure} and \citet{Evans2002Order} and \citet{rime2010exchange}, speculative sentiment and positions would be part of the price discovery process if informed and would disript this process if not.
-Therefore,}if speculators are informed their activity drives exchange rates towards equilibrium and prices should follow a random walk after these extremes \added[remark=3]{of speculative sentiment or position}; if speculators are uninformed and their activity can be regarded as noise, the most extreme speculative sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when reversals are most likely to take place.
+Therefore,}if speculators are informed their activity drives exchange rates towards equilibrium and prices should follow a random walk after these extremes \added{of speculative sentiment or position}; if speculators are uninformed and their activity can be regarded as noise, the most extreme speculative sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when reversals are most likely to take place.
 
@@
 Risk-reversal skew is the relative price of currency put and call options for the same strike price.}\delete{, as well as CFTC data on speculative positions.} The risk-reversal skew is the relationship between a call option and the equivalent put.
-If the exchange rate is expected to follow a random walk with drift towards the 1 month forward \added[remark=4]{the same maturity as the option}, implied volatility on the put and the call for the same delta should be equal.
+If the exchange rate is expected to follow a random walk with drift towards the 1 month forward \added{the same maturity as the option}, implied volatility on the put and the call for the same delta should be equal.
 Any deviation from this equivalence is an indication that a divergence from the random walk with drift is implied in the market price of options.

### 2018-01-16 MDPI/MDPI_template/FinStudies.tex

@@
 \section{Measuring speculation}
-\replaced{This study uses a unique series of option}{A set of} \emph{risk-reversal skews}\added{are used} to measure \delete{of} speculative sentiment\footnote{\deleted{It appears that this is the only period for which the data currently exist.
-The risk-reversal data is very difficult to obtain in an over-the-counter market like foreign exchange where there is no centralised exchange to record and collect information, generally being the property of quoting institutions.}  The risk-reversals are the average prices quoted at 10:00 GMT by major banks on the Thomson-Reuters composite option pages (OPT1)  for the period between January 2 1996 and September 6 2002.
-The risk-reversals are the spread of implied volatility for a 1 month 25 delta at-the-money call option relative to the equivalent put.
-Risk-reversal skew is the relative price of currency put and call options for the same strike price.}\delete{, as well as CFTC data on speculative positions.} The risk-reversal skew is the relationship between a call option and the equivalent put.
-If the exchange rate is expected to follow a random walk with drift towards the 1 month forward \added{the same maturity as the option}, implied volatility on the put and the call for the same delta should be equal.
-Any deviation from this equivalence is an indication that a divergence from the random walk with drift is implied in the market price of options.
-
-\added{It is common practice to back market expectations out of information drawn from option prices.
-\footnote{If the risk-reversal for a 1-month USD/CAD is quoted at 0.15-0.28\% and the implied volatility is 8.50\%, the market-maker would be willing to buy the 1-month 25 delta USD-put-CAD-call at 8.65\% and sell the USD-call-CAD-put at 8.50\%.
+\added{It is common practice to back market expectations out of information drawn from option prices.}  \footnote{If the risk-reversal for a 1-month USD/CAD is quoted at 0.15-0.28\% and the implied volatility is 8.50\%, the market-maker would be willing to buy the 1-month 25 delta USD-put-CAD-call at 8.65\% and sell the USD-call-CAD-put at 8.50\%.
 The dealer pays 0.15\%.

### 2018-01-16 MDPI/MDPI_template/FinStudies.tex

@@
 \added{Speculation can be stabilising or destabilising.} If speculation is based on information it is a stabilising force, providing liquidity and helping to ensure that markets swiftly find equilibrium.
-\citet{KeynesHedge} and \citet{HicksHedge} \deleted{for example} emphasise the role that speculation plays in facilitating financial market transactions \added{by providing liquidity that allows hedgers and other market participants to swiftly find counterparties with the need for large prices concessions.}.
-Speculators offer a service for a fee that is determined by the return on their activity.
-\replaced{While speculators may be trading on information or noise, informed speculators will tend to drive out the uninformed noise-traders according to \citet{FriedmanPositive}.}{\citet{FriedmanPositive} argued that even where there are a mixture of infomed and unnformed speculators, the informed woudl tend to buy when prices were below intrinsic value and sell when above, making profits, stabilising the market and driving the loss-making uninformed speculators out of teh market.} \deleted{However, the later} \citet[p.
+\citet{KeynesHedge} and \citet{HicksHedge} \deleted{for example} emphasise the role that speculation plays in facilitating financial market transactions \added{by providing liquidity that allows hedgers and other market participants to swiftly find counterparties without the need for large prices concessions.} Speculators offer a service for a fee that is determined by the return on their activity.
+\replaced{\citet{FriedmanPositive} argued that even where there are a mixture of infomed and unnformed speculators, the informed woudl tend to buy when prices were below intrinsic value and sell when above, making profits, stabilising the market and driving the loss-making uninformed speculators out of business.}{While speculators may be trading on information or noise, informed speculators will tend to drive out the uninformed noise-traders according to \citet{FriedmanPositive}.} \deleted{However, the later} \citet[p.
 101]{Keynes1936} \added{also had a later, more negative view of speculation.
-Here he} regarded speculation as a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd'\added{.
-Uninformed speculators may dominate} and \citet{Delong1990noise} provides a model of cases where speculative noise provides a constraint on informed traders benefiting from their knowledge.
+Here he} regarded speculation as a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd''\added{.
+Uninformed speculators may now dominate} and \citet{Delong1990noise} provide a model \deleted{of cases} where speculative noise provides a constraint on informed traders benefiting from their knowledge.
 Now speculation is an uninformed and destabilising force.
@@
 The debate about the relative merit of these two concepts of speculation continues to the current era where there has been much discussion over the role that speculation played in the boom and bust of commodities prices in the period between 2006 and 2007.
-A common story that has encouraged congressional investigation says that wild speculation drove oil and other commodities prices well above their fundamental value causing harship for commodity consumers and a subsequent collapse that was equally disruptive for the producders.
+A common story that has encouraged Congressional investigation says that wild speculation drove oil and other commodities prices well above their fundamental value causing harship for commodity consumers and a subsequent collapse that was equally disruptive for the producders.
 \citet{PindyckSpec} show a broad movement of commodities prices not linked to fundaments that they argue is caused by speculation.
-\citet{TangSpec} suggest that this speculative disruption was exacerbated by speculation on commodity indices.
-However, \citet{ScottSpec} do not find a link between speculative positions and excess price movements in commodities.
-\citet{Jacksspec} documents the history of attacks on speculation following price volatility.}
+\citet{TangSpec} suggest that this disruption was exacerbated by speculation on commodity indices.
+However, \citet{ScottSpec} do not find a link between speculative positions and excess price movements in commodities and \citet{Jacksspec} documents a long history of attacks on speculation following price volatility.}
 
@@
 
-\replaced{More recently the debate has move towards the way that noise-trading speculators may affect welfare and how that could be contained.
+\added{More recently the debate has move towards the way that noise-trading speculators may affect welfare and how that could be contained.
 \citet{Tobin1982} and \citet{Tobvin} developed the argument for the eponymous transaction tax that would seek to limit the influence of exchange rate speculation.
@@
 \section{Measuring speculation}
-\added{It is common practice to back market expectations out of information drawn from option prices.}  \footnote{If the risk-reversal for a 1-month USD/CAD is quoted at 0.15-0.28\% and the implied volatility is 8.50\%, the market-maker would be willing to buy the 1-month 25 delta USD-put-CAD-call at 8.65\% and sell the USD-call-CAD-put at 8.50\%.
+\added{It is common practice to back market expectations out of information drawn from option prices.
+This is where the expectatiosn and other information goes.
+Using the risk-reversal skew is a simplified version of that.}  \footnote{If the risk-reversal for a 1-month USD/CAD is quoted at 0.15-0.28\% and the implied volatility is 8.50\%, the market-maker would be willing to buy the 1-month 25 delta USD-put-CAD-call at 8.65\% and sell the USD-call-CAD-put at 8.50\%.
 The dealer pays 0.15\%.

### 2018-01-19 MDPI/MDPI_template/FinStudies.tex
Updated major reviewsions.

--- 2018-01-16 MDPI/MDPI_template/FinStudies.tex
@@
 \added{Speculation can be stabilising or destabilising.} If speculation is based on information it is a stabilising force, providing liquidity and helping to ensure that markets swiftly find equilibrium.
-\citet{KeynesHedge} and \citet{HicksHedge} \deleted{for example} emphasise the role that speculation plays in facilitating financial market transactions \added{by providing liquidity that allows hedgers and other market participants to swiftly find counterparties with the need for large prices concessions.}.
-Speculators offer a service for a fee that is determined by the return on their activity.
-\replaced{While speculators may be trading on information or noise, informed speculators will tend to drive out the uninformed noise-traders according to \citet{FriedmanPositive}.}{\citet{FriedmanPositive} argued that even where there are a mixture of infomed and unnformed speculators, the informed woudl tend to buy when prices were below intrinsic value and sell when above, making profits, stabilising the market and driving the loss-making uninformed speculators out of teh market.} \deleted{However, the later} \citet[p.
+\citet{KeynesHedge} and \citet{HicksHedge} \deleted{for example} emphasise the role that speculation plays in facilitating financial market transactions \added{by providing liquidity that allows hedgers and other market participants to swiftly find counterparties without the need for large prices concessions.
+In this version, speculators} \deleted{Speculators} offer a service for a fee that is determined by the return on their activity.
+\replaced{\citet{FriedmanPositive} argued that even where there is a mixture of infomed and unnformed speculators, the informed woudl tend to buy when prices were below intrinsic value and sell when above, making profits, stabilising the market and driving the loss-making uninformed speculators out of business.}{While speculators may be trading on information or noise, informed speculators will tend to drive out the uninformed noise-traders according to \citet{FriedmanPositive}.} \deleted{However, the later} \citet[p.
 101]{Keynes1936} \added{also had a later, more negative view of speculation.
-Here he} regarded speculation as a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd'\added{.
-Uninformed speculators may dominate} and \citet{Delong1990noise} provides a model of cases where speculative noise provides a constraint on informed traders benefiting from their knowledge.
+Here he} regarded it as a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd''\added{.
+Uninformed speculators may now dominate} and \citet{Delong1990noise} provide a model \deleted{of cases} where speculative noise \replaced{constrains}{provides a constraint on} informed traders \added{from} benefiting from their knowledge.
 Now speculation is an uninformed and destabilising force.
@@
 
-The debate about the relative merit of these two concepts of speculation continues to the current era where there has been much discussion over the role that speculation played in the boom and bust of commodities prices in the period between 2006 and 2007.
-A common story that has encouraged congressional investigation says that wild speculation drove oil and other commodities prices well above their fundamental value causing harship for commodity consumers and a subsequent collapse that was equally disruptive for the producders.
-\citet{PindyckSpec} show a broad movement of commodities prices not linked to fundaments that they argue is caused by speculation.
-\citet{TangSpec} suggest that this speculative disruption was exacerbated by speculation on commodity indices.
-However, \citet{ScottSpec} do not find a link between speculative positions and excess price movements in commodities.
-\citet{Jacksspec} documents the history of attacks on speculation following price volatility.}
+The debate about the relative merit of these two concepts of speculation continues.
+There has been particular focus on commodities markets and foreign exchange markets.
+\citet{Jacksspec} documents a long history of attacks on speculation following price volatility.} This has been particularly evident in commodities and foreign exchange markets.
+There has been much discussion over the role that speculation played in the boom and bust of commodities prices in the period between 2006 and 2007.
+\citet{TangSpec} suggest that this disruption was exacerbated by newly estabilished commodity indices that encoraged speculation.
 
-This paper evaluates the relative importance of these two concepts of speculation by investigating the relationship between the intensity of speculation and the weight of speculative positions relative to subsequent movement of foreign exchange prices.\footnote{\label{FX}The foreign exchange market is chosen for this study because it is the largest and most liquid financial market in the world.
+There have also been investigations of the link between speclulation or noise trading and price misalignments or volatility across a range of other asset marketsA.
+\citet{Verma} examine the link between noise trading and equity markets; \citet{Hongjun} asks whether speculation in aggregate is just random noise lacking influence on prices.
+Howevever, foreign exchange has been a particular focus as a consequence of its liquidity, volatility and the way that speculative attacks appear to have infringed on government soverignty.
+See \citet{Dornbusch1988} in geneal and \citet{Eichengreen1988} for discussion of the speculative attacks on the European exchange rate system.
+\citet{Eichengreen1994} use a set of fundamentals to show that speculative attacks drove exchange rates away from their fundamental value.
+However, others have made the case that speculative is a key part of the forign exchange market, providing liquidity and contribuing to stability.
+See \citet{Mayer1985} and \citet{McKinnon1986}.
+
+There is no consensus and the the debate enters policy circles with the suggestion that a tax on financial transactions could reduce speculation and thereby limit misalignment and volatility.
+\citet[pp.158]{TobinTax} and \citet{Tobin1988a} say that such a tax would "throw some sand in the well-greased wheels." of the foreign exchange market.
+At face value the tax has the advantage of improving market efficiency and raising funds for other welfare-enhancing acitvity.
+However, this is only the case if speculation is uninformed and destabilising.
+If it is informed and contributes to stability the tax will come at a cost.
+
+This paper evaluates the relative importance of these two concepts of speculation by investigating the relationship between the intensity of speculation and the weight of speculative positions relative to subsequent movement of foreign exchange prices.
+\added{The foreign exchange market is appropriate this study because it is central to the call for the introduction of a transaction tax as a consequence of it being the largest and most liquid financial market in the world.
 The latest survey taken by the Bank for International Settlements (BIS) in April 2016 estimated average daily turnover at \$5.1trn with average daily spot transactions of \$1.7trn See \citet{BISFX2016}.
@@
 An analysis by \citet{Steely2013} of the bid-ask spreads for USD-JPY, GBP-USD and EUR-USD for the period January 2001 to 2005 finds that the average spread for each is 0.342\%, 0.305\% and 0.326\% respectively.
-This would be the cost for a round-trip (buying and selling immediately).}  \added{The intenisty of speculation is measured with the use of option risk-reversal skew.
-As these become more extreme it signals that expectations of buyers and sellers in the foreign exchange options market have diverged significantly from the forward rate.
-Weight of speculative positions is measured by the net long or short of speculative accounts in the futures market.
-If the movement of prices towards equilibrium, \emph{price discovery} is facilitated by order flow from informed traders (as has been asserted theoetically by \citet{Glosten1985Bid} and \citet{Kyle1985Continuous} while being demonstrted empirically by \citet{Lyons1995Microstructure} and \citet{Evans2002Order} and \citet{rime2010exchange}, speculative sentiment and positions would be part of the price discovery process if informed and would disript this process if not.
-Therefore,}if speculators are informed their activity drives exchange rates towards equilibrium and prices should follow a random walk after these extremes \added{of speculative sentiment or position}; if speculators are uninformed and their activity can be regarded as noise, the most extreme speculative sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when reversals are most likely to take place.
+This would be the cost for a round-trip (buying and selling immediately).This makes it relatively attractive for speculators.
+
+The intenisty of speculation is measured in two ways: speculative sentiment and speculative positions.
+Speculative sentiment is identified with the use of option risk-reversal skew.
+As these become more extreme it signals that expectations of future value in the foreign exchange options market have diverged significantly from the appropraite forward rate.
+The Weight of speculative positions is identified with US regulatory data of the psotions held by type of participant in the futures market.
+Net long and net short are open positions held by spculators relative to overall activit in the market.
+See Section \ref{secref:measure} for full data details.}
 
@@
 
-The comparison of informed and uninformed traders is a central component of microstructure research.
-At the heart of the \citet{Kyle1985Continuous} model of market-making is the interaction of orders from the informed and the uninformed.
-On the assumption that uniformed orders arrive in a random fashion, the weight of orders reveals the direction of private information and should encourage adjustment of the bid-ask spread as part of the process of price discovery.
-\citet{glosten1985bid} presents a model where the bid-ask spread is a function of adverse selection, being caused by the risk that orders come from informed traders.
-
 \citet[p.529]{BlackNoise} asserted that the interaction of informed and uninformed activity was essential to the smooth functioning of financial markets.
@@
 
-\replaced{More recently the debate has move towards the way that noise-trading speculators may affect welfare and how that could be contained.
-\citet{Tobin1982} and \citet{Tobvin} developed the argument for the eponymous transaction tax that would seek to limit the influence of exchange rate speculation.
-\citet{ERW} use a series of fundamentals to show how speculation casued unwarranted devaluations during the period of the European Exchange Rate Mechanism (ERM).
-\citet{Tobin3} use this experience to re-state the call for a tax to reduce speculation.
-However, if speculation is informed and part of the process of price discovery this tax can be counterproductive as it will reduce liquidity and may cause evengreater volatility.
-This is a point made by \citet{Mayer} and \citet{McKinnon}.
+\added{This study is about short-term movements in foreign exchange prices.
+The event windows are between 1 week and 4 weeks in length.
+While the interpretation of economic fundamentals can drive prices, market posotions and liquidity trades will also be important.
+Even at longer horizons it is often very difficult to connect exchange rate movements to specific events.
+\citet{Meese1983Empirical} find that forecasts using exchange rate models cannot beat a random walk while \citet{ChinnModel} updated the study to find very similar results.
+.  
+The foreign exchange market is primarily a dealer-orientated market.
+It is largely consistent with the theortical microstructure models that have been presented by the likes of \citet{Kyle1985Continuous} and \citet{glosten1985bid}.
+The interaction of informed and uninformed traders drives the price-discovery process: if the uninformed traders are random, the market-maker earns the bid-ask spread; if informed traders arrive they drive the price towards equilibrium.}
 
-The call for a tax on speculative activity in financial markets arose again in the middle of the first decade of the new century as a result of surge in commodities prices that was experienced in the period 2005 to 2007.
-Focused mostly on oil but encompassing many other energy, metal and food commodities, one argument attributed the rise in prices again to speculative actvity.
-See..}.
+If the movement of the price of foreign exchange towards equilibrium, \emph{price discovery} is facilitated by order flow from informed traders (as has been asserted theoetically by \citet{Glosten1985Bid} and \citet{Kyle1985Continuous} while being demonstrted empirically by \citet{Lyons1995Microstructure} and \citet{Evans2002Order} and \citet{rime2010exchange}, speculative sentiment and positions would be part of the price discovery process if informed and would disript this process if not.
+Therefore,}if speculators are informed their activity drives exchange rates towards equilibrium and prices should follow a random walk after these extremes of speculative sentiment or position; if speculators are uninformed and their activity can be regarded as noise, the most extreme speculative sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when subsequent price reversals are most likely to take place.
 
-\section{Measuring speculation}
-\added{It is common practice to back market expectations out of information drawn from option prices.}  \footnote{If the risk-reversal for a 1-month USD/CAD is quoted at 0.15-0.28\% and the implied volatility is 8.50\%, the market-maker would be willing to buy the 1-month 25 delta USD-put-CAD-call at 8.65\% and sell the USD-call-CAD-put at 8.50\%.
+\section{Measuring speculation}\label{secref:measure}
+\added{There are two steps: identify speculation and assess what happens when it becomes extreme.
+If speculation is uninformed, extremes should cause a divergence from fundamental value and a price reversal should be anticipated; if informed and part of the price-discovery process, extreme speculation achieves a new equilibrium and the following price performance should be a random walk.
+
+The measurement of speculative intensity must identify how far speculators believe that the market will move in a particular direction.
+In the absence of a daily survey of speculators, this study uses a unique series of option\emph{risk-reversal skew} as an measure of speculative sentiment.
+The risk-reversal data is very difficult to obtain in an over-the-counter market like foreign exchange where there is no centralised exchange to record and collect information, generally being the property of quoting institutions.
+The risk-reversals are the average prices quoted at 10:00 GMT by major banks on the Thomson-Reuters composite option pages (OPT1)  for the period between January 2 1996 and September 6 2002.
+The risk-reversals are the spread of implied volatility for a 1 month 25 delta at-the-money call option relative to the equivalent put.
+Risk-reversal skew is the relative price of currency put and call options for the same strike price.} The risk-reversal skew is the relationship between a call option and the equivalent put.
+Equivalence here being measured by the delta or relationship between movement of the underlying price and the movement of the option price.
+While the \citet{Black1973Options} model is the basis for valuing options where the price of a non-dividend paying stock trades at a discount to the forward rate equal to the risk free rate, the forward rate in a currency option is determined by the relative interest rate differential between the two currencies.
+The \citet{GarmanKohlhagan} adjustment that accounts for the difference in interest payments on the two currencies is used.
+The spot discount is determined therefore by the relative interest rate differential for the appropriate maturity.
+1 month in this case.
+This model allows the market practice of pricing options from the volatility implied by the model.
+In other words, the Garman-Kohlhagan is used to extract the volatility that is implied by a specific foreign exchange option price given the strike and other parameters.
+
+It is common practice to back market expectations out of information drawn from option prices.
+See \citet{Malickoption} and \citet{Pochin} for the general approach that combines the implied volatility, risk-reversal and strangles to infer a mixed distrubtion that can be inferreed as the risk-neutral probabiolity density function for a particular asset.
+See \citet{Gaati} an application on the Yen-dollar exchange rate where expectations are affected by central bank intervention.
+Using the risk-reversal is a simplified version of that where the skew is used to infer positive and expectations relative to the appropriate forward rate.
+See ???
+for previous uses of this data.}
+
+If the risk-reversal for a 1-month USD/CAD is quoted at 0.15-0.28\% and the implied volatility is 8.50\%, the market-maker would be willing to buy the 1-month 25 delta USD-put-CAD-call at 8.65\% and sell the USD-call-CAD-put at 8.50\%.
 The dealer pays 0.15\%.
@@
 See \citet{Global} for a technical review.
-}  Therefore, the relationship between implied volatility for calls and implied volatilities for puts for the same delta and maturity strike is an indication of how much more expensive calls are relative to the equivalent put or as the market bias towards puts over calls and as an indication of market sentiment or related market positioning.
+Therefore, the relationship between implied volatility for calls and implied volatilities for puts for the same delta and maturity strike is an indication of how much more expensive calls are relative to the equivalent put or as the market bias towards puts over calls and as an indication of market sentiment or related market positioning.
 The larger the absolute level of the risk-reversal skew, the greater the relative price and, if the deviation is based on information, subsequent price action should be zero on average; if the deviation is based on noise, this movement should be reversed.

### 2018-01-19 MDPI/MDPI_template/FinStudies.tex
Updaed some parts of the paper.

@@
 \section{Introduction}
-\added{Speculation can be stabilising or destabilising.} If speculation is based on information it is a stabilising force, providing liquidity and helping to ensure that markets swiftly find equilibrium.
+\added{Speculation can be viewed as stabilising or destabilising.
+If speculation is informed, it is a stabilising force, providing liquidity and helping to ensure that markets swiftly find equilibrium.
+\citet{KeynesHedge} and \citet{HicksHedge} emphasise the role that speculation plays in facilitating financial market transactions by providing liquidity that allows hedgers and other market participants to swiftly find counterparties without the need for large price concessions.
+In this version, speculators offer a service for a fee that is determined by the return on their activity.
+\citet{FriedmanPositive} argued that even whew there is a mixture of informed and uninformed speculators, the informed would tend to buy when prices were below intrinsic value and sell when above, making profits, limiting volatility and driving the loss-making uninformed speculators out of business.
+However, \citet[p.
+101]{Keynes1936} also had a later, more negative view of speculations.
+Here he regarded it as a myopic, sentient-driven activity, dominated by the desire to ``beat the gun`` or ``outwit the crowd.
+Uninformed speculators may now dominate and \citet{Delong1990noise} provide a model where speculative noise constrains informed traded from benefiting from their knowledge.
+Now speculation is an uninformed and destabilising force.
+We have therefore two concepts of speculation:  one that is random and disruptive and one that is informed and beneficial.
+
+%\added{Speculation can be viewed as stabilising or destabilising.} If speculation is based on information it is a stabilising force, providing liquidity and helping to ensure that markets swiftly find equilibrium.
 \citet{KeynesHedge} and \citet{HicksHedge} \deleted{for example} emphasise the role that speculation plays in facilitating financial market transactions \added{by providing liquidity that allows hedgers and other market participants to swiftly find counterparties without the need for large prices concessions.
 In this version, speculators} \deleted{Speculators} offer a service for a fee that is determined by the return on their activity.
-\replaced{\citet{FriedmanPositive} argued that even where there is a mixture of infomed and unnformed speculators, the informed woudl tend to buy when prices were below intrinsic value and sell when above, making profits, stabilising the market and driving the loss-making uninformed speculators out of business.}{While speculators may be trading on information or noise, informed speculators will tend to drive out the uninformed noise-traders according to \citet{FriedmanPositive}.} \deleted{However, the later} \citet[p.
+\added{\citet{FriedmanPositive} argued that even where there is a mixture of informed and uninformed speculators, the informed would tend to buy when prices were below intrinsic value and sell when above, making profits, limiting volatility and driving the loss-making uninformed speculators out of business.}\deleted{While speculators may be trading on information or noise, informed speculators will tend to drive out the uninformed noise-traders according to \citet{FriedmanPositive}.} However, \deleted{the later} \citet[p.
 101]{Keynes1936} \added{also had a later, more negative view of speculation.
@@
 Now speculation is an uninformed and destabilising force.
-\added{We have therefore two concepts of speculation: one that is random and disruptive and one that is informed and beneicial in its provison of liqidity.
+\added{We have therefore two concepts of speculation: one that is random and disruptive and one that is informed and beneficial in its proviso of liquidity.
 
@@
 There has been particular focus on commodities markets and foreign exchange markets.
-\citet{Jacksspec} documents a long history of attacks on speculation following price volatility.} This has been particularly evident in commodities and foreign exchange markets.
+\citet{Jacksspec} documents a long history of attacks on speculation following price volatility.
+This has been particularly evident in commodities and foreign exchange markets.
 There has been much discussion over the role that speculation played in the boom and bust of commodities prices in the period between 2006 and 2007.
-\citet{TangSpec} suggest that this disruption was exacerbated by newly estabilished commodity indices that encoraged speculation.
+\citet{TangSpec} suggest that this disruption was exacerbated by newly established commodity indices that encouraged speculation.
 
-There have also been investigations of the link between speclulation or noise trading and price misalignments or volatility across a range of other asset marketsA.
-\citet{Verma} examine the link between noise trading and equity markets; \citet{Hongjun} asks whether speculation in aggregate is just random noise lacking influence on prices.
-Howevever, foreign exchange has been a particular focus as a consequence of its liquidity, volatility and the way that speculative attacks appear to have infringed on government soverignty.
-See \citet{Dornbusch1988} in geneal and \citet{Eichengreen1988} for discussion of the speculative attacks on the European exchange rate system.
+There have also been investigations of the link between speculation or noise trading and price misalignments or volatility across a range of other asset markets.
+\citet{Verma} examine the link between noise trading and equity markets; \citet{Hongjun} argues that speculation is not random and influential in aggregation but that behavioural biases will cause systematic deviations from intrinsic value.
+Foreign exchange has been a particular focus as a consequence of its liquidity, volatility and the way that speculative attacks appear to have infringed on government sovereignty.
+See \citet{Dornbusch1988} in general and \citet{Eichengreen1988} for discussion of the speculative attacks on the European exchange rate system.
 \citet{Eichengreen1994} use a set of fundamentals to show that speculative attacks drove exchange rates away from their fundamental value.
-However, others have made the case that speculative is a key part of the forign exchange market, providing liquidity and contribuing to stability.
+However, others have made the case that speculative is a key part of the foreign exchange market, providing liquidity and contributing to stability.
 See \citet{Mayer1985} and \citet{McKinnon1986}.
@@
 \citet[pp.158]{TobinTax} and \citet{Tobin1988a} say that such a tax would "throw some sand in the well-greased wheels." of the foreign exchange market.
-At face value the tax has the advantage of improving market efficiency and raising funds for other welfare-enhancing acitvity.
+At face value the tax has the advantage of improving market efficiency and raising funds for other welfare-enhancing activity.
 However, this is only the case if speculation is uninformed and destabilising.
-If it is informed and contributes to stability the tax will come at a cost.
+If it is informed and contributes to stability the tax will come at a cost.}
 
@@
 
-The intenisty of speculation is measured in two ways: speculative sentiment and speculative positions.
+The intensity of speculation is measured in two ways: speculative sentiment and speculative positions.
 Speculative sentiment is identified with the use of option risk-reversal skew.
-As these become more extreme it signals that expectations of future value in the foreign exchange options market have diverged significantly from the appropraite forward rate.
-The Weight of speculative positions is identified with US regulatory data of the psotions held by type of participant in the futures market.
-Net long and net short are open positions held by spculators relative to overall activit in the market.
+As these become more extreme it signals that expectations of future value in the foreign exchange options market have diverged significantly from the appropriate forward rate.
+The Weight of speculative positions is identified with US regulatory data of the positions held by type of participant in the futures market.
+Net long and net short are open positions held by speculators relative to overall activity in the market.
 See Section \ref{secref:measure} for full data details.}
@@
 Those trading without information are frequently called \emph{noise-traders} or \emph{liquidity traders}.
-However, the terms are not fixed and  \citet[p.
+\added{The first implies some random quality that flow from the lack of information.
+However, the psychologists find that the cognitive short-cuts or \emph{heuristics} that are used to make decisions under uncertainty tend to have biases \citet{Kahneman1979Prospect} and numerous studies purport to find these biases in financial asset prices.
+For example, \citet{DeBondtOver} find evidence that \emph{representativeness} causes stock prices to overshoot and \citet{Coval2005} identifies evidence of behavioural biases among Chicago Board of Trade proprietary traders.
+Liquidity trades are generally those carrying out transactions for reasons other than value.
+This can be passive investment funds buying stocks as money is drawn from investors or hedging activity that does not depend on price.
+The terms are not fixed and  \citet[p.
 1]{GortonNoise} categorise noise-traders as all those trading in markets for ``non-information-based reasons''.
@@
 
-\added{This study is about short-term movements in foreign exchange prices.
+\added{The foreign exchange market is primarily a dealer-orientated market.
+It is largely consistent with the theoretical microstructure models that have been presented by the likes of \citet{Kyle1985Continuous} and \citet{glosten1985bid}.
+The interaction of informed and uninformed traders drives the price-discovery process: if the uninformed traders are random, the market-maker earns the bid-ask spread; if informed traders arrive they drive the price towards equilibrium.
+This study is about short-term movements in foreign exchange prices.
 The event windows are between 1 week and 4 weeks in length.
-While the interpretation of economic fundamentals can drive prices, market posotions and liquidity trades will also be important.
+While the interpretation of economic fundamentals can drive prices, market positions and liquidity trades will also be important.
 Even at longer horizons it is often very difficult to connect exchange rate movements to specific events.
 \citet{Meese1983Empirical} find that forecasts using exchange rate models cannot beat a random walk while \citet{ChinnModel} updated the study to find very similar results.
-.  
-The foreign exchange market is primarily a dealer-orientated market.
-It is largely consistent with the theortical microstructure models that have been presented by the likes of \citet{Kyle1985Continuous} and \citet{glosten1985bid}.
-The interaction of informed and uninformed traders drives the price-discovery process: if the uninformed traders are random, the market-maker earns the bid-ask spread; if informed traders arrive they drive the price towards equilibrium.}
-
-If the movement of the price of foreign exchange towards equilibrium, \emph{price discovery} is facilitated by order flow from informed traders (as has been asserted theoetically by \citet{Glosten1985Bid} and \citet{Kyle1985Continuous} while being demonstrted empirically by \citet{Lyons1995Microstructure} and \citet{Evans2002Order} and \citet{rime2010exchange}, speculative sentiment and positions would be part of the price discovery process if informed and would disript this process if not.
-Therefore,}if speculators are informed their activity drives exchange rates towards equilibrium and prices should follow a random walk after these extremes of speculative sentiment or position; if speculators are uninformed and their activity can be regarded as noise, the most extreme speculative sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when subsequent price reversals are most likely to take place.
+\citet{Lyons1995Microstructure}, \citet{Evans2002Order} and \citet{rime2010exchange} show that order flow can explain exchange rate movements much more effectively than fundamentals.
+The initiating or forcing orders drive the price.
+Therefore, deliberate speculative should drive the price.
+Speculative sentiment changes and speculative positions changes would be part of the price discovery process if informed and would disrupt this process if not.
+Therefore,if speculators are informed their activity drives exchange rates towards equilibrium and prices should follow a random walk after these extremes of speculative sentiment or position; if speculators are uninformed and their activity can be regarded as noise, the most extreme speculative sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when subsequent price reversals are most likely to take place.
+}
 
@@
 It is common practice to back market expectations out of information drawn from option prices.
-See \citet{Malickoption} and \citet{Pochin} for the general approach that combines the implied volatility, risk-reversal and strangles to infer a mixed distrubtion that can be inferreed as the risk-neutral probabiolity density function for a particular asset.
+See \citet{Malickoption} and \citet{Pochin} for the general approach that combines the implied volatility, risk-reversal and strangles to infer a mixed distribution that can be inferred as the risk-neutral probability density function for a particular asset.
 See \citet{Gaati} an application on the Yen-dollar exchange rate where expectations are affected by central bank intervention.
 Using the risk-reversal is a simplified version of that where the skew is used to infer positive and expectations relative to the appropriate forward rate.
-See ???
-for previous uses of this data.}
+See \citet{Eitrheim} and \citet{ECBRR} for previous uses of this data.}
 
@@
 The Commodity Futures Trading Commission (CFTC) collect weekly information about the open positions held by private entities in the US derivative markets.
-A sub-set of this information is released to the public each week as the \emph{Commitment of Traders Report} (CoT).\footnote{The \citet{cot} require traders to categorise themselves as being either commercial (C) or non-commercial (NC).
+A sub-set of this information is released to the public each week as the \emph{Commitment of Traders Report} (CoT).The \citet{cot} require traders to categorise themselves as being either commercial (C) or non-commercial (NC).
 Commercial traders must have some underlying business interest in the security, commodity or instrument that they are trading.
@@
 See \citet{FuturesSanders} and \citet{FuturesWang} for examples of previous use of this data in this fashion.
-
 The data used run from September 30 1998 when the data started to be released on a weekly basis and continue up to December 31 2008.
@@
 The absence of these data seems to be a result of problems with the CFTC database.
-The contracts are for EUR 100,000; GBP 62,500; JPY 12,500,000; CHF 100,000 and CAD 100,000.}
+The contracts are for EUR 100,000; GBP 62,500; JPY 12,500,000; CHF 100,000 and CAD 100,000.
+Two new series are created from the raw data.
+These are labelled S1 and S2.
+S1 measures net non-commercial positions relative to the total non-commercial positions and S2 measures non-commercial positions relative to open interest.
 
+\begin{equation}
+S1 = \frac{NCL - NCS}{NCL + NCS + NCSP}
+\end{equation}
+
+Where NCL are the non-commercial long positions (holding non-US currency); NCS and non-commercial short positions and NCSP and the non-commercial spreading positions (spread long and short between different contract maturities).
+This will measure within the speculative group how biased they are towards holding long or short positions.
+
+\begin{equation}
+S2 = \frac{NCL - NCS}{OI}
+\end{equation}
+
+Where NCL are the non-commercial long positions (holding non-US currency); OI is the open interest in the contract.
+This will measure how biased speculators are to holding long or short positions relative to the whole market.
+As such S2 is the preferred measure as it picks up not only direction but also the prominence of speculators relative to other agents.
+
+S2 is the preferred measure as it
+
+\added{\subsection{The Event Study}
+An event study will be used to identify what happens to foreign exchange prices around the time that speculative sentiment or speculative positions are extreme.
+If speculation is uninformed when speculative sentiment or speculative positions are extremely bias to one direction, there should be the greatest divergence from fundamental value and therefore the greatest risk of reversal; if speculation is informed these extremes are part of the adjustment to a new equilibrium.
 An event study will analyse the performance of security of asset prices around an event.
 In this case the event is the extreme of speculative sentiment or extreme speculative activity.
-See \citet{Dolly1933} for the original version of the event study method and \citet{FamaFisherJensenRoll} for an overview of the method used in this study.
+
+\citet{Dolly1933} is credited with the original event study method while \citet{FamaFisherJensenRoll} developed the method to investigate the reaction of equity prices to news of a stock split.
+An event study will look at the evolution of asset prices in a window around an \emph{event}.
+One of the major challenges in this process is to disentangle the effect of the event from all the other information that is around at that time.
+This problem is usually addressed in two ways: modelling the expected value of the asset and viewing deviations from the expected value as being influenced by the event; using a large sample that, on the condition that there are no confounding factors at that time, would tend to dilute the effect of other information.
+In this case, the expected value should be the forward rate but given that we are looking at very short time horizons and interest rates for these currencies are relatively low, it is reasonable to take the expected value as zero and to investigate positive or negative exchange rate return.
+The samples in this case depend on the how the \emph{extreme} sentiment and positions are defined.
+As is seen in Table \ref{tabref:RR1} taking the $5^{th}$ and $95^{th}$ percentile as the extreme for the risk-reversal skew will give around 100 samples for each exchange rate; Table \ref{tabref:SP1} shows that the $95^{th}$ percentile gives just less than 50 samples for the extreme positions.
+The results do not seem to be sensitive to the definition of extreme.
+}
 
@@
 Alternative measures produced very similar results.
-Alternative quantiles from 1\% and 99\% to 20\% and 80\% have been assessed; Figures \ref{fig:ES1}, \ref{fig:ES2}, \ref{fig:ES3} and \ref{fig:ES4} have much wider windows and do not find any alternative to the view that extreme speculation is followed by a random walk.
-
+Alternative quantiles from 1\% and 99\% to 20\% and 80\% have been assessed; Figures \ref{fig:ES5}, \ref{fig:ES2}, \ref{fig:ES3} and \ref{fig:ES4} have much wider windows and do not find any alternative to the view that extreme speculation is followed by a random walk.
 It is encouraging that two very different measures of speculative activity produce very similar results and that the results hold over different time frames: one week for the option data and one week for the regulatory position data.
 It is less clear whether the results would be reproduced across markets for other financial assets.
+
 \added{The foreign exchange market has a particular microstructure that is dominated by over-the-counter transactions in the spot market and this may influence the results.
-Further reseach might look at risk-reversals in financial markets that are traded on electronic exchanges to see if these results are replicated.
-The data on speculative positions are from the excahnge traded futures market.} \deleted{The results do not rule out uninformed speculation.
+Further research might look at risk-reversals in financial markets that are traded on electronic exchanges to see if these results can be replicated.
+The data on speculative positions are from the exchange traded futures market.} \deleted{The results do not rule out uninformed speculation.
 The Friedman effect from the Noise-trader model, whereby the informed speculators take advantage of the space created by price pressure from the uniformed, depends on there being \emph{noise-trader risk} or some limits to arbitrage.
@@
 The weight of positions from hedging activity could be influential.
-Therefore future research might also investigate the link between these figures and other measures of short-term sentiment if they exist.}
+Therefore future research might also investigate the link between these figures and other measures of short-term sentiment if they exist.}  However, a number of studies have found that exchange rates tend to co-move with survey expectations (see for example \citet{FrankelFroot1987} and \citet{Frankel1987Using}.
+
 \deleted{This is important because there is no unambiguous view of speculative activity and because speculation is an important issues in a number of policy debates.
@@
 These results suggest that it could reduce liquidity and increase volatility.}
+
+It is a common refrain that speculation is destabilising and that it will drive prices away from equilibrium.
+Testing this assertion is difficult because speculation is difficult to measure and equilibrium cannot be viewed.
+However, if speculation is destabilising, when speculators have the most extreme views of the future and when they are most involved in the market should lead to the greatest divergence from equilibrium and the greatest risk that prices will reverse.
+Speculation is measured in two ways:  intensity of sentiment and positions.
+In each case, foreign exchange prices move with speculative sentiment and speculative positions.
+this is consistent with the deliberate orders move prices.
+This would be consistent with speculation being part of the process of price discovery.
+However, in each case the extreme does not tell us anything about the future direction of the prices.
+
+Caveats:  futures are small?  risk reversal does not measure expectations, foreign exchange is different,
 

### 2018-01-22 MDPI/MDPI_template/FinStudiesbackup.tex
Updated backupfiles a the main document.

--- 2018-01-19 MDPI/MDPI_template/FinStudies.tex
@@
 \section{Introduction}
-\added{Speculation can be viewed as stabilising or destabilising.
+Speculation can be viewed as stabilising or destabilising.
 If speculation is informed, it is a stabilising force, providing liquidity and helping to ensure that markets swiftly find equilibrium.
@@
 In this version, speculators offer a service for a fee that is determined by the return on their activity.
-\citet{FriedmanPositive} argued that even whew there is a mixture of informed and uninformed speculators, the informed would tend to buy when prices were below intrinsic value and sell when above, making profits, limiting volatility and driving the loss-making uninformed speculators out of business.
+\citet{FriedmanPositive} argued that even when speculation is a mixture of informed and uninformed, the informed would tend to buy when prices were below intrinsic value and sell when above, making profits, limiting volatility and driving the loss-making uninformed speculators out of business.
 However, \citet[p.
-101]{Keynes1936} also had a later, more negative view of speculations.
-Here he regarded it as a myopic, sentient-driven activity, dominated by the desire to ``beat the gun`` or ``outwit the crowd.
+101]{Keynes1936} also had a later, more negative view of speculation.
+Here he regarded it as a myopic, sentient-driven activity, dominated by the desire to ``beat the gun`` or ``outwit the crowd``.
 Uninformed speculators may now dominate and \citet{Delong1990noise} provide a model where speculative noise constrains informed traded from benefiting from their knowledge.
@@
 
-%\added{Speculation can be viewed as stabilising or destabilising.} If speculation is based on information it is a stabilising force, providing liquidity and helping to ensure that markets swiftly find equilibrium.
-\citet{KeynesHedge} and \citet{HicksHedge} \deleted{for example} emphasise the role that speculation plays in facilitating financial market transactions \added{by providing liquidity that allows hedgers and other market participants to swiftly find counterparties without the need for large prices concessions.
-In this version, speculators} \deleted{Speculators} offer a service for a fee that is determined by the return on their activity.
-\added{\citet{FriedmanPositive} argued that even where there is a mixture of informed and uninformed speculators, the informed would tend to buy when prices were below intrinsic value and sell when above, making profits, limiting volatility and driving the loss-making uninformed speculators out of business.}\deleted{While speculators may be trading on information or noise, informed speculators will tend to drive out the uninformed noise-traders according to \citet{FriedmanPositive}.} However, \deleted{the later} \citet[p.
-101]{Keynes1936} \added{also had a later, more negative view of speculation.
-Here he} regarded it as a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd''\added{.
-Uninformed speculators may now dominate} and \citet{Delong1990noise} provide a model \deleted{of cases} where speculative noise \replaced{constrains}{provides a constraint on} informed traders \added{from} benefiting from their knowledge.
-Now speculation is an uninformed and destabilising force.
-\added{We have therefore two concepts of speculation: one that is random and disruptive and one that is informed and beneficial in its proviso of liquidity.
+The debate about the relative merit of these two concepts of speculation continues.
+This has been particularly evident in commodities and foreign exchange markets.
+\citet{PindyckSpec} find suspicious co-movement of commodities' prices that they attribute to speculation.
+There has been much discussion over the role that speculation played in the boom and bust of commodities prices in the period between 2003 and 2007.
+This is usually called \emph{The Masters' Thesis} after Fund Manger Michael Masters testimony to the US Senate Committee on Homeland Security and Government Affairs in May 2008 that linked commodity index investors, rising food and energy prices and the so-called \emph{financialisation} of commodity markets.
+\citet{TangSpec} suggest that price disruption was exacerbated by newly established commodity indices that increased the amount of speculation in the commodity markets.
+However, \citet{ScottSpec} dispute this thesis, citing the working of the futures market and the relationship between speculative activity and prices.
+\citet{Jacksspec} document the long history of speculation being blamed for price volatilities.
 
-The debate about the relative merit of these two concepts of speculation continues.
-There has been particular focus on commodities markets and foreign exchange markets.
-\citet{Jacksspec} documents a long history of attacks on speculation following price volatility.
-This has been particularly evident in commodities and foreign exchange markets.
-There has been much discussion over the role that speculation played in the boom and bust of commodities prices in the period between 2006 and 2007.
-\citet{TangSpec} suggest that this disruption was exacerbated by newly established commodity indices that encouraged speculation.
-
-There have also been investigations of the link between speculation or noise trading and price misalignments or volatility across a range of other asset markets.
-\citet{Verma} examine the link between noise trading and equity markets; \citet{Hongjun} argues that speculation is not random and influential in aggregation but that behavioural biases will cause systematic deviations from intrinsic value.
+Across a range of other financial markets there have been investigations of the link between speculation or noise trading and price misalignments or volatility.
+\citet{Verma} examine noise trading in equity markets; \citet{Hongjun} argues that speculation is not random and influential in aggregation but that behavioural biases will cause systematic deviations from intrinsic value.
 Foreign exchange has been a particular focus as a consequence of its liquidity, volatility and the way that speculative attacks appear to have infringed on government sovereignty.
 See \citet{Dornbusch1988} in general and \citet{Eichengreen1988} for discussion of the speculative attacks on the European exchange rate system.
-\citet{Eichengreen1994} use a set of fundamentals to show that speculative attacks drove exchange rates away from their fundamental value.
+\citet{Eichengreen1994} use a set of economic data to show that speculative attacks drove European Exchange Rate Mechanism devaluations rather than fundamentals.
 However, others have made the case that speculative is a key part of the foreign exchange market, providing liquidity and contributing to stability.
@@
 However, this is only the case if speculation is uninformed and destabilising.
-If it is informed and contributes to stability the tax will come at a cost.}
+If it is informed and contributes to stability the tax will come at a cost.
 
+The motivation is to contribute to this debate and to specifically make the case that a tax on transactions in the foreign exchange market would not be costless.
 This paper evaluates the relative importance of these two concepts of speculation by investigating the relationship between the intensity of speculation and the weight of speculative positions relative to subsequent movement of foreign exchange prices.
-\added{The foreign exchange market is appropriate this study because it is central to the call for the introduction of a transaction tax as a consequence of it being the largest and most liquid financial market in the world.
+The foreign exchange market is appropriate this study because it is central to the call for the introduction of a transaction tax as a consequence of it being the largest and most liquid financial market in the world.
 The latest survey taken by the Bank for International Settlements (BIS) in April 2016 estimated average daily turnover at \$5.1trn with average daily spot transactions of \$1.7trn See \citet{BISFX2016}.
@@
 Speculative sentiment is identified with the use of option risk-reversal skew.
-As these become more extreme it signals that expectations of future value in the foreign exchange options market have diverged significantly from the appropriate forward rate.
+As these become more extreme they signal that expectations of future value in the foreign exchange options market have diverged significantly from the appropriate forward rate.
 The Weight of speculative positions is identified with US regulatory data of the positions held by type of participant in the futures market.
-Net long and net short are open positions held by speculators relative to overall activity in the market.
-See Section \ref{secref:measure} for full data details.}
+Net long and net short are open positions held by speculators relative to total speculator activity of to total overall activity in the market.
+See Section \ref{secref:measure} for full data details.
 
@@
 Those trading without information are frequently called \emph{noise-traders} or \emph{liquidity traders}.
-\added{The first implies some random quality that flow from the lack of information.
+The first implies some random quality that flow from the lack of information.
 However, the psychologists find that the cognitive short-cuts or \emph{heuristics} that are used to make decisions under uncertainty tend to have biases \citet{Kahneman1979Prospect} and numerous studies purport to find these biases in financial asset prices.
@@
 
-\added{The foreign exchange market is primarily a dealer-orientated market.
+The foreign exchange market is primarily a dealer-orientated market.
 It is largely consistent with the theoretical microstructure models that have been presented by the likes of \citet{Kyle1985Continuous} and \citet{glosten1985bid}.
-The interaction of informed and uninformed traders drives the price-discovery process: if the uninformed traders are random, the market-maker earns the bid-ask spread; if informed traders arrive they drive the price towards equilibrium.
-This study is about short-term movements in foreign exchange prices.
-The event windows are between 1 week and 4 weeks in length.
-While the interpretation of economic fundamentals can drive prices, market positions and liquidity trades will also be important.
-Even at longer horizons it is often very difficult to connect exchange rate movements to specific events.
-\citet{Meese1983Empirical} find that forecasts using exchange rate models cannot beat a random walk while \citet{ChinnModel} updated the study to find very similar results.
+The interaction of informed and uninformed traders drives the price-discovery process: if the uninformed traders are random, the market-maker earns the bid-ask spread; if informed traders arrive to realise their informational advantage they cause an inventory imbalance for the market-maker and drive the price towards equilibrium.
+Thereby there is a link between information, orders and price.
+
 \citet{Lyons1995Microstructure}, \citet{Evans2002Order} and \citet{rime2010exchange} show that order flow can explain exchange rate movements much more effectively than fundamentals.
@@
 Therefore,if speculators are informed their activity drives exchange rates towards equilibrium and prices should follow a random walk after these extremes of speculative sentiment or position; if speculators are uninformed and their activity can be regarded as noise, the most extreme speculative sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when subsequent price reversals are most likely to take place.
-}
 
 \section{Measuring speculation}\label{secref:measure}
-\added{There are two steps: identify speculation and assess what happens when it becomes extreme.
+There are two steps: identify speculation and assess what happens when it becomes extreme.
 If speculation is uninformed, extremes should cause a divergence from fundamental value and a price reversal should be anticipated; if informed and part of the price-discovery process, extreme speculation achieves a new equilibrium and the following price performance should be a random walk.
 
-The measurement of speculative intensity must identify how far speculators believe that the market will move in a particular direction.
+The measurement of speculative sentiment should identify how far speculators believe that the market will move in a particular direction.
+This study is about short-term movements in foreign exchange prices.
+The event windows are between 1 week and 4 weeks in length.
+While the interpretation of economic fundamentals can drive prices, market positions and liquidity trades will also be important.
+Even at longer horizons it is often very difficult to connect exchange rate movements to specific events.
+\citet{Meese1983Empirical} find that forecasts using exchange rate models cannot beat a random walk while \citet{ChinnModel} updated the study to find very similar more recent results.
 In the absence of a daily survey of speculators, this study uses a unique series of option\emph{risk-reversal skew} as an measure of speculative sentiment.
@@
 The risk-reversals are the spread of implied volatility for a 1 month 25 delta at-the-money call option relative to the equivalent put.
-Risk-reversal skew is the relative price of currency put and call options for the same strike price.} The risk-reversal skew is the relationship between a call option and the equivalent put.
+Risk-reversal skew is the relative price of currency put and call options for the same strike price.
+The risk-reversal skew is the relationship between a call option and the equivalent put.
 Equivalence here being measured by the delta or relationship between movement of the underlying price and the movement of the option price.
@@
 In other words, the Garman-Kohlhagan is used to extract the volatility that is implied by a specific foreign exchange option price given the strike and other parameters.
+It is usual to quote foreign exchange  options in the form of implied volatility.
 
-It is common practice to back market expectations out of information drawn from option prices.
+There is a widespread use of option prices to uncover market expectations.
 See \citet{Malickoption} and \citet{Pochin} for the general approach that combines the implied volatility, risk-reversal and strangles to infer a mixed distribution that can be inferred as the risk-neutral probability density function for a particular asset.
@@
 Using the risk-reversal is a simplified version of that where the skew is used to infer positive and expectations relative to the appropriate forward rate.
-See \citet{Eitrheim} and \citet{ECBRR} for previous uses of this data.}
+See \citet{Eitrheim} and \citet{ECBRR} for previous use implied volatility data for this type of exercise.
 
@@
 Non-commercial accounts are generally considered to be speculators.
-See \citet{FuturesSanders} and \citet{FuturesWang} for examples of previous use of this data in this fashion.
 The data used run from September 30 1998 when the data started to be released on a weekly basis and continue up to December 31 2008.
@@
 These are labelled S1 and S2.
-S1 measures net non-commercial positions relative to the total non-commercial positions and S2 measures non-commercial positions relative to open interest.
+S1 measures net non-commercial positions relative to the total non-commercial positions and S2 measures non-commercial positions relative to open interest.See \citet{FuturesSanders} and \citet{FuturesWang} for examples of previous use of this data to identify speculation.
 
@@
 
-S2 is the preferred measure as it
-
-\added{\subsection{The Event Study}
+\subsection{The Event Study}
 An event study will be used to identify what happens to foreign exchange prices around the time that speculative sentiment or speculative positions are extreme.
@@
 This problem is usually addressed in two ways: modelling the expected value of the asset and viewing deviations from the expected value as being influenced by the event; using a large sample that, on the condition that there are no confounding factors at that time, would tend to dilute the effect of other information.
-In this case, the expected value should be the forward rate but given that we are looking at very short time horizons and interest rates for these currencies are relatively low, it is reasonable to take the expected value as zero and to investigate positive or negative exchange rate return.
-The samples in this case depend on the how the \emph{extreme} sentiment and positions are defined.
+
+The event is an extreme reading of risk-reversal or speculative positions.
+The extreme is defined as being either the top or bottom $5^{th}$ or $1^{st}$ percentile of all the series.
+The expected value should be the forward rate but given that we are looking at very short time horizons and interest rates for these currencies are relatively low, it is reasonable to take the expected value as zero and to investigate positive or negative exchange rate return.
+It is standard practice to call this the \emph{abnormal return} to signify that it is different from the expected value.
+Returns are cumulated through the event window and the average recorded to give the cumulative abnormal return for the whole window (CARW) and cumulative abnormal return for after the event (CARA).
+The number of samples depend on the how the \emph{extreme} sentiment and positions are defined.
 As is seen in Table \ref{tabref:RR1} taking the $5^{th}$ and $95^{th}$ percentile as the extreme for the risk-reversal skew will give around 100 samples for each exchange rate; Table \ref{tabref:SP1} shows that the $95^{th}$ percentile gives just less than 50 samples for the extreme positions.
 The results do not seem to be sensitive to the definition of extreme.
-}
+The event windows are between 1 and 4 days for the risk-reversal and between 2 and 4 weeks for the speculative positions.
+The results do not depend on the event window.
 
@@
 & Lo & 122 & 0.0451 & 0.0996 &122 &-0.3097* &0.2903* & & 21 &-0.07523 &-0.0780 &21 & -0.4120 &
--0.1290  \\
+0.1290  \\
 \multirow{2}{*}{EURCHF}
@@
 Just over half the cases with the four-day window show a reversal (8 of 18), but only 2 of these are statistically significant.
+The level of statistical significance does not make any difference.
 Exchange rate prices are as likely to go up as down after an extreme and there does not appear to be any information about the future from knowledge of these extremes.
@@
 There is evidence here of momentum and of prices following the movement of speculator positions.
+Once again, it appears that orders from speculators drive the price.
 
@@
 If foreign exchange speculation is uninformed noise, extreme speculation, whether measured by sentiment or weight of activity, will coincide with deviations from fundamental value; if speculation is informed, speculation is part of the process of price discovery and extremes will provide no information about future prices.
-An event study showed that speculation is associated with price movement: exchange rates move in the direction of speculative sentiment and activity.
-However, once extreme \added{speculative sentiment or extreme speculative positions are established}, prices are as likely to continue as to reverse.
-Speculative sentiment and speculative activity influence foreign exchange prices as indicated by the covariance of speculative measures with returns.
-These results show that foreign exchange speculation is an important part of the price discovery process.
-The random walk after the extreme is indicative of foreign exchange prices absorbing most of the information that is available.
+It appears that speculation drives prices towards equilibrium and that speculation is informed.
+An event study showed that exchange rates move in the direction of speculative sentiment and activity.
+However, once extreme speculative sentiment or extreme speculative positions are established, prices are as likely to continue as to reverse.
+The random moves after the extreme is indicative of foreign exchange prices having absorbed available information: the next information and movement is random.
 
-The results are not a function of the design of the event study.
+The results do not depend on the design of the event study.
 The event window and the quantiles that are chosen to measure the extreme can be changed.
@@
 Alternative quantiles from 1\% and 99\% to 20\% and 80\% have been assessed; Figures \ref{fig:ES5}, \ref{fig:ES2}, \ref{fig:ES3} and \ref{fig:ES4} have much wider windows and do not find any alternative to the view that extreme speculation is followed by a random walk.
-It is encouraging that two very different measures of speculative activity produce very similar results and that the results hold over different time frames: one week for the option data and one week for the regulatory position data.
-It is less clear whether the results would be reproduced across markets for other financial assets.
 
-\added{The foreign exchange market has a particular microstructure that is dominated by over-the-counter transactions in the spot market and this may influence the results.
-Further research might look at risk-reversals in financial markets that are traded on electronic exchanges to see if these results can be replicated.
-The data on speculative positions are from the exchange traded futures market.} \deleted{The results do not rule out uninformed speculation.
-The Friedman effect from the Noise-trader model, whereby the informed speculators take advantage of the space created by price pressure from the uniformed, depends on there being \emph{noise-trader risk} or some limits to arbitrage.
-The effect is not evident in this study.}  However, the foreign exchange market is a particularly large and liquid market, other asset classes may have alternative institutional features that mean a different relationship or balance between the informed and uninformed and therefore different results.
+It is possible to raise some questions about the measures of speculation that are used here, though the fact that two different measures show very similar results for daily and weekly data is encouraging.
+For example, it may be disputed that there is a link between risk reversal and market sentiment.
+However, the co-movement of this measure of sentiment with the exchange rate is consistent with the findings of numerous studies that have used actual survey data to measure sentiment and expectations, see for example \citet{FrankelFroot1987} and \citet{Frankel1987Using}.
+Nonetheless, future research might try to consider the link between risk-reversals and other measures of expectations.
+It may also be questioned whether speculative positions in the foreign exchange futures market can be used to understand speculation in the much larger spot market.
+There are some private sources of information about spot market speculation and future research might try to establish whether they show the same results.
 
-\added{It is possible to dispute the link between risk reversal and market sentiment.
-It is not only the direct change in expectations about the future that affect  risk-reversals.
-The weight of positions from hedging activity could be influential.
-Therefore future research might also investigate the link between these figures and other measures of short-term sentiment if they exist.}  However, a number of studies have found that exchange rates tend to co-move with survey expectations (see for example \citet{FrankelFroot1987} and \citet{Frankel1987Using}.
-
-\deleted{This is important because there is no unambiguous view of speculative activity and because speculation is an important issues in a number of policy debates.
-Most notably, the \emph{Tobin Tax} would put a charge against financial transactions, particularly in foreign exchange.
-Part of the argument depends on speculation being uninformed as the reduced activity that is caused by the tax would, it is argued, limit volatility and allow more long-term fundamental evaluations to become prominent.
-However, if speculation is informed and part of the price discovery process, this tax may have the opposite effect to what is envisaged.
-These results suggest that it could reduce liquidity and increase volatility.}
-
-It is a common refrain that speculation is destabilising and that it will drive prices away from equilibrium.
-Testing this assertion is difficult because speculation is difficult to measure and equilibrium cannot be viewed.
-However, if speculation is destabilising, when speculators have the most extreme views of the future and when they are most involved in the market should lead to the greatest divergence from equilibrium and the greatest risk that prices will reverse.
-Speculation is measured in two ways:  intensity of sentiment and positions.
-In each case, foreign exchange prices move with speculative sentiment and speculative positions.
-this is consistent with the deliberate orders move prices.
-This would be consistent with speculation being part of the process of price discovery.
-However, in each case the extreme does not tell us anything about the future direction of the prices.
-
-Caveats:  futures are small?  risk reversal does not measure expectations, foreign exchange is different,
+The foreign exchange market has a particular microstructure that is dominated by over-the-counter transactions in the spot market.
+It is not clear whether these results could be used to make assertions about other markets where there are alternative institutions.
+Future research might look at risk-reversals in financial markets that are traded on electronic exchanges to see if these results can be replicated.
+The data on speculative positions are from the exchange traded futures market.
 

### 2018-02-12 MDPI/MDPI_template/FinStudies3.tex
Completed third review.

--- 2018-01-22 MDPI/MDPI_template/FinStudiesbackup.tex
@@
 \subsection{The Event Study}
-An event study will be used to identify what happens to foreign exchange prices around the time that speculative sentiment or speculative positions are extreme.
+It is common to use series that measure market sentiment or market postitions or even proxies like moving averages as \emph{trading-rules} that aim to achieve abnormal returns.
+See for example, \citet{Olson}, \citet{ImprovedMA}, \citet{Malz2} and \citet{DeBrock}.
+It is assumed that the signals will identify times when the market is inefficient.
+However, in our case there is a direct test of the informational content of extreme speculative sentiment and speculative positions.
+A special case of the event study method will be used to identify what happens to foreign exchange prices around the time that speculative sentiment or speculative positions are extreme.
 If speculation is uninformed when speculative sentiment or speculative positions are extremely bias to one direction, there should be the greatest divergence from fundamental value and therefore the greatest risk of reversal; if speculation is informed these extremes are part of the adjustment to a new equilibrium.
